### 2020-01-24 written/software_design.tex
software design notes

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=black,
}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{tikz-uml}
\usetikzlibrary{positioning}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Emergent property inference python package design} \\
Sean Bittner \\
December 19, 2019 \\
\end{center}

\section{Introduction}
Throughout the emergent property inference (EPI) project \cite{bittner2019interrogating}, we hosted all development code on the cunningham-lab github repo: \href{https://github.com/cunningham-lab/dsn}{dsn} (EPI was formerly DSN), which also relies on the cunningham-lab repo \href{https://github.com/cunningham-lab/tf_util}{tf\_util}, which has customized tensorflow code for normalizing flows, and Hessians of deep generative models.  
Here, we design a new python software package for EPI that is more accessible for our user-base of theoretical neuroscientists. 
While functioning as a software development library for our users with high programming skill, this software shall also run seemlessly on NCAP through a graphical user interface.

\section{Requirements}
EPI is designed to be a tool for theoretical neuroscientists to automate inference in complex models of neural circuits.  For more details, see \cite{bittner2019interrogating}.

The epi python package
\begin{itemize}
\item shall run emergent property inference given
\begin{itemize}
\item a model specification
\begin{itemize}
\item parameter names and continuous ranges
\item emergent property statistics as a differentiable function of parameters
\end{itemize}
\item an emergent property value
\item (optional) a specific normalizing flow architecture 
\item (optional) and specific optimization hyperparameters.
\end{itemize}
\item shall return the distribution sampler upon convergence.
\item shall provide log probabilities, and Hessians throughout the support of the EPI distribution.
\item shall visualize optimization diagnostics.
\item shall visualize the EPI distribution and its properties.
\item shall run on NCAP.
\end{itemize}

\section{Dependency selection}
The EPI method relies on automatic differentiation, so we should use one of the three established open source python packages: PyTorch, Tensorflow, or Jax.
PyTorch has risen in popularity with respect to Tensorflow, possibly due to PyTorch's early implementation and release of a simplifying/intuitive eager execution mode.
The Tensorflow 2.0 release puts these two packages on roughly equal footing.
Jax, a python library for autodiff and efficient optimization with standard python/numpy code, is gaining popularity.

\subsection{Normalizing flows support}
In the end, Tensorflow is the only option with an established normalizing flows API -- one that includes real NVP, MAF, spline flows -- making it the most supportive for our purposes: \href{https://www.tensorflow.org/probability/api_docs/python/tfp/bijectors/Bijector}{tf.bijectors}.
PyTorch has a TransformedDistributions class, which doesn't yet support the popular normalizing flow architectures.
PyTorch and Jax have a few options on github, but their level of support is contingent on independent developers (\href{https://github.com/ericjang/nf-jax}{nf-jax}, \href{https://github.com/kamenbliznashki/normalizing_flows}{kamenbliznashki/normalizing\_flows}, \href{https://github.com/karpathy/pytorch-normalizing-flows}{karpathy/pytorch-normalizing-flows}).  If we want to go with these repos, we'll need to have faith in their current implementation, and future maintenance.

\subsection{Model implementation requirement}
The EPI package is designed to receive a differentiable function producing the emergent property statistics from the model parameters.  
In the event that our user-base is largely unfamiliar with Tensorflow/PyTorch, Jax is an interesting alternative, in that we could lower the barriers-to-entry for our software package.
Specifically, users could submit native python functions through which Jax can take gradients.
Otherwise, we must require our users to write Tensorflow or PyTorch functions, or find a reliable way of taking Tensorflow or PyTorch gradients through native python.

To gain an idea of the programming experience of our userbase, I asked 20 Theory Center members working on theoretical modeling projects whether they a.) used python and b.) if so, knew PyTorch or Tensorflow.  The results suggest a 70\% level of familiarity with at least one of PyTorch or Tensorflow.

\begin{table}[h]
\caption{Theoretical neuroscientist programming experience}
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
no python & python (no autograd) & python autograd\\ \hline
3 & 3 & 14 \\ \hline
\end{tabular}
\end{center}
\end{table}
\subsection{Tensorflow implementation rationale}
The lack of a well-supported normalizing flows library for PyTorch and Jax makes these option unattractive.  Given the relatively high python autograd proficiency of our user-base, I conclude that it makes sense to design the package to accept a Tensorflow function defining the emergent property statistics.

\clearpage
\section{EPI package design}
EPI is a method designed to run for a broad space of theoretical models and behaviors.  EPI's implementation changes according to the details of each model.  It makes sense to have our users construct a model object (defined by the Model class), which has a set of defined parameters and an eps() (emergent property statistics) function.  With these class members, we can then run the method epi(), which will return an optimized distribution object (Distribution class).

We create an Architecture class as well, to shield the user from Tensorflow, except through the supplied eps() function.  
As a result, the user can construct simplified python objects (constructors of these classes) according to their use case, run the EPI method, and have a simplified python distribution object returned.
The normalizing flows library and autograd functionality of Tensorflow are used by the package, yet not exposed to the user.

\subsection{EPI class diagram}
\begin{tikzpicture} 
\begin{umlpackage}{epi}
\umlclass{Model}{ 
  \# \_name : string \\ 
  \# \_parameters : list of parameters \\ 
  }
  { 
  + set\_name() \\
  + set\_parameters() \\
  + set\_eps() \\
  + eps() \\
  + epi() \\
  + load\_epi\_dist() \\
}

\umlclass[x=7, y=0]{Parameter}{ 
  \# \_name : string \\ 
  \# \_range : [float, float] \\ 
  }
  { 
  + set\_name() \\
  + set\_range() \\
}

\umlclass[x=0, y=-6]{Architecture}{ 
  \# \_type : string \\ 
  \# \_num\_layers : pos int \\ 
  \# \_post\_affine : bool \\
  }
  { 
  + set\_type() \\
  + set\_num\_layers() \\
  + set\_post\_affine() \\
  + to\_string() \\
  + to\_model() \\
}

\umlclass[x=-1, y=-12]{CouplingArch}{ 
  \# \_type : string \\ 
  \# \_repeats : pos int \\ 
  \# \_num\_couplings : pos int \\ 
  \# \_layers\_per\_coupling : pos int \\ 
  \# \_units\_per\_layer : pos int \\ 
  }
  { 
  + set\_repeats() \\
  + set\_num\_couplings() \\
  + set\_layers\_per\_coupling() \\
  + set\_units\_per\_layer() \\
  + to\_string() \\
  + to\_model() \\
}

\umlclass[x=6, y=-12]{AutoregressiveArch}{ 
  \# \_type : string \\ 
  \# \_repeats : pos int \\ 
  \# \_layers\_per\_conditional : pos int \\ 
  \# \_units\_per\_layer : pos int \\ 
  }
  { 
  + set\_repeats() \\
  + set\_layers\_per\_conditional() \\
  + set\_units\_per\_layer() \\
  + to\_string() \\
  + to\_model() \\
}

\umlclass[x=6, y=-6]{Distribution}{ 
  \# \_D : int \\ 
  \# \_parameters : list of parameters \\ 
  }
  { 
  + sample() \\
  + log\_prob() \\
  + grad() \\
  + hessian() \\
}

\umlunicompo{Model}{Parameter}
\umlunicompo{Distribution}{Parameter}
\umldep{Model}{Architecture}
\umldep{Model}{Distribution}
\umlinherit{CouplingArch}{Architecture}
\umlinherit{AutoregressiveArch}{Architecture}
\end{umlpackage}

\end{tikzpicture}

\subsection{Key functions}
\begin{verbatim}
Model.eps(): parameters (list of tf.Tensors) -> T_x (tf.Tensor)
\end{verbatim}
Tensorflow computation of the the emergent property statistics (eps) of the model  from an arbitrary length list of parameters.  The user can provide this function during initialization or through the Model.set\_eps() method.

\begin{verbatim}
Model.epi(): mu (np.array), arch (Architecture), hps, ... -> 
             epi_dist (Distribution)
\end{verbatim}
Run EPI on the model (which has set parameters and an Tensorflow eps() function) for emergent property value ``mu" with optional parameters for architecture ``arch" and hyperparameters ``hps".

\begin{verbatim}
Architecture.to_model(): self -> tf.Model
\end{verbatim}
Convert Architecture class to a realization of a Tensorflow normalizing flow distribution with the given architecture.

\section{Testing}
\subsection{Unit testing}
We will unit test the API functionality with pytest and codecov.  Package stability on multiple OS's and builds will be assessed with Travis CI, running these unit tests and coverage diagnostics.

\subsection{Validation}
Algorithm validation can happen be done on exponential families with known solutions, and simple systems with derivable contours like the oscillating 2D linear dynamical system.

\section{Documentation}
We will use ReadTheDocs for automatic documentation generation and hosting (no gitpage).

\section{Style}
We will use black as our automatic style formatter.

\bibliography{epi_software_docs}
\bibliographystyle{unsrt}
\end{document}

### 2020-01-24 setup.py
added pytest-cov to setup.py inst req

@@ -7,6 +7,6 @@ setup(name='epi',
       description='Emergent property inference.',
       author='Sean Bittner',
       author_email='srb2201@columbia.edu',
-      install_requires=[],
+      install_requires=['pytest-cov'],
       packages=['epi'],
      )

### 2020-01-24 epi/temp.py
try bash uploader to invoke code cov

@@ -1,0 +1,3 @@
+def foo():
+    print('foo'):
+    return None

### 2020-01-24 epi/error_formatters.py
more norm flow tests

@@ -12,6 +12,7 @@ def format_type_err_msg(obj, arg_name: str, arg, correct_type) -> str:
         arg_type.__name__,
     )
 
+
 def format_arch_type_err_msg(arch_type: str) -> str:
     if arch_type == "ar":
         arch_class = "AutoregressiveArch"

### 2020-01-24 epi/normalizing_flows.py

@@ -1,6 +1,14 @@
 """ Normalizing flow architecture class definitions for param distributions. """
 
+import numpy as np
+import tensorflow as tf
+import tensorflow_probability as tfp
+
+tfb = tfp.bijectors
+tfd = tfp.distributions
+
 from epi.error_formatters import format_type_err_msg, format_arch_type_err_msg
+
 
 class Arch:
     def __init__(self, arch_type, num_layers, post_affine=True):
@@ -8,37 +16,39 @@ class Arch:
         self._set_num_layers(num_layers)
         self._set_post_affine(post_affine)
         self.trainable_variables = []
-        
-    def _set_arch_type(self, arch_type): # Make this noninherited?
-        arch_types = ['planar']
-        if (type(arch_type) is not str):
-            raise TypeError(format_type_err_msg(self, 'arch_type', arch_type, str))
-        if (arch_type not in arch_types):
+
+    def _set_arch_type(self, arch_type):  # Make this noninherited?
+        arch_types = ["planar"]
+        if type(arch_type) is not str:
+            raise TypeError(format_type_err_msg(self, "arch_type", arch_type, str))
+        if arch_type not in arch_types:
             raise ValueError(format_arch_type_err_msg(arch_type))
         self.arch_type = arch_type
 
     def _set_num_layers(self, num_layers):
-        if (type(num_layers) is not int):
-            raise TypeError(format_type_err_msg(self, 'num_layers', num_layers, int))
-        elif (num_layers < 1):
-            raise ValueError('Arch num_layers arg %d must be greater than 0.' % num_layers)
+        if type(num_layers) is not int:
+            raise TypeError(format_type_err_msg(self, "num_layers", num_layers, int))
+        elif num_layers < 1:
+            raise ValueError(
+                "Arch num_layers arg %d must be greater than 0." % num_layers
+            )
         self.num_layers = num_layers
-    
+
     def _set_post_affine(self, post_affine):
-        if (type(post_affine) is not bool):
-            raise TypeError(format_type_err_msg(self, 'post_affine', post_affine, bool))
+        if type(post_affine) is not bool:
+            raise TypeError(format_type_err_msg(self, "post_affine", post_affine, bool))
         self.post_affine = post_affine
-        
+
     def to_string(self,):
-        arch_string = '%d_%s' % (self.num_layers, self.arch_type)
-        if (self.post_affine):
-            arch_string += '_PA'
[…]

### 2020-01-24 epi/temp.py

@@ -1,3 +1,0 @@
-def foo():
-    print('foo'):
-    return None

### 2020-01-24 tests/test_normalizing_flows.py

@@ -1,7 +1,8 @@
 """ Test normalizing flow architectures. """
 
-from epi.normalizing_flows import Arch
+from epi.normalizing_flows import Arch, CouplingArch
 from pytest import raises
+
 
 def test_Arch_init():
     """Test architecture initialization."""
@@ -11,56 +12,106 @@ def test_Arch_init():
 
     # Check setters.
     A = Arch(arch_type, num_layers, post_affine)
-    assert(A.arch_type == 'planar')
-    assert(A.num_layers == 2)
-    assert(A.post_affine)
+    assert A.arch_type == "planar"
+    assert A.num_layers == 2
+    assert A.post_affine
 
     # Check type checking.
     with raises(TypeError):
-        A = Arch(0, num_layers, post_affine) 
+        A = Arch(0, num_layers, post_affine)
     with raises(TypeError):
-        A = Arch({}, num_layers, post_affine) 
+        A = Arch({}, num_layers, post_affine)
 
     with raises(TypeError):
-        A = Arch(arch_type, 2., post_affine) 
+        A = Arch(arch_type, 2.0, post_affine)
     with raises(TypeError):
-        A = Arch(arch_type, 'foo', post_affine) 
+        A = Arch(arch_type, "foo", post_affine)
     with raises(TypeError):
-        A = Arch(arch_type, {}, post_affine) 
+        A = Arch(arch_type, {}, post_affine)
 
     with raises(TypeError):
-        A = Arch(arch_type, num_layers, 0) 
+        A = Arch(arch_type, num_layers, 0)
     with raises(TypeError):
-        A = Arch(arch_type, num_layers, 'False') 
+        A = Arch(arch_type, num_layers, "False")
 
     # Check value checking
     with raises(ValueError):
-        A = Arch('ar', num_layers, post_affine) 
+        A = Arch("ar", num_layers, post_affine)
     with raises(ValueError):
-        A = Arch('coupling', num_layers, post_affine) 
+        A = Arch("coupling", num_layers, post_affine)
     with raises(ValueError):
-        A = Arch('foo', num_layers, post_affine) 
+        A = Arch("foo", num_layers, post_affine)
 
     with raises(ValueError):
-        A = Arch(arch_type, 0, post_affine) 
+        A = Arch(arch_type, 0, post_affine)
     with raises(ValueError):
[…]

### 2020-01-24 setup.py
tensorflow and numpy deps in setup.py

@@ -7,6 +7,9 @@ setup(name='epi',
       description='Emergent property inference.',
       author='Sean Bittner',
       author_email='srb2201@columbia.edu',
-      install_requires=['pytest-cov'],
+      install_requires=['tensorflow=2.1.0',
+                        'tensorflow-probability=0.9.0',
+                        'numpy',
+                        'pytest-cov'],
       packages=['epi'],
      )

### 2020-01-24 setup.py
== instead of = in setup.py

@@ -7,8 +7,8 @@ setup(name='epi',
       description='Emergent property inference.',
       author='Sean Bittner',
       author_email='srb2201@columbia.edu',
-      install_requires=['tensorflow=2.1.0',
-                        'tensorflow-probability=0.9.0',
+      install_requires=['tensorflow==2.1.0',
+                        'tensorflow-probability==0.9.0',
                         'numpy',
                         'pytest-cov'],
       packages=['epi'],

### 2020-10-01 written/V1_model/V1_model.tex
added recent notes

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}

\newcommand{\myname}{Sean Bittner}
\newcommand{\myandrew}{srb2201@columbia.edu}
\newcommand{\myhwnum}{12}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
 %
\pagestyle{fancyplain}
\rhead{\fancyplain{}{\myname\\ \myandrew}}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Input/output properties of neuron-type population model of V1} \\
Sean Bittner \\
May 12, 2020 \\
\end{center}

\section{Introduction}
In section 3.3 of our bioRxiv preprint \cite{bittner2019interrogating}, we analyze the input/output properties of a 4 neuron-type model of V1.  
We identify degeneracies and sensitivities of neuron-type responses w.r.t specific inputs by visual inspection of the EPI distributions.  
In my opinion, these ``novel insights" are not particularly interesting.
Furthermore, these insights are easily obtained using an existing method -- approximate Bayesian computation (ABC).

In this write-up, I present a new line of inquiry in the same V1 circuit model as the paper.
Rather than conditioning on levels of increase of each neuron-type, we condition on the stabilization of each neuron-type.
EPI then presents us with a distribution of stimulation to this V1 circuit that does not change much a particular neuron-type response.

First, I show EPI converges to a distribution covering the set of samples produced by ABC.
This important sanity check, along with some ABC bias visualization, will make nice appendix figures.
Then, I show how EPI readily identifies the most efficacious patterns of stimulation.
Finally, we confirm two distinct mechanisms of E-population stabilization using EPI on the V1 model with inactivated populations.

\section{V1 model (same as bioRxiv post)}
We consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (Fig. 1A). 
Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:

\begin{equation}
\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
\end{equation}

\begin{center}
\textbf{Figure 1: V1 model} \\
\includegraphics[scale=.4]{figs/V1_responses.pdf}
\end{center}

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic}.
The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.

We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
Maintaining the steady state of a neuron-type population amounts to the emergent property 
\begin{equation}
\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.
\end{equation}
In the following analyses, we chose $\sigma=0.25$.

\section{EPI agrees with ABC}
To get an idea of what distribution of parameters ($dh$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
 \[f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.\]
 In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.25$ of $dx_\alpha = 0$.
 In Figure 2, we see that the distributions obtained via EPI (colored by $\log q_\theta(z)$) are visually similar to those obtained via ABC (colored by $dx_\alpha$). \\

 \textbf{Figure 2}: EPI (left) vs ABC (right).  Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$. The importance of $v_S$ and $v_P$ are explained in section 4.2. \\
\begin{center}
$q_\theta(dh \mid \mathcal{B}(E,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(E,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_E.pdf} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_E.png} \\
\clearpage
$q_\theta(dh \mid \mathcal{B}(P,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(P,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_P.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_P.png} \\
$q_\theta(dh \mid \mathcal{B}(S,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(S,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_S.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_S.png} \\
$q_\theta(dh \mid \mathcal{B}(V,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(V,\sigma_{ABC}=0.25)$. \\
\includegraphics[scale=0.25]{figs/V1_drdh_EPI_V.png} 
\includegraphics[scale=0.25]{figs/V1_drdh_ABC_V.png} \\
\end{center}

There are some subtle differences between the EPI and ABC distributions, but some difference is to be expected.
Rather than simply attributing these differences to imperfection of the EPI optimization routine, we consider the the effect of bias in $dx_\alpha$ from ABC (and lack thereof from EPI).

When $\epsilon > 0$ in ABC, the samples are \textit{not} from the posterior distribution.
For $\epsilon > 0$, the ``posterior" predictive means of ABC $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ may be far from zero.
In Figure 3, we see that with ABC, there is an increasingly negative bias in $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ for greater error tolerances across all neuron-types.  
Additionally with ABC, there is no precise control of the variance $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha^2 \right]$, which may be undesireable.

In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4). 
The regions of the ABC pairplots which have $|dx_\alpha| \approx 0.5$ are far less probable in the EPI distributions (by design), and can account for most mismatch observed.
As a note, there is a discontinuity in Figure 2, $EPI(E, \sigma=0.25)$ in the $dh_E$-$dh_S$ plane.
I'd like to see if I can run the optimization for more random seeds so that we obtain a more connected distribution.

\begin{center}
\textbf{Figure 3: Biased ABC} \\
\includegraphics[scale=0.4]{figs/V1_ABC_Tx.pdf}
\end{center}

\clearpage

\begin{center}
\textbf{Figure 4: Unbiased EPI} \\
\includegraphics[scale=0.4]{figs/V1_EPI_Tx.pdf}
\end{center}

\section{EPI-enabled analyses}
\subsection{Dimensions of sensitivity}
Each arrow in the EPI distributions (Fig. 2, left) points in the direction of maximal sensitivity indicated by the hessian $\partial^2 \log q_\theta(dh) / \partial dh \partial dh^\top$ evaluated at the $dh$ at the base of the arrow.
These patterns of input elicit a greater $dx_\alpha^2$ than (at least) 98.8\% of random stimulation patterns.
Each sensitivity vector is representative of a different stabilization mechanism.

\subsection{Confiriming distinct mechanisms of stabilization}
The sensitivity arrows on the EPI distribution for E-population stability (Fig. 2, top, left) indicate two distinct stability mechanisms for increased $dh_E$.
In one case, $dh_P$ must increase along with increased $dh_E$ (Fig. 2, top, left black arrow $v_P$).
In the other case, $dh_S$ must increase along with increased $dh_E$ (Fig. 2, top, left gray arrow $v_S$).
These inferred distributions and sensitivity vectors $v_P$ and $v_S$ motivate the following hypothesis:

H: If the P- and S-populations are inactivated, the only E-stabilization mechanism with increasing $dh_E$ will be that of the other population (S- or P-population, respectively) .

We can test this hypothesis with EPI, since it is optimized to identify the distribution of all inputs resulting in E-population stability.
We ran EPI with the same emergent property as Eq 2 ($\alpha=E$, $\sigma=0.25$) for two different inactivation scenarios.
In the P-inactivated case, we changed the baseline input to $b =b_{\not P} \triangleq \left[ 1, -5, 1 ,1.25 \right]^\top$, and to $b =b_{\not S} \triangleq \left[ 1, 1, -5 ,1.25 \right]^\top$ in the S-inactivated case.
To test for the existence of stabilization mechanisms corresponding to $v_P$ and $v_S$, we measured the most-sensitive dimension of the Hessian at all points in the inferred distribution.  
We declare that the P- or S-inactivated circuit lacks a stability mechanism, if no sensitivity vectors throughout the distribution are aligned with the corresponding sensitivity vector.

\begin{center}
\textbf{Figure 5}: Stability mechanisms of inactivated circuits.
\includegraphics[scale=.5]{figs/cos_dists.pdf}
\end{center}

Confirming our hypothesis, we found that only the $v_S$ stability mechanism is present in the P-inactivated circuit (Figure 5A), and that only the $v_P$ stability mechanism is present in the S-inactivated circuit (Figure 5B).
As expected, the samples of the inferred distributions aligned with the stability mechanism correspond to the sloped regions orthogonal to $v_S$ and $v_P$ (Figure 6).

\clearpage

\begin{center}
\textbf{Figure 6}: Inactivated circuit stability distributions.
\includegraphics[scale=.4]{figs/EPI_inactivation.pdf}
\end{center}

\section{Summary}
We have improved on our previous EPI analysis of V1 in the following ways:
\begin{enumerate}
\item We utilize a novel ability of EPI (Hessian evaluation).
\item Our scientific conclusion is about mechanisms of stability rather than high-level characteristics of responses.
\item We have validation of structure via ABC.
\item We characterize the issue of bias in ABC (and lack thereof in EPI).
\item The emergent property of ``neuron-type stabilization" follows the ethos of the definition of emergent property better than ``neuron-type population rate increase of $y$".
\end{enumerate}

I recommend substituting this work for section 3.3 in the bioRxiv manuscript. I'd be happy to write up a draft of such a section.

%\section{Isolated modulation paths using EPI}
%Let's say we want to find a stimulation pattern that changes the $\alpha$-population steady-state response $dx_\alpha$ as much as possible while keeping the rest the same.
%One can not simply stimulate the $\alpha$-population, since one must take into account the effects of the recurrent connectivity.
%Furthermore, such an input pattern producing isolated modulation may change nonlinearly.
%
%Beginning at $dh = \textbf{0}$, we seek a sequence of differential inputs that will increasingly change $dx_\alpha$, while maintaining $dx_\beta$, $\beta \in \{E, P, S, V\} \setminus	\{\alpha\}$.
%At each step in the path of inputs, the direction is selected to maximize change in $dx_\alpha$ and minimize change in $dx_{\beta \neq \alpha}$.
%These input selections can be guided by the EPI distributions $q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ that result in the modulation of the $\alpha$ population within a given range.
%
%We leverage the availability of a functional representation of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ to calculate the the hessians of the EPI distributions at each step of $dh$ in the path
%\begin{equation}
%\text{Hess}_\alpha(dh) = \frac{\partial^2 \log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))}{\partial dh \partial dh^\top}.
%\end{equation}
%
% We select the optimal direction of isolated modulation as vector v, which points in the maximally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$, and minimally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\beta, 0.2))$, $\beta \neq \alpha$:
% \begin{equation}
% v_\alpha^*(dh) = \argmin_v v^\top \text{Hess}_\alpha(dh) v - \sum_{\beta \neq \alpha} v^\top \text{Hess}_\beta(dh) v,
% \end{equation}
%  where $v\top v$ = 1.  $v_\alpha(dh)$ is obtained via an eigendecomposition:
%  \begin{equation}
%  v_\alpha^*(dh) = \argmin_v v^\top M_\alpha(dh) v
% \end{equation}
% \begin{equation}
%M_\alpha(dh) = \text{Hess}_\alpha(dh) - \sum_{\beta \neq \alpha} \text{Hess}_\beta(dh).
% \end{equation}
% The solution is the eigenvector with smallest eigenvalue of $M_\alpha(dh)$.
% 
% \includegraphics[scale=0.4]{figs/hess_paths.pdf}
%

\bibliography{epi}
\bibliographystyle{unsrt}

\end{document}

### 2020-10-01 written/scaling/epi_scaling.tex
added recent notes

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    citecolor=black,
}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large EPI scaling} \\
Sean Bittner \\
September 17, 2020 \\
\end{center}

\section{Introduction}
A relative strength of EPI is its scalability in parameter dimension $|z|$. 
This write-up compares the performance of EPI to SMC-ABC and SNPE as $|z|$ is increased.  
The application is to condition arbitrary rank-2 RNNs on a regime of stable amplification.

\textbf{SMC-ABC}: 
Traditional approaches to likelihood-free inference -- approximate Bayesian computation (ABC) methods -- randomly sample parameters $z$ until a suitable set is obtained.
State-of-the-art ABC methods leverage sequential monte-carlo (SMC) sampling techniques to obtain parameter sets more efficiently.
To obtain more parameter samples, SMC-ABC must be run from scratch again.
ABC methods do not confer log probabilities of samples.

\textbf{SNPE}: Like EPI, sequential neural posterior estimation (SNPE) uses deep learning to produce flexible posterior approximations.
Like traditional Bayesian inference methods, SNPE conditions directly on the statistics of data.
This differs from EPI, where posteriors are conditioned on emergent properties (moment constraints on the posterior predictive distribution).
Peculiarities of SNPE (density estimation approach, two deep networks) make scaling in $z$ prohibitive.

\textbf{Rank-2 RNN}: The model we use is a rank-2 RNN with N neurons:
\begin{equation*}
U = \begin{bmatrix} u_1 & u_2 \end{bmatrix}, V = \begin{bmatrix} v_1 & v_2 \end{bmatrix}, J = UV^\top, \text{\hspace{2cm}}
u_1, u_2, v_1, v_2 \in \left[-1, 1 \right]^N
\end{equation*}
\begin{equation*}
\tau \dot{r} = -r + Jr
\end{equation*}
From \cite{bondanelli2020coding}, we know necessary and sufficient conditions for RNNs to exhibit stable amplification.  These two conditions are on the maximal real eigenvalue of $J$ ($\text{real}(\lambda_1)$) and the maximal eigenvalue of $J^s = \frac{J + J\top}{2} (\lambda^s_1)$:
\[\text{real}(\lambda_1) < 1, \lambda^s_1 > 1\]

In our analysis, we seek to condition rank-2 networks of increasing size on a regime of stable amplification.  
Networks with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will yield moderate (quantifiable) amplification.

\underline{EPI emergent property}: EPI can naturally condition on this emergent property.
\[\mathbb{E}\begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \\ (\text{real}(\lambda_1)-0.5)^2 \\ (\lambda^s_1 -1.5)^2  \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \\ 0.25^2 \\ 0.25^2 \end{bmatrix}\]

\underline{SNPE observation}: SNPE cannot condition on the variance of observations across posterior.  Thus, we condition on an observation $x_0$ located at the mean of our desired emergent property.
\[x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \]

\underline{SMC-ABC criteria}: ABC methods define tolerance $\epsilon$ and distance for observed data $x_0$.\\
\begin{center}
$\epsilon = 0.5$, $l2$ distance, and  $x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}$
\end{center}

\clearpage

\section{Results}
SMC-ABC was run with \href{https://pyabc.readthedocs.io/en/latest/index.html}{pyabc} package.  SNPE was run with \href{http://www.mackelab.org/delfi/}{delfi} package.  EPI was run with revision branch of \href{https://github.com/cunningham-lab/epi/tree/revision}{epi} package.  All methods use a uniform prior form -1 to 1 for each parameter.

\begin{figure}[h]
\begin{center}
\includegraphics[scale=.4]{epi_both.png} \\
\end{center}
\caption{\small EPI scales with $z$ to high dimensions. 
Square - algorithm never converged across all experiments.
Convergence definitions: 
EPI (blue) - satisfies all moment constraints, 
SNPE (orange)- produces at least 1/10,000 parameter samples in the bounds of emergent property (mean +- 0.5), 
and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=.3]{EPI_scaling3.png}
\caption{\small Emergent property fidelity.  Top row: Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
Gray star indicates emergent property mean, and gray dashed lines indicate two standard deviations corresponding to the variance constraint.
For $N <= 6$ where SMC-ABC converges, samples are not diverse (path degeneracies).  
For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
Bottom row: simulations of network parameters resulting from each method ($\tau=100ms$).  Each trace corresponds to simulation of one $z$.  
Almost all EPI networks, few SNPE networks, and all SMC-ABC networks exhibit stable amplification.  
There is little variation in SMC-ABC network simulations.}
\end{center}
\end{figure}

\bibliography{epi_scaling}
\bibliographystyle{unsrt}
\end{document}

### 2020-10-01 written/software/software_design.tex
added recent notes

--- 2020-01-24 written/software_design.tex
@@
 \umlclass{Model}{ 
-  \# \_name : string \\ 
-  \# \_parameters : list of parameters \\ 
+  + name : string \\ 
+  + parameters : list of parameters \\ 
   }
   { 
-  + set\_name() \\
-  + set\_parameters() \\
+  \# \_set\_name() \\
+  \# \_set\_parameters() \\
   + set\_eps() \\
@@
 \umlclass[x=7, y=0]{Parameter}{ 
-  \# \_name : string \\ 
-  \# \_range : [float, float] \\ 
+  + name : string \\ 
+  + bounds : [float, float] \\ 
   }
   { 
-  + set\_name() \\
-  + set\_range() \\
+  \# \_set\_name() \\
+  \# \_set\_bounds() \\
 }
@@
 \umlclass[x=0, y=-6]{Architecture}{ 
-  \# \_type : string \\ 
-  \# \_num\_layers : pos int \\ 
-  \# \_post\_affine : bool \\
+  + arch\_type : string \\ 
+  + num\_layers : pos int \\ 
+  + post\_affine : bool \\
   }
   { 
-  + set\_type() \\
-  + set\_num\_layers() \\
-  + set\_post\_affine() \\
+  + \_\_call\_\_() \\
+  + \_set\_type() \\
+  + \_set\_num\_layers() \\
+  + \_set\_post\_affine() \\
   + to\_string() \\
@@
 \umlclass[x=-1, y=-12]{CouplingArch}{ 
-  \# \_type : string \\ 
-  \# \_repeats : pos int \\ 
-  \# \_num\_couplings : pos int \\ 
-  \# \_layers\_per\_coupling : pos int \\ 
-  \# \_units\_per\_layer : pos int \\ 
+  + D : pos int \\ 
+  + num\_couplings : pos int \\
+  + num\_units : pos int \\  
   }
   { 
-  + set\_repeats() \\
-  + set\_num\_couplings() \\
-  + set\_layers\_per\_coupling() \\
-  + set\_units\_per\_layer() \\
+  + \_\_call\_\_() \\
+  \# \_set\_D() \\
+  \# \_set\_num\_couplings() \\
+  \# \_set\_num\_units() \\
   + to\_string() \\
-  + to\_model() \\
 }
@@
 \umlclass[x=6, y=-12]{AutoregressiveArch}{ 
-  \# \_type : string \\ 
-  \# \_repeats : pos int \\ 
-  \# \_layers\_per\_conditional : pos int \\ 
-  \# \_units\_per\_layer : pos int \\ 
+  + D : pos int \\ 
+  + num\_couplings : pos int \\
+  + num\_units : pos int \\  
   }
   { 
-  + set\_repeats() \\
-  + set\_layers\_per\_conditional() \\
-  + set\_units\_per\_layer() \\
+  + \_\_call\_\_() \\
+  \# \_set\_D() \\
+  \# \_set\_num\_ars() \\
+  \# \_set\_num\_units() \\
   + to\_string() \\
-  + to\_model() \\
 }

### 2020-10-01 scripts/V1_Circuit/SV_flip_epi.py
new sv flip

@@ -1,4 +1,4 @@
-import os
+import os 
 import argparse
 import numpy as np
 import tensorflow as tf
@@ -24,25 +24,23 @@ ub_dh = 1.*np.ones((2,))
 
 h = Parameter("h", 4, lb=lb_h, ub=ub_h)
 dh = Parameter("dh", 2, lb=lb_dh, ub=ub_dh)
-#epsilon = Parameter("epsilon", 1, lb=0., ub=1.)
 
 # Define model
-name = "V1Circuit"
+sigma_eps = 0.1
+name = "V1Circuit_SVflip_sigeps=%.2f" % sigma_eps
 parameters = [h, dh]
 model = Model(name, parameters)
 
 X_INIT = tf.constant(np.random.normal(1.0, 0.01, (1, 4, 1)).astype(np.float32))
 
 # Define eps
-diff_prod_mean = -0.5
+diff_prod_mean = -0.25
 diff_sum_mean = 0.
-def SV_flip(h, dh):#, epsilon):
+def SV_flip(h, dh):
     h = h[:, :, None]
     dh = tf.concat((dh, tf.zeros_like(dh, dtype=tf.float32)), axis=1)[:, :, None]
-    #epsilon = epsilon[:, :, None]
 
     n = 2.
-    epsilon = 0.05
     dt = 0.005
     T = 100
     tau = 0.02
@@ -57,12 +55,12 @@ def SV_flip(h, dh):#, epsilon):
 
     def f1(y):
         omega = tf.random.normal(y.shape, 0., 1.)
-        noise = epsilon*omega
+        noise = sigma_eps*omega
         return (-y + (tf.nn.relu(tf.matmul(W, y) + h + noise) ** n)) / tau
 
     def f2(y):
         omega = tf.random.normal(y.shape, 0., 1.)
-        noise = epsilon*omega
+        noise = sigma_eps*omega
         return (-y + (tf.nn.relu(tf.matmul(W, y) + h + dh + noise) ** n)) / tau
 
     ss1 = euler_sim(f1, x_init, dt, T)
@@ -82,7 +80,7 @@ def SV_flip(h, dh):#, epsilon):
 model.set_eps(SV_flip)
 
 # Emergent property values.
-mu = np.array([diff_prod_mean, diff_sum_mean, 0.25**2, 0.25**2])
+mu = np.array([diff_prod_mean, diff_sum_mean, 0.125**2, 0.125**2])
 
 # 3. Run EPI.
 q_theta, opt_data, epi_path, failed = model.epi(

### 2020-10-01 scripts/V1_Circuit/SV_flip_epi.py
drdh

@@ -26,7 +26,7 @@ h = Parameter("h", 4, lb=lb_h, ub=ub_h)
 dh = Parameter("dh", 2, lb=lb_dh, ub=ub_dh)
 
 # Define model
-sigma_eps = 0.1
+sigma_eps = 0.25
 name = "V1Circuit_SVflip_sigeps=%.2f" % sigma_eps
 parameters = [h, dh]
 model = Model(name, parameters)
@@ -34,7 +34,7 @@ model = Model(name, parameters)
 X_INIT = tf.constant(np.random.normal(1.0, 0.01, (1, 4, 1)).astype(np.float32))
 
 # Define eps
-diff_prod_mean = -0.25
+diff_prod_mean = -0.5
 diff_sum_mean = 0.
 def SV_flip(h, dh):
     h = h[:, :, None]
@@ -80,7 +80,7 @@ def SV_flip(h, dh):
 model.set_eps(SV_flip)
 
 # Emergent property values.
-mu = np.array([diff_prod_mean, diff_sum_mean, 0.125**2, 0.125**2])
+mu = np.array([diff_prod_mean, diff_sum_mean, 0.25**2, 0.25**2])
 
 # 3. Run EPI.
 q_theta, opt_data, epi_path, failed = model.epi(

### 2020-10-01 scripts/V1_Circuit/drdh_epi.py

@@ -23,9 +23,9 @@ c0 = 10.**args.logc0
 bnmom = args.bnmom
 random_seed = args.random_seed
 
-#sleep_dur = ord(alpha)/11. + 3.*epsilon + np.abs(args.logc0) + random_seed/5.
-#print('short stagger sleep of', sleep_dur, flush=True)
-#time.sleep(sleep_dur)
+sleep_dur = ord(alpha)/11. + beta/2. + 3.*epsilon + np.abs(args.logc0) + random_seed/17.
+print('short stagger sleep of', sleep_dur, flush=True)
+time.sleep(sleep_dur)
 
 # 1. Specify the V1 model for EPI.
 D = 4
@@ -44,7 +44,7 @@ X_INIT = tf.constant(np.random.normal(1.0, 0.01, (1, 4, 1)).astype(np.float32))
 inc_val = 0.
 npzfile = np.load('SV_mode.npz')
 z_mode = npzfile['z_mode1']
-h = z_mode[:,:4]
+h = z_mode[0,:4]
 dr = V1_dr_eps(alpha, inc_val, epsilon, h=h)
 model.set_eps(dr)
 

### 2020-10-02 scripts/V1_Circuit/drdh_epi.py
drdh

@@ -25,7 +25,7 @@ random_seed = args.random_seed
 
 sleep_dur = ord(alpha)/11. + beta/2. + 3.*epsilon + np.abs(args.logc0) + random_seed/17.
 print('short stagger sleep of', sleep_dur, flush=True)
-time.sleep(sleep_dur)
+#time.sleep(sleep_dur)
 
 # 1. Specify the V1 model for EPI.
 D = 4
@@ -42,9 +42,11 @@ model = Model(name, parameters)
 X_INIT = tf.constant(np.random.normal(1.0, 0.01, (1, 4, 1)).astype(np.float32))
 
 inc_val = 0.
-npzfile = np.load('SV_mode.npz')
-z_mode = npzfile['z_mode1']
-h = z_mode[0,:4]
+
+npzfile = np.load('SVflips_hs.npz')
+print(npzfile)
+h = npzfile['b_cross'][0]
+
 dr = V1_dr_eps(alpha, inc_val, epsilon, h=h)
 model.set_eps(dr)
 

### 2020-10-05 written/software/software_design.tex

@@
   + arch\_type : string \\ 
-  + num\_layers : pos int \\ 
+  + D : pos int \\ 
+  + num\_stages :  int \\ 
+  + num\_units :  int \\ 
+  + num\_layers :  int \\
+  + batch\_norm : bool \\
+  + bn\_momentum : float \\ 
   + post\_affine : bool \\
+  + bounds : tuple \\
+  + random\_seed : int \\
   }
   { 
-  + \_\_call\_\_() \\
-  + \_set\_type() \\
-  + \_set\_num\_layers() \\
-  + \_set\_post\_affine() \\
-  + to\_string() \\
-  + to\_model() \\
-}
-
-\umlclass[x=-1, y=-12]{CouplingArch}{ 
-  + D : pos int \\ 
-  + num\_couplings : pos int \\
-  + num\_units : pos int \\  
-  }
-  { 
-  + \_\_call\_\_() \\
-  \# \_set\_D() \\
-  \# \_set\_num\_couplings() \\
-  \# \_set\_num\_units() \\
-  + to\_string() \\
-}
-
-\umlclass[x=6, y=-12]{AutoregressiveArch}{ 
-  + D : pos int \\ 
-  + num\_couplings : pos int \\
-  + num\_units : pos int \\  
-  }
-  { 
-  + \_\_call\_\_() \\
-  \# \_set\_D() \\
-  \# \_set\_num\_ars() \\
-  \# \_set\_num\_units() \\
+  \# setters
+  + sample(N : int) \\
+  + initialize(init\_type, init\_params) \\
   + to\_string() \\
@@
 \umldep{Model}{Distribution}
-\umlinherit{CouplingArch}{Architecture}
-\umlinherit{AutoregressiveArch}{Architecture}
 \end{umlpackage}

### 2020-10-07 notebooks/episync.sh
updating sync file

@@ -1,1 +1,2 @@
 rsync -azhe ssh --progress srb2201@habanero.rcs.columbia.edu:/rigel/stats/users/srb2201/epi/scripts/$1/data $1/
+rsync -azhe ssh --progress srb2201@terremoto.rcs.columbia.edu:/moto/stats/users/srb2201/epi/scripts/$1/data $1/

### 2020-10-07 written/V1_model/V1_model.tex
latest writing for V1 section

@@
 \begin{center}                  % Center the following lines
-{\Large Input/output properties of neuron-type population model of V1} \\
-Sean Bittner \\
-May 12, 2020 \\
+{\Large Draft of new V1 section} \\
+Sean Bittner, Agostina Palmigiano \\
+October 6, 2020 \\
 \end{center}
 
-\section{Introduction}
-In section 3.3 of our bioRxiv preprint \cite{bittner2019interrogating}, we analyze the input/output properties of a 4 neuron-type model of V1.  
-We identify degeneracies and sensitivities of neuron-type responses w.r.t specific inputs by visual inspection of the EPI distributions.  
-In my opinion, these ``novel insights" are not particularly interesting.
-Furthermore, these insights are easily obtained using an existing method -- approximate Bayesian computation (ABC).
+\section{EPI clarifies the implications of contrast-response flipping on [noise quenching, inhibition stabilization] in V1}\label{results_V1}
+Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
+However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+\ref{fig:SV_flip}A) \cite{pfeffer2013inhibition}.
+Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
+Here, we use EPI to analyze the posteriors of a stochastic nonlinear dynamical model of V1 conditioned on the emergent property of contrast-dependent S-V flipping.
+We then [use this info to make a statement about ...]
 
-In this write-up, I present a new line of inquiry in the same V1 circuit model as the paper.
-Rather than conditioning on levels of increase of each neuron-type, we condition on the stabilization of each neuron-type.
-EPI then presents us with a distribution of stimulation to this V1 circuit that does not change much a particular neuron-type response.
-
-First, I show EPI converges to a distribution covering the set of samples produced by ABC.
-This important sanity check, along with some ABC bias visualization, will make nice appendix figures.
-Then, I show how EPI readily identifies the most efficacious patterns of stimulation.
-Finally, we confirm two distinct mechanisms of E-population stabilization using EPI on the V1 model with inactivated populations.
-
-\section{V1 model (same as bioRxiv post)}
-We consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (Fig. 1A). 
-Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:
+Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and sum of recurrent inputs $Wx$, external inputs $h$, and external noise $\epsilon \sim \mathcal{N}(0, \sigma_{\epsilon}^2)$:
 
 \begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
+\tau \frac{dx}{dt} = -x + [W x+ h + \epsilon]_+^n.
 \end{equation}
 
+We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
+The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
+Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.
+
+With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
+Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
+All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
+While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
+\ref{fig:V1_EPI}B right).
+Currently available approaches to deriving the steady state response of the system are limited.
+
+\clearpage
+\begin{figure}[h]
+\caption{\small \textbf{A}.
+Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.
+Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
+\textbf{B}.
+EPI posterior $q_\theta(z \mid \mathcal{B}_{S-V})$ for S-V flipping.
+The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$.
+This posterior is bimodal and concentrated in planes $dh_E > 0$ and $dh_E < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.
+Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two stds).
+\textbf{C}. Posterior predictive distribution of $d_{S-V}(h)$ of each mode shows that the $z_1$-mode produces V-to-S flipping with increasing contrast and the $z_2$-mode produces S-to-V flipping.
+\textbf{D}.
+Model simulations at the mode $z_1$ at $b_1$ (left), $b_1 + dh_1^{(c)}$ (middle), and steady state solutions for varying levels of contrast (right).
+Shaded area is one standard deviation according to randomness of $\epsilon$.
+\textbf{E}.
+Same as D.
+for $z_2$.
+\textbf{F}.
+...
+ }\label{fig:SV_flip}
 \begin{center}
-\textbf{Figure 1: V1 model} \\
-\includegraphics[scale=.4]{figs/V1_responses.pdf}
+\includegraphics[scale=.31]{figs/Fig3/Fig3.pdf}
 \end{center}
+\end{figure}
+\clearpage
 
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic}.
-The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
+TODO Summarize S-V flipping phenomena and why it's a good EP
+TODO Explain that we can't really fit the whole data-set of responses yet because the model is quite constrained.
+TODO Motivate this input model, where $dh^{(c)}$ is the direction of increasing contrast.
 
-We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
-We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
-We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
-Maintaining the steady state of a neuron-type population amounts to the emergent property 
 \begin{equation}
-\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.
+ \begin{bmatrix} h_E \\ h_P \\ h_S \\ h_V \end{bmatrix} =  \begin{bmatrix} b_E \\ b_P \\ b_S \\ b_V \end{bmatrix} +  \begin{bmatrix} dh^{(c)}_E \\ dh^{(c)}_P \\ 0 \\ 0 \end{bmatrix} 
 \end{equation}
-In the following analyses, we chose $\sigma=0.25$.
 
-\section{EPI agrees with ABC}
-To get an idea of what distribution of parameters ($dh$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
-We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
- \[f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.\]
- In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.25$ of $dx_\alpha = 0$.
- In Figure 2, we see that the distributions obtained via EPI (colored by $\log q_\theta(z)$) are visually similar to those obtained via ABC (colored by $dx_\alpha$). \\
+Since we are unsure what either the baseline input $b$ and contrast-dependent change in input $dh^{(c)}$ should be, we treat them as free parameters when running EPI.
+\begin{equation}
+ z =  \begin{bmatrix} b_E & b_P & b_S & b_V & dh^{(c)}_E & dh^{(c)}_P \end{bmatrix}^\top
+\end{equation}
+We consider positive baseline inputs $b_\alpha \in \left[0, 3 \right]$ and small contrast-dependent changes in input $|dh_\alpha| \leq 1$.  
+To find parameters resulting in S-V flipping, we focus on models driven by nominal amounts of external noise $\sigma_{\eta} = 0.1$.
 
- \textbf{Figure 2}: EPI (left) vs ABC (right).
+We formulate the emergent property of S-V flipping with a statistic measuring he difference in steady state ($x_ss$) between the S- and V-population at a given input:
+\begin{equation}
+d_{S-V}(h) = x_{ss,S}(h) - x_{ss,V}(h).
+\end{equation}
+For S and V to flip their steady states, the difference between the S- and V-population rates at $h = b$ must have opposite sign from $h = b + dh$.
+Therefore, we stipulate the emergent property of S-V flipping to require the product between $d_{S-V}(b)$ and $d_{S-V}(b+dh)$ to be appreciably negative.
+Second we stipulate that the differences between S and V in each input condition cancel out on average ($d_{S-V}(b) + d_{S-V}(b+dh)$ is 0 on average).
+The means and variances of the emergent property statistics were sensibly chosen based on some inexpensive model simulations within the parameter bounds.
+
+\begin{equation}
+\mathcal{B}_{S-V} \triangleq \mathbb{E}\begin{bmatrix} d_{S-V}(b)d_{S-V}(b+dh) \\ \left( d_{S-V}(b)d_{S-V}(b+dh) - (-0.25) \right)^2 \\  d_{S-V}(b) + d_{S-V}(b+dh) \\  \left( d_{S-V}(b) + d_{S-V}(b+dh)\right)^2 \end{bmatrix} =
+\begin{bmatrix} -0.25 \\ 0.125^2 \\ 0 \\ 0.125^2 \end{bmatrix}
+\end{equation}
+
+We ran EPI to inspect the structure of the posterior distribution of $z$ conditioned on S-V flipping shown in Fig 2.  
+It is clear from this visualization that S-V flipping in this V1-model is sensitive with respect to some parameter settings and robust with respect to others.
+Additionally, the posterior is bimodal: there is one mode in each of the hyperplanes $dh_E < 0$ and $dh_E > 0$.
+
+The structure of the V1 model S-V flipping posterior yields the following insights:
+\begin{enumerate}
+\item The marginal distributions of $q_\theta(h_P \mid \mathcal{B}_{S-V})$ and  $q_\theta(dh_P \mid \mathcal{B}_{S-V})$ are approximately uniform along their allowed range.  
+The approximate uniformity of the $P-population$ parameter marginal distributions shows that the P-population plays little role in S-V flipping. 
+\item The pairwise marginal distribution of $q_\theta(h_S, h_V \mid \mathcal{B}_{S-V})$ shows strong correlation between these parameters.
+The strong correlation between $h_S$ and $h_V$ in the posterior reveals that S-V flipping is sensitive with respect to the baseline inputs to the S- and V-populations.
+\textbf{Augment this to show that multidim info from Hessian is useful.}
+The Hessian provided by EPI at the modes indicates that coordinated increases of $h_S$ and $h_V$ by a ratio of $\frac{h_S}{h_V} = 0.701$ will preserve S-V flipping, while changes in an orthogonal dimension will disrupt it.
+
+\item This distribution is multimodal.
+\end{enumerate}
+
+%\section{Stability analysis}
+%The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
+%
+%We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
+%We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
+%We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
+%Maintaining the steady state of a neuron-type population amounts to the emergent property 
+%\begin{equation}
+%\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
+%\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.\\
+%\end{equation}
+%In the following analyses, we chose $\sigma=0.25$.
+%
+%\section{EPI agrees with ABC}
+%To get an idea of what distribution of parameters ($dh$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
+%We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
+% \[f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.\]
+% In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.25$ of $dx_\alpha = 0$.
+% In Figure 2, we see that the distributions obtained via EPI (colored by $\log q_\theta(z)$) are visually similar to those obtained via ABC (colored by $dx_\alpha$). \\
+% 
+% \textbf{Figure 2}: EPI (left) vs ABC (right).
 Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$.
@@
 \\
+%
+%There are some subtle differences between the EPI and ABC distributions, but some difference is to be expected.
+%Rather than simply attributing these differences to imperfection of the EPI optimization routine, we consider the the effect of bias in $dx_\alpha$ from ABC (and lack thereof from EPI).
+%
+%When $\epsilon > 0$ in ABC, the samples are \textit{not} from the posterior distribution.
+%For $\epsilon > 0$, the ``posterior" predictive means of ABC $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ may be far from zero.
+%In Figure 3, we see that with ABC, there is an increasingly negative bias in $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ for greater error tolerances across all neuron-types.  
+%Additionally with ABC, there is no precise control of the variance $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha^2 \right]$, which may be undesireable.
+%
+%In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
+%The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).
+
+\section{V1 Supplemental section}\label{methods_V1}
+
+We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1.
+Specifically, Billeh et al.
+\cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
+\begin{equation}
+M = \begin{bmatrix} 
+0.36 & -0.48 & -0.31 & -0.28 \\
+1.49 & -0.68 & -0.50 & -0.18 \\
+0.86 & -0.42 & -0.15 & -0.32 \\
+1.31 & -0.41 & -0.52 & -0.37 \end{bmatrix}
+\end{equation}
+and connection probability
+\begin{equation}
+C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
+0.395 & .451 & 0.857 & 0.02 \\
+0.182 & 0.03 & 0.082 & 0.625 \\
+0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
+\end{equation}
+
+Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
+\begin{equation}
+W = C \odot M = \begin{bmatrix} 
+0.0576 & -0.197 & -0.131 & -0.0244 \\
+0.589 & -0.307 & -0.429 & -0.00360 \\
+0.157 & -0.0126  & -0.0123 & -0.200 \\
+0.138 & -0.0902 & -0.400 & -0.0104 \end{bmatrix}.
+\end{equation}
+
+\begin{figure}[h]
+\caption{\small \textbf{A}. ...
+ }\label{fig:drdh}
 \begin{center}
-$q_\theta(dh \mid \mathcal{B}(E,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(E,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_E.pdf} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_E.png} \\
-\clearpage
-$q_\theta(dh \mid \mathcal{B}(P,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(P,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_P.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_P.png} \\
-$q_\theta(dh \mid \mathcal{B}(S,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(S,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_S.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_S.png} \\
-$q_\theta(dh \mid \mathcal{B}(V,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(V,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_V.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_V.png} \\
+\includegraphics[scale=.3]{figs/FigSabc/FigSabc.pdf}
 \end{center}
+\end{figure}
 
-There are some subtle differences between the EPI and ABC distributions, but some difference is to be expected.
-Rather than simply attributing these differences to imperfection of the EPI optimization routine, we consider the the effect of bias in $dx_\alpha$ from ABC (and lack thereof from EPI).
-
-When $\epsilon > 0$ in ABC, the samples are \textit{not} from the posterior distribution.
-For $\epsilon > 0$, the ``posterior" predictive means of ABC $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ may be far from zero.
-In Figure 3, we see that with ABC, there is an increasingly negative bias in $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ for greater error tolerances across all neuron-types.  
-Additionally with ABC, there is no precise control of the variance $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha^2 \right]$, which may be undesireable.
-
-In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
-The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4). 
-The regions of the ABC pairplots which have $|dx_\alpha| \approx 0.5$ are far less probable in the EPI distributions (by design), and can account for most mismatch observed.
-As a note, there is a discontinuity in Figure 2, $EPI(E, \sigma=0.25)$ in the $dh_E$-$dh_S$ plane.
-I'd like to see if I can run the optimization for more random seeds so that we obtain a more connected distribution.
-
+\begin{figure}[h]
+\caption{\small \textbf{A}. ...
+ }\label{fig:drdh}
 \begin{center}
-\textbf{Figure 3: Biased ABC} \\
-\includegraphics[scale=0.4]{figs/V1_ABC_Tx.pdf}
+\includegraphics[scale=.12]{figs/FigSX/FigSX.pdf}
 \end{center}
-
-\clearpage
-
-\begin{center}
-\textbf{Figure 4: Unbiased EPI} \\
-\includegraphics[scale=0.4]{figs/V1_EPI_Tx.pdf}
-\end{center}
-
-\section{EPI-enabled analyses}
-\subsection{Dimensions of sensitivity}
-Each arrow in the EPI distributions (Fig.
-2, left) points in the direction of maximal sensitivity indicated by the hessian $\partial^2 \log q_\theta(dh) / \partial dh \partial dh^\top$ evaluated at the $dh$ at the base of the arrow.
-These patterns of input elicit a greater $dx_\alpha^2$ than (at least) 98.8\% of random stimulation patterns.
-Each sensitivity vector is representative of a different stabilization mechanism.
-
-\subsection{Confiriming distinct mechanisms of stabilization}
-The sensitivity arrows on the EPI distribution for E-population stability (Fig. 2, top, left) indicate two distinct stability mechanisms for increased $dh_E$.
-In one case, $dh_P$ must increase along with increased $dh_E$ (Fig. 2, top, left black arrow $v_P$).
-In the other case, $dh_S$ must increase along with increased $dh_E$ (Fig. 2, top, left gray arrow $v_S$).
-These inferred distributions and sensitivity vectors $v_P$ and $v_S$ motivate the following hypothesis:
-
-H: If the P- and S-populations are inactivated, the only E-stabilization mechanism with increasing $dh_E$ will be that of the other population (S- or P-population, respectively) .
-
-We can test this hypothesis with EPI, since it is optimized to identify the distribution of all inputs resulting in E-population stability.
-We ran EPI with the same emergent property as Eq 2 ($\alpha=E$, $\sigma=0.25$) for two different inactivation scenarios.
-In the P-inactivated case, we changed the baseline input to $b =b_{\not P} \triangleq \left[ 1, -5, 1 ,1.25 \right]^\top$, and to $b =b_{\not S} \triangleq \left[ 1, 1, -5 ,1.25 \right]^\top$ in the S-inactivated case.
-To test for the existence of stabilization mechanisms corresponding to $v_P$ and $v_S$, we measured the most-sensitive dimension of the Hessian at all points in the inferred distribution.  
-We declare that the P- or S-inactivated circuit lacks a stability mechanism, if no sensitivity vectors throughout the distribution are aligned with the corresponding sensitivity vector.
-
-\begin{center}
-\textbf{Figure 5}: Stability mechanisms of inactivated circuits.
-\includegraphics[scale=.5]{figs/cos_dists.pdf}
-\end{center}
-
-Confirming our hypothesis, we found that only the $v_S$ stability mechanism is present in the P-inactivated circuit (Figure 5A), and that only the $v_P$ stability mechanism is present in the S-inactivated circuit (Figure 5B).
-As expected, the samples of the inferred distributions aligned with the stability mechanism correspond to the sloped regions orthogonal to $v_S$ and $v_P$ (Figure 6).
-
-\clearpage
-
-\begin{center}
-\textbf{Figure 6}: Inactivated circuit stability distributions.
-\includegraphics[scale=.4]{figs/EPI_inactivation.pdf}
-\end{center}
-
-\section{Summary}
-We have improved on our previous EPI analysis of V1 in the following ways:
-\begin{enumerate}
-\item We utilize a novel ability of EPI (Hessian evaluation).
-\item Our scientific conclusion is about mechanisms of stability rather than high-level characteristics of responses.
-\item We have validation of structure via ABC.
-\item We characterize the issue of bias in ABC (and lack thereof in EPI).
-\item The emergent property of ``neuron-type stabilization" follows the ethos of the definition of emergent property better than ``neuron-type population rate increase of $y$".
-\end{enumerate}
-
-I recommend substituting this work for section 3.3 in the bioRxiv manuscript. I'd be happy to write up a draft of such a section.
-
-%\section{Isolated modulation paths using EPI}
-%Let's say we want to find a stimulation pattern that changes the $\alpha$-population steady-state response $dx_\alpha$ as much as possible while keeping the rest the same.
-%One can not simply stimulate the $\alpha$-population, since one must take into account the effects of the recurrent connectivity.
-%Furthermore, such an input pattern producing isolated modulation may change nonlinearly.
-%
-%Beginning at $dh = \textbf{0}$, we seek a sequence of differential inputs that will increasingly change $dx_\alpha$, while maintaining $dx_\beta$, $\beta \in \{E, P, S, V\} \setminus	\{\alpha\}$.
-%At each step in the path of inputs, the direction is selected to maximize change in $dx_\alpha$ and minimize change in $dx_{\beta \neq \alpha}$.
-%These input selections can be guided by the EPI distributions $q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ that result in the modulation of the $\alpha$ population within a given range.
-%
-%We leverage the availability of a functional representation of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ to calculate the the hessians of the EPI distributions at each step of $dh$ in the path
-%\begin{equation}
-%\text{Hess}_\alpha(dh) = \frac{\partial^2 \log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))}{\partial dh \partial dh^\top}.
-%\end{equation}
-%
-% We select the optimal direction of isolated modulation as vector v, which points in the maximally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$, and minimally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\beta, 0.2))$, $\beta \neq \alpha$:
-% \begin{equation}
-% v_\alpha^*(dh) = \argmin_v v^\top \text{Hess}_\alpha(dh) v - \sum_{\beta \neq \alpha} v^\top \text{Hess}_\beta(dh) v,
-% \end{equation}
-%  where $v\top v$ = 1.  $v_\alpha(dh)$ is obtained via an eigendecomposition:
-%  \begin{equation}
-%  v_\alpha^*(dh) = \argmin_v v^\top M_\alpha(dh) v
-% \end{equation}
-% \begin{equation}
-%M_\alpha(dh) = \text{Hess}_\alpha(dh) - \sum_{\beta \neq \alpha} \text{Hess}_\beta(dh).
-% \end{equation}
-% The solution is the eigenvector with smallest eigenvalue of $M_\alpha(dh)$.
-% 
-% \includegraphics[scale=0.4]{figs/hess_paths.pdf}
-%
+\end{figure}
 

### 2020-11-02 written/eLifeRevision/Bittner_eLife2020.tex
eLife revision placeholder

\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{floatrow}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\oddsidemargin0cm
\topmargin-2cm
\textwidth16.5cm
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}
\linespread{1.5}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%

\begin{document}
\medskip

\thispagestyle{plain}
{\Large Interrogating theoretical models of neural computation with deep inference} \\
Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3}$, Chunyu A. Duan$^{4}$, Carlos D. Brody$^{2,3,5}$, \\
Kenneth D. Miller$^{1}$, and John P. Cunningham$^{6}$.

{\small
$^{1}$Department of Neuroscience, Columbia University, \\
$^{2}$Princeton Neuroscience Institute, \\
$^{3}$Princeton University, \\
$^{4}$Institute of Neuroscience, Chinese Academy of Sciences, \\
$^{5}$Howard Hughes Medical Institute, \\
$^{6}$Department of Statistics, Columbia University
}

\linenumbers
\section{Abstract}
A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or in terms of neural activity -- and thus can offer insights into neural computation.
The operation of these circuits, like all models, critically depends on the choices of model parameters.
Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.  
We bring recent machine learning techniques -- the use of deep generative models for probabilistic inference -- to bear on this problem, learning distributions of parameters that produce the specified properties of computation.   
Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task. 
More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.

\section{Introduction}
The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing \cite{abbott2008theoretical}.  
A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations.
The function of this system is governed by the choice of model parameters, which when configured in a particular way, give rise to a measurable signature of a computation.   
The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about these particular parameter configurations?  
The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.

Consider the idealized practice: one carefully designs a model and analytically derives how model parameters govern the computation.  
Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.  
This creates an unfavorable tradeoff.  On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.  
On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.

% now transition to ML...
Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.  
In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.  
One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data. 
Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating parameters (sometimes treated as hidden states) in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}. 
These works have used deep neural networks to expand the expressivity and accuracy of statistical models of neural data \cite{paninski2018neural}.

However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.  
First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.  
Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.  
Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model. 
Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.

% now we've constructed the tension that theory models need this and statistical models know how to do this... now connect!
To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.  
First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.    
Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.

Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to infinite neurons).
First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{gutierrez2013multiple, litwin2016inhibitory, duan2018collicular, mastrogiuseppe2018linking}, and more generally, these results point to the value of deep inference for the interrogation of biologically relevant models.

%-------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\subsection{Motivating emergent property inference of theoretical models} \label{results_motivating}

Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  The hub neuron (hub) couples with either the fast or slow population or both.

Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.  This emergent property is shown in Figure 1A at a frequency of 0.53Hz.

Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.  Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.  Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.

\begin{figure}
\begin{center}
\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
\end{center}
\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  A. For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.  In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.  B. The EPI distribution of STG model parameters producing network syncing.  Samples are colored by log probability density.  Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).  Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $v_1$ (solid) and $v_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.  Simulated activity is shown for three samples (stars). (Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  $v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  C. Deep probability distributions map a latent random variable $w$ through a deep neural network with weights and biases $\theta$ to parameters $z = f_\theta(w)$ distributed as $q_\theta(z)$. D. EPI optimization: To learn the EPI distribution $q_\theta(z)$ of model parameters that produce an emergent property, the emergent property statistics $T(x)$ are set in expectation over model parameter samples $z \sim q_\theta(z)$ and model simulations $x \sim p(x \mid z)$ to emergent property values $\mu$.}
 \label{fig:STG}
\end{figure}

Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig. \ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.  
This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of network syncing (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig. 1B, contours) and have diminished or preserved network syncing, respectively (Figure 1B inset and activity traces) (see Section \ref{methods_STG}).

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}

Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  In the running STG example, the model activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} C_m \frac{dx}{dt} = -h(x; z) = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right] 
\end{equation} 
where $C_m$=1nF, and $h_{\text{leak}}$, $h_{Ca}$, $h_K$, $h_{\text{hyp}}$, $h_{\text{elec}}$, and $h_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $x$ and $z = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).

Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(x)$, $\omega_{\text{f2}}(x)$, etc. -- is close to an intermediate frequency of 0.53Hz.  
Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(x)$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
\begin{equation}\label{eq:EP}
 \mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
 \begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
 \end{equation}
  which completes the quantification of the emergent property.

Third, we perform emergent property inference: we find a distribution over parameter configurations $z$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.  
This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_\theta(z) : \theta \in \Theta \right\}$, defined by a deep generative distribution
of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).  
This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
Then, mathematically, we must solve the following optimization program: 
 \begin{equation} \label{eq:EPI}
\begin{split}
&\argmax_{q_\theta \in \mathcal{Q}} H(q_\theta(z)) \\
 &\text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu, \\
\end{split}
\end{equation}
where $T(x), \mu$ are defined as in Equation \ref{eq:EP}, and $p(x|z)$ is the intractable distribution of data from the model, $x$, given that model's parameters $z$ (we access samples from this distribution by running the model forward).   The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness. The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.

EPI optimizes the weights and biases $\theta$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
The optimization is complete when the sampled models with parameters $z \sim q_\theta$ produce activity consistent with the specified emergent property (Fig. S4).
Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).

In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
We now prove out the value of EPI by using it to investigate and produce novel insights about three prominent models in neuroscience.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comprehensive input-responsivity in a nonlinear sensory system} \label{results_V1}

\begin{figure}
\begin{center}
\includegraphics[scale=0.54]{figures/fig2/fig2.pdf}
\end{center}
\caption{\small Hypothesis generation through EPI in a V1 model.  A. Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).  B. Linear response predictions become inaccurate with greater input strength.  V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$. Stars indicate the linear response prediction.  C. EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$. Supporting evidence for the four generated hypotheses are indicated by gray boxes with labels H1, H2, H3, and H4. The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin). }
 \label{fig:V1_EPI}
\end{figure}

Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}. 
Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig. \ref{fig:V1_EPI}A) \cite{pfeffer2013inhibition}.
Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools. 
Here, we go beyond linear theory by systematically generating and evaluating hypotheses of circuit model function using EPI distributions of neuron-type inputs producing various neuron-type population responses.

Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$. Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:

\begin{equation}
\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
\end{equation}

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.  
Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.

With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig. \ref{fig:V1_EPI}B right).  
Currently available approaches to deriving the steady state response of the system are limited.

To get a more comprehensive picture of the input-responsivity of each neuron-type beyond linear theory, we used EPI to learn a distribution of the differential inputs to each population $dh$ that produce an increase of $y$ in the rate of each neuron-type population $\alpha \in \{E, P, S, V \}$.  
We want to know the differential inputs $dh$ that result in a differential steady state $dx_{\alpha}$ (the change in $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$) of value $y$ with some small, arbitrarily chosen amount of variance  $0.01^2$.   
These statements amount to the emergent property 
\begin{equation}
\mathcal{B}(\alpha, y) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ (dx_{\alpha} - y)^2 \end{bmatrix} ~~=~~ \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
\end{equation}
We maintain the notation $\mathcal{B}(\cdot)$ throughout the rest of the study as short hand for emergent property, which represents a different signature of computation in each application.

Using EPI, we inferred the distribution of $dh$ shown in Figure \ref{fig:V1_EPI}C producing $\mathcal{B}(\alpha, y)$.  
Columns correspond to inferred distributions of excitatory ($\alpha=E$, red), parvalbumin ($\alpha=P$, blue), somatostatin ($\alpha=S$, red) and VIP ($\alpha=V$, green) neuron-type response increases, while each row corresponds to increase amounts of $y \in \{0.1, 0.5\}$.  For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_\theta(dh \mid \mathcal{B}(\alpha, y))$.  The inferred distributions immediately suggest four hypotheses: \\

{\addtolength{\leftskip}{10 mm}
H1: as is intuitive, each neuron-type's firing rate should be sensitive to that neuron-type's direct input (e.g. Fig. \ref{fig:V1_EPI}C H1 gray boxes indicate low variance in $dh_E$ when $\alpha=E$. Same observation in all inferred distributions); \\
H2: the E- and P-populations should be largely unaffected by input to the V-population (Fig. \ref{fig:V1_EPI}C H2 gray boxes indicate high variance in $dh_V$ when $\alpha \in \{E, P \}$); \\
H3: the S-population should be largely unaffected by input to the P-population  (Fig. \ref{fig:V1_EPI}C H3 gray boxes indicate high variance in $dh_P$ when $\alpha =S$); \\
H4: there should be a nonmonotonic response of the V-population with input to the E-population (Fig. \ref{fig:V1_EPI}C H4 gray boxes indicate that negative $dh_E$ should result in small $dx_{V}$, but positive $dh_E$ should elicit a larger $dx_{V}$);

}

\begin{figure}
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=4cm}}]{figure}[\FBwidth]
{\caption{Confirming EPI generated hypotheses in V1. A. Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$. B-D Same plots for the P-, S-, and V-populations.  Labels H1, H2, H3, and H4 indicate which curves confirm which hypotheses.}\label{fig:V1_HT}}
{\includegraphics[scale=0.6]{figures/fig3/fig3.pdf}}
\end{figure}
We evaluate these hypotheses by taking perturbations in individual neuron-type input $\delta h_\alpha$ away from the modes of the inferred distributions at $y=0.1$
\begin{equation}
dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)).
\end{equation}
Here $\delta x_{\alpha}$ is the change in steady state response of the system with input $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ compared to $h = b + dh^*$, where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$. The EPI-generated  hypotheses are confirmed (for details, see Section \ref{methods_V1}):

{\addtolength{\leftskip}{10 mm}
H1: the neuron-type responses are sensitive to their direct inputs (Fig. \ref{fig:V1_HT}A black, \ref{fig:V1_HT}B blue, \ref{fig:V1_HT}C red, \ref{fig:V1_HT}D green); \\
H2: the E- and P-populations are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
H3: the S-population is not affected by $\delta h_P$ (Fig. \ref{fig:V1_HT}C blue); \\
H4: the V-population exhibits a nonmonotonic response to $\delta h_E$ (Fig. \ref{fig:V1_HT}D black), and is in fact the only population to do so (Fig. \ref{fig:V1_HT}A-C black).

}

These hypotheses were in stark contrast to what was available to us via traditional analytical linear prediction (Fig. \ref{fig:V1_EPI}C, magenta, see Section \ref{methods_V1}).

Here, we examined the neuron-type responsivity of this model of V1 with scientifically motivated choice of connectivity $W$.
With EPI, we could just as easily have examined the distribution of such $W$'s consistent with some response characteristics for a fixed input $h$ or another emergent property such as inhibition stabilization.
Most importantly, this analysis is a proof-of-concept demonstrating the valuable ability to condition parameters of interest of a neural circuit model on some chosen emergent property.
To this point, we have shown the utility of EPI on relatively low-level emergent properties like network syncing and differential neuron-type population responses.  
In the remainder of the study, we focus on using EPI to understand models of more abstract cognitive function.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}a). Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
\begin{equation}
x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\epsilon = 0.05$ and $\zeta = 0.5$.  The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
\begin{equation}
\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
\end{equation}
with gaussian noise of variance $\sigma^2$ = 1.
The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
\end{center}
\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  A. Rapid task switching behavioral paradigm (see text). B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti.  Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  C. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy. E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
\label{fig:SC}
\end{figure}

To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
\begin{equation}
\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
\end{equation}
Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.

We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
To make sense of these inferred distributions, we followed the approach of Duan et al. by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig. \ref{fig:SC}C). These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair. 
The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.

We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig. \ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
These hypotheses are confirmed by forward simulation of the SC model (Fig. \ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Linking RNN connectivity to error} \label{results_RNN}
So far, each model we have studied was designed from fundamental biophysical principles, genetically- or functionally-defined neuron types.  
At a more abstract level of modeling, recurrent neural networks (RNNs) are high-dimensional dynamical models of computation that are becoming increasingly popular in neuroscience research \cite{barak2017recurrent}. 
In theoretical neuroscience, RNN dynamics usually follow the equation
\begin{equation}
\frac{dx}{dt} = -x + W \phi(x) + h,
\end{equation}
where $x$ is the network activity, $W$ is the network connectivity, $\phi(\cdot) = \tanh(\cdot)$, and $h$ is the input to the system.
Such RNNs are trained to do a task from a systems neuroscience experiment, and then the unit activations of the trained RNN are compared to recorded neural activity.
Fully-connected RNNs with tens of thousands of parameters are challenging to characterize \cite{sussillo2013opening}, especially making statistical inferences about their parameterization.
Alternatively, we considered a rank-1, $N$-neuron RNN with connectivity consisting of the sum of a random and a structured component:
\begin{equation}
W = g\chi + \frac{1}{N}mn^\top.
\end{equation}
The random component $g \chi$ has strength $g$, and random component weights are Gaussian distributed $\chi_{i,j} \sim \mathcal{N}(0, \frac{1}{N})$.  
The structured component $\frac{1}{N}m n\top$ has entries of $m$ and $n$ drawn from Gaussian distributions $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$.
Recent theoretical work derives the low-dimensional response properties of low-rank networks from statistical parameterizations of their connectivity, such as $z = [g, M_m, M_n]$ \cite{mastrogiuseppe2018linking}.
We used EPI to infer the parameterizations of rank-1 RNNs solving an example task, enabling discovery of properties of connectivity that result in different types of error in the computation.

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{figures/fig5/fig5.pdf}
\end{center}
\caption{\small Sources of error in an RNN solving a simple task.  A. (left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$. (right) Error in this computation can come from over- or under-estimating the posterior mean or variance. B. EPI distribution of rank-1 RNNs executing Gaussian posterior conditioning.  Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$  C.  Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.  Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.  D. The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.  Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
\label{fig:RNN}
\end{figure}

The task we consider is Gaussian posterior conditioning: calculate the parameters of a posterior distribution induced by a prior $p(\mu_y) = \mathcal{N}(\mu_0=4, \sigma_0^2=1)$ and a likelihood $p(y | \mu_y) = \mathcal{N}(\mu_y,\sigma^2_y = 1)$, given a single observation $y$.  
Conjugacy offers the result analytically; $p(\mu_y | y ) = \mathcal{N}\left(\mu_{post},\sigma_{post}^2\right)$, where:   
\begin{equation}
\mu_{\text{post}} = \frac{\frac{\mu_0}{\sigma_0^2} + \frac{y}{\sigma_y^2}}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}
~~~~~~~~~~
 \sigma^2_{\text{post}} = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}.
\end{equation} 
To solve this Gaussian posterior conditioning task, the RNN response to a constant input $h = y r + (n-M_n)$ must equal the posterior mean along readout vector $r$ , where
\begin{equation}
\kappa_r =  \frac{1}{N} \sum_{j=1}^N r_j \phi(x_j).
\end{equation}
Additionally, the amount of chaotic variance $\Delta_T$ must equal the posterior variance.
Theory for low-rank RNNs allows us to express $\kappa_r$ and $\Delta_T$ in terms of each other through a solvable system of nonlinear equations  (see Section \ref{methods_LRRNN}) \cite{mastrogiuseppe2018linking}.
This theory facilitates the mathematical formalization of task execution into an emergent property, where the emergent property statistics of the RNN activity are $\kappa_r$ and $\Delta_T$, and the emergent property values are the ground truth posterior mean $\mu_{\text{post}}$ and variance $\sigma^2_{\text{post}}$:
\begin{equation}
\mathbb{E} \begin{bmatrix} \kappa_r \\ \Delta_T \\ (\kappa_r-\mu_{\text{post}})^2 \\ (\Delta_T^2-\sigma^2_{\text{post}})^2 \end{bmatrix} ~~=~~ \begin{bmatrix} \mu_{\text{post}} \\ \sigma^2_{\text{post}} \\ 0.1 \\ 0.1 \end{bmatrix}.
\end{equation}
We chose a substantial amount of variance in these emergent property statistics, so that the inferred distribution resulted in RNNs with a variety of errors in their solutions to the gaussian posterior conditioning problem.

EPI was used to learn distributions of RNN connectivity properties $z = [g, M_m, M_n]$ executing Gaussian posterior conditioning given an input of $y=2$, where the true posterior is $\mu_{\text{post}}=3$ and $\sigma_{\text{post}} = 0.5$ (Fig. \ref{fig:RNN}A). 
We examined the nature of the over- and under-estimation of the posterior means (Fig. \ref{fig:RNN}B left) and variances (Fig. \ref{fig:RNN}B right) in the inferred distributions (300 samples).  
The symmetry in the $M_m$-$M_n$ plane, suggests a degeneracy in the product of $M_m$ and $M_n$ (Fig. \ref{fig:RNN}B).  
Indeed, $M_m M_n$ strongly determines the posterior mean ($r=0.62$,  $p<10^{-4}$). 
Furthermore, the random strength $g$ strongly determines the chaotic variance ($r=0.56$,  $p<10^{-4}$).
Neither of these observations were obvious from what mathematical analysis is available in networks of this type (see Section \ref{methods_LRRNN}).  
While the link between random strength $g$ and chaotic variance $\Delta_T$ (and resultingly posterior variance in this problem) is well-known \cite{sompolinsky1988chaos}, the distribution admits a novel hypothesis: the estimation of the posterior mean by the RNN increases with $M_m M_n$.

We tested this prediction by taking parameters $z_1$ and $z_2$ as representative samples from the positive and negative $M_m$-$M_n$ quadrants, respectively.
Instead of using the theoretical predictions shown in Figure \ref{fig:RNN}B, we simulated finite-size realizations of these networks with 2,000 neurons (e.g. Fig. \ref{fig:RNN}C).  
We perturbed these parameter choices by $M_m M_n$ clarifying that the posterior mean can be directly controlled in this way (Fig. \ref{fig:RNN}D; $p<10^{-4}$), see Section \ref{methods_LRRNN}).
Thus, EPI confers a clear picture of error in this computation: the product of the low rank vector means $M_m$ and $M_n$ modulates the estimated posterior mean while the random strength $g$ modulates the estimated posterior variance.
This novel procedure of inference on reduced parameterizations of RNNs conditioned on the emergent property of task execution is generalizable to other settings modeled in \cite{mastrogiuseppe2018linking} like noisy integration and context-dependent decision making (Fig. S5).

\section{Discussion}
\subsection{EPI is a general tool for theoretical neuroscience} 
Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).  
We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.

In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
Although, theoretical approaches (e.g. \cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.

\subsection{Novel hypotheses from EPI} 
In neuroscience, machine learning has primarily been used to reveal structure in large-scale neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}). 
 Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
 However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.  
In this work, we replaced this opaque procedure of parameter identification in theoretical neuroscience with emergent property inference, opening the door to careful inference in careful models of neural computation.

Biologically realistic models of neural circuits often prove formidable to analyze.
Two main factors contribute to the difficulty of this endeavor.
First, in most neural circuit models, the number of parameters scales quadratically with the number of neurons, limiting analysis of its parameter space. 
Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.

In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}. 
By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out. 
Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.

In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.

EPI leverages deep learning technology for neuroscientific inquiry in a categorically different way than approaches focused on training neural networks to execute behavioral tasks \cite{richards2019deep}. 
These works focus on examining optimized deep neural networks while considering the objective function, learning rule, and architecture used.
This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
These approaches can be used complementarily to enhance the practice of theoretical neuroscience.

\textbf{Acknowledgements}: \\
This work was funded by NSF Graduate Research Fellowship,  DGE-1644869, McKnight Endowment Fund, NIH NINDS 5R01NS100066, Simons Foundation 542963, NSF NeuroNex Award, DBI-1707398, The Gatsby Charitable Foundation, Simons Collaboration on the Global Brain Postdoctoral Fellowship, Chinese Postdoctoral Science Foundation, and International Exchange Program Fellowship. 
Helpful conversations were had with Francesca Mastrogiuseppe, Srdjan Ostojic, James Fitzgerald, Stephen Baccus, Dhruva Raman, Liam Paninski, and Larry Abbott.

\textbf{Data availability statement}: \\
The datasets generated during and/or analysed during the current study are available from the corresponding author upon reasonable request.

\textbf{Code availability statement}: \\
The software written for the current study is available from the corresponding author upon reasonable request.

\bibliography{eLife2020}
\bibliographystyle{unsrt}

\newpage

\section{Methods}

\subsection{Emergent property inference (EPI)}\label{methods_EPI}
Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.  
Consider model parameterization $z$ and data $x$ which has an intractable likelihood $p(x \mid z)$ defined by a model simulator of which samples are available $x \sim p(x \mid z)$.  
EPI optimizes a distribution $q_\theta(z)$ (itself parameterized by $\theta$) of model parameters $z$ to produce an emergent property of interest $\mathcal{B}$,
\begin{equation}
\mathcal{B} ~~\triangleq~~  \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
\end{equation}
Precisely, the emergent property statistics $T(x)$ must equal the emergent property values $\mu$, in expectation over the EPI distribution of parameters $q_\theta(z)$ and the distribution of simulated activity $p(x \mid z)$.  This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.

With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_\theta(z)$.
 In deep probability distributions, a simple random variable $w \sim q_0(w)$ is mapped deterministically via a sequence of deep neural network layers ($f_1$, .. $f_l$) parameterized by weights and biases $\theta$ to the support of the distribution of interest:
\begin{equation}
z = f_{\theta}(\omega) = f_l(..f_1(w)).
\end{equation}
Given a simulator defined by a theoretical model $x \sim p(x \mid z)$ and some emergent property of interest $\mathcal{B}$, $q_\theta(z)$ is optimized via the neural network parameters $\theta$ to find a maximally entropic distribution $q_{\theta}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
\begin{equation} \label{eq:opt}
\begin{split}
q_\theta^*(z) &= \argmax_{q_\theta \in Q} H(q_\theta(z)) \\
 &  \text{s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu. \\
 \end{split}
\end{equation} 
Since we are optimizing parameters $\theta$ of our deep probability distribution with respect to the entropy $H(q_\theta(z))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.  Entropy of $q_\theta(z)$ can be expressed as an expectation of the negative log density of parameter samples $z$ over the randomness in the parameterless initial distribution $q_0$:
\begin{equation}
H(q_\theta(z)) = \int - q_\theta(z) \log(q_\theta(z)) dz = \mathbb{E}_{z \sim q_\theta}\left[-\log(q_\theta(z)) \right] = \mathbb{E}_{w \sim q_0}\left[-\log(q_\theta(f_\theta(w))) \right].
\end{equation}
Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $z$:
\begin{equation}
\nabla_\theta H(q_\theta(z)) = \mathbb{E}_{w \sim q_0}\left[- \nabla_\theta \log(q_\theta(f_\theta(w))) \right].
\end{equation}
In EPI, MEFNs are purposed towards variational learning of model parameter distributions.  
A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(x \mid z)$.
Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
But, unlike LFVI, EPI uses a single deep network to learn a distribution and is optimized to produce an emergent property, rather than condition on data points.  Optimizing the EPI objective is a technological challenge, the details of which we elaborate in Section \ref{methods_AL_opt}.  
 Before going through those details, we ground this optimization in a toy example.

 We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.

\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
\begin{equation} 
\tau \frac{dx}{dt} = Ax
\end{equation}
with
\begin{equation}
A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
\end{equation}
To run EPI with the dynamics matrix elements as the free parameters $z = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(x)$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system. 
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
\begin{equation}
 \mathbb{E}\left[T(x) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \mu.
 \end{equation} 
\begin{figure}
\begin{center}
\includegraphics[scale=0.58]{figures/figS1/figS1.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S1: A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.  Dashed lines indicate the parameter axes. C. Entropy throughout the optimization.  At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.  D. Emergent property moments throughout optimization.  At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
\end{flushleft}
\end{figure}

Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $z$, since the eigenvalues can be calculated using the quadratic formula: 
\begin{equation}
\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
\end{equation}

Importantly, even though $\mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_\theta$ directly.  This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\mu$ of an exponential family distribution \cite{wainwright2008graphical}.  Instead, we used EPI to approximate this distribution (Fig. S1B). We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$. (see Section \ref{methods_NF}).

Even this relatively simple system has nontrivial (though intuitively sensible) structure in the parameter distribution.  To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
 Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS2/figS2.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S2: A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
\end{flushleft}
\end{figure}

To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
\begin{equation}
\text{imag}(\lambda_1) = \begin{cases}
                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
                             0 & \text{otherwise } \\
                         \end{cases}.
\end{equation}

When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
\begin{equation}
\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
\end{equation}
Since we constrained $\mathbb{E}_{z \sim q_\theta}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig. S2B). 
With $\sigma_{1,4} = \mathbb{E}_{z \sim q_\theta}(|a_1 a_4 - E_{q_\theta}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted). 
This validates the curved structure of the inferred distribution learned through EPI.  
We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS3/figS3.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S3: Sampled dynamical systems $z \sim q_\theta(z)$ and their simulated activity from $x(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability. A. Each dimension of the simulated trajectories throughout time.  B  The simulated trajectories in phase space.}
\end{flushleft}
\end{figure}

\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
To optimize $q_\theta(z)$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
\begin{equation} \label{eq:AL}
L(\theta; \eta, c) = -H(q_\theta) + \eta^\top R(\theta) + \frac{c}{2}||R(\theta)||^2
\end{equation}
where $R(\theta) = \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x) - \mu \right] \right]$, $\eta \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\mu| = |T(x)|$,  and $c$ is the penalty coefficient. 
These Lagrange multipliers are closely related to the natural parameters of exponential families (see Section \ref{methods_VI}).
Deep neural network weights and biases $\theta$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
$\eta$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\eta$, $c$) for a given number of stochastic optimization iterations. 
A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
The p-value of $\mathbb{E}[||R(\theta_{k+1})||] > \gamma \mathbb{E} \left[||R(\theta_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
The other update rule is $\eta_{k+1} = \eta_k + c_k \frac{1}{n} \sum_{i=1}^n (T(x^{(i)}) - \mu)$ given a batch size $n$.
Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.

The intention is that $c$ and $\eta$ start at values encouraging entropic growth early in optimization.  
With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
In the oscillating 2D LDS example, each augmented Lagrangian epoch ran for 2,000 iterations (Fig. S1C-D).
Notice the initial entropic growth, and subsequent reduction upon each update of $\eta$ and $c$.
The momentum parameters of the Adam optimizer were reset at the end of each augmented Lagrangian epoch.

Rather than starting optimization from some $\theta$ drawn from a randomized distribution, we found that initializing $q_\theta(z)$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.  The parameters of the Gaussain initialization were chosen on an application-specific basis.  
Throughout the study, we chose isotropic Gaussian initializations with mean $\mu_{\text{init}}$ at the center of the distribution support and some standard deviation $\sigma_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_V1}).

To assess whether EPI distribution $q_\theta(z)$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
The algorithm has converged when a null hypothesis test of constraint violations $R(\theta)_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.  
This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\theta)^i$ are made from $k$ resamplings of $z$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
$k$ is determined by a fraction of the batch size $\nu$, which varies according to the application.
In the linear two-dimensional system example, we used a batch size of $n = 1000$ and set $\nu = 0.1$ resulting in convergence after the ninth epoch of optimization. (Fig. S1C-D black dotted line).

When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
Next, one should consider the computational cost of the gradient calculations. 
In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_2DLDS}) for the emergent property statistic given the model parameters.  
On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g. Section \ref{methods_STG}).  In such cases, optimization will be expensive.

\subsubsection{Normalizing flows}\label{methods_NF}
Deep probability models typically consist of several layers of fully connected neural networks.  When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.  For $z' = f(z)$,

\begin{equation}
q(z') = q(f^{-1}(z')) \left| \det \frac{\partial f^{-1}(z')}{\partial z'} \right| = q(z) \left| \det \frac{\partial f(z)}{\partial z} \right|^{-1}.
\end{equation}

However, this computation has cubic complexity in dimensionality for fully connected layers.  By restricting our layers to normalizing flows \cite{rezende2015variational} -- bijective functions with fast log determinant Jacobian computations, we can tractably optimize deep generative models with objectives that are a function of sample density, like entropy. Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.

\subsubsection{Emergent property inference as variational inference in an exponential family}\label{methods_VI}
Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
In Bayesian inference a prior belief about model parameters $z$ is formalized into a prior distribution $p(z)$, and the statistical model capturing the effect of $z$ on observed data points $x$ is formalized in the likelihood distribution $p(x \mid z)$.
In Bayesian inference, we obtain a posterior distribution $p(z \mid x)$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
\begin{equation}
p(z \mid x) = \frac{p(x \mid z) p(z)}{p(x)}.
\end{equation}
The posterior disribution is analytically available when the prior is conjugate with the likehood.
However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.

As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions. A maximum entropy distribution of form:
\begin{equation}
\begin{split}
p^*(z) &= \argmax_{p \in \mathcal{P}} H(p(z)) \\
 &  \text{s.t.  } \mathbb{E}_{z \sim p}\left[T(z)\right] = \mu. \\
 \end{split}
\end{equation} 
will have probability density in the exponential family:
\begin{equation}
p^*(z) \propto \exp(\eta^\top T(z)).
\end{equation}
The mappings between the mean parameterization $\mu$ and the natural parameterization $\eta$ are formally hard to identify \cite{wainwright2008graphical}.

Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid x)$.  We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(x)$.  The log normalizing constant $A(\eta(x))$ remains unchanged:
\begin{equation}
\begin{split}
p(z \mid x) = b(z) \exp{\left( \eta(x)^\top T(z) - A(\eta(x)) \right)} = \exp{\left( \begin{bmatrix} \eta(x) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(x)) \right)} \\= \exp{\left(\tilde{\eta(x)}^\top \tilde{T}(z) - A(\eta(x)) \right)} 
\end{split}.
\end{equation}
Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
\begin{equation}
q_\theta^* = \argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)).
\end{equation}
$q_\theta(z)$ is the variational approximation to the posterior with variational parameters $\theta$.  We can write this KL divergence in terms of entropy of the variational approximation:
\begin{equation}
KL(q_\theta \mid \mid p(z \mid x)) = \mathbb{E}_{z \sim q_\theta} \left[ \log (q_\theta(z)) \right] - \mathbb{E}_{z \sim q_\theta} \left[ \log (p(z \mid x)) \right]
\end{equation}
\begin{equation}
 = -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) - A(\eta(x)) \right].
\end{equation}
As far as the variational optimization is concerned, the log normalizing constant is independent of $q_\theta(z)$, so it can be dropped
\begin{equation}
\argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)) =  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) \right].
\end{equation}
 Further, we can write the objective in terms of the first moment of the sufficient statistics $\mu = \mathbb{E}_{z \sim p(z \mid x)}\left[T(z) \right]$:
 \begin{equation}
=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right] + \tilde{\eta}(x)^\top \mu,
 \end{equation}
which simplifies to
  \begin{equation}
=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right].
 \end{equation}.

In comparison, in emergent property inference (EPI), we solve the following problem:
\begin{equation}
q_\theta^*(z) = \argmax_{q_\theta \in Q} H(q_\theta(z)),   \text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
\end{equation}
The Lagrangian objective (without augmentation) is
\begin{equation}
q_\theta^* = \argmin_{q_\theta \in Q} - H(q_\theta) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_\theta} \left[\tilde{T}(z) \right] - \mu \right).
\end{equation}
Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(x)$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).

We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
Accordingly, we replace the notation of $p(z \mid x)$ with $p(z \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).

\subsection{Theoretical models}\label{methods_theoretical_models}
In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  Here, we provide the details of each model  and the related analyses.

\subsubsection{Stomatogastric ganglion}\label{methods_STG}
We analyze how the parameters $z = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right].
\end{equation} 
The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization,
electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $z$.  
The capacitance of the cell membrane was set to $C_m = 1nF$. Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
\begin{equation}  h_{leak}(x; z) = g_{leak} (x_\alpha - V_{leak}) 
\end{equation} 
\begin{equation}  h_{elec}(x; z) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
\end{equation} 
\begin{equation}  h_{syn}(x; z) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
\begin{equation}  h_{Ca}(x; z) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
\end{equation} 
\begin{equation}  h_K(x; z) = g_K N (x_\alpha - V_K) 
\end{equation} 
\begin{equation}  h_{hyp}(x; z) = g_h H(x_\alpha - V_{hyp}).
\end{equation} 
The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.  Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.

Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
 \end{equation}
where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.

Finally, there is a synaptic gating variable as well:
\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
\end{equation}
When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.

In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.  In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible. The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.  Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$

Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.  High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.  Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.  The signal was subsequently mean centered.  After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.

Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.  Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.  The ``frequency identification" vector is 
\begin{equation}
v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
\end{equation}
The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.

Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.  The emergent property statistics are the first and second moments of the firing frequencies. The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
\begin{equation}
E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
\end{equation}
for the EPI distribution shown in Fig. \ref{fig:STG}B.
Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  The optimization converged after four epochs (Fig. S4).

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{figures/figS4/figS4.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S4: EPI optimization of the STG model producing network syncing. A. Entropy throughout optimization. B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  Since $q_\theta(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.    C. The second moment emergent property statistics converge to the emergent property values.}
\end{flushleft}
\end{figure}

For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $z = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$. We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.

We calculated the Hessian at the mode of the inferred EPI distribution.  
The Hessian of a probability model is the second order gradient of the log probability density $\log q_\theta(z)$ with respect to the parameters $z$:  $\frac{\partial^2 \log q_\theta(z)}{\partial z \partial z^\top}$.  
With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced. 
In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig. \ref{fig:STG}B, inset).
The contours of Figure \ref{fig:STG} were calculated as error in $T(x)$ from $\mu$ in both the first and second moment emergent property statistics.

\subsubsection{Primary visual cortex}\label{methods_V1}
The dynamics of each neural populations average rate
$x = [ x_E, x_P, x_S, x_V]^\top$
are given by:
\begin{equation}
\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
\end{equation}

By consolidating information from many experimental datasets, Billeh et al. \cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
\begin{equation}
M = \begin{bmatrix} 0.36 & 0.48 & 0.31 & 0.28 \\
1.49 & 0.68 & 0.50 & 0.18 \\
0.86 & 0.42 & 0.15 & 0.32 \\
1.31 & 0.41 & 0.52 & 0.37 \end{bmatrix}
\end{equation}
and connection probability
\begin{equation}
C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
0.395 & .451 & 0.857 & 0.02 \\
0.182 & 0.03 & 0.082 & 0.625 \\
0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
\end{equation}

Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
\begin{equation}
W_{\text{full}} = C \odot M = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
0.395 & .451 & 0.857 & 0.02 \\
0.182 & 0.03 & 0.082 & 0.625 \\
0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
\end{equation}

Theoretical work on these systems considers a subset of the effective connectivities \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}
\begin{equation} \label{eq:V1_W}
W = \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & 0 \\
                                W_{PE} & W_{PP} & W_{PS} & 0 \\
                                W_{SE} & 0 & 0 & W_{SV} \\
                                W_{VE} & W_{VP} &  W_{VS} &  0 \end{bmatrix}.
\end{equation}
In coherence with this work, we only keep the entries of $W_{\text{full}}$ corresponding to parameters in Equation \ref{eq:V1_W}.

We look at how this four-dimensional nonlinear dynamical model of V1 responds to different inputs, and compare the predictions of the linear response to the approximate posteriors obtained through EPI.  The input to the system is the sum of a baseline input $b = [1, 1, 1, 1]^\top$ and a differential input $dh$:
\begin{equation}
h = b + dh.
\end{equation}
All simulations of this system had $T=100$ time points, a time step $dt = 5$ms, and time constant $\tau = 20$ms.  The system was initialized to a random draw $x(0)_i \sim \mathcal{N}(1, 0.01)$.

We can describe the dynamics of this system more generally by
\begin{equation}
\dot{x}_i = -x_i + f(u_i)
\end{equation}
where the input to each neuron is
\begin{equation}
u_i = \sum_j W_{ij} x_j + h_i.
\end{equation}
Let $F_{ij} = \gamma_i \delta(i,j)$, where $\gamma_i = f'(u_i)$.  Then, the linear response is
\begin{equation}
\frac{dx_{ss}}{dh} = F(W\frac{dx_{ss}}{dh} + I)
\end{equation}
which is calculable by
\begin{equation}
\frac{dx_{ss}}{dh} = (F^{-1} - W)^{-1}.
\end{equation}
This calculation is used to produce the magenta lines in Figure \ref{fig:V1_EPI}C, which show the linearly predicted inputs that generate a response from two standard deviations (of $\mathcal{B}$) below and above $y$.

The emergent property we considered was the first and second moments of the change in steady state rate $dx_{ss}$ between the baseline input $h= b$ and $h = b + dh$.  We use the following notation to indicate that the emergent property statistics were set to the following values:
\begin{equation}
\mathcal{B}(\alpha, y) ~~\triangleq~~ 
\mathbb{E} \begin{bmatrix} dx_{\alpha,ss} \\ (dx_{\alpha,ss} - y)^2 \end{bmatrix} = \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
\end{equation}

In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)$.
The differential responses $\delta x_{\alpha,ss}$ are examined at perturbed inputs  $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$ and $\delta x$ is evaluated at 101 equally spaced samples of $\delta h_\alpha$ from -15 to 15.

We measured the linear regression slope between neuron-types of $\delta x$ and $\delta h$ to confirm the hypotheses H1-H3 (H4 is simply observing the nonmonotonicity) and report the p values for tests of non-zero slope.

{\addtolength{\leftskip}{10 mm}
H1: the neuron-type ressponses are sensitive to their direct inputs. 
E-population: $\beta=1.62$, $p<10^{-4}$ (Fig. \ref{fig:V1_HT}A black),
P-population: $\beta=1.06$, $p<10^{-4}$ (Fig. \ref{fig:V1_HT}B blue), S-population: $\beta=6.80$, $p<10^{-4}$  (Fig. \ref{fig:V1_HT}C red), V-population: $\beta=6.41$, $p<10^{-4}$  (Fig. \ref{fig:V1_HT}D green). \\
H2: the E-population ($\beta=0$, $p=1$) and P-populations ($\beta=0$, $p=1$) are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
H3: the S-population is not affected by $\delta h_P$ ($\beta=0$, $p=1$) (Fig. \ref{fig:V1_HT}C blue); \\

}

For each $\mathcal{B}(\alpha, y)$ with $\alpha \in \{E, P, S, V\}$ and $y \in \{0.1, 0.5\}$, we ran EPI using a real NVP architecture of four masks layers with two hidden layers of 10 units,  mapped to a support of $z_i \in \left[-5, 5\right]$ with no batch normalization.  We used an augmented Lagrangian coefficient of $c_0 = 10^5$, a batch size $n=1000$, set $\nu = 0.5$.
The EPI distributions shown in Fig. \ref{fig:V1_EPI} are the converged distributions with maximum entropy across random seeds.

We set the parameters of the Gaussian initialization $\mu_\text{init}$ and $\Sigma_\text{init}$ to the mean and covariance of random samples $z^{(i)} \sim \mathcal{U}(-5, 5)$ that produced emergent property statistic $dx_{\alpha,ss}$ within a bound $\epsilon$ of the emergent property value $y$. $\epsilon=0.01$ was set to be one standard deviation of the emergent property value according to the emergent property value $0.01^2$ of the variance emergent property statistic.

\subsubsection{Superior colliculus}\label{methods_SC}
In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
\begin{equation}
x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\alpha \in \{LP, LA, RA, RP\}$ $\epsilon = 0.05$ and $\zeta = 0.5$ control the position and shape of the nonlinearity, respectively.

We order the elements of $x$ and $u$ in the following manner
\begin{equation}
x = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} u = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix}.
\end{equation}

 The internal variables follow dynamics:
\begin{equation}
\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
\end{equation}
with time constant $\tau = 0.09s$ and Gaussian noise $\sigma dB$ controlled by the magnitude of $\sigma=1.0$.  The weight matrix has 8 parameters $sW_P$, $sW_A$, $vW_{PA}$, $vW_{AP}$, $hW_P$, $hW_A$, $dW_{PA}$, and $dW_{AP}$ (Fig. 4B):
\begin{equation}
W = \begin{bmatrix} sW_P & vW_{PA} & hW_P & dW_{PA}  \\ vW_{AP}  & sW_A & dW_{AP}  & hW_A \\ hW_P & dW_{PA}  & sW_P & vW_{PA}  \\ dW_{AP}  & hW_A & vW_{AP}  & sW_A \end{bmatrix}.
\end{equation}

The system receives five inputs throughout each trial, which has a total length of 1.8s.
\begin{equation}
h = h_{\text{rule}} + h_{\text{choice-period}} + h_{\text{light}}.
\end{equation}

There are rule-based inputs depending on the condition,
\begin{equation}h_{\text{P,rule}}(t) = \begin{cases}
                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
                            0,              & \text{otherwise}
                         \end{cases}
\end{equation}
\begin{equation} h_{\text{A,rule}}(t) = \begin{cases}
                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
                            0,              & \text{otherwise}
                         \end{cases}
\end{equation}
a choice-period input,
\begin{equation} h_{\text{choice}}(t) = \begin{cases}
                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
                            0,              & \text{otherwise}
                         \end{cases}
\end{equation}
and an input to the right or left-side depending on where the light stimulus is delivered.     
\begin{equation}  h_{\text{light}}(t) = \begin{cases}
                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } t > 1.2s \text{ and Left} \\
                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } t > 1.2s \text{ and Right} \\
                            0,              & t \leq 1.2s
                         \end{cases} .
\end{equation}
The input parameterization was fixed to $I_{\text{P,rule}} = 10 $,  $I_{\text{A,rule}} = 10$,  $I_{\text{choice}} = 2$,  and $I_{\text{light}} = 1$.

To produce an accuracy rate of $p_{LP}$ in the Left, Pro condition, let $\hat{p}_i$ be the empirical average steady state response (final $x_{LP}$ at end of task) over M=500 Gaussian noise draws for a given SC model parameterization $z_i$:
\begin{equation}
 \hat{p}_i = \mathbb{E}_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] = \frac{1}{M}\sum_{j=1}^M x_{LP}(s=L, c=P, z=z_i, \sigma dB_j)
 \end{equation}
 where stimulus $s \in \{L, R\}$, cue $c \in \{P, A\}$, and $\sigma dB_j$ is the Gaussian noise on trial $j$.
As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
For the first emergent property statistic, the average over EPI samples (from $q_\theta(z)$) is set to the desired value $p_{LP}$:
\begin{equation}
\mathbb{E}_{z_i \sim q_\phi} \left[ \mathbb{E}_{\sigma dB} \left[ x_{LP,\text{ss}} \mid s=L, c=P, z=z_i \right] \right] = \mathbb{E}_{z_i \sim q_\phi} \left[ \hat{p}_i \right] = p_{LP}.
\end{equation}

For the next emergent property statistic, we ask that the variance of the steady state responses across Gaussian draws, is the Bernoulli variance for the empirical rate $\hat{p}_i$:
\begin{equation}
\mathbb{E}_{z \sim q\phi} \left[ \sigma^2_{err} \right] = 0
\end{equation}
where the Bernoulli variance error $\sigma^2_{err}$ for the Pro task, left condition is
\begin{equation}
\sigma^2_{err} = Var_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] - \hat{p}_i(1 - \hat{p}_i).
\end{equation}

We have an additional constraint that the Pro neuron on the opposite hemisphere should have the opposite value (0 and 1).  We can enforce this with another constraint:
\begin{equation}
\mathbb{E}_{z \sim q\phi} \left[ d_P \right] = 1,
\end{equation}
where the distance between Pro neuron steady states $d_P$ in the Pro condition is
\begin{equation}
d_P =  \mathbb{E}_{\sigma dB} \left[ (x_{LP} - x_{RP})^2  \mid s=L, c=P, z=z_i \right]
\end{equation}
The emergent property statistics only need to be measured during the Left stimulus condition of the Pro and Anti tasks, since the network is symmetrically parameterized.
In total, the emergent property of rapid task switching at accuracy level $p$ was defined as
\begin{equation}
\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
\end{equation}

Since the maximum variance of a random variable bounded from 0 to 1 is the Bernoulli variance $\hat{p}(1-\hat{p})$, and the maximum squared difference between to variables bounded from 0 to 1 is 1, we do not need to control the second moment of these test statistics. 
These variables are dynamical system states and can only exponentially decay (or saturate) to 0 (or 1), so the Bernoulli variance error and squared difference constraints cannot be satisfied exactly in simulation.  
This is important to be mindful of when evaluating the convergence criteria.  
Instead of using our usual hypothesis testing criteria for convergence to the emergent property, we set a slack variable threshold only for these technically infeasible emergent property values to 0.05.

Using EPI to learn distributions of dynamical systems producing Bernoulli responses at a given rate (with small variance around that rate) was more challenging than expected.  
There is a pathology in this optimization setup, where the learned distribution of weights is bimodal attributing a fraction $p$ of the samples to an expansive mode (which always sends $x_{LP}$ to 1), and a fraction $1-p$ to a decaying mode (which always sends $x_{LP}$ to 0).  
This pathology was avoided using an inequality constraint prohibiting parameter samples that resulted in low variance of responses across noise.

\begin{table}[h]
\begin{tabular}{l|l|l|l|l}
$\lambda$  & $\hat{p}$ & $q_\theta(z)$ & $r$ & p-value \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $7.56 \times 10^{-01}$ & $p<10^{-4}$ \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $4.59 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $3.76 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline

$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $4.80 \times 10^{-02}$ & $p<.01$ \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $2.08 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $4.84 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(90\%))$ & $4.25 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline

$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(50\%))$ & $-7.57 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $-6.73 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $-4.86 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $-1.43 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $-1.93 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline

$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $-7.60 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $-2.73 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $-2.74 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
\end{tabular}
\caption{\small Table of significant correlation values from Fig. \ref{fig:SC}E.}
\end{table}

For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.

We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  Correlations were measured from 5,000 samples of $q_\theta(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.

\subsubsection{Rank-1 RNN}\label{methods_LRRNN}
Extensive research on random fully-connected recurrent neural networks has resulted in foundational theories of their activity \cite{sompolinsky1988chaos, brunel2000dynamics}.
Furthermore, independent research on training these models to perform computations suggests that learning occurs through low-rank perturbations to the connectivity (e.g. \cite{jaeger2004harnessing, sussillo2009generating}).
Recent theoretical work extends theory for random neural networks \cite{sompolinsky1988chaos} to those with added low-rank structure \cite{mastrogiuseppe2018linking}.
In Section \ref{results_RNN}, we used this theory to enable EPI on RNN parameters conditioned on the emergent property of task execution.

Such RNNs have the following dynamics: 
\begin{equation}
\frac{dx}{dt} = -x + W \phi(x) + h,
\end{equation}
where $x$ is network activity, $W$ is the connectivity weight matrix, $\phi(\cdot) = \tanh(\cdot)$ is the input-output function, and $h$ is the input to the system.
In a rank-1 RNN (which was sufficiently complex for the Gaussian posterior conditioning task), $W$ is the sum of a random component with strength $g$ and a structured component determined by the outer product of vectors $m$ and $n$:
\begin{equation}
W = g\chi + \frac{1}{N}mn^\top,
\end{equation}
where  $\chi_{ij} \sim \mathcal{N}(0, \frac{1}{N})$, and the entries of $m$ and $n$ are distributed as $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$. 
For EPI, we consider $z = [g, M_m, M_n]$, which are the parameters governing the connectivity properties of the RNN.

From such a parameterization $z$, the theory of Mastrogiuseppe et al. produces solutions for variables describing the low dimensional response properties of the RNN.
These ``dynamic mean field" (DMF) variables (e.g. the activity along a vector $\kappa_v$, the total variance $\Delta_0$, structured variance $\Delta_\infty$, and the chaotic variance $\Delta_T$) are derived to be functions of one another and connectivity parameters $z$.
The collection of these derived functions results in a system of equations, whose solution must be obtained through a nonlinear system of equations solver.
The iterative steps of this system of equations solver are differentiable, so we take gradients through this solve process.
The DMF variables provide task-relevant information about the RNN's response to task inputs.

In the Gaussian posterior conditioning example, $\kappa_r$ and $\Delta_T$ are DMF variables used as task-relevant emergent property statistics $\mu_{\text{post}}$ and $\sigma^2_{\text{post}}$.
Specifically, we solve for the DMF variables $\kappa_r$, $\kappa_n$, $\Delta_0$ and $\Delta_\infty$, where the readout is nominally chosen to point in the unit orthant $r = [1, ..., 1]^\top$.  The consistency equations for these variables in the presence of a constant input $h = y r - (n - M_n)$ can be derived following \cite{mastrogiuseppe2018linking}:
{\small
\begin{equation}
\begin{split}
\kappa_r = G_1(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_m \kappa_n + y \\
\kappa_n = G_2(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_n \langle \left[ \phi_i \right] \rangle + \langle \left[ \phi_i' \right] \rangle \\
\frac{\Delta_0^2-\Delta_\infty^2}{2} = G_3(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = g^2 \left( \int \mathcal{D}z \Phi^2(\kappa_r + \sqrt{\Delta_0}z) - \int \mathcal{D}z \int \mathcal{D}x \Phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z)  \right) \\
+ (\kappa_n^2  + 1)(\Delta_0-\Delta_\infty) \\
\Delta_\infty = G_4(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty)  = g^2 \int \mathcal{D}z \left[ \int \mathcal{D}x \phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z \right]^2 + \kappa_n^2 + 1
\end{split} 
\end{equation}
}
where here $z$ is a gaussian integration variable. We can solve these equations by simulating the following Langevin dynamical system to a steady state:

\begin{equation}
\begin{split}
l(t) = \frac{\Delta_0(t)^2-\Delta_\infty(t)^2}{2} \\
\Delta_0(t) = \sqrt{2l(t) + \Delta_\infty(t)^2} \\
\frac{d \kappa_r(t)}{dt}  = -\kappa_r(t) + G_1(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
\frac{d \kappa_n(t)}{dt} = -\kappa_n + G_2(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
\frac{d l(t)}{dt}  = -l(t) + G_3(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
\frac{d \Delta_\infty(t)}{dt}  = -\Delta_\infty(t) + G_4(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t))
\end{split}.
\end{equation}
Then, the chaotic variance, which is necessary for the Gaussian posterior conditioning example, is simply calculated via $\Delta_T = \Delta_0 - \Delta_\infty$.

We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization. 
We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
 The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.

To examine the effect of product $M_m M_n$ on the posterior mean, $\mu_{\text{post}}$ we took perturbations in $M_m M_n$ away from two representative parameters  $z_1$ and $z_2$ in 21 equally space increments from -1 to 1.  For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.  In Fig. \ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.  The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.  For perturbations away from $z_1$ the correlation was 0.995 with $p<10^{-4}$, and for perturbations away from $z_2$ the correlation was 0.983 with $p<10^{-4}$.

In addition to the Gaussian posterior conditioning example in Section \ref{results_RNN}, we modeled two tasks from Mastrogiuseppe et al.: noisy detection and context-dependent discrimination.  We used the same theoretical equations and task setups described in their study.

%TODO Need to explain the warm starting for the aficionados.

%TODO explain the density network architectures used.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
\end{center}
{\small Fig. S5: A. EPI for rank-1 networks doing noisy discrimination. B. EPI for rank-2 networks doing context-dependent discrimination.  See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
\end{figure}

\end{document}

### 2020-11-06 written/SC_model/SC_model.tex
SC model written

--- 2020-11-02 written/eLifeRevision/Bittner_eLife2020.tex
@@
+% You should title the file with a .tex extension (hw1.tex, for example)
 \documentclass[11pt]{article}
 
+\usepackage{hyperref}
 \usepackage{amsmath}
@@
 \usepackage{tikz}
+\usepackage{systeme}
 \usepackage{graphicx}
-\usepackage{lineno}
-\usepackage{floatrow}
-
 \DeclareMathOperator*{\argmin}{argmin}
@@
 \oddsidemargin0cm
-\topmargin-2cm
-\textwidth16.5cm
+\topmargin-2cm     %I recommend adding these three lines to increase the 
+\textwidth16.5cm   %amount of usable space on the page (and save trees)
 \textheight23.5cm
@@
 \renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}
-\linespread{1.5}
+
+\newcommand{\myname}{Sean Bittner}
+\newcommand{\myandrew}{srb2201@columbia.edu}
+\newcommand{\myhwnum}{12}
 
@@
 \DeclarePairedDelimiter\abs{\lvert}{\rvert}%
+ %
+\pagestyle{fancyplain}
+\rhead{\fancyplain{}{\myname\\ \myandrew}}
 
 \begin{document}
-\medskip
+
+\medskip                        % Skip a "medium" amount of space
+                                % (latex determines what medium is)
+                                % Also try: \bigskip, \littleskip
 
 \thispagestyle{plain}
-{\Large Interrogating theoretical models of neural computation with deep inference} \\
-Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3}$, Chunyu A. Duan$^{4}$, Carlos D. Brody$^{2,3,5}$, \\
-Kenneth D. Miller$^{1}$, and John P. Cunningham$^{6}$.
+\begin{center}                  % Center the following lines
+{\Large Draft of new SC section} \\
+Sean Bittner \\
+November 4, 2020 \\
+\end{center}
 
-{\small
-$^{1}$Department of Neuroscience, Columbia University, \\
-$^{2}$Princeton Neuroscience Institute, \\
-$^{3}$Princeton University, \\
-$^{4}$Institute of Neuroscience, Chinese Academy of Sciences, \\
-$^{5}$Howard Hughes Medical Institute, \\
-$^{6}$Department of Statistics, Columbia University
-}
-
-\linenumbers
-\section{Abstract}
-A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
-Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or in terms of neural activity -- and thus can offer insights into neural computation.
-The operation of these circuits, like all models, critically depends on the choices of model parameters.
-Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
-However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-We bring recent machine learning techniques -- the use of deep generative models for probabilistic inference -- to bear on this problem, learning distributions of parameters that produce the specified properties of computation.
-Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
-We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
-We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
-More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
-
-\section{Introduction}
-The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing \cite{abbott2008theoretical}.
-A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations.
-The function of this system is governed by the choice of model parameters, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about these particular parameter configurations?  
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
-
-Consider the idealized practice: one carefully designs a model and analytically derives how model parameters govern the computation.  
-Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
-Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.  
-This creates an unfavorable tradeoff.
-On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
-On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-
-% now transition to ML...
-Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.  
-In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
-One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating parameters (sometimes treated as hidden states) in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
-These works have used deep neural networks to expand the expressivity and accuracy of statistical models of neural data \cite{paninski2018neural}.
-
-However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.  
-First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.  
-Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
-Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
-Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model. 
-Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
-
-% now we've constructed the tension that theory models need this and statistical models know how to do this... now connect!
-To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
-This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
-First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
-Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
-Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
-Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to infinite neurons).
-First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
-The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{gutierrez2013multiple, litwin2016inhibitory, duan2018collicular, mastrogiuseppe2018linking}, and more generally, these results point to the value of deep inference for the interrogation of biologically relevant models.
-
-%-------------------
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Results}
-\subsection{Motivating emergent property inference of theoretical models} \label{results_motivating}
-
-Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
-First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
-To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
-Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
-A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
-Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).
-The hub neuron (hub) couples with either the fast or slow population or both.
-
-Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
-To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-This emergent property is shown in Figure 1A at a frequency of 0.53Hz.
-
-Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
-Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.
-In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
-\end{center}
-\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.
-A.
-For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.
-In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$.
-Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.
-Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.
-B.
-The EPI distribution of STG model parameters producing network syncing.
-Samples are colored by log probability density.
-Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
-Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $v_1$ (solid) and $v_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-Simulated activity is shown for three samples (stars).
-(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.
-$v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).
-C.
-Deep probability distributions map a latent random variable $w$ through a deep neural network with weights and biases $\theta$ to parameters $z = f_\theta(w)$ distributed as $q_\theta(z)$.
-D.
-EPI optimization: To learn the EPI distribution $q_\theta(z)$ of model parameters that produce an emergent property, the emergent property statistics $T(x)$ are set in expectation over model parameter samples $z \sim q_\theta(z)$ and model simulations $x \sim p(x \mid z)$ to emergent property values $\mu$.}
- \label{fig:STG}
-\end{figure}
-
-Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
-This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
-The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of network syncing (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig.
-1B, contours) and have diminished or preserved network syncing, respectively (Figure 1B inset and activity traces) (see Section \ref{methods_STG}).
-
-\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
-
-Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
-First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.
-In the running STG example, the model activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
-\begin{equation} C_m \frac{dx}{dt} = -h(x; z) = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right] 
-\end{equation} 
-where $C_m$=1nF, and $h_{\text{leak}}$, $h_{Ca}$, $h_K$, $h_{\text{hyp}}$, $h_{\text{elec}}$, and $h_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $x$ and $z = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).
-
-Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(x)$, $\omega_{\text{f2}}(x)$, etc.
--- is close to an intermediate frequency of 0.53Hz.
-Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(x)$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-\begin{equation}\label{eq:EP}
- \mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
- \begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
- \end{equation}
-  which completes the quantification of the emergent property.
-
-Third, we perform emergent property inference: we find a distribution over parameter configurations $z$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_\theta(z) : \theta \in \Theta \right\}$, defined by a deep generative distribution
-of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
-Then, mathematically, we must solve the following optimization program: 
- \begin{equation} \label{eq:EPI}
-\begin{split}
-&\argmax_{q_\theta \in \mathcal{Q}} H(q_\theta(z)) \\
- &\text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu, \\
-\end{split}
-\end{equation}
-where $T(x), \mu$ are defined as in Equation \ref{eq:EP}, and $p(x|z)$ is the intractable distribution of data from the model, $x$, given that model's parameters $z$ (we access samples from this distribution by running the model forward).
-The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
-Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
-This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
-Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness.
-The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.
-
-EPI optimizes the weights and biases $\theta$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-The optimization is complete when the sampled models with parameters $z \sim q_\theta$ produce activity consistent with the specified emergent property (Fig. S4).
-Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
-Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).
-
-In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
-Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
-Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
-We now prove out the value of EPI by using it to investigate and produce novel insights about three prominent models in neuroscience.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comprehensive input-responsivity in a nonlinear sensory system} \label{results_V1}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.54]{figures/fig2/fig2.pdf}
-\end{center}
-\caption{\small Hypothesis generation through EPI in a V1 model.
-A.
-Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.
-Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
-B.
-Linear response predictions become inaccurate with greater input strength.
-V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$.
-Stars indicate the linear response prediction.
-C.
-EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$.
-Supporting evidence for the four generated hypotheses are indicated by gray boxes with labels H1, H2, H3, and H4.
-The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin).
-}
- \label{fig:V1_EPI}
-\end{figure}
-
-Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
-Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1_EPI}A) \cite{pfeffer2013inhibition}.
-Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we go beyond linear theory by systematically generating and evaluating hypotheses of circuit model function using EPI distributions of neuron-type inputs producing various neuron-type population responses.
-
-Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:
-
-\begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
-\end{equation}
-
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
-The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
-Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.
-
-With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
-Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
-All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
-While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
-\ref{fig:V1_EPI}B right).
-Currently available approaches to deriving the steady state response of the system are limited.
-
-To get a more comprehensive picture of the input-responsivity of each neuron-type beyond linear theory, we used EPI to learn a distribution of the differential inputs to each population $dh$ that produce an increase of $y$ in the rate of each neuron-type population $\alpha \in \{E, P, S, V \}$.
-We want to know the differential inputs $dh$ that result in a differential steady state $dx_{\alpha}$ (the change in $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$) of value $y$ with some small, arbitrarily chosen amount of variance  $0.01^2$.
-These statements amount to the emergent property 
-\begin{equation}
-\mathcal{B}(\alpha, y) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ (dx_{\alpha} - y)^2 \end{bmatrix} ~~=~~ \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
-\end{equation}
-We maintain the notation $\mathcal{B}(\cdot)$ throughout the rest of the study as short hand for emergent property, which represents a different signature of computation in each application.
-
-Using EPI, we inferred the distribution of $dh$ shown in Figure \ref{fig:V1_EPI}C producing $\mathcal{B}(\alpha, y)$.  
-Columns correspond to inferred distributions of excitatory ($\alpha=E$, red), parvalbumin ($\alpha=P$, blue), somatostatin ($\alpha=S$, red) and VIP ($\alpha=V$, green) neuron-type response increases, while each row corresponds to increase amounts of $y \in \{0.1, 0.5\}$.
-For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_\theta(dh \mid \mathcal{B}(\alpha, y))$.
-The inferred distributions immediately suggest four hypotheses: \\
-
-{\addtolength{\leftskip}{10 mm}
-H1: as is intuitive, each neuron-type's firing rate should be sensitive to that neuron-type's direct input (e.g.
-Fig.
-\ref{fig:V1_EPI}C H1 gray boxes indicate low variance in $dh_E$ when $\alpha=E$.
-Same observation in all inferred distributions); \\
-H2: the E- and P-populations should be largely unaffected by input to the V-population (Fig. \ref{fig:V1_EPI}C H2 gray boxes indicate high variance in $dh_V$ when $\alpha \in \{E, P \}$); \\
-H3: the S-population should be largely unaffected by input to the P-population  (Fig. \ref{fig:V1_EPI}C H3 gray boxes indicate high variance in $dh_P$ when $\alpha =S$); \\
-H4: there should be a nonmonotonic response of the V-population with input to the E-population (Fig.
-\ref{fig:V1_EPI}C H4 gray boxes indicate that negative $dh_E$ should result in small $dx_{V}$, but positive $dh_E$ should elicit a larger $dx_{V}$);
-
-}
-
-\begin{figure}
-\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=4cm}}]{figure}[\FBwidth]
-{\caption{Confirming EPI generated hypotheses in V1.
-A.
-Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$.
-B-D Same plots for the P-, S-, and V-populations.
-Labels H1, H2, H3, and H4 indicate which curves confirm which hypotheses.}\label{fig:V1_HT}}
-{\includegraphics[scale=0.6]{figures/fig3/fig3.pdf}}
-\end{figure}
-We evaluate these hypotheses by taking perturbations in individual neuron-type input $\delta h_\alpha$ away from the modes of the inferred distributions at $y=0.1$
-\begin{equation}
-dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)).
-\end{equation}
-Here $\delta x_{\alpha}$ is the change in steady state response of the system with input $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ compared to $h = b + dh^*$, where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$.
-The EPI-generated  hypotheses are confirmed (for details, see Section \ref{methods_V1}):
-
-{\addtolength{\leftskip}{10 mm}
-H1: the neuron-type responses are sensitive to their direct inputs (Fig. \ref{fig:V1_HT}A black, \ref{fig:V1_HT}B blue, \ref{fig:V1_HT}C red, \ref{fig:V1_HT}D green); \\
-H2: the E- and P-populations are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
-H3: the S-population is not affected by $\delta h_P$ (Fig. \ref{fig:V1_HT}C blue); \\
-H4: the V-population exhibits a nonmonotonic response to $\delta h_E$ (Fig. \ref{fig:V1_HT}D black), and is in fact the only population to do so (Fig. \ref{fig:V1_HT}A-C black).
-
-}
-
-These hypotheses were in stark contrast to what was available to us via traditional analytical linear prediction (Fig. \ref{fig:V1_EPI}C, magenta, see Section \ref{methods_V1}).
-
-Here, we examined the neuron-type responsivity of this model of V1 with scientifically motivated choice of connectivity $W$.
-With EPI, we could just as easily have examined the distribution of such $W$'s consistent with some response characteristics for a fixed input $h$ or another emergent property such as inhibition stabilization.
-Most importantly, this analysis is a proof-of-concept demonstrating the valuable ability to condition parameters of interest of a neural circuit model on some chosen emergent property.
-To this point, we have shown the utility of EPI on relatively low-level emergent properties like network syncing and differential neuron-type population responses.  
-In the remainder of the study, we focus on using EPI to understand models of more abstract cognitive function.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
+\section{Identifying neural mechanisms of flexible task switching} \label{results_SC}
 In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
@@
 \end{equation}
-where $\epsilon = 0.05$ and $\zeta = 0.5$.  The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
+where $\epsilon = 0.05$ and $\zeta = 0.5$.  
+The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
 \begin{equation}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
+\includegraphics[scale=0.5]{figs/fig4.pdf}
 \end{center}
 \caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
-A.
-Rapid task switching behavioral paradigm (see text).
-B.
-Model of superior colliculus (SC).
-Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti.
+A. Rapid task switching behavioral paradigm (see text). 
+B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g.
-$vW_{PA}$ is a vertical weight from an Anti to a Pro population.
+Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
 C.
@@
 Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
-D.
-The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
-E.
-The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
+D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
+E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
 \label{fig:SC}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Linking RNN connectivity to error} \label{results_RNN}
-So far, each model we have studied was designed from fundamental biophysical principles, genetically- or functionally-defined neuron types.  
-At a more abstract level of modeling, recurrent neural networks (RNNs) are high-dimensional dynamical models of computation that are becoming increasingly popular in neuroscience research \cite{barak2017recurrent}.
-In theoretical neuroscience, RNN dynamics usually follow the equation
-\begin{equation}
-\frac{dx}{dt} = -x + W \phi(x) + h,
-\end{equation}
-where $x$ is the network activity, $W$ is the network connectivity, $\phi(\cdot) = \tanh(\cdot)$, and $h$ is the input to the system.
-Such RNNs are trained to do a task from a systems neuroscience experiment, and then the unit activations of the trained RNN are compared to recorded neural activity.
-Fully-connected RNNs with tens of thousands of parameters are challenging to characterize \cite{sussillo2013opening}, especially making statistical inferences about their parameterization.
-Alternatively, we considered a rank-1, $N$-neuron RNN with connectivity consisting of the sum of a random and a structured component:
-\begin{equation}
-W = g\chi + \frac{1}{N}mn^\top.
-\end{equation}
-The random component $g \chi$ has strength $g$, and random component weights are Gaussian distributed $\chi_{i,j} \sim \mathcal{N}(0, \frac{1}{N})$.  
-The structured component $\frac{1}{N}m n\top$ has entries of $m$ and $n$ drawn from Gaussian distributions $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$.
-Recent theoretical work derives the low-dimensional response properties of low-rank networks from statistical parameterizations of their connectivity, such as $z = [g, M_m, M_n]$ \cite{mastrogiuseppe2018linking}.
-We used EPI to infer the parameterizations of rank-1 RNNs solving an example task, enabling discovery of properties of connectivity that result in different types of error in the computation.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.75]{figures/fig5/fig5.pdf}
-\end{center}
-\caption{\small Sources of error in an RNN solving a simple task.
-A.
-(left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$.
-(right) Error in this computation can come from over- or under-estimating the posterior mean or variance.
-B.
-EPI distribution of rank-1 RNNs executing Gaussian posterior conditioning.
-Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$  C.
-Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.
-Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.
-D.
-The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.
-Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
-\label{fig:RNN}
-\end{figure}
-
-The task we consider is Gaussian posterior conditioning: calculate the parameters of a posterior distribution induced by a prior $p(\mu_y) = \mathcal{N}(\mu_0=4, \sigma_0^2=1)$ and a likelihood $p(y | \mu_y) = \mathcal{N}(\mu_y,\sigma^2_y = 1)$, given a single observation $y$.
-Conjugacy offers the result analytically; $p(\mu_y | y ) = \mathcal{N}\left(\mu_{post},\sigma_{post}^2\right)$, where:   
-\begin{equation}
-\mu_{\text{post}} = \frac{\frac{\mu_0}{\sigma_0^2} + \frac{y}{\sigma_y^2}}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}
-~~~~~~~~~~
- \sigma^2_{\text{post}} = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}.
-\end{equation} 
-To solve this Gaussian posterior conditioning task, the RNN response to a constant input $h = y r + (n-M_n)$ must equal the posterior mean along readout vector $r$ , where
-\begin{equation}
-\kappa_r =  \frac{1}{N} \sum_{j=1}^N r_j \phi(x_j).
-\end{equation}
-Additionally, the amount of chaotic variance $\Delta_T$ must equal the posterior variance.
-Theory for low-rank RNNs allows us to express $\kappa_r$ and $\Delta_T$ in terms of each other through a solvable system of nonlinear equations  (see Section \ref{methods_LRRNN}) \cite{mastrogiuseppe2018linking}.
-This theory facilitates the mathematical formalization of task execution into an emergent property, where the emergent property statistics of the RNN activity are $\kappa_r$ and $\Delta_T$, and the emergent property values are the ground truth posterior mean $\mu_{\text{post}}$ and variance $\sigma^2_{\text{post}}$:
-\begin{equation}
-\mathbb{E} \begin{bmatrix} \kappa_r \\ \Delta_T \\ (\kappa_r-\mu_{\text{post}})^2 \\ (\Delta_T^2-\sigma^2_{\text{post}})^2 \end{bmatrix} ~~=~~ \begin{bmatrix} \mu_{\text{post}} \\ \sigma^2_{\text{post}} \\ 0.1 \\ 0.1 \end{bmatrix}.
-\end{equation}
-We chose a substantial amount of variance in these emergent property statistics, so that the inferred distribution resulted in RNNs with a variety of errors in their solutions to the gaussian posterior conditioning problem.
-
-EPI was used to learn distributions of RNN connectivity properties $z = [g, M_m, M_n]$ executing Gaussian posterior conditioning given an input of $y=2$, where the true posterior is $\mu_{\text{post}}=3$ and $\sigma_{\text{post}} = 0.5$ (Fig.
-\ref{fig:RNN}A).
-We examined the nature of the over- and under-estimation of the posterior means (Fig. \ref{fig:RNN}B left) and variances (Fig. \ref{fig:RNN}B right) in the inferred distributions (300 samples).  
-The symmetry in the $M_m$-$M_n$ plane, suggests a degeneracy in the product of $M_m$ and $M_n$ (Fig. \ref{fig:RNN}B).  
-Indeed, $M_m M_n$ strongly determines the posterior mean ($r=0.62$,  $p<10^{-4}$). 
-Furthermore, the random strength $g$ strongly determines the chaotic variance ($r=0.56$,  $p<10^{-4}$).
-Neither of these observations were obvious from what mathematical analysis is available in networks of this type (see Section \ref{methods_LRRNN}).  
-While the link between random strength $g$ and chaotic variance $\Delta_T$ (and resultingly posterior variance in this problem) is well-known \cite{sompolinsky1988chaos}, the distribution admits a novel hypothesis: the estimation of the posterior mean by the RNN increases with $M_m M_n$.
-
-We tested this prediction by taking parameters $z_1$ and $z_2$ as representative samples from the positive and negative $M_m$-$M_n$ quadrants, respectively.
-Instead of using the theoretical predictions shown in Figure \ref{fig:RNN}B, we simulated finite-size realizations of these networks with 2,000 neurons (e.g. Fig. \ref{fig:RNN}C).  
-We perturbed these parameter choices by $M_m M_n$ clarifying that the posterior mean can be directly controlled in this way (Fig. \ref{fig:RNN}D; $p<10^{-4}$), see Section \ref{methods_LRRNN}).
-Thus, EPI confers a clear picture of error in this computation: the product of the low rank vector means $M_m$ and $M_n$ modulates the estimated posterior mean while the random strength $g$ modulates the estimated posterior variance.
-This novel procedure of inference on reduced parameterizations of RNNs conditioned on the emergent property of task execution is generalizable to other settings modeled in \cite{mastrogiuseppe2018linking} like noisy integration and context-dependent decision making (Fig.
-S5).
-
-\section{Discussion}
-\subsection{EPI is a general tool for theoretical neuroscience} 
-Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
-We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
-We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).
-We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
-Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.
-
-In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
-These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
-In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
-Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
-Although, theoretical approaches (e.g.
-\cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
-
-\subsection{Novel hypotheses from EPI} 
-In neuroscience, machine learning has primarily been used to reveal structure in large-scale neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
- Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
- However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
-In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
-The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we replaced this opaque procedure of parameter identification in theoretical neuroscience with emergent property inference, opening the door to careful inference in careful models of neural computation.
-
-Biologically realistic models of neural circuits often prove formidable to analyze.
-Two main factors contribute to the difficulty of this endeavor.
-First, in most neural circuit models, the number of parameters scales quadratically with the number of neurons, limiting analysis of its parameter space. 
-Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
-For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-
-In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}.
-By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out.
-Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
-For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
-This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.
-
-In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
-Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
-This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
-While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
-The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.
-
-EPI leverages deep learning technology for neuroscientific inquiry in a categorically different way than approaches focused on training neural networks to execute behavioral tasks \cite{richards2019deep}.
-These works focus on examining optimized deep neural networks while considering the objective function, learning rule, and architecture used.
-This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
-These approaches can be used complementarily to enhance the practice of theoretical neuroscience.
-
-\textbf{Acknowledgements}: \\
-This work was funded by NSF Graduate Research Fellowship,  DGE-1644869, McKnight Endowment Fund, NIH NINDS 5R01NS100066, Simons Foundation 542963, NSF NeuroNex Award, DBI-1707398, The Gatsby Charitable Foundation, Simons Collaboration on the Global Brain Postdoctoral Fellowship, Chinese Postdoctoral Science Foundation, and International Exchange Program Fellowship.
-Helpful conversations were had with Francesca Mastrogiuseppe, Srdjan Ostojic, James Fitzgerald, Stephen Baccus, Dhruva Raman, Liam Paninski, and Larry Abbott.
-
-\textbf{Data availability statement}: \\
-The datasets generated during and/or analysed during the current study are available from the corresponding author upon reasonable request.
-
-\textbf{Code availability statement}: \\
-The software written for the current study is available from the corresponding author upon reasonable request.
-
-\bibliography{eLife2020}
-\bibliographystyle{unsrt}
-
-\newpage
-
-\section{Methods}
-
-\subsection{Emergent property inference (EPI)}\label{methods_EPI}
-Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
-Consider model parameterization $z$ and data $x$ which has an intractable likelihood $p(x \mid z)$ defined by a model simulator of which samples are available $x \sim p(x \mid z)$.  
-EPI optimizes a distribution $q_\theta(z)$ (itself parameterized by $\theta$) of model parameters $z$ to produce an emergent property of interest $\mathcal{B}$,
-\begin{equation}
-\mathcal{B} ~~\triangleq~~  \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
-\end{equation}
-Precisely, the emergent property statistics $T(x)$ must equal the emergent property values $\mu$, in expectation over the EPI distribution of parameters $q_\theta(z)$ and the distribution of simulated activity $p(x \mid z)$.
-This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
-
-With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_\theta(z)$.
- In deep probability distributions, a simple random variable $w \sim q_0(w)$ is mapped deterministically via a sequence of deep neural network layers ($f_1$, ..
-$f_l$) parameterized by weights and biases $\theta$ to the support of the distribution of interest:
-\begin{equation}
-z = f_{\theta}(\omega) = f_l(..f_1(w)).
-\end{equation}
-Given a simulator defined by a theoretical model $x \sim p(x \mid z)$ and some emergent property of interest $\mathcal{B}$, $q_\theta(z)$ is optimized via the neural network parameters $\theta$ to find a maximally entropic distribution $q_{\theta}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
-\begin{equation} \label{eq:opt}
-\begin{split}
-q_\theta^*(z) &= \argmax_{q_\theta \in Q} H(q_\theta(z)) \\
- &  \text{s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu. \\
- \end{split}
-\end{equation} 
-Since we are optimizing parameters $\theta$ of our deep probability distribution with respect to the entropy $H(q_\theta(z))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-Entropy of $q_\theta(z)$ can be expressed as an expectation of the negative log density of parameter samples $z$ over the randomness in the parameterless initial distribution $q_0$:
-\begin{equation}
-H(q_\theta(z)) = \int - q_\theta(z) \log(q_\theta(z)) dz = \mathbb{E}_{z \sim q_\theta}\left[-\log(q_\theta(z)) \right] = \mathbb{E}_{w \sim q_0}\left[-\log(q_\theta(f_\theta(w))) \right].
-\end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $z$:
-\begin{equation}
-\nabla_\theta H(q_\theta(z)) = \mathbb{E}_{w \sim q_0}\left[- \nabla_\theta \log(q_\theta(f_\theta(w))) \right].
-\end{equation}
-In EPI, MEFNs are purposed towards variational learning of model parameter distributions.  
-A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
-Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(x \mid z)$.
-Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
-Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
-But, unlike LFVI, EPI uses a single deep network to learn a distribution and is optimized to produce an emergent property, rather than condition on data points.
-Optimizing the EPI objective is a technological challenge, the details of which we elaborate in Section \ref{methods_AL_opt}.
- Before going through those details, we ground this optimization in a toy example.
-
- We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
-We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
-These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.
-
-\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
-\begin{equation} 
-\tau \frac{dx}{dt} = Ax
-\end{equation}
-with
-\begin{equation}
-A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
-\end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $z = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(x)$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
- $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
-\begin{equation}
- \mathbb{E}\left[T(x) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \mu.
- \end{equation} 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.58]{figures/figS1/figS1.pdf}
-\end{center}
-\begin{flushleft}
-{\small Fig.
-S1: A.
-Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.
-B.
-The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.
-Dashed lines indicate the parameter axes.
-C.
-Entropy throughout the optimization.
-At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
-D.
-Emergent property moments throughout optimization.
-At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
-\end{flushleft}
-\end{figure}
-
-Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $z$, since the eigenvalues can be calculated using the quadratic formula:
-\begin{equation}
-\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
-\end{equation}
-
-Importantly, even though $\mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_\theta$ directly.
-This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\mu$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we used EPI to approximate this distribution (Fig.
-S1B).
-We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
-(see Section \ref{methods_NF}).
-
-Even this relatively simple system has nontrivial (though intuitively sensible) structure in the parameter distribution.
-To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
-In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
-the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig.
-S2A).
-The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
- Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figS2/figS2.pdf}
-\end{center}
-\begin{flushleft}
-{\small Fig.
-S2: A.
-Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$.
-B.
-Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\end{flushleft}
-\end{figure}
-
-To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
-\begin{equation}
-\text{imag}(\lambda_1) = \begin{cases}
-                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
-                             0 & \text{otherwise } \\
-                         \end{cases}.
-\end{equation}
-
-When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
-\begin{equation}
-\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
-\end{equation}
-Since we constrained $\mathbb{E}_{z \sim q_\theta}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
-S2B).
-With $\sigma_{1,4} = \mathbb{E}_{z \sim q_\theta}(|a_1 a_4 - E_{q_\theta}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
-This validates the curved structure of the inferred distribution learned through EPI.  
-We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
-Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figS3/figS3.pdf}
-\end{center}
-\begin{flushleft}
-{\small Fig.
-S3: Sampled dynamical systems $z \sim q_\theta(z)$ and their simulated activity from $x(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
-A.
-Each dimension of the simulated trajectories throughout time.
-B  The simulated trajectories in phase space.}
-\end{flushleft}
-\end{figure}
-
-\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-To optimize $q_\theta(z)$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
-\begin{equation} \label{eq:AL}
-L(\theta; \eta, c) = -H(q_\theta) + \eta^\top R(\theta) + \frac{c}{2}||R(\theta)||^2
-\end{equation}
-where $R(\theta) = \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x) - \mu \right] \right]$, $\eta \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\mu| = |T(x)|$,  and $c$ is the penalty coefficient.
-These Lagrange multipliers are closely related to the natural parameters of exponential families (see Section \ref{methods_VI}).
-Deep neural network weights and biases $\theta$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
-$\eta$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\eta$, $c$) for a given number of stochastic optimization iterations.
-A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
-For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
-The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
-The p-value of $\mathbb{E}[||R(\theta_{k+1})||] > \gamma \mathbb{E} \left[||R(\theta_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-The other update rule is $\eta_{k+1} = \eta_k + c_k \frac{1}{n} \sum_{i=1}^n (T(x^{(i)}) - \mu)$ given a batch size $n$.
-Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
-
-The intention is that $c$ and $\eta$ start at values encouraging entropic growth early in optimization.  
-With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
-In the oscillating 2D LDS example, each augmented Lagrangian epoch ran for 2,000 iterations (Fig. S1C-D).
-Notice the initial entropic growth, and subsequent reduction upon each update of $\eta$ and $c$.
-The momentum parameters of the Adam optimizer were reset at the end of each augmented Lagrangian epoch.
-
-Rather than starting optimization from some $\theta$ drawn from a randomized distribution, we found that initializing $q_\theta(z)$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
-The parameters of the Gaussain initialization were chosen on an application-specific basis.
-Throughout the study, we chose isotropic Gaussian initializations with mean $\mu_{\text{init}}$ at the center of the distribution support and some standard deviation $\sigma_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_V1}).
-
-To assess whether EPI distribution $q_\theta(z)$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
-The algorithm has converged when a null hypothesis test of constraint violations $R(\theta)_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
-This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
-The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\theta)^i$ are made from $k$ resamplings of $z$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
-$k$ is determined by a fraction of the batch size $\nu$, which varies according to the application.
-In the linear two-dimensional system example, we used a batch size of $n = 1000$ and set $\nu = 0.1$ resulting in convergence after the ninth epoch of optimization. (Fig. S1C-D black dotted line).
-
-When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
-First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
-Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
-If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
-Next, one should consider the computational cost of the gradient calculations. 
-In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_2DLDS}) for the emergent property statistic given the model parameters.  
-On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
-Section \ref{methods_STG}).
-In such cases, optimization will be expensive.
-
-\subsubsection{Normalizing flows}\label{methods_NF}
-Deep probability models typically consist of several layers of fully connected neural networks.
-When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
-For $z' = f(z)$,
-
-\begin{equation}
-q(z') = q(f^{-1}(z')) \left| \det \frac{\partial f^{-1}(z')}{\partial z'} \right| = q(z) \left| \det \frac{\partial f(z)}{\partial z} \right|^{-1}.
-\end{equation}
-
-However, this computation has cubic complexity in dimensionality for fully connected layers.
-By restricting our layers to normalizing flows \cite{rezende2015variational} -- bijective functions with fast log determinant Jacobian computations, we can tractably optimize deep generative models with objectives that are a function of sample density, like entropy.
-Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.
-Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
-
-\subsubsection{Emergent property inference as variational inference in an exponential family}\label{methods_VI}
-Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In Bayesian inference a prior belief about model parameters $z$ is formalized into a prior distribution $p(z)$, and the statistical model capturing the effect of $z$ on observed data points $x$ is formalized in the likelihood distribution $p(x \mid z)$.
-In Bayesian inference, we obtain a posterior distribution $p(z \mid x)$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
-\begin{equation}
-p(z \mid x) = \frac{p(x \mid z) p(z)}{p(x)}.
-\end{equation}
-The posterior disribution is analytically available when the prior is conjugate with the likehood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
-
-As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
-A maximum entropy distribution of form:
-\begin{equation}
-\begin{split}
-p^*(z) &= \argmax_{p \in \mathcal{P}} H(p(z)) \\
- &  \text{s.t.  } \mathbb{E}_{z \sim p}\left[T(z)\right] = \mu. \\
- \end{split}
-\end{equation} 
-will have probability density in the exponential family:
-\begin{equation}
-p^*(z) \propto \exp(\eta^\top T(z)).
-\end{equation}
-The mappings between the mean parameterization $\mu$ and the natural parameterization $\eta$ are formally hard to identify \cite{wainwright2008graphical}.
-
-Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid x)$.
-We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(x)$.
-The log normalizing constant $A(\eta(x))$ remains unchanged:
-\begin{equation}
-\begin{split}
-p(z \mid x) = b(z) \exp{\left( \eta(x)^\top T(z) - A(\eta(x)) \right)} = \exp{\left( \begin{bmatrix} \eta(x) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(x)) \right)} \\= \exp{\left(\tilde{\eta(x)}^\top \tilde{T}(z) - A(\eta(x)) \right)}
-\end{split}.
-\end{equation}
-Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
-\begin{equation}
-q_\theta^* = \argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)).
-\end{equation}
-$q_\theta(z)$ is the variational approximation to the posterior with variational parameters $\theta$.  We can write this KL divergence in terms of entropy of the variational approximation:
-\begin{equation}
-KL(q_\theta \mid \mid p(z \mid x)) = \mathbb{E}_{z \sim q_\theta} \left[ \log (q_\theta(z)) \right] - \mathbb{E}_{z \sim q_\theta} \left[ \log (p(z \mid x)) \right]
-\end{equation}
-\begin{equation}
- = -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) - A(\eta(x)) \right].
-\end{equation}
-As far as the variational optimization is concerned, the log normalizing constant is independent of $q_\theta(z)$, so it can be dropped
-\begin{equation}
-\argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)) =  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) \right].
-\end{equation}
- Further, we can write the objective in terms of the first moment of the sufficient statistics $\mu = \mathbb{E}_{z \sim p(z \mid x)}\left[T(z) \right]$:
- \begin{equation}
-=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right] + \tilde{\eta}(x)^\top \mu,
- \end{equation}
-which simplifies to
-  \begin{equation}
-=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right].
- \end{equation}.
-
-In comparison, in emergent property inference (EPI), we solve the following problem:
-\begin{equation}
-q_\theta^*(z) = \argmax_{q_\theta \in Q} H(q_\theta(z)),   \text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
-\end{equation}
-The Lagrangian objective (without augmentation) is
-\begin{equation}
-q_\theta^* = \argmin_{q_\theta \in Q} - H(q_\theta) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_\theta} \left[\tilde{T}(z) \right] - \mu \right).
-\end{equation}
-Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(x)$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
-
-We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
-Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
-However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-Accordingly, we replace the notation of $p(z \mid x)$ with $p(z \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
-
-\subsection{Theoretical models}\label{methods_theoretical_models}
-In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  Here, we provide the details of each model  and the related analyses.
-
-\subsubsection{Stomatogastric ganglion}\label{methods_STG}
-We analyze how the parameters $z = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
-Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
-\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right].
-\end{equation} 
-The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization,
-electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $z$.  
-The capacitance of the cell membrane was set to $C_m = 1nF$.
-Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
-\begin{equation}  h_{leak}(x; z) = g_{leak} (x_\alpha - V_{leak}) 
-\end{equation} 
-\begin{equation}  h_{elec}(x; z) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
-\end{equation} 
-\begin{equation}  h_{syn}(x; z) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
-\begin{equation}  h_{Ca}(x; z) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
-\end{equation} 
-\begin{equation}  h_K(x; z) = g_K N (x_\alpha - V_K) 
-\end{equation} 
-\begin{equation}  h_{hyp}(x; z) = g_h H(x_\alpha - V_{hyp}).
-\end{equation} 
-The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.
-The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$.
-$g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.
-The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.
-The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.
-Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.
-
-Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
-\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
-\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
-\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
-\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
-\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
-\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
-\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
- \end{equation}
-where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
-$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.
-
-Finally, there is a synaptic gating variable as well:
-\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
-\end{equation}
-When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-
-In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.
-In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible.
-The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.
-Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.
-Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$
-
-Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.
-High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.
-Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.
-The signal was subsequently mean centered.
-After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.
-
-Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.
-Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-The ``frequency identification" vector is
-\begin{equation}
-v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
-\end{equation}
-The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.
-
-Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.
-The emergent property statistics are the first and second moments of the firing frequencies.
-The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
-\begin{equation}
-E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
-\end{equation}
-for the EPI distribution shown in Fig. \ref{fig:STG}B.
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).
-The optimization converged after four epochs (Fig.
-S4).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.45]{figures/figS4/figS4.pdf}
-\end{center}
-\begin{flushleft}
-{\small Fig.
-S4: EPI optimization of the STG model producing network syncing.
-A.
-Entropy throughout optimization.
-B.
-The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.
-Since $q_\theta(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
-C.
-The second moment emergent property statistics converge to the emergent property values.}
-\end{flushleft}
-\end{figure}
-
-For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $z = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
-
-We calculated the Hessian at the mode of the inferred EPI distribution.  
-The Hessian of a probability model is the second order gradient of the log probability density $\log q_\theta(z)$ with respect to the parameters $z$:  $\frac{\partial^2 \log q_\theta(z)}{\partial z \partial z^\top}$.
-With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
-The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
-We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
-Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
-The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
-In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
-\ref{fig:STG}B, inset).
-The contours of Figure \ref{fig:STG} were calculated as error in $T(x)$ from $\mu$ in both the first and second moment emergent property statistics.
-
-\subsubsection{Primary visual cortex}\label{methods_V1}
-The dynamics of each neural populations average rate
-$x = [ x_E, x_P, x_S, x_V]^\top$
-are given by:
-\begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
-\end{equation}
-
-By consolidating information from many experimental datasets, Billeh et al. \cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
-\begin{equation}
-M = \begin{bmatrix} 0.36 & 0.48 & 0.31 & 0.28 \\
-1.49 & 0.68 & 0.50 & 0.18 \\
-0.86 & 0.42 & 0.15 & 0.32 \\
-1.31 & 0.41 & 0.52 & 0.37 \end{bmatrix}
-\end{equation}
-and connection probability
-\begin{equation}
-C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
-0.395 & .451 & 0.857 & 0.02 \\
-0.182 & 0.03 & 0.082 & 0.625 \\
-0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
-\end{equation}
-
-Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
-\begin{equation}
-W_{\text{full}} = C \odot M = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
-0.395 & .451 & 0.857 & 0.02 \\
-0.182 & 0.03 & 0.082 & 0.625 \\
-0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
-\end{equation}
-
-Theoretical work on these systems considers a subset of the effective connectivities \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}
-\begin{equation} \label{eq:V1_W}
-W = \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & 0 \\
-                                W_{PE} & W_{PP} & W_{PS} & 0 \\
-                                W_{SE} & 0 & 0 & W_{SV} \\
-                                W_{VE} & W_{VP} &  W_{VS} &  0 \end{bmatrix}.
-\end{equation}
-In coherence with this work, we only keep the entries of $W_{\text{full}}$ corresponding to parameters in Equation \ref{eq:V1_W}.
-
-We look at how this four-dimensional nonlinear dynamical model of V1 responds to different inputs, and compare the predictions of the linear response to the approximate posteriors obtained through EPI.
-The input to the system is the sum of a baseline input $b = [1, 1, 1, 1]^\top$ and a differential input $dh$:
-\begin{equation}
-h = b + dh.
-\end{equation}
-All simulations of this system had $T=100$ time points, a time step $dt = 5$ms, and time constant $\tau = 20$ms.  The system was initialized to a random draw $x(0)_i \sim \mathcal{N}(1, 0.01)$.
-
-We can describe the dynamics of this system more generally by
-\begin{equation}
-\dot{x}_i = -x_i + f(u_i)
-\end{equation}
-where the input to each neuron is
-\begin{equation}
-u_i = \sum_j W_{ij} x_j + h_i.
-\end{equation}
-Let $F_{ij} = \gamma_i \delta(i,j)$, where $\gamma_i = f'(u_i)$.  Then, the linear response is
-\begin{equation}
-\frac{dx_{ss}}{dh} = F(W\frac{dx_{ss}}{dh} + I)
-\end{equation}
-which is calculable by
-\begin{equation}
-\frac{dx_{ss}}{dh} = (F^{-1} - W)^{-1}.
-\end{equation}
-This calculation is used to produce the magenta lines in Figure \ref{fig:V1_EPI}C, which show the linearly predicted inputs that generate a response from two standard deviations (of $\mathcal{B}$) below and above $y$.
-
-The emergent property we considered was the first and second moments of the change in steady state rate $dx_{ss}$ between the baseline input $h= b$ and $h = b + dh$.
-We use the following notation to indicate that the emergent property statistics were set to the following values:
-\begin{equation}
-\mathcal{B}(\alpha, y) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha,ss} \\ (dx_{\alpha,ss} - y)^2 \end{bmatrix} = \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
-\end{equation}
-
-In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)$.
-The differential responses $\delta x_{\alpha,ss}$ are examined at perturbed inputs  $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$ and $\delta x$ is evaluated at 101 equally spaced samples of $\delta h_\alpha$ from -15 to 15.
-
-We measured the linear regression slope between neuron-types of $\delta x$ and $\delta h$ to confirm the hypotheses H1-H3 (H4 is simply observing the nonmonotonicity) and report the p values for tests of non-zero slope.
-
-{\addtolength{\leftskip}{10 mm}
-H1: the neuron-type ressponses are sensitive to their direct inputs. 
-E-population: $\beta=1.62$, $p<10^{-4}$ (Fig. \ref{fig:V1_HT}A black),
-P-population: $\beta=1.06$, $p<10^{-4}$ (Fig.
-\ref{fig:V1_HT}B blue), S-population: $\beta=6.80$, $p<10^{-4}$  (Fig.
-\ref{fig:V1_HT}C red), V-population: $\beta=6.41$, $p<10^{-4}$  (Fig.
-\ref{fig:V1_HT}D green).
-\\
-H2: the E-population ($\beta=0$, $p=1$) and P-populations ($\beta=0$, $p=1$) are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
-H3: the S-population is not affected by $\delta h_P$ ($\beta=0$, $p=1$) (Fig. \ref{fig:V1_HT}C blue); \\
-
-}
-
-For each $\mathcal{B}(\alpha, y)$ with $\alpha \in \{E, P, S, V\}$ and $y \in \{0.1, 0.5\}$, we ran EPI using a real NVP architecture of four masks layers with two hidden layers of 10 units,  mapped to a support of $z_i \in \left[-5, 5\right]$ with no batch normalization.
-We used an augmented Lagrangian coefficient of $c_0 = 10^5$, a batch size $n=1000$, set $\nu = 0.5$.
-The EPI distributions shown in Fig. \ref{fig:V1_EPI} are the converged distributions with maximum entropy across random seeds.
-
-We set the parameters of the Gaussian initialization $\mu_\text{init}$ and $\Sigma_\text{init}$ to the mean and covariance of random samples $z^{(i)} \sim \mathcal{U}(-5, 5)$ that produced emergent property statistic $dx_{\alpha,ss}$ within a bound $\epsilon$ of the emergent property value $y$.
-$\epsilon=0.01$ was set to be one standard deviation of the emergent property value according to the emergent property value $0.01^2$ of the variance emergent property statistic.
-
-\subsubsection{Superior colliculus}\label{methods_SC}
+\section{Supplementary}\label{methods_SC}
 In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
@@
 
-\subsubsection{Rank-1 RNN}\label{methods_LRRNN}
-Extensive research on random fully-connected recurrent neural networks has resulted in foundational theories of their activity \cite{sompolinsky1988chaos, brunel2000dynamics}.
-Furthermore, independent research on training these models to perform computations suggests that learning occurs through low-rank perturbations to the connectivity (e.g.
-\cite{jaeger2004harnessing, sussillo2009generating}).
-Recent theoretical work extends theory for random neural networks \cite{sompolinsky1988chaos} to those with added low-rank structure \cite{mastrogiuseppe2018linking}.
-In Section \ref{results_RNN}, we used this theory to enable EPI on RNN parameters conditioned on the emergent property of task execution.
-
-Such RNNs have the following dynamics: 
-\begin{equation}
-\frac{dx}{dt} = -x + W \phi(x) + h,
-\end{equation}
-where $x$ is network activity, $W$ is the connectivity weight matrix, $\phi(\cdot) = \tanh(\cdot)$ is the input-output function, and $h$ is the input to the system.
-In a rank-1 RNN (which was sufficiently complex for the Gaussian posterior conditioning task), $W$ is the sum of a random component with strength $g$ and a structured component determined by the outer product of vectors $m$ and $n$:
-\begin{equation}
-W = g\chi + \frac{1}{N}mn^\top,
-\end{equation}
-where  $\chi_{ij} \sim \mathcal{N}(0, \frac{1}{N})$, and the entries of $m$ and $n$ are distributed as $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$. 
-For EPI, we consider $z = [g, M_m, M_n]$, which are the parameters governing the connectivity properties of the RNN.
-
-From such a parameterization $z$, the theory of Mastrogiuseppe et al. produces solutions for variables describing the low dimensional response properties of the RNN.
-These ``dynamic mean field" (DMF) variables (e.g.
-the activity along a vector $\kappa_v$, the total variance $\Delta_0$, structured variance $\Delta_\infty$, and the chaotic variance $\Delta_T$) are derived to be functions of one another and connectivity parameters $z$.
-The collection of these derived functions results in a system of equations, whose solution must be obtained through a nonlinear system of equations solver.
-The iterative steps of this system of equations solver are differentiable, so we take gradients through this solve process.
-The DMF variables provide task-relevant information about the RNN's response to task inputs.
-
-In the Gaussian posterior conditioning example, $\kappa_r$ and $\Delta_T$ are DMF variables used as task-relevant emergent property statistics $\mu_{\text{post}}$ and $\sigma^2_{\text{post}}$.
-Specifically, we solve for the DMF variables $\kappa_r$, $\kappa_n$, $\Delta_0$ and $\Delta_\infty$, where the readout is nominally chosen to point in the unit orthant $r = [1, ..., 1]^\top$.
-The consistency equations for these variables in the presence of a constant input $h = y r - (n - M_n)$ can be derived following \cite{mastrogiuseppe2018linking}:
-{\small
-\begin{equation}
-\begin{split}
-\kappa_r = G_1(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_m \kappa_n + y \\
-\kappa_n = G_2(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_n \langle \left[ \phi_i \right] \rangle + \langle \left[ \phi_i' \right] \rangle \\
-\frac{\Delta_0^2-\Delta_\infty^2}{2} = G_3(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = g^2 \left( \int \mathcal{D}z \Phi^2(\kappa_r + \sqrt{\Delta_0}z) - \int \mathcal{D}z \int \mathcal{D}x \Phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z)  \right) \\
-+ (\kappa_n^2  + 1)(\Delta_0-\Delta_\infty) \\
-\Delta_\infty = G_4(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty)  = g^2 \int \mathcal{D}z \left[ \int \mathcal{D}x \phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z \right]^2 + \kappa_n^2 + 1
-\end{split} 
-\end{equation}
-}
-where here $z$ is a gaussian integration variable. We can solve these equations by simulating the following Langevin dynamical system to a steady state:
-
-\begin{equation}
-\begin{split}
-l(t) = \frac{\Delta_0(t)^2-\Delta_\infty(t)^2}{2} \\
-\Delta_0(t) = \sqrt{2l(t) + \Delta_\infty(t)^2} \\
-\frac{d \kappa_r(t)}{dt}  = -\kappa_r(t) + G_1(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d \kappa_n(t)}{dt} = -\kappa_n + G_2(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d l(t)}{dt}  = -l(t) + G_3(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d \Delta_\infty(t)}{dt}  = -\Delta_\infty(t) + G_4(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t))
-\end{split}.
-\end{equation}
-Then, the chaotic variance, which is necessary for the Gaussian posterior conditioning example, is simply calculated via $\Delta_T = \Delta_0 - \Delta_\infty$.
-
-We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization.
-We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
- The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.
-
-To examine the effect of product $M_m M_n$ on the posterior mean, $\mu_{\text{post}}$ we took perturbations in $M_m M_n$ away from two representative parameters  $z_1$ and $z_2$ in 21 equally space increments from -1 to 1.
-For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.
-In Fig.
-\ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.
-The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.
-For perturbations away from $z_1$ the correlation was 0.995 with $p<10^{-4}$, and for perturbations away from $z_2$ the correlation was 0.983 with $p<10^{-4}$.
-
-In addition to the Gaussian posterior conditioning example in Section \ref{results_RNN}, we modeled two tasks from Mastrogiuseppe et al.: noisy detection and context-dependent discrimination.
-We used the same theoretical equations and task setups described in their study.
-
-%TODO Need to explain the warm starting for the aficionados.
-
-%TODO explain the density network architectures used.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
-\end{center}
-{\small Fig.
-S5: A.
-EPI for rank-1 networks doing noisy discrimination.
-B.
-EPI for rank-2 networks doing context-dependent discrimination.
-See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
-\end{figure}
+\bibliography{epi}
+\bibliographystyle{unsrt}
 

### 2020-11-06 written/SC_model/SC_model.tex
sc update tetx

@@
 \end{equation}
-with gaussian noise of variance $\sigma^2$ = 1.
+with gaussian noise of variance $\sigma^2$ = 0.2.
 The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.
+Here, we use EPI to determine the changes in network connectivity $z = [sW, vW, dW, hW]^{\top}$ resulting in greater levels of rapid task switching accuracy.
 
@@
 
-To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
-At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
-Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
-Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
+To quantify the emergent property of rapid task switching , we considered the requirements of this model in this behavioral paradigm.
+At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must be greater than the Pro population in the opposite hemisphere.
+Thus, we can formulate rapid task switching accuracy as the average number of successful trials over gaussian noise.
 \begin{equation}

### 2020-11-07 written/SC_model/SC_model.tex
SC opto update

@@
 \DeclareMathOperator*{\argmax}{argmax}
+\usepackage{bm}
 
@@
 \ref{fig:SC}B).
-The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
+The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
+The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
 \begin{equation}
-x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
+\begin{split}
+\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + \epsilon d\mathbf{B} \\
+\mathbf{x} &= f(\mathbf{u})
+\end{split}
 \end{equation}
-where $\epsilon = 0.05$ and $\zeta = 0.5$.  
-The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
-\begin{equation}
-\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
-\end{equation}
-with gaussian noise of variance $\sigma^2$ = 0.2.
-The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the changes in network connectivity $z = [sW, vW, dW, hW]^{\top}$ resulting in greater levels of rapid task switching accuracy.
+with white noise of variance $\epsilon^2 = 0.2^2$.
+The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
+Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.
 
@@
 
-To quantify the emergent property of rapid task switching , we considered the requirements of this model in this behavioral paradigm.
-At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must be greater than the Pro population in the opposite hemisphere.
-Thus, we can formulate rapid task switching accuracy as the average number of successful trials over gaussian noise.
-\begin{equation}
-\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
+We define rapid task switching behavior as accurate execution of each task.
+Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
+We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.
+\begin{equation}\label{eq:EP}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 5\%^2 \\ 5\%^2  \end{bmatrix}
+\end{split}
 \end{equation}
-Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.
+A variance of $5\%$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.
 
-We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
-To make sense of these inferred distributions, we followed the approach of Duan et al.
+We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
+There is strong pairwise correlation between each parameter of the connectivity matrix.
+Shading the distribution by $p_P$, we see that positive shifts in the parameters result in lower accuracy in the Pro task.
+Indeed, this difference in accuracy is apparent when we simulate this model for greater parameter values (black star) and lower parameter values (gray star) (Fig. \ref{fig:SC}D).
+
+To make sense of this inferred distributions, we followed the approach of Duan et al.
 by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-\ref{fig:SC}C).
+SX).
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
 The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
-
-We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig.
-\ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
-Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
-Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
-There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
-Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
-These hypotheses are confirmed by forward simulation of the SC model (Fig.
-\ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.
+We found that $\lambda_{\text{all}}$ is anti-correlated with $p_P$ and $\lambda_{\text{task}}$ is correlated with $p_P$.
+TODO interpret further.  Maybe push further on the optogenetics analysis.
 
 \section{Supplementary}\label{methods_SC}
-In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
-They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
-Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
 \begin{equation}
-x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
+\mathbf{x}_\alpha = f(\mathbf{u}) = \left(\frac{1}{2}\tanh\left(\frac{\mathbf{u}_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
 \end{equation}
-where $\alpha \in \{LP, LA, RA, RP\}$ $\epsilon = 0.05$ and $\zeta = 0.5$ control the position and shape of the nonlinearity, respectively.
+where $\epsilon = 0.05$ and $\zeta = 0.5$.
 
-We order the elements of $x$ and $u$ in the following manner
+The accuracies of $p_P$ and $p_A$ are calculated as
 \begin{equation}
-x = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} u = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix}.
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
 \end{equation}
+and 
+\begin{equation}
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+\end{equation}
+given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function.
 
- The internal variables follow dynamics:
+The Heaviside step function is approximated as
 \begin{equation}
-\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
+\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
 \end{equation}
-with time constant $\tau = 0.09s$ and Gaussian noise $\sigma dB$ controlled by the magnitude of $\sigma=1.0$.
-The weight matrix has 8 parameters $sW_P$, $sW_A$, $vW_{PA}$, $vW_{AP}$, $hW_P$, $hW_A$, $dW_{PA}$, and $dW_{AP}$ (Fig.
-4B):
-\begin{equation}
-W = \begin{bmatrix} sW_P & vW_{PA} & hW_P & dW_{PA}  \\ vW_{AP}  & sW_A & dW_{AP}  & hW_A \\ hW_P & dW_{PA}  & sW_P & vW_{PA}  \\ dW_{AP}  & hW_A & vW_{AP}  & sW_A \end{bmatrix}.
-\end{equation}
+where $\beta = 100$.
 
@@
 \begin{equation}
-h = h_{\text{rule}} + h_{\text{choice-period}} + h_{\text{light}}.
+\mathbf{h}(t) = \mathbf{h}_{\text{constant}} + \mathbf{h}_{\text{P,bias}} + \mathbf{h}_{\text{rule}}(t) + \mathbf{h}_{\text{choice-period}}(t) + \mathbf{h}_{\text{light}}(t).
 \end{equation}
@@
 There are rule-based inputs depending on the condition,
-\begin{equation}h_{\text{P,rule}}(t) = \begin{cases}
+\begin{equation} 
+\mathbf{h}_{\text{constant}} = I_{\text{constant}} [1, 1, 1, 1]^\top
+\end{equation}
+\begin{equation} 
+\mathbf{h}_{\text{P,bias}} = I_{\text{P,rule}} [1, 0, 1, 0]^\top
+\end{equation}
+\begin{equation} \mathbf{h}_{\text{P,rule}}(t) = \begin{cases}
                            I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
@@
 \end{equation}
-\begin{equation} h_{\text{A,rule}}(t) = \begin{cases}
+\begin{equation} \mathbf{h}_{\text{A,rule}}(t) = \begin{cases}
                            I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
@@
 a choice-period input,
-\begin{equation} h_{\text{choice}}(t) = \begin{cases}
+\begin{equation} \mathbf{h}_{\text{choice}}(t) = \begin{cases}
                            I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
@@
 and an input to the right or left-side depending on where the light stimulus is delivered.     
-\begin{equation}  h_{\text{light}}(t) = \begin{cases}
+\begin{equation}  \mathbf{h}_{\text{light}}(t) = \begin{cases}
                            I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } t > 1.2s \text{ and Left} \\
@@
 \end{equation}
-The input parameterization was fixed to $I_{\text{P,rule}} = 10 $,  $I_{\text{A,rule}} = 10$,  $I_{\text{choice}} = 2$,  and $I_{\text{light}} = 1$.
-
-To produce an accuracy rate of $p_{LP}$ in the Left, Pro condition, let $\hat{p}_i$ be the empirical average steady state response (final $x_{LP}$ at end of task) over M=500 Gaussian noise draws for a given SC model parameterization $z_i$:
-\begin{equation}
- \hat{p}_i = \mathbb{E}_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] = \frac{1}{M}\sum_{j=1}^M x_{LP}(s=L, c=P, z=z_i, \sigma dB_j)
- \end{equation}
- where stimulus $s \in \{L, R\}$, cue $c \in \{P, A\}$, and $\sigma dB_j$ is the Gaussian noise on trial $j$.
-As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
-For the first emergent property statistic, the average over EPI samples (from $q_\theta(z)$) is set to the desired value $p_{LP}$:
-\begin{equation}
-\mathbb{E}_{z_i \sim q_\phi} \left[ \mathbb{E}_{\sigma dB} \left[ x_{LP,\text{ss}} \mid s=L, c=P, z=z_i \right] \right] = \mathbb{E}_{z_i \sim q_\phi} \left[ \hat{p}_i \right] = p_{LP}.
-\end{equation}
-
-For the next emergent property statistic, we ask that the variance of the steady state responses across Gaussian draws, is the Bernoulli variance for the empirical rate $\hat{p}_i$:
-\begin{equation}
-\mathbb{E}_{z \sim q\phi} \left[ \sigma^2_{err} \right] = 0
-\end{equation}
-where the Bernoulli variance error $\sigma^2_{err}$ for the Pro task, left condition is
-\begin{equation}
-\sigma^2_{err} = Var_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] - \hat{p}_i(1 - \hat{p}_i).
-\end{equation}
-
-We have an additional constraint that the Pro neuron on the opposite hemisphere should have the opposite value (0 and 1).  We can enforce this with another constraint:
-\begin{equation}
-\mathbb{E}_{z \sim q\phi} \left[ d_P \right] = 1,
-\end{equation}
-where the distance between Pro neuron steady states $d_P$ in the Pro condition is
-\begin{equation}
-d_P =  \mathbb{E}_{\sigma dB} \left[ (x_{LP} - x_{RP})^2  \mid s=L, c=P, z=z_i \right]
-\end{equation}
-The emergent property statistics only need to be measured during the Left stimulus condition of the Pro and Anti tasks, since the network is symmetrically parameterized.
-In total, the emergent property of rapid task switching at accuracy level $p$ was defined as
-\begin{equation}
-\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
-\end{equation}
-
-Since the maximum variance of a random variable bounded from 0 to 1 is the Bernoulli variance $\hat{p}(1-\hat{p})$, and the maximum squared difference between to variables bounded from 0 to 1 is 1, we do not need to control the second moment of these test statistics.
-These variables are dynamical system states and can only exponentially decay (or saturate) to 0 (or 1), so the Bernoulli variance error and squared difference constraints cannot be satisfied exactly in simulation.
-This is important to be mindful of when evaluating the convergence criteria.  
-Instead of using our usual hypothesis testing criteria for convergence to the emergent property, we set a slack variable threshold only for these technically infeasible emergent property values to 0.05.
-
-Using EPI to learn distributions of dynamical systems producing Bernoulli responses at a given rate (with small variance around that rate) was more challenging than expected.  
-There is a pathology in this optimization setup, where the learned distribution of weights is bimodal attributing a fraction $p$ of the samples to an expansive mode (which always sends $x_{LP}$ to 1), and a fraction $1-p$ to a decaying mode (which always sends $x_{LP}$ to 0).
-This pathology was avoided using an inequality constraint prohibiting parameter samples that resulted in low variance of responses across noise.
-
-\begin{table}[h]
-\begin{tabular}{l|l|l|l|l}
-$\lambda$  & $\hat{p}$ & $q_\theta(z)$ & $r$ & p-value \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $7.56 \times 10^{-01}$ & $p<10^{-4}$ \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $4.59 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $3.76 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $4.80 \times 10^{-02}$ & $p<.01$ \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $2.08 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $4.84 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(90\%))$ & $4.25 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(50\%))$ & $-7.57 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $-6.73 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $-4.86 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $-1.43 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $-1.93 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $-7.60 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $-2.73 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $-2.74 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-\end{tabular}
-\caption{\small Table of significant correlation values from Fig. \ref{fig:SC}E.}
-\end{table}
-
-For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
-We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
-The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.
-
-We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  
-Correlations were measured from 5,000 samples of $q_\theta(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
+The input parameterization was fixed to $I_{\text{constant}} = 0.75$, $I_{\text{P,bias}} = 0.5 $, $I_{\text{P,rule}} = 0.6$,  $I_{\text{A,rule}} = 0.6$,  $I_{\text{choice}} = 0.25$,  and $I_{\text{light}} = 0.5$.
 

### 2020-11-07 epi/SC_Circuit_4.py

@@ -107,7 +107,7 @@ def unwrap(z):
 
 
 def SC_acc(sW, vW, dW, hW):
-    N = 200
+    N = 500
     Wrow1 = tf.stack([sW, vW, dW, hW], axis=2)
     Wrow2 = tf.stack([vW, sW, hW, dW], axis=2)
     Wrow3 = tf.stack([dW, hW, sW, vW], axis=2)
@@ -164,3 +164,71 @@ def SC_acc_var(sW, vW, dW, hW):
     T_x = tf.concat((p, p_var), axis=1)
 
     return T_x
+
+C_opto = 4
+I_opto = tf.concat((I_LP, I_LA, I_LP, I_LA), axis=2)
+opto_strength = 0.3
+eta = np.ones((T, 1, C_opto, 1, 1), dtype=np.float32)
+eta[np.logical_and(0.8 <= t, t <= 1.2), :, 2:, :, :] = opto_strength
+
+def SC_sim_opto(sW, vW, dW, hW):
+    N = 200
+    Wrow1 = tf.stack([sW, vW, dW, hW], axis=2)
+    Wrow2 = tf.stack([vW, sW, hW, dW], axis=2)
+    Wrow3 = tf.stack([dW, hW, sW, vW], axis=2)
+    Wrow4 = tf.stack([hW, dW, vW, sW], axis=2)
+
+    W = tf.stack([Wrow1, Wrow2, Wrow3, Wrow4], axis=2)
+
+    # initial conditions
+    # M,C,4,1
+    state_shape = (sW.shape[0], C_opto, 4, N)
+    v0 = 0.1 * tf.ones(state_shape, dtype=DTYPE)
+    v0 = v0 + 0.005*tf.random.normal(v0.shape, 0., 1.)
+    u0 = beta * tf.math.atanh(2 * v0 - 1) - theta
+
+    v = v0
+    u = u0
+    v_t_list = [v]
+    u_t_list = [u]
+    for i in range(1, T):
+        du = (dt / tau) * (-u + tf.matmul(W, v) + I_opto[i] + sigma * tf.random.normal(state_shape, 0., 1.))
+        u = u + du
+        v = eta[i] * (0.5 * tf.tanh((u - theta) / beta) + 0.5)
+        v_t_list.append(v)
+        u_t_list.append(u)
+
+    u_t = tf.stack(u_t_list, axis=0)
+    v_t = tf.stack(v_t_list, axis=0)
[…]

### 2020-11-09 written/SC_model/SC_model.tex
sc model stuff
sc update

@@
 There is strong pairwise correlation between each parameter of the connectivity matrix.
-Shading the distribution by $p_P$, we see that positive shifts in the parameters result in lower accuracy in the Pro task.
-Indeed, this difference in accuracy is apparent when we simulate this model for greater parameter values (black star) and lower parameter values (gray star) (Fig. \ref{fig:SC}D).
-
-To make sense of this inferred distributions, we followed the approach of Duan et al.
+To make sense of this inferred distribution, we followed the approach of Duan et al.
 by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-SX).
+\ref{fig:SC_Eigs}A).
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
 The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
-We found that $\lambda_{\text{all}}$ is anti-correlated with $p_P$ and $\lambda_{\text{task}}$ is correlated with $p_P$.
-TODO interpret further.  Maybe push further on the optogenetics analysis.
+Strong anticorrelation between $p_P$ and $\lambda_{\text{all}}$ (Fig.
+\ref{fig:SC_Eigs}B) suggests that accuracy in the Pro task is greater when all parameters decrease along the strongly correlated posterior.
+Indeed, this difference in accuracy is apparent when we simulate this model for greater parameter values (black star) and lower parameter values (gray star) (Fig. \ref{fig:SC}D). 
+In contrast, intermediate parameter values result in greater $p_A$ than extremal parameter vaues (Fig. \ref{fig:SC_EPI2}).
+While $p_P$ is well predicted by the remaining Schur eigenvalues (correlated with $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ and anticorrelated with $\lambda_{\text{side}}$), $p_A$ is not well predicted by any of them (Fig.
+\ref{fig:SC_Eigs}B).
+
+Finally, we sought to
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.75]{figs/figSX1.pdf}
+\end{center}
+\caption{\small Sentence 1 .
+A. Rapid task switching behavioral paradigm (see text). 
+B. blah}
+\label{fig:SC_Eigs}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.75]{figs/figSX2.pdf}
+\end{center}
+\caption{\small Sentence 1 .
+A. Rapid task switching behavioral paradigm (see text). 
+B. blah}
+\label{fig:SC_EPI2}
+\end{figure}
+
 \bibliography{epi}

### 2020-11-23 written/V1_model/V1_model.tex
checkpoint V1 model

@@
 \usepackage{tikz}
+\usepackage{systeme}
 \usepackage{graphicx}
@@
 
-\section{EPI clarifies the implications of contrast-response flipping on [noise quenching, inhibition stabilization] in V1}\label{results_V1}
+\section{Exploratory analysis of V1 with EPI produces a novel theory}\label{results_V1}
 Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
 Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
@@
 Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we use EPI to analyze the posteriors of a stochastic nonlinear dynamical model of V1 conditioned on the emergent property of contrast-dependent S-V flipping.
-We then [use this info to make a statement about ...]
+Here, we use EPI to gain a comprehensive understanding of how external circuit inputs govern S-V flipping and modulate variability.
 
-Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and sum of recurrent inputs $Wx$, external inputs $h$, and external noise $\epsilon \sim \mathcal{N}(0, \sigma_{\epsilon}^2)$:
-
+We consider contrast responses of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:SV_flip}A) with a state comprised of each neuron-type population's average membrane potential $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (mV).
+Specifically, we examine the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, generalized to have inhibitory multiplicity (see Section \ref{methods_V1}).
+Each population of neuron-type $\alpha$ receives recurrent input $\left(W f_r(x)\right)_\alpha$ from synaptic projections, where $W$ is the connectivity matrix and $f_r = \left[\cdot\right]_+^2$ is the rate nonlinearity.
+$W$ is calculated from recent experiments measuring post-synaptic potentials and connectivity rates in mice  \cite{allen2018layer, billeh2019systematic}.
+Additionally, each population experiences a positive baseline input of $b_\alpha$, a contrast ($c$)-dependent input $h_\alpha$, and a stochastic input $\epsilon_\alpha$.
+%Since contrast is only considered to change input to the E- and P-populations in V1 (\textbf{TODO} AP to add citation), $h_E = h_P = c h^{(c)}$ and $h_S = h_V = 0\text{mV}$.
+Little is known about the scales of these baseline and contrast-dependent inputs, so we treat these as free parameters in our application of EPI:
  \begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h + \epsilon]_+^n.
+ z =  \begin{bmatrix} b_E & b_P & b_S & b_V & h^{(c)} \end{bmatrix}^\top.
 \end{equation}
 
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
-The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
-Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.
-
-With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
-Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
-All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
-While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
-\ref{fig:V1_EPI}B right).
-Currently available approaches to deriving the steady state response of the system are limited.
+From first principles, we may attempt to fit or infer $z$ based on recorded V1 population responses at varying contrasts (\textbf{TODO} AP to add citation).
+However, such a class of constrained models rarely produce good fits that generalize well.
+Instead, we can use EPI to condition $z$ on one of the most salient properties of these data-sets called ``S-V flipping" --  a winner-take-all relationship between the S- and V-populations.
+At low contrast values, the V-population dominates the S-population, and at greater contrast values, the S-population dominates the V-population.
+We formulate the emergent property of S-V flipping with a normalized statistic measuring the difference in steady state ($x_{ss}$) between the S- and V-population at a given contrast:
+\begin{equation}
+d_{S,V}(c; z) = \frac{f_r(x_{ss,S}(c; z)) - f_r(x_{ss,V}(c; z))}{|f_r(x_{ss}(c; z))|_2}
+\end{equation}
 
@@
 The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$.
-This posterior is bimodal and concentrated in planes $dh_E > 0$ and $dh_E < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.
-Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two stds).
-\textbf{C}. Posterior predictive distribution of $d_{S-V}(h)$ of each mode shows that the $z_1$-mode produces V-to-S flipping with increasing contrast and the $z_2$-mode produces S-to-V flipping.
+This posterior is bimodal and concentrated in planes $h^{(c)} > 0$ and $h^{(c)} < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.
+Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two standard deviations).
+\textbf{C}. Posterior predictive distribution of $d_{S,V}$ for each mode.  The $z_1$-mode produces V-to-S flipping with increasing contrast.
 \textbf{D}.
-Model simulations at the mode $z_1$ at $b_1$ (left), $b_1 + dh_1^{(c)}$ (middle), and steady state solutions for varying levels of contrast (right).
-Shaded area is one standard deviation according to randomness of $\epsilon$.
+Model simulations at the mode $z_1$ at $c=0$ (solid) and $c=1$ (dashed).
 \textbf{E}.
-Same as D.
-for $z_2$.
+Mean system response at varying contrasts for $z_1$.
+Error bars show noise in rate.
 \textbf{F}.
-...
+Plot of noise with contrast from E.
+Error bars are standard error measured across simulations.
  }\label{fig:SV_flip}
@@
 
-TODO Summarize S-V flipping phenomena and why it's a good EP
-TODO Explain that we can't really fit the whole data-set of responses yet because the model is quite constrained.
-TODO Motivate this input model, where $dh^{(c)}$ is the direction of increasing contrast.
+For S and V to flip their steady states, the difference between the S- and V-population rates at no contrast ($c=0$) must have opposite sign at full contrast ($c=1$).
+Therefore, we stipulate the emergent property of S-V flipping to require the product of $d_{S,V}(c=0)$ and $d_{S,V}(c=1)$ to be appreciably negative.
+The means and variance of this emergent property statistics were chosen such that the posterior predictive distribution always produced S-V flipping.
 
 \begin{equation}
- \begin{bmatrix} h_E \\ h_P \\ h_S \\ h_V \end{bmatrix} =  \begin{bmatrix} b_E \\ b_P \\ b_S \\ b_V \end{bmatrix} +  \begin{bmatrix} dh^{(c)}_E \\ dh^{(c)}_P \\ 0 \\ 0 \end{bmatrix} 
+\begin{split}
+\mathcal{B}_{S,V} \triangleq \text{\hspace{.5cm}} \mathbb{E}\left[ d_{S,V}(0)d_{S,V}(1) \right] &= -0.25 \\ 
+\text{Var}\left[ d_{S,V}(0)d_{S,V}(1)\right] &=  0.125^2
+\end{split}
 \end{equation}
 
-Since we are unsure what either the baseline input $b$ and contrast-dependent change in input $dh^{(c)}$ should be, we treat them as free parameters when running EPI.
-\begin{equation}
- z =  \begin{bmatrix} b_E & b_P & b_S & b_V & dh^{(c)}_E & dh^{(c)}_P \end{bmatrix}^\top
-\end{equation}
-We consider positive baseline inputs $b_\alpha \in \left[0, 3 \right]$ and small contrast-dependent changes in input $|dh_\alpha| \leq 1$.  
-To find parameters resulting in S-V flipping, we focus on models driven by nominal amounts of external noise $\sigma_{\eta} = 0.1$.
+We ran EPI to obtain the parameter distribution of neuron-type population input parameters $z$ that produce S-V flipping (Fig. \ref{fig:SV_flip}B).
+We considered positive baseline inputs $b_\alpha$ from 0 to 25mV and contrast-dependent changes in $h^{(c)}$ from -10 to 10mV.
+The shape and structure of this distribution reveal some important properties of this model.
+\begin{enumerate}
+\item Two types of $z$ result in S-V flipping. (There are two modes separated by the hyperplane $h^{(c)} = 0$.)
+\item S-V flipping is sensitive to anticorrelated changes in $b_S$ and $b_V$.
+(The pairwise marginal distribution of $q_\theta(b_S, b_V \mid \mathcal{B}_{S,V})$ reveals strong correlation between these parameters.)
+\item S-V flipping is most degenerate with respect to input to the P population.
+(The marginal distributions of $q_\theta(h_P \mid \mathcal{B}_{S,V})$ and  $q_\theta(dh_P \mid \mathcal{B}_{S,V})$ are approximately uniform along their allowed range.)
+\end{enumerate}
+Simulation from the posterior reveals that the biologically plausible mode ($h^{(c)} > 0$) matches experiments by producing V-to-S flipping with contrast.
+Alternatively, the biologically implausible mode ($h^{(c)} > 0$) produces S-to-V flipping with contrast -- the opposite effect from experiments (Fig. \ref{fig:SV_flip}C).
 
-We formulate the emergent property of S-V flipping with a statistic measuring he difference in steady state ($x_ss$) between the S- and V-population at a given input:
-\begin{equation}
-d_{S-V}(h) = x_{ss,S}(h) - x_{ss,V}(h).
-\end{equation}
-For S and V to flip their steady states, the difference between the S- and V-population rates at $h = b$ must have opposite sign from $h = b + dh$.
-Therefore, we stipulate the emergent property of S-V flipping to require the product between $d_{S-V}(b)$ and $d_{S-V}(b+dh)$ to be appreciably negative.
-Second we stipulate that the differences between S and V in each input condition cancel out on average ($d_{S-V}(b) + d_{S-V}(b+dh)$ is 0 on average).
-The means and variances of the emergent property statistics were sensibly chosen based on some inexpensive model simulations within the parameter bounds.
+We examined the responses of this circuit model to varying contrasts using the highest log-probability parameterization ($z_1$) of the plausible mode.
+Simulations show the stochastic responses of the S- and V-population are indeed flipped at no and full contrast (Fig. \ref{fig:SV_flip}D).
+This flipping occurs gradually with graded increases in contrast (Fig. \ref{fig:SV_flip}E).
+Interestingly, the responses are noisier at intermediate contrasts (Fig. \ref{fig:SV_flip}F).
+This suggests a new hypothesis explaining phenomena of variability in V1: noise is greater in stabilized supralinear networks during winner-take-all competition between the S- and V- populations, but is quenched as either the S- or V-population overtakes the other.
+
+\section{V1 Supplemental section}\label{methods_V1}
+
+The dynamics are driven by the rectified and exponentiated sum of recurrent inputs $Wx$, external inputs $h$, and external noise $\epsilon \sim \mathcal{N}(0, \sigma_{\epsilon}^2)$:
 
 \begin{equation}
-\mathcal{B}_{S-V} \triangleq \mathbb{E}\begin{bmatrix} d_{S-V}(b)d_{S-V}(b+dh) \\ \left( d_{S-V}(b)d_{S-V}(b+dh) - (-0.25) \right)^2 \\  d_{S-V}(b) + d_{S-V}(b+dh) \\  \left( d_{S-V}(b) + d_{S-V}(b+dh)\right)^2 \end{bmatrix} =
-\begin{bmatrix} -0.25 \\ 0.125^2 \\ 0 \\ 0.125^2 \end{bmatrix}
+T \frac{dx}{dt} = -x + x_{\text{rest}} +Wf_r(x) + b + h + \epsilon
 \end{equation}
 
-We ran EPI to inspect the structure of the posterior distribution of $z$ conditioned on S-V flipping shown in Fig 2.  
-It is clear from this visualization that S-V flipping in this V1-model is sensitive with respect to some parameter settings and robust with respect to others.
-Additionally, the posterior is bimodal: there is one mode in each of the hyperplanes $dh_E < 0$ and $dh_E > 0$.
+where $T$ is a diagonal matrix with $\tau_E = 20\text{ms}$ and $\tau_P = \tau_S = \tau_V = 10\text{ms}$, $x_{\text{rest}} = -65\text{mV}$, and $f_r(x) = \left[x - x_{\text{rest}} \right]_+^2$.
 
-The structure of the V1 model S-V flipping posterior yields the following insights:
-\begin{enumerate}
-\item The marginal distributions of $q_\theta(h_P \mid \mathcal{B}_{S-V})$ and  $q_\theta(dh_P \mid \mathcal{B}_{S-V})$ are approximately uniform along their allowed range.  
-The approximate uniformity of the $P-population$ parameter marginal distributions shows that the P-population plays little role in S-V flipping. 
-\item The pairwise marginal distribution of $q_\theta(h_S, h_V \mid \mathcal{B}_{S-V})$ shows strong correlation between these parameters.
-The strong correlation between $h_S$ and $h_V$ in the posterior reveals that S-V flipping is sensitive with respect to the baseline inputs to the S- and V-populations.
-\textbf{Augment this to show that multidim info from Hessian is useful.}
-The Hessian provided by EPI at the modes indicates that coordinated increases of $h_S$ and $h_V$ by a ratio of $\frac{h_S}{h_V} = 0.701$ will preserve S-V flipping, while changes in an orthogonal dimension will disrupt it.
+The noise is an modeled as an Ornstein-Uhlenbeck process:
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
+\end{equation}
 
-\item This distribution is multimodal.
-\end{enumerate}
+where $\tau_{\text{noise}}$ is slow at 50ms, $\sigma_E = 0.2\text{mV}$, $\sigma_P = \sigma_S = \sigma_V = 0.1\text{mV}$, and $dB$ is a standard Wiener process.
+
+We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
+The input to this circuit is comprised of a nominal baseline input $b$ and additive contrast-dependent input $h$
+\begin{equation}
+h = c \begin{bmatrix} h^{(c)} \\ h^{(c)} \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1.
+Specifically, Billeh et al.
+\cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
+\begin{equation}
+M = \begin{bmatrix} 
+0.36 & -0.48 & -0.31 & -0.28 \\
+1.49 & -0.68 & -0.50 & -0.18 \\
+0.86 & -0.42 & -0.15 & -0.32 \\
+1.31 & -0.41 & -0.52 & -0.37 \end{bmatrix}
+\end{equation}
+and connection probability
+\begin{equation}
+C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
+0.395 & .451 & 0.857 & 0.02 \\
+0.182 & 0.03 & 0.082 & 0.625 \\
+0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
+\end{equation}
+
+Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
+\begin{equation}
+W = C \odot M = \begin{bmatrix} 
+0.0576 & -0.197 & -0.131 & -0.0244 \\
+0.589 & -0.307 & -0.429 & -0.00360 \\
+0.157 & -0.0126  & -0.0123 & -0.200 \\
+0.138 & -0.0902 & -0.400 & -0.0104 \end{bmatrix}.
+\end{equation}
+
+%\begin{figure}[h]
+%\caption{\small \textbf{A}. ...
+% }\label{fig:drdh}
+%\begin{center}
+%\includegraphics[scale=.3]{figs/FigSabc/FigSabc.pdf}
+%\end{center}
+%\end{figure}
+%
+%s
+%\begin{figure}[h]
+%\caption{\small \textbf{A}. ...
+% }\label{fig:drdh}
+%\begin{center}
+%\includegraphics[scale=.15]{figs/FigSX/FigSX.pdf}
+%\end{center}
+%\end{figure}
+
+\bibliography{epi}
+\bibliographystyle{unsrt}
+
+\end{document}
+
+% With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
+% Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
+% All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
+% While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
+\ref{fig:V1_EPI}B right).
+% Currently available approaches to deriving the steady state response of the system are limited.
 
@@
 %The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).
-
-\section{V1 Supplemental section}\label{methods_V1}
-
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1.
-Specifically, Billeh et al.
-\cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
-\begin{equation}
-M = \begin{bmatrix} 
-0.36 & -0.48 & -0.31 & -0.28 \\
-1.49 & -0.68 & -0.50 & -0.18 \\
-0.86 & -0.42 & -0.15 & -0.32 \\
-1.31 & -0.41 & -0.52 & -0.37 \end{bmatrix}
-\end{equation}
-and connection probability
-\begin{equation}
-C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
-0.395 & .451 & 0.857 & 0.02 \\
-0.182 & 0.03 & 0.082 & 0.625 \\
-0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
-\end{equation}
-
-Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
-\begin{equation}
-W = C \odot M = \begin{bmatrix} 
-0.0576 & -0.197 & -0.131 & -0.0244 \\
-0.589 & -0.307 & -0.429 & -0.00360 \\
-0.157 & -0.0126  & -0.0123 & -0.200 \\
-0.138 & -0.0902 & -0.400 & -0.0104 \end{bmatrix}.
-\end{equation}
-
-\begin{figure}[h]
-\caption{\small \textbf{A}. ...
- }\label{fig:drdh}
-\begin{center}
-\includegraphics[scale=.3]{figs/FigSabc/FigSabc.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\caption{\small \textbf{A}. ...
- }\label{fig:drdh}
-\begin{center}
-\includegraphics[scale=.12]{figs/FigSX/FigSX.pdf}
-\end{center}
-\end{figure}
-
-\bibliography{epi}
-\bibliographystyle{unsrt}
-
-\end{document}

### 2020-11-25 written/SC_model/SC_model.tex
latest updates

--- 2020-11-07 written/SC_model/SC_model.tex
@@
 Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
-C.
-The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
-Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
-D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
-E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
+C. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
+D. Simulations of SC networks with parameterizations indicated by the stars in panel C.  Shading  shows standard deviations across trials.
+E. The effect of delay period inactivation on accuracy versus Schur eigenvalues.
+F. Same as D with delay period inactivation.}
 \label{fig:SC}
@@
 We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
-There is strong pairwise correlation between each parameter of the connectivity matrix.
-Shading the distribution by $p_P$, we see that positive shifts in the parameters result in lower accuracy in the Pro task.
-Indeed, this difference in accuracy is apparent when we simulate this model for greater parameter values (black star) and lower parameter values (gray star) (Fig. \ref{fig:SC}D).
-
-To make sense of this inferred distributions, we followed the approach of Duan et al.
+To make sense of this inferred distribution, we followed the approach of Duan et al.
 by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-SX).
+\ref{fig:SC_Eigs}A).
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
 The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.
-We found that $\lambda_{\text{all}}$ is anti-correlated with $p_P$ and $\lambda_{\text{task}}$ is correlated with $p_P$.
-TODO interpret further.  Maybe push further on the optogenetics analysis.
+
+In agreement with experimental results form Duan et al., we found the optogenetic silencing during the delay period consistently decreased performance in the Anti task by 15-30\%, but had no consistent effect on the Pro task (Fig.
+\ref{fig:SC}E).
+For example, the networks indicated by the black star suffers a steep drop in accuracy in the Anti task with delay period inactivation (Fig. \ref{fig:SC}F).  
+Networks resilient to delay period inactivation had greater $\lambda_{\text{task}}$ and $\lambda_{\text{task}}$ and lower $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ (Fig. \ref{fig:SC}E).
 
@@
 \begin{equation}
-\mathbf{x}_\alpha = f(\mathbf{u}) = \left(\frac{1}{2}\tanh\left(\frac{\mathbf{u}_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
+\mathbf{x}_\alpha = f(\mathbf{u}) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
 \end{equation}
@@
 \begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
 \end{equation}
@@
 \begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{x \sim p(x \mid z)}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
 \end{equation}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.9]{figs/figSX1.pdf}
+\end{center}
+\caption{\small Sentence 1 .
+A.
+The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
+Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
+B. Task accuracy vs. Schur eigenvalue.}
+\label{fig:SC_Eigs}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.6]{figs/figSX2.pdf}
+\end{center}
+\caption{\small
+Same as Fig. 1C shaded by Anti task accuracy.}
+\label{fig:SC_EPI2}
+\end{figure}
+
 \bibliography{epi}

### 2020-11-25 written/V1_model/V1_model.tex
latest updates

--- 2020-10-01 written/V1_model/V1_model.tex
@@
 \begin{center}                  % Center the following lines
-{\Large Input/output properties of neuron-type population model of V1} \\
-Sean Bittner \\
-May 12, 2020 \\
+{\Large Draft of new V1 section} \\
+Sean Bittner, Agostina Palmigiano \\
+November 24, 2020 \\
 \end{center}
 
-\section{Introduction}
-In section 3.3 of our bioRxiv preprint \cite{bittner2019interrogating}, we analyze the input/output properties of a 4 neuron-type model of V1.  
-We identify degeneracies and sensitivities of neuron-type responses w.r.t specific inputs by visual inspection of the EPI distributions.  
-In my opinion, these ``novel insights" are not particularly interesting.
-Furthermore, these insights are easily obtained using an existing method -- approximate Bayesian computation (ABC).
+\section{Exploratory analysis of V1 with EPI produces a novel theory}\label{results_V1}
+Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
+However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+\ref{fig:SV_flip}A) \cite{pfeffer2013inhibition}.
+Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
+Here, we use EPI to elucidate the mechanisms of neuron-type stability in V1 at different levels of contrast.
 
-In this write-up, I present a new line of inquiry in the same V1 circuit model as the paper.
-Rather than conditioning on levels of increase of each neuron-type, we condition on the stabilization of each neuron-type.
-EPI then presents us with a distribution of stimulation to this V1 circuit that does not change much a particular neuron-type response.
+We consider contrast responses of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:Fig3}A) with a state comprised of each neuron-type population's rate $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Each population of neuron-type $\alpha$ receives recurrent input $\left(W f_r(x)\right)_\alpha$ from synaptic projections, where $W$ is the connectivity matrix and $f_r = \left[\cdot\right]_+^2$ is the rate nonlinearity.
+Driven by a baseline input $\mathbf{b}$, the circuit evolves form an initial condition $\mathbf{x}(0)$ to a steady-state solution (Fig. \ref{fig:Fig3}B, solid).
+When slow noise $\mathbf{\epsilon}$ is introduced, circuit activity fluctuates around this solution (Fig.
+\ref{fig:Fig3}B dashed), and the model is then a stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} (see Section \ref{methods_V1}).
+As contrast is enhanced, input to the E- and P-populations via $h^{(c)}$ increases causing the steady state solution (and fluctuations thereabout) to change  (Fig. \ref{fig:Fig3}C).
+In this analysis, we consider $W$,$b$, and $h^{(c)}$ that have been fit to contrast responses in mouse V1 using the deterministic model (TODO cite, see Section \ref{methods_V1}).
+As contrast changes, so does the degree of stochastic variability of each neuron-type (Fig. \ref{fig:Fig3}D).
+Moreover, the circuit transitions into an inhibition-stabilized network (ISN) with increasing contrast (Fig. \ref{fig:Fig3}D inset).
+While the ISN-regime and it's effects are well understood in E-I networks, it's role in the stability of this SSSN with inhibitory multiplicity has not been explored.
 
-First, I show EPI converges to a distribution covering the set of samples produced by ABC.
-This important sanity check, along with some ABC bias visualization, will make nice appendix figures.
-Then, I show how EPI readily identifies the most efficacious patterns of stimulation.
-Finally, we confirm two distinct mechanisms of E-population stabilization using EPI on the V1 model with inactivated populations.
+To study the role of inhibition stabilization in this network, we explore the changes in input 
+\begin{equation}
+\mathbf{z} = \begin{bmatrix} dh_E & dh_P & dh_S & dh_V \end{bmatrix}^{\top}
+\end{equation}
+resulting in each neuron-type's stability around it's mean rate.
+The emergent property of ``neuron-type stability" for population $\alpha$  at contrast level $c$ is defined as 
+\begin{equation}\label{eq:EP}
+\begin{split}
+\mathcal{X}(\alpha, c) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  0  ~~\triangleq~~ \bm{\mu}  \\ 
+\text{Var}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.05^2 \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
+\end{split}
+\end{equation}
 
-\section{V1 model (same as bioRxiv post)}
-We consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (Fig. 1A). 
-Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:
+\clearpage
+\begin{figure}[h]
+\caption{\small \textbf{A}.
+Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.
+Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
+\textbf{B}.
+EPI posterior $q_\theta(z \mid \mathcal{B}_{S-V})$ for S-V flipping.
+The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$.
+This posterior is bimodal and concentrated in planes $h^{(c)} > 0$ and $h^{(c)} < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.
+Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two standard deviations).
+\textbf{C}. Posterior predictive distribution of $d_{S,V}$ for each mode.  The $z_1$-mode produces V-to-S flipping with increasing contrast.
+\textbf{D}.
+Model simulations at the mode $z_1$ at $c=0$ (solid) and $c=1$ (dashed).
+\textbf{E}.
+Mean system response at varying contrasts for $z_1$.
+Error bars show noise in rate.
+\textbf{F}.
+Plot of noise with contrast from E.
+Error bars are standard error measured across simulations.
+ }\label{fig:Fig3}
+\begin{center}
+\includegraphics[scale=.18]{figs/Fig3/Fig3.pdf}
+\end{center}
+\end{figure}
+\clearpage
+
+\section{V1 Supplemental section}\label{methods_V1}
+
+The dynamics are driven by the rectified and exponentiated sum of recurrent inputs $Wx$, external inputs $h$, and external noise:
 
 \begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
+\tau \frac{dx}{dt} = -x +Wf_r(x) + b + h + \epsilon
 \end{equation}
 
+where $\tau = 1 \text{ms}$, and $f_r(x) = \left[x \right]_+^2$.
+
+The noise is modeled as an Ornstein-Uhlenbeck process:
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
+\end{equation}
+
+where $\tau_{\text{noise}}$ is slow at 5ms, $\sigma_\alpha =  0.01 $, and $dB$ is a standard Wiener process.
+
+We considered the stability of this stochastic stabilized supralinear network using connectivity ($W$) and input ($b$ and $h^{(c)}$) parameters fit to the deterministic model (TODO cite)
+
+\begin{equation}
+W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
+W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
+W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
+W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
+ \begin{bmatrix} .785 & -.102 & -1.24 & -.306 \\
+.809 & -.100 & -.620 & -.264 \\
+.833 & -.000128 & -.0000483 & -.742 \\
+.708 & -.306 & -.452 & -.0000605 \\
+ \end{bmatrix},
+\end{equation}
+
+\begin{equation}
+b =  \begin{bmatrix} b_E \\ b_P \\ b_S \\ b_V \end{bmatrix} =
+ \begin{bmatrix} .590 \\ .504 \\ .515 \\ .670 \end{bmatrix} ,
+\end{equation} 
+and
+\begin{equation} 
+h^{(c)} = \begin{bmatrix} h^{(c)}_E \\ h^{(c)}_P \\ 0 \\ 0 \end{bmatrix} = 
+\begin{bmatrix} .595 \\ .396 \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+This ISN coefficient $\gamma(x; z)$ is calculated by simulating
+
+The expectation and variance in the emergent property definition are taken over both the posterior distribution and the stochasticity of the SSSN:
+\begin{equation}
+= \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  = \mathbb{E}_{\mathbf{z}\sim q_\theta} \left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \begin{bmatrix} x_\alpha(\mathbf{x}; \mathbf{z},c) - x_\alpha(\mathbf{x}; \mathbf{0},c) \end{bmatrix}\right].
+\end{equation}
+
+\begin{figure}[h]
+\caption{\small SX 1
+ }\label{fig:SX1}
 \begin{center}
-\textbf{Figure 1: V1 model} \\
-\includegraphics[scale=.4]{figs/V1_responses.pdf}
+\includegraphics[scale=.5]{figs/FigSX/Fig_SX1.pdf}
 \end{center}
+\end{figure}
 
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic}.
+\begin{figure}[h]
+\caption{\small SX 2
+ }\label{fig:SX2}
+\begin{center}
+\includegraphics[scale=.5]{figs/FigSX/Fig_SX2.pdf}
+\end{center}
+\end{figure}
+
+\bibliography{epi}
+\bibliographystyle{unsrt}
+
+ With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
+ Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
+ All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
+ While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
+\ref{fig:V1_EPI}B right).
+ Currently available approaches to deriving the steady state response of the system are limited.
+
+%\section{Stability analysis}
 The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
@@
 \mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.
+\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.\\
 \end{equation}
@@
 \\
-\begin{center}
-$q_\theta(dh \mid \mathcal{B}(E,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(E,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_E.pdf} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_E.png} \\
-\clearpage
-$q_\theta(dh \mid \mathcal{B}(P,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(P,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_P.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_P.png} \\
-$q_\theta(dh \mid \mathcal{B}(S,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(S,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_S.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_S.png} \\
-$q_\theta(dh \mid \mathcal{B}(V,\sigma=0.25))$ \hspace{1.2in} $dh \sim ABC(V,\sigma_{ABC}=0.25)$. \\
-\includegraphics[scale=0.25]{figs/V1_drdh_EPI_V.png} 
-\includegraphics[scale=0.25]{figs/V1_drdh_ABC_V.png} \\
-\end{center}
 
@@
 The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).
-The regions of the ABC pairplots which have $|dx_\alpha| \approx 0.5$ are far less probable in the EPI distributions (by design), and can account for most mismatch observed.
-As a note, there is a discontinuity in Figure 2, $EPI(E, \sigma=0.25)$ in the $dh_E$-$dh_S$ plane.
-I'd like to see if I can run the optimization for more random seeds so that we obtain a more connected distribution.
-
-\begin{center}
-\textbf{Figure 3: Biased ABC} \\
-\includegraphics[scale=0.4]{figs/V1_ABC_Tx.pdf}
-\end{center}
-
-\clearpage
-
-\begin{center}
-\textbf{Figure 4: Unbiased EPI} \\
-\includegraphics[scale=0.4]{figs/V1_EPI_Tx.pdf}
-\end{center}
-
-\section{EPI-enabled analyses}
-\subsection{Dimensions of sensitivity}
-Each arrow in the EPI distributions (Fig.
-2, left) points in the direction of maximal sensitivity indicated by the hessian $\partial^2 \log q_\theta(dh) / \partial dh \partial dh^\top$ evaluated at the $dh$ at the base of the arrow.
-These patterns of input elicit a greater $dx_\alpha^2$ than (at least) 98.8\% of random stimulation patterns.
-Each sensitivity vector is representative of a different stabilization mechanism.
-
-\subsection{Confiriming distinct mechanisms of stabilization}
-The sensitivity arrows on the EPI distribution for E-population stability (Fig. 2, top, left) indicate two distinct stability mechanisms for increased $dh_E$.
-In one case, $dh_P$ must increase along with increased $dh_E$ (Fig. 2, top, left black arrow $v_P$).
-In the other case, $dh_S$ must increase along with increased $dh_E$ (Fig. 2, top, left gray arrow $v_S$).
-These inferred distributions and sensitivity vectors $v_P$ and $v_S$ motivate the following hypothesis:
-
-H: If the P- and S-populations are inactivated, the only E-stabilization mechanism with increasing $dh_E$ will be that of the other population (S- or P-population, respectively) .
-
-We can test this hypothesis with EPI, since it is optimized to identify the distribution of all inputs resulting in E-population stability.
-We ran EPI with the same emergent property as Eq 2 ($\alpha=E$, $\sigma=0.25$) for two different inactivation scenarios.
-In the P-inactivated case, we changed the baseline input to $b =b_{\not P} \triangleq \left[ 1, -5, 1 ,1.25 \right]^\top$, and to $b =b_{\not S} \triangleq \left[ 1, 1, -5 ,1.25 \right]^\top$ in the S-inactivated case.
-To test for the existence of stabilization mechanisms corresponding to $v_P$ and $v_S$, we measured the most-sensitive dimension of the Hessian at all points in the inferred distribution.  
-We declare that the P- or S-inactivated circuit lacks a stability mechanism, if no sensitivity vectors throughout the distribution are aligned with the corresponding sensitivity vector.
-
-\begin{center}
-\textbf{Figure 5}: Stability mechanisms of inactivated circuits.
-\includegraphics[scale=.5]{figs/cos_dists.pdf}
-\end{center}
-
-Confirming our hypothesis, we found that only the $v_S$ stability mechanism is present in the P-inactivated circuit (Figure 5A), and that only the $v_P$ stability mechanism is present in the S-inactivated circuit (Figure 5B).
-As expected, the samples of the inferred distributions aligned with the stability mechanism correspond to the sloped regions orthogonal to $v_S$ and $v_P$ (Figure 6).
-
-\clearpage
-
-\begin{center}
-\textbf{Figure 6}: Inactivated circuit stability distributions.
-\includegraphics[scale=.4]{figs/EPI_inactivation.pdf}
-\end{center}
-
-\section{Summary}
-We have improved on our previous EPI analysis of V1 in the following ways:
-\begin{enumerate}
-\item We utilize a novel ability of EPI (Hessian evaluation).
-\item Our scientific conclusion is about mechanisms of stability rather than high-level characteristics of responses.
-\item We have validation of structure via ABC.
-\item We characterize the issue of bias in ABC (and lack thereof in EPI).
-\item The emergent property of ``neuron-type stabilization" follows the ethos of the definition of emergent property better than ``neuron-type population rate increase of $y$".
-\end{enumerate}
-
-I recommend substituting this work for section 3.3 in the bioRxiv manuscript. I'd be happy to write up a draft of such a section.
-
-%\section{Isolated modulation paths using EPI}
-%Let's say we want to find a stimulation pattern that changes the $\alpha$-population steady-state response $dx_\alpha$ as much as possible while keeping the rest the same.
-%One can not simply stimulate the $\alpha$-population, since one must take into account the effects of the recurrent connectivity.
-%Furthermore, such an input pattern producing isolated modulation may change nonlinearly.
-%
-%Beginning at $dh = \textbf{0}$, we seek a sequence of differential inputs that will increasingly change $dx_\alpha$, while maintaining $dx_\beta$, $\beta \in \{E, P, S, V\} \setminus	\{\alpha\}$.
-%At each step in the path of inputs, the direction is selected to maximize change in $dx_\alpha$ and minimize change in $dx_{\beta \neq \alpha}$.
-%These input selections can be guided by the EPI distributions $q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ that result in the modulation of the $\alpha$ population within a given range.
-%
-%We leverage the availability of a functional representation of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$ to calculate the the hessians of the EPI distributions at each step of $dh$ in the path
-%\begin{equation}
-%\text{Hess}_\alpha(dh) = \frac{\partial^2 \log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))}{\partial dh \partial dh^\top}.
-%\end{equation}
-%
-% We select the optimal direction of isolated modulation as vector v, which points in the maximally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\alpha, 0.2))$, and minimally decreasing dimensions of $\log q_\theta(dh \mid \mathcal{B}(\beta, 0.2))$, $\beta \neq \alpha$:
-% \begin{equation}
-% v_\alpha^*(dh) = \argmin_v v^\top \text{Hess}_\alpha(dh) v - \sum_{\beta \neq \alpha} v^\top \text{Hess}_\beta(dh) v,
-% \end{equation}
-%  where $v\top v$ = 1.  $v_\alpha(dh)$ is obtained via an eigendecomposition:
-%  \begin{equation}
-%  v_\alpha^*(dh) = \argmin_v v^\top M_\alpha(dh) v
-% \end{equation}
-% \begin{equation}
-%M_\alpha(dh) = \text{Hess}_\alpha(dh) - \sum_{\beta \neq \alpha} \text{Hess}_\beta(dh).
-% \end{equation}
-% The solution is the eigenvector with smallest eigenvalue of $M_\alpha(dh)$.
-% 
-% \includegraphics[scale=0.4]{figs/hess_paths.pdf}
-%
-
-\bibliography{epi}
-\bibliographystyle{unsrt}
 

### 2020-11-25 written/eLifeRevision/Bittner_eLife2020.tex
latest updates

@@
 \usepackage{floatrow}
+\usepackage{bm}
 
@@
 
-Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
+%Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
+Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference (Fig. \ref{fig:STG}A).  
 First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
@@
 Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
-A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
+%A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
+A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}C, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
 Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  
@@
 
-Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
-To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-This emergent property is shown in Figure 1A at a frequency of 0.53Hz.
-
-Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
+%Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
+Second, once the model is selected, one defines the emergent property, the measurable behavior of scientific interest.  
+%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
+To continue our running STG example, one such emergent property is the phenomenon of \emph{unified intermediacy} -- in certain parameter regimes, the frequency of all neurons match at an intermediate firing rate.
+This emergent property is shown in Figure \ref{fig:STG}D at a frequency of 0.53Hz.
+
+%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
+Third, qualitative parameter analysis ensues: since mathematical analysis of frequency is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.  
 Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
@@
 \caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
+%A.
+For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.
+%In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
+%Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
+%Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.  
+%B. The EPI distribution of STG model parameters producing network syncing.  
+%Samples are colored by log probability density.  
+%Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
+%Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
+%Simulated activity is shown for three samples (stars). 
+%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
+%$v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  
+%C.
+Deep probability distributions map a latent random variable $\mathbf{w}$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = f_{\bm{\theta}}(\mathbf{w})$ distributed as $q_{\bm{\theta}}(\mathbf{z})$.
+%D.
+EPI optimization: To learn the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ of model parameters that produce an emergent property, the emergent property statistics $T(\mathbf{x})$ are set in expectation over model parameter samples $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and model simulations $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ to emergent property values $\bm{\mu}$.
 A.
-For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.
+For a choice of model (STG) and emergent property (unified intermediacy), emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing unified intermediacy.
 In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
 Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.
-B.
-The EPI distribution of STG model parameters producing network syncing.
+Deep probability distributions map a simple random variable $\mathbf{w}$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = f_{\bm{\theta}}(\mathbf{w})$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+B. The EPI distribution of STG model parameters producing network syncing.  
 Samples are colored by log probability density.  
 Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
-Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $v_1$ (solid) and $v_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-Simulated activity is shown for three samples (stars).
-(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.
+Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
+Simulated activity is shown for three samples (stars)in panel D. 
 $v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  
-C.
-Deep probability distributions map a latent random variable $w$ through a deep neural network with weights and biases $\theta$ to parameters $z = f_\theta(w)$ distributed as $q_\theta(z)$.
-D.
-EPI optimization: To learn the EPI distribution $q_\theta(z)$ of model parameters that produce an emergent property, the emergent property statistics $T(x)$ are set in expectation over model parameter samples $z \sim q_\theta(z)$ and model simulations $x \sim p(x \mid z)$ to emergent property values $\mu$.}
+D. The emergent property of unified intermediacy, in which all neurons are firing close to the same intermediate frequency.
+Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution. }
  \label{fig:STG}
@@
 
+%Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
+\ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
 Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
+\ref{fig:STG}A) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution.
 This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
-The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of network syncing (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
+%In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
+In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for unified intermediacy (the mode; Fig.
+\ref{fig:STG}B, yellow star), and how it decays based on changes away from the mode.
+The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of unified intermediacy (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
 Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig.
-1B, contours) and have diminished or preserved network syncing, respectively (Figure 1B inset and activity traces) (see Section \ref{methods_STG}).
+\ref{fig:STG}B contours) and have diminished or preserved network syncing, respectively (Fig.
+\ref{fig:STG}D activity traces, Fig.
+S TODO) (see Section \ref{methods_STG}).
 
 \subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
-
 Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
 First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
-In the running STG example, the model activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
-\begin{equation} C_m \frac{dx}{dt} = -h(x; z) = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right] 
-\end{equation} 
-where $C_m$=1nF, and $h_{\text{leak}}$, $h_{Ca}$, $h_K$, $h_{\text{hyp}}$, $h_{\text{elec}}$, and $h_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $x$ and $z = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).
+In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
+\begin{equation} C_m \frac{d\mathbf{x}}{dt} = -\mathbf{h}(\mathbf{x}; \mathbf{z}) = - \left[ \mathbf{h}_{leak}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{Ca}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_K(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{hyp}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{elec}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{syn}(\mathbf{x}; \mathbf{z})\right]
+\end{equation} 
+where $C_m$=1nF, and $\mathbf{h}_{\text{leak}}$, $\mathbf{h}_{Ca}$, $\mathbf{h}_K$, $\mathbf{h}_{\text{hyp}}$, $\mathbf{h}_{\text{elec}}$, and $\mathbf{h}_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).
 
 Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(x)$, $\omega_{\text{f2}}(x)$, etc.
+Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
 -- is close to an intermediate frequency of 0.53Hz.
-Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(x)$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
+Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
+
+%\begin{equation}\label{eq:EP}
+ %\mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
+ %\begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
+ %\end{equation}
 \begin{equation}\label{eq:EP}
- \mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
- \begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
- \end{equation}
-  which completes the quantification of the emergent property.
-
-Third, we perform emergent property inference: we find a distribution over parameter configurations $z$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_\theta(z) : \theta \in \Theta \right\}$, defined by a deep generative distribution
-of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.53 \\ 0.53 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
+ \text{Var}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
+\end{split}
+\end{equation}
+
+%which completes the quantification of the emergent property.
+The emergent property statistics $T(\mathbf{x}; \mathbf{z})$ along with their constrained means $\bm{\mu}$ and variances $\bm{\sigma}^2$ define the emergent property denoted $\mathcal{X}$.
+
+Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.
+This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
+This deep distribution is represented in Figure \ref{fig:STG}B (see Section \ref{methods_EPI}).  
 Then, mathematically, we must solve the following optimization program: 
+%\begin{equation} \label{eq:EPI}
+%\begin{split}
+%&\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
+%&\text{  s.t.  } \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right] \right] = \bm{\mu}, \\
+%\end{split}
+%\end{equation}
 \begin{equation} \label{eq:EPI}
 \begin{split}
-&\argmax_{q_\theta \in \mathcal{Q}} H(q_\theta(z)) \\
- &\text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu, \\
+q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) ~~=~~ &\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
+&\text{  s.t.  } ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
 \end{split}
 \end{equation}
-where $T(x), \mu$ are defined as in Equation \ref{eq:EP}, and $p(x|z)$ is the intractable distribution of data from the model, $x$, given that model's parameters $z$ (we access samples from this distribution by running the model forward).
-The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
+%where $T(\mathbf{x}), \bm{\mu}$ are defined as in Equation \ref{eq:EP}, and $p(\mathbf{x} | \mathbf{z})$ is the intractable distribution of data from the model, $\mathbf{x}$, given that model's parameters $\mathbf{z}$ (we access samples from this distribution by running the model forward).
+where $T(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
+%The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
 Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
@@
 
-EPI optimizes the weights and biases $\theta$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-The optimization is complete when the sampled models with parameters $z \sim q_\theta$ produce activity consistent with the specified emergent property (Fig. S4).
-Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
+EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
+The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
+%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
+Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}).
 Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).
@@
 In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
-Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
-Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
+%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
+Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $T(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
+%Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
+Equipped with this method, we may examine structure in posterior distributions or make comparisons between posteriors conditioned at different levels of the same emergent property statistic.
 We now prove out the value of EPI by using it to investigate and produce novel insights about three prominent models in neuroscience.
@@
 \caption{\small Hypothesis generation through EPI in a V1 model.  
-A.
-Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.
+A. Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   
 Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).  
-B.
-Linear response predictions become inaccurate with greater input strength.
-V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$.
+B. Linear response predictions become inaccurate with greater input strength.  V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$. 
 Stars indicate the linear response prediction.  
-C.
-EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$.
+C. EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$. 
 Supporting evidence for the four generated hypotheses are indicated by gray boxes with labels H1, H2, H3, and H4. 
-The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin).
-}
+The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin). }
  \label{fig:V1_EPI}
@@
 Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) -- composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
 \ref{fig:V1_EPI}A) \cite{pfeffer2013inhibition}.
@@
 {\caption{Confirming EPI generated hypotheses in V1. 
-A.
-Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$.
+A. Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$. 
 B-D Same plots for the P-, S-, and V-populations.  
@@
 \caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
-A.
-Rapid task switching behavioral paradigm (see text).
-B.
-Model of superior colliculus (SC).
-Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti.
+A. Rapid task switching behavioral paradigm (see text). 
+B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g.
-$vW_{PA}$ is a vertical weight from an Anti to a Pro population.
+Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
 C.
@@
 \end{center}
-\caption{\small Sources of error in an RNN solving a simple task.
-A.
-(left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$.
+\caption{\small Sources of error in an RNN solving a simple task.  A. (left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$. 
 (right) Error in this computation can come from over- or under-estimating the posterior mean or variance. 
@@
 EPI distribution of rank-1 RNNs executing Gaussian posterior conditioning.
-Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$  C.
-Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.
+Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$
+C.  Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.  
 Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.  
-D.
-The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.
+D. The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.  
 Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
@@
 \begin{flushleft}
-{\small Fig.
-S1: A.
-Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.
+{\small Fig. S1: A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
 B.
@@
 Dashed lines indicate the parameter axes.
-C.
-Entropy throughout the optimization.
+C. Entropy throughout the optimization.  
 At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
-D.
-Emergent property moments throughout optimization.
+D. Emergent property moments throughout optimization.  
 At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
@@
 the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig.
-S2A).
+$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). 
 The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
@@
 \begin{flushleft}
-{\small Fig.
-S2: A.
-Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$.
-B.
-Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+{\small Fig. S2: A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
 \end{flushleft}
@@
 Deep probability models typically consist of several layers of fully connected neural networks.  
-When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
-For $z' = f(z)$,
+When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.  For $z' = f(z)$,
 
@@
 \end{equation}
-The posterior disribution is analytically available when the prior is conjugate with the likehood.
+The posterior distribution is analytically available when the prior is conjugate with the likelihood.
 However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
@@
 \end{equation}
-$q_\theta(z)$ is the variational approximation to the posterior with variational parameters $\theta$.  We can write this KL divergence in terms of entropy of the variational approximation:
+$q_\theta(z)$ is the variational approximation to the posterior with variational parameters $\theta$.  
+We can write this KL divergence in terms of entropy of the variational approximation:
 \begin{equation}
@@
 \subsection{Theoretical models}\label{methods_theoretical_models}
-In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  Here, we provide the details of each model  and the related analyses.
+In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  
+Here, we provide the details of each model  and the related analyses.
 
@@
 \end{equation} 
-The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization,
-electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $z$.  
+The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $z$.
 The capacitance of the cell membrane was set to $C_m = 1nF$.
@@
 The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  
-The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$.
-$g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.
+The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  
 The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  
@@
 Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  
-The optimization converged after four epochs (Fig.
-S4).
+The optimization converged after four epochs (Fig. S4).
 
@@
 \begin{flushleft}
-{\small Fig.
-S4: EPI optimization of the STG model producing network syncing.
-A.
-Entropy throughout optimization.
-B.
-The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.
+{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
+A. Entropy throughout optimization. 
+B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
 Since $q_\theta(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.    
-C.
-The second moment emergent property statistics converge to the emergent property values.}
+C. The second moment emergent property statistics converge to the emergent property values.}
 \end{flushleft}
@@
 For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $z = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
+We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_\theta(z)$ to produce an isotropic 
+Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
 
@@
 \end{equation}
-All simulations of this system had $T=100$ time points, a time step $dt = 5$ms, and time constant $\tau = 20$ms.  The system was initialized to a random draw $x(0)_i \sim \mathcal{N}(1, 0.01)$.
+All simulations of this system had $T=100$ time points, a time step $dt = 5$ms, and time constant $\tau = 20$ms.  
+The system was initialized to a random draw $x(0)_i \sim \mathcal{N}(1, 0.01)$.
 
@@
 \end{equation}
-Let $F_{ij} = \gamma_i \delta(i,j)$, where $\gamma_i = f'(u_i)$.  Then, the linear response is
+Let $F_{ij} = \gamma_i \delta(i,j)$, where $\gamma_i = f'(u_i)$.  
+Then, the linear response is
 \begin{equation}
@@
 }
-where here $z$ is a gaussian integration variable. We can solve these equations by simulating the following Langevin dynamical system to a steady state:
+where here $z$ is a gaussian integration variable. 
+We can solve these equations by simulating the following Langevin dynamical system to a steady state:
 
@@
 For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.  
-In Fig.
-\ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.
+In Fig. \ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.  
 The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.  

### 2020-11-26 written/V1_model/V1_model.tex
V1 model updates

@@
 composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:SV_flip}A) \cite{pfeffer2013inhibition}.
+\ref{fig:Fig3}A) \cite{pfeffer2013inhibition}.
 Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
@@
 
-We consider contrast responses of a nonlinear dynamical V1 circuit model (Fig.
-\ref{fig:Fig3}A) with a state comprised of each neuron-type population's rate $x = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population of neuron-type $\alpha$ receives recurrent input $\left(W f_r(x)\right)_\alpha$ from synaptic projections, where $W$ is the connectivity matrix and $f_r = \left[\cdot\right]_+^2$ is the rate nonlinearity.
-Driven by a baseline input $\mathbf{b}$, the circuit evolves form an initial condition $\mathbf{x}(0)$ to a steady-state solution (Fig. \ref{fig:Fig3}B, solid).
+We considered the contrast responses of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:Fig3}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via nonlinearity $f_r = \left[\right]^2_+$ (see Section \ref{methods_V1}).
+The deterministic circuit model evolves form an initial condition $\mathbf{x}(0)$ to a steady-state solution (Fig. \ref{fig:Fig3}B, solid).
 When slow noise $\mathbf{\epsilon}$ is introduced, circuit activity fluctuates around this solution (Fig.
-\ref{fig:Fig3}B dashed), and the model is then a stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} (see Section \ref{methods_V1}).
-As contrast is enhanced, input to the E- and P-populations via $h^{(c)}$ increases causing the steady state solution (and fluctuations thereabout) to change  (Fig. \ref{fig:Fig3}C).
-In this analysis, we consider $W$,$b$, and $h^{(c)}$ that have been fit to contrast responses in mouse V1 using the deterministic model (TODO cite, see Section \ref{methods_V1}).
+\ref{fig:Fig3}B dashed), and the model is then a stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical}
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +f_r(W\mathbf{x} + \mathbf{h})
+\end{equation}
+At full contrast, input to the E- and P-populations increases 
+\begin{equation}
+\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c,
+\end{equation}
+changing the steady state solution (and fluctuations thereabout)  (Fig. \ref{fig:Fig3}C).
+In this analysis, we fixed $W$,$\mathbf{h}_b$, and $\mathbf{h}_c$, which have been fit to contrast responses in mouse V1 using the deterministic model (TODO cite, see Section \ref{methods_V1}).
 As contrast changes, so does the degree of stochastic variability of each neuron-type (Fig. \ref{fig:Fig3}D).
-Moreover, the circuit transitions into an inhibition-stabilized network (ISN) with increasing contrast (Fig. \ref{fig:Fig3}D inset).
-While the ISN-regime and it's effects are well understood in E-I networks, it's role in the stability of this SSSN with inhibitory multiplicity has not been explored.
+Moreover, the circuit transitions into an inhibition-stabilized network (ISN) ($\gamma(\mathbf{x}; \mathbf{z}) < 0$) with increasing contrast (Fig. \ref{fig:Fig3}D inset).
+While the ISN-regime and its effects are understood well in E-I networks, its role in this SSSN with inhibitory multiplicity has not been explored.
 
-To study the role of inhibition stabilization in this network, we explore the changes in input 
+To study the influence of circuit input on stabilization mechanisms in this stochastic model, we asked what changes in circuit input $d\mathbf{h}$ have little effect on the responses of each neuron-type.
 \begin{equation}
-\mathbf{z} = \begin{bmatrix} dh_E & dh_P & dh_S & dh_V \end{bmatrix}^{\top}
+\mathbf{z} = \begin{bmatrix} dh_E & dh_P & dh_S & dh_V \end{bmatrix}^{\top}.
 \end{equation}
-resulting in each neuron-type's stability around it's mean rate.
-The emergent property of ``neuron-type stability" for population $\alpha$  at contrast level $c$ is defined as 
+To formulate ``fluctuation constraints" for population $\alpha$  at contrast level $c$ as an emergent property in this stochastic model, we prescribed that the mean change in activity with $dh_\alpha$ is 0 and the variance over both EPI posterior and model stochasticity is a small amount
 \begin{equation}\label{eq:EP}
 \begin{split}
-\mathcal{X}(\alpha, c) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  0  ~~\triangleq~~ \bm{\mu}  \\ 
-\text{Var}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.05^2 \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
+\mathcal{X}(\alpha, c) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  0  ~~\triangleq~~ \mu  \\ 
+\text{Var}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.05^2 \end{bmatrix} ~~\triangleq~~ \sigma^2 .
 \end{split}
 \end{equation}
+By conditioning on this emergent property, we constrained stochastic fluctuations of circuit models sampled from the EPI posterior to be unbiased and vary to a fixed extent from the steady state solution at $d\mathbf{h}=\mathbf{0}$.
+
+Visualization of the resulting $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(\alpha, c))$ shows that the EPI posteriors largely agree with sets of parameters obtained using ABC (Fig. \ref{fig:SX1}).
+Focusing our attention on $P$-fluctuation constraints at 0\% and 100\% contrast ($\mathcal{X}(P, 0\%)$ and $\mathcal{X}(P, 100\%)$, respectively), we gain insight about how $d\mathbf{h}$ governs mechanisms of stability.
+At 0\% contrast, the SSSN is not inhibition-stabilized, and only local perturbations of positive $dh_E$ and negative $dh_P$ place the circuit in the ISN regime, while $P$-population fluctuations are kept near the 0\% contrast level (Fig.
+\ref{fig:Fig3}E).
+At 100\% contrast, where $h_E$ and $h_P$ are much greater, the SSSN is inhibition-stabilized, and $q_\theta(z \mid \mathcal{X}(P, 100\%)$ reveals distinct patterns of $d\mathbf{h}$ that obey the $P$-fluctuation constraint (Fig.
+\ref{fig:Fig3}F).
+As $dh_V$ decreases, either $dh_P$ and $dh_S$ must increase or decrease together.
+When $dh_P$ and $dh_S$ increase together, the circuit leaves the ISN regime, and when they are decreased, the strength of inhibition stabilization grows.
+The pattern in $d\mathbf{h}$ producing transitions out of the ISN regime is unapparent when running traditional ABC, even when SSSN noise is reduced to a negligible order of magnitude (Fig.
+\ref{fig:SX2}).
 
@@
 Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
-\textbf{B}.
-EPI posterior $q_\theta(z \mid \mathcal{B}_{S-V})$ for S-V flipping.
-The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$.
-This posterior is bimodal and concentrated in planes $h^{(c)} > 0$ and $h^{(c)} < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.
-Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two standard deviations).
-\textbf{C}. Posterior predictive distribution of $d_{S,V}$ for each mode.  The $z_1$-mode produces V-to-S flipping with increasing contrast.
-\textbf{D}.
-Model simulations at the mode $z_1$ at $c=0$ (solid) and $c=1$ (dashed).
-\textbf{E}.
-Mean system response at varying contrasts for $z_1$.
-Error bars show noise in rate.
-\textbf{F}.
-Plot of noise with contrast from E.
-Error bars are standard error measured across simulations.
+\textbf{B}. V1 circuit model simulations at 0\% contrast without (SSN, solid) and with (SSSN, solid) slow noise.
+\textbf{C}. Simulations at 100\% contrast.
+\textbf{D}. Average responses of the SSSN for all contrasts.  Error bars show the standard deviation of stochastic fluctuations.
+\textbf{E}. EPI posterior conditioned on P-fluctuation constraints at 0\% contrast.
+Parameter samples are colored by their ISN coefficient.
+Bottom-left inset is the posterior predictive distribution of $dx_P$.
+\textbf{F}. Same as E. at 100\% contrast.
  }\label{fig:Fig3}
@@
 
-The dynamics are driven by the rectified and exponentiated sum of recurrent inputs $Wx$, external inputs $h$, and external noise:
+The dynamics are driven by the rectified and squared sum of recurrent inputs $W\mathbf{x}$, external inputs $\mathbf{h}$, and external noise:
 
 \begin{equation}
-\tau \frac{dx}{dt} = -x +Wf_r(x) + b + h + \epsilon
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +f_r(W\mathbf{x} + \mathbf{h} + \mathbf{\epsilon})
 \end{equation}
-
-where $\tau = 1 \text{ms}$, and $f_r(x) = \left[x \right]_+^2$.
+where the input is comprised of a baseline and contrast-dependent input 
+\begin{equation}
+ \mathbf{h} =  \mathbf{h}_b +  c\mathbf{h}_c
+\end{equation} 
+$\tau = 1 \text{ms}$, $f_r(\mathbf{x}) = \left[\mathbf{x} \right]_+^2$, and $c \in \left[0, 1\right]$.
 
@@
 
-We considered the stability of this stochastic stabilized supralinear network using connectivity ($W$) and input ($b$ and $h^{(c)}$) parameters fit to the deterministic model (TODO cite)
+We considered the stability of this stochastic stabilized supralinear network with respect to an additive differential input $d\mathbf{h}$
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +f_r(W\mathbf{x} + \mathbf{h} + d\mathbf{h}+ \epsilon).
+\end{equation}
+
+We used connectivity ($W$) and input ($\mathbf{h}_b$ and $\mathbf{h}_c$) parameters fit to the deterministic model (TODO cite)
 
@@
 \begin{equation}
-b =  \begin{bmatrix} b_E \\ b_P \\ b_S \\ b_V \end{bmatrix} =
+\mathbf{h}_b =  \begin{bmatrix} h_{b,E} \\ h_{b,P} \\ h_{b,S} \\ h_{b,V} \end{bmatrix} =
  \begin{bmatrix} .590 \\ .504 \\ .515 \\ .670 \end{bmatrix} ,
@@
 \begin{equation} 
-h^{(c)} = \begin{bmatrix} h^{(c)}_E \\ h^{(c)}_P \\ 0 \\ 0 \end{bmatrix} = 
+\mathbf{h}_c = \begin{bmatrix} h_{c,E} \\ h_{c,P} \\ h_{c,S} \\ h_{c,V} \end{bmatrix} = 
 \begin{bmatrix} .595 \\ .396 \\ 0 \\ 0 \end{bmatrix}.
@@
 
-This ISN coefficient $\gamma(x; z)$ is calculated by simulating
+The expectation and variance in the emergent property are taken over both the posterior distribution and the stochasticity of the SSSN:
+\begin{equation}
+= \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  = \mathbb{E}_{\mathbf{z}\sim q_\theta(\mathbf{z})} \left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \begin{bmatrix} x_\alpha(\mathbf{x}; \mathbf{z},c) - x_\alpha(\mathbf{x}; \mathbf{0},c) \end{bmatrix}\right].
+\end{equation}
+Thus, over the EPI posterior predictive distribution, the mean and variance of the slow fluctuations about the steady state solution are constrained to $\mu = 0$ and $\sigma^2 = 0.05^2$
 
-The expectation and variance in the emergent property definition are taken over both the posterior distribution and the stochasticity of the SSSN:
+This ISN coefficient of the network at a given input is calculated by simulating the system to a steady state in the absence of noise
 \begin{equation}
-= \mathbb{E}_{\mathbf{z}}\begin{bmatrix} dx_\alpha(\mathbf{x}; \mathbf{z},c) \end{bmatrix}  = \mathbb{E}_{\mathbf{z}\sim q_\theta} \left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \begin{bmatrix} x_\alpha(\mathbf{x}; \mathbf{z},c) - x_\alpha(\mathbf{x}; \mathbf{0},c) \end{bmatrix}\right].
+    \gamma(\mathbf{x}; \mathbf{z}) = 1 - 2 \left[u_E \right]_+ W_{EE}
+\end{equation}
+where
+\begin{equation}
+    \mathbf{u} = W\mathbf{x} + \mathbf{h} + d\mathbf{h}.
 \end{equation}
 
+When the ISN coefficient is negative, the network is inhibition-stabilized.  
+To get an idea of what distribution of parameters ($d\mathbf{h}$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
+We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
+ \begin{equation}
+ f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.
+ \end{equation}
+ In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.05$ of $dx_\alpha = 0$.
+
 \begin{figure}[h]
-\caption{\small SX 1
+\caption{\small Supplemental Figure: A.
+Top row: EPI posteriors $q_\theta(z \mid \mathcal{X}(\alpha, c)$ for neuron-type $\alpha$, 0\% contrast, $\epsilon=0.01$ visualized as samples colored by their log probability.
+Bottom-left of pairplots indicate that the posterior predictive distribution obeys the emergent property constraints.
+Bottom row: Sets of parameters produced using the ABC approach.
+Bottom-left of pairplots shows statistics of accepted parameters (black) and the ABC posterior predictive distribution given that set (purple).
+B. Same as A at 100\% contrast.
  }\label{fig:SX1}
@@
 \begin{figure}[h]
-\caption{\small SX 2
+\caption{\small Supplemental Figure: Same as Figure 2 (will be supp fig) with $\sigma_{\epsilon} = 0.001$.
  }\label{fig:SX2}
@@
 
- With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
- Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
- All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
- While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
-\ref{fig:V1_EPI}B right).
- Currently available approaches to deriving the steady state response of the system are limited.
-
-%\section{Stability analysis}
-The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
-
-We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
-We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
-We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
-Maintaining the steady state of a neuron-type population amounts to the emergent property 
-\begin{equation}
-\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.\\
-\end{equation}
-In the following analyses, we chose $\sigma=0.25$.
-
-\section{EPI agrees with ABC}
-To get an idea of what distribution of parameters ($dh$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
-We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
- \[f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.\]
- In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.25$ of $dx_\alpha = 0$.
- In Figure 2, we see that the distributions obtained via EPI (colored by $\log q_\theta(z)$) are visually similar to those obtained via ABC (colored by $dx_\alpha$). \\
-
- \textbf{Figure 2}: EPI (left) vs ABC (right).
-Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$.
-The importance of $v_S$ and $v_P$ are explained in section 4.2.
-\\
-
-There are some subtle differences between the EPI and ABC distributions, but some difference is to be expected.
-Rather than simply attributing these differences to imperfection of the EPI optimization routine, we consider the the effect of bias in $dx_\alpha$ from ABC (and lack thereof from EPI).
-
-When $\epsilon > 0$ in ABC, the samples are \textit{not} from the posterior distribution.
-For $\epsilon > 0$, the ``posterior" predictive means of ABC $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ may be far from zero.
-In Figure 3, we see that with ABC, there is an increasingly negative bias in $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ for greater error tolerances across all neuron-types.  
-Additionally with ABC, there is no precise control of the variance $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha^2 \right]$, which may be undesireable.
-
-In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
-The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).
-
 \end{document}

### 2020-11-26 written/SC_model/SC_model.tex
SC model updates

@@
 Sean Bittner \\
-November 4, 2020 \\
+November 25, 2020 \\
 \end{center}
@@
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
+Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
 C. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
@@
 \ref{fig:SC}E).
-For example, the networks indicated by the black star suffers a steep drop in accuracy in the Anti task with delay period inactivation (Fig. \ref{fig:SC}F).  
-Networks resilient to delay period inactivation had greater $\lambda_{\text{task}}$ and $\lambda_{\text{task}}$ and lower $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ (Fig. \ref{fig:SC}E).
+For example, both networks suffer a steep drop in accuracy in the Anti task with delay period inactivation (Fig. \ref{fig:SC}F).  
+Networks resilient to delay period inactivation had greater $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$ and lower $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ (Fig. \ref{fig:SC}E).
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.6]{figs/figSX2.pdf}
-\end{center}
-\caption{\small
-Same as Fig. 1C shaded by Anti task accuracy.}
-\label{fig:SC_EPI2}
-\end{figure}
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.6]{figs/figSX2.pdf}
+%\end{center}
+%\caption{\small
+%Same as Fig. 1C shaded by Anti task accuracy.}
+%\label{fig:SC_EPI2}
+%\end{figure}
 

### 2020-11-30 written/eLifeRevision/Bittner_eLife2020.tex
elife revision

@@
 The operation of these circuits, like all models, critically depends on the choices of model parameters.
-Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
-However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-We bring recent machine learning techniques -- the use of deep generative models for probabilistic inference -- to bear on this problem, learning distributions of parameters that produce the specified properties of computation.
-Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
+When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
+%Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
+%However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
+We bring the use of deep generative models for probabilistic inference to bear on this problem, learning distributions of parameters that produce the specified properties of computation.
+By learning parameter distributions that produce computations -- an emergent property, we introduce a novel methodology for exploratory analyses and hypothesis testing that is particularly well-suited to the stochastic dynamical systems models predominant in our field of theoretical neuroscience.
+%Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
 We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
 We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
-More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
+While much use of deep learning in theoretical neuroscience focuses on drawing analogies between optimized neural architectures and the brain, this work illustrates how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
+%More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
 
@@
 The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about these particular parameter configurations?  
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
+%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
+The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model.
 
@@
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.  
-This creates an unfavorable tradeoff.
+%This creates an unfavorable tradeoff.
 On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
-On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-
-% now transition to ML...
-Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.  
-In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
-One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
+%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
+Alternatively, Bayesian inference can be run to obtain model parameters likely to produce some observations, and local sensitivity analyses can be performed at individual parameter choices of the obtained posterior.
+%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
+Since most neural circuit models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood central to the probabilistic modeling toolkit.
+Therefore, the most popular approaches to the inverse problem have been likelihood-free methods namely approximate Bayesian computation (ABC) \cite{liepe2014framework}, in which a set of model parameters analogous to a posterior is obtained via simulation and rejection.
+
+% now transition to deep ML...
+%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
+%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
+Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
+In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
+%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
+One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
 Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating parameters (sometimes treated as hidden states) in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
@@
 
-However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.  
-First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.  
-Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
-Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
-Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model. 
-Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
+% What's wrong with existing methodology:
+Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
+First, principled Bayesian approaches using MCMC or VAEs do not work absent a tractable likelihood function, which theoretical models of neural computation generally lack.
+The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except for the theoretical 0-distance scenario), lack parameter probabilities, and only confer sensitivity analyses of an alternative likelihood to the simulator-defined likelihood of ABC \cite{raue2009structural}.
+Second is the undesirable trade-off between the flexibility and tractability of the approximated posterior distribution.
+While sampling-based approaches like ABC and MCMC can produce flexible posterior approximations, they must be run continually for increasing samples.
+While VAE approaches result in tractable posterior sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions, which restrict the flexibility of the posterior approximation.
+And third, the posterior predictive distribution of all Bayesian approaches to the inverse problem are left unconstrained.  
+This is well understood when considering Box's loop and the role of posterior predictive checks in the development and critique of scientific models \cite{gelman2013philosophy, blei2014build}.
+However, unconstrained posterior predictive distributions introduce a conceptual degree of freedom to the inverse problem that may be unnecessary and undesirable given the scientific motivation.
+
+%However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.
+%First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.
+%Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
+%Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
+%Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model.
+%Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
 
@@
 This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
-First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
-Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+%First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
+%Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
+%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
+%Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
+%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
+%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
+First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
+Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
+Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
 Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
 Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
 Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
@@
 These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to infinite neurons).
-First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
+%First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
+First, we use EPI to elucidate the mechanisms of inhibition stabilization with varying contrast in a stochastic nonlinear dynamical model of primary visual cortex with inhibitory multiplicity.
+%Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
+Second, we discover connectivity patterns in superior colliculus resulting in resilience to optogenetic perturbation by using EPI to condition on rapid task switching.
 Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  

### 2020-12-07 written/eLifeRevision/Bittner_eLife2020.tex
introduction update minus reference updates

@@
 %On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-Alternatively, Bayesian inference can be run to obtain model parameters likely to produce some observations, and local sensitivity analyses can be performed at individual parameter choices of the obtained posterior.
+Alternatively, statistical inference can be run to obtain model parameters likely to produce some model output, and local sensitivity analyses can be performed at inferred parameter values.
 %Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
 Since most neural circuit models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood central to the probabilistic modeling toolkit.
-Therefore, the most popular approaches to the inverse problem have been likelihood-free methods namely approximate Bayesian computation (ABC) \cite{liepe2014framework}, in which a set of model parameters analogous to a posterior is obtained via simulation and rejection.
+Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{liepe2014framework}, in which a set of reasonable parameters estimates is obtained via simulation and rejection.
 
@@
 Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
-First, principled Bayesian approaches using MCMC or VAEs do not work absent a tractable likelihood function, which theoretical models of neural computation generally lack.
-The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except for the theoretical 0-distance scenario), lack parameter probabilities, and only confer sensitivity analyses of an alternative likelihood to the simulator-defined likelihood of ABC \cite{raue2009structural}.
+First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
+Thus, standard approaches from statistical inference are unavailable.
+The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except in the unrealistic $0$-distance scenario), lack parameter probabilities, and only confer sensitivity analyses of an alternative likelihood to the simulator-defined likelihood of ABC \cite{raue2009structural}.
 Second is the undesirable trade-off between the flexibility and tractability of the approximated posterior distribution.
-While sampling-based approaches like ABC and MCMC can produce flexible posterior approximations, they must be run continually for increasing samples.
-While VAE approaches result in tractable posterior sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions, which restrict the flexibility of the posterior approximation.
-And third, the posterior predictive distribution of all Bayesian approaches to the inverse problem are left unconstrained.  
+While sampling-based approaches like ABC and Markov chain Monte Carlo (MCMC) can produce flexible posterior approximations, they must be run continually for increasing samples.
+While VAE approaches can result in tractable posterior sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions, which restrict the flexibility of the posterior approximation.
+And third, you can never make assumptions of what inferred model parameters will predict.
 This is well understood when considering Box's loop and the role of posterior predictive checks in the development and critique of scientific models \cite{gelman2013philosophy, blei2014build}.
-However, unconstrained posterior predictive distributions introduce a conceptual degree of freedom to the inverse problem that may be unnecessary and undesirable given the scientific motivation.
+Uncertainty about the properties of inferred model predictions introduce a conceptual degree of freedom to the inverse problem that may be unnecessary and undesirable given the scientific motivation.
 

### 2020-12-07 epi/normalizing_flows.py
enable affine as 0 stages

@@ -245,7 +245,7 @@ class NormalizingFlow(tf.keras.Model):
     def _set_num_stages(self, num_stages):
         if type(num_stages) is not int:
             raise TypeError(format_type_err_msg(self, "num_stages", num_stages, int))
-        elif num_stages < 1:
+        elif num_stages < 0:
             raise ValueError(
                 "NormalizingFlow num_stages %d must be greater than 0." % num_stages
             )
@@ -292,8 +292,6 @@ class NormalizingFlow(tf.keras.Model):
             if type(bijector).__name__ == "BatchNormalization":
                 bijector.batchnorm.moving_mean.assign(np.zeros((self.D,)))
                 bijector.batchnorm.moving_variance.assign(np.ones((self.D,)))
-                print(bijector.batchnorm.moving_mean)
-                print(bijector.batchnorm.moving_variance)
         return None
 
     def _set_post_affine(self, post_affine):

### 2020-12-11 written/SC_model/SC_model.tex
SC model analysis update

@@
 {\Large Draft of new SC section} \\
-Sean Bittner \\
-November 25, 2020 \\
+Sean Bittner and Alex Piet \\
+December 11, 2020 \\
 \end{center}
@@
 In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}a).
+\ref{fig:SC}A).
 Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
@@
 \begin{split}
-\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + \epsilon d\mathbf{B} \\
+\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
 \mathbf{x} &= f(\mathbf{u})
@@
 \begin{center}
-\includegraphics[scale=0.5]{figs/fig4.pdf}
+\includegraphics[scale=0.48]{figs/fig4.pdf}
 \end{center}
-\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
+\caption{\small 
 A. Rapid task switching behavioral paradigm (see text). 
@@
 Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
-C. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
-D. Simulations of SC networks with parameterizations indicated by the stars in panel C.  Shading  shows standard deviations across trials.
-E. The effect of delay period inactivation on accuracy versus Schur eigenvalues.
-F. Same as D with delay period inactivation.}
+C. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice u=of simulations.
+D.
+Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.
+Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (opto strength 0.7).
+Shading indicates standard deviation across trials.
+E. The effect of delay period inactivation on accuracy versus dynamics eigenvalues.
+}
 \label{fig:SC}
@@
 
-We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
-To make sense of this inferred distribution, we followed the approach of Duan et al.
-by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-\ref{fig:SC_Eigs}A).
+We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C, Fig. \ref{fig:SX5}).
+Each parameter is predictive of accuracy in either task (Fig. \ref{fig:SX1}), and may have the same ($vW$ and $dW$) or opposite ($sW$ and $hW$) effects on $p_P$ and $p_A$.
+To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $v_i$ for all $W$ parameterized by $z$ (Fig.
+\ref{fig:SX2}A).
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.
 
-In agreement with experimental results form Duan et al., we found the optogenetic silencing during the delay period consistently decreased performance in the Anti task by 15-30\%, but had no consistent effect on the Pro task (Fig.
-\ref{fig:SC}E).
-For example, both networks suffer a steep drop in accuracy in the Anti task with delay period inactivation (Fig. \ref{fig:SC}F).  
-Networks resilient to delay period inactivation had greater $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$ and lower $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ (Fig. \ref{fig:SC}E).
+Smaller $\lambda_{\text{all}}$ and greater $\lambda_{\text{side}}$ predict greater accuracy in both tasks (Fig.
+\ref{fig:SX2}B) suggesting that wide separation between Left and Right populations at the end of stimulus presentation and fast decaying dynamics during the choice period yield better performance in both tasks.
+Since the Pro neurons receive a positive bias input $h_{P, bias}$, $\lambda_{\text{task}}$ is not as important to performance in the Pro task as in the Anti task (Fig. \ref{fig:SX2}B).
+The diagonal mode, hypothesized to be crucial for processing in the Anti task, was ultimately anticorrelated with $p_A$, 
+This suggests that $p_A$ increases with dynamics having stronger task and stimulus representations and faster decaying dynamics in alternate dimensions.
+
+In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
+\ref{fig:SX3}), e.g.
+(Fig.
+\ref{fig:SC}D, bottom).
+Interestingly, the trends in eigenvalues that result in greater $p_A$ increase the error from delay period inactivation in the Anti task ($p_A - p_{A, opto}$) (Fig. \ref{fig:SC}E.
+Despite these opposing effects, $p_A$ is uncorrelated with ($p_A - p_{A, opto}$), while $p_P$ is correlated (Fig. \ref{fig:SX4}).
+This shows that the degree of error introduced by delay period inactivation is controlled by the dynamical mechanisms controlling success in the Pro task (namely $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$).
+Such comparable dynamics between the non-inactivated Pro task and the delay period inactivated Anti task are salient in simulations (Fig. \ref{fig:SC}D top-left and bottom-right).
 
@@
 \begin{equation} 
-\mathbf{h}_{\text{P,bias}} = I_{\text{P,rule}} [1, 0, 1, 0]^\top
+\mathbf{h}_{\text{P,bias}} = I_{\text{P,bias}} [1, 0, 1, 0]^\top
 \end{equation}
@@
 \begin{center}
-\includegraphics[scale=0.9]{figs/figSX1.pdf}
+\includegraphics[scale=0.6]{figs/figSX5.pdf}
 \end{center}
-\caption{\small Sentence 1 .
-A.
-The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
-Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
-B. Task accuracy vs. Schur eigenvalue.}
-\label{fig:SC_Eigs}
+\caption{\small The EPI parameter distribution of Fig. 1C with reference to a unity line.
+}
+\label{fig:SX5}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.85]{figs/figSX1.pdf}
+\end{center}
+\caption{\small Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p value.
+}
+\label{fig:SX1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.3]{figs/figSX2.pdf}
+\end{center}
+\caption{\small 
+A. Invariant eigenvectors of connectivity matrix $W$.
+B. Eigenvalues of connectivities of EPI distribution versus task accuracies.}
+\label{fig:SX2}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.6]{figs/figSX4.pdf}
+\end{center}
+\caption{\small 
+Difference in performance of each task during inactivation.
+Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).
+We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
+}
+\label{fig:SX3}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.75]{figs/figSX3.pdf}
+\end{center}
+\caption{\small Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.7.
+}
+\label{fig:SX4}
 \end{figure}

### 2020-12-13 written/eLifeRevision/Bittner_eLife2020.tex
EPI as VI with mf gauss llhd

@@
 \begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.53 \\ 0.53 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
- \text{Var}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.53 \\ 0.53 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
+ \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
 \end{split}
@@
 q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) ~~=~~ &\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
-&\text{  s.t.  } ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
+&\text{  s.t.
+} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
 \end{split}
@@
 Columns correspond to inferred distributions of excitatory ($\alpha=E$, red), parvalbumin ($\alpha=P$, blue), somatostatin ($\alpha=S$, red) and VIP ($\alpha=V$, green) neuron-type response increases, while each row corresponds to increase amounts of $y \in \{0.1, 0.5\}$.
-For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_\theta(dh \mid \mathcal{B}(\alpha, y))$.
+For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_{\bm{\theta}}(dh \mid \mathcal{B}(\alpha, y))$.
 The inferred distributions immediately suggest four hypotheses: \\
@@
 \begin{equation}
-dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)).
+dh^* = z^* = \argmax_{z} \log q_{\bm{\theta}}(z \mid \mathcal{B}(\alpha, 0.1)).
 \end{equation}
@@
 \subsection{Emergent property inference (EPI)}\label{methods_EPI}
-Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
-Consider model parameterization $z$ and data $x$ which has an intractable likelihood $p(x \mid z)$ defined by a model simulator of which samples are available $x \sim p(x \mid z)$.  
-EPI optimizes a distribution $q_\theta(z)$ (itself parameterized by $\theta$) of model parameters $z$ to produce an emergent property of interest $\mathcal{B}$,
-\begin{equation}
-\mathcal{B} ~~\triangleq~~  \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
-\end{equation}
-Precisely, the emergent property statistics $T(x)$ must equal the emergent property values $\mu$, in expectation over the EPI distribution of parameters $q_\theta(z)$ and the distribution of simulated activity $p(x \mid z)$.
+%Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
+%Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
+%EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{B}$,
+%\begin{equation}
+%\mathcal{B} ~~\triangleq~~  \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right] \right] = \mu.
+%\end{equation}
+%Precisely, the emergent property statistics $T(\mathbf{x})$ must equal the emergent property values $\bm{\mu}$, in expectation over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and the distribution of simulated activity $p(\mathbf{x} \mid \mathbf{z})$.
+%This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
+Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
+EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{X}$ defined by the means and variances of emergent property statistics $f(\mathbf{x}; \mathbf{z})$
+ \begin{equation}
+\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\end{equation}
+Precisely, the emergent property statistics $f(\mathbf{x})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and stochasticity of the data given the parameters defined by the model $p(\mathbf{x} \mid \mathbf{z})$.
 This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
 
-With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_\theta(z)$.
- In deep probability distributions, a simple random variable $w \sim q_0(w)$ is mapped deterministically via a sequence of deep neural network layers ($f_1$, ..
-$f_l$) parameterized by weights and biases $\theta$ to the support of the distribution of interest:
-\begin{equation}
-z = f_{\theta}(\omega) = f_l(..f_1(w)).
-\end{equation}
-Given a simulator defined by a theoretical model $x \sim p(x \mid z)$ and some emergent property of interest $\mathcal{B}$, $q_\theta(z)$ is optimized via the neural network parameters $\theta$ to find a maximally entropic distribution $q_{\theta}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
+%With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
+% In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
+$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
+%\begin{equation}
+%\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
+%\end{equation}
+%Given a simulator defined by a theoretical model $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ and some emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
+%\begin{equation} \label{eq:opt}
+%\begin{split}
+%q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
+% &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
+% \end{split}
+%\end{equation} 
+%Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
+%Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0$:
+%\begin{equation}
+%H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{w \sim q_0}\left[-\log(q_{\bm{\theta}}(f_{\bm{\theta}}(w))) \right].
+%\end{equation}
+%Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $\mathbf{z}$:
+%\begin{equation}
+%\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{w \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(f_{\bm{\theta}}(w))) \right].
+%\end{equation}
+%In EPI, MEFNs are purposed towards variational learning of model parameter distributions.  
+%A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
+%Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(\mathbf{x} \mid \mathbf{z})$.
+%Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
+%Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
+%But, unlike LFVI, EPI uses a single deep network to learn a distribution and is optimized to produce an emergent property, rather than condition on data points.  
+%Optimizing the EPI objective is a technological challenge, the details of which we elaborate in Section \ref{methods_AL_opt}.  
+%Before going through those details, we ground this optimization in a toy example.
+
+With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
+ In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
+$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
+\begin{equation}
+\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
+\end{equation}
+Given a simulator defined by a theoretical model $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ and some emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
 \begin{equation} \label{eq:opt}
 \begin{split}
-q_\theta^*(z) &= \argmax_{q_\theta \in Q} H(q_\theta(z)) \\
- &  \text{s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu. \\
+q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
+ &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
  \end{split}
 \end{equation} 
-Since we are optimizing parameters $\theta$ of our deep probability distribution with respect to the entropy $H(q_\theta(z))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-Entropy of $q_\theta(z)$ can be expressed as an expectation of the negative log density of parameter samples $z$ over the randomness in the parameterless initial distribution $q_0$:
-\begin{equation}
-H(q_\theta(z)) = \int - q_\theta(z) \log(q_\theta(z)) dz = \mathbb{E}_{z \sim q_\theta}\left[-\log(q_\theta(z)) \right] = \mathbb{E}_{w \sim q_0}\left[-\log(q_\theta(f_\theta(w))) \right].
-\end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $z$:
-\begin{equation}
-\nabla_\theta H(q_\theta(z)) = \mathbb{E}_{w \sim q_0}\left[- \nabla_\theta \log(q_\theta(f_\theta(w))) \right].
+Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
+Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0$:
+\begin{equation}
+H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
+\end{equation}
+Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $\mathbf{z}$:
+\begin{equation}
+\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
 \end{equation}
 In EPI, MEFNs are purposed towards variational learning of model parameter distributions.
+
+% We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
+%We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
+%These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.
+
+ \subsubsection{Related work}\label{methods_related}
+
+TODO: rewrite this whole section.
+
 A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
-Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(x \mid z)$.
+Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(\mathbf{x} \mid \mathbf{z})$.
 Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
@@
 Before going through those details, we ground this optimization in a toy example.
-
  We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
@@
 
+\subsubsection{Normalizing flows}\label{methods_NF}
+%Deep probability models typically consist of several layers of fully connected neural networks.  
+Deep probability distributions are comprised of multiple layers of fully connected neural networks.  
+%When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
+For $\mathbf{z}' = g(\mathbf{z})$,
+When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
+For $\mathbf{z}_i = g_i(\mathbf{z_{i-1}})$,
+%\begin{equation}
+%q(\mathbf{z}') = q(f^{-1}(\mathbf{z}')) \left| \det \frac{\partial f^{-1}(\mathbf{z}')}{\partial \mathbf{z}'} \right| = q(\mathbf{z}) \left| \det \frac{\partial f(\mathbf{z})}{\partial \mathbf{z}} \right|^{-1}.
+%\end{equation}
+\begin{equation}
+p(\mathbf{z}_i) = p(g_i^{-1}(\mathbf{z}_i)) \left| \det \frac{\partial g_i^{-1}(\mathbf{z}_i)}{\partial \mathbf{z}_i} \right| = p(\mathbf{z}_{i-1}) \left| \det \frac{\partial g_i(\mathbf{z}_{i-1})}{\partial \mathbf{z}_{i-1}} \right|^{-1}.
+\end{equation}
+
+However, this computation has cubic complexity in dimensionality for fully connected layers.  
+By restricting our layers to normalizing flows \cite{rezende2015variational} -- bijective functions with fast log determinant Jacobian computations, we can tractably optimize deep generative models with objectives that are a function of sample density, like entropy.
+%Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
+%Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
+TODO: (clean up) We use Real NVP because it's a coupling architecture, which is fast to run either forwards (probability with samples) and backwards (prrobability or hessian).
+Normalizing flow architectures for deep probability distributions used in EPI are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
+
+\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
+%To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
+%\begin{equation} \label{eq:AL}
+%L(\bm{\theta}; \eta, c) = -H(q_{\bm{\theta}}) + \eta^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
+%\end{equation}
+%where $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}) - \bm{\mu} \right] \right]$, $\eta \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}| = |T(\mathbf{x})|$,  and $c$ is the penalty coefficient.
+%These Lagrange multipliers are closely related to the natural parameters of exponential families (see Section \ref{methods_VI}).
+%Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
+%$\eta$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\eta$, $c$) for a given number of stochastic optimization iterations.
+%A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
+%For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
+%The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
+%The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
+%The other update rule is $\eta_{k+1} = \eta_k + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu})$ given a batch size $n$.
+%Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
+To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
+\begin{equation} \label{eq:AL}
+L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
+\end{equation}
+where $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
+These Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
+Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
+$\bm{\eta}_{\text{opt}}$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
+A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
+For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
+The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
+The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
+The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu})$ given a batch size $n$.
+Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
+
+The intention is that $c$ and $\bm{\eta}_{\text{opt}}$ start at values encouraging entropic growth early in optimization.  
+With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
+This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
+In the oscillating 2D LDS example, each augmented Lagrangian epoch ran for 2,000 iterations (Fig. S1C-D).
+Notice the initial entropic growth, and subsequent reduction upon each update of $\bm{\eta}_{\text{opt}}$ and $c$.
+The momentum parameters of the Adam optimizer were reset at the end of each augmented Lagrangian epoch.
+
+Rather than starting optimization from some $\bm{\theta}$ drawn from a randomized distribution, we found that initializing $q_{\bm{\theta}}(\mathbf{z})$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
+The parameters of the Gaussain initialization were chosen on an application-specific basis.  
+Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some standard deviation $\bm{\sigma}_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_V1}).
+
+To assess whether EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
+The algorithm has converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
+This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
+The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made from $k$ resamplings of $\mathbf{z}$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
+$k$ is determined by a fraction of the batch size $\nu$, which varies according to the application.
+In the linear two-dimensional system example, we used a batch size of $n = 1000$ and set $\nu = 0.1$ resulting in convergence after the ninth epoch of optimization. (Fig. S1C-D black dotted line).
+
+When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
+First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
+Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
+If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
+Next, one should consider the computational cost of the gradient calculations. 
+In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_2DLDS}) for the emergent property statistic given the model parameters.  
+On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
+Section \ref{methods_STG}).
+In such cases, optimization will be expensive.
+
+\subsubsection{Maximum entropy distributions and exponential families}\label{methods_VI}
+%As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
+%A maximum entropy distribution of form:
+%\begin{equation} \label{eq:max_ent}
+%\begin{split}
+%p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
+% &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}. \\
+% \end{split}
+%\end{equation} 
+%will have probability density in the exponential family:
+%\begin{equation}
+%p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
+%\end{equation}
+%The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
+Maximum entropy distributions have a fundamental link to exponential family distributions. 
+A maximum entropy distribution of form:
+\begin{equation} \label{eq:max_ent}
+\begin{split}
+p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
+ &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
+ \end{split}
+\end{equation} 
+will have probability density in the exponential family:
+\begin{equation}
+p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
+\end{equation}
+The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
+
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP}
+ \begin{equation}
+\mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\end{equation}
+The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
+ \begin{equation}
+\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
+\end{equation}
+As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
+\begin{equation} \label{eq:moments}
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
+\end{equation}
+which are constrained to the chosen means and variances
+\begin{equation}
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
+\end{equation}
+
+\subsubsection{EPI as variational inference}\label{methods_exp_fam_lh}
+%Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
+In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
+In Bayesian inference, we obtain a posterior distribution $p(z \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+\begin{equation}
+p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
+\end{equation}
+The posterior distribution is analytically available when the prior is conjugate with the likelihood.
+However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
+
+%Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid \mathbf{x})$.  
+%We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(\mathbf{x})$.
+%The log normalizing constant $A(\eta(\mathbf{x}))$ remains unchanged:
+%\begin{equation}
+%\begin{split}
+%p(z \mid \mathbf{x}) = b(z) \exp{\left( \eta(\mathbf{x})^\top T(z) - A(\eta(\mathbf{x})) \right)} = \exp{\left( \begin{bmatrix} \eta(\mathbf{x}) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(\mathbf{x})) \right)} \\= \exp{\left(\tilde{\eta(\mathbf{x})}^\top \tilde{T}(z) - A(\eta(\mathbf{x})) \right)}
+%\end{split}.
+%\end{equation}
+%Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
+%\begin{equation}
+%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})).
+%\end{equation}
+%$q_{\bm{\theta}}(z)$ is the variational approximation to the posterior with variational parameters $\bm{\theta}$.  
+%We can write this KL divergence in terms of entropy of the variational approximation:
+%\begin{equation}
+%KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) = \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(z)) \right] - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (p(z \mid \mathbf{x})) \right]
+%\end{equation}
+%\begin{equation}
+% = -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) - A(\eta(\mathbf{x})) \right].
+%\end{equation}
+%As far as the variational optimization is concerned, the log normalizing constant is independent of $q_{\bm{\theta}}(z)$, so it can be dropped
+%\begin{equation}
+%\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) =  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) \right].
+%\end{equation}
+% Further, we can write the objective in terms of the first moment of the sufficient statistics $\bm{\mu} = \mathbb{E}_{z \sim p(z \mid \mathbf{x})}\left[T(z) \right]$:
+% \begin{equation}
+%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right] + \tilde{\eta}(\mathbf{x})^\top \bm{\mu},
+% \end{equation}
+%which simplifies to
+%  \begin{equation}
+%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right].
+% \end{equation}.
+%
+%In comparison, in emergent property inference (EPI), we solve the following problem:
+%\begin{equation}
+%q_{\bm{\theta}}^*(z) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(z)),   \text{  s.t.
+} \mathbb{E}_{z \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid z)}\left[T(\mathbf{x})\right] \right] = \bm{\mu}.
+%\end{equation}
+%The Lagrangian objective (without augmentation) is
+%\begin{equation}
+%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_{\bm{\theta}}} \left[\tilde{T}(z) \right] - \bm{\mu} \right).
+%\end{equation}
+%Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(\mathbf{x})$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
+%
+%We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
+%Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
+%However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
+%Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
+%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
+
+In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$
+\begin{equation}
+q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
+\end{equation}
+The KL divergence can be written in terms of entropy of the variational approximation:
+\begin{equation}
+KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
+\end{equation}
+\begin{equation}
+= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
+\end{equation}
+Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
+This is usually framed as maximizing the evidence lower bound (ELBO)
+\begin{equation}
+\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
+\end{equation}
+
+Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
+\begin{equation}
+p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
+\end{equation}
+where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
+The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
+\begin{equation}
+\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
+\end{equation}
+where
+\begin{equation}
+\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
+\end{equation}
+\begin{equation}
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
+\end{equation}
+The variational objective is then
+\begin{equation}
+\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
+\end{equation}
+
+Comparing this to the Lagrangian objective (without augmentation) of EPI, we see they are the same
+\begin{equation}
+\begin{split}
+q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
+ &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
+\end{split}
+\end{equation}
+where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
+Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
+The data $\mathbf{x}$ used by this implicit variational inference program would be that generated by the adapting variational approximation $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z}) q_{\bm{\theta}}(\mathbf{z})$, and the likelihood parameters $\bm{\eta}_{f}$ of EPI optimization epoch $k$ are predicated by $\bm{\eta}_{\text{opt},k}$ .
+However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
+Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
+Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
+
 \subsubsection{Example: 2D LDS}\label{methods_2DLDS}
@@
 \begin{equation} 
-\tau \frac{dx}{dt} = Ax
+\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
 \end{equation}
@@
 \end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $z = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(x)$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
+To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
  $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
@@
 \begin{equation}
- \mathbb{E}\left[T(x) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \mu.
+ \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
  \end{equation} 
@@
 
-Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $z$, since the eigenvalues can be calculated using the quadratic formula:
+Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $\mathbf{z}$, since the eigenvalues can be calculated using the quadratic formula:
 \begin{equation}
@@
 
-Importantly, even though $\mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_\theta$ directly.
-This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\mu$ of an exponential family distribution \cite{wainwright2008graphical}.
+Importantly, even though $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}$ directly.
+This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
 Instead, we used EPI to approximate this distribution (Fig.
@@
 \end{equation}
-Since we constrained $\mathbb{E}_{z \sim q_\theta}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
+Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
 S2B).
-With $\sigma_{1,4} = \mathbb{E}_{z \sim q_\theta}(|a_1 a_4 - E_{q_\theta}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
+With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
 This validates the curved structure of the inferred distribution learned through EPI.  
@@
 {\small Fig.
-S3: Sampled dynamical systems $z \sim q_\theta(z)$ and their simulated activity from $x(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+S3: Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
 A.
@@
 \end{figure}
-
-\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-To optimize $q_\theta(z)$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
-\begin{equation} \label{eq:AL}
-L(\theta; \eta, c) = -H(q_\theta) + \eta^\top R(\theta) + \frac{c}{2}||R(\theta)||^2
-\end{equation}
-where $R(\theta) = \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x) - \mu \right] \right]$, $\eta \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\mu| = |T(x)|$,  and $c$ is the penalty coefficient.
-These Lagrange multipliers are closely related to the natural parameters of exponential families (see Section \ref{methods_VI}).
-Deep neural network weights and biases $\theta$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
-$\eta$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\eta$, $c$) for a given number of stochastic optimization iterations.
-A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
-For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
-The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
-The p-value of $\mathbb{E}[||R(\theta_{k+1})||] > \gamma \mathbb{E} \left[||R(\theta_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-The other update rule is $\eta_{k+1} = \eta_k + c_k \frac{1}{n} \sum_{i=1}^n (T(x^{(i)}) - \mu)$ given a batch size $n$.
-Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
-
-The intention is that $c$ and $\eta$ start at values encouraging entropic growth early in optimization.  
-With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
-In the oscillating 2D LDS example, each augmented Lagrangian epoch ran for 2,000 iterations (Fig. S1C-D).
-Notice the initial entropic growth, and subsequent reduction upon each update of $\eta$ and $c$.
-The momentum parameters of the Adam optimizer were reset at the end of each augmented Lagrangian epoch.
-
-Rather than starting optimization from some $\theta$ drawn from a randomized distribution, we found that initializing $q_\theta(z)$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
-The parameters of the Gaussain initialization were chosen on an application-specific basis.  
-Throughout the study, we chose isotropic Gaussian initializations with mean $\mu_{\text{init}}$ at the center of the distribution support and some standard deviation $\sigma_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_V1}).
-
-To assess whether EPI distribution $q_\theta(z)$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
-The algorithm has converged when a null hypothesis test of constraint violations $R(\theta)_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
-This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
-The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\theta)^i$ are made from $k$ resamplings of $z$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
-$k$ is determined by a fraction of the batch size $\nu$, which varies according to the application.
-In the linear two-dimensional system example, we used a batch size of $n = 1000$ and set $\nu = 0.1$ resulting in convergence after the ninth epoch of optimization. (Fig. S1C-D black dotted line).
-
-When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
-First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
-Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
-If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
-Next, one should consider the computational cost of the gradient calculations. 
-In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_2DLDS}) for the emergent property statistic given the model parameters.  
-On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
-Section \ref{methods_STG}).
-In such cases, optimization will be expensive.
-
-\subsubsection{Normalizing flows}\label{methods_NF}
-Deep probability models typically consist of several layers of fully connected neural networks.  
-When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.  For $z' = f(z)$,
-
-\begin{equation}
-q(z') = q(f^{-1}(z')) \left| \det \frac{\partial f^{-1}(z')}{\partial z'} \right| = q(z) \left| \det \frac{\partial f(z)}{\partial z} \right|^{-1}.
-\end{equation}
-
-However, this computation has cubic complexity in dimensionality for fully connected layers.  
-By restricting our layers to normalizing flows \cite{rezende2015variational} -- bijective functions with fast log determinant Jacobian computations, we can tractably optimize deep generative models with objectives that are a function of sample density, like entropy.
-Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
-Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
-
-\subsubsection{Emergent property inference as variational inference in an exponential family}\label{methods_VI}
-Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In Bayesian inference a prior belief about model parameters $z$ is formalized into a prior distribution $p(z)$, and the statistical model capturing the effect of $z$ on observed data points $x$ is formalized in the likelihood distribution $p(x \mid z)$.
-In Bayesian inference, we obtain a posterior distribution $p(z \mid x)$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
-\begin{equation}
-p(z \mid x) = \frac{p(x \mid z) p(z)}{p(x)}.
-\end{equation}
-The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
-
-As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
-A maximum entropy distribution of form:
-\begin{equation}
-\begin{split}
-p^*(z) &= \argmax_{p \in \mathcal{P}} H(p(z)) \\
- &  \text{s.t.  } \mathbb{E}_{z \sim p}\left[T(z)\right] = \mu. \\
- \end{split}
-\end{equation} 
-will have probability density in the exponential family:
-\begin{equation}
-p^*(z) \propto \exp(\eta^\top T(z)).
-\end{equation}
-The mappings between the mean parameterization $\mu$ and the natural parameterization $\eta$ are formally hard to identify \cite{wainwright2008graphical}.
-
-Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid x)$.  
-We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(x)$.
-The log normalizing constant $A(\eta(x))$ remains unchanged:
-\begin{equation}
-\begin{split}
-p(z \mid x) = b(z) \exp{\left( \eta(x)^\top T(z) - A(\eta(x)) \right)} = \exp{\left( \begin{bmatrix} \eta(x) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(x)) \right)} \\= \exp{\left(\tilde{\eta(x)}^\top \tilde{T}(z) - A(\eta(x)) \right)}
-\end{split}.
-\end{equation}
-Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
-\begin{equation}
-q_\theta^* = \argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)).
-\end{equation}
-$q_\theta(z)$ is the variational approximation to the posterior with variational parameters $\theta$.  
-We can write this KL divergence in terms of entropy of the variational approximation:
-\begin{equation}
-KL(q_\theta \mid \mid p(z \mid x)) = \mathbb{E}_{z \sim q_\theta} \left[ \log (q_\theta(z)) \right] - \mathbb{E}_{z \sim q_\theta} \left[ \log (p(z \mid x)) \right]
-\end{equation}
-\begin{equation}
- = -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) - A(\eta(x)) \right].
-\end{equation}
-As far as the variational optimization is concerned, the log normalizing constant is independent of $q_\theta(z)$, so it can be dropped
-\begin{equation}
-\argmin_{q_\theta \in Q} KL(q_\theta \mid \mid p(z \mid x)) =  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top  \tilde{T}(z) \right].
-\end{equation}
- Further, we can write the objective in terms of the first moment of the sufficient statistics $\mu = \mathbb{E}_{z \sim p(z \mid x)}\left[T(z) \right]$:
- \begin{equation}
-=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right] + \tilde{\eta}(x)^\top \mu,
- \end{equation}
-which simplifies to
-  \begin{equation}
-=  \argmin_{q_\theta \in Q} -H(q_\theta) - \mathbb{E}_{z \sim q_\theta} \left[ \tilde{\eta}(x)^\top \left(  \tilde{T}(z) -\mu \right) \right].
- \end{equation}.
-
-In comparison, in emergent property inference (EPI), we solve the following problem:
-\begin{equation}
-q_\theta^*(z) = \argmax_{q_\theta \in Q} H(q_\theta(z)),   \text{  s.t.  } \mathbb{E}_{z \sim q_\theta}\left[ \mathbb{E}_{x\sim p(x \mid z)}\left[T(x)\right] \right] = \mu.
-\end{equation}
-The Lagrangian objective (without augmentation) is
-\begin{equation}
-q_\theta^* = \argmin_{q_\theta \in Q} - H(q_\theta) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_\theta} \left[\tilde{T}(z) \right] - \mu \right).
-\end{equation}
-Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(x)$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
-
-We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
-Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
-However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-Accordingly, we replace the notation of $p(z \mid x)$ with $p(z \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
 
@@
 B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
-Since $q_\theta(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.    
+Since $q_{\bm{\theta}}(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
 C. The second moment emergent property statistics converge to the emergent property values.}
@@
 For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $z = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_\theta(z)$ to produce an isotropic 
+We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic 
 Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
@@
 We calculated the Hessian at the mode of the inferred EPI distribution.  
-The Hessian of a probability model is the second order gradient of the log probability density $\log q_\theta(z)$ with respect to the parameters $z$:  $\frac{\partial^2 \log q_\theta(z)}{\partial z \partial z^\top}$.
+The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(z)$ with respect to the parameters $z$:  $\frac{\partial^2 \log q_{\bm{\theta}}(z)}{\partial z \partial z^\top}$.
 With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
@@
 
-In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = z^* = \argmax_{z} \log q_\theta(z \mid \mathcal{B}(\alpha, 0.1)$.
+In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = z^* = \argmax_{z} \log q_{\bm{\theta}}(z \mid \mathcal{B}(\alpha, 0.1)$.
 The differential responses $\delta x_{\alpha,ss}$ are examined at perturbed inputs  $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$ and $\delta x$ is evaluated at 101 equally spaced samples of $\delta h_\alpha$ from -15 to 15.
@@
 As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
-For the first emergent property statistic, the average over EPI samples (from $q_\theta(z)$) is set to the desired value $p_{LP}$:
+For the first emergent property statistic, the average over EPI samples (from $q_{\bm{\theta}}(z)$) is set to the desired value $p_{LP}$:
 \begin{equation}
@@
 \begin{tabular}{l|l|l|l|l}
-$\lambda$  & $\hat{p}$ & $q_\theta(z)$ & $r$ & p-value \\ \hline
+$\lambda$  & $\hat{p}$ & $q_{\bm{\theta}}(z)$ & $r$ & p-value \\ \hline
 $\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
@@
 For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
-We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
+We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
 The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.
@@
 We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  
-Correlations were measured from 5,000 samples of $q_\theta(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
+Correlations were measured from 5,000 samples of $q_{\bm{\theta}}(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
 
@@
 We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization.
-We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_\theta(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
+We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
 The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.

### 2020-12-14 written/eLifeRevision/Bittner_eLife2020.tex
STG section update

@@
 %Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
-Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference (Fig. \ref{fig:STG}A).  
+Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
 First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
@@
 %A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
-A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}C, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
+A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
 Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  
@@
 %Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
-Second, once the model is selected, one defines the emergent property, the measurable behavior of scientific interest.  
+Second, once the model is selected, one defines the emergent phenomena of scientific interest.
 %To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-To continue our running STG example, one such emergent property is the phenomenon of \emph{unified intermediacy} -- in certain parameter regimes, the frequency of all neurons match at an intermediate firing rate.
-This emergent property is shown in Figure \ref{fig:STG}D at a frequency of 0.53Hz.
+In the STG example, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model \ref{fig:STG}B.
+%One such emergent property is the phenomenon of \emph{unified intermediacy} -- in certain parameter regimes, the frequency of all neurons match at an intermediate firing rate.
+An interesting emergent property of this stochastic model is when the hub neuron fires at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
+%This emergent property is shown in Figure \ref{fig:STG}D at a frequency of 0.53Hz.
+This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
 
 %Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
-Third, qualitative parameter analysis ensues: since mathematical analysis of frequency is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.  
-Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
+%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
+%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
+% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
+Third, parameter analyses ensue: brute-force parameter sweeps, ABC sampling, and sensitivity analyses are all routinely used to reason about what parameter configurations lead to an emergent property.
 In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
-\end{center}
-\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
 %A.
@@
 EPI optimization: To learn the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ of model parameters that produce an emergent property, the emergent property statistics $T(\mathbf{x})$ are set in expectation over model parameter samples $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and model simulations $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ to emergent property values $\bm{\mu}$.
-A.
-For a choice of model (STG) and emergent property (unified intermediacy), emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing unified intermediacy.
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
+\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
+\textbf{A}. Conductance-based biophysical model of the STG subcircuit.
 In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
 Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-Deep probability distributions map a simple random variable $\mathbf{w}$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = f_{\bm{\theta}}(\mathbf{w})$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
-B. The EPI distribution of STG model parameters producing network syncing.  
+Parameters are represented by the vector $\mathbf{z}$ and data by the vector $\mathbf{x}$.
+\textbf{B}. Simulated activity form the STG model at $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
+\textbf{C}. The emergent property of unified intermediacy, in which all neurons are firing close to the same intermediate frequency.
+Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution. 
+\textbf{D}.
+For a choice of model and emergent property, emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing middle hub frequency.
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+\textbf{E}. The EPI distribution of STG model parameters producing network syncing.  
 Samples are colored by log probability density.  
@@
 Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-Simulated activity is shown for three samples (stars)in panel D. 
+Simulated activity is shown for three samples (stars). 
+%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
 $v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}). 
-D. The emergent property of unified intermediacy, in which all neurons are firing close to the same intermediate frequency.
-Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution. }
+\textbf{F} Simulations from parameters in E.  
+(Top) The predictive distribution of the posterior obeys the constraints stipulated by the emergent property.
+(Bottom) Simulations at the starred parameter values.
+ }
+\end{center}
 \label{fig:STG}
@@
 Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}A) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution.
+\ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
+\ref{fig:STG}E).
 This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
-%In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}B yellow star), and how network syncing decays based on changes away from the mode.
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for unified intermediacy (the mode; Fig.
-\ref{fig:STG}B, yellow star), and how it decays based on changes away from the mode.
+In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}E yellow star), and how network syncing decays based on changes away from the mode.
 The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of unified intermediacy (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
@@
 \ref{fig:STG}B contours) and have diminished or preserved network syncing, respectively (Fig.
-\ref{fig:STG}D activity traces, Fig.
+\ref{fig:STG}F activity traces, Fig.
 S TODO) (see Section \ref{methods_STG}).
@@
 Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
-First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
+%First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
+First, we consider the model as a coupled set of stochastic differential equations \cite{gutierrez2013multiple}.  
 In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
-\begin{equation} C_m \frac{d\mathbf{x}}{dt} = -\mathbf{h}(\mathbf{x}; \mathbf{z}) = - \left[ \mathbf{h}_{leak}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{Ca}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_K(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{hyp}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{elec}(\mathbf{x}; \mathbf{z}) + \mathbf{h}_{syn}(\mathbf{x}; \mathbf{z})\right]
+\begin{equation} 
+\begin{split}
+C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) = - [ &\mathbf{h}_{leak}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{Ca}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_K(\mathbf{x}(t); \mathbf{z}) \\
+ & + \mathbf{h}_{hyp}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{elec}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{syn}(\mathbf{x}(t); \mathbf{z})] 
+ \end{split}
 \end{equation} 
@@
 
-Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
+%Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
+Second, we define the emergent property, which as above is middle ``hub neuron frequency": oscillation of the hub neuron at an intermediate frequency (Figure \ref{fig:STG}C).
+%Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
 -- is close to an intermediate frequency of 0.53Hz.
-Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
+%Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
+Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$is close to an intermediate frequency of 0.55Hz.  
+Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
 
@@
 \begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.53 \\ 0.53 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
- \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{f1}}(\mathbf{x}; \mathbf{z}) \\ \omega_{\text{f2}}(\mathbf{x}; \mathbf{z}) \\ \vdots \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
+ \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
 \end{split}
@@
 %which completes the quantification of the emergent property.
-The emergent property statistics $T(\mathbf{x}; \mathbf{z})$ along with their constrained means $\bm{\mu}$ and variances $\bm{\sigma}^2$ define the emergent property denoted $\mathcal{X}$.
+The emergent property statistic $f(\mathbf{x}; \mathbf{z}) = \omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ along with its constrained mean $\bm{\mu}$ and variance $\bm{\sigma}^2$ define the emergent property denoted $\mathcal{X}$.
 
@@
 This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}B (see Section \ref{methods_EPI}).  
+This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
 Then, mathematically, we must solve the following optimization program: 
@@
 &\text{  s.t.
-} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
+} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
 \end{split}
@@
 %where $T(\mathbf{x}), \bm{\mu}$ are defined as in Equation \ref{eq:EP}, and $p(\mathbf{x} | \mathbf{z})$ is the intractable distribution of data from the model, $\mathbf{x}$, given that model's parameters $\mathbf{z}$ (we access samples from this distribution by running the model forward).
-where $T(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
+where $f(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
 %The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
@@
 %Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
-Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $T(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
+%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
+Specifically, EPI (TODO insert interpretation).
 %Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.

### 2020-12-17 epi/models.py
nf saving/loading with splines

@@ -1048,6 +1048,7 @@ class Model(object):
             "num_stages": nf.num_stages,
             "num_layers": nf.num_layers,
             "num_units": nf.num_units,
+            "elemwise_fn": nf.elemwise_fn,
             "batch_norm": nf.batch_norm,
             "bn_momentum": nf.bn_momentum,
             "post_affine": nf.post_affine,
@@ -1126,6 +1127,7 @@ class Model(object):
             num_stages=arch["num_stages"],
             num_layers=arch["num_layers"],
             num_units=arch["num_units"],
+            elemwise_fn="spline",
             batch_norm=arch["batch_norm"],
             bn_momentum=arch["bn_momentum"],
             post_affine=arch["post_affine"],
@@ -1365,7 +1367,7 @@ class Distribution(object):
         del z  # Get rid of dummy variable.
         return hess_z.numpy()
 
-    @tf.function
+    #@tf.function
     def _hessian(self, z):
         with tf.GradientTape(persistent=True) as tape:
             log_q_z = self.nf.trans_dist.log_prob(z)

### 2020-12-17 scripts/SC_Circuit/SC_acc_diff.py
sc scripts

@@ -23,12 +23,11 @@ c0 = 10.**args.logc0
 bnmom = args.bnmom
 random_seed = args.random_seed
 
-M = 200
-N = 100
+M = 100
 
 # 1. Specify the V1 model for EPI.
-lb = -20.
-ub = 20.
+lb = -1.
+ub = 1.
 
 sW_P = Parameter("sW_P", 1, lb=lb, ub=ub)
 sW_A = Parameter("sW_A", 1, lb=lb, ub=ub)

### 2020-12-17 scripts/SC_Circuit/SC_acc_epi_var.py

@@ -60,8 +60,8 @@ q_theta, opt_data, epi_path, failed = model.epi(
     num_units=50,
     post_affine=True,
     batch_norm=True,
-    bn_momentum=0.99,
-    K=4,
+    bn_momentum=0.0,
+    K=15,
     N=M,
     num_iters=2000,
     lr=1e-3,
@@ -73,7 +73,7 @@ q_theta, opt_data, epi_path, failed = model.epi(
     #init_params=init_params,
     verbose=True,
     stop_early=True,
-    log_rate=50,
+    log_rate=100,
     save_movie_data=True,
 )
 

### 2020-12-17 scripts/SC_Circuit/cpu_SC_acc_epi_var.sh

@@ -6,4 +6,4 @@
 #SBATCH --mem-per-cpu=50gb
 
 source activate epi
-python3 SC_acc_epi.py --p $1 --beta $2 --logc0 $3 --bnmom $4 --random_seed $5
+python3 SC_acc_epi_var.py --p $1 --beta $2 --logc0 $3 --mu_std $4 --random_seed $5

### 2020-12-17 scripts/SC_Circuit/sc_sweep_var.sh

@@ -2,15 +2,18 @@
 
 # This shell script does it linearly, but could you run each of these 
 # python scripts independently on different instances?
-for beta in 2. 4.
+for p in 0.75
 do
-  for logc0 in 3
+  for beta in 4.
   do
-    for bnmom in 0.99
+    for logc0 in 2
     do
-      for rs in 1 2
+      for mu_std in 0.05
       do
-        sbatch cpu_SC_acc_diff.sh $beta $logc0 $bnmom $rs
+        for rs in 1 2 3 4
+        do
+          sbatch cpu_SC_acc_epi_var.sh $p $beta $logc0 $mu_std $rs
+        done
       done
     done
   done

### 2020-12-17 scripts/SC_Circuit/SC_acc_epi_var.py
new SC scripts

@@ -12,6 +12,7 @@ DTYPE = tf.float32
 parser = argparse.ArgumentParser()
 parser.add_argument('--p', type=float, default=0.5) # neuron type
 parser.add_argument('--beta', type=float, default=4.) # aug lag hp
+parser.add_argument('--elemwise_fn', type=str, default="affine") # aug lag hp
 parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
 parser.add_argument('--mu_std', type=float, default=0.1) 
 parser.add_argument('--random_seed', type=int, default=1)
@@ -19,6 +20,7 @@ args = parser.parse_args()
 
 p = args.p
 AL_beta = args.beta
+elemwise_fn = args.elemwise_fn
 c0 = 10.**args.logc0
 mu_std = args.mu_std
 random_seed = args.random_seed
@@ -58,8 +60,9 @@ q_theta, opt_data, epi_path, failed = model.epi(
     num_stages=3,
     num_layers=2,
     num_units=50,
-    post_affine=True,
-    batch_norm=True,
+    elemwise_fn=elemwise_fn,
+    post_affine=False,
+    batch_norm=False,
     bn_momentum=0.0,
     K=15,
     N=M,

### 2020-12-17 scripts/SC_Circuit/cpu_SC_acc_epi_var.sh

@@ -6,4 +6,4 @@
 #SBATCH --mem-per-cpu=50gb
 
 source activate epi
-python3 SC_acc_epi_var.py --p $1 --beta $2 --logc0 $3 --mu_std $4 --random_seed $5
+python3 SC_acc_epi_var.py --p $1 --beta $2 --elemwise_fn $3 --logc0 $4 --mu_std $5 --random_seed $6

### 2020-12-17 scripts/SC_Circuit/sc_sweep_var.sh

@@ -6,13 +6,16 @@ for p in 0.75
 do
   for beta in 4.
   do
-    for logc0 in 2
+    for elemwise_fn in affine spline
     do
-      for mu_std in 0.05
+      for logc0 in 2
       do
-        for rs in 1 2 3 4
+        for mu_std in 0.05
         do
-          sbatch cpu_SC_acc_epi_var.sh $p $beta $logc0 $mu_std $rs
+          for rs in 1
+          do
+            sbatch cpu_SC_acc_epi_var.sh $p $beta $elemwise_fn $logc0 $mu_std $rs
+          done
         done
       done
     done

### 2020-12-17 epi/models.py
convenient func updates for models.py

@@ -584,6 +584,40 @@ class Model(object):
                             dfs.append(df)
         return pd.concat(dfs)
 
+    def get_max_H(self, mu, nu, paths=None):
+        epi_df = self.get_epi_df()
+        if paths is None:
+            paths = epi_df['path'].unique()
+        
+        best_Hs = []
+        convergeds = []
+        best_ks = []
+        for i, path in enumerate(paths):
+            print('checking convergence', i)
+            epi_df2 = epi_df[epi_df['path'] == path]
+            df_row = epi_df2.iloc[0]
+            init = df_row['init']
+            init_params = {"mu":init["mu"], "Sigma":init["Sigma"]}
+            nf = self._df_row_to_nf(df_row)
+            aug_lag_hps = self._df_row_to_al_hps(df_row)
+            best_k, converged, best_H = self.get_convergence_epoch(
+                    init_params, nf, mu, aug_lag_hps, alpha=0.05, nu=nu)
+            best_Hs.append(best_H)
+            convergeds.append(converged)
+            best_ks.append(best_k)
+
+        bestHs = np.array(best_Hs)
+        best_ks = np.array(best_ks)
+
+        best_Hs = np.array([x if x is not None else np.nan for x in best_Hs])
+        ind = np.nanargmax(best_Hs)
+
+        best_k = int(best_ks[ind])
+        path = paths[ind]
+        best_H = best_Hs[ind]
+
+        return path, best_k
+
     def epi_opt_movie(self, path):
         """Generate video of EPI optimization.
 
@@ -1104,11 +1138,12 @@ class Model(object):
         # if final index is set, this EPI opt has been run before.
         return epi_path, exists
 
-    def get_epi_dist(self, df_row):
+    def get_epi_dist(self, df_row, k=None):
         init = df_row["init"]
         ep = df_row["EP"]
 
-        k = int(df_row["k"])
+        if k is None:
+            k = int(df_row["k"])
         init_params = {"mu":init["mu"], "Sigma":init["Sigma"]}
         nf = self._df_row_to_nf(df_row)
         mu = ep["mu"]
@@ -1174,8 +1209,11 @@ class Model(object):
         return q_theta
 
     def get_convergence_epoch(
-        self, init_params, nf, mu, aug_lag_hps, alpha=0.05, nu=0.1, mu_test=None
[…]

### 2020-12-17 written/SC_model/SC_model.tex

@@
 Sean Bittner and Alex Piet \\
-December 11, 2020 \\
+December 14, 2020 \\
 \end{center}
@@
 The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
+\textbf{The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.}
 The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
@@
 Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
-We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.
+\textbf{We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.}
 \begin{equation}\label{eq:EP}
@@
 
-In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
+\textbf{In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
 \ref{fig:SX3}), e.g.
@@
 \ref{fig:SC}D, bottom).
+This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.}
 Interestingly, the trends in eigenvalues that result in greater $p_A$ increase the error from delay period inactivation in the Anti task ($p_A - p_{A, opto}$) (Fig. \ref{fig:SC}E.
 Despite these opposing effects, $p_A$ is uncorrelated with ($p_A - p_{A, opto}$), while $p_P$ is correlated (Fig. \ref{fig:SX4}).
-This shows that the degree of error introduced by delay period inactivation is controlled by the dynamical mechanisms controlling success in the Pro task (namely $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$).
+\textbf{This shows that the connectivity mechanisms that decrease $p_A$ during delay period inactivation are the same mechanisms producing greater $p_P$ during no activation (namely $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$).}
 Such comparable dynamics between the non-inactivated Pro task and the delay period inactivated Anti task are salient in simulations (Fig. \ref{fig:SC}D top-left and bottom-right).

### 2020-12-17 epi/SSSN.py
SSSN func updates from desktop

@@ -69,7 +69,7 @@ def SSSN_sim_traj(eps):
         return x_t
     return _SSSN_sim_traj
 
-def SSSN_sim(eps):
+def SSSN_sim(eps, N=1):
     sigma_eps = eps*np.array([1., 1., 1., 1.], np.float32)
     sigma_eps = sigma_eps[None,None,:,None]
     def _SSSN_sim(h):
@@ -143,6 +143,18 @@ def SSSN_sim_sigma_c1(sigma):
         
     x_ss = euler_sim_stoch(f, y_init, dt, T)
     return x_ss
+
+def ISN_coeff(dh, H):
+    sssn_sim = SSSN_sim(0.)
+    h = H + dh
+    h_E = h[:,0]
+
+    r_ss = sssn_sim(h)
+    u_E = tf.linalg.matvec(r_ss[:,0,:4], W_mat[0,:])
+    u_E = u_E + h_E
+    u_E = tf.nn.relu(u_E)
+    isn_coeff = 1.-2.*(u_E)*W_mat[0,0]
+    return isn_coeff
 
 """def SSSN_sim_tfp(h):
     h = h[:,None,:,None]

### 2020-12-18 written/eLifeRevision/Bittner_eLife2020.tex
STG section update

@@
 Samples are colored by log probability density.  
-Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
+Distribution contours of hub neuron frequency from mean of .55 Hz are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
+Frequencies are averages over the stochasticity of the model.
 Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.

### 2020-12-19 written/eLifeRevision/Bittner_eLife2020.tex
STG updates + STG methods

@@
  %\end{equation}
-\begin{equation}\label{eq:EP}
+\begin{equation}\label{eq:EP_STG}
 \begin{split}
@@
 
-Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP}.
+Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
 This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
@@
 
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP}
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}
  \begin{equation}
@@
 which are constrained to the chosen means and variances
-\begin{equation}
+\begin{equation} \label{eq:mu_opt}
 \bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
@@
 \subsubsection{Stomatogastric ganglion}\label{methods_STG}
-We analyze how the parameters $z = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $x = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
-Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
-\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(x; z) + h_{Ca}(x; z) + h_K(x; z) + h_{hyp}(x; z) + h_{elec}(x; z) + h_{syn}(x; z)\right].
-\end{equation} 
-The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $z$.
+%We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
+%Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
+%\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(\mathbf{x}; \mathbf{z}) + h_{Ca}(\mathbf{x}; \mathbf{z}) + h_K(\mathbf{x}; \mathbf{z}) + h_{hyp}(\mathbf{x}; \mathbf{z}) + h_{elec}(\mathbf{x}; \mathbf{z}) + h_{syn}(\mathbf{x}; \mathbf{z})\right].
+%\end{equation} 
+%The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $\mathbf{z}$.
+%The capacitance of the cell membrane was set to $C_m = 1nF$.
+Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
+%\begin{equation}  h_{leak}(\mathbf{x}; \mathbf{z}) = g_{leak} (x_\alpha - V_{leak}) 
+%\end{equation} 
+%\begin{equation}  h_{elec}(\mathbf{x}; \mathbf{z}) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
+%\end{equation} 
+%\begin{equation}  h_{syn}(\mathbf{x}; \mathbf{z}) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
+%\begin{equation}  h_{Ca}(\mathbf{x}; \mathbf{z}) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
+%\end{equation} 
+%\begin{equation}  h_K(\mathbf{x}; \mathbf{z}) = g_K N (x_\alpha - V_K) 
+%\end{equation} 
+%\begin{equation}  h_{hyp}(\mathbf{x}; \mathbf{z}) = g_h H(x_\alpha - V_{hyp}).
+%\end{equation} 
+%The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  
+%The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  
+%The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  
+%The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.  
+%Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.
+%
+%Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
+%\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
+%\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
+%\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
+%\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
+%\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
+%\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
+%\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
+% \end{equation}
+%where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
+%$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.  
+%
+%Finally, there is a synaptic gating variable as well:
+%\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
+%\end{equation}
+%When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
+%
+%In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.  
+%In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible. 
+%The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
+%Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.  
+%Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$
+%
+%Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.  
+%High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.  
+%Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.  
+%The signal was subsequently mean centered.  
+%After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.
+%
+%Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.  
+%Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
+%The ``frequency identification" vector is 
+%\begin{equation}
+%v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
+%\end{equation}
+%The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.
+%
+%Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.  
+%The emergent property statistics are the first and second moments of the firing frequencies. 
+%The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
+%\begin{equation}
+%E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
+%\end{equation}
+%for the EPI distribution shown in Fig. \ref{fig:STG}B.
+%Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  
+%The optimization converged after four epochs (Fig. S4).
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
+%\end{center}
+%\begin{flushleft}
+%{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
+%A. Entropy throughout optimization. 
+%B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
+%Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
+%C. The second moment emergent property statistics converge to the emergent property values.}
+%\end{flushleft}
+%\end{figure}
+%
+%For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
+%We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce an isotropic 
+%Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
+%
+%We calculated the Hessian at the mode of the inferred EPI distribution.  
+%The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
+%With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
+%In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
+%The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
+%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
+%Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
+%The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
+%In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
+%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
+\ref{fig:STG}B, inset).
+%The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
+
+We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of intermediate hub frequency in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
+Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following stochastic differential equation:
+\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(\mathbf{x}; \mathbf{z}) + h_{Ca}(\mathbf{x}; \mathbf{z}) + h_K(\mathbf{x}; \mathbf{z}) + h_{hyp}(\mathbf{x}; \mathbf{z}) + h_{elec}(\mathbf{x}; \mathbf{z}) + h_{syn}(\mathbf{x}; \mathbf{z})\right] + dB.
+\end{equation} 
+The input current of each neuron is the sum of the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents as well as gaussian noise $dB$.
+Each current component is a function of all membrane potentials and the conductance parameters $\mathbf{z}$.
+
 The capacitance of the cell membrane was set to $C_m = 1nF$.
 Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
-\begin{equation}  h_{leak}(x; z) = g_{leak} (x_\alpha - V_{leak}) 
-\end{equation} 
-\begin{equation}  h_{elec}(x; z) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
-\end{equation} 
-\begin{equation}  h_{syn}(x; z) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
-\begin{equation}  h_{Ca}(x; z) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
-\end{equation} 
-\begin{equation}  h_K(x; z) = g_K N (x_\alpha - V_K) 
-\end{equation} 
-\begin{equation}  h_{hyp}(x; z) = g_h H(x_\alpha - V_{hyp}).
+\begin{equation}  h_{leak}(\mathbf{x}; \mathbf{z}) = g_{leak} (x_\alpha - V_{leak}) 
+\end{equation} 
+\begin{equation}  h_{elec}(\mathbf{x}; \mathbf{z}) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
+\end{equation} 
+\begin{equation}  h_{syn}(\mathbf{x}; \mathbf{z}) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
+\begin{equation}  h_{Ca}(\mathbf{x}; \mathbf{z}) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
+\end{equation} 
+\begin{equation}  h_K(\mathbf{x}; \mathbf{z}) = g_K N (x_\alpha - V_K) 
+\end{equation} 
+\begin{equation}  h_{hyp}(\mathbf{x}; \mathbf{z}) = g_h H(x_\alpha - V_{hyp}).
 \end{equation} 
@@
 When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-
-In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.  
-In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible. 
+Gaussian noise of variance $\epsilon^2 = (1 \times 10^{-12})^2$ amps makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
+
+In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 300$ time steps of $dt = 25ms$.  
 The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
-Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.  
-Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$
-
-Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.  
-High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.  
-Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.  
-The signal was subsequently mean centered.  
-After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.
-
-Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.  
-Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-The ``frequency identification" vector is 
-\begin{equation}
-v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
-\end{equation}
-The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.
-
-Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.  
-The emergent property statistics are the first and second moments of the firing frequencies. 
-The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
-\begin{equation}
-E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
-\end{equation}
-for the EPI distribution shown in Fig. \ref{fig:STG}B.
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  
-The optimization converged after four epochs (Fig. S4).
+We used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution to measure frequency from simulated time series
+\begin{equation}
+\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top..
+\end{equation}
+
+To measure spiking frequency, we processed simulated membrane potentials with a relu (spike extraction) and low-pass filter with averaging window of size 20, then took the frequency with the maximum absolute value of the complex exponential basis coefficients of the processed time-series.
+The first 20 temporal samples of the simulation are ignored to account for initial transients.
+
+To differentiate through the maximum frequency identification, we used a soft-argmax
+Let $X_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
+The soft-argmax is then calculated using temperature parameter $\beta=100$
+\begin{equation}
+\psi_\alpha = \text{softmax}(\beta|X_\alpha|\odot i),
+\end{equation}
+where $i = [0, 1, ..., 100]$.
+The frequency is then calculated as 
+\begin{equation}
+\omega_\alpha = 0.01\psi_\alpha \text{Hz}.
+\end{equation}
+
+Intermediate hub frequency, like all other emergent properties in this work, is defined by the mean and variance of the emergent property statistics.
+In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ to capture variation in frequency between 0.5Hz and 0.6Hz (Equation \ref{eq:EP_STG}).
+As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \\ \left( \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) - 0.55 \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} 0.55 \\ 0.025^2 \end{bmatrix}.
+\end{equation}
+Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 5,000 iterations(see Section \ref{methods_AL_opt}).  
+The optimization converged after five epochs (Fig. S4).
 
@@
 \begin{center}
-\includegraphics[scale=0.45]{figures/figS4/figS4.pdf}
+\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
 \end{center}
@@
 B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
-Since $q_{\bm{\theta}}(z)$ failed to produce enough samples yielding $\omega_{\text{f1}}(x)$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
+Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
 C. The second moment emergent property statistics converge to the emergent property values.}
@@
 
-For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $z = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic 
-Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
-
-We calculated the Hessian at the mode of the inferred EPI distribution.  
-The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(z)$ with respect to the parameters $z$:  $\frac{\partial^2 \log q_{\bm{\theta}}(z)}{\partial z \partial z^\top}$.
-With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
-The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
-We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
-Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
-The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
-In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
+For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 25 units per layer.
+The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$. 
+We did not include $g_{\text{synA}} < 0.01$, since conductances that low make the circuit simulations numerically unstable.
+We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, set $\nu = 0.25$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
+This initialization had much greater entropy and a different emergent property than the the returned EPI posterior.
+
+TODO write about specifics of the Hessian analysis.
+%We calculated the Hessian at the mode of the inferred EPI distribution.  
+%The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
+%With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
+%In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
+%The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
+%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
+%Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
+%The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
+%In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
+%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
 \ref{fig:STG}B, inset).
-The contours of Figure \ref{fig:STG} were calculated as error in $T(x)$ from $\mu$ in both the first and second moment emergent property statistics.
+%The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
 
@@
 
-In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = z^* = \argmax_{z} \log q_{\bm{\theta}}(z \mid \mathcal{B}(\alpha, 0.1)$.
+In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = \mathbf{z}^* = \argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{B}(\alpha, 0.1)$.
 The differential responses $\delta x_{\alpha,ss}$ are examined at perturbed inputs  $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$ and $\delta x$ is evaluated at 101 equally spaced samples of $\delta h_\alpha$ from -15 to 15.

### 2020-12-20 written/eLifeRevision/Bittner_eLife2020.tex
SC methods

@@
 \begin{split}
-C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) = - [ &\mathbf{h}_{leak}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{Ca}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_K(\mathbf{x}(t); \mathbf{z}) \\
- & + \mathbf{h}_{hyp}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{elec}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{syn}(\mathbf{x}(t); \mathbf{z})] 
+C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
  \end{split}
 \end{equation} 
-where $C_m$=1nF, and $\mathbf{h}_{\text{leak}}$, $\mathbf{h}_{Ca}$, $\mathbf{h}_K$, $\mathbf{h}_{\text{hyp}}$, $\mathbf{h}_{\text{elec}}$, and $\mathbf{h}_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).
+where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG}).
 
 %Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Second, we define the emergent property, which as above is middle ``hub neuron frequency": oscillation of the hub neuron at an intermediate frequency (Figure \ref{fig:STG}C).
+Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
 %Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
@@
 \subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
+%In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
+\ref{fig:SC}a).
+%Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
+%Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
+%We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig.
+\ref{fig:SC}B).
+%The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
+%\begin{equation}
+%x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
+%\end{equation}
+%where $\epsilon = 0.05$ and $\zeta = 0.5$.  
+%The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
+%\begin{equation}
+%\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
+%\end{equation}
+%with gaussian noise of variance $\sigma^2$ = 1.
+%The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
+%Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
+%\end{center}
+%\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
+%A. Rapid task switching behavioral paradigm (see text). 
+%B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
+%Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
+%Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
+%C.
+The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
+Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
+%D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
+%E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
+%\label{fig:SC}
+%\end{figure}
+%
+%To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
+%At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
+%Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
+%Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
+%\begin{equation}
+%\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
+%\end{equation}
+%Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.
+%
+%We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
+%To make sense of these inferred distributions, we followed the approach of Duan et al.
+by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
+\ref{fig:SC}C).
+%These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
+%The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
+%
+%We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig.
+\ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
+%Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
+%Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
+%There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
+%Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
+%These hypotheses are confirmed by forward simulation of the SC model (Fig.
+\ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.
+
 In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}a).
+\ref{fig:SC}A).
 Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
@@
 \ref{fig:SC}B).
-The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
-\begin{equation}
-x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
-\end{equation}
-where $\epsilon = 0.05$ and $\zeta = 0.5$.  
-The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
-\begin{equation}
-\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
-\end{equation}
-with gaussian noise of variance $\sigma^2$ = 1.
-The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.
+The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
+The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
+The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
+\begin{equation}
+\begin{split}
+\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
+\mathbf{x} &= f(\mathbf{u})
+\end{split}
+\end{equation}
+with white noise of variance $\epsilon^2 = 0.2^2$.
+The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
+Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.
 
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
+\includegraphics[scale=0.48]{figures/fig4/fig4.pdf}
 \end{center}
-\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
+\caption{\small 
 A. Rapid task switching behavioral paradigm (see text). 
@@
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
-C.
-The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
-Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
-D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
-E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
+Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
+C. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice u=of simulations.
+D.
+Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.
+Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (opto strength 0.7).
+Shading indicates standard deviation across trials.
+E.
+Difference in performance of each task during inactivation.
+Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).
+We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
+F. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
+}
 \label{fig:SC}
@@
 
-To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
-At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
-Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
-Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
-\begin{equation}
-\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
-\end{equation}
-Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.
-
-We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
-To make sense of these inferred distributions, we followed the approach of Duan et al.
-by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-\ref{fig:SC}C).
+We define rapid task switching behavior as accurate execution of each task.
+Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
+We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.
+\begin{equation}\label{eq:EP}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 5\%^2 \\ 5\%^2  \end{bmatrix}
+\end{split}
+\end{equation}
+A variance of $5\%$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.
+
+We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
+Some parameters were predictive of accuracy while others were not (Fig. \ref{fig:SC1}), and often had different effects on $p_P$ and $p_A$.
+To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $v_i$ for all $W$ parameterized by $z$ (Fig.
+\ref{fig:SC2}A).
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.
-
-We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig.
-\ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
-Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
-Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
-There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
-Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
-These hypotheses are confirmed by forward simulation of the SC model (Fig.
-\ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.
+
+Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
+This shows that strong task representations and hemispheral dominance in the dynamics result in better execution of the Pro task.
+By visualizing these four variables together by $p_A$ (Fig.
+\ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
+Thus, stronger hemispheral dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.
+
+In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
+\ref{fig:SC}E) e.g.
+(Fig.
+\ref{fig:SC}D, bottom).
+This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
+Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task (Fig.
+\ref{fig:SC4}).
+Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation (Fig. \ref{fig:SC}F)).
 
@@
 \begin{center}
-\includegraphics[scale=0.58]{figures/figS1/figS1.pdf}
+\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
 \end{center}
 \begin{flushleft}
-{\small Fig. S1: A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
+\caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
 B.
@@
 \end{flushleft}
+\label{fig:LDS1}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figS2/figS2.pdf}
+\includegraphics[scale=0.5]{figures/figLDS2/figLDS2.pdf}
 \end{center}
 \begin{flushleft}
-{\small Fig. S2: A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+\caption{\small (LDS2): A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
 B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
 \end{flushleft}
+\label{fig:LDS2}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figS3/figS3.pdf}
+\includegraphics[scale=0.5]{figures/figLDS3/figLDS3.pdf}
 \end{center}
 \begin{flushleft}
-{\small Fig.
-S3: Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
 A.
@@
 \end{flushleft}
+\label{fig:LDS3}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
+\includegraphics[scale=0.6]{figures/figSTG1/figSTG1.pdf}
 \end{center}
 \begin{flushleft}
-{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
+\caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
 A. Entropy throughout optimization. 
-B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
-Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
-C. The second moment emergent property statistics converge to the emergent property values.}
+B. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
 \end{flushleft}
+\label{fig:STG1}
 \end{figure}
@@
 \subsubsection{Superior colliculus}\label{methods_SC}
+%In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
+%They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
+%Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
+%\begin{equation}
+%x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
+%\end{equation}
+%where $\alpha \in \{LP, LA, RA, RP\}$ $\epsilon = 0.05$ and $\zeta = 0.5$ control the position and shape of the nonlinearity, respectively.
+%
+%We order the elements of $x$ and $u$ in the following manner
+%\begin{equation}
+%x = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} u = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix}.
+%\end{equation}
+%
+% The internal variables follow dynamics:
+%\begin{equation}
+%\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
+%\end{equation}
+%with time constant $\tau = 0.09s$ and Gaussian noise $\sigma dB$ controlled by the magnitude of $\sigma=1.0$.
+The weight matrix has 8 parameters $sW_P$, $sW_A$, $vW_{PA}$, $vW_{AP}$, $hW_P$, $hW_A$, $dW_{PA}$, and $dW_{AP}$ (Fig.
+4B):
+%\begin{equation}
+%W = \begin{bmatrix} sW_P & vW_{PA} & hW_P & dW_{PA}  \\ vW_{AP}  & sW_A & dW_{AP}  & hW_A \\ hW_P & dW_{PA}  & sW_P & vW_{PA}  \\ dW_{AP}  & hW_A & vW_{AP}  & sW_A \end{bmatrix}.
+%\end{equation}
+%
+%The system receives five inputs throughout each trial, which has a total length of 1.8s.
+%\begin{equation}
+%h = h_{\text{rule}} + h_{\text{choice-period}} + h_{\text{light}}.
+%\end{equation}
+%
+%There are rule-based inputs depending on the condition,
+%\begin{equation}h_{\text{P,rule}}(t) = \begin{cases}
+%                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
+%                            0,              & \text{otherwise}
+%                         \end{cases}
+%\end{equation}
+%\begin{equation} h_{\text{A,rule}}(t) = \begin{cases}
+%                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
+%                            0,              & \text{otherwise}
+%                         \end{cases}
+%\end{equation}
+%a choice-period input,
+%\begin{equation} h_{\text{choice}}(t) = \begin{cases}
+%                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
+%                            0,              & \text{otherwise}
+%                         \end{cases}
+%\end{equation}
+%and an input to the right or left-side depending on where the light stimulus is delivered.     
+%\begin{equation}  h_{\text{light}}(t) = \begin{cases}
+%                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } t > 1.2s \text{ and Left} \\
+%                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } t > 1.2s \text{ and Right} \\
+%                            0,              & t \leq 1.2s
+%                         \end{cases} .
+%\end{equation}
+%The input parameterization was fixed to $I_{\text{P,rule}} = 10 $,  $I_{\text{A,rule}} = 10$,  $I_{\text{choice}} = 2$,  and $I_{\text{light}} = 1$.
+%
+%To produce an accuracy rate of $p_{LP}$ in the Left, Pro condition, let $\hat{p}_i$ be the empirical average steady state response (final $x_{LP}$ at end of task) over M=500 Gaussian noise draws for a given SC model parameterization $z_i$:
+%\begin{equation}
+% \hat{p}_i = \mathbb{E}_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] = \frac{1}{M}\sum_{j=1}^M x_{LP}(s=L, c=P, z=z_i, \sigma dB_j)
+% \end{equation}
+% where stimulus $s \in \{L, R\}$, cue $c \in \{P, A\}$, and $\sigma dB_j$ is the Gaussian noise on trial $j$.
+%As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
+%For the first emergent property statistic, the average over EPI samples (from $q_{\bm{\theta}}(z)$) is set to the desired value $p_{LP}$:
+%\begin{equation}
+%\mathbb{E}_{z_i \sim q_\phi} \left[ \mathbb{E}_{\sigma dB} \left[ x_{LP,\text{ss}} \mid s=L, c=P, z=z_i \right] \right] = \mathbb{E}_{z_i \sim q_\phi} \left[ \hat{p}_i \right] = p_{LP}.
+%\end{equation}
+%
+%For the next emergent property statistic, we ask that the variance of the steady state responses across Gaussian draws, is the Bernoulli variance for the empirical rate $\hat{p}_i$:
+%\begin{equation}
+%\mathbb{E}_{z \sim q\phi} \left[ \sigma^2_{err} \right] = 0
+%\end{equation}
+%where the Bernoulli variance error $\sigma^2_{err}$ for the Pro task, left condition is
+%\begin{equation}
+%\sigma^2_{err} = Var_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] - \hat{p}_i(1 - \hat{p}_i).
+%\end{equation}
+%
+%We have an additional constraint that the Pro neuron on the opposite hemisphere should have the opposite value (0 and 1).  We can enforce this with another constraint:
+%\begin{equation}
+%\mathbb{E}_{z \sim q\phi} \left[ d_P \right] = 1,
+%\end{equation}
+%where the distance between Pro neuron steady states $d_P$ in the Pro condition is
+%\begin{equation}
+%d_P =  \mathbb{E}_{\sigma dB} \left[ (x_{LP} - x_{RP})^2  \mid s=L, c=P, z=z_i \right]
+%\end{equation}
+%The emergent property statistics only need to be measured during the Left stimulus condition of the Pro and Anti tasks, since the network is symmetrically parameterized.
+%In total, the emergent property of rapid task switching at accuracy level $p$ was defined as
+%\begin{equation}
+%\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
+%\end{equation}
+%
+%Since the maximum variance of a random variable bounded from 0 to 1 is the Bernoulli variance $\hat{p}(1-\hat{p})$, and the maximum squared difference between to variables bounded from 0 to 1 is 1, we do not need to control the second moment of these test statistics.
+%These variables are dynamical system states and can only exponentially decay (or saturate) to 0 (or 1), so the Bernoulli variance error and squared difference constraints cannot be satisfied exactly in simulation.
+%This is important to be mindful of when evaluating the convergence criteria.  
+%Instead of using our usual hypothesis testing criteria for convergence to the emergent property, we set a slack variable threshold only for these technically infeasible emergent property values to 0.05.
+%
+%Using EPI to learn distributions of dynamical systems producing Bernoulli responses at a given rate (with small variance around that rate) was more challenging than expected.  
+%There is a pathology in this optimization setup, where the learned distribution of weights is bimodal attributing a fraction $p$ of the samples to an expansive mode (which always sends $x_{LP}$ to 1), and a fraction $1-p$ to a decaying mode (which always sends $x_{LP}$ to 0).
+%This pathology was avoided using an inequality constraint prohibiting parameter samples that resulted in low variance of responses across noise.
+%
+%
+%\begin{table}[h]
+%\begin{tabular}{l|l|l|l|l}
+%$\lambda$  & $\hat{p}$ & $q_{\bm{\theta}}(z)$ & $r$ & p-value \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $7.56 \times 10^{-01}$ & $p<10^{-4}$ \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $4.59 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $3.76 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%
+%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $4.80 \times 10^{-02}$ & $p<.01$ \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $2.08 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $4.84 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(90\%))$ & $4.25 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%
+%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(50\%))$ & $-7.57 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $-6.73 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $-4.86 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $-1.43 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $-1.93 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%
+%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $-7.60 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $-2.73 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $-2.74 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
+%\end{tabular}
+%\caption{\small Table of significant correlation values from Fig. \ref{fig:SC}E.}
+%\end{table}
+%
+%For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
+%We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
+%The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.
+%
+%We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  
+%Correlations were measured from 5,000 samples of $q_{\bm{\theta}}(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
+
 In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
@@
 \begin{equation}
-x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
-\end{equation}
-where $\alpha \in \{LP, LA, RA, RP\}$ $\epsilon = 0.05$ and $\zeta = 0.5$ control the position and shape of the nonlinearity, respectively.
-
-We order the elements of $x$ and $u$ in the following manner
-\begin{equation}
-x = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} u = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix}.
-\end{equation}
-
- The internal variables follow dynamics:
-\begin{equation}
-\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
-\end{equation}
-with time constant $\tau = 0.09s$ and Gaussian noise $\sigma dB$ controlled by the magnitude of $\sigma=1.0$.
-The weight matrix has 8 parameters $sW_P$, $sW_A$, $vW_{PA}$, $vW_{AP}$, $hW_P$, $hW_A$, $dW_{PA}$, and $dW_{AP}$ (Fig.
-4B):
-\begin{equation}
-W = \begin{bmatrix} sW_P & vW_{PA} & hW_P & dW_{PA}  \\ vW_{AP}  & sW_A & dW_{AP}  & hW_A \\ hW_P & dW_{PA}  & sW_P & vW_{PA}  \\ dW_{AP}  & hW_A & vW_{AP}  & sW_A \end{bmatrix}.
-\end{equation}
-
-The system receives five inputs throughout each trial, which has a total length of 1.8s.
-\begin{equation}
-h = h_{\text{rule}} + h_{\text{choice-period}} + h_{\text{light}}.
-\end{equation}
-
-There are rule-based inputs depending on the condition,
-\begin{equation}h_{\text{P,rule}}(t) = \begin{cases}
+x_\alpha = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
+\end{equation}
+where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
+
+We order the neural populations of $x$ and $u$ in the following manner
+\begin{equation}
+\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
+\end{equation}
+
+which evolve according to
+\begin{equation}
+\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + \sigma d\mathbf{B}.
+\end{equation}
+with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $dB$of variance $\sigma^2=0.2$.  
+The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
+\begin{equation}
+W = \begin{bmatrix} sW & vW & hW & dW \\ vW & sW & dW  & hW \\ hW & dW & sW & vW \\ dW  & hW & vW  & sW \end{bmatrix}.
+\end{equation}
+
+The circuit receives four different inputs throughout each trial, which has a total length of 1.8s.
+\begin{equation}
+\mathbf{h} = \mathbf{h}_{\text{constant}}  + \mathbf{h}_{\text{P,bias}} + \mathbf{h}_{\text{rule}} +  \mathbf{h}_{\text{choice-period}} + \mathbf{h}_{\text{light}}.
+\end{equation}
+
+There is a constant input to every population,
+\begin{equation}\mathbf{h}_{\text{constant}} = I_{\text{constant}} [1, 1, 1, 1]\top,
+\end{equation}
+ a bias to the Pro populations
+\begin{equation}\mathbf{h}_{\text{P,bias}} = I_{\text{P,bias}} [1, 0, 1, 0]\top,
+\end{equation}
+rule-based input depending on the condition
+\begin{equation}\mathbf{h}_{\text{P,rule}}(t) = \begin{cases}
                            I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
@@
 \end{equation}
-\begin{equation} h_{\text{A,rule}}(t) = \begin{cases}
+\begin{equation} \mathbf{h}_{\text{A,rule}}(t) = \begin{cases}
                            I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
                             0,              & \text{otherwise}
-                         \end{cases}
-\end{equation}
-a choice-period input,
-\begin{equation} h_{\text{choice}}(t) = \begin{cases}
+                         \end{cases},
+\end{equation}
+a choice-period input
+\begin{equation} \mathbf{h}_{\text{choice}}(t) = \begin{cases}
                            I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
                             0,              & \text{otherwise}
-                         \end{cases}
-\end{equation}
-and an input to the right or left-side depending on where the light stimulus is delivered.     
-\begin{equation}  h_{\text{light}}(t) = \begin{cases}
-                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } t > 1.2s \text{ and Left} \\
-                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } t > 1.2s \text{ and Right} \\
-                            0,              & t \leq 1.2s
-                         \end{cases} .
-\end{equation}
-The input parameterization was fixed to $I_{\text{P,rule}} = 10 $,  $I_{\text{A,rule}} = 10$,  $I_{\text{choice}} = 2$,  and $I_{\text{light}} = 1$.
-
-To produce an accuracy rate of $p_{LP}$ in the Left, Pro condition, let $\hat{p}_i$ be the empirical average steady state response (final $x_{LP}$ at end of task) over M=500 Gaussian noise draws for a given SC model parameterization $z_i$:
-\begin{equation}
- \hat{p}_i = \mathbb{E}_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] = \frac{1}{M}\sum_{j=1}^M x_{LP}(s=L, c=P, z=z_i, \sigma dB_j)
- \end{equation}
- where stimulus $s \in \{L, R\}$, cue $c \in \{P, A\}$, and $\sigma dB_j$ is the Gaussian noise on trial $j$.
-As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
-For the first emergent property statistic, the average over EPI samples (from $q_{\bm{\theta}}(z)$) is set to the desired value $p_{LP}$:
-\begin{equation}
-\mathbb{E}_{z_i \sim q_\phi} \left[ \mathbb{E}_{\sigma dB} \left[ x_{LP,\text{ss}} \mid s=L, c=P, z=z_i \right] \right] = \mathbb{E}_{z_i \sim q_\phi} \left[ \hat{p}_i \right] = p_{LP}.
-\end{equation}
-
-For the next emergent property statistic, we ask that the variance of the steady state responses across Gaussian draws, is the Bernoulli variance for the empirical rate $\hat{p}_i$:
-\begin{equation}
-\mathbb{E}_{z \sim q\phi} \left[ \sigma^2_{err} \right] = 0
-\end{equation}
-where the Bernoulli variance error $\sigma^2_{err}$ for the Pro task, left condition is
-\begin{equation}
-\sigma^2_{err} = Var_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] - \hat{p}_i(1 - \hat{p}_i).
-\end{equation}
-
-We have an additional constraint that the Pro neuron on the opposite hemisphere should have the opposite value (0 and 1).  We can enforce this with another constraint:
-\begin{equation}
-\mathbb{E}_{z \sim q\phi} \left[ d_P \right] = 1,
-\end{equation}
-where the distance between Pro neuron steady states $d_P$ in the Pro condition is
-\begin{equation}
-d_P =  \mathbb{E}_{\sigma dB} \left[ (x_{LP} - x_{RP})^2  \mid s=L, c=P, z=z_i \right]
-\end{equation}
-The emergent property statistics only need to be measured during the Left stimulus condition of the Pro and Anti tasks, since the network is symmetrically parameterized.
-In total, the emergent property of rapid task switching at accuracy level $p$ was defined as
-\begin{equation}
-\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
-\end{equation}
-
-Since the maximum variance of a random variable bounded from 0 to 1 is the Bernoulli variance $\hat{p}(1-\hat{p})$, and the maximum squared difference between to variables bounded from 0 to 1 is 1, we do not need to control the second moment of these test statistics.
-These variables are dynamical system states and can only exponentially decay (or saturate) to 0 (or 1), so the Bernoulli variance error and squared difference constraints cannot be satisfied exactly in simulation.
-This is important to be mindful of when evaluating the convergence criteria.  
-Instead of using our usual hypothesis testing criteria for convergence to the emergent property, we set a slack variable threshold only for these technically infeasible emergent property values to 0.05.
-
-Using EPI to learn distributions of dynamical systems producing Bernoulli responses at a given rate (with small variance around that rate) was more challenging than expected.  
-There is a pathology in this optimization setup, where the learned distribution of weights is bimodal attributing a fraction $p$ of the samples to an expansive mode (which always sends $x_{LP}$ to 1), and a fraction $1-p$ to a decaying mode (which always sends $x_{LP}$ to 0).
-This pathology was avoided using an inequality constraint prohibiting parameter samples that resulted in low variance of responses across noise.
-
-\begin{table}[h]
-\begin{tabular}{l|l|l|l|l}
-$\lambda$  & $\hat{p}$ & $q_{\bm{\theta}}(z)$ & $r$ & p-value \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $7.56 \times 10^{-01}$ & $p<10^{-4}$ \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $4.59 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $3.76 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $4.80 \times 10^{-02}$ & $p<.01$ \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $2.08 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $4.84 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(90\%))$ & $4.25 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(50\%))$ & $-7.57 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $-6.73 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $-4.86 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $-1.43 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $-1.93 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $-7.60 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $-2.73 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $-2.74 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-\end{tabular}
-\caption{\small Table of significant correlation values from Fig. \ref{fig:SC}E.}
-\end{table}
-
-For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
-We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
-The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.
-
-We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  
-Correlations were measured from 5,000 samples of $q_{\bm{\theta}}(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
+                         \end{cases},
+\end{equation}
+and an input to the right or left-side depending on where the light stimulus is delivered    
+\begin{equation}  \mathbf{h}_{\text{light}}(t) = \begin{cases}
+                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Left} \\
+                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Right} \\
+                            0,              & \text{otherwise}
+                         \end{cases}.
+\end{equation}
+The input parameterization was fixed to $I_{\text{constant}} = 0.75$, $I_{\text{P,bias}} = 0.5 $, $I_{\text{P,rule}} = 0.6$,  $I_{\text{A,rule}} = 0.6$,  $I_{\text{choice}} = 0.25$,  and $I_{\text{light}} = 0.5$.
+
+The accuracies of $p_P$ and $p_A$ are calculated as
+\begin{equation}
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+\end{equation}
+and 
+\begin{equation}
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+\end{equation}
+given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function.
+
+The Heaviside step function is approximated as
+\begin{equation}
+\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
+\end{equation}
+where $\beta = 100$.
+
+As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 5\%^2 \\ 5\%^2 \end{bmatrix}.
+\end{equation}
+Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
+The optimization converged after six epochs (Fig. \ref{fig:SC5}).
+
+For EPI in Fig \label{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
+The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
+We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce an isotropic gaussian with mean 0 and variance $2.5^2$.
+Accuracies were estimated over 200 trials of random gaussian noise, which was sampled independently for each drawn parameter $\mathbf{z}$ and each iteration of the EPI optimization.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.85]{figures/figSC1/figSC1.pdf}
+\end{center}
+\caption{\small (SC1): Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p value.
+}
+\label{fig:SC1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.3]{figures/figSC2/figSC2.pdf}
+\end{center}
+\caption{\small (SC2):  
+A. Invariant eigenvectors of connectivity matrix $W$.
+B. Eigenvalues of connectivities of EPI distribution versus task accuracies.}
+\label{fig:SC2}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.6]{figures/figSC3/figSC3.pdf}
+\end{center}
+\caption{\small (SC3): 
+A. Connectitivy eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
+B. Same for Anti task.
+}
+\label{fig:SC3}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.75]{figures/figSC4/figSC4.pdf}
+\end{center}
+\caption{\small (SC4):  Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.8.
+}
+\label{fig:SC4}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.6]{figures/figSC5/figSC5.pdf}
+\end{center}
+\caption{\small (SC5): Caption...
+}
+\label{fig:SC5}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
-\end{center}
-{\small Fig.
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
+%\end{center}
+%{\small Fig.
 S5: A.
@@
 See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
-\end{figure}
+%\end{figure}
 

### 2020-12-31 written/eLifeRevision/Bittner_eLife2020.tex
SC5 caption
tex updates

@@
 {\Large Interrogating theoretical models of neural computation with deep inference} \\
-Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3}$, Chunyu A. Duan$^{4}$, Carlos D. Brody$^{2,3,5}$, \\
-Kenneth D. Miller$^{1}$, and John P. Cunningham$^{6}$.
+Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3,4}$, Chunyu A. Duan$^{5}$, Carlos D. Brody$^{2,3,6}$, \\
+Kenneth D. Miller$^{1}$, and John P. Cunningham$^{7}$.
 
@@
 $^{3}$Princeton University, \\
-$^{4}$Institute of Neuroscience, Chinese Academy of Sciences, \\
-$^{5}$Howard Hughes Medical Institute, \\
-$^{6}$Department of Statistics, Columbia University
+$^{4}$Allen Institute for Brain Science, \\
+$^{5}$Institute of Neuroscience, Chinese Academy of Sciences, \\
+$^{6}$Howard Hughes Medical Institute, \\
+$^{7}$Department of Statistics, Columbia University
 }
@@
 We bring the use of deep generative models for probabilistic inference to bear on this problem, learning distributions of parameters that produce the specified properties of computation.
-By learning parameter distributions that produce computations -- an emergent property, we introduce a novel methodology for exploratory analyses and hypothesis testing that is particularly well-suited to the stochastic dynamical systems models predominant in our field of theoretical neuroscience.
+By learning parameter distributions that produce computations -- an emergent property, we introduce a novel methodology that is particularly well-suited to the stochastic dynamical systems models predominant in our field of theoretical neuroscience.
 %Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
 We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
-We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
+%We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
+We then use it to reveal the key factors of variability in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and scale inference of large low-rank RNN's exhibiting stable amplification.
 While much use of deep learning in theoretical neuroscience focuses on drawing analogies between optimized neural architectures and the brain, this work illustrates how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
@@
 Thus, standard approaches from statistical inference are unavailable.
-The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except in the unrealistic $0$-distance scenario), lack parameter probabilities, and only confer sensitivity analyses of an alternative likelihood to the simulator-defined likelihood of ABC \cite{raue2009structural}.
+The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except in the unrealistic $0$-distance scenario), and lack parameter probabilities.
 Second is the undesirable trade-off between the flexibility and tractability of the approximated posterior distribution.
@@
 Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to infinite neurons).
+These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to 100 neurons).
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-First, we use EPI to elucidate the mechanisms of inhibition stabilization with varying contrast in a stochastic nonlinear dynamical model of primary visual cortex with inhibitory multiplicity.
+First, we use EPI to understand the characteristics of noise that govern Fano factor in a stochastic four neuron-type model of primary visual cortex.
 %Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-Second, we discover connectivity patterns in superior colliculus resulting in resilience to optogenetic perturbation by using EPI to condition on rapid task switching.
-Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
-The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{gutierrez2013multiple, litwin2016inhibitory, duan2018collicular, mastrogiuseppe2018linking}, and more generally, these results point to the value of deep inference for the interrogation of biologically relevant models.
+Second, we discover connectivity patterns in superior colliculus resilient to optogenetic perturbation by using EPI to condition on rapid task switching.
+The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{litwin2016inhibitory, duan2018collicular}.
+%Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
+Third, we emphasize the methodological advancement of EPI by inferring high-dimensional distributions of RNN connectivities exhibiting stable amplification.
+These results point to the value of deep inference for the interrogation of biologically relevant models.
 
@@
 Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-Parameters are represented by the vector $\mathbf{z}$ and data by the vector $\mathbf{x}$.
-\textbf{B}. Simulated activity form the STG model at $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
-\textbf{C}. The emergent property of unified intermediacy, in which all neurons are firing close to the same intermediate frequency.
-Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution. 
+Parameters are represented by the vector $\mathbf{z}$ and membrane potentials by the vector $\mathbf{x}$.
+The evolution of this model's activity $\mathbf{x}(t)$ is predicated by differential equations.
+\textbf{B}.
+Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic.
+Spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
+\textbf{C}. The emergent property of intermediate hub frequency, in which the hb neuron fires at a rate between the fast and slow frequencies.
+Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution (Panel E).
 \textbf{D}.
-For a choice of model and emergent property, emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing middle hub frequency.
+For a choice of model and emergent property, emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing intermediate hub frequency.
 Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
-\textbf{E}. The EPI distribution of STG model parameters producing network syncing.  
+\textbf{E}. The EPI distribution of STG model parameters producing intermediate hub frequency.
 Samples are colored by log probability density.  
@@
 %(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
-$v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}). 
 \textbf{F} Simulations from parameters in E.  
@@
 \end{center}
-\label{fig:STG}
+\label{fig:STG_EPI}
 \end{figure}
@@
 Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
+\ref{fig:STG_EPI}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
 \ref{fig:STG}E).
@@
 \end{center}
-\caption{\small (SC5): Caption...
+\caption{\small (SC5): EPI optimization of the SC model producing rapid task switching.
+A. Entropy throughout optimization. 
+B. The emergent property statistic means and variances converge to their constraints at 12,000 iterations following the sixth augmented Lagrangian epoch.
 }

### 2021-01-02 written/eLifeRevision/Bittner_eLife2020.tex
latest revision

@@
 While VAE approaches can result in tractable posterior sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions, which restrict the flexibility of the posterior approximation.
-And third, you can never make assumptions of what inferred model parameters will predict.
+And third, one can never assume what inferred model parameters may predict.
 This is well understood when considering Box's loop and the role of posterior predictive checks in the development and critique of scientific models \cite{gelman2013philosophy, blei2014build}.
@@
 Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to 100 neurons).
+These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to hundreds of neurons).
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
@@
 \includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
+\end{center}
 \caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
@@
 Spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
-\textbf{C}. The emergent property of intermediate hub frequency, in which the hb neuron fires at a rate between the fast and slow frequencies.
+\textbf{C}. The emergent property of intermediate hub frequency, in which the hub neuron fires at a rate between the fast and slow frequencies.
 Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution (Panel E).
@@
 Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced.
 \textbf{E}. The EPI distribution of STG model parameters producing intermediate hub frequency.
@@
 (Top) The predictive distribution of the posterior obeys the constraints stipulated by the emergent property.
+The black and gray dashed lines show the mean and two standard deviations according the emergent property, respectively.
 (Bottom) Simulations at the starred parameter values.
  }
-\end{center}
-\label{fig:STG_EPI}
+ \label{fig:STG}
 \end{figure}
@@
 Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG_EPI}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
+\ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
 \ref{fig:STG}E).
@@
 %First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
-First, we consider the model as a coupled set of stochastic differential equations \cite{gutierrez2013multiple}.  
+First, we consider the model as a coupled set of differential equations \cite{gutierrez2013multiple}.  
 In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
@@
 \begin{split}
-C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
+C_m \frac{d\mathbf{x}(t)}{dt} = -h(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
  \end{split}
@@
 %Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$is close to an intermediate frequency of 0.55Hz.  
+Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to an intermediate frequency of 0.55Hz.  
 Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
@@
 where $f(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
+According to the emergent property of interest, $f(\mathbf{x}, \mathbf{z})$ may contain multiple statistics, in which case the mean and variance vectors $\bm{\mu}$ and $\bm{\sigma}^2$ match this dimension.
 %The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
@@
 %Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
-Specifically, EPI (TODO insert interpretation).
+Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_exp_fam_lh}).
+A key advantage of EPI over established Bayesian inference is that the predictions made my the inferred distribution are constrained to produce the specified emergent property.
 %Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
 Equipped with this method, we may examine structure in posterior distributions or make comparisons between posteriors conditioned at different levels of the same emergent property statistic.
-We now prove out the value of EPI by using it to investigate and produce novel insights about three prominent models in neuroscience.
+In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using it to investigate and produce novel insights into two prominent models in neuroscience. 
+Subsequently in Section \ref{results_LRRNN}, we show EPI's superiority in parameter scalability and fidelity of the posterior predictive distribution by conditioning on stable amplification in low-rank RNNs.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comprehensive input-responsivity in a nonlinear sensory system} \label{results_V1}
+\subsection{EPI reveals how noise across neural population types governs Fano factor in a stochastic inhibition stabilized network} \label{results_V1}
 
 \begin{figure}
+\vspace{-1cm}
 \begin{center}
-\includegraphics[scale=0.54]{figures/fig2/fig2.pdf}
+\includegraphics[scale=0.2]{figures/fig2/fig2.pdf}
 \end{center}
-\caption{\small Hypothesis generation through EPI in a V1 model.  
-A. Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   
-Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).  
-B. Linear response predictions become inaccurate with greater input strength.  V1 model simulations for input (solid) $h=b$ and (dashed) $h = b + dh$. 
-Stars indicate the linear response prediction.  
-C. EPI distributions on differential input $dh$ conditioned on differential response $\mathcal{B}(\alpha, y)$. 
-Supporting evidence for the four generated hypotheses are indicated by gray boxes with labels H1, H2, H3, and H4. 
-The linear prediction from two standard deviations away from $y$ (from negative to positive) is overlaid in magenta (very small, near origin). }
- \label{fig:V1_EPI}
+\vspace{-1cm}
+\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
+\textbf{A}.
+Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).
+Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
+Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
+Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
+\textbf{B}. Responses of the deterministic smodel ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
+\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
+\textbf{D}. EPI posterior of noise parameteres $\mathbf{z}$ conditioned on realistic E-population Fano factors.
+The posterior predictive distribution is show on the bottom-left.
+and the mode of the distribution is starred.
+\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
+Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
+The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
+(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
+\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
+\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
+}
+ \label{fig:V1}
 \end{figure}
@@
 Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
 Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) -- composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1_EPI}A) \cite{pfeffer2013inhibition}.
+However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
 Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we go beyond linear theory by systematically generating and evaluating hypotheses of circuit model function using EPI distributions of neuron-type inputs producing various neuron-type population responses.
-
-Specifically, we consider a four-dimensional circuit model with dynamical state given by the firing rate $x$ of each neuron-type population $x = \left[x_E, x_P , x_S, x_V \right]^\top$. 
-Given a time constant of $\tau = 20$ ms and a power $n = 2$, the dynamics are driven by the rectified and exponentiated sum of recurrent ($Wx$) and external $h$ inputs:
-
-\begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
-\end{equation}
-
-We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).
-The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top$ and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.
-Throughout subsequent analyses, the baseline input is $b = \left[ 1 ,1,1,1\right]^\top$.
-
-With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
-Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
-All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
-While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig.
-\ref{fig:V1_EPI}B right).
-Currently available approaches to deriving the steady state response of the system are limited.
-
-To get a more comprehensive picture of the input-responsivity of each neuron-type beyond linear theory, we used EPI to learn a distribution of the differential inputs to each population $dh$ that produce an increase of $y$ in the rate of each neuron-type population $\alpha \in \{E, P, S, V \}$.
-We want to know the differential inputs $dh$ that result in a differential steady state $dx_{\alpha}$ (the change in $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$) of value $y$ with some small, arbitrarily chosen amount of variance  $0.01^2$.
-These statements amount to the emergent property 
-\begin{equation}
-\mathcal{B}(\alpha, y) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ (dx_{\alpha} - y)^2 \end{bmatrix} ~~=~~ \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
-\end{equation}
-We maintain the notation $\mathcal{B}(\cdot)$ throughout the rest of the study as short hand for emergent property, which represents a different signature of computation in each application.
-
-Using EPI, we inferred the distribution of $dh$ shown in Figure \ref{fig:V1_EPI}C producing $\mathcal{B}(\alpha, y)$.  
-Columns correspond to inferred distributions of excitatory ($\alpha=E$, red), parvalbumin ($\alpha=P$, blue), somatostatin ($\alpha=S$, red) and VIP ($\alpha=V$, green) neuron-type response increases, while each row corresponds to increase amounts of $y \in \{0.1, 0.5\}$.
-For each pair of parameters, we show the two-dimensional marginal distribution of samples colored by $\log q_{\bm{\theta}}(dh \mid \mathcal{B}(\alpha, y))$.
-The inferred distributions immediately suggest four hypotheses: \\
-
-{\addtolength{\leftskip}{10 mm}
-H1: as is intuitive, each neuron-type's firing rate should be sensitive to that neuron-type's direct input (e.g.
-Fig.
-\ref{fig:V1_EPI}C H1 gray boxes indicate low variance in $dh_E$ when $\alpha=E$.
-Same observation in all inferred distributions); \\
-H2: the E- and P-populations should be largely unaffected by input to the V-population (Fig. \ref{fig:V1_EPI}C H2 gray boxes indicate high variance in $dh_V$ when $\alpha \in \{E, P \}$); \\
-H3: the S-population should be largely unaffected by input to the P-population  (Fig. \ref{fig:V1_EPI}C H3 gray boxes indicate high variance in $dh_P$ when $\alpha =S$); \\
-H4: there should be a nonmonotonic response of the V-population with input to the E-population (Fig.
-\ref{fig:V1_EPI}C H4 gray boxes indicate that negative $dh_E$ should result in small $dx_{V}$, but positive $dh_E$ should elicit a larger $dx_{V}$);
-
-}
-
-\begin{figure}
-\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=4cm}}]{figure}[\FBwidth]
-{\caption{Confirming EPI generated hypotheses in V1. 
-A. Differential responses $\delta x_E$ by the E-population to changes in individual input $\delta h_\alpha \hat{u}_\alpha$ away from the mode of the EPI distribution $dh^*$. 
-B-D Same plots for the P-, S-, and V-populations.  
-Labels H1, H2, H3, and H4 indicate which curves confirm which hypotheses.}\label{fig:V1_HT}}
-{\includegraphics[scale=0.6]{figures/fig3/fig3.pdf}}
-\end{figure}
-We evaluate these hypotheses by taking perturbations in individual neuron-type input $\delta h_\alpha$ away from the modes of the inferred distributions at $y=0.1$
-\begin{equation}
-dh^* = z^* = \argmax_{z} \log q_{\bm{\theta}}(z \mid \mathcal{B}(\alpha, 0.1)).
-\end{equation}
-Here $\delta x_{\alpha}$ is the change in steady state response of the system with input $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ compared to $h = b + dh^*$, where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$.
-The EPI-generated  hypotheses are confirmed (for details, see Section \ref{methods_V1}):
-
-{\addtolength{\leftskip}{10 mm}
-H1: the neuron-type responses are sensitive to their direct inputs (Fig. \ref{fig:V1_HT}A black, \ref{fig:V1_HT}B blue, \ref{fig:V1_HT}C red, \ref{fig:V1_HT}D green); \\
-H2: the E- and P-populations are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
-H3: the S-population is not affected by $\delta h_P$ (Fig. \ref{fig:V1_HT}C blue); \\
-H4: the V-population exhibits a nonmonotonic response to $\delta h_E$ (Fig. \ref{fig:V1_HT}D black), and is in fact the only population to do so (Fig. \ref{fig:V1_HT}A-C black).
-
-}
-
-These hypotheses were in stark contrast to what was available to us via traditional analytical linear prediction (Fig. \ref{fig:V1_EPI}C, magenta, see Section \ref{methods_V1}).
-
-Here, we examined the neuron-type responsivity of this model of V1 with scientifically motivated choice of connectivity $W$.
-With EPI, we could just as easily have examined the distribution of such $W$'s consistent with some response characteristics for a fixed input $h$ or another emergent property such as inhibition stabilization.
-Most importantly, this analysis is a proof-of-concept demonstrating the valuable ability to condition parameters of interest of a neural circuit model on some chosen emergent property.
-To this point, we have shown the utility of EPI on relatively low-level emergent properties like network syncing and differential neuron-type population responses.  
-In the remainder of the study, we focus on using EPI to understand models of more abstract cognitive function.
+Here, we use EPI to characterize the properties of slow noise in a stochastic version of this model, which result in biologically realistic responses.
+
+We considered the contrast response of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via nonlinearity $\phi = \left[\right]^2_+$ (see Section \ref{methods_V1}).
+The circuit model evolves form an initial condition $\mathbf{x}(0) \sim \mathcal{U}(\left[10, 25\right])$ with time constant $\tau = 1$ms according to a contrast-dependent input $\mathbf{h}$ and  slow noise $\bm{\epsilon}$ of time constant $\tau_{\text{noise}} = 5$ms.
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical}  generalized to have inhibitory multiplicity 
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}).
+\end{equation}
+As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h}_b$ via $\mathbf{h}_c$
+\begin{equation}
+\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c,
+\end{equation}
+where $h_{c,E}, h_{c,P}  > 0$ and $h_{c,S}, h_{c,V}  = 0$.
+In this analysis, we fixed $W$,$\mathbf{h}_b$, and $\mathbf{h}_c$ to values fit to mean contrast responses in mice with the deterministic model \cite{palmigiano2020structure} ($\bm{\epsilon}=\mathbf{0}$, Fig.
+\ref{fig:V1}B, see Section \ref{methods_V1}).
+At all contrasts, the E-population of this SSSN is unstable without recurrent inhibitory feedback.
+At 50\% contrast, only the P-population exhibits the paradoxical effect (\ref{fig:V1}C, Fig. \ref{fig:V1_1}), so the network is P-stabilized.
+
+The slow noise of the SSSN is an an Ornstein-Uhlenbeck process
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB,
+\end{equation}
+parameterized by $\sigma_\alpha$, which can be different for each neuron type,
+\begin{equation}
+\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top.
+\end{equation} 
+For this SSSN, we are interested in the parameters of slow noise that produce realistic stochastic fluctuations.
+Here, we quantify this emergent property as having an excitatory population Fano factor near 1:
+\begin{equation}\label{eq:EP_V1}
+\begin{split}
+\mathcal{X} ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 1 \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  0.125^2,
+\end{split}
+\end{equation}
+where $F_\alpha(\mathbf{x}; \mathbf{z})$ is the Fano factor of the $\alpha$-population.
+
+We ran EPI to obtain a posterior $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, where each parameter $\mathbf{z}$ produces biologically realistic levels of E-population variability (Fig.
+\ref{fig:V1}D).
+From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig. \ref{fig:V1}D, top-left), we can see that $F_E(\mathbf{x}; \mathbf{z})$ is sensitive to the combination of $\sigma_E$ and $\sigma_P$.
+In fact, the posterior obtained through EPI offers exactly how this sensitivity changes along this ridge of the posterior (Fig. \ref{fig:V1}E).
+$\sigma_S$ and $\sigma_V$ are degenerate with respect to $F_E(\mathbf{x}; \mathbf{z})$ evidenced by the uniform distribution in those dimensions of the posterior (Fig. \ref{fig:V1}D, bottom-right).
+Together, this posterior indicates a parametric manifold of degeneracy with respect to Fano factor: the ridge visualized in the $\sigma_E$-$\sigma_P$ marginal (Fig.
+\ref{fig:V1_2}) and the dimensions of $\sigma_S$ and $\sigma_V$.
+
+Greater $\sigma_E$ and $\sigma_P$ confer greater Fano factor, and the Fano factors of each neuron-type are strongly correlated across the posterior (Fig \ref{fig:V1}F), showing that Fano factor of each neuron-type can be modulated globally via $\sigma_E$ and $\sigma_P$.
+Furthermore, across the entire posterior distribution of noise parameterizations, we find that when contrast is increased above 50\%, variability is quenched for all neuron types (Fig \ref{fig:V1}G).
+In summary, we used EPI to obtain a posterior of SSSNs producing realistic Fano factors, which allowed degenerate manifold identification via sample visualization, fast sensitivity measurements via Hessian evaluation, and predictions of variability quenching.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
+\subsection{EPI identifies neural mechanisms of flexible task switching} \label{results_SC}
 %In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
@@
 \ref{fig:SC}B).
-The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
+The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
 The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
@@
 \tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
-\mathbf{x} &= f(\mathbf{u})
+\mathbf{x} &= \phi(\mathbf{u})
 \end{split}
 \end{equation}
-with white noise of variance $\epsilon^2 = 0.2^2$.
+with white noise of variance $0.2^2$.
 The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
@@
 \begin{center}
-\includegraphics[scale=0.48]{figures/fig4/fig4.pdf}
+\includegraphics[scale=0.48]{figures/fig3/fig3.pdf}
 \end{center}
 \caption{\small 
-A. Rapid task switching behavioral paradigm (see text). 
-B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
+\textbf{A}. Rapid task switching behavioral paradigm (see text). 
+\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
 Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
-C. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice u=of simulations.
-D.
+\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
+\textbf{D}.
 Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.
-Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (opto strength 0.7).
+Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).
 Shading indicates standard deviation across trials.
-E.
+\textbf{E}.
 Difference in performance of each task during inactivation.
@@
 We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
-F. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
+\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
 }
@@
 \end{equation}
-A variance of $5\%$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.
-
-We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
+A variance of $5\%^2$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.
+
+We ran EPI to obtain SC model connectivity parameters $\mathbf{z}$ producing rapid task switching (Fig. \ref{fig:SC}C).
 Some parameters were predictive of accuracy while others were not (Fig. \ref{fig:SC1}), and often had different effects on $p_P$ and $p_A$.
-To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $v_i$ for all $W$ parameterized by $z$ (Fig.
+To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
 \ref{fig:SC2}A).
@@
 \ref{fig:SC4}).
-Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation (Fig. \ref{fig:SC}F)).
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Linking RNN connectivity to error} \label{results_RNN}
-So far, each model we have studied was designed from fundamental biophysical principles, genetically- or functionally-defined neuron types.  
-At a more abstract level of modeling, recurrent neural networks (RNNs) are high-dimensional dynamical models of computation that are becoming increasingly popular in neuroscience research \cite{barak2017recurrent}.
-In theoretical neuroscience, RNN dynamics usually follow the equation
-\begin{equation}
-\frac{dx}{dt} = -x + W \phi(x) + h,
-\end{equation}
-where $x$ is the network activity, $W$ is the network connectivity, $\phi(\cdot) = \tanh(\cdot)$, and $h$ is the input to the system.
-Such RNNs are trained to do a task from a systems neuroscience experiment, and then the unit activations of the trained RNN are compared to recorded neural activity.
-Fully-connected RNNs with tens of thousands of parameters are challenging to characterize \cite{sussillo2013opening}, especially making statistical inferences about their parameterization.
-Alternatively, we considered a rank-1, $N$-neuron RNN with connectivity consisting of the sum of a random and a structured component:
-\begin{equation}
-W = g\chi + \frac{1}{N}mn^\top.
-\end{equation}
-The random component $g \chi$ has strength $g$, and random component weights are Gaussian distributed $\chi_{i,j} \sim \mathcal{N}(0, \frac{1}{N})$.  
-The structured component $\frac{1}{N}m n\top$ has entries of $m$ and $n$ drawn from Gaussian distributions $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$.
-Recent theoretical work derives the low-dimensional response properties of low-rank networks from statistical parameterizations of their connectivity, such as $z = [g, M_m, M_n]$ \cite{mastrogiuseppe2018linking}.
-We used EPI to infer the parameterizations of rank-1 RNNs solving an example task, enabling discovery of properties of connectivity that result in different types of error in the computation.
+Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation in the Anti task (Fig. \ref{fig:SC}F)).
+
+In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
+This posterior revealed the mechanisms leading to greater accuarcy in each task as well as those increasing resiliency to perturbation in the Anti task.
+Importantly, every connectivity from this inferred distribution predicts fragility and robustness of performance in the Anti and Pro tasks, respectively.
+EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well chosen model.
+
+\subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
+
+Here, we are interested in the scalability of EPI in number of parameters ($|\mathbf{z}|$).  We consider rank-2 RNN with N neurons of connectivity
+\begin{equation}
+ W = UV^\top + g\chi
+ \end{equation}
+ and dynamics
+ \begin{equation}
+ \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
+ \end{equation}
+where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.
+
+We want to learn distributions of connectivity that produce stable amplification.
+Two conditions are both necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}.
+These conditions are inequalities on $\text{real}(\lambda_1)$ and $\lambda^s_1$ the maximal real eigenvalue of $W$ and the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$, respectively.
+
+In our analysis, we seek to condition rank-2 networks of increasing size on a regime of stable amplification.  
+Networks with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will yield moderate amplification.
+ EPI can naturally condition on this emergent property
+\begin{equation}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
+\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix} .
+\end{split}
+\end{equation}
+
+In contrast, SNPE cannot condition on the variance of observations across posterior.  '
+Thus, we condition on an observation $x_0$ located at the mean of our desired emergent property.
+\begin{equation}\label{eq:SNPE_stab_amp_x0}
+x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
+\end{equation}
+ABC methods define tolerance $\epsilon$ and distance for observed data $x_0$.
+Here, we chose $\epsilon = 0.5$, an $l-2$ distance, and the same choice for $x_0$ as in Equation \ref{eq:SNPE_stab_amp_x0}.
+
+EPI is capable of scaling to higher dimensional parameter spaces than ABC and SNPE.
+EPI consistently produces the same posterior predictive distribution independent of the dimensionality.
+SMC produces a limited variety of parameters due to the nature of its proposal generation algorithm, yet all parameters obtained produce stable amplification.
+SNPE's posterior predictive distribution is not necessarily close to the conditioning point, and is very dependent on dimensionality.
 
@@
 \begin{center}
-\includegraphics[scale=0.75]{figures/fig5/fig5.pdf}
+\includegraphics[scale=1.4]{figures/fig4/fig4.pdf}
 \end{center}
-\caption{\small Sources of error in an RNN solving a simple task.  A. (left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$. 
-(right) Error in this computation can come from over- or under-estimating the posterior mean or variance. 
-B.
+\caption{\small 
+\textbf{A}. Recurrent neural network.
+\textbf{B}. EPI scales with $z$ to high dimensions. 
+Convergence definitions: 
+EPI (blue) - satisfies all moment constraints, 
+SNPE (orange)- produces at least 2/$n_{\text{train}}$ parameter samples are in the bounds of emergent property (mean +- 0.5), 
+and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.
+\textbf{C}. Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
+Gray star indicates emergent property mean, and gray dashed lines indicate two standard deviations corresponding to the variance constraint.
+For $N <= 6$ where SMC-ABC converges, samples are not diverse (path degeneracies).  
+For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
+\textbf{D}. Simulations of network parameters resulting from each method ($\tau=100ms$).  Each trace corresponds to simulation of one $z$.  
+\textbf{E}. Ratio of obtained samples producing stable amplification.
+}
+\label{fig:LRRNN}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Linking RNN connectivity to error} \label{results_RNN}
+%So far, each model we have studied was designed from fundamental biophysical principles, genetically- or functionally-defined neuron types.  
+%At a more abstract level of modeling, recurrent neural networks (RNNs) are high-dimensional dynamical models of computation that are becoming increasingly popular in neuroscience research \cite{barak2017recurrent}.
+%In theoretical neuroscience, RNN dynamics usually follow the equation
+%\begin{equation}
+%\frac{dx}{dt} = -x + W \phi(x) + h,
+%\end{equation}
+%where $x$ is the network activity, $W$ is the network connectivity, $\phi(\cdot) = \tanh(\cdot)$, and $h$ is the input to the system.
+%Such RNNs are trained to do a task from a systems neuroscience experiment, and then the unit activations of the trained RNN are compared to recorded neural activity.
+%Fully-connected RNNs with tens of thousands of parameters are challenging to characterize \cite{sussillo2013opening}, especially making statistical inferences about their parameterization.
+%Alternatively, we considered a rank-1, $N$-neuron RNN with connectivity consisting of the sum of a random and a structured component:
+%\begin{equation}
+%W = g\chi + \frac{1}{N}mn^\top.
+%\end{equation}
+%The random component $g \chi$ has strength $g$, and random component weights are Gaussian distributed $\chi_{i,j} \sim \mathcal{N}(0, \frac{1}{N})$.  
+%The structured component $\frac{1}{N}m n\top$ has entries of $m$ and $n$ drawn from Gaussian distributions $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$.
+%Recent theoretical work derives the low-dimensional response properties of low-rank networks from statistical parameterizations of their connectivity, such as $z = [g, M_m, M_n]$ \cite{mastrogiuseppe2018linking}.
+%We used EPI to infer the parameterizations of rank-1 RNNs solving an example task, enabling discovery of properties of connectivity that result in different types of error in the computation.
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.75]{figures/fig5/fig5.pdf}
+%\end{center}
+%\caption{\small Sources of error in an RNN solving a simple task.  A. (left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$. 
+%(right) Error in this computation can come from over- or under-estimating the posterior mean or variance. 
+%B.
 EPI distribution of rank-1 RNNs executing Gaussian posterior conditioning.
 Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$
-C.  Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.  
-Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.  
-D. The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.  
-Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
-\label{fig:RNN}
-\end{figure}
-
-The task we consider is Gaussian posterior conditioning: calculate the parameters of a posterior distribution induced by a prior $p(\mu_y) = \mathcal{N}(\mu_0=4, \sigma_0^2=1)$ and a likelihood $p(y | \mu_y) = \mathcal{N}(\mu_y,\sigma^2_y = 1)$, given a single observation $y$.
-Conjugacy offers the result analytically; $p(\mu_y | y ) = \mathcal{N}\left(\mu_{post},\sigma_{post}^2\right)$, where:   
-\begin{equation}
-\mu_{\text{post}} = \frac{\frac{\mu_0}{\sigma_0^2} + \frac{y}{\sigma_y^2}}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}
-~~~~~~~~~~
- \sigma^2_{\text{post}} = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}.
-\end{equation} 
-To solve this Gaussian posterior conditioning task, the RNN response to a constant input $h = y r + (n-M_n)$ must equal the posterior mean along readout vector $r$ , where
-\begin{equation}
-\kappa_r =  \frac{1}{N} \sum_{j=1}^N r_j \phi(x_j).
-\end{equation}
-Additionally, the amount of chaotic variance $\Delta_T$ must equal the posterior variance.
-Theory for low-rank RNNs allows us to express $\kappa_r$ and $\Delta_T$ in terms of each other through a solvable system of nonlinear equations  (see Section \ref{methods_LRRNN}) \cite{mastrogiuseppe2018linking}.
-This theory facilitates the mathematical formalization of task execution into an emergent property, where the emergent property statistics of the RNN activity are $\kappa_r$ and $\Delta_T$, and the emergent property values are the ground truth posterior mean $\mu_{\text{post}}$ and variance $\sigma^2_{\text{post}}$:
-\begin{equation}
-\mathbb{E} \begin{bmatrix} \kappa_r \\ \Delta_T \\ (\kappa_r-\mu_{\text{post}})^2 \\ (\Delta_T^2-\sigma^2_{\text{post}})^2 \end{bmatrix} ~~=~~ \begin{bmatrix} \mu_{\text{post}} \\ \sigma^2_{\text{post}} \\ 0.1 \\ 0.1 \end{bmatrix}.
-\end{equation}
-We chose a substantial amount of variance in these emergent property statistics, so that the inferred distribution resulted in RNNs with a variety of errors in their solutions to the gaussian posterior conditioning problem.
-
-EPI was used to learn distributions of RNN connectivity properties $z = [g, M_m, M_n]$ executing Gaussian posterior conditioning given an input of $y=2$, where the true posterior is $\mu_{\text{post}}=3$ and $\sigma_{\text{post}} = 0.5$ (Fig.
+%C.  Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.  
+%Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.  
+%D. The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.  
+%Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
+%\label{fig:RNN}
+%\end{figure}
+%
+%The task we consider is Gaussian posterior conditioning: calculate the parameters of a posterior distribution induced by a prior $p(\mu_y) = \mathcal{N}(\mu_0=4, \sigma_0^2=1)$ and a likelihood $p(y | \mu_y) = \mathcal{N}(\mu_y,\sigma^2_y = 1)$, given a single observation $y$.
+%Conjugacy offers the result analytically; $p(\mu_y | y ) = \mathcal{N}\left(\mu_{post},\sigma_{post}^2\right)$, where:   
+%\begin{equation}
+%\mu_{\text{post}} = \frac{\frac{\mu_0}{\sigma_0^2} + \frac{y}{\sigma_y^2}}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}
+%~~~~~~~~~~
+% \sigma^2_{\text{post}} = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}.
+%\end{equation} 
+%To solve this Gaussian posterior conditioning task, the RNN response to a constant input $h = y r + (n-M_n)$ must equal the posterior mean along readout vector $r$ , where
+%\begin{equation}
+%\kappa_r =  \frac{1}{N} \sum_{j=1}^N r_j \phi(x_j).
+%\end{equation}
+%Additionally, the amount of chaotic variance $\Delta_T$ must equal the posterior variance.
+%Theory for low-rank RNNs allows us to express $\kappa_r$ and $\Delta_T$ in terms of each other through a solvable system of nonlinear equations  (see Section \ref{methods_LRRNN}) \cite{mastrogiuseppe2018linking}.
+%This theory facilitates the mathematical formalization of task execution into an emergent property, where the emergent property statistics of the RNN activity are $\kappa_r$ and $\Delta_T$, and the emergent property values are the ground truth posterior mean $\mu_{\text{post}}$ and variance $\sigma^2_{\text{post}}$:
+%\begin{equation}
+%\mathbb{E} \begin{bmatrix} \kappa_r \\ \Delta_T \\ (\kappa_r-\mu_{\text{post}})^2 \\ (\Delta_T^2-\sigma^2_{\text{post}})^2 \end{bmatrix} ~~=~~ \begin{bmatrix} \mu_{\text{post}} \\ \sigma^2_{\text{post}} \\ 0.1 \\ 0.1 \end{bmatrix}.
+%\end{equation}
+%We chose a substantial amount of variance in these emergent property statistics, so that the inferred distribution resulted in RNNs with a variety of errors in their solutions to the gaussian posterior conditioning problem.
+%
+%EPI was used to learn distributions of RNN connectivity properties $z = [g, M_m, M_n]$ executing Gaussian posterior conditioning given an input of $y=2$, where the true posterior is $\mu_{\text{post}}=3$ and $\sigma_{\text{post}} = 0.5$ (Fig.
 \ref{fig:RNN}A).
-We examined the nature of the over- and under-estimation of the posterior means (Fig. \ref{fig:RNN}B left) and variances (Fig. \ref{fig:RNN}B right) in the inferred distributions (300 samples).  
-The symmetry in the $M_m$-$M_n$ plane, suggests a degeneracy in the product of $M_m$ and $M_n$ (Fig. \ref{fig:RNN}B).  
-Indeed, $M_m M_n$ strongly determines the posterior mean ($r=0.62$,  $p<10^{-4}$). 
-Furthermore, the random strength $g$ strongly determines the chaotic variance ($r=0.56$,  $p<10^{-4}$).
-Neither of these observations were obvious from what mathematical analysis is available in networks of this type (see Section \ref{methods_LRRNN}).  
-While the link between random strength $g$ and chaotic variance $\Delta_T$ (and resultingly posterior variance in this problem) is well-known \cite{sompolinsky1988chaos}, the distribution admits a novel hypothesis: the estimation of the posterior mean by the RNN increases with $M_m M_n$.
-
-We tested this prediction by taking parameters $z_1$ and $z_2$ as representative samples from the positive and negative $M_m$-$M_n$ quadrants, respectively.
-Instead of using the theoretical predictions shown in Figure \ref{fig:RNN}B, we simulated finite-size realizations of these networks with 2,000 neurons (e.g. Fig. \ref{fig:RNN}C).  
-We perturbed these parameter choices by $M_m M_n$ clarifying that the posterior mean can be directly controlled in this way (Fig. \ref{fig:RNN}D; $p<10^{-4}$), see Section \ref{methods_LRRNN}).
-Thus, EPI confers a clear picture of error in this computation: the product of the low rank vector means $M_m$ and $M_n$ modulates the estimated posterior mean while the random strength $g$ modulates the estimated posterior variance.
-This novel procedure of inference on reduced parameterizations of RNNs conditioned on the emergent property of task execution is generalizable to other settings modeled in \cite{mastrogiuseppe2018linking} like noisy integration and context-dependent decision making (Fig.
+%We examined the nature of the over- and under-estimation of the posterior means (Fig. \ref{fig:RNN}B left) and variances (Fig. \ref{fig:RNN}B right) in the inferred distributions (300 samples).  
+%The symmetry in the $M_m$-$M_n$ plane, suggests a degeneracy in the product of $M_m$ and $M_n$ (Fig. \ref{fig:RNN}B).  
+%Indeed, $M_m M_n$ strongly determines the posterior mean ($r=0.62$,  $p<10^{-4}$). 
+%Furthermore, the random strength $g$ strongly determines the chaotic variance ($r=0.56$,  $p<10^{-4}$).
+%Neither of these observations were obvious from what mathematical analysis is available in networks of this type (see Section \ref{methods_LRRNN}).  
+%While the link between random strength $g$ and chaotic variance $\Delta_T$ (and resultingly posterior variance in this problem) is well-known \cite{sompolinsky1988chaos}, the distribution admits a novel hypothesis: the estimation of the posterior mean by the RNN increases with $M_m M_n$.
+%
+%We tested this prediction by taking parameters $z_1$ and $z_2$ as representative samples from the positive and negative $M_m$-$M_n$ quadrants, respectively.
+%Instead of using the theoretical predictions shown in Figure \ref{fig:RNN}B, we simulated finite-size realizations of these networks with 2,000 neurons (e.g. Fig. \ref{fig:RNN}C).  
+%We perturbed these parameter choices by $M_m M_n$ clarifying that the posterior mean can be directly controlled in this way (Fig. \ref{fig:RNN}D; $p<10^{-4}$), see Section \ref{methods_LRRNN}).
+%Thus, EPI confers a clear picture of error in this computation: the product of the low rank vector means $M_m$ and $M_n$ modulates the estimated posterior mean while the random strength $g$ modulates the estimated posterior variance.
+%This novel procedure of inference on reduced parameterizations of RNNs conditioned on the emergent property of task execution is generalizable to other settings modeled in \cite{mastrogiuseppe2018linking} like noisy integration and context-dependent decision making (Fig.
 S5).
-
+%
+%\section{Discussion}
+%\subsection{EPI is a general tool for theoretical neuroscience} 
+%Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
+%We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
+%We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).
+%We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
+%Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.
+%
+%In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
+%These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
+%In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
+%Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
+%Although, theoretical approaches (e.g.
+\cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
+
+%\subsection{Novel hypotheses from EPI} 
 \section{Discussion} 
-\subsection{EPI is a general tool for theoretical neuroscience} 
-Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
-We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
-We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).
-We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
-Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.
-
-In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
-These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
-In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
-Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
-Although, theoretical approaches (e.g.
-\cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
-
-\subsection{Novel hypotheses from EPI} 
+NOTE: \textbf{This is the old discussion section.  I will rewrite this based on our discussion of the rest of the draft.}
+
 In neuroscience, machine learning has primarily been used to reveal structure in large-scale neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
@@
 When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-Gaussian noise of variance $\epsilon^2 = (1 \times 10^{-12})^2$ amps makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
+Gaussian noise of variance $(1 \times 10^{-12})^2$ amps makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
 
@@
 \subsubsection{Primary visual cortex}\label{methods_V1}
-The dynamics of each neural populations average rate
-$x = [ x_E, x_P, x_S, x_V]^\top$
-are given by:
-\begin{equation}
-\tau \frac{dx}{dt} = -x + [W x+ h]_+^n.
-\end{equation}
-
-By consolidating information from many experimental datasets, Billeh et al. \cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
-\begin{equation}
-M = \begin{bmatrix} 0.36 & 0.48 & 0.31 & 0.28 \\
-1.49 & 0.68 & 0.50 & 0.18 \\
-0.86 & 0.42 & 0.15 & 0.32 \\
-1.31 & 0.41 & 0.52 & 0.37 \end{bmatrix}
-\end{equation}
-and connection probability
-\begin{equation}
-C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
-0.395 & .451 & 0.857 & 0.02 \\
-0.182 & 0.03 & 0.082 & 0.625 \\
-0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
-\end{equation}
-
-Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
-\begin{equation}
-W_{\text{full}} = C \odot M = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
-0.395 & .451 & 0.857 & 0.02 \\
-0.182 & 0.03 & 0.082 & 0.625 \\
-0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
-\end{equation}
-
-Theoretical work on these systems considers a subset of the effective connectivities \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}
-\begin{equation} \label{eq:V1_W}
-W = \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & 0 \\
-                                W_{PE} & W_{PP} & W_{PS} & 0 \\
-                                W_{SE} & 0 & 0 & W_{SV} \\
-                                W_{VE} & W_{VP} &  W_{VS} &  0 \end{bmatrix}.
-\end{equation}
-In coherence with this work, we only keep the entries of $W_{\text{full}}$ corresponding to parameters in Equation \ref{eq:V1_W}.
-
-We look at how this four-dimensional nonlinear dynamical model of V1 responds to different inputs, and compare the predictions of the linear response to the approximate posteriors obtained through EPI.
-The input to the system is the sum of a baseline input $b = [1, 1, 1, 1]^\top$ and a differential input $dh$:
-\begin{equation}
-h = b + dh.
-\end{equation}
-All simulations of this system had $T=100$ time points, a time step $dt = 5$ms, and time constant $\tau = 20$ms.  
-The system was initialized to a random draw $x(0)_i \sim \mathcal{N}(1, 0.01)$.
-
-We can describe the dynamics of this system more generally by
-\begin{equation}
-\dot{x}_i = -x_i + f(u_i)
-\end{equation}
-where the input to each neuron is
-\begin{equation}
-u_i = \sum_j W_{ij} x_j + h_i.
-\end{equation}
-Let $F_{ij} = \gamma_i \delta(i,j)$, where $\gamma_i = f'(u_i)$.  
-Then, the linear response is
-\begin{equation}
-\frac{dx_{ss}}{dh} = F(W\frac{dx_{ss}}{dh} + I)
-\end{equation}
-which is calculable by
-\begin{equation}
-\frac{dx_{ss}}{dh} = (F^{-1} - W)^{-1}.
-\end{equation}
-This calculation is used to produce the magenta lines in Figure \ref{fig:V1_EPI}C, which show the linearly predicted inputs that generate a response from two standard deviations (of $\mathcal{B}$) below and above $y$.
-
-The emergent property we considered was the first and second moments of the change in steady state rate $dx_{ss}$ between the baseline input $h= b$ and $h = b + dh$.  
-We use the following notation to indicate that the emergent property statistics were set to the following values:
-\begin{equation}
-\mathcal{B}(\alpha, y) ~~\triangleq~~ 
-\mathbb{E} \begin{bmatrix} dx_{\alpha,ss} \\ (dx_{\alpha,ss} - y)^2 \end{bmatrix} = \begin{bmatrix} y \\ 0.01^2 \end{bmatrix}.
-\end{equation}
-
-In the final analysis for this model, we sweep the input one neuron at a time away from the mode of each inferred distributions $dh^* = \mathbf{z}^* = \argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{B}(\alpha, 0.1)$.
-The differential responses $\delta x_{\alpha,ss}$ are examined at perturbed inputs  $h = b + dh^* + \delta h_\alpha \hat{u}_\alpha$ where $\hat{u}_\alpha$ is a unit vector in the dimension of $\alpha$ and $\delta x$ is evaluated at 101 equally spaced samples of $\delta h_\alpha$ from -15 to 15.
-
-We measured the linear regression slope between neuron-types of $\delta x$ and $\delta h$ to confirm the hypotheses H1-H3 (H4 is simply observing the nonmonotonicity) and report the p values for tests of non-zero slope.
-
-{\addtolength{\leftskip}{10 mm}
-H1: the neuron-type ressponses are sensitive to their direct inputs. 
-E-population: $\beta=1.62$, $p<10^{-4}$ (Fig. \ref{fig:V1_HT}A black),
-P-population: $\beta=1.06$, $p<10^{-4}$ (Fig.
-\ref{fig:V1_HT}B blue), S-population: $\beta=6.80$, $p<10^{-4}$  (Fig.
-\ref{fig:V1_HT}C red), V-population: $\beta=6.41$, $p<10^{-4}$  (Fig.
-\ref{fig:V1_HT}D green).
-\\
-H2: the E-population ($\beta=0$, $p=1$) and P-populations ($\beta=0$, $p=1$) are not affected by $\delta h_V$ (Fig. \ref{fig:V1_HT}A green, \ref{fig:V1_HT}B green); \\
-H3: the S-population is not affected by $\delta h_P$ ($\beta=0$, $p=1$) (Fig. \ref{fig:V1_HT}C blue); \\
-
+Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
+
+\begin{equation}
+W_{\text{fit}} =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
+W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
+W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
+W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
+ \begin{bmatrix} 2.18 & -1.19 & -.594 & -.229 \\
+ 1.66 & -.651 & -.680 & -.242 \\
+ .895 & -5.22 \times 10^{-3} & -1.51 \times 10^{-4}   & -.761 \\
+ 3.34 &  -2.31 & -.254  & -2.52 \times 10^{-4} \\
+ \end{bmatrix},
+\end{equation}
+
+\begin{equation}
+\mathbf{h}_{b,\text{fit}} =
+ \begin{bmatrix} .416 \\ .429 \\ .491 \\ .486 \end{bmatrix} ,
+\end{equation} 
+and
+\begin{equation} 
+\mathbf{h}_{c,\text{fit}} = 
+\begin{bmatrix} .359 \\ .403 \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+To obtain rates on a realistic scale (100-fold greater), we map these fitted parameters to an equivalence class
+
+\begin{equation}
+W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
+W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
+W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
+W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
+ \begin{bmatrix} .218 & -.119 & -.0594 & -.0229 \\
+ .166 & -.0651 & -.068 & -.0242 \\
+ .0895 & -5.22 \times 10^{-4} & -1.51 \times 10^{-5}   & -.0761 \\
+ .334 &  -.231 & -.0254  & -2.52 \times 10^{-5} \\
+ \end{bmatrix},
+\end{equation}
+
+\begin{equation}
+\mathbf{h}_b =  \begin{bmatrix} h_{b,E} \\ h_{b,P} \\ h_{b,S} \\ h_{b,V} \end{bmatrix} =
+ \begin{bmatrix} 4.16 \\ 4.29 \\ 4.91 \\ 4.86 \end{bmatrix} ,
+\end{equation} 
+and
+\begin{equation} 
+\mathbf{h}_c = \begin{bmatrix} h_{c,E} \\ h_{c,P} \\ h_{c,S} \\ h_{c,V} \end{bmatrix} = 
+\begin{bmatrix} 3.59 \\ 4.03 \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+Since the E-population of this network increases exponentially in the abscense of  recurrent inhibitory feedback, we may also observe a paradoxical effect in the inhibitory populations (which is present in E-I networks).
+At 50\% contrast (Fig. \ref{fig:V1}B, dots), this network exhibits a paradoxical effect in the P-population (Fig. \ref{fig:V1}C), but no others (Fig. \ref{fig:V1_1}).
+That is, for a small increase in $h_P$, $\mathbb{E}_t\left[x_P\right]$ decreases.
+
+Fano factor is calculated as the temporal variance divided by the temporal mean following some time $t_{ss}$ following dynamical evolution from the initial state at $\mathbf{x}(t=0)$.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=.5]{figures/figV1_1/figV1_1.pdf}
+\end{center}
+\caption{\small Supplemental Figure: 
+(Left) Simulations for small increases in neuron-type population input.
+Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
+(Right) Average and standard deviation of stochastic fluctuations of responses.
  }
-
-For each $\mathcal{B}(\alpha, y)$ with $\alpha \in \{E, P, S, V\}$ and $y \in \{0.1, 0.5\}$, we ran EPI using a real NVP architecture of four masks layers with two hidden layers of 10 units,  mapped to a support of $z_i \in \left[-5, 5\right]$ with no batch normalization.
-We used an augmented Lagrangian coefficient of $c_0 = 10^5$, a batch size $n=1000$, set $\nu = 0.5$.
-The EPI distributions shown in Fig. \ref{fig:V1_EPI} are the converged distributions with maximum entropy across random seeds.
-
-We set the parameters of the Gaussian initialization $\mu_\text{init}$ and $\Sigma_\text{init}$ to the mean and covariance of random samples $z^{(i)} \sim \mathcal{U}(-5, 5)$ that produced emergent property statistic $dx_{\alpha,ss}$ within a bound $\epsilon$ of the emergent property value $y$.
-$\epsilon=0.01$ was set to be one standard deviation of the emergent property value according to the emergent property value $0.01^2$ of the variance emergent property statistic.
+ \label{fig:V1_1}
+\end{figure}
+
+\begin{figure}[h]
+\caption{\small Supplemental Figure: 
+Fano factors along the ridge of the posterior in Fig. \ref{fig:V1}E.
+ }
+ \label{fig:V1_2}
+\begin{center}
+\includegraphics[scale=.5]{figures/figV1_2/figV1_2.pdf}
+\end{center}
+\end{figure}
 
@@
 \begin{equation}
-x_\alpha = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
+x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
 \end{equation}
 where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
+During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
+\begin{equation}
+x_\alpha = (1-\gamma)\phi(u_\alpha).
+\end{equation}
 
@@
 \begin{equation}
-\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + \sigma d\mathbf{B}.
-\end{equation}
-with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $dB$of variance $\sigma^2=0.2$.  
+\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
+\end{equation}
+with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2$.  
 The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
@@
 
-For EPI in Fig \label{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
+For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
 The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
@@
 
-\subsubsection{Rank-1 RNN}\label{methods_LRRNN}
-Extensive research on random fully-connected recurrent neural networks has resulted in foundational theories of their activity \cite{sompolinsky1988chaos, brunel2000dynamics}.
-Furthermore, independent research on training these models to perform computations suggests that learning occurs through low-rank perturbations to the connectivity (e.g.
+\subsubsection{Rank-2 RNN}\label{methods_RNN}
+Traditional approaches to likelihood-free inference -- approximate Bayesian computation (ABC) methods -- randomly sample parameters $\mathbf{z}$ until a suitable set is obtained.
+State-of-the-art ABC methods leverage sequential monte-carlo (SMC) sampling techniques to obtain parameter sets more efficiently.
+To obtain more parameter samples, SMC-ABC must be run from scratch again.
+ABC methods do not confer log probabilities of samples.
+ Like EPI, sequential neural posterior estimation (SNPE) uses deep learning to produce flexible posterior approximations.
+Like traditional Bayesian inference methods, SNPE conditions directly on the statistics of data.
+This differs from EPI, where posteriors are conditioned on emergent properties (moment constraints on the posterior predictive distribution).
+Peculiarities of SNPE (density estimation approach, two deep networks) make scaling in $\mathbf{z}$ prohibitive.
+
+SMC-ABC has many hyperparameters, of which pyABC selects automatically by running some initial diagnostics upon initialization.
+In concurrence with the literacture, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
+We searched over many SNPE hyperparameter choices: $n_{\text{train}} \in [2,000, 10,000, 100,000]$ is the number of simulations run per training epoch, and $n_{\text{mades}} \in [2, 3]$ is the number of masked autoregressive density estimators in the deep parameter distribution architecture.
+The greater $n_{\text{train}}$, the longer each epoch will take, but the more likely SNPE may converge during that epoch.
+Greater $n_{\text{mades}}$ increases the flexibility of the deep parameter distribution of SNPE, but slows optimization.
+For the timing plot, we show the fastest among all of these choices, and for the convergence plot, we show the best convergence among all of these choices.
+During optimization, we used $n_{\text{atom}}$=100 atomic proposals as is recommended.
+
+%\subsubsection{Rank-1 RNN}\label{methods_LRRNN}
+%Extensive research on random fully-connected recurrent neural networks has resulted in foundational theories of their activity \cite{sompolinsky1988chaos, brunel2000dynamics}.
+%Furthermore, independent research on training these models to perform computations suggests that learning occurs through low-rank perturbations to the connectivity (e.g.
 \cite{jaeger2004harnessing, sussillo2009generating}).
-Recent theoretical work extends theory for random neural networks \cite{sompolinsky1988chaos} to those with added low-rank structure \cite{mastrogiuseppe2018linking}.
-In Section \ref{results_RNN}, we used this theory to enable EPI on RNN parameters conditioned on the emergent property of task execution.
-
-Such RNNs have the following dynamics: 
-\begin{equation}
-\frac{dx}{dt} = -x + W \phi(x) + h,
-\end{equation}
-where $x$ is network activity, $W$ is the connectivity weight matrix, $\phi(\cdot) = \tanh(\cdot)$ is the input-output function, and $h$ is the input to the system.
-In a rank-1 RNN (which was sufficiently complex for the Gaussian posterior conditioning task), $W$ is the sum of a random component with strength $g$ and a structured component determined by the outer product of vectors $m$ and $n$:
-\begin{equation}
-W = g\chi + \frac{1}{N}mn^\top,
-\end{equation}
-where  $\chi_{ij} \sim \mathcal{N}(0, \frac{1}{N})$, and the entries of $m$ and $n$ are distributed as $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$. 
-For EPI, we consider $z = [g, M_m, M_n]$, which are the parameters governing the connectivity properties of the RNN.
-
-From such a parameterization $z$, the theory of Mastrogiuseppe et al. produces solutions for variables describing the low dimensional response properties of the RNN.
-These ``dynamic mean field" (DMF) variables (e.g.
+%Recent theoretical work extends theory for random neural networks \cite{sompolinsky1988chaos} to those with added low-rank structure \cite{mastrogiuseppe2018linking}.
+%In Section \ref{results_RNN}, we used this theory to enable EPI on RNN parameters conditioned on the emergent property of task execution.
+%
+%Such RNNs have the following dynamics: 
+%\begin{equation}
+%\frac{dx}{dt} = -x + W \phi(x) + h,
+%\end{equation}
+%where $x$ is network activity, $W$ is the connectivity weight matrix, $\phi(\cdot) = \tanh(\cdot)$ is the input-output function, and $h$ is the input to the system.
+%In a rank-1 RNN (which was sufficiently complex for the Gaussian posterior conditioning task), $W$ is the sum of a random component with strength $g$ and a structured component determined by the outer product of vectors $m$ and $n$:
+%\begin{equation}
+%W = g\chi + \frac{1}{N}mn^\top,
+%\end{equation}
+%where  $\chi_{ij} \sim \mathcal{N}(0, \frac{1}{N})$, and the entries of $m$ and $n$ are distributed as $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$. 
+%For EPI, we consider $z = [g, M_m, M_n]$, which are the parameters governing the connectivity properties of the RNN.
+%
+%From such a parameterization $z$, the theory of Mastrogiuseppe et al. produces solutions for variables describing the low dimensional response properties of the RNN.
+%These ``dynamic mean field" (DMF) variables (e.g.
 the activity along a vector $\kappa_v$, the total variance $\Delta_0$, structured variance $\Delta_\infty$, and the chaotic variance $\Delta_T$) are derived to be functions of one another and connectivity parameters $z$.
-The collection of these derived functions results in a system of equations, whose solution must be obtained through a nonlinear system of equations solver.
-The iterative steps of this system of equations solver are differentiable, so we take gradients through this solve process.
-The DMF variables provide task-relevant information about the RNN's response to task inputs.
-
-In the Gaussian posterior conditioning example, $\kappa_r$ and $\Delta_T$ are DMF variables used as task-relevant emergent property statistics $\mu_{\text{post}}$ and $\sigma^2_{\text{post}}$.
-Specifically, we solve for the DMF variables $\kappa_r$, $\kappa_n$, $\Delta_0$ and $\Delta_\infty$, where the readout is nominally chosen to point in the unit orthant $r = [1, ..., 1]^\top$.  
-The consistency equations for these variables in the presence of a constant input $h = y r - (n - M_n)$ can be derived following \cite{mastrogiuseppe2018linking}:
-{\small
-\begin{equation}
-\begin{split}
-\kappa_r = G_1(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_m \kappa_n + y \\
-\kappa_n = G_2(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_n \langle \left[ \phi_i \right] \rangle + \langle \left[ \phi_i' \right] \rangle \\
-\frac{\Delta_0^2-\Delta_\infty^2}{2} = G_3(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = g^2 \left( \int \mathcal{D}z \Phi^2(\kappa_r + \sqrt{\Delta_0}z) - \int \mathcal{D}z \int \mathcal{D}x \Phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z)  \right) \\
-+ (\kappa_n^2  + 1)(\Delta_0-\Delta_\infty) \\
-\Delta_\infty = G_4(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty)  = g^2 \int \mathcal{D}z \left[ \int \mathcal{D}x \phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z \right]^2 + \kappa_n^2 + 1
-\end{split} 
-\end{equation}
-}
-where here $z$ is a gaussian integration variable. 
-We can solve these equations by simulating the following Langevin dynamical system to a steady state:
-
-\begin{equation}
-\begin{split}
-l(t) = \frac{\Delta_0(t)^2-\Delta_\infty(t)^2}{2} \\
-\Delta_0(t) = \sqrt{2l(t) + \Delta_\infty(t)^2} \\
-\frac{d \kappa_r(t)}{dt}  = -\kappa_r(t) + G_1(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d \kappa_n(t)}{dt} = -\kappa_n + G_2(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d l(t)}{dt}  = -l(t) + G_3(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-\frac{d \Delta_\infty(t)}{dt}  = -\Delta_\infty(t) + G_4(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t))
-\end{split}.
-\end{equation}
-Then, the chaotic variance, which is necessary for the Gaussian posterior conditioning example, is simply calculated via $\Delta_T = \Delta_0 - \Delta_\infty$.
-
-We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization.
-We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
-The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.
-
-To examine the effect of product $M_m M_n$ on the posterior mean, $\mu_{\text{post}}$ we took perturbations in $M_m M_n$ away from two representative parameters  $z_1$ and $z_2$ in 21 equally space increments from -1 to 1.
-For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.  
-In Fig. \ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.  
-The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.  
-For perturbations away from $z_1$ the correlation was 0.995 with $p<10^{-4}$, and for perturbations away from $z_2$ the correlation was 0.983 with $p<10^{-4}$.
-
-In addition to the Gaussian posterior conditioning example in Section \ref{results_RNN}, we modeled two tasks from Mastrogiuseppe et al.: noisy detection and context-dependent discrimination.  
-We used the same theoretical equations and task setups described in their study.
-
-%TODO Need to explain the warm starting for the aficionados.
-
-%TODO explain the density network architectures used.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
-%\end{center}
-%{\small Fig.
-S5: A.
-EPI for rank-1 networks doing noisy discrimination.
-B.
-EPI for rank-2 networks doing context-dependent discrimination.
-See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
-%\end{figure}
-
+%The collection of these derived functions results in a system of equations, whose solution must be obtained through a nonlinear system of equations solver.
+%The iterative steps of this system of equations solver are differentiable, so we take gradients through this solve process.
+%The DMF variables provide task-relevant information about the RNN's response to task inputs.
+%
+%In the Gaussian posterior conditioning example, $\kappa_r$ and $\Delta_T$ are DMF variables used as task-relevant emergent property statistics $\mu_{\text{post}}$ and $\sigma^2_{\text{post}}$.
+%Specifically, we solve for the DMF variables $\kappa_r$, $\kappa_n$, $\Delta_0$ and $\Delta_\infty$, where the readout is nominally chosen to point in the unit orthant $r = [1, ..., 1]^\top$.  
+%The consistency equations for these variables in the presence of a constant input $h = y r - (n - M_n)$ can be derived following \cite{mastrogiuseppe2018linking}:
+%{\small
+%\begin{equation}
+%\begin{split}
+%\kappa_r = G_1(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_m \kappa_n + y \\
+%\kappa_n = G_2(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_n \langle \left[ \phi_i \right] \rangle + \langle \left[ \phi_i' \right] \rangle \\
+%\frac{\Delta_0^2-\Delta_\infty^2}{2} = G_3(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = g^2 \left( \int \mathcal{D}z \Phi^2(\kappa_r + \sqrt{\Delta_0}z) - \int \mathcal{D}z \int \mathcal{D}x \Phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z)  \right) \\
+%+ (\kappa_n^2  + 1)(\Delta_0-\Delta_\infty) \\
+%\Delta_\infty = G_4(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty)  = g^2 \int \mathcal{D}z \left[ \int \mathcal{D}x \phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z \right]^2 + \kappa_n^2 + 1
+%\end{split} 
+%\end{equation}
+%}
+%where here $z$ is a gaussian integration variable. 
+%We can solve these equations by simulating the following Langevin dynamical system to a steady state:
+%
+%\begin{equation}
+%\begin{split}
+%l(t) = \frac{\Delta_0(t)^2-\Delta_\infty(t)^2}{2} \\
+%\Delta_0(t) = \sqrt{2l(t) + \Delta_\infty(t)^2} \\
+%\frac{d \kappa_r(t)}{dt}  = -\kappa_r(t) + G_1(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
+%\frac{d \kappa_n(t)}{dt} = -\kappa_n + G_2(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
+%\frac{d l(t)}{dt}  = -l(t) + G_3(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
+%\frac{d \Delta_\infty(t)}{dt}  = -\Delta_\infty(t) + G_4(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t))
+%\end{split}.
+%\end{equation}
+%Then, the chaotic variance, which is necessary for the Gaussian posterior conditioning example, is simply calculated via $\Delta_T = \Delta_0 - \Delta_\infty$.
+%
+%We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization.
+%We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
+%The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.
+%
+%To examine the effect of product $M_m M_n$ on the posterior mean, $\mu_{\text{post}}$ we took perturbations in $M_m M_n$ away from two representative parameters  $z_1$ and $z_2$ in 21 equally space increments from -1 to 1.
+%For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.  
+%In Fig. \ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.  
+%The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.  
+%For perturbations away from $z_1$ the correlation was 0.995 with $p<10^{-4}$, and for perturbations away from $z_2$ the correlation was 0.983 with $p<10^{-4}$.
+%
+%In addition to the Gaussian posterior conditioning example in Section \ref{results_RNN}, we modeled two tasks from Mastrogiuseppe et al.: noisy detection and context-dependent discrimination.  
+%We used the same theoretical equations and task setups described in their study.
 \end{document}

### 2021-01-06 written/eLifeRevision/Bittner_eLife2020.tex
methods re-org

@@
 %Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
-Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_exp_fam_lh}).
+Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
 A key advantage of EPI over established Bayesian inference is that the predictions made my the inferred distribution are constrained to produce the specified emergent property.
@@
 \subsection{Emergent property inference (EPI)}\label{methods_EPI}
-%Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
 %Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
-%EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{B}$,
-%\begin{equation}
-%\mathcal{B} ~~\triangleq~~  \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right] \right] = \mu.
-%\end{equation}
-%Precisely, the emergent property statistics $T(\mathbf{x})$ must equal the emergent property values $\bm{\mu}$, in expectation over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and the distribution of simulated activity $p(\mathbf{x} \mid \mathbf{z})$.
-%This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.Emergent property inference (EPI) learns distributions of theoretical model parameters that produce emergent properties of interest by combining ideas from maximum entropy flow networks (MEFNs) \cite{loaiza2017maximum} and likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}.
-Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
-EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{X}$ defined by the means and variances of emergent property statistics $f(\mathbf{x}; \mathbf{z})$
- \begin{equation}
-\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}
-Precisely, the emergent property statistics $f(\mathbf{x})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and stochasticity of the data given the parameters defined by the model $p(\mathbf{x} \mid \mathbf{z})$.
-This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
-
+%EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{X}$ defined by the means and variances of emergent property statistics $f(\mathbf{x}; \mathbf{z})$
+% \begin{equation}
+%\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+%\end{equation}
+%Precisely, the emergent property statistics $f(\mathbf{x})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and stochasticity of the data given the parameters defined by the model $p(\mathbf{x} \mid \mathbf{z})$.
+%This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
+%
+%
 %With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
@@
 %q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[T(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
+% &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
 % \end{split}
@@
 %\begin{equation}
-%H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{w \sim q_0}\left[-\log(q_{\bm{\theta}}(f_{\bm{\theta}}(w))) \right].
+%H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
 %\end{equation}
@@
 %\begin{equation}
-%\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{w \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(f_{\bm{\theta}}(w))) \right].
+%\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
 %\end{equation}
 %In EPI, MEFNs are purposed towards variational learning of model parameter distributions.
+
+Determining the combinations of model parameters that can produce observed data or a desired output is a key part of scientific practice.
+Solving inverse problems is especially important in neuroscience, since we require complex models to describe the complex phenomena of neural computations.
+While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models conditioned upon the emergent phenomena they produce.
+Here, we introduce a novel method for statistical inference, which finds distributions of parameter solutions that only produce the desired emergent property.
+This method seamlessly handles neural circuit models with stochastic nonlinear dynamical generative processes, which are predominant in theoretical neuroscience.
+
+Consider model parameterization $\mathbf{z}$, which is a collection of scientifically interesting variables that govern the complex simulation of data $\mathbf{x}$.
+For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
+In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
+Even so, we are not so much scientifically interested in reasoning about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
+In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
+With EPI, we learn distributions of $\mathbf{z}$ that results in an average and variance of $f(\mathbf{x}; \mathbf{z})$, denoted $\bm{\mu}$ and $\bm{\sigma}^2$.
+We refer to the collection of these statistical moments as an emergent property.
+Such emergent properties $\mathcal{X}$ are defined through choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), $\bm{\mu}$, and $\bm{\sigma}^2$
+ \begin{equation}
+\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\end{equation}
+Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.
+
+In EPI, deep probability distributions are used as posterior approximations $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+ In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
+$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
+\begin{equation}
+\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
+\end{equation}
+Such deep probability distributions embed the posterior distribution in a deep network.
+Once optimized, this deep network representation has remarkably useful properties: immediate posterior sampling, and immediate probability, gradient, and Hessian evaluation at any parameter choice.
+
+Given a choice of model $p(\mathbf{x} \mid \mathbf{z})$ and emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
+\begin{equation} \label{eq:opt}
+\begin{split}
+q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) = q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
+ &  \text{s.t.
+} \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\\
+ \end{split}
+\end{equation} 
+Entropy is chosen as the normative selection principle, since we want the posterior to only contain structure predicated by the emergent property \cite{jaynes1957information, elsayed2017structure}.
+This choice of selection principle is also that of standard Bayesian inference, and we derive an exact relation between EPI and variational inference (see Section \ref{methods_ME_EF}).
+However, a key difference is that variational inference and other Bayesian methods do not constrain the predictions of their inferred posteriors.
+This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
+
+In the remainder of Section \ref{methods_EPI}, we will explain the finer details and motivation of the EPI method. 
+First, we explain related approaches and what EPI introduces to this domain (Section \ref{methods_related}).
+Second, we describe the special class of deep probability distributions used in EPI called normalizing flows (Section \ref{methods_NF}).  
+Next, we explain the constrained optimization technique used to solve Equation \ref{eq:opt} (Section \ref{methods_AL_opt}).
+Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
+Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain the relation between EPI and variational inference (Section \ref{methods_VI}).
+
+ \subsubsection{Related approaches}\label{methods_related}
+Inverse problems across scientific fields have been approached in a variety of ways.
+The goals of such approaches are generally to a.) identify parametric sensitivities and manifolds of degeneracy with respect to some model output (structural identifiability analysis), and to reason about the parameters that are most probable to have produced a model output (statistical inference).
+While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.
+
+To understand sensitivity around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
+A popular efficient approach for systems of ODEs has been neural ODE adjoint \cite{chen2018neural} and its stochastic adaptation \cite{li2020scalable}.
+Casting identifiability as a statistical estimation problem, the profile likeilhood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
+An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point \cite{raman2017delineating}.
+Global structual non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
+
+MCMC approaches
+
+ABC approaches
+
+NF LFI approaches
+
 %A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
@@
 %Before going through those details, we ground this optimization in a toy example.
-
-With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
- In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
-$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-\begin{equation}
-\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
-\end{equation}
-Given a simulator defined by a theoretical model $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ and some emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
-\begin{equation} \label{eq:opt}
-\begin{split}
-q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
- &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
- \end{split}
-\end{equation} 
-Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0$:
-\begin{equation}
-H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $\mathbf{z}$:
-\begin{equation}
-\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-In EPI, MEFNs are purposed towards variational learning of model parameter distributions.
-
 % We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
@@
 %These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.
-
- \subsubsection{Related work}\label{methods_related}
-
-TODO: rewrite this whole section.
-
-A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
-Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(\mathbf{x} \mid \mathbf{z})$.
-Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
-Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
-But, unlike LFVI, EPI uses a single deep network to learn a distribution and is optimized to produce an emergent property, rather than condition on data points.  
-Optimizing the EPI objective is a technological challenge, the details of which we elaborate in Section \ref{methods_AL_opt}.  
-Before going through those details, we ground this optimization in a toy example.
- We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
-We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
-These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.
 
@@
 
+optimization with respect to entropy, and to confer immediate probability evaluations after optimization.
+
 \subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-%To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
-%\begin{equation} \label{eq:AL}
-%L(\bm{\theta}; \eta, c) = -H(q_{\bm{\theta}}) + \eta^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
-%\end{equation}
-%where $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}) - \bm{\mu} \right] \right]$, $\eta \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}| = |T(\mathbf{x})|$,  and $c$ is the penalty coefficient.
-%These Lagrange multipliers are closely related to the natural parameters of exponential families (see Section \ref{methods_VI}).
-%Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
-%$\eta$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\eta$, $c$) for a given number of stochastic optimization iterations.
-%A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
-%For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
-%The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
-%The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-%The other update rule is $\eta_{k+1} = \eta_k + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu})$ given a batch size $n$.
-%Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
+Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
+Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
+\begin{equation}
+H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
+\end{equation}
+Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
+\begin{equation}
+\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
+\end{equation}
+
 To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
@@
 In such cases, optimization will be expensive.
-
-\subsubsection{Maximum entropy distributions and exponential families}\label{methods_VI}
-%As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
-%A maximum entropy distribution of form:
-%\begin{equation} \label{eq:max_ent}
-%\begin{split}
-%p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}. \\
-% \end{split}
-%\end{equation} 
-%will have probability density in the exponential family:
-%\begin{equation}
-%p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-%\end{equation}
-%The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-Maximum entropy distributions have a fundamental link to exponential family distributions. 
-A maximum entropy distribution of form:
-\begin{equation} \label{eq:max_ent}
-\begin{split}
-p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
- &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
- \end{split}
-\end{equation} 
-will have probability density in the exponential family:
-\begin{equation}
-p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-\end{equation}
-The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}
- \begin{equation}
-\mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}
-The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
- \begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
-\end{equation}
-As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
-\begin{equation} \label{eq:moments}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
-\end{equation}
-which are constrained to the chosen means and variances
-\begin{equation} \label{eq:mu_opt}
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
-\end{equation}
-
-\subsubsection{EPI as variational inference}\label{methods_exp_fam_lh}
-%Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
-In Bayesian inference, we obtain a posterior distribution $p(z \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
-\begin{equation}
-p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
-\end{equation}
-The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
-
-%Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid \mathbf{x})$.  
-%We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(\mathbf{x})$.
-%The log normalizing constant $A(\eta(\mathbf{x}))$ remains unchanged:
-%\begin{equation}
-%\begin{split}
-%p(z \mid \mathbf{x}) = b(z) \exp{\left( \eta(\mathbf{x})^\top T(z) - A(\eta(\mathbf{x})) \right)} = \exp{\left( \begin{bmatrix} \eta(\mathbf{x}) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(\mathbf{x})) \right)} \\= \exp{\left(\tilde{\eta(\mathbf{x})}^\top \tilde{T}(z) - A(\eta(\mathbf{x})) \right)}
-%\end{split}.
-%\end{equation}
-%Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})).
-%\end{equation}
-%$q_{\bm{\theta}}(z)$ is the variational approximation to the posterior with variational parameters $\bm{\theta}$.  
-%We can write this KL divergence in terms of entropy of the variational approximation:
-%\begin{equation}
-%KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) = \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(z)) \right] - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (p(z \mid \mathbf{x})) \right]
-%\end{equation}
-%\begin{equation}
-% = -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) - A(\eta(\mathbf{x})) \right].
-%\end{equation}
-%As far as the variational optimization is concerned, the log normalizing constant is independent of $q_{\bm{\theta}}(z)$, so it can be dropped
-%\begin{equation}
-%\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) =  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) \right].
-%\end{equation}
-% Further, we can write the objective in terms of the first moment of the sufficient statistics $\bm{\mu} = \mathbb{E}_{z \sim p(z \mid \mathbf{x})}\left[T(z) \right]$:
-% \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right] + \tilde{\eta}(\mathbf{x})^\top \bm{\mu},
-% \end{equation}
-%which simplifies to
-%  \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right].
-% \end{equation}.
-%
-%In comparison, in emergent property inference (EPI), we solve the following problem:
-%\begin{equation}
-%q_{\bm{\theta}}^*(z) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(z)),   \text{  s.t.
-} \mathbb{E}_{z \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid z)}\left[T(\mathbf{x})\right] \right] = \bm{\mu}.
-%\end{equation}
-%The Lagrangian objective (without augmentation) is
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_{\bm{\theta}}} \left[\tilde{T}(z) \right] - \bm{\mu} \right).
-%\end{equation}
-%Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(\mathbf{x})$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
-%
-%We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
-%Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
-%However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-%Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
-
-In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$
-\begin{equation}
-q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
-\end{equation}
-The KL divergence can be written in terms of entropy of the variational approximation:
-\begin{equation}
-KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
-\end{equation}
-\begin{equation}
-= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
-\end{equation}
-Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
-This is usually framed as maximizing the evidence lower bound (ELBO)
-\begin{equation}
-\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
-\end{equation}
-
-Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
-\begin{equation}
-p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
-\end{equation}
-where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
-The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
-\begin{equation}
-\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
-\end{equation}
-where
-\begin{equation}
-\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
-\end{equation}
-\begin{equation}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
-\end{equation}
-The variational objective is then
-\begin{equation}
-\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
-\end{equation}
-
-Comparing this to the Lagrangian objective (without augmentation) of EPI, we see they are the same
-\begin{equation}
-\begin{split}
-q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
- &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
-\end{split}
-\end{equation}
-where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
-Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
-The data $\mathbf{x}$ used by this implicit variational inference program would be that generated by the adapting variational approximation $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z}) q_{\bm{\theta}}(\mathbf{z})$, and the likelihood parameters $\bm{\eta}_{f}$ of EPI optimization epoch $k$ are predicated by $\bm{\eta}_{\text{opt},k}$ .
-However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
 
@@
 
+\subsubsection{Maximum entropy distributions and exponential families}\label{methods_ME_EF}
+%As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
+%A maximum entropy distribution of form:
+%\begin{equation} \label{eq:max_ent}
+%\begin{split}
+%p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
+% &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}. \\
+% \end{split}
+%\end{equation} 
+%will have probability density in the exponential family:
+%\begin{equation}
+%p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
+%\end{equation}
+%The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
+Maximum entropy distributions have a fundamental link to exponential family distributions. 
+A maximum entropy distribution of form:
+\begin{equation} \label{eq:max_ent}
+\begin{split}
+p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
+ &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
+ \end{split}
+\end{equation} 
+will have probability density in the exponential family:
+\begin{equation}
+p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
+\end{equation}
+The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
+
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}
+ \begin{equation}
+\mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\end{equation}
+The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
+ \begin{equation}
+\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
+\end{equation}
+As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
+\begin{equation} \label{eq:moments}
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
+\end{equation}
+which are constrained to the chosen means and variances
+\begin{equation} \label{eq:mu_opt}
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
+\end{equation}
+
+\subsubsection{EPI as variational inference}\label{methods_VI}
+%Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
+In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
+In Bayesian inference, we obtain a posterior distribution $p(z \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+\begin{equation}
+p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
+\end{equation}
+The posterior distribution is analytically available when the prior is conjugate with the likelihood.
+However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
+
+%Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid \mathbf{x})$.  
+%We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(\mathbf{x})$.
+%The log normalizing constant $A(\eta(\mathbf{x}))$ remains unchanged:
+%\begin{equation}
+%\begin{split}
+%p(z \mid \mathbf{x}) = b(z) \exp{\left( \eta(\mathbf{x})^\top T(z) - A(\eta(\mathbf{x})) \right)} = \exp{\left( \begin{bmatrix} \eta(\mathbf{x}) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(\mathbf{x})) \right)} \\= \exp{\left(\tilde{\eta(\mathbf{x})}^\top \tilde{T}(z) - A(\eta(\mathbf{x})) \right)}
+%\end{split}.
+%\end{equation}
+%Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
+%\begin{equation}
+%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})).
+%\end{equation}
+%$q_{\bm{\theta}}(z)$ is the variational approximation to the posterior with variational parameters $\bm{\theta}$.  
+%We can write this KL divergence in terms of entropy of the variational approximation:
+%\begin{equation}
+%KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) = \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(z)) \right] - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (p(z \mid \mathbf{x})) \right]
+%\end{equation}
+%\begin{equation}
+% = -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) - A(\eta(\mathbf{x})) \right].
+%\end{equation}
+%As far as the variational optimization is concerned, the log normalizing constant is independent of $q_{\bm{\theta}}(z)$, so it can be dropped
+%\begin{equation}
+%\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) =  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) \right].
+%\end{equation}
+% Further, we can write the objective in terms of the first moment of the sufficient statistics $\bm{\mu} = \mathbb{E}_{z \sim p(z \mid \mathbf{x})}\left[T(z) \right]$:
+% \begin{equation}
+%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right] + \tilde{\eta}(\mathbf{x})^\top \bm{\mu},
+% \end{equation}
+%which simplifies to
+%  \begin{equation}
+%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right].
+% \end{equation}.
+%
+%In comparison, in emergent property inference (EPI), we solve the following problem:
+%\begin{equation}
+%q_{\bm{\theta}}^*(z) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(z)),   \text{  s.t.
+} \mathbb{E}_{z \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid z)}\left[T(\mathbf{x})\right] \right] = \bm{\mu}.
+%\end{equation}
+%The Lagrangian objective (without augmentation) is
+%\begin{equation}
+%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_{\bm{\theta}}} \left[\tilde{T}(z) \right] - \bm{\mu} \right).
+%\end{equation}
+%Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(\mathbf{x})$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
+%
+%We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
+%Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
+%However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
+%Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
+%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
+
+In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$
+\begin{equation}
+q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
+\end{equation}
+The KL divergence can be written in terms of entropy of the variational approximation:
+\begin{equation}
+KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
+\end{equation}
+\begin{equation}
+= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
+\end{equation}
+Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
+This is usually framed as maximizing the evidence lower bound (ELBO)
+\begin{equation}
+\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
+\end{equation}
+
+Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
+\begin{equation}
+p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
+\end{equation}
+where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
+The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
+\begin{equation}
+\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
+\end{equation}
+where
+\begin{equation}
+\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
+\end{equation}
+\begin{equation}
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
+\end{equation}
+The variational objective is then
+\begin{equation}
+\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
+\end{equation}
+
+Comparing this to the Lagrangian objective (without augmentation) of EPI, we see they are the same
+\begin{equation}
+\begin{split}
+q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
+ &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
+\end{split}
+\end{equation}
+where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
+Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
+The data $\mathbf{x}$ used by this implicit variational inference program would be that generated by the adapting variational approximation $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z}) q_{\bm{\theta}}(\mathbf{z})$, and the likelihood parameters $\bm{\eta}_{f}$ of EPI optimization epoch $k$ are predicated by $\bm{\eta}_{\text{opt},k}$ .
+However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
+Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
+Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
+
 \subsection{Theoretical models}\label{methods_theoretical_models}

### 2021-01-07 written/eLifeRevision/Bittner_eLife2020.tex
intro citation injection

@@
 %On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-Alternatively, statistical inference can be run to obtain model parameters likely to produce some model output, and local sensitivity analyses can be performed at inferred parameter values.
+Alternatively, we can gain insight into these complex models by identifying all of the parameters consistent with the emergent phenomena of interest.
+By examining the structure of the full space of possible parameters, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+
+The preferred formalism for parameter identification in science, statistical inference, has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
 %Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
-Since most neural circuit models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood central to the probabilistic modeling toolkit.
-Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{liepe2014framework}, in which a set of reasonable parameters estimates is obtained via simulation and rejection.
+However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
+Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which a set of reasonable parameters estimates is obtained via simulation and rejection.
 
@@
 First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
-Thus, standard approaches from statistical inference are unavailable.
-The parameter sets obtained from likelihood-free ABC lack a formalized link to Bayesian inference (except in the unrealistic $0$-distance scenario), and lack parameter probabilities.
-Second is the undesirable trade-off between the flexibility and tractability of the approximated posterior distribution.
-While sampling-based approaches like ABC and Markov chain Monte Carlo (MCMC) can produce flexible posterior approximations, they must be run continually for increasing samples.
-While VAE approaches can result in tractable posterior sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions, which restrict the flexibility of the posterior approximation.
+Neuroscientists therefore resort to using approximate Bayesian computation, which requires a rejection heuristic, scales poorly, and only obtains sets of non-rejected parameters lacking probabilities.
+Second, is the undesirable trade-off between the flexibility and sampling speed of the approximated posterior distributions.
+Sampling-based approaches to statistical inference (e.g. ABC and Markov chain Monte Carlo (MCMC)) have flexibility in approximation, but must be executed continually for increasing samples.
+While variational approaches often result in fast sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions.  
+These simple distributions (e.g. mean-field gaussians) restrict the flexibility of the posterior approximation.
 And third, one can never assume what inferred model parameters may predict.
@@
 Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to abstract cognitive function), and network scale (from four to hundreds of neurons).
+These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to decision making), and network scale (from four to hundreds of neurons).
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.

### 2021-01-12 written/eLifeRevision/Bittner_eLife2020.tex
2D lds up-to-date

@@
 EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
+The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig.
+\ref{fig:STG1}).
 %Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
@@
 Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
-A key advantage of EPI over established Bayesian inference is that the predictions made my the inferred distribution are constrained to produce the specified emergent property.
+A key advantage of EPI over established Bayesian inference is that the predictions made by the inferred distribution are constrained to produce the specified emergent property.
 %Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
@@
 Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
-\textbf{B}. Responses of the deterministic smodel ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
+\textbf{B}. Responses of the deterministic model ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
 \textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
-\textbf{D}. EPI posterior of noise parameteres $\mathbf{z}$ conditioned on realistic E-population Fano factors.
+\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on realistic E-population Fano factors.
 The posterior predictive distribution is show on the bottom-left.
@@
 
-The slow noise of the SSSN is an an Ornstein-Uhlenbeck process
+The slow noise of the SSSN is an Ornstein-Uhlenbeck process
 \begin{equation}
@@
 The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
-The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
+The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.
 The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
@@
 Difference in performance of each task during inactivation.
-Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).
+Inactivation level $\gamma$ scales from no inactivation (0) to full inactivation (1).
 We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
@@
 Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
-This shows that strong task representations and hemispheral dominance in the dynamics result in better execution of the Pro task.
+This shows that strong task representations and hemispheric\begin{center}
+•
+\end{center} dominance in the dynamics result in better execution of the Pro task.
 By visualizing these four variables together by $p_A$ (Fig.
 \ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
-Thus, stronger hemispheral dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.
-
-In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
+Thus, stronger hemispheric dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.
+
+In agreement with experimental results from Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
 \ref{fig:SC}E) e.g.
@@
 In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This posterior revealed the mechanisms leading to greater accuarcy in each task as well as those increasing resiliency to perturbation in the Anti task.
+This posterior revealed the mechanisms leading to greater accuracy in each task as well as those increasing resiliency to perturbation in the Anti task.
 Importantly, every connectivity from this inferred distribution predicts fragility and robustness of performance in the Anti and Pro tasks, respectively.
-EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well chosen model.
+EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well-chosen model.
 
@@
  \begin{equation}
- \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
+ \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} + W\mathbf{x}
  \end{equation}
@@
 
-In neuroscience, machine learning has primarily been used to reveal structure in large-scale neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
+In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
  Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
@@
 
+\textbf{TODO} *merge this point in*
+
+While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.
+Talk about Wyktor's work here \cite{mlynarski2020statistical}.
+
 \textbf{Acknowledgements}: \\
@@
 \textbf{Data availability statement}: \\
-The datasets generated during and/or analysed during the current study are available from the corresponding author upon reasonable request.
+The datasets generated during and/or analyzed during the current study are available from the corresponding author upon reasonable request.
 
@@
 In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
-Even so, we are not so much scientifically interested in reasoning about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
+Even so, we do not scientifically reason about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
 In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
@@
 $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-\begin{equation}
+\begin{equation} \label{eq:deep_transform}
 \mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
@@
  \subsubsection{Related approaches}\label{methods_related}
-Inverse problems across scientific fields have been approached in a variety of ways.
-The goals of such approaches are generally to a.) identify parametric sensitivities and manifolds of degeneracy with respect to some model output (structural identifiability analysis), and to reason about the parameters that are most probable to have produced a model output (statistical inference).
-While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.
-
-To understand sensitivity around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
+ When Bayesian inference problems lack conjugacy, scientists use approximate inference methods like variational inference (VI) \cite{saul1998mean} and Markov chain Monte Carlo (MCMC) \cite{metropolis1953equation, hastings1970monte}.
+After optimization, variational methods return a parameterized posterior distribution, which we can analyze.
+Also, the variational approximating distribution class is often chosen such that it permits fast sampling.
+In contrast MCMC methods only produce samples from the approximated posterior distribution.
+No parameterized distribution is estimated, and additional samples are always generated with the same sampling complexity.
+Inference in models defined by systems of differential has been demonstrated with MCMC \cite{girolami2011riemann}, although this approach requires tractable likelihoods.
+Advancements have leveraged structure in stochastic differential equation models to improve likelihood approximations, thus expanding the domain of applicable models \cite{golightly2011bayesian}.
+
+Likelihood-free (or ``simulation-based") inference (LFI) \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
+The most prevalent approach to LFI is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
+The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
+Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
+SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:LRRNN}).
+Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again with a new population of initialized particles to obtain additional samples.
+
+For scientific model analysis, we seek a posterior distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
+For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
+EPI executes likelihood-free inference using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational posterior approximation.
+The deep neural network of EPI defines the parametric form of the posterior approximation.
+Furthermore, the EPI distribution is constrained to produce an emergent property.
+In other words, the summary statistics of the posterior predictive distribution are fixed to have certain first and second moments.
+EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
+The analytic relationship between EPI and variational inference is explained in Secton \ref{methods_VI}.
+
+ We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}).%, preprint posted simultaneously with this preprint).
+We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
+
+The method EPI differs from SNPE in some key ways.
+SNPE belongs to a ``sequential" class of recently developed LFI methods in which two neural networks are used for posterior inference.
+This first neural network is a normalizing flow used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
+A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE \cite{hermans2020likelihood}).
+In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
+This contrasts with EPI and SNPE, which afford a normalizing flow approximation to the posterior, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
+The second neural network in this sequential class of methods is the amortizer.
+This network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
+These methods are optimized on a conditional density (or ratio) estimation objective on a sequentially adapting finite sample-based approximation to the posterior.
+
+The approximating fidelity of the first neural network in sequential approaches is optimized to generalize across the entire distribution it is conditioned upon.
+This optimization towards generalization of sequential methods can reduce the accuracy at the singular posterior of interest.
+Whereas in EPI, the entire expressivity of the normalizing flow is dedicated to learning a single distribution as well as possible.
+While amortization is not possible in EPI parameterized by the mean parameter $\mu$ (due to the inverse mapping problem \cite{wainwright2008graphical}), we have shown this two-network amortization approach to be effective in exponential family distributions defined by their natural parameterization \cite{bittner2019approximating}.
+
+Structural identifiability analysis involves the measurement of sensitivity and unidentifiabilities in natural models.
+Around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
 A popular efficient approach for systems of ODEs has been neural ODE adjoint \cite{chen2018neural} and its stochastic adaptation \cite{li2020scalable}.
-Casting identifiability as a statistical estimation problem, the profile likeilhood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
-An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point \cite{raman2017delineating}.
-Global structual non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
-
-MCMC approaches
-
-ABC approaches
-
-NF LFI approaches
+Casting identifiability as a statistical estimation problem, the profile likelihood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
+An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point by following degenerate dimensions of the fisher information matrix \cite{raman2017delineating}.
+Global structural non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
+With EPI, we have all the benefits given by a statistical inference method plus the ability to query the gradient or Hessian of the inferred distribution at any chosen parameter value.
+
+%\cite{dhruva/goldman papers}
 
@@
 %Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(\mathbf{x} \mid \mathbf{z})$.
-%Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an addtional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
+%Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an additional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
 %Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
@@
 %Deep probability models typically consist of several layers of fully connected neural networks.  
-Deep probability distributions are comprised of multiple layers of fully connected neural networks.  
+Deep probability distributions are comprised of multiple layers of fully connected neural networks (Equation \label{eq:deep_transform}).
 %When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
@@
 However, this computation has cubic complexity in dimensionality for fully connected layers.  
-By restricting our layers to normalizing flows \cite{rezende2015variational} -- bijective functions with fast log determinant Jacobian computations, we can tractably optimize deep generative models with objectives that are a function of sample density, like entropy.
+By restricting our layers to normalizing flows \cite{rezende2015variational, papamakarios2019normalizing} -- bijective functions with fast log determinant Jacobian computations, which confer a fast calculation of the sample log probability.
+Fast log probability calculation confers efficient optimization of the maximum entropy objective (see Section \ref{methods_AL_opt}).
 %Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
 %Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
-TODO: (clean up) We use Real NVP because it's a coupling architecture, which is fast to run either forwards (probability with samples) and backwards (prrobability or hessian).
-Normalizing flow architectures for deep probability distributions used in EPI are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
-
-optimization with respect to entropy, and to confer immediate probability evaluations after optimization.
+We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
+Fast probability evaluation in turn facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
+Glow permutations were used in between coupling stages \cite{kingma2018glow}.
+This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
+Normalizing flow architectures for deep probability distributions used in EPI are specified by the number of coupling stages, neural network layers and units per layer for conditioning functions.
 
 \subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
-\begin{equation}
-H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
-\begin{equation}
-\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-
 To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
@@
 These Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
-Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with its standard parameterization \cite{kingma2014adam}.
-$\bm{\eta}_{\text{opt}}$ is initialized to the zero vector and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
+Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
+
+To take gradients with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, it can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
+\begin{equation}
+H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
+\end{equation}
+Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
+\begin{equation}
+\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
+\end{equation}
+
+The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
 A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
-For example, the initial value of $c$ was $c_0 = 10^{-3}$ during EPI with the oscillating 2D LDS (Fig. S1C).
 The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
 The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu})$ given a batch size $n$.
+The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
 Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
@@
 With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
-In the oscillating 2D LDS example, each augmented Lagrangian epoch ran for 2,000 iterations (Fig. S1C-D).
-Notice the initial entropic growth, and subsequent reduction upon each update of $\bm{\eta}_{\text{opt}}$ and $c$.
-The momentum parameters of the Adam optimizer were reset at the end of each augmented Lagrangian epoch.
+This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see example in Section \ref{methods_2DLDS}).
+The momentum parameters of the Adam optimizer are reset at the end of each augmented Lagrangian epoch.
 
 Rather than starting optimization from some $\bm{\theta}$ drawn from a randomized distribution, we found that initializing $q_{\bm{\theta}}(\mathbf{z})$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
-The parameters of the Gaussain initialization were chosen on an application-specific basis.  
-Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some standard deviation $\bm{\sigma}_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_V1}).
-
-To assess whether EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
+The parameters of the Gaussian initialization were chosen on an application-specific basis.  
+Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some standard deviation $\bm{\sigma}_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
+
+To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
 The algorithm has converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
@@
 The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made from $k$ resamplings of $\mathbf{z}$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
-$k$ is determined by a fraction of the batch size $\nu$, which varies according to the application.
-In the linear two-dimensional system example, we used a batch size of $n = 1000$ and set $\nu = 0.1$ resulting in convergence after the ninth epoch of optimization. (Fig. S1C-D black dotted line).
 
@@
 Next, one should consider the computational cost of the gradient calculations. 
-In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_2DLDS}) for the emergent property statistic given the model parameters.  
+In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_RNN}) for the emergent property statistic given the model parameters.  
 On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
 Section \ref{methods_STG}).
-In such cases, optimization will be expensive.
+In such cases, backpropagating gradients through the SDE evolution will be expensive.
 
 \subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
+To gain intuition for EPI optimization, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. \ref{fig:LDS1}A):
 \begin{equation} 
@@
 \end{equation}
-with
-\begin{equation}
-A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
-\end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
- $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
-\begin{equation}
- \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
- \end{equation} 
+where
+\begin{equation}
+A = \begin{bmatrix} a_{1,1} & a_{1,2} \\ a_{2,1} & a_{2,2} \end{bmatrix},
+\end{equation}
+$\tau = 1$s, and $\mathbf{z} = [a_{1,1}, a_{1,2}, a_{2,1}, a_{2,2}]^\top$.
 \begin{figure}
 \begin{center}
-\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
+\includegraphics[scale=0.7]{figures/figLDS1/figLDS1.pdf}
 \end{center}
-\begin{flushleft}
 \caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
-B.
-The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.
-Dashed lines indicate the parameter axes.
-C. Entropy throughout the optimization.  
-At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
-D. Emergent property moments throughout optimization.  
-At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
-\end{flushleft}
+B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces oscillations around 1Hz.  
+Inset show distribution of eigenvalues across the EPI posterior.
+C. (Left) Entropy throughout the optimization with different $c_0$.
+Individual optimizations were run with the same architecture and random seed.
+Norm of emergent property statistic errors throughout optimization.  
+}
 \label{fig:LDS1}
@@
 
+To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz.
+Thus, our statistics vector $f(\mathbf{x}; \mathbf{z})$ for this emergent property is two dimensional, and after specifying the variance constraint is
+\begin{equation}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} \text{real}(\lambda_1 (\mathbf{x}; \mathbf{z})) \\ \text{imag}(\lambda_1 (\mathbf{x}; \mathbf{z})) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0 \\ 2 \pi \omega \end{bmatrix}  \\
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} \text{real}(\lambda_1 (\mathbf{x}; \mathbf{z}))\\ \text{imag}(\lambda_1 (\mathbf{x}; \mathbf{z})) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.5^2 \\ (\frac{\pi}{5})^2 \end{bmatrix}.
+\end{split}
+\end{equation}
+
+Notation of mean and variance taken with respect to $\mathbf{x}$ is omitted, since this model is deterministic, and EPI generalizes naturally to this scenario.
+
+ \begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{figures/figLDS2/figLDS2.pdf}
+\end{center}
+\caption{\small (LDS2): A. Probability contours in the $a_{1,1}$-$a_{2,2}$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+B. Probability contours in the $a_{1,2}$-$a_{2,1}$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\label{fig:LDS2}
+\end{figure}
+
 Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $\mathbf{z}$, since the eigenvalues can be calculated using the quadratic formula:
 \begin{equation}
-\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
-\end{equation}
-
-Importantly, even though $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}$ directly.
+\lambda = \frac{(\frac{a_{1,1} + a_{2,2}}{\tau}) \pm \sqrt{(\frac{a_{1,1}+a_{2,2}}{\tau})^2 + 4(\frac{a_{1,2} a_{2,1} - a_{1,1} a_{2,2}}{\tau})}}{2}.
+\end{equation}
+
+Importantly, even though $f(\mathbf{x}; \mathbf{z})$ is analytically available and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}(z \mid \mathcal{X})$ directly.  
 This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we used EPI to approximate this distribution (Fig.
-S1B).
-We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
+Instead, we can use EPI to approximate this distribution (Fig. \ref{fig:LDS1}B). 
+We used a real-NVP normalizing flow architecture with three masks, two neural network layers of 50 units per mask, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
 (see Section \ref{methods_NF}).
+
+The effect of $c_0$ on the EPI optimization is shown in Figure \ref{fig:LDS1}C.
+Each EPI optimization was run for enough augmented Lagrangian epochs until convergence, with 1,000 iterations per epoch.  
+We see that low values of $c_0$ (e.g. $10^{-5}$) result in fast entropic growth to the uniform distribution at the outset of optimization. 
+We know this, since the dashed black line indicates the maximum entropy of a distribution with the given bounds. 
+Initial entropic growth can be desireable based on the application, however one should expect a greater number of epochs until convergence compared to a greater value of $c_0$ (e.g. $10^{-2}$).
+
+However, an important drawback of selecting a high `c0` is that you may miss interesting structure in the learned distribution.
 
@@
 To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
-In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
-the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). 
-The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
+In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} = 0$, dotted black line at
+the standard deviation $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.5$, and the dotted gray line at twice the standard deviation
+$\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 1$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
+The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_{1,1}$ and $a_{2,2}$.
 Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figLDS2/figLDS2.pdf}
+\includegraphics[scale=0.7]{figures/figLDS3/figLDS3.pdf}
 \end{center}
-\begin{flushleft}
-\caption{\small (LDS2): A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\end{flushleft}
-\label{fig:LDS2}
+\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+A. Each dimension of the simulated trajectories throughout time.  
+B  The simulated trajectories in phase space.}
+\label{fig:LDS3}
 \end{figure}
 
-To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
+To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} = 0$, we have
 \begin{equation}
 \text{imag}(\lambda_1) = \begin{cases}
-                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
+                             \sqrt{\frac{a_{1,1} a_{2,2} - a_{1,2} a_{2,1}}{\tau}},  & \text{if } a_{1,1} a_{2,2} < a_{1,2} a_{2,1} \\
                              0 & \text{otherwise } \\
@@
 
-When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
-\begin{equation}
-\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
-\end{equation}
-Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
-S2B).
-With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
+When $\tau=1$ and $a_{1,1} a_{2,2} > a_{1,2} a_{2,1}$ (center of distribution above), we have the following equation for the other two dimensions:
+\begin{equation}
+\text{imag}(\lambda_1)^2 = a_{1,1} a_{2,2} - a_{1,2} a_{2,1}
+\end{equation}
+Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_{1,1} a_{2,2} - a_{1,2} a_{2,1} = (2 \pi)^2$ assuming $a_{1,1} a_{2,2} = \mathbb{E}_z\left[a_{1,1} a_{2,2}\right]$ (Fig.
+\ref{fig:LDS2}B).
 This validates the curved structure of the inferred distribution learned through EPI.  
-We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
 Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figLDS3/figLDS3.pdf}
-\end{center}
-\begin{flushleft}
-\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
-A.
-Each dimension of the simulated trajectories throughout time.
-B  The simulated trajectories in phase space.}
-\end{flushleft}
-\label{fig:LDS3}
-\end{figure}
+As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. \ref{fig:LDS3}).
 
@@
 \end{center}
-\begin{flushleft}
 \caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
@@
 B. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
-\end{flushleft}
 \label{fig:STG1}
@@
 We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, set $\nu = 0.25$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
-This initialization had much greater entropy and a different emergent property than the the returned EPI posterior.
+This initialization had much greater entropy and a different emergent property than the returned EPI posterior.
 
@@
 
-Since the E-population of this network increases exponentially in the abscense of  recurrent inhibitory feedback, we may also observe a paradoxical effect in the inhibitory populations (which is present in E-I networks).
+Since the E-population of this network increases exponentially in the absense of  recurrent inhibitory feedback, we may also observe a paradoxical effect in the inhibitory populations (which is present in E-I networks).
 At 50\% contrast (Fig. \ref{fig:V1}B, dots), this network exhibits a paradoxical effect in the P-population (Fig. \ref{fig:V1}C), but no others (Fig. \ref{fig:V1_1}).
@@
 
-Fano factor is calculated as the temporal variance divided by the temporal mean following some time $t_{ss}$ following dynamical evolution from the initial state at $\mathbf{x}(t=0)$.
+Fano factor is calculated as the temporal variance divided by the temporal mean following sometime $t_{ss}$ following dynamical evolution from the initial state at $\mathbf{x}(t=0)$.
 
@@
 \caption{\small (SC3): 
-A. Connectitivy eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
+A. Connectivity eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
 B. Same for Anti task.
@@
 Traditional approaches to likelihood-free inference -- approximate Bayesian computation (ABC) methods -- randomly sample parameters $\mathbf{z}$ until a suitable set is obtained.
-State-of-the-art ABC methods leverage sequential monte-carlo (SMC) sampling techniques to obtain parameter sets more efficiently.
+State-of-the-art ABC methods leverage sequential Monte Carlo (SMC) sampling techniques to obtain parameter sets more efficiently.
 To obtain more parameter samples, SMC-ABC must be run from scratch again.
@@
 SMC-ABC has many hyperparameters, of which pyABC selects automatically by running some initial diagnostics upon initialization.
-In concurrence with the literacture, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
+In concurrence with the literature, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
 We searched over many SNPE hyperparameter choices: $n_{\text{train}} \in [2,000, 10,000, 100,000]$ is the number of simulations run per training epoch, and $n_{\text{mades}} \in [2, 3]$ is the number of masked autoregressive density estimators in the deep parameter distribution architecture.

### 2021-01-13 epi/models.py
default param change

@@ -206,11 +206,11 @@ class Model(object):
         arch_type="coupling",
         num_stages=3,
         num_layers=2,
-        num_units=None,
+        num_units=50,
         elemwise_fn="affine",
-        batch_norm=True,
+        batch_norm=False,
         bn_momentum=0.,
-        post_affine=False,
+        post_affine=True,
         random_seed=1,
         init_type=None,  # "iso_gauss",
         init_params=None,  # {"loc": 0.0, "scale": 1.0},

### 2021-01-19 scripts/SSSN/cpu_drdh.sh
terre scripts

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --account=stats
+#SBATCH --job-name=drdh_epi
+#SBATCH -c 1
+#SBATCH --time=3:00:00
+#SBATCH --mem-per-cpu=25gb
+
+source activate epi
+python3 drdh_epi.py --alpha $1 --beta $2 --contrast $3 --logc0 $4 --random_seed $5

### 2021-01-19 scripts/SSSN/cpu_ff.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --account=stats
+#SBATCH --job-name=drdh_epi
+#SBATCH -c 1
+#SBATCH --time=10:00:00
+#SBATCH --mem-per-cpu=25gb
+
+source activate epi
+python3 ff_epi.py --alpha E --ind $1 --ff_mean $2 --ff_std $3 --beta $4 --logc0 $5 --random_seed $6

### 2021-01-19 scripts/SSSN/ff_epi.py

@@ -1,0 +1,94 @@
+import os
+import argparse
+import numpy as np
+import tensorflow as tf
+from epi.models import Parameter, Model
+from epi.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_Fano
+
+# Parse script command-line parameters.
+parser = argparse.ArgumentParser()
+parser.add_argument('--alpha', type=str, default='E') # neuron type
+parser.add_argument('--ind', type=int, default=62) # neuron type
+parser.add_argument('--lim', type=float, default=0.25) # neuron type
+parser.add_argument('--ff_mean', type=float, default=0.05) # neuron type
+parser.add_argument('--ff_std', type=float, default=0.025) # neuron type
+parser.add_argument('--beta', type=float, default=4.) # aug lag hp
+parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
+parser.add_argument('--random_seed', type=int, default=1)
+args = parser.parse_args()
+
+alpha = args.alpha
+ind = args.ind
+lim = args.lim
+ff_mean = args.ff_mean
+ff_std = args.ff_std
+beta = args.beta
+c0 = 10.**args.logc0
+random_seed = args.random_seed
+
+contrast = 0.5
+W_mat = load_SSSN_variable('W', ind=ind)
+hb = load_SSSN_variable('hb', ind=ind).numpy()
+hc = load_SSSN_variable('hc', ind=ind).numpy()
+_h = (hb + contrast*hc)
+
+neuron_inds = {'E':0, 'P':1, 'S':2, 'V':3}
+neuron_ind = neuron_inds[alpha]
+
+M = 100
+
+if ind == 49:
+    sigma_eps = np.array([.015, .015, .015, .015])[None,:]
+elif ind == 62:
+    sigma_eps = np.array([0.00137919, 0.00202632, 0.00074385, 0.00485482])[None,:]
+
+# 1. Specify the V1 model for EPI.
+D = 4
+lb = _h - lim
+ub = _h + lim
+
+h = Parameter("h", D, lb=lb, ub=ub)
+
+# Define model
+name = "SSSN_%s_ff_%.2E_%.2E_ind=%d" % (alpha, ff_mean, ff_std, ind)
+parameters = [h]
+model = Model(name, parameters)
+
+dt = 0.0005
+T = 150
+N = 50
+
[…]

### 2021-01-19 scripts/SSSN/ff_epi_sigma.py

@@ -1,0 +1,95 @@
+import os
+import argparse
+import numpy as np
+import tensorflow as tf
+from epi.models import Parameter, Model
+from epi.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_Fano_sigma
+
+# Parse script command-line parameters.
+parser = argparse.ArgumentParser()
+parser.add_argument('--alpha', type=str, default='E') # neuron type
+parser.add_argument('--ind', type=int, default=62) # neuron type
+parser.add_argument('--lim', type=float, default=None) # neuron type
+parser.add_argument('--ff_mean', type=float, default=1.) # neuron type
+parser.add_argument('--ff_std', type=float, default=0.025) # neuron type
+parser.add_argument('--beta', type=float, default=4.) # aug lag hp
+parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
+parser.add_argument('--random_seed', type=int, default=1)
+args = parser.parse_args()
+
+alpha = args.alpha
+ind = args.ind
+lim = args.lim
+ff_mean = args.ff_mean
+ff_std = args.ff_std
+beta = args.beta
+c0 = 10.**args.logc0
+random_seed = args.random_seed
+
+if lim is None:
+    if ind == 49:
+        lim = 0.03
+    elif ind == 62:
+        lim = 0.005
+    else:
+        raise NotImplementedError()
+
+contrast = 0.5
+W_mat = load_SSSN_variable('W', ind=ind)
+hb = load_SSSN_variable('hb', ind=ind).numpy()
+hc = load_SSSN_variable('hc', ind=ind).numpy()
+h = (hb[None,:] + contrast*hc[None,:])
+
+neuron_inds = {'E':0, 'P':1, 'S':2, 'V':3}
+neuron_ind = neuron_inds[alpha]
+
+M = 100
+# 1. Specify the V1 model for EPI.
+D = 4
+lb = np.zeros((D,))
+ub = lim*np.ones((D,))
+sigma_eps = Parameter("sigma_eps", D, lb=lb, ub=ub)
+
+# Define model
+name = "SSSN_ff_sigma_%s_%.2E_%.2E_ind=%d" % (alpha, ff_mean, ff_std, ind)
+parameters = [sigma_eps]
+model = Model(name, parameters)
+
+dt = 0.0005
+T = 150
+N = 100
+
[…]

### 2021-01-19 scripts/SSSN/ff_sigma_sweep.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for ind in 49
+do
+  for ff_mean in 1.
+  do
+    for ff_std in 0.125
+    do
+      for beta in 2. 4.
+      do
+        for logc0 in 1 2
+        do
+          for rs in 1 2 3
+          do
+            sbatch cpu_ff_sigma.sh $ind $ff_mean $ff_std $beta $logc0 $rs
+          done
+        done
+      done
+    done
+  done
+done

### 2021-01-19 scripts/STG_Circuit/cpu_stg_epi.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --account=stats
+#SBATCH --job-name=STG_epi
+#SBATCH -c 1
+#SBATCH --time=11:00:00
+#SBATCH --mem-per-cpu=50gb
+
+source activate epi
+python3 stg_epi_fast.py --freq $1 --mu_std $2 --elemwise_fn $3 --num_layers $4 --logc0 $5 --random_seed $6

### 2021-01-19 scripts/STG_Circuit/stg_epi_fast.py

@@ -1,0 +1,91 @@
+import os
+import argparse
+import numpy as np
+import tensorflow as tf
+from epi.models import Parameter, Model
+from epi.STG_Circuit import NetworkFreq
+import time
+
+DTYPE = tf.float32
+
+# Parse script command-line parameters.
+parser = argparse.ArgumentParser()
+parser.add_argument('--freq', type=float, default=0.55) # frequency for mu
+parser.add_argument('--mu_std', type=float, default=0.05) # std in mu constraint
+parser.add_argument('--elemwise_fn', type=str, default="spline") # lower bound on g_el
+parser.add_argument('--num_layers', type=int, default=2) # aug lag hp
+parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
+parser.add_argument('--random_seed', type=int, default=1)
+args = parser.parse_args()
+
+freq = args.freq
+mu_std = args.mu_std
+elemwise_fn = args.elemwise_fn
+num_layers = args.num_layers
+c0 = 10.**args.logc0
+random_seed = args.random_seed
+
+g_el_lb = 4.
+beta = 2.
+
+sigma_I = 1.e-12
+#sleep_dur = np.abs(args.logc0) + random_seed/5. + beta/3.
+#print('short stagger sleep of', sleep_dur, flush=True)
+#time.sleep(sleep_dur)
+
+# 1. Specify the V1 model for EPI.
+D = 2 
+g_el = Parameter("g_el", 1, lb=g_el_lb, ub=8.)
+g_synA = Parameter("g_synA", 1, lb=0.01, ub=4.)
+
+# Define model
+name = "STG_sigmaI=%.2E" % sigma_I
+parameters = [g_el, g_synA]
+model = Model(name, parameters)
+
+# Emergent property values.
+mu = np.array([freq, mu_std**2])
+
+init_type = 'abc'
+abc_std = mu_std
+init_params = {'num_keep':500,
+               'means':np.array([freq]),
+               'stds':np.array([abc_std,])}
+
+dt = 0.025
+T = 300
+network_freq = NetworkFreq(dt, T, sigma_I, mu)
+
+model.set_eps(network_freq)
+
+# 3. Run EPI.
+q_theta, opt_data, epi_path, failed = model.epi(
+    mu,
+    arch_type="coupling",
+    num_stages=2,
+    num_layers=num_layers,
+    num_units=25,
+    post_affine=True,
[…]

### 2021-01-19 scripts/STG_Circuit/stg_sweep.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for freq in 0.55
+  do
+  for mu_std in 0.025
+  do
+    for elemwise_fn in affine
+    do
+      for num_layers in 2
+      do
+        for logc0 in 5
+        do
+          for rs in {6..15}
+          do
+            sbatch cpu_stg_epi.sh $freq $mu_std $elemwise_fn $num_layers $logc0 $rs
+          done
+        done
+      done
+    done
+  done
+done

### 2021-01-19 scripts/V1_Circuit/SV_flip_sweep.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for c0 in -2 0 2
+do
+  for rs in {1..3}
+  do
+    echo "Running SV Flip, c0 = $c0, rs = $rs"
+    sbatch cpu_SV_flip.sh $c0 $rs
+  done
+done

### 2021-01-19 scripts/V1_Circuit/drdh_epi.py

@@ -1,0 +1,83 @@
+import os
+import argparse
+import numpy as np
+import tensorflow as tf
+from epi.models import Parameter, Model
+from epi.example_eps import V1_dr_eps
+import time
+
+# Parse script command-line parameters.
+parser = argparse.ArgumentParser()
+parser.add_argument('--alpha', type=str, default='E') # neuron type
+parser.add_argument('--beta', type=float, default=4.) # aug lag hp
+parser.add_argument('--epsilon', type=float, default=0.) # noise
+parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
+parser.add_argument('--bnmom', type=float, default=.99) # log10 of c_0
+parser.add_argument('--random_seed', type=int, default=1)
+args = parser.parse_args()
+
+alpha = args.alpha
+beta = args.beta
+epsilon = args.epsilon
+c0 = 10.**args.logc0
+bnmom = args.bnmom
+random_seed = args.random_seed
+
+sleep_dur = ord(alpha)/11. + beta/2. + 3.*epsilon + np.abs(args.logc0) + random_seed/17.
+print('short stagger sleep of', sleep_dur, flush=True)
+time.sleep(sleep_dur)
+
+# 1. Specify the V1 model for EPI.
+D = 4
+lb = -5.*np.ones((D,))
+ub = 5.*np.ones((D,))
+
+dh = Parameter("dh", D, lb=lb, ub=ub)
+
+# Define model
+name = "V1_Circuit_%s_eps=%.2f" % (alpha, epsilon)
+parameters = [dh]
+model = Model(name, parameters)
+
+X_INIT = tf.constant(np.random.normal(1.0, 0.01, (1, 4, 1)).astype(np.float32))
+
+inc_val = 0.
+
+npzfile = np.load('SVflips_hs.npz')
+print(npzfile)
+h = npzfile['b_cross'][0]
+
+dr = V1_dr_eps(alpha, inc_val, epsilon, h=h)
+model.set_eps(dr)
+
+# Emergent property values.
+mu = np.array([inc_val, 0.5**2])
+
+# 3. Run EPI.
+q_theta, opt_data, epi_path, failed = model.epi(
+    mu,
+    arch_type='coupling',
+    num_stages=3,
+    num_layers=2,
+    num_units=50,
+    post_affine=True,
+    batch_norm=True,
+    bn_momentum=bnmom,
+    K=15,
+    N=500,
+    num_iters=2500,
[…]

### 2021-01-19 scripts/V1_Circuit/drdh_sweep.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for alpha in P
+do
+  for beta in 2.
+  do
+    for logc0 in 0
+    do
+      for bnmom in 0.999
+      do
+        for rs in {1..5}
+        do
+          sbatch cpu_drdh.sh $alpha $beta 0.4 $logc0 $bnmom $rs
+        done
+      done
+    done
+  done
+done

### 2021-01-19 scripts/rnn_epi_timing_batch.sh

@@ -3,7 +3,11 @@ source activate epi
 
 # This shell script does it linearly, but could you run each of these 
 # python scripts independently on different instances?
+<<<<<<< Updated upstream:scripts/rnn_epi_timing_batch.sh
 for n in 2 4
+=======
+for N in 7
+>>>>>>> Stashed changes:scripts/stab_amp/stab_amp_smc_sweep.sh
 do
   for T in 10
   do

### 2021-01-19 scripts/stab_amp/cpu_stab_amp_epi.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --account=stats
+#SBATCH --job-name=stab_amp_epi
+#SBATCH -c 1
+#SBATCH --time=2:00:00
+#SBATCH --mem-per-cpu=2gb
+
+source activate epi
+python3 stab_amp_epi.py --N $1 --c0 $2 --rs $3

### 2021-01-19 scripts/stab_amp/cpu_stab_amp_smc.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#SBATCH --account=stats
+#SBATCH --job-name=sa_smc
+#SBATCH -c 1
+#SBATCH --time=6:00:00
+#SBATCH --mem-per-cpu=1gb
+
+source activate epi
+python3 stab_amp_smc.py --N $1 --rs $2

### 2021-01-19 scripts/stab_amp/stab_amp_epi.py

@@ -1,0 +1,98 @@
+"""Run EPI on Rank2 RNN. """
+
+from epi.models import Model, Parameter
+from epi.example_eps import linear2D_freq
+from epi.util import sample_aug_lag_hps
+import numpy as np
+import tensorflow as tf
+import tensorflow_probability as tfp
+import argparse
+
+tf.test.is_gpu_available(
+    cuda_only=False, min_cuda_compute_capability=None
+)
+
+DTYPE = np.float32
+
+# Get random seed.
+parser = argparse.ArgumentParser()
+parser.add_argument('--N', type=int)
+parser.add_argument('--c0', type=float)
+parser.add_argument('--rs', type=int)
+args = parser.parse_args()
+
+print('Running epi for RNN N=%d, c0=%f, rs=%d stable amplification.' % (args.N, args.c0, args.rs))
+N = args.N
+c0 = args.c0
+rs = args.rs
+
+r = 2 # rank-2 networks
+
+num_epochs = 20
+
+# 1. Define model: dxd matrix
+D = int(N*r)
+
+# Set up the bound vectors.
+lb = -np.ones((D,))
+ub = np.ones((D,))
+
+# Define the parameter A.
+U = Parameter("U", D, lb=lb, ub=ub)
+V = Parameter("V", D, lb=lb, ub=ub)
+parameters = [U, V]
+
+# Define the model matrix.
+M = Model("Rank2Net", parameters)
+
+# 2. Define the emergent property:
+Js_eig_max_mean = 1.5
+J_eig_realmax_mean = 0.5
+mu = np.array([Js_eig_max_mean, 
+               0.25**2, 
+               J_eig_realmax_mean,
+               0.25**2], dtype=DTYPE)
+
+
+def stable_amplification_r2(U, V):
+    U = tf.reshape(U, (-1, N, 2))
+    V = tf.reshape(V, (-1, N, 2))
+    J = tf.matmul(U, tf.transpose(V, [0,2,1]))
+    J = J + tf.random.normal(J.shape, 0., 0.01)
+    Js = (J + tf.transpose(J, [0, 2, 1])) / 2.
+    Js_eigs = tf.linalg.eigvalsh(Js)
+    Js_eig_max = tf.reduce_max(Js_eigs, axis=1)
+    
+    # Take eig of low rank similar mat
[…]

### 2021-01-19 scripts/stab_amp/stab_amp_epi_sweep.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for N in 2 5 10 15 20 25
+do
+  for c0 in 3
+  do
+    for rs in {1..5}
+    do
+      echo "Running N = $N, c0 = $c0, rs = $rs"
+      sbatch cpu_stab_amp_epi.sh $N $c0 $rs
+    done
+  done
+done

### 2021-01-19 scripts/stab_amp/stab_amp_smc.py

@@ -1,0 +1,120 @@
+"""SNPE: RNN stable amplification. """
+
+import numpy as np
+import os
+import pickle
+import tempfile
+import pandas as pd
+import time
+#import matplotlib.pyplot as plt
+import argparse
+
+import pyabc
+import scipy.stats as st
+
+DTYPE = np.float32
+
+# Get random seed.
+parser = argparse.ArgumentParser()
+parser.add_argument('--N', type=int)
+parser.add_argument('--rs', type=int, default=1)
+args = parser.parse_args()
+
+N = args.N
+rs = args.rs
+
+sleep_time = N*0.5 + rs*0.05
+time.sleep(sleep_time)
+
+print('Running SNPE on RNN conditioned on stable amplification with:')
+print('N = %d, seed=%d' % (N, rs))
+
+base_path = os.path.join("data", "smc")
+save_dir = "SMC_RNN_stab_amp_N=%d_rs=%d" % (N, rs)
+
+save_path = os.path.join(base_path, save_dir)
+if not os.path.exists(save_path):
+    os.makedirs(save_path)
+
+if os.path.exists(os.path.join(base_path, save_dir, "optim.pkl")):
+    print("SMC optimization already run. Exitting.")
+    exit()    
+
+def model(parameter):
+    u1 = np.array([parameter["U%d1" % i] for i in range(1, N+1)])
+    u2 = np.array([parameter["U%d2" % i] for i in range(1, N+1)])
+    v1 = np.array([parameter["V%d1" % i] for i in range(1, N+1)])
+    v2 = np.array([parameter["V%d2" % i] for i in range(1, N+1)])
+
+    U = np.stack((u1, u2), axis=1)
+    V = np.stack((v1, v2), axis=1)
+
+    J = np.matmul(U, np.transpose(V))
+    J = J + np.random.normal(0., 0.01, J.shape)
+    Js = (J + np.transpose(J)) / 2.
+    Js_eigs = np.linalg.eigvalsh(Js)
+    Js_eig_max = np.max(Js_eigs, axis=0)
+
+    # Take eig of low rank similar mat
+    Jr = np.matmul(np.transpose(V), U) + 0.01*np.eye(2)
+    Jr_tr = np.trace(Jr)
+    sqrt_term = np.square(Jr_tr) + -4.*np.linalg.det(Jr)
+    if sqrt_term < 0.:
+        sqrt_term = 0.
[…]

### 2021-01-19 scripts/stab_amp/stab_amp_snpe_sweep.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+# This shell script does it linearly, but could you run each of these 
+# python scripts independently on different instances?
+for N in 20
+do
+  for n_train in 10000
+  do
+    for n_mades in 2
+    do
+      for n_atoms in 100
+      do
+        for rs in {4..10}
+        do
+          echo "Running N = $N, n_train $n_train, n_mades $n_mades, n_atoms = $n_atoms, rs = $rs"
+          sbatch gpu_stab_amp_snpe.sh $N $n_train $n_mades $n_atoms $rs
+        done
+      done
+    done
+  done
+done

### 2021-01-19 written/SC_model/SC_model/SC_model.tex

--- 2020-12-11 written/SC_model/SC_model.tex
@@
 The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
+The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
 The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
@@
 Shading indicates standard deviation across trials.
-E. The effect of delay period inactivation on accuracy versus dynamics eigenvalues.
+E.
+Difference in performance of each task during inactivation.
+Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).
+We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
+F. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
 }
@@
 
-We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C, Fig. \ref{fig:SX5}).
-Each parameter is predictive of accuracy in either task (Fig. \ref{fig:SX1}), and may have the same ($vW$ and $dW$) or opposite ($sW$ and $hW$) effects on $p_P$ and $p_A$.
+We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
+Some parameters were predictive of accuracy while others were not (Fig. \ref{fig:SX1}), and often had different effects on $p_P$ and $p_A$.
 To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $v_i$ for all $W$ parameterized by $z$ (Fig.
@@
 
-Smaller $\lambda_{\text{all}}$ and greater $\lambda_{\text{side}}$ predict greater accuracy in both tasks (Fig.
+Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
+This shows that strong task representations and hemispheral dominance in the dynamics result in better execution of the Pro task.
+These same connectivity variables are anticorrelated with Anti accuracy across the EPI posterior.
+By visualizing these four variables together with $p_A$ (Fig.
+\ref{fig:SX2}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
+Thus, stronger hemispheral dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.
+
+%Smaller $\lambda_{\text{all}}$ and greater $\lambda_{\text{side}}$ predict greater accuracy in both tasks (Fig.
 \ref{fig:SX2}B) suggesting that wide separation between Left and Right populations at the end of stimulus presentation and fast decaying dynamics during the choice period yield better performance in both tasks.
-Since the Pro neurons receive a positive bias input $h_{P, bias}$, $\lambda_{\text{task}}$ is not as important to performance in the Pro task as in the Anti task (Fig. \ref{fig:SX2}B).
-The diagonal mode, hypothesized to be crucial for processing in the Anti task, was ultimately anticorrelated with $p_A$, 
-This suggests that $p_A$ increases with dynamics having stronger task and stimulus representations and faster decaying dynamics in alternate dimensions.
+%Since the Pro neurons receive a positive bias input $h_{P, bias}$, $\lambda_{\text{task}}$ is not as important to performance in the Pro task as in the Anti task (Fig. \ref{fig:SX2}B).
+%The diagonal mode, hypothesized to be crucial for processing in the Anti task, was ultimately anticorrelated with $p_A$, 
+%This suggests that $p_A$ increases with dynamics having stronger task and stimulus representations and faster decaying dynamics in alternate dimensions.
 
 In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
-\ref{fig:SX3}), e.g.
+\ref{fig:SC}E) e.g.
 (Fig.
 \ref{fig:SC}D, bottom).
-Interestingly, the trends in eigenvalues that result in greater $p_A$ increase the error from delay period inactivation in the Anti task ($p_A - p_{A, opto}$) (Fig. \ref{fig:SC}E.
-Despite these opposing effects, $p_A$ is uncorrelated with ($p_A - p_{A, opto}$), while $p_P$ is correlated (Fig. \ref{fig:SX4}).
-This shows that the degree of error introduced by delay period inactivation is controlled by the dynamical mechanisms controlling success in the Pro task (namely $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$).
-Such comparable dynamics between the non-inactivated Pro task and the delay period inactivated Anti task are salient in simulations (Fig. \ref{fig:SC}D top-left and bottom-right).
+This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
+Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task.
+Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation.
+%Interestingly, the trends in eigenvalues that result in greater $p_A$ increase the error from delay period inactivation in the Anti task ($p_A - p_{A, opto}$) (Fig. \ref{fig:SC}E.
+%Despite these opposing effects, $p_A$ is uncorrelated with ($p_A - p_{A, opto}$), while $p_P$ is correlated (Fig. \ref{fig:SX4}).
+%\textbf{This shows that the connectivity mechanisms that decrease $p_A$ during delay period inactivation are the same mechanisms producing greater $p_P$ during no activation (namely $\lambda_{\text{task}}$ and $\lambda_{\text{side}}$).}
+%Such comparable dynamics between the non-inactivated Pro task and the delay period inactivated Anti task are salient in simulations (Fig. \ref{fig:SC}D top-left and bottom-right).
 
@@
 \begin{center}
-\includegraphics[scale=0.6]{figs/figSX5.pdf}
-\end{center}
-\caption{\small The EPI parameter distribution of Fig. 1C with reference to a unity line.
-}
-\label{fig:SX5}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[scale=0.85]{figs/figSX1.pdf}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figs/figSX4.pdf}
+\includegraphics[scale=0.6]{figs/figSX3.pdf}
 \end{center}
 \caption{\small 
-Difference in performance of each task during inactivation.
-Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).
-We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
+A. Connectitivy eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
+B. Same for Anti task.
 }
@@
 \begin{center}
-\includegraphics[scale=0.75]{figs/figSX3.pdf}
+\includegraphics[scale=0.75]{figs/figSX4.pdf}
 \end{center}
-\caption{\small Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.7.
+\caption{\small Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.8.
 }
@@
 \end{figure}
-
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.6]{figs/figSX2.pdf}
-%\end{center}
-%\caption{\small
-%Same as Fig. 1C shaded by Anti task accuracy.}
-%\label{fig:SC_EPI2}
-%\end{figure}
 

### 2021-01-19 written/eLifeRevision/Bittner_eLife2020.tex

@@
 EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig.
-\ref{fig:STG1}).
+The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
 %Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
@@
  \begin{equation}
- \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} + W\mathbf{x}
+ \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
  \end{equation}
@@
 \subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI optimization, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. \ref{fig:LDS1}A):
+To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
 \begin{equation} 
@@
 \end{equation}
-where
-\begin{equation}
-A = \begin{bmatrix} a_{1,1} & a_{1,2} \\ a_{2,1} & a_{2,2} \end{bmatrix},
-\end{equation}
-$\tau = 1$s, and $\mathbf{z} = [a_{1,1}, a_{1,2}, a_{2,1}, a_{2,2}]^\top$.
+with
+\begin{equation}
+A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
+\end{equation}
+To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
+ $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
+To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
+\begin{equation}
+ \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
+ \end{equation} 
 \begin{figure}
 \begin{center}
-\includegraphics[scale=0.7]{figures/figLDS1/figLDS1.pdf}
+\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
 \end{center}
+\begin{flushleft}
 \caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
-B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces oscillations around 1Hz.  
-Inset show distribution of eigenvalues across the EPI posterior.
-C. (Left) Entropy throughout the optimization with different $c_0$.
-Individual optimizations were run with the same architecture and random seed.
-Norm of emergent property statistic errors throughout optimization.  
-}
+B.
+The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.
+Dashed lines indicate the parameter axes.
+C. Entropy throughout the optimization.  
+At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
+D. Emergent property moments throughout optimization.  
+At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
+\end{flushleft}
 \label{fig:LDS1}
@@
 
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz.
-Thus, our statistics vector $f(\mathbf{x}; \mathbf{z})$ for this emergent property is two dimensional, and after specifying the variance constraint is
-\begin{equation}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} \text{real}(\lambda_1 (\mathbf{x}; \mathbf{z})) \\ \text{imag}(\lambda_1 (\mathbf{x}; \mathbf{z})) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0 \\ 2 \pi \omega \end{bmatrix}  \\
- \text{Var}_{\mathbf{z}}\begin{bmatrix} \text{real}(\lambda_1 (\mathbf{x}; \mathbf{z}))\\ \text{imag}(\lambda_1 (\mathbf{x}; \mathbf{z})) \end{bmatrix}  &~~=~~  \begin{bmatrix} 0.5^2 \\ (\frac{\pi}{5})^2 \end{bmatrix}.
-\end{split}
-\end{equation}
-
-Notation of mean and variance taken with respect to $\mathbf{x}$ is omitted, since this model is deterministic, and EPI generalizes naturally to this scenario.
+Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $\mathbf{z}$, since the eigenvalues can be calculated using the quadratic formula:
+\begin{equation}
+\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
+\end{equation}
+
+Importantly, even though $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}$ directly.
+This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
+Instead, we used EPI to approximate this distribution (Fig.
+S1B).
+We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
+(see Section \ref{methods_NF}).
+
+Even this relatively simple system has nontrivial (though intuitively sensible) structure in the parameter distribution.  
+To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
+In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
+the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
+$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). 
+The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
+Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figLDS2/figLDS2.pdf}
+\includegraphics[scale=0.5]{figures/figLDS2/figLDS2.pdf}
 \end{center}
-\caption{\small (LDS2): A. Probability contours in the $a_{1,1}$-$a_{2,2}$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-B. Probability contours in the $a_{1,2}$-$a_{2,1}$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\begin{flushleft}
+\caption{\small (LDS2): A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\end{flushleft}
 \label{fig:LDS2}
@@
 
-Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $\mathbf{z}$, since the eigenvalues can be calculated using the quadratic formula:
-\begin{equation}
-\lambda = \frac{(\frac{a_{1,1} + a_{2,2}}{\tau}) \pm \sqrt{(\frac{a_{1,1}+a_{2,2}}{\tau})^2 + 4(\frac{a_{1,2} a_{2,1} - a_{1,1} a_{2,2}}{\tau})}}{2}.
-\end{equation}
-
-Importantly, even though $f(\mathbf{x}; \mathbf{z})$ is analytically available and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}(z \mid \mathcal{X})$ directly.  
-This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we can use EPI to approximate this distribution (Fig. \ref{fig:LDS1}B). 
-We used a real-NVP normalizing flow architecture with three masks, two neural network layers of 50 units per mask, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
-(see Section \ref{methods_NF}).
-
-The effect of $c_0$ on the EPI optimization is shown in Figure \ref{fig:LDS1}C.
-Each EPI optimization was run for enough augmented Lagrangian epochs until convergence, with 1,000 iterations per epoch.  
-We see that low values of $c_0$ (e.g. $10^{-5}$) result in fast entropic growth to the uniform distribution at the outset of optimization. 
-We know this, since the dashed black line indicates the maximum entropy of a distribution with the given bounds. 
-Initial entropic growth can be desireable based on the application, however one should expect a greater number of epochs until convergence compared to a greater value of $c_0$ (e.g. $10^{-2}$).
-
-However, an important drawback of selecting a high `c0` is that you may miss interesting structure in the learned distribution.
-
-Even this relatively simple system has nontrivial (though intuitively sensible) structure in the parameter distribution.  
-To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
-In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} = 0$, dotted black line at
-the standard deviation $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.5$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 1$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
-The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_{1,1}$ and $a_{2,2}$.
-Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
+To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
+\begin{equation}
+\text{imag}(\lambda_1) = \begin{cases}
+                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
+                             0 & \text{otherwise } \\
+                         \end{cases}.
+\end{equation}
+
+When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
+\begin{equation}
+\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
+\end{equation}
+Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
+S2B).
+With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
+This validates the curved structure of the inferred distribution learned through EPI.  
+We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
+Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
+As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figLDS3/figLDS3.pdf}
+\includegraphics[scale=0.5]{figures/figLDS3/figLDS3.pdf}
 \end{center}
-\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
-A. Each dimension of the simulated trajectories throughout time.  
+\begin{flushleft}
+\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+A.
+Each dimension of the simulated trajectories throughout time.
 B  The simulated trajectories in phase space.}
+\end{flushleft}
 \label{fig:LDS3}
 \end{figure}
-
-To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} = 0$, we have
-\begin{equation}
-\text{imag}(\lambda_1) = \begin{cases}
-                             \sqrt{\frac{a_{1,1} a_{2,2} - a_{1,2} a_{2,1}}{\tau}},  & \text{if } a_{1,1} a_{2,2} < a_{1,2} a_{2,1} \\
-                             0 & \text{otherwise } \\
-                         \end{cases}.
-\end{equation}
-
-When $\tau=1$ and $a_{1,1} a_{2,2} > a_{1,2} a_{2,1}$ (center of distribution above), we have the following equation for the other two dimensions:
-\begin{equation}
-\text{imag}(\lambda_1)^2 = a_{1,1} a_{2,2} - a_{1,2} a_{2,1}
-\end{equation}
-Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_{1,1} a_{2,2} - a_{1,2} a_{2,1} = (2 \pi)^2$ assuming $a_{1,1} a_{2,2} = \mathbb{E}_z\left[a_{1,1} a_{2,2}\right]$ (Fig.
-\ref{fig:LDS2}B).
-This validates the curved structure of the inferred distribution learned through EPI.  
-Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. \ref{fig:LDS3}).
 
@@
 \end{center}
+\begin{flushleft}
 \caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
@@
 B. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
+\end{flushleft}
 \label{fig:STG1}

### 2021-01-19 written/scaling/epi_scaling.tex

@@
 \begin{equation*}
-U = \begin{bmatrix} u_1 & u_2 \end{bmatrix}, V = \begin{bmatrix} v_1 & v_2 \end{bmatrix}, J = UV^\top, \text{\hspace{2cm}}
-u_1, u_2, v_1, v_2 \in \left[-1, 1 \right]^N
-\end{equation*}
-\begin{equation*}
+\begin{split}
+U = \begin{bmatrix} u_1 & u_2 \end{bmatrix}, V = \begin{bmatrix} v_1 & v_2 \end{bmatrix} \\ J = UV^\top + g\chi \\
+u_1, u_2, v_1, v_2 \in \left[-1, 1 \right]^N,~~ g = 0.01 \\
 \tau \dot{r} = -r + Jr
+\end{split}
 \end{equation*}
@@
 
-\clearpage
-
 \section{Results}
@@
 \begin{center}
-\includegraphics[scale=.4]{epi_both.png} \\
+\includegraphics[scale=.5]{figs/epi_both.png} \\
 \end{center}
 \caption{\small EPI scales with $z$ to high dimensions. 
-Square - algorithm never converged across all experiments.
 Convergence definitions: 
 EPI (blue) - satisfies all moment constraints, 
-SNPE (orange)- produces at least 1/10,000 parameter samples in the bounds of emergent property (mean +- 0.5), 
+SNPE (orange)- produces at least 2/$n_{\text{train}}$ parameter samples are in the bounds of emergent property (mean +- 0.5), 
 and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.}
@@
 
+\textbf{Figure 1 points}:
+\begin{enumerate}
+\item EPI is capable of scaling to high dimensional parameter spaces.
+\item SMC has many hyperparameters, of which pyABC selects automatically by running some initial diagnostics upon initialization.
+In concurrence with the literacture, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
+\item We searched over many SNPE hyperparameter choices: $n_{\text{train}} \in [2,000, 10,000, 100,000]$ is the number of simulations run per training epoch, and $n_{\text{mades}} \in [2, 3]$ is the number of masked autoregressive density estimators in the deep parameter distribution architecture.
+The greater $n_{\text{train}}$, the longer each epoch will take, but the more likely SNPE may converge during that epoch.
+Greater $n_{\text{mades}}$ increases the flexibility of the deep parameter distribution of SNPE, but slows optimization.
+For the timing plot, we show the fastest among all of these choices, and for the convergence plot, we show the best convergence among all of these choices.
+During optimization, we used $n_{\text{atom}}$=100 atomic proposals as is recommended.
+\end{enumerate}
+
 \begin{figure}[h]
 \begin{center}
-\includegraphics[scale=.3]{EPI_scaling3.png}
+\includegraphics[scale=.75]{figs/StableAmplification_convergence.pdf}
 \caption{\small Emergent property fidelity.  Top row: Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
@@
 
+\textbf{Figure 2 points}:
+\begin{enumerate}
+\item EPI consistently produces the same distribution posterior predictive independent of the dimensionality.
+\item SMC produces a limited variety of parameters due to the nature of its proposal generation algorithm, yet all parameters obtained produce stable amplification.
+\item SNPE's posterior predictive distribution is not necessarily close to the conditioning point, and is very dependent on dimensionality.
+\end{enumerate}
+
 \bibliography{epi_scaling}

### 2021-01-19 epi/SSSN.py
fano to std

@@ -183,26 +183,24 @@ def get_Fano(alpha, sigma_eps, W_mat, N=100, dt=0.0005, T=150, T_ss=100, mu=0.01
         return T_x
     return Fano
 
-def get_Fano_sigma(alpha, W_mat, h, N=100, dt=0.0005, T=150, T_ss=100, mu=0.01, k=100.):
+def get_stddev_sigma(alpha, W_mat, h, N=100, dt=0.0005, T=150, T_ss=100, mu=0.01, k=100.):
     if not (alpha == 'all'):
         alpha_ind = neuron_inds[alpha]
 
     sssn_sim_traj = SSSN_sim_traj_sigma(h, W_mat, N=N, dt=dt, T=T)
-    def Fano(sigma_eps):
+    def get_stddev(sigma_eps):
         if (alpha == 'all'):
             x_t = k*sssn_sim_traj(sigma_eps)[:,:,:4,T_ss:]
         else:
             x_t = k*sssn_sim_traj(sigma_eps)[:,:,alpha_ind,T_ss:]
-        _means = tf.math.reduce_mean(x_t, axis=-1)
-        _vars = tf.square(tf.math.reduce_std(x_t, axis=-1))
-        fano = _vars / _means 
-        vars_mean = tf.reduce_mean(fano, axis=1)
+        stddevs = tf.math.reduce_std(x_t, axis=-1)
+        stddevs_mean = tf.reduce_mean(stddevs, axis=1)
         if (alpha == 'all'):
-            T_x = tf.concat((vars_mean, tf.square(vars_mean - mu)), axis=1)
+            T_x = tf.concat((stddevs_mean, tf.square(stddevs_mean - mu)), axis=1)
         else:
-            T_x = tf.stack((vars_mean, tf.square(vars_mean - mu)), axis=1)
+            T_x = tf.stack((stddevs_mean, tf.square(stddevs_mean - mu)), axis=1)
         return T_x
-    return Fano
+    return get_stddev
 
 def plot_contrast_response(c, x, title='', ax=None, linestyle='-', colors=None, fontsize=14):
     if colors is None:

### 2021-01-19 scripts/SSSN/cpu_drdh.sh
scripts

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH --account=stats
-#SBATCH --job-name=dvdh_epi
+#SBATCH --job-name=drdh_epi
 #SBATCH -c 1
 #SBATCH --time=3:00:00
-#SBATCH --mem-per-cpu=5gb
+#SBATCH --mem-per-cpu=25gb
 
 source activate epi
-python3 dvdh_epi.py --alpha $1 --beta $2 --contrast $3 --logc0 $4 --random_seed $5
+python3 drdh_epi.py --alpha $1 --beta $2 --contrast $3 --logc0 $4 --random_seed $5

### 2021-01-19 scripts/SSSN/ff_epi.py

@@ -9,7 +9,11 @@ from epi.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_Fano
 parser = argparse.ArgumentParser()
 parser.add_argument('--alpha', type=str, default='E') # neuron type
 parser.add_argument('--ind', type=int, default=62) # neuron type
+<<<<<<< HEAD
+parser.add_argument('--lim', type=float, default=0.25) # neuron type
+=======
 parser.add_argument('--lim', type=float, default=0.2) # neuron type
+>>>>>>> f0cc7e04db7b25d4e5bb912dab11d9366a17b44c
 parser.add_argument('--ff_mean', type=float, default=0.05) # neuron type
 parser.add_argument('--ff_std', type=float, default=0.025) # neuron type
 parser.add_argument('--beta', type=float, default=4.) # aug lag hp
@@ -56,7 +60,11 @@ model = Model(name, parameters)
 
 dt = 0.0005
 T = 150
+<<<<<<< HEAD
+N = 50
+=======
 N = 5
+>>>>>>> f0cc7e04db7b25d4e5bb912dab11d9366a17b44c
 
 fano = get_Fano(alpha, sigma_eps, W_mat, N=N, dt=dt, T=T, T_ss=T-50, mu=ff_mean)
 model.set_eps(fano)
@@ -74,9 +82,9 @@ q_theta, opt_data, epi_path, failed = model.epi(
     post_affine=False,
     batch_norm=False,
     bn_momentum=0.0,
-    K=2,
+    K=6,
     N=M,
-    num_iters=200,
+    num_iters=1000,
     lr=1e-3,
     c0=c0,
     beta=beta,

### 2021-01-19 scripts/SSSN/ff_epi_sigma.py

@@ -26,9 +26,15 @@ beta = args.beta
 c0 = 10.**args.logc0
 random_seed = args.random_seed
 
+<<<<<<< HEAD
+if lim is None:
+    if ind == 49:
+        lim = 0.03
+=======
 if lim is not None:
     if ind == 49:
         lim = 0.05
+>>>>>>> f0cc7e04db7b25d4e5bb912dab11d9366a17b44c
     elif ind == 62:
         lim = 0.005
     else:
@@ -56,8 +62,8 @@ parameters = [sigma_eps]
 model = Model(name, parameters)
 
 dt = 0.0005
-T = 100
-N = 5
+T = 150
+N = 100
 
 fano = get_Fano_sigma(alpha, W_mat, h, N=N, dt=dt, T=T, T_ss=T-50, mu=ff_mean)
 model.set_eps(fano)
@@ -75,9 +81,9 @@ q_theta, opt_data, epi_path, failed = model.epi(
     post_affine=False,
     batch_norm=False,
     bn_momentum=0.0,
-    K=2,
+    K=10,
     N=M,
-    num_iters=200,
+    num_iters=2000,
     lr=1e-3,
     c0=c0,
     beta=beta,
@@ -85,7 +91,7 @@ q_theta, opt_data, epi_path, failed = model.epi(
     random_seed=random_seed,
     verbose=True,
     stop_early=True,
-    log_rate=1,
+    log_rate=100,
     save_movie_data=True,
 )
 

### 2021-01-19 scripts/STG_Circuit/cpu_stg_epi.sh

@@ -6,4 +6,4 @@
 #SBATCH --mem-per-cpu=50gb
 
 source activate epi
-python3 stg_epi_fast.py --freq $1 --mu_std $2 --g_el_lb $3 --beta $4 --logc0 $5 --random_seed $6
+python3 stg_epi_fast.py --freq $1 --mu_std $2 --elemwise_fn $3 --num_layers $4 --logc0 $5 --random_seed $6

### 2021-01-19 scripts/STG_Circuit/stg_epi_fast.py

@@ -12,19 +12,28 @@ DTYPE = tf.float32
 parser = argparse.ArgumentParser()
 parser.add_argument('--freq', type=float, default=0.55) # frequency for mu
 parser.add_argument('--mu_std', type=float, default=0.05) # std in mu constraint
+<<<<<<< HEAD
+parser.add_argument('--elemwise_fn', type=str, default="spline") # lower bound on g_el
+parser.add_argument('--num_layers', type=int, default=2) # aug lag hp
+=======
 parser.add_argument('--g_el_lb', type=float, default=0.01) # lower bound on g_el
 parser.add_argument('--beta', type=float, default=4.) # aug lag hp
+>>>>>>> f0cc7e04db7b25d4e5bb912dab11d9366a17b44c
 parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
 parser.add_argument('--random_seed', type=int, default=1)
 args = parser.parse_args()
 
 freq = args.freq
 mu_std = args.mu_std
-g_el_lb = args.g_el_lb
-beta = args.beta
+elemwise_fn = args.elemwise_fn
+num_layers = args.num_layers
 c0 = 10.**args.logc0
 random_seed = args.random_seed
 
+g_el_lb = 4.
+beta = 2.
+
+sigma_I = 1.e-12
 #sleep_dur = np.abs(args.logc0) + random_seed/5. + beta/3.
 #print('short stagger sleep of', sleep_dur, flush=True)
 #time.sleep(sleep_dur)
@@ -35,7 +44,7 @@ g_el = Parameter("g_el", 1, lb=g_el_lb, ub=8.)
 g_synA = Parameter("g_synA", 1, lb=0.01, ub=4.)
 
 # Define model
-name = "STG"
+name = "STG_sigmaI=%.2E" % sigma_I
 parameters = [g_el, g_synA]
 model = Model(name, parameters)
 
@@ -48,7 +57,6 @@ init_params = {'num_keep':500,
                'means':np.array([freq]),
                'stds':np.array([abc_std,])}
 
-sigma_I = 5e-13
 dt = 0.025
 T = 300
 network_freq = NetworkFreq(dt, T, sigma_I, mu)
@@ -58,16 +66,13 @@ model.set_eps(network_freq)
 # 3. Run EPI.
 q_theta, opt_data, epi_path, failed = model.epi(
     mu,
-    arch_type='coupling',
-    num_stages=3,
-    num_layers=2,
+    arch_type="coupling",
[…]

### 2021-01-19 scripts/STG_Circuit/stg_sweep.sh

@@ -4,17 +4,17 @@
 # python scripts independently on different instances?
 for freq in 0.55
   do
-  for mu_std in 0.025 0.05
+  for mu_std in 0.025
   do
-    for g_el_lb in 0.01 4.
+    for elemwise_fn in affine
     do
-      for beta in 4.
+      for num_layers in 2
       do
-        for logc0 in 2
+        for logc0 in 5
         do
-          for rs in 1
+          for rs in {6..15}
           do
-            sbatch cpu_stg_epi.sh $freq $mu_std $g_el_lb $beta $logc0 $rs
+            sbatch cpu_stg_epi.sh $freq $mu_std $elemwise_fn $num_layers $logc0 $rs
           done
         done
       done

### 2021-01-19 scripts/stab_amp/cpu_stab_amp_smc.sh

@@ -2,7 +2,7 @@
 #SBATCH --account=stats
 #SBATCH --job-name=sa_smc
 #SBATCH -c 1
-#SBATCH --time=2:00:00
+#SBATCH --time=6:00:00
 #SBATCH --mem-per-cpu=1gb
 
 source activate epi

### 2021-01-19 scripts/stab_amp/stab_amp_epi.py

@@ -80,8 +80,9 @@ q_theta, opt_data, save_path, failed = M.epi(
     batch_norm=False,
     lr=1e-3, 
     N = 100,
+    num_iters = 1000,
     c0=c0,
-    beta = 10.,
+    beta = 4.,
     nu=1.,
     K=num_epochs,
     verbose=True,

### 2021-01-19 scripts/stab_amp/stab_amp_epi_sweep.sh

@@ -2,11 +2,11 @@
 
 # This shell script does it linearly, but could you run each of these 
 # python scripts independently on different instances?
-for N in 100
+for N in 2 5 10 15 20 25
 do
   for c0 in 3
   do
-    for rs in 2 3 4 5
+    for rs in {1..5}
     do
       echo "Running N = $N, c0 = $c0, rs = $rs"
       sbatch cpu_stab_amp_epi.sh $N $c0 $rs

### 2021-01-19 scripts/stab_amp/stab_amp_smc.py

@@ -29,6 +29,17 @@ time.sleep(sleep_time)
 print('Running SNPE on RNN conditioned on stable amplification with:')
 print('N = %d, seed=%d' % (N, rs))
 
+base_path = os.path.join("data", "smc")
+save_dir = "SMC_RNN_stab_amp_N=%d_rs=%d" % (N, rs)
+
+save_path = os.path.join(base_path, save_dir)
+if not os.path.exists(save_path):
+    os.makedirs(save_path)
+
+if os.path.exists(os.path.join(base_path, save_dir, "optim.pkl")):
+    print("SMC optimization already run. Exitting.")
+    exit()    
+
 def model(parameter):
     u1 = np.array([parameter["U%d1" % i] for i in range(1, N+1)])
     u2 = np.array([parameter["U%d2" % i] for i in range(1, N+1)])
@@ -102,13 +113,6 @@ optim = {'history':history,
         'converged':converged,
         'total_sims':total_sims,
         }
-
-base_path = os.path.join("data", "smc")
-save_dir = "SMC_RNN_stab_amp_N=%d_rs=%d" % (N, rs)
-
-save_path = os.path.join(base_path, save_dir)
-if not os.path.exists(save_path):
-    os.makedirs(save_path)
 
 print('Saving', save_path, '...')
 with open(os.path.join(base_path, save_dir, "optim.pkl"), "wb") as f:

### 2021-01-19 epi/SSSN.py
SSSN refactor

@@ -64,6 +64,8 @@ tau = tau[None,None,:,None]
 tau_noise = 0.005*np.array([1., 1., 1., 1.], np.float32)
 tau_noise = tau_noise[None,None,:,None]
 
+sigma_fac = np.sqrt(1. + (tau/tau_noise))
+
 # Dim is [M,N,|r|,T]
 def SSSN_sim_traj(sigma_eps, W_mat, N=1, dt=0.0005, T=150, x_init=None):
     sigma_eps = sigma_eps[:,None,:,None]
@@ -85,7 +87,7 @@ def SSSN_sim_traj(sigma_eps, W_mat, N=1, dt=0.0005, T=150, x_init=None):
             B = tf.random.normal(eps.shape, 0., np.sqrt(dt))
 
             dx = (-x + (tf.nn.relu(tf.matmul(W, x) + h + eps)**n)) / tau
-            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*B/dt)) / tau_noise
+            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*sigma_fac*B/dt)) / tau_noise
 
             return tf.concat((dx, deps), axis=2)
 
@@ -111,7 +113,7 @@ def SSSN_sim_traj_sigma(h, W_mat, N=1, dt=0.0005, T=150):
             B = tf.random.normal(eps.shape, 0., np.sqrt(dt))
 
             dx = (-x + (tf.nn.relu(tf.matmul(W, x) + h + eps)**n)) / tau
-            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*B/dt)) / tau_noise
+            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*sigma_fac*B/dt)) / tau_noise
 
             return tf.concat((dx, deps), axis=2)
 
@@ -138,7 +140,7 @@ def SSSN_sim(sigma_eps, W_mat, N=1, dt=0.0005, T=150):
             B = tf.random.normal(eps.shape, 0., np.sqrt(dt))
 
             dx = (-x + (tf.nn.relu(tf.matmul(W, x) + h + eps)**n)) / tau
-            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*B/dt)) / tau_noise
+            deps = (-eps + (np.sqrt(2.*tau_noise)*sigma_eps*sigma_fac*B/dt)) / tau_noise
             
             return tf.concat((dx, deps), axis=2)
             

### 2021-01-19 scripts/SSSN/ff_epi_sigma.py

@@ -3,15 +3,15 @@ import argparse
 import numpy as np
 import tensorflow as tf
 from epi.models import Parameter, Model
-from epi.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_Fano_sigma
+from epi.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_stddev_sigma
 
 # Parse script command-line parameters.
 parser = argparse.ArgumentParser()
 parser.add_argument('--alpha', type=str, default='E') # neuron type
 parser.add_argument('--ind', type=int, default=62) # neuron type
-parser.add_argument('--lim', type=float, default=None) # neuron type
-parser.add_argument('--ff_mean', type=float, default=1.) # neuron type
-parser.add_argument('--ff_std', type=float, default=0.025) # neuron type
+parser.add_argument('--lim', type=float, default=0.005) # neuron type
+parser.add_argument('--f_mean', type=float, default=5.) # neuron type
+parser.add_argument('--f_std', type=float, default=0.25) # neuron type
 parser.add_argument('--beta', type=float, default=4.) # aug lag hp
 parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
 parser.add_argument('--random_seed', type=int, default=1)
@@ -20,25 +20,11 @@ args = parser.parse_args()
 alpha = args.alpha
 ind = args.ind
 lim = args.lim
-ff_mean = args.ff_mean
-ff_std = args.ff_std
+f_mean = args.f_mean
+f_std = args.f_std
 beta = args.beta
 c0 = 10.**args.logc0
 random_seed = args.random_seed
-
-<<<<<<< HEAD
-if lim is None:
-    if ind == 49:
-        lim = 0.03
-=======
-if lim is not None:
-    if ind == 49:
-        lim = 0.05
->>>>>>> f0cc7e04db7b25d4e5bb912dab11d9366a17b44c
-    elif ind == 62:
-        lim = 0.005
-    else:
-        raise NotImplementedError()
 
 contrast = 0.5
 W_mat = load_SSSN_variable('W', ind=ind)
@@ -57,7 +43,7 @@ ub = lim*np.ones((D,))
 sigma_eps = Parameter("sigma_eps", D, lb=lb, ub=ub)
 
 # Define model
[…]

### 2021-02-02 written/eLifeRevision/Bittner_eLife2020.tex
latest tex

@@
 \begin{center}
-\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}
+\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.2]{figures/fig2/fig2.pdf}
+\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.48]{figures/fig3/fig3.pdf}
+\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
 \end{center}
@@
 
-Here, we are interested in the scalability of EPI in number of parameters ($|\mathbf{z}|$).  We consider rank-2 RNN with N neurons of connectivity
+Here, we are interested in the scalability of EPI in number of parameters $|\mathbf{z}|$.  We consider rank-2 RNN with N neurons of connectivity
 \begin{equation}
@@
 \begin{center}
-\includegraphics[scale=1.4]{figures/fig4/fig4.pdf}
+\includegraphics[scale=.8]{figures/fig4/fig4.pdf}
 \end{center}
@@
 This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
-Normalizing flow architectures for deep probability distributions used in EPI are specified by the number of coupling stages, neural network layers and units per layer for conditioning functions.
+
+In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$.
+We specify the architecture used in each application by the number of Real-NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.
 
 \subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
+To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented Lagrangian method.
+The following objective is minimized:
 \begin{equation} \label{eq:AL}
@@
 \end{equation}
-where $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
-These Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
-Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
+where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
+The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
+Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
+Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
 
@@
 The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
-Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.
+Throughout the study, $\gamma = 0.25$, while $\beta$ was chosen to be either 2 or 4.  The batch size of EPI also varied according to application.
 
@@
 
-To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
-The algorithm has converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
+To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we assess whether each individual constraint on the means and variances of $f(\mathbf{x}; \mathbf{z})$ is satisfied.
+We consider the EPI to have converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
 This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
-The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made from $k$ resamplings of $\mathbf{z}$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.
+The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented Lagrangian epoch.
 
@@
 \begin{center}
-\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
+\includegraphics[scale=0.8]{figures/figLDS1/figLDS1.pdf}
 \end{center}
@@
 
-For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 25 units per layer.
+For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
 The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$. 
 We did not include $g_{\text{synA}} < 0.01$, since conductances that low make the circuit simulations numerically unstable.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, set $\nu = 0.25$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
+We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, $\beta = 2$, $N_{\text{test}} = 100$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
 This initialization had much greater entropy and a different emergent property than the returned EPI posterior.
@@
 \begin{center}
-\includegraphics[scale=0.85]{figures/figSC1/figSC1.pdf}
+\includegraphics[scale=0.7]{figures/figSC1/figSC1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=1.3]{figures/figSC2/figSC2.pdf}
+\includegraphics[scale=.7]{figures/figSC2/figSC2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figures/figSC3/figSC3.pdf}
+\includegraphics[scale=1.]{figures/figSC3/figSC3.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.75]{figures/figSC4/figSC4.pdf}
+\includegraphics[scale=0.6]{figures/figSC4/figSC4.pdf}
 \end{center}

### 2021-02-02 epi/SSSN.py
core epi func updates

@@ -212,7 +212,6 @@ def plot_contrast_response(c, x, title='', ax=None, linestyle='-', colors=None, 
         fig, ax = plt.subplots(1,1)
     for i in range(4):
         ax.plot(100*c, x[:,i], linestyle, c=colors[i], lw=4)
-    ax.set_ylim([0., 80])
     ticksize = fontsize-4
     ax.set_xlabel('contrast (%)', fontsize=fontsize)
     ax.set_ylabel('rate (Hz)', fontsize=fontsize)

### 2021-02-02 epi/models.py

@@ -619,7 +619,8 @@ class Model(object):
             nf = self._df_row_to_nf(df_row)
             aug_lag_hps = self._df_row_to_al_hps(df_row)
             best_k, converged, best_H = self.get_convergence_epoch(
-                    init_params, nf, mu, aug_lag_hps, alpha=0.05, nu=nu)
+                    init_params, nf, mu, aug_lag_hps, 
+                    alpha=0.05, nu=nu)
             best_Hs.append(best_H)
             convergeds.append(converged)
             best_ks.append(best_k)
@@ -1209,6 +1210,7 @@ class Model(object):
         optimizer = tf.keras.optimizers.Adam(aug_lag_hps.lr)
         checkpoint = tf.train.Checkpoint(optimizer=optimizer, model=nf)
         ckpt_dir, exists = self.get_epi_path(init_params, nf, mu, aug_lag_hps)
+        print('ckpt_dir', ckpt_dir)
         if not exists:
             return None
         ckpt_state = tf.train.get_checkpoint_state(ckpt_dir)
@@ -1232,7 +1234,8 @@ class Model(object):
         return q_theta
 
     def get_convergence_epoch(
-        self, init_params, nf, mu, aug_lag_hps, alpha=0.05, nu=0.1, mu_test=None, sampling_seed=0,
+        self, init_params, nf, mu, aug_lag_hps, alpha=0.05, nu=1., 
+        mu_test=None, start_k=0, sampling_seed=0,
     ):
 
         np.random.seed(sampling_seed)
@@ -1243,6 +1246,7 @@ class Model(object):
         else:
             _mu = np_column_vec(mu).astype(np.float32).T
         N_test = int(nu * aug_lag_hps.N)
+        print(N_test, _mu)
 
         optimizer = tf.keras.optimizers.Adam(aug_lag_hps.lr)
         checkpoint = tf.train.Checkpoint(optimizer=optimizer, model=nf)
@@ -1260,12 +1264,15 @@ class Model(object):
         best_k = None
         best_H = None
         converged = False
-        for k in range(num_ckpts):
+        for k in range(start_k,num_ckpts):
             status = checkpoint.restore(ckpts[k])
             status.expect_partial()
 
             m = _mu.shape[1]
             z, log_q_z = nf(self.M_test * N_test)
+            H = -np.mean(log_q_z.numpy())
+            if converged and H < best_H:
+                continue
             T_x = self.eps(z)
[…]

### 2021-02-02 epi/util.py

@@ -124,14 +124,20 @@ def np_column_vec(x):
     return x
 
 
-def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1.):
+def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1., check_last_k=None):
     paths = epi_df['path'].unique()
     best_Hs = []
     best_ks = []
     df_rows = []
+    print('num paths', len(paths))
+    any_converged = False
     for path in paths:
         epi_df2 = epi_df[epi_df['path'] == path]
         print('path:', path)
+        if check_last_k is None:
+            start_k = 0
+        else:
+            start_k = int(epi_df2['k'].max()) - check_last_k + 1
         df_row = epi_df2.iloc[0]
         init = df_row['init']
         init_params = {"mu":init["mu"], "Sigma":init["Sigma"]}
@@ -144,15 +150,22 @@ def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1.):
             aug_lag_hps, 
             alpha=alpha, 
             nu=nu,
+            start_k=start_k
         )
+        if (not any_converged) and converged:
+            any_converged = True
         best_Hs.append(best_H)
         best_ks.append(best_k)
         df_rows.append(df_row)
+        print('best k', best_k)
+        print('best H', best_H)
         
     bestHs = np.array(best_Hs)
     best_ks = np.array(best_ks)
 
     best_Hs = np.array([x if x is not None else np.nan for x in best_Hs])
+    if not any_converged:
+        return None, None, None
     ind = np.nanargmax(best_Hs)
 
     path = paths[ind]
@@ -163,6 +176,38 @@ def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1.):
     dist = model._get_epi_dist(best_k, init_params, nf, mu, aug_lag_hps)
     
     return dist, path, best_k
+
+def get_conditional_mode(dist, ind, val, z0=None, lr=1e-6, num_steps=100):
+    if z0 is None:
+        z0 = (dist.nf.lb + dist.nf.ub) / 2.
+        z0[ind] = val
+    z = tf.Variable(initial_value=z0[None,:], dtype=tf.float32, trainable=True)
+    
+    log_q_z = dist.log_prob(z.numpy())
+    
+    zs = [z[0].numpy()]
+    log_q_zs = [log_q_z]
[…]

### 2021-02-02 epi/SC_Circuit_4.py
get rid of SC draft directory

@@ -1,6 +1,9 @@
 import tensorflow as tf
 import numpy as np
 import scipy
+from scipy.special import expit
+import matplotlib.pyplot as plt
+import os
 
 DTYPE = tf.float32
 
@@ -172,6 +175,7 @@ C_opto = 4
 I_opto = tf.concat((I_LP, I_LA, I_LP, I_LA), axis=2)
 
 def SC_sim_opto(strength, period):
+    print('str', strength)
     eta = np.ones((T, 1, C_opto, 1, 1), dtype=np.float32)
     if period == 'delay':
         eta[np.logical_and(0.8 <= t, t <= 1.2), :, 2:, :, :] = strength
@@ -258,6 +262,19 @@ def SC_acc_diff(strength, period):
 
     return _SC_acc_diff
 
+def z_to_W(z):
+    sW = z[:,0]
+    vW = z[:,1]
+    dW = z[:,2]
+    hW = z[:,3]
+
+    Wrow1 = tf.stack([sW, vW, dW, hW], axis=1)
+    Wrow2 = tf.stack([vW, sW, hW, dW], axis=1)
+    Wrow3 = tf.stack([dW, hW, sW, vW], axis=1)
+    Wrow4 = tf.stack([hW, dW, vW, sW], axis=1)
+    W = tf.stack([Wrow1, Wrow2, Wrow3, Wrow4], axis=1)
+    return W
+
 def get_schur_eigs(W):
     # returns 
     T, Z = scipy.linalg.schur(W)
@@ -278,3 +295,42 @@ def get_schur_eigs(W):
     #print(z_inds)
     #print(T)
     return eigs
+
+c_LP = '#3B8023'
+c_LA = '#EA8E4C'
+c_RA = '#F4C673'
+c_RP = '#81C176'
+
+def plot_SC_responses(v_t, fname, figsize=(6,4)):
+    M = v_t.shape[1]
+    T_x = np.mean(expit(100.*(v_t[-1,:,:,0,:] - v_t[-1,:,:,3,:])), axis=2)
+    
+    percfont = {'family': 'arial',
+                'weight': 'light',
+                'size': 20,
+            }
+    neuron_labels = ['LP', 'LA', 'RA', 'RP']
+    colors = [c_LP, c_LA, c_RA, c_RP]
+    C_titles = ['Pro, left trials', 'Anti, left trials']
+    for m in range(M):
[…]

### 2021-02-02 scripts/SSSN/ff_epi_sigma.py
SSSN cleanup

@@ -49,7 +49,7 @@ model = Model(name, parameters)
 
 dt = 0.0005
 T = 150
-N = 100
+N = 25
 
 stddev = get_stddev_sigma(alpha, W_mat, h, N=N, dt=dt, T=T, T_ss=T-50, mu=f_mean)
 model.set_eps(stddev)

### 2021-02-02 scripts/STG_Circuit/stg_hessian.py
final cleanup

@@ -1,0 +1,85 @@
+import os
+import argparse
+import numpy as np
+import tensorflow as tf
+from epi.models import Parameter, Model
+import time
+import matplotlib.pyplot as plt
+from epi.STG_Circuit import Simulate, Simulate_all, NetworkFreq
+from epi.util import pairplot, plot_T_x
+
+# 1. Specify the V1 model for EPI.
+D = 2
+g_el = Parameter("g_el", 1, lb=4., ub=8.)
+g_synA = Parameter("g_synA", 1, lb=0.01, ub=4.)
+
+# Define model
+name = "STG"
+parameters = [g_el, g_synA]
+model = Model(name, parameters)
+
+freq = 0.55
+
+mu = np.array([freq])
+
+def hessian(z, f):
+    """Calculates the Hessian.
+
+    :param z: Samples from distribution.
+    :type z: np.ndarray
+    :param f: Function to compute hessian of
+    :type f: func
+    :returns: Hessian of log probability with respect to z.
+    :rtype: np.ndarray
+    """
+    z = _set_z_type(z)
+    z = tf.Variable(initial_value=z, trainable=True)
+    hess_z = _hessian(z, f)
+    del z  # Get rid of dummy variable.
+    return hess_z.numpy()
+
+def _hessian(z, f):
+    with tf.GradientTape(persistent=True) as tape:
+        f_z = f(z)
+        st = time.time()
+        print('calculating the gradient')
+        dfdz = tape.gradient(f_z, z)
+        print('calculating the gradient done')
+        print(time.time() - st)
+        print(dfdz)
+    return tape.batch_jacobian(dfdz, z)
+
+def _set_z_type(z):
+    if type(z) is list:
+        z = np.ndarray(z)
+    z = z.astype(np.float32)
+    return z
+
+dt = 0.025
+T = 300
+sigma_I = 0.
+
+network_freq = NetworkFreq(dt, T, sigma_I, mu)
+Ds = [param.D for param in model.parameters]
+def f(z):
+    ind = 0
+    zs = []
+    for D in Ds:
+        zs.append(z[:, ind : (ind + D)])
+        ind += D
+    return network_freq(*zs)
+
+@tf.function
+def temp(z):
+    T_x = f(z)
+    return T_x[:,0]
+z_star = np.array([[6.1071167, 1.3093035]], dtype=np.float32)
+
+start_time = time.time()
+print('Calculating hessian')
[…]

### 2021-02-02 written/TCM/TCM_report.tex

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\usepackage{hyperref}
\usepackage{soul}
\usepackage{natbib}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}

\newcommand{\myname}{Sean Bittner}
\newcommand{\myandrew}{srb2201@columbia.edu}
\newcommand{\myhwnum}{12}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}
\setlength{\bibsep}{0pt plus 0.3ex}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%

\pagestyle{fancyplain}
\rhead{\fancyplain{}{\myname\\ \myandrew}}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Thesis committee report} \\
Sean Bittner, January 16, 2021
\end{center}

\textbf{Summary}:
During my last thesis committee meeting in May, 2020, I presented progress on the emergent property inference (EPI) project.  
To refresh, EPI is a likelihood-free inference (LFI) method, that uses deep probability distributions (normalizing flows) to obtain flexible posterior approximations.  
I presented the EPI method and my work on neuroscientific applications of EPI, which come from this manuscript \cite{bittner2019interrogating}.
Our conclusion was that I should focus on a.) improving the science done with this method and b.) providing a comparative analysis between EPI and other LFI techniques.
The advancements I've made since our last meeting are listed below.

\underline{Advancements since May, 2020}:
\begin{enumerate}
\item \textbf{STG}: This introductory example based on an STG subcircuit \cite{gutierrez2013multiple} has been revisited: we introduced stochasticity to make a technical point, and changed the emergent property to improve the pedagogical value of the analysis.
\item \textbf{V1}: We now focus on how parameters of noise govern variability in the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have inhibitory multiplicity.
\item \textbf{SC}: By request from our collaborators, we changed the modeling of network responses from winner-take-all to comparative decision making.
Intriguingly, EPI inferred connectivities postdict optogenetic response properties published by the Brody Lab.
\item \textbf{RNNs}: We infer connectivities of low-rank RNN's exhibiting stable amplification using criteria defined in \cite{bondanelli2020coding}.  
Here, we demonstrate the superior scaling properties of EPI compared to sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2020training}.
\end{enumerate}

Each model in the paper was re-analyzed using the new {\color{blue} \href{https://github.com/cunningham-lab/epi}{epi} software package}, which takes advantage of modern advancements in normalizing flows.

In September, I had the opportunity to present our work on EPI at the Bernstein Conference Workshop on ``Inferring and Testing Optimality in Perception and Neurons".
Upon feedback from the committee and some final touches, we will resubmit our manuscript to eLife.

\textbf{Professional ambitions}:
This past summer (June-Aug 2020) I had a research internship at Facebook Reality Labs (with the CTRL-Labs Science Team).
I used recently developed techniques from automatic speech recognition to substantially improve EMG typing accuracy.
I'm excited about the challenging research problems in BCI development, and I've applied for a Research Scientist position at Facebook with the CTRL-Labs group.
I will be interviewed soon, the start date would be flexible, and my preferred graduation date is still May 2021.
Here is my proposal for the thesis outline.

\textbf{Proposed thesis outline} \\
\underline{Title}: Bridging data, theory, and experiment in computational neuroscience. \\
\underline{Chapter 1}: Understanding movement generation using normative models: \\ 
$~~~~~~~~~~~~~~~~~~~$ tangling \cite{russo2018motor} and divergence \cite{russo2020neural}. \\
\underline{Chapter 2}: Amortizing statistical inference in exponential family models of neural activity \cite{bittner2019approximating}. \\
\underline{Chapter 3}: Inference in theoretical models of neural computation \cite{bittner2019interrogating}. \\

In our meeting, I'll summarize the latest advancements to the EPI project, and describe my contributions to the work listed in the thesis outline.

\bibliography{tcm}
\bibliographystyle{unsrt}
\vspace{3mm}

\end{document}

### 2021-02-02 written/eLifeRevision/Bittner_eLife2020.tex
methods edits and figure resizing
intro edits from mid-Jan

@@
 %However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-We bring the use of deep generative models for probabilistic inference to bear on this problem, learning distributions of parameters that produce the specified properties of computation.
-By learning parameter distributions that produce computations -- an emergent property, we introduce a novel methodology that is particularly well-suited to the stochastic dynamical systems models predominant in our field of theoretical neuroscience.
+We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
+Our novel method solves the inverse problem by identifying the full space of parameters producing the emergent property.
 %Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
@@
 We then use it to reveal the key factors of variability in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and scale inference of large low-rank RNN's exhibiting stable amplification.
-While much use of deep learning in theoretical neuroscience focuses on drawing analogies between optimized neural architectures and the brain, this work illustrates how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
+This work illustrates how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
 %More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
@@
 \section{Introduction}
-The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing \cite{abbott2008theoretical}.
-A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations.
-The function of this system is governed by the choice of model parameters, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about these particular parameter configurations?  
+The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
+A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
+The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
+The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
 %The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model.
-
-Consider the idealized practice: one carefully designs a model and analytically derives how model parameters govern the computation.
+The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.
+
+Consider the idealized practice: one carefully designs a model and analytically derives how computational properties determine model parameters.
 Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
@@
 %On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-Alternatively, we can gain insight into these complex models by identifying all of the parameters consistent with the emergent phenomena of interest.
-By examining the structure of the full space of possible parameters, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-
-The preferred formalism for parameter identification in science, statistical inference, has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
+Alternatively, we can gain insight into these complex models by identifying the full distribution of parameters consistent with specified emergent phenomena.
+By solving the inverse problem in this way, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+
+The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
 %Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
 However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
-Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which a set of reasonable parameters estimates is obtained via simulation and rejection.
+Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
 
@@
 Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
-In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
+In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
 %One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
 One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating parameters (sometimes treated as hidden states) in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
-These works have used deep neural networks to expand the expressivity and accuracy of statistical models of neural data \cite{paninski2018neural}.
+Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
+These works have used deep neural networks to expand the domain of neural data sets amenable to statistical modeling \cite{paninski2018neural}.
 
@@
 First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
-Neuroscientists therefore resort to using approximate Bayesian computation, which requires a rejection heuristic, scales poorly, and only obtains sets of non-rejected parameters lacking probabilities.
-Second, is the undesirable trade-off between the flexibility and sampling speed of the approximated posterior distributions.
-Sampling-based approaches to statistical inference (e.g. ABC and Markov chain Monte Carlo (MCMC)) have flexibility in approximation, but must be executed continually for increasing samples.
-While variational approaches often result in fast sampling and sensitivity measurements post-optimization, existing approaches have relied on simplified classes of distributions.  
-These simple distributions (e.g. mean-field gaussians) restrict the flexibility of the posterior approximation.
-And third, one can never assume what inferred model parameters may predict.
-This is well understood when considering Box's loop and the role of posterior predictive checks in the development and critique of scientific models \cite{gelman2013philosophy, blei2014build}.
-Uncertainty about the properties of inferred model predictions introduce a conceptual degree of freedom to the inverse problem that may be unnecessary and undesirable given the scientific motivation.
+Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
+Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
+Sampling-based inference approaches (e.g.
+ABC and Markov chain Monte Carlo (MCMC) \cite{hastings1970monte, metropolis1953equation}) confer flexible approximations, yet scale poorly in number of parameters.
+While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
+Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
+Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
 
@@
 First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
-Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
+Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
 Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
+Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
 Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
@@
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-First, we use EPI to understand the characteristics of noise that govern Fano factor in a stochastic four neuron-type model of primary visual cortex.
+First, we use EPI to understand the characteristics of noise that govern variability in a stochastic four neuron-type model of primary visual cortex.
 %Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
@@
 %Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
-Third, we emphasize the methodological advancement of EPI by inferring high-dimensional distributions of RNN connectivities exhibiting stable amplification.
+Third, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
 These results point to the value of deep inference for the interrogation of biologically relevant models.
@@
 \begin{center}
-\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
+\includegraphics[scale=0.8]{figures/fig2/fig2.pdf}
 \end{center}
@@
 This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
-
 In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
@@
 \begin{center}
-\includegraphics[scale=.7]{figures/figSC2/figSC2.pdf}
+\includegraphics[scale=0.7]{figures/figSC2/figSC2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=1.]{figures/figSC3/figSC3.pdf}
+\includegraphics[scale=1.0]{figures/figSC3/figSC3.pdf}
 \end{center}

### 2021-02-03 written/eLifeRevision/Bittner_eLife2020.tex
figure reformatting

@@
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-First, we use EPI to understand the characteristics of noise that govern variability in a stochastic four neuron-type model of primary visual cortex.
+First, we use EPI to understand the characteristics of noise across mulitple neuron-type populations that govern variability in a model of primary visual cortex.
 %Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-Second, we discover connectivity patterns in superior colliculus resilient to optogenetic perturbation by using EPI to condition on rapid task switching.
+Then, we use EPI to infer multiple regimes of superior colliculus connectivity that perform rapid task switching.
 The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{litwin2016inhibitory, duan2018collicular}.
 %Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.  
-Third, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
+Finally, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
 These results point to the value of deep inference for the interrogation of biologically relevant models.
@@
 The evolution of this model's activity $\mathbf{x}(t)$ is predicated by differential equations.
-\textbf{B}.
-Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic.
-Spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
+\textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
+In this example, spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
 \textbf{C}. The emergent property of intermediate hub frequency, in which the hub neuron fires at a rate between the fast and slow frequencies.
+This emergent property is defined by a mean and variance on the emergent property statistic.
 Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution (Panel E).
-\textbf{D}.
-For a choice of model and emergent property, emergent property inference (EPI) learns a distribution of the model parameters $\mathbf{z} = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing intermediate hub frequency.
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ distributed as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+\textbf{D}. For a choice of model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$.
 In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced.
-\textbf{E}. The EPI distribution of STG model parameters producing intermediate hub frequency.
+The EPI posterior distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$. 
+\textbf{E}. The EPI posterior producing intermediate hub frequency.
 Samples are colored by log probability density.  
-Distribution contours of hub neuron frequency from mean of .55 Hz are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Frequencies are averages over the stochasticity of the model.
+Distribution contours of average hub neuron frequency from mean of .55 Hz are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
 Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-Simulated activity is shown for three samples (stars). 
 %(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
 \textbf{F} Simulations from parameters in E.  
-(Top) The predictive distribution of the posterior obeys the constraints stipulated by the emergent property.
+(Top) The predictive distribution of the posterior obeys the emergent property.
 The black and gray dashed lines show the mean and two standard deviations according the emergent property, respectively.
@@
 In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}E yellow star), and how network syncing decays based on changes away from the mode.
-The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of unified intermediacy (Fig. \ref{fig:STG}B solid ($v_1$) and dashed  ($v_2$) black arrows).
+The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of intermediate hub frequency (Fig.
+\ref{fig:STG}E solid ($v_1$) and dashed  ($v_2$) black arrows).
 Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig.
-\ref{fig:STG}B contours) and have diminished or preserved network syncing, respectively (Fig.
-\ref{fig:STG}F activity traces, Fig.
-S TODO) (see Section \ref{methods_STG}).
+\ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
+\ref{fig:STG}F activity traces) (see Section \ref{methods_STG}).
 
@@
 Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2016density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
+This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
 This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
@@
 
-EPI optimizes the weights and biases $\bm{\theta}$ of the deep neural network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
+EPI optimizes the weights and biases $\bm{\theta}$ of the deep network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
 The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figLDS2/figLDS2.pdf}
+\includegraphics[scale=0.8]{figures/figLDS2/figLDS2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figLDS3/figLDS3.pdf}
+\includegraphics[scale=0.8]{figures/figLDS3/figLDS3.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figures/figSTG1/figSTG1.pdf}
+\includegraphics[scale=0.8]{figures/figSTG1/figSTG1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=.5]{figures/figV1_2/figV1_2.pdf}
+\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figSC1/figSC1.pdf}
+\includegraphics[scale=0.8]{figures/figSC1/figSC1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figSC2/figSC2.pdf}
+\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=1.0]{figures/figSC3/figSC3.pdf}
+\includegraphics[scale=0.8]{figures/figSC3/figSC3.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figures/figSC4/figSC4.pdf}
+\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.6]{figures/figSC5/figSC5.pdf}
+\includegraphics[scale=0.8]{figures/figSC5/figSC5.pdf}
 \end{center}

### 2021-02-03 written/eLifeRevision/Bittner_eLife2020.tex
V1 updates

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI reveals how noise across neural population types governs Fano factor in a stochastic inhibition stabilized network} \label{results_V1}
+\subsection{EPI reveals how neuron-type specific noise governs variability in the stochastic stabilized supralinear network} \label{results_V1}
 
@@
 \vspace{-1cm}
-\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
+\caption{\small Emergent property inference in the stochastic stabilized supralinear network (SSSN)
 \textbf{A}.
@@
 Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
-Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
+Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
 Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
-\textbf{B}. Responses of the deterministic model ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
-\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
-\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on realistic E-population Fano factors.
-The posterior predictive distribution is show on the bottom-left.
-and the mode of the distribution is starred.
-\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
-Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
+\textbf{B}. Steady-state responses of the SSN model (deterministic, $\bm{\sigma}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
+\textbf{C}.
+Transient network responses of the SSSN model at 50 \% contrast.
+(Left) Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise realization.
+(Right) Mean (solid line) and standard deviation (shading) of responses.
+\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on E-population variability.
+The posterior predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
+\textbf{E}.
+(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posteriors $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+Each black dot shows the mode at each $\sigma_P$.
 The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
-(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
-\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
-\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
+\textbf{F}.
+The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each posterior $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
 }
@@
 Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we use EPI to characterize the properties of slow noise in a stochastic version of this model, which result in biologically realistic responses.
-
-We considered the contrast response of a nonlinear dynamical V1 circuit model (Fig.
+Here, we use EPI to analyze V1 models of greater complexity in order to characterize properties of slow noise governing circuit variability.
+
+We considered the response properties of a nonlinear dynamical V1 circuit model (Fig.
 \ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via nonlinearity $\phi = \left[\right]^2_+$ (see Section \ref{methods_V1}).
-The circuit model evolves form an initial condition $\mathbf{x}(0) \sim \mathcal{U}(\left[10, 25\right])$ with time constant $\tau = 1$ms according to a contrast-dependent input $\mathbf{h}$ and  slow noise $\bm{\epsilon}$ of time constant $\tau_{\text{noise}} = 5$ms.
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical}  generalized to have inhibitory multiplicity 
+Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via supralinear nonlinearity $\phi = \left[\right]^2_+$.
+The input is also comprised of a slow noise component $\bm{\epsilon} \sim OU\left(\tau_{\text{noise}}, \Sigma \right)$ of time scale $\tau_{\text{noise}} > \tau$ and covariance $\Sigma = \text{diag}\left(\bm{\sigma^2}\right)$ (see Section \ref{methods_V1})
 \begin{equation}
@@
 \end{equation}
-As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h}_b$ via $\mathbf{h}_c$
-\begin{equation}
-\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c,
-\end{equation}
-where $h_{c,E}, h_{c,P}  > 0$ and $h_{c,S}, h_{c,V}  = 0$.
-In this analysis, we fixed $W$,$\mathbf{h}_b$, and $\mathbf{h}_c$ to values fit to mean contrast responses in mice with the deterministic model \cite{palmigiano2020structure} ($\bm{\epsilon}=\mathbf{0}$, Fig.
-\ref{fig:V1}B, see Section \ref{methods_V1}).
-At all contrasts, the E-population of this SSSN is unstable without recurrent inhibitory feedback.
-At 50\% contrast, only the P-population exhibits the paradoxical effect (\ref{fig:V1}C, Fig. \ref{fig:V1_1}), so the network is P-stabilized.
-
-The slow noise of the SSSN is an Ornstein-Uhlenbeck process
-\begin{equation}
-\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB,
-\end{equation}
-parameterized by $\sigma_\alpha$, which can be different for each neuron type,
-\begin{equation}
-\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top.
-\end{equation} 
-For this SSSN, we are interested in the parameters of slow noise that produce realistic stochastic fluctuations.
-Here, we quantify this emergent property as having an excitatory population Fano factor near 1:
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have inhibitory multiplicity, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
+Both modeling advancements introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
+Here, we consider fixed weights $W$ and input $\mathbf{h}$ according to a fit of the deterministic model to contrast responses \cite{palmigiano2020structure} (Fig.
+\ref{fig:V1}B), and study the effect of noise parameterization $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on fluctuations at 50\% contrast.
+
+For this SSSN, we are interested in how noise variability across neural populations governs stochastic fluctuations in the E-population.
+Here, we quantify different levels $y$ of E-population variability with the emergent property
 \begin{equation}\label{eq:EP_V1}
 \begin{split}
-\mathcal{X} ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 1 \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} F_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  0.125^2,
+\mathcal{X}(y) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ y \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2,
 \end{split}
 \end{equation}
-where $F_\alpha(\mathbf{x}; \mathbf{z})$ is the Fano factor of the $\alpha$-population.
-
-We ran EPI to obtain a posterior $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, where each parameter $\mathbf{z}$ produces biologically realistic levels of E-population variability (Fig.
-\ref{fig:V1}D).
-From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig. \ref{fig:V1}D, top-left), we can see that $F_E(\mathbf{x}; \mathbf{z})$ is sensitive to the combination of $\sigma_E$ and $\sigma_P$.
-In fact, the posterior obtained through EPI offers exactly how this sensitivity changes along this ridge of the posterior (Fig. \ref{fig:V1}E).
-$\sigma_S$ and $\sigma_V$ are degenerate with respect to $F_E(\mathbf{x}; \mathbf{z})$ evidenced by the uniform distribution in those dimensions of the posterior (Fig. \ref{fig:V1}D, bottom-right).
-Together, this posterior indicates a parametric manifold of degeneracy with respect to Fano factor: the ridge visualized in the $\sigma_E$-$\sigma_P$ marginal (Fig.
-\ref{fig:V1_2}) and the dimensions of $\sigma_S$ and $\sigma_V$.
-
-Greater $\sigma_E$ and $\sigma_P$ confer greater Fano factor, and the Fano factors of each neuron-type are strongly correlated across the posterior (Fig \ref{fig:V1}F), showing that Fano factor of each neuron-type can be modulated globally via $\sigma_E$ and $\sigma_P$.
-Furthermore, across the entire posterior distribution of noise parameterizations, we find that when contrast is increased above 50\%, variability is quenched for all neuron types (Fig \ref{fig:V1}G).
-In summary, we used EPI to obtain a posterior of SSSNs producing realistic Fano factors, which allowed degenerate manifold identification via sample visualization, fast sensitivity measurements via Hessian evaluation, and predictions of variability quenching.
+where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
+
+We ran EPI to obtain a posterior distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
+From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
+\ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
+Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions of the posterior (Fig.
+\ref{fig:V1}D, bottom-right).
+Together, these observations imply a parametric manifold of degeneracy with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ in the $\sigma_E$-$\sigma_P$ marginal (Fig.
+\ref{fig:V1}E).
+The dimensions of sensitivity conferred by EPI and this plain visual structure suggest a quadratic relationship in the emergent property statistic $s_E(\mathbf{x}; \mathbf{z})$ and parameters $\mathbf{z}$, which is preserved at a greater level of variability $\mathcal{X}(10\text{ Hz})$ (Fig.
+\ref{fig:V1}E).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ (Fig \ref{fig:V1}F, $p=0$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two posteriors (Fig.
+\ref{fig:V1_3}, $p=.402$).
+
+While a quadratic relationship in $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ is potentially derivable by extending the derivation in Section \ref{methods_V1} to the case of $\tau \neq \tau_{\text{noise}}$, the coefficients in front of each quadratic term would be unruly, and likely escape comprehensible analysis.
+This makes EPI an attractive tool for revealing the characteristics of noise governing variability and for answering other questions in this complex model.
+Intriguingly, this circuit exhibited a paradoxical effect in the P-population, and no other inhibitory types at 50\% contrast (Fig. \ref{fig:V1_3}) implying that the E-population is P-stabilized.
+Future work motivated by our analysis here, may uncover a relationship between the neuron-type mediating stability and the factors governing circuit variability.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI identifies neural mechanisms of flexible task switching} \label{results_SC}
+\subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
 %In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
@@
 Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
-We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.
+We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
 \begin{equation}\label{eq:EP}
@@
 \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 5\%^2 \\ 5\%^2  \end{bmatrix}
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 7.5\%^2 \\ 7.5\%^2  \end{bmatrix}
 \end{split}
 \end{equation}
-A variance of $5\%^2$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.
+A variance of $7.5\%^2$  in each task will confer a posterior producing performances ranging from about $60\%-90\%$, allowing us to examine the properties of connectivity that yield better performance.
 
@@
 For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$. 
-We did not include $g_{\text{synA}} < 0.01$, since conductances that low make the circuit simulations numerically unstable.
-We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, $\beta = 2$, $N_{\text{test}} = 100$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
-This initialization had much greater entropy and a different emergent property than the returned EPI posterior.
-
-TODO write about specifics of the Hessian analysis.
+The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$, initialized to a gaussian approximation of samples returned by a preliminary ABC search.
+We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
+EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
+The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
+
 %We calculated the Hessian at the mode of the inferred EPI distribution.  
@@
 \subsubsection{Primary visual cortex}\label{methods_V1}
+In the stochastic stabilized supralinear network, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}),
+\end{equation}
+where the noise is an Ornstein-Uhlenbeck process
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
+\end{equation}
+with $\tau_{\text{noise}} = 5\text{ms} > \tau = 1\text{ms}$.
+As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
 Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
+\end{center}
+\caption{\small (V1 1)
+(Left) Simulations for small increases in neuron-type population input.
+Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
+(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
+ }
+ \label{fig:V1_1}
+\end{figure}
+
+\begin{figure}[h]
+\caption{\small (V1 2)
+Posterior predictive distributions of the sum of squares of each pair of noise parameters.
+ }
+ \label{fig:V1_2}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
+\end{center}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+\caption{\small (V1 3)
+EPI posterior for $\mathcal{X}(10\text{ Hz})$.
+ }
+ \label{fig:V1_3}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
+\end{center}
+\end{figure}
+
 \begin{equation}
@@
 
-Since the E-population of this network increases exponentially in the absense of  recurrent inhibitory feedback, we may also observe a paradoxical effect in the inhibitory populations (which is present in E-I networks).
-At 50\% contrast (Fig. \ref{fig:V1}B, dots), this network exhibits a paradoxical effect in the P-population (Fig. \ref{fig:V1}C), but no others (Fig. \ref{fig:V1_1}).
-That is, for a small increase in $h_P$, $\mathbb{E}_t\left[x_P\right]$ decreases.
-
-Fano factor is calculated as the temporal variance divided by the temporal mean following sometime $t_{ss}$ following dynamical evolution from the initial state at $\mathbf{x}(t=0)$.
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[scale=.5]{figures/figV1_1/figV1_1.pdf}
-\end{center}
-\caption{\small Supplemental Figure: 
-(Left) Simulations for small increases in neuron-type population input.
-Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
-(Right) Average and standard deviation of stochastic fluctuations of responses.
- }
- \label{fig:V1_1}
-\end{figure}
-
-\begin{figure}[h]
-\caption{\small Supplemental Figure: 
-Fano factors along the ridge of the posterior in Fig. \ref{fig:V1}E.
- }
- \label{fig:V1_2}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
-\end{center}
-\end{figure}
+Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{ Hz}, 25\text{ Hz}\right]$.
+Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
+\begin{equation}
+s_E(\mathbf{x}; \mathbf{z}) =\mathbb{E}_{x}\left[\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]}\right]
+\end{equation}
+
+For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
+The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.0, 0.5]$.
+EPI optimization was run using 3 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
+The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
+
+\subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
+TODO Agostina and I are putting this together now.
 

### 2021-02-12 written/eLifeRevision/Bittner_eLife2020.tex
latest updates

@@
 The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.
-The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
+The dynamics evolve with timescale $\tau= 90\text{ms}$ governed by connectivity weights $W$
 \begin{equation}
@@
 \end{center}
-\caption{\small 
+\caption{\footnotesize 
 \textbf{A}. Rapid task switching behavioral paradigm (see text). 
@@
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
-\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
-\textbf{D}.
-Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.
-Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).
-Shading indicates standard deviation across trials.
-\textbf{E}.
-Difference in performance of each task during inactivation.
-Inactivation level $\gamma$ scales from no inactivation (0) to full inactivation (1).
-We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
-\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
+\textbf{C}. The EPI posterior distribution of rapid task switching networks.  
+Red and purple stars ($\mathbf{z}_1$ and $\mathbf{z}_2$) indicate different connectivity regimes with different sensitivity vectors $\mathbf{v}_1$ and $\mathbf{v}_2$.
+(Middle-left) Posterior predictive distribution of task accuracies.
+(Bottom-left) Task accuracy along dimensions of sensitivity in each connectivity regime.
+\textbf{D}. Means (solid) and standard deviations (shaded) of each population across random simulated trials.
+Top plots show Pro (top) and Anti (bottom) responses for connectivity $\mathbf{z}_1$.
+Bottom rows show the same $\mathbf{z}_2$.
+\textbf{E}. The EPI posterior predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
+\textbf{F}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
+\textbf{G}. Accuracy with delay period inactivation along each connectivity regime's dimension of sensitivity.
 }
@@
 We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
-\begin{equation}\label{eq:EP}
+\begin{equation}\label{eq:SC_EP}
 \begin{split}
@@
 \end{equation}
-A variance of $7.5\%^2$  in each task will confer a posterior producing performances ranging from about $60\%-90\%$, allowing us to examine the properties of connectivity that yield better performance.
-
-We ran EPI to obtain SC model connectivity parameters $\mathbf{z}$ producing rapid task switching (Fig. \ref{fig:SC}C).
-Some parameters were predictive of accuracy while others were not (Fig. \ref{fig:SC1}), and often had different effects on $p_P$ and $p_A$.
-To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
-\ref{fig:SC2}A).
-These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-
-Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
-This shows that strong task representations and hemispheric\begin{center}
-•
-\end{center} dominance in the dynamics result in better execution of the Pro task.
-By visualizing these four variables together by $p_A$ (Fig.
-\ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
-Thus, stronger hemispheric dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.
-
-In agreement with experimental results from Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig.
-\ref{fig:SC}E) e.g.
-(Fig.
-\ref{fig:SC}D, bottom).
+A variance of $7.5\%^2$  in each task will confer a posterior producing performances ranging from about $60\%-90\%$, allowing us to examine the properties of connectivity that yield better performance in each task.
+Notably, this is our first example using EPI to condition on multiple emergent property statistics ($|f(\mathbf{x}; \mathbf{z})| = 2$).
+
+We ran EPI to obtain the posterior connectivities $\mathbf{z}$ producing rapid task switching (Fig. \ref{fig:SC}C).
+The inferred parameters generate a distribution of task accuracies (Fig. \ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
+The nonlinear patterns of connectivity that govern each task accuracy (Fig. \ref{fig:SC1}A-B) are not fully captured by linear prediction (Fig. \ref{fig:SC1}C).
+For example, the patterns in connectivity increasing Pro accuracy change dramatically after crossing a threshold of $sW$ (Fig. \ref{fig:SC1}A $sW$-$hW$ marginal).
+Not only has EPI captured this complex nonlinear posterior, it offers probabilistic tools for understanding the different regimes of model behavior.
+
+To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
+\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
+Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
+We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
+The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
+\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
+\ref{fig:SC}D, bottom-left, \ref{fig:SC2}).
+In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy tracks monotonic increases in Pro accuracy.
+These responses make intuitive sense, recognizing that $\mathbf{v}_1$ (unlike $\mathbf{v}_2$) points strongly in the direction of connectivity eigenvalue $\lambda_{\text{diag}}$, which is strongly anticorrelated with $p_A$ (Fig.
+\ref{fig:SC3}, \ref{fig:SC4}, see Section \ref{methods_SC}).
+
+In agreement with experimental results from Duan et al., we found optogenetic inactivation during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
+\ref{fig:SC}E)).
 This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
-Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task (Fig.
-\ref{fig:SC4}).
-Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation in the Anti task (Fig. \ref{fig:SC}F)).
+Similarities across Pro and Anti trials in choice period responses following delay period inactivation (Fig.
+\ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
+\ref{fig:SC}F).
+The strong anticorrelation between $p_P$ and $p_{A, opto}$ across posterior connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
+Indeed, in regimes 1 and 2 whose sensitivity dimension elicits very different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
+\ref{fig:SC}F).
 
 In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This posterior revealed the mechanisms leading to greater accuracy in each task as well as those increasing resiliency to perturbation in the Anti task.
-Importantly, every connectivity from this inferred distribution predicts fragility and robustness of performance in the Anti and Pro tasks, respectively.
-EPI allows us to conclude that since \textit{all} parameters of this model producing rapid task switching make such an experimentally verified prediction, we have a well-chosen model.
+This posterior revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
+EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, we have a well-chosen model in that regard.
+Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
 
 \subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
-
-Here, we are interested in the scalability of EPI in number of parameters $|\mathbf{z}|$.  We consider rank-2 RNN with N neurons of connectivity
+Here, we study the scalability of EPI in number of parameters $|\mathbf{z}|$ by inferring the connectivities of recurrent neural networks (RNNs, Fig. \ref{fig:LRRNN}A). 
+ We consider a rank-2 RNN with N neurons of connectivity
 \begin{equation}
@@
 where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.
-
-We want to learn distributions of connectivity that produce stable amplification.
-Two conditions are both necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}.
-These conditions are inequalities on $\text{real}(\lambda_1)$ and $\lambda^s_1$ the maximal real eigenvalue of $W$ and the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$, respectively.
+We  infer connectivity distributions $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   producing stable amplification.
+RNN's exhibiting stable amplification amplify responses to input along some dimensions, and are stable across all dimensions.
+Two conditions are both necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
+$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$.
 
@@
  EPI can naturally condition on this emergent property
-\begin{equation}
+\begin{equation}\label{eq:EP_LRRNN}
 \begin{split}
@@
 
-In contrast, SNPE cannot condition on the variance of observations across posterior.  '
-Thus, we condition on an observation $x_0$ located at the mean of our desired emergent property.
+For comparison, we infer rank-2 RNN connectivites with alternative approaches to likelihood free inference.
+ABC methods define a distance tolerance $\epsilon$ observed data $x_0$ for which we keep sampled parameters.
+To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
 \begin{equation}\label{eq:SNPE_stab_amp_x0}
@@
 \end{equation}
-ABC methods define tolerance $\epsilon$ and distance for observed data $x_0$.
-Here, we chose $\epsilon = 0.5$, an $l-2$ distance, and the same choice for $x_0$ as in Equation \ref{eq:SNPE_stab_amp_x0}.
-
-EPI is capable of scaling to higher dimensional parameter spaces than ABC and SNPE.
-EPI consistently produces the same posterior predictive distribution independent of the dimensionality.
-SMC produces a limited variety of parameters due to the nature of its proposal generation algorithm, yet all parameters obtained produce stable amplification.
-SNPE's posterior predictive distribution is not necessarily close to the conditioning point, and is very dependent on dimensionality.
+located at the mean of our desired emergent property.
+We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the state-of-the-art ABC approach.
+SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
+In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
+Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
+
+As we scale the number of neurons $N$ in the RNN, and thus the dimensionality of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI has superior scaling properties (Fig.
+\ref{fig:LRRNN}B).
+SMC-ABC and SNPE become intractable around 25 and 90 dimensions respectively, while EPI can infer 1000-dimensional distributions in about 1 day.
+No matter the number of neurons, EPI always produces the same distribution of emergent property statistics $\text{real}(\lambda_1)$ and $\lambda_1^s$ (Fig.
+\ref{fig:LRRNN}C, blue), and high variation in response profiles \ref{fig:LRRNN}D, red).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification, are less varied \ref{fig:LRRNN}C, red) and largely produce similar responses \ref{fig:LRRNN}D, red).
+When using SNPE the inferred parameters are widely varied \ref{fig:LRRNN}C, orange), but often produce non amplified or unstable responses \ref{fig:LRRNN}D, orange).
+In conclusion, we found that deep likelihood-free inference techniques are capable of scaling to higher dimensional inference than SMC-ABC.
+However, only EPI can scale to high dimensions while reproducing the emergent property.
 
@@
 When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-Gaussian noise of variance $(1 \times 10^{-12})^2$ amps makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
-
-In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 300$ time steps of $dt = 25ms$.  
+Gaussian noise $d\mathbf{B}$ of variance $(1 \times 10^{-12})^2$ A$^2$ makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
+
+In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 300$ time steps of $dt = 25\text{ms}$.  
 The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
@@
 
-%We calculated the Hessian at the mode of the inferred EPI distribution.  
-%The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
-%With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-%In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
-%The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
+We calculated the Hessian at the mode of the inferred EPI distribution.  
+The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
+With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (strongly negative eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
+In Figure \ref{fig:STG}D, the eigenvectors of the Hessian $v_1$ (solid) and $v_2$ (dashed) are shown evaluated at the mode of the distribution.  
+The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
+Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space are chosen arbitrarily.
 %We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
@@
 
-\begin{figure}[h]
+\begin{figure}
 \begin{center}
@@
 
-\begin{figure}[h]
+\begin{figure}
 \caption{\small (V1 2)
@@
 
+In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
+Specifically, we calculated
+\begin{equation}
+\begin{split}
+\mathbf{z}^*(\sigma_{P,\text{fixed}}) = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. } \sigma_P = \sigma_{P,\text{fixed}}\
+\end{split}
+\end{equation}
+At each mode $\mathbf{z}^*$, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
+
 \subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.9]{figures/figSC1/figSC1.pdf}
+\end{center}
+\caption{\small (SC1): \textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
+\textbf{B}. Same as A colored by Anti task accuracy.  
+\textbf{C}. Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p-value.
+}
+\label{fig:SC1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
+\end{center}
+\caption{\small (SC2):  
+\textbf{A}.
+Simulations in network regime $\mathbf{z}_1$ (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
+\textbf{B}. Same as A for network regime $\mathbf{z}_2$.
+}
+\label{fig:SC2}
+\end{figure}
+
 We order the neural populations of $x$ and $u$ in the following manner
@@
 \end{equation}
-with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2$.  
+with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.  
 The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
@@
 
-The accuracies of $p_P$ and $p_A$ are calculated as
-\begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
-\end{equation}
-and 
-\begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
-\end{equation}
-given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function.
-
-The Heaviside step function is approximated as
-\begin{equation}
-\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
-\end{equation}
-where $\beta = 100$.
-
-As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 5\%^2 \\ 5\%^2 \end{bmatrix}.
-\end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after six epochs (Fig. \ref{fig:SC5}).
-
-For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce an isotropic gaussian with mean 0 and variance $2.5^2$.
-Accuracies were estimated over 200 trials of random gaussian noise, which was sampled independently for each drawn parameter $\mathbf{z}$ and each iteration of the EPI optimization.
-
 \begin{figure}
 \begin{center}
-\includegraphics[scale=0.8]{figures/figSC1/figSC1.pdf}
-\end{center}
-\caption{\small (SC1): Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p value.
-}
-\label{fig:SC1}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
-\end{center}
-\caption{\small (SC2):  
-A. Invariant eigenvectors of connectivity matrix $W$.
-B. Eigenvalues of connectivities of EPI distribution versus task accuracies.}
-\label{fig:SC2}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC3/figSC3.pdf}
+\includegraphics[scale=0.9]{figures/figSC3/figSC3.pdf}
 \end{center}
 \caption{\small (SC3):  
-A. Connectivity eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
-B. Same for Anti task.
+\textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
+\textbf{B}. Eigenvalues of connectivities of EPI distribution versus task accuracies.
 }
@@
 \end{figure}
+
+The accuracies of $p_P$ and $p_A$ are calculated as
+\begin{equation}
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+\end{equation}
+and 
+\begin{equation}
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+\end{equation}
+given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function, and the accuracy is averaged over 200 independent trials.  The Heaviside step function is approximated as
+\begin{equation}
+\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
+\end{equation}
+where $\beta = 100$.
 
@@
 \end{center}
-\caption{\small (SC4):  Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.8.
+\caption{\small (SC4):  \textbf{A}.
+Pairplots of eigenvalues of connectivity matrices in EPI distribution colored by Pro task accuracy.
+Red and purple stars and arrows correspond to eigenvalues and sensitivity directions $\mathbf{z}_1$, $\mathbf{z}_2$, $\mathbf{v}_1$, and $\mathbf{v}_2$.
+\textbf{B}.
+Same colored by Anti task accuracy.
 }
@@
 \end{center}
-\caption{\small (SC5): EPI optimization of the SC model producing rapid task switching.
-A. Entropy throughout optimization. 
-B. The emergent property statistic means and variances converge to their constraints at 12,000 iterations following the sixth augmented Lagrangian epoch.
+\caption{\small (SC5): 
+\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
+\textbf{B}. Connectivity eigenvalues versus the task error induced by delay period inactivation.
+\textbf{C}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
 }
@@
 \end{figure}
+
+Writing the EPI posterior as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 7.5\%^2 \\ 7.5\%^2 \end{bmatrix}.
+\end{equation}
+Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
+The optimization converged after six epochs (Fig. \ref{fig:SC5}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{figures/figSC6/figSC6.pdf}
+\end{center}
+\caption{\small (SC6): 
+A. Entropy throughout optimization. 
+B. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+}
+\label{fig:SC5}
+\end{figure}
+
+For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
+The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
+We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
+The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 25$ at greatest entropy across random seeds.
+
+To make sense of this inferred distribution, we identified two modes used to represent the two regimes of connectivity in this posterior:
+\begin{equation}
+\begin{split}
+\mathbf{z}_1 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. }  hw = -1.25\text{, }  sW > 0
+\end{split}
+\end{equation}
+and 
+\begin{equation}
+\begin{split}
+\mathbf{z}_2 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. } hw = -1.25\text{, } sW < 0
+\end{split}
+\end{equation}
+
+To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
+\ref{fig:SC3}A).
+These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
+We found significant trends across the EPI posterior connectivities: the eigenvalues $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ were correlated with $p_P$, while $\lambda_{\text{all}}$ was anticorrelated with $p_P$.
+$\lambda_{\text{all}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ were all significantly anticorrelated with $p_A$.
+
+Under this decomposition, we can re-visualize the posterior in eigenvalue space (Fig. \ref{fig:SC4}).
+Furthermore, we can project the dimensions of sensitivity into eigenvalue space as well, giving us a more intuitive sense of how connectivity affects computation in each regime.
+We see that sensitivity dimensions $\mathbf{v}_1$ and $\mathbf{v}_2$, which cause $p_P$ to increase and a regime dependent change in $p_A$, both point in the direction of increasing $\lambda_{\text{side}}$ and decreasing $\lambda_{\text{task}}$.
+These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
+As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
+A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
+
+During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.
+We found that $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ were significantly anticorrelated with Anti error during delay period inactivation.
+Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.
 

### 2021-02-16 written/eLifeRevision/Bittner_eLife2020.tex
section 3.5 + discussion

@@
 Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via supralinear nonlinearity $\phi = \left[\right]^2_+$.
-The input is also comprised of a slow noise component $\bm{\epsilon} \sim OU\left(\tau_{\text{noise}}, \Sigma \right)$ of time scale $\tau_{\text{noise}} > \tau$ and covariance $\Sigma = \text{diag}\left(\bm{\sigma^2}\right)$ (see Section \ref{methods_V1})
+The input is also comprised of a slow noise component $\bm{\epsilon} \sim OU\left(\tau_{\text{noise}}, \bm{\sigma} \right)$ of time scale $\tau_{\text{noise}} > \tau$ and variance parameters $\bm{\sigma}$ (see Section \ref{methods_V1})
 \begin{equation}
@@
 This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have inhibitory multiplicity, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
-Both modeling advancements introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
+Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
 Here, we consider fixed weights $W$ and input $\mathbf{h}$ according to a fit of the deterministic model to contrast responses \cite{palmigiano2020structure} (Fig.
@@
 The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.
+Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching behavior.
 
@@
 
-We ran EPI to obtain the posterior connectivities $\mathbf{z}$ producing rapid task switching (Fig. \ref{fig:SC}C).
-The inferred parameters generate a distribution of task accuracies (Fig. \ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
+The EPI inferred parameters  (Fig.
+\ref{fig:SC}C) generate a distribution of task accuracies (Fig.
+\ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
 The nonlinear patterns of connectivity that govern each task accuracy (Fig. \ref{fig:SC1}A-B) are not fully captured by linear prediction (Fig. \ref{fig:SC1}C).
@@
 \ref{fig:SC}F).
-The strong anticorrelation between $p_P$ and $p_{A, opto}$ across posterior connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
-Indeed, in regimes 1 and 2 whose sensitivity dimension elicits very different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
+The strong anticorrelation between $p_P$ and $p_{A, opto}$ across posterior connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
+Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
 \ref{fig:SC}F).
@@
 This posterior revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
-EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, we have a well-chosen model in that regard.
+EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, the model is well-chosen in that regard.
 Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
@@
 Here, we study the scalability of EPI in number of parameters $|\mathbf{z}|$ by inferring the connectivities of recurrent neural networks (RNNs, Fig. \ref{fig:LRRNN}A). 
- We consider a rank-2 RNN with N neurons of connectivity
+ We consider a rank-2 RNN with N neurons having connectivity
 \begin{equation}
@@
  \end{equation}
-where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.
+where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$ and $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$.
+The random component has strength $g=0.01$ and $\chi_{i,j} \sim \mathcal{N}(0, 1)$.  
 We  infer connectivity distributions $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   producing stable amplification.
@@
 $\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$.
-
-In our analysis, we seek to condition rank-2 networks of increasing size on a regime of stable amplification.  
-Networks with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will yield moderate amplification.
- EPI can naturally condition on this emergent property
-\begin{equation}\label{eq:EP_LRRNN}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
-\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix} .
-\end{split}
-\end{equation}
-
-For comparison, we infer rank-2 RNN connectivites with alternative approaches to likelihood free inference.
-ABC methods define a distance tolerance $\epsilon$ observed data $x_0$ for which we keep sampled parameters.
-To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
-\begin{equation}\label{eq:SNPE_stab_amp_x0}
-x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
-\end{equation}
-located at the mean of our desired emergent property.
-We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the state-of-the-art ABC approach.
-SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
-In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
-Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
-
-As we scale the number of neurons $N$ in the RNN, and thus the dimensionality of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI has superior scaling properties (Fig.
-\ref{fig:LRRNN}B).
-SMC-ABC and SNPE become intractable around 25 and 90 dimensions respectively, while EPI can infer 1000-dimensional distributions in about 1 day.
-No matter the number of neurons, EPI always produces the same distribution of emergent property statistics $\text{real}(\lambda_1)$ and $\lambda_1^s$ (Fig.
-\ref{fig:LRRNN}C, blue), and high variation in response profiles \ref{fig:LRRNN}D, red).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification, are less varied \ref{fig:LRRNN}C, red) and largely produce similar responses \ref{fig:LRRNN}D, red).
-When using SNPE the inferred parameters are widely varied \ref{fig:LRRNN}C, orange), but often produce non amplified or unstable responses \ref{fig:LRRNN}D, orange).
-In conclusion, we found that deep likelihood-free inference techniques are capable of scaling to higher dimensional inference than SMC-ABC.
-However, only EPI can scale to high dimensions while reproducing the emergent property.
 
@@
 For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
-\textbf{D}. Simulations of network parameters resulting from each method ($\tau=100ms$).  Each trace corresponds to simulation of one $z$.  
-\textbf{E}. Ratio of obtained samples producing stable amplification.
+\textbf{D}.
+Simulations of network parameters resulting from each method ($\tau=100ms$).
+Each trace corresponds to simulation of one $z$.
+(Below) Ratio of obtained samples producing stable amplification.
 }
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Linking RNN connectivity to error} \label{results_RNN}
-%So far, each model we have studied was designed from fundamental biophysical principles, genetically- or functionally-defined neuron types.  
-%At a more abstract level of modeling, recurrent neural networks (RNNs) are high-dimensional dynamical models of computation that are becoming increasingly popular in neuroscience research \cite{barak2017recurrent}.
-%In theoretical neuroscience, RNN dynamics usually follow the equation
-%\begin{equation}
-%\frac{dx}{dt} = -x + W \phi(x) + h,
-%\end{equation}
-%where $x$ is the network activity, $W$ is the network connectivity, $\phi(\cdot) = \tanh(\cdot)$, and $h$ is the input to the system.
-%Such RNNs are trained to do a task from a systems neuroscience experiment, and then the unit activations of the trained RNN are compared to recorded neural activity.
-%Fully-connected RNNs with tens of thousands of parameters are challenging to characterize \cite{sussillo2013opening}, especially making statistical inferences about their parameterization.
-%Alternatively, we considered a rank-1, $N$-neuron RNN with connectivity consisting of the sum of a random and a structured component:
-%\begin{equation}
-%W = g\chi + \frac{1}{N}mn^\top.
-%\end{equation}
-%The random component $g \chi$ has strength $g$, and random component weights are Gaussian distributed $\chi_{i,j} \sim \mathcal{N}(0, \frac{1}{N})$.  
-%The structured component $\frac{1}{N}m n\top$ has entries of $m$ and $n$ drawn from Gaussian distributions $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$.
-%Recent theoretical work derives the low-dimensional response properties of low-rank networks from statistical parameterizations of their connectivity, such as $z = [g, M_m, M_n]$ \cite{mastrogiuseppe2018linking}.
-%We used EPI to infer the parameterizations of rank-1 RNNs solving an example task, enabling discovery of properties of connectivity that result in different types of error in the computation.
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.75]{figures/fig5/fig5.pdf}
-%\end{center}
-%\caption{\small Sources of error in an RNN solving a simple task.  A. (left) A rank-1 RNN executing a Gaussian posterior conditioning computation on $\mu_y$. 
-%(right) Error in this computation can come from over- or under-estimating the posterior mean or variance. 
-%B.
-EPI distribution of rank-1 RNNs executing Gaussian posterior conditioning.
-Samples are colored by (left) posterior mean $\mu_{\text{post}}=\kappa_r$ and (right) posterior variance $\sigma^2_{\text{post}}=\Delta_T$
-%C.  Finite-size network simulations of 2,000 neurons with parameters $z_1$ and $z_2$ sampled from the inferred distribution.  
-%Activity along readout $\kappa_r$ (cyan) is stable despite chaotic fluctuations.  
-%D. The posterior mean computed by RNNs parameterized by $z_1$ and $z_2$ perturbed in the dimension of the product of $M_m$ and $M_n$.  
-%Means and standard errors are shown across 10 realizations of 2,000-neuron networks.}
-%\label{fig:RNN}
-%\end{figure}
-%
-%The task we consider is Gaussian posterior conditioning: calculate the parameters of a posterior distribution induced by a prior $p(\mu_y) = \mathcal{N}(\mu_0=4, \sigma_0^2=1)$ and a likelihood $p(y | \mu_y) = \mathcal{N}(\mu_y,\sigma^2_y = 1)$, given a single observation $y$.
-%Conjugacy offers the result analytically; $p(\mu_y | y ) = \mathcal{N}\left(\mu_{post},\sigma_{post}^2\right)$, where:   
-%\begin{equation}
-%\mu_{\text{post}} = \frac{\frac{\mu_0}{\sigma_0^2} + \frac{y}{\sigma_y^2}}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}
-%~~~~~~~~~~
-% \sigma^2_{\text{post}} = \frac{1}{\frac{1}{\sigma_0^2} + \frac{1}{\sigma_y^2}}.
-%\end{equation} 
-%To solve this Gaussian posterior conditioning task, the RNN response to a constant input $h = y r + (n-M_n)$ must equal the posterior mean along readout vector $r$ , where
-%\begin{equation}
-%\kappa_r =  \frac{1}{N} \sum_{j=1}^N r_j \phi(x_j).
-%\end{equation}
-%Additionally, the amount of chaotic variance $\Delta_T$ must equal the posterior variance.
-%Theory for low-rank RNNs allows us to express $\kappa_r$ and $\Delta_T$ in terms of each other through a solvable system of nonlinear equations  (see Section \ref{methods_LRRNN}) \cite{mastrogiuseppe2018linking}.
-%This theory facilitates the mathematical formalization of task execution into an emergent property, where the emergent property statistics of the RNN activity are $\kappa_r$ and $\Delta_T$, and the emergent property values are the ground truth posterior mean $\mu_{\text{post}}$ and variance $\sigma^2_{\text{post}}$:
-%\begin{equation}
-%\mathbb{E} \begin{bmatrix} \kappa_r \\ \Delta_T \\ (\kappa_r-\mu_{\text{post}})^2 \\ (\Delta_T^2-\sigma^2_{\text{post}})^2 \end{bmatrix} ~~=~~ \begin{bmatrix} \mu_{\text{post}} \\ \sigma^2_{\text{post}} \\ 0.1 \\ 0.1 \end{bmatrix}.
-%\end{equation}
-%We chose a substantial amount of variance in these emergent property statistics, so that the inferred distribution resulted in RNNs with a variety of errors in their solutions to the gaussian posterior conditioning problem.
-%
-%EPI was used to learn distributions of RNN connectivity properties $z = [g, M_m, M_n]$ executing Gaussian posterior conditioning given an input of $y=2$, where the true posterior is $\mu_{\text{post}}=3$ and $\sigma_{\text{post}} = 0.5$ (Fig.
-\ref{fig:RNN}A).
-%We examined the nature of the over- and under-estimation of the posterior means (Fig. \ref{fig:RNN}B left) and variances (Fig. \ref{fig:RNN}B right) in the inferred distributions (300 samples).  
-%The symmetry in the $M_m$-$M_n$ plane, suggests a degeneracy in the product of $M_m$ and $M_n$ (Fig. \ref{fig:RNN}B).  
-%Indeed, $M_m M_n$ strongly determines the posterior mean ($r=0.62$,  $p<10^{-4}$). 
-%Furthermore, the random strength $g$ strongly determines the chaotic variance ($r=0.56$,  $p<10^{-4}$).
-%Neither of these observations were obvious from what mathematical analysis is available in networks of this type (see Section \ref{methods_LRRNN}).  
-%While the link between random strength $g$ and chaotic variance $\Delta_T$ (and resultingly posterior variance in this problem) is well-known \cite{sompolinsky1988chaos}, the distribution admits a novel hypothesis: the estimation of the posterior mean by the RNN increases with $M_m M_n$.
-%
-%We tested this prediction by taking parameters $z_1$ and $z_2$ as representative samples from the positive and negative $M_m$-$M_n$ quadrants, respectively.
-%Instead of using the theoretical predictions shown in Figure \ref{fig:RNN}B, we simulated finite-size realizations of these networks with 2,000 neurons (e.g. Fig. \ref{fig:RNN}C).  
-%We perturbed these parameter choices by $M_m M_n$ clarifying that the posterior mean can be directly controlled in this way (Fig. \ref{fig:RNN}D; $p<10^{-4}$), see Section \ref{methods_LRRNN}).
-%Thus, EPI confers a clear picture of error in this computation: the product of the low rank vector means $M_m$ and $M_n$ modulates the estimated posterior mean while the random strength $g$ modulates the estimated posterior variance.
-%This novel procedure of inference on reduced parameterizations of RNNs conditioned on the emergent property of task execution is generalizable to other settings modeled in \cite{mastrogiuseppe2018linking} like noisy integration and context-dependent decision making (Fig.
-S5).
-%
+In our analysis, we seek to condition rank-2 networks of increasing size on a regime of stable amplification.  
+Networks with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will yield moderate amplification.
+ EPI can naturally condition on this emergent property
+\begin{equation}\label{eq:EP_LRRNN}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
+\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix} .
+\end{split}
+\end{equation}
+
+For comparison, we infer rank-2 RNN connectivites with alternative approaches to likelihood free-inference.
+ABC methods define a tolerance $\epsilon$ from observed data $x_0$ for which we keep sampled parameters.
+To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
+\begin{equation}\label{eq:SNPE_stab_amp_x0}
+x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
+\end{equation}
+located at the mean of our desired emergent property.
+We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the most efficient ABC approach.
+SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
+In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
+Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
+
+As we scale the number of neurons $N$ in the RNN, and thus the dimensionality of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI has superior scaling properties (Fig.
+\ref{fig:LRRNN}B).
+SMC-ABC and SNPE become intractable around 25 and 90 dimensions respectively, while EPI can infer 1000-dimensional distributions in about 1 day.
+No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ of $\mathcal{X}$(Fig.
+\ref{fig:LRRNN}C, blue), and high variation in response profiles \ref{fig:LRRNN}D, blue).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification, are less varied \ref{fig:LRRNN}C, red) and largely produce similar responses \ref{fig:LRRNN}D, red).
+When using SNPE the inferred parameters are widely varied \ref{fig:LRRNN}C, orange), but often produce non-amplified or unstable responses \ref{fig:LRRNN}D, orange).
+In conclusion, we found that deep likelihood-free inference techniques are capable of scaling to higher dimensional inference than SMC-ABC.
+However, only EPI can scale to high dimensions while reproducing the emergent property.
+
 %\section{Discussion}
@@
 \section{Discussion}
-NOTE: \textbf{This is the old discussion section.  I will rewrite this based on our discussion of the rest of the draft.}
-
-In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, druckmann2007novel, byron2009gaussian, park2011bayesian, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
+
+In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{paninski2018neural}.
 Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
@@
 The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we replaced this opaque procedure of parameter identification in theoretical neuroscience with emergent property inference, opening the door to careful inference in careful models of neural computation.
-
-Biologically realistic models of neural circuits often prove formidable to analyze.
-Two main factors contribute to the difficulty of this endeavor.
-First, in most neural circuit models, the number of parameters scales quadratically with the number of neurons, limiting analysis of its parameter space. 
-Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
-For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-
-In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}.
-By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out.
-Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
-For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
-This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.
-
-In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
-Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
-This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
-While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
-The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.
+In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards careful inference in careful models of neural computation (see Section \ref{methods_related}).
+
+Specifically, approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential} has been the standard approach to parameter inference in neural circuit models lacking tractable likelihoods.
+ABC methods do not confer probabilities on accepted parameters, require an acceptance threshold chosen to trade-off inference quality with tractability, do not scale efficiently to high-dimensional parameter spaces, and require independent techniques to analyze sensitivity for local parameter choices \cite{sisson2018handbook}.
+In contrast, EPI allows probability evaluations at any point in parameter space, conditions posteriors on the natural quantification of emergent properties, scales to high dimensional parameter spaces, and naturally admits sensitivity quantification via fast evaluations of the posterior Hessian.
+
+Technically, EPI is a maximum entropy method, which learns parameter distributions that are as random as possible given that they produce the emergent property.
+Conceptually, maximally random distributions given some constraints are useful for understanding parametric sensitivity.
+This is well understood in Bayesian inference, where maximum entropy is the chosen normative principle.
+This is emphasized by an innovative formalism unifying top-down maximum entropy normative models with bottom-up statistical models \cite{mlynarski2020statistical}.
+Indeed, EPI is an adaptive variational inference program, and may be considered to have a Bayesian uniform prior (see Section \ref{methods_VI}).
+
+%Biologically realistic models of neural circuits often prove formidable to analyze.
+%Two main factors contribute to the difficulty of this endeavor.
+%First, in most neural circuit models, the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space. 
+%Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
+%For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
+
+Biologically realistic models of neural circuits often prove formidable to analyze for two main reasons.
+A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
+We see in Section \ref{results_LRRNN} that EPI scales seemlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
+EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
+
+A secondary challenge is that the structure of the parametric regimes governing emergent properties is intricate.  
+For example, even in low dimensional circuits, models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
+With EPI, we use deep probabillity distributions to capture the complex nonlinear parameter distributions governing model behavior.
+In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing curcuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
+In Section \ref{results_SC}, we identified two regimes of SC connectivity resulting in rapid task switching, and found that the full distribution of rapid task switching networks reproduced an experimental result.
+
+%In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}.
+%By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out.
+%Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
+%For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
+%This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.
+
+%In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
+%Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
+%This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
+%While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
+%The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.
 
@@
 This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
-These approaches can be used complementarily to enhance the practice of theoretical neuroscience.
-
-\textbf{TODO} *merge this point in*
-
-While much research in computational neuroscience has focused on optimizing neural architectures to process information and accomplish tasks \cite{richards2019deep}, structure in parameter space of the set of optimized solutions is rarely discussed and lacks a probabilistic treatment.
-Talk about Wyktor's work here \cite{mlynarski2020statistical}.
+All of these approaches can be used complementarily to enhance the practice of theoretical neuroscience.
 
@@
 \textbf{Code availability statement}: \\
-The software written for the current study is available from the corresponding author upon reasonable request.
+All software written for the current study is available at https://github.com/cunningham-lab/epi.
 
@@
 \subsubsection{Primary visual cortex}\label{methods_V1}
-In the stochastic stabilized supralinear network, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
+In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
 \begin{equation}
@@
 \end{equation}
-where the noise is an Ornstein-Uhlenbeck process
-\begin{equation}
-\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
+where the noise is an Ornstein-Uhlenbeck process $\bm{\epsilon} \sim OU(\tau_{\text{noise}}, \bm{\sigma})$
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\tilde{\sigma}_\alpha dB
 \end{equation}
 with $\tau_{\text{noise}} = 5\text{ms} > \tau = 1\text{ms}$.
+The noisy process is parameterized as
+\begin{equation}
+\tilde{\sigma}_\alpha = \sigma_\alpha \sqrt{1 + \frac{\tau}{\tau_{\text{noise}}}},
+\end{equation}
+so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
 As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
@@
 For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
-The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.0, 0.5]$.
-EPI optimization was run using 3 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
+The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
+EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
 The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.

### 2021-02-26 written/eLifeRevision/Bittner_eLife2020.tex
abstract update

@@
 The operation of these circuits, like all models, critically depends on the choices of model parameters.
-When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
+%When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
+%We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
+We present a novel technique designed to infer parameter distributions that produce mathematically defined emergent properties of computation, rather than sets of experimental data.
+We find that conditioning parameter distributions on an emergent property rather than datasets is natural and often advantageous in theoretical practice.
+Also, this approach allows us to greatly simplify the deep inference architecture used and to scale inference to high dimensions.
+We bring this deep inference technique to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
+First, we show a worked example analyzing parametric sensitivity in the stomatogastric ganglion, emphasizing the general applicability of this approach to complex models with complex emergent properties.
+Then, in a model of primary visual cortex where mathematical analysis becomes untenable, we identify a nonlinear manifold revealing the rule of parametric degeneracy with respect to variability.
+Next, we gain a mechanistic understanding of how superior colliculus connectivity produces rapid task switching in multiple parametric regimes.
+Finally, we scale inference of RNN connectivity exhibiting stable amplification to 1,000-dimensional parameter spaces.
+These analyses illustrate how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
+
+%-------- old sentences -----------
+
 %Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
 %However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
-Our novel method solves the inverse problem by identifying the full space of parameters producing the emergent property.
 %Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
-We motivate this methodology with a worked example analyzing sensitivity in the stomatogastric ganglion.  
 %We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
-We then use it to reveal the key factors of variability in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and scale inference of large low-rank RNN's exhibiting stable amplification.
-This work illustrates how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
 %More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
-%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
 The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.
@@
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-%This creates an unfavorable tradeoff.
-On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
-%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
 Alternatively, we can gain insight into these complex models by identifying the full distribution of parameters consistent with specified emergent phenomena.
@@
 The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
 However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
@@
 
-% now transition to deep ML...
-%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
-%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
 Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
 In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
-%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
 One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
@@
 
-% What's wrong with existing methodology:
 Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
@@
 
+To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
+This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
+First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
+Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
+Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
+Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
+Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
+We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
+
+Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
+These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to decision making), and network scale (from four to hundreds of neurons).
+First, we use EPI to understand the characteristics of noise across mulitple neuron-type populations that govern variability in a model of primary visual cortex.
+Then, we use EPI to infer multiple regimes of superior colliculus connectivity that perform rapid task switching.
+The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{litwin2016inhibitory, duan2018collicular}.
+Finally, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
+These results point to the value of deep inference for the interrogation of biologically relevant models.
+
+%-------- old sentences -----------
+
+%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
+
+%This creates an unfavorable tradeoff.
+On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
+%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
+
+%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
+
+% now transition to deep ML...
+%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
+%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
+%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
+
 %However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.
@@
 
-% now we've constructed the tension that theory models need this and statistical models know how to do this... now connect!
-To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
-This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
 %First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
@@
 %We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
-Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
-Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
-Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to decision making), and network scale (from four to hundreds of neurons).
+
 %First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-First, we use EPI to understand the characteristics of noise across mulitple neuron-type populations that govern variability in a model of primary visual cortex.
 %Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-Then, we use EPI to infer multiple regimes of superior colliculus connectivity that perform rapid task switching.
-The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{litwin2016inhibitory, duan2018collicular}.
 %Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.
-Finally, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
-These results point to the value of deep inference for the interrogation of biologically relevant models.
-
-%-------------------
 

### 2021-03-07 written/eLifeRevision/Bittner_eLife2020.tex
focus on EP vs data through 3.2

@@
 Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or in terms of neural activity -- and thus can offer insights into neural computation.
-The operation of these circuits, like all models, critically depends on the choices of model parameters.
+The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
 %When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
 %We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
-We present a novel technique designed to infer parameter distributions that produce mathematically defined emergent properties of computation, rather than sets of experimental data.
-We find that conditioning parameter distributions on an emergent property rather than datasets is natural and often advantageous in theoretical practice.
-Also, this approach allows us to greatly simplify the deep inference architecture used and to scale inference to high dimensions.
-We bring this deep inference technique to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
+We present a novel technique designed to infer parameter distributions that produce mathematically defined emergent properties of computation.
+We detail how the parameter distributions of theoretical neuroscience problems are better defined by the emergent properties they produce, rather than conditioned upon artificial datasets.
+This approach is not only well-motivated, but allows technical simplifications to statistical inference in mechanistic models and scaling to high dimensions.
+With this method, we bring deep inference to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
 First, we show a worked example analyzing parametric sensitivity in the stomatogastric ganglion, emphasizing the general applicability of this approach to complex models with complex emergent properties.
@@
 The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
-A neural computation is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
+A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
 The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
@@
 Alternatively, we can gain insight into these complex models by identifying the full distribution of parameters consistent with specified emergent phenomena.
-By solving the inverse problem in this way, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-
-The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
-Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-
-Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
-In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
-One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
-These works have used deep neural networks to expand the domain of neural data sets amenable to statistical modeling \cite{paninski2018neural}.
-
-Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
-First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
-Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
-Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
-Sampling-based inference approaches (e.g.
+By solving the inverse problem in this way, scientific analysis of complex biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+%By solving the inverse problem in this way, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+
+%The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
+%However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
+%Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+% 
+%Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
+%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
+%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
+%Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
+%These works have used deep neural networks to expand the domain of neural data sets amenable to statistical modeling \cite{paninski2018neural}.
+
+One preferred formalism for parameter identification is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
+Recent work has used variational autoencoders (VAEs) \cite{kingma2013auto, rezende2014stochastic} to interrogate hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}, thus expanding the domain of neural data sets amenable to statistical modeling.
+However, most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood necessary for statistical inference.
+Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference (LFI) methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+A flourishing new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional LFI approaches.
+However, as we detail, all of these approaches require good datasets for the scientific question at hand.
+
+%
+%Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
+%First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
+%Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
+%Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
+%Sampling-based inference approaches (e.g.
 ABC and Markov chain Monte Carlo (MCMC) \cite{hastings1970monte, metropolis1953equation}) confer flexible approximations, yet scale poorly in number of parameters.
-While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
-Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
-Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
-
-To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
-This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
-First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
-Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
-Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
+%While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
+%Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
+%Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
+
+%To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
+%This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
+%First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
+%Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
+%Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
+%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
+%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
+%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
+
+This work seeks to clarify an important incongruity between theoretical approaches to neuroscience and existing statistical inference methodology.
+In theoretical neuroscience, we are concerned with the computational properties -- the \emph{emergent phenomena} -- of our models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, not noisy observed datasets \cite{paninski2018neural}.
+To use the aforementioned inference paradigm, scientists must shoehorn such mathematical criteria into an artificial dataset compatible with existing statistical approaches.
+Theorists are therefore barred from using the probabilistic modeling toolkit for science, unless they reformulate their inverse problem to fit an evidence accumulation framework.
+
+These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
+As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets  exhibiting some mathematical criteria.
+EPI fixates on the inverse problem by constraining the predictions of the inferred parameter distribution to produce the emergent property exactly.
 Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-Equipped with this methodology, we then investigated three models of current importance in theoretical neuroscience.
-These models were chosen to demonstrate generality through ranges of biological realism (from conductance-based biophysics to recurrent neural networks), neural system function (from pattern generation to decision making), and network scale (from four to hundreds of neurons).
-First, we use EPI to understand the characteristics of noise across mulitple neuron-type populations that govern variability in a model of primary visual cortex.
-Then, we use EPI to infer multiple regimes of superior colliculus connectivity that perform rapid task switching.
-The novel scientific insights offered by EPI contextualize and clarify the previous studies exploring these models \cite{litwin2016inhibitory, duan2018collicular}.
-Finally, we emphasize the scalability of EPI by inferring high-dimensional distributions of RNNs exhibiting stable amplification.
-These results point to the value of deep inference for the interrogation of biologically relevant models.
+By using deep probability distributions EPI flexibly captures parametric structure in mechanistic models.
+This technique is designed to identify the full parameter space producing an emergent property and facilitate the seemless structural analysis of the inferred parameter distribution.
+Thus, EPI provides neuroscientists with an efficient, versatile probabilistic modeling toolkit designed explicitly for theoretical inverse problems.
+
+Equipped with this method, we bring deep inference to bear on theoretical neuroscience to an unprecedented extent.
+Throughout this work, we showcase the capabilities of EPI on four neural circuit models across ranges of biological realism, neural system function, and network scale.
+First, we show EPI's ability to capture subtle, nonlinear parametric structure in a stomatogastric ganglion subcircuit model \cite{gutierrez2013multiple}.
+In a model of primary visual cortex \cite{palmigiano2020structure}, we show how to gain insight from multiple inferred distributions.
+Next, we used EPI to idenitfy and structurally characterize multiple parametric regimes of superior colliculus activity in a model of task switching \cite{duan2018collicular}.
+Finally, we emphasize the superior scalability of EPI compared to other LFI techniques by inferring high-dimensional distributions of RNN connectivities that exhibit amplified, yet stable responses -- a hallmark of cortical sensory systems \cite{more, bondanelli2020coding}.
+
+Most importantly of all in this work, we present novel theories of neural computation borne from EPI analysis.
+We identified an unknown parametric rule of variability with respect to inhibitory neuron type in a V1 model, where analytic techniques became untenable.
+Furthermore, we identified multiple regimes of SC connectivity which confer rapid task switching, and used the structural analytic tools of EPI to gain a mechanistic understanding of circuit responses.
+These valuable theoretical insights illustrate the value of deep inference for the interrogation of neural circuit models.
 
@@
 % Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-Third, parameter analyses ensue: brute-force parameter sweeps, ABC sampling, and sensitivity analyses are all routinely used to reason about what parameter configurations lead to an emergent property.
+Third, parameter analyses ensue: brute-force parameter sweeps, ABC sampling \cite{liepe2014framework}, and sensitivity analyses \cite{raue2009structural} are all routinely used to reason about what parameter configurations lead to an emergent property.
 In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
@@
 \ref{fig:STG}E).
-This distribution -- represented for clarity as samples from the distribution -- is then a scientifically meaningful and mathematically tractable object.  
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for network syncing (the mode; Figure \ref{fig:STG}E yellow star), and how network syncing decays based on changes away from the mode.
-The eigenvectors (of the Hessian of the distribution at the mode) quantitatively formalize the robustness of intermediate hub frequency (Fig.
-\ref{fig:STG}E solid ($v_1$) and dashed  ($v_2$) black arrows).
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) agree with error contours (Fig.
+This distribution -- represented for clarity as samples from the distribution -- is then the most random parameter distribution producing the emergent property. 
+In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
+Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) (Fig.
+\ref{fig:STG}E, arrows) agree with error contours (Fig.
 \ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
@@
 \subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
-Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
+Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions.
 %First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
@@
 %Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to an intermediate frequency of 0.55Hz.  
-Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
-
-%\begin{equation}\label{eq:EP}
- %\mathbb{E}\left[T(x) \right] ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \omega_{\text{f1}}(x) \\ \vdots \\ (\omega_{\text{f1}}(x) - 0.53)^2 \\ \vdots \end{bmatrix} ~~=~~  
- %\begin{bmatrix} 0.53 \\ \vdots \\ 0.025^2 \\ \vdots \end{bmatrix} ~~\triangleq~~ \mu,
- %\end{equation}
+Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz.  
+%Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
+Mathematically, we achieve this with two constraints: by setting the mean hub frequency of model generated activity to 0.55Hz.
+\begin{equation}\label{eq:EP_STG1}
+ \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix}
+\end{equation}
+and requiring that the variance around this mean is small
+\begin{equation}\label{eq:EP_STG2}
+\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix}.
+\end{equation}
+This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the compatible parameter space.
+In general, an emergent property
 \begin{equation}\label{eq:EP_STG}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~  \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix} ~~\triangleq~~ \bm{\mu}  \\
- \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] &~~\triangleq~~ \text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix} ~~\triangleq~~ \bm{\sigma}^2 .
-\end{split}
-\end{equation}
-
-%which completes the quantification of the emergent property.
-The emergent property statistic $f(\mathbf{x}; \mathbf{z}) = \omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ along with its constrained mean $\bm{\mu}$ and variance $\bm{\sigma}^2$ define the emergent property denoted $\mathcal{X}$.
+\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
+\end{equation}
+defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
+The choice of $\bm{\sigma}^2$ predicates the degree of variability in around mean $\bm{\mu}$ that is consistent with the emergent property.
 
@@
 This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}C (see Section \ref{methods_EPI}).  
-Then, mathematically, we must solve the following optimization program: 
+This deep distribution is represented in Figure \ref{fig:STG}D (see Section \ref{methods_EPI}).
+Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we choose a normative selection principle imposing no additional structure beyond the production of the emergent property \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}, which is the same normative principle of Bayesian inference (see Section \ref{methods_VI}).
+
+%Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
+%This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
+%This deep distribution is represented in Figure \ref{fig:STG}E (see Section \ref{methods_EPI}).  
+%Then, mathematically, we must solve the following optimization program: 
 %\begin{equation} \label{eq:EPI}
 %\begin{split}
-%&\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
-%&\text{  s.t.  } \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right] \right] = \bm{\mu}, \\
+%q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) ~~=~~ &\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
+%&\text{  s.t.
+} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
 %\end{split}
 %\end{equation}
-\begin{equation} \label{eq:EPI}
-\begin{split}
-q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) ~~=~~ &\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
-&\text{  s.t.
-} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
-\end{split}
-\end{equation}
-%where $T(\mathbf{x}), \bm{\mu}$ are defined as in Equation \ref{eq:EP}, and $p(\mathbf{x} | \mathbf{z})$ is the intractable distribution of data from the model, $\mathbf{x}$, given that model's parameters $\mathbf{z}$ (we access samples from this distribution by running the model forward).
-where $f(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
-According to the emergent property of interest, $f(\mathbf{x}, \mathbf{z})$ may contain multiple statistics, in which case the mean and variance vectors $\bm{\mu}$ and $\bm{\sigma}^2$ match this dimension.
-%The purpose of each element in this program is detailed in Figure \ref{fig:STG}D.
-Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
-This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
-Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness. 
-The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.
-
-EPI optimizes the weights and biases $\bm{\theta}$ of the deep network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
-%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
-Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}).
-Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).
-
-In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
-%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by the emergent property statistics and values, respectively (see Section \ref{methods_VI}).
-%Specifically, EPI executes variational inference in an exponential family model, the sufficient statistics and mean parameter of which are defined by $f(\mathbf{x})$, $\bm{\mu}$ and $\bm{\sigma}$ (see Section \ref{methods_VI}).
-Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
-A key advantage of EPI over established Bayesian inference is that the predictions made by the inferred distribution are constrained to produce the specified emergent property.
-%Equipped with this method, we may examine structure in such inferred distributions or make comparisons between inferred distributions at graded values of emergent property statistics.
-Equipped with this method, we may examine structure in posterior distributions or make comparisons between posteriors conditioned at different levels of the same emergent property statistic.
-In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using it to investigate and produce novel insights into two prominent models in neuroscience. 
-Subsequently in Section \ref{results_LRRNN}, we show EPI's superiority in parameter scalability and fidelity of the posterior predictive distribution by conditioning on stable amplification in low-rank RNNs.
+%where $f(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
+%According to the emergent property of interest, $f(\mathbf{x}, \mathbf{z})$ may contain multiple statistics, in which case the mean and variance vectors $\bm{\mu}$ and $\bm{\sigma}^2$ match this dimension.
+%Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
+%This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
+%Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness. 
+%The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.
+
+%EPI optimizes the weights and biases $\bm{\theta}$ of the deep network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
+%The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
+%%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
+%Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}).
+%Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).
+
+%In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
+%Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
+%A key advantage of EPI over established Bayesian inference is that the predictions made by the inferred distribution are constrained to produce the specified emergent property.
+
+The probabilities of the distribution inferred from EPI are the densities of these parameters in the most random distribution producing the emergent property.
+While existing approaches to probabilistic structural identifiability analysis use the lens of evidence accumulation \cite{raue2009structural, gonccalves2019training}, sensitivity and robustness of parameter space dimensions with respect to emergent phenomena can be directly quantified with EPI.
+Sensitivity quantifications are measured by the second order derivative of EPI probability along the parameteric dimension of interest: to what extent is the emergent property maintained or diminished along this dimension?
+Once an EPI distribution has been inferred, this second order derivative requires trivial computation (as long as the correct architecture class is chosen, see Section \ref{methods_NF}).
+Equipped with this method, we may examine structure in the resulting parameter distributions or make comparisons between distributions conditioned at different levels of the same emergent property statistic.
+In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using these techniques to investigate and produce novel scientific insight.
 

### 2021-03-15 written/eLifeRevision/Bittner_eLife2020.tex
edits

@@
 \section{Abstract}
-A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
-Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or in terms of neural activity -- and thus can offer insights into neural computation.
-The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
 %When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
 %We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
-We present a novel technique designed to infer parameter distributions that produce mathematically defined emergent properties of computation.
-We detail how the parameter distributions of theoretical neuroscience problems are better defined by the emergent properties they produce, rather than conditioned upon artificial datasets.
-This approach is not only well-motivated, but allows technical simplifications to statistical inference in mechanistic models and scaling to high dimensions.
+A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
+Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
+The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
+A key process in neuroscientific modeling is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
+While statistical inference has proven effective on a broad variety of neuroscientific datasets, we clarify an important incongruity between theoretical approaches to neuroscience and this probabilistic methodology.
+Theoretical neuroscience is focused on computational properties and how they emerge from biological mechanisms, rather than noisy experimental datasets and their quantified structure.
+In this work, we present a novel technique to directly infer circuit model parameters producing these computational properties.
+This method tailors deep inference, the use of deep neural networks for statistical inference, to the nature of theoretical inverse problems, enabling scaling to high dimensions and considerable technical simplifications.
 With this method, we bring deep inference to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
-First, we show a worked example analyzing parametric sensitivity in the stomatogastric ganglion, emphasizing the general applicability of this approach to complex models with complex emergent properties.
-Then, in a model of primary visual cortex where mathematical analysis becomes untenable, we identify a nonlinear manifold revealing the rule of parametric degeneracy with respect to variability.
-Next, we gain a mechanistic understanding of how superior colliculus connectivity produces rapid task switching in multiple parametric regimes.
-Finally, we scale inference of RNN connectivity exhibiting stable amplification to 1,000-dimensional parameter spaces.
+First, we emphasize the general applicability of this approach by inferring channel conductance parameters that produce a distinctive spiking frequency in a biophysical model of the stomatogastric ganglion.
+Then, in a model of primary visual cortex with multiple neuron-types, where  analysis becomes untenable as more neuron-types are included, we use this technique to discover how noise properties govern excitatory population variability.
+Next, in a model of superior colliculus, we identify and characterize two distinct regimes of connectivity that facilitate switching between opposite tasks throughout interleaved trials. 
+Finally, we scale inference to 1,000-dimensional parameter spaces of RNN connectivity that exhibit stable, yet amplified responses.
 These analyses illustrate how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
@@
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-Alternatively, we can gain insight into these complex models by identifying the full distribution of parameters consistent with specified emergent phenomena.
+Alternatively, we can gain insight into these complex models by identifying the distribution of parameters that produce a specific computational property.
 By solving the inverse problem in this way, scientific analysis of complex biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.

### 2021-03-16 scripts/RNN/cpu_stab_amp_snpe.sh
snpe script updates

@@ -3,13 +3,7 @@
 #SBATCH --job-name=rnn_snpe
 #SBATCH -c 1
 #SBATCH --time=6:00:00
-#SBATCH --mem-per-cpu=1gb
+#SBATCH --mem-per-cpu=2gb
 
-echo $1
-echo $2
-echo $3
-echo $4
-echo $5
-echo $6
 source activate sbi
-python3 stab_amp_snpe.py --N $1 --num_sims $2 --num_batch $3 --num_transforms $4 --num_atoms $5 --g $6 --K $7 --rs $8
+python3 stab_amp_snpe.py --N $1 --num_sims $2 --num_batch $3 --num_transforms $4 --num_atoms $5 --g $6 --K $7 --rs $8 --max_rounds 100

### 2021-03-16 scripts/RNN/stab_amp_snpe.py

@@ -110,32 +110,33 @@ for r in range(max_rounds):
         break
     proposal = posterior.set_default_x(x_0)
 
-zs = [] 
-xs = []
-log_probs = [] 
-distances = []
-M = 1000
-for posterior in posteriors: 
-    z = posterior.sample((M,), x=x_0)
-    x = simulator(z).numpy()
-    log_prob = posterior.log_prob(z, x=x_0)
-    median_x = np.median(x, axis=0)
-    distance = np.linalg.norm(median_x - x_0.numpy())
+    zs = [] 
+    xs = []
+    log_probs = [] 
+    distances = []
+    M = 1000
+    for posterior in posteriors: 
+        z = posterior.sample((M,), x=x_0)
+        x = simulator(z).numpy()
+        log_prob = posterior.log_prob(z, x=x_0)
+        median_x = np.median(x, axis=0)
+        distance = np.linalg.norm(median_x - x_0.numpy())
 
-    zs.append(z.numpy())
-    xs.append(x)
-    log_probs.append(log_prob.numpy())
-    distances.append(distance)
+        zs.append(z.numpy())
+        xs.append(x)
+        log_probs.append(log_prob.numpy())
+        distances.append(distance)
 
-optim = {'summary':inference._summary, 
-         'round_val_log_probs':np.array(round_val_log_probs), 
-         'zs':np.array(zs), 
-         'xs':np.array(xs), 
-         'log_probs':np.array(log_probs),
-         'distances':np.array(distances),
-         'times':times}
+        optim = {'summary':inference._summary, 
+                 'round_val_log_probs':np.array(round_val_log_probs), 
+                 'zs':np.array(zs), 
+                 'xs':np.array(xs), 
+                 'log_probs':np.array(log_probs),
+                 'distances':np.array(distances),
+                 'times':times}
 
-print('Saving', save_path, '...', flush=True)
-with open(os.path.join(save_path, "optim.pkl"), "wb") as f:
-    pickle.dump(optim, f)
+    print('Saving', save_path, '...', flush=True)
+    with open(os.path.join(save_path, "optim.pkl"), "wb") as f:
+        pickle.dump(optim, f)
+
 print('done.', flush=True)

### 2021-03-16 scripts/RNN/stab_amp_snpe_sweep.sh

@@ -2,15 +2,15 @@
 
 # This shell script does it linearly, but could you run each of these 
 # python scripts independently on different instances?
-for N in 50 100 250
+for N in 2
 do
-  for g in 0.0001
+  for g in 0.0001 0.01 0.1 0.25
   do
-    for K in 1
+    for K in 1 5 100
     do
-      for num_sims in 1000
+      for num_sims in 2000
       do
-        for num_batch in 50 250
+        for num_batch in 100
         do
           for num_atoms in 100
           do

### 2021-03-16 scripts/RNN/stab_amp_epi.py
stab amp epi update

@@ -20,15 +20,15 @@ parser = argparse.ArgumentParser()
 parser.add_argument('--N', type=int)
 parser.add_argument('--g', type=float, default=0.01)
 parser.add_argument('--K', type=int, default=1)
-parser.add_argument('--c0', type=float)
-parser.add_argument('--rs', type=int)
+parser.add_argument('--c0', type=float, default=0.)
+parser.add_argument('--rs', type=int, default=1)
 args = parser.parse_args()
 
 print('Running epi for RNN N=%d, c0=%f, rs=%d stable amplification.' % (args.N, args.c0, args.rs))
 N = args.N
 g = args.g
 K = args.K
-c0 = args.c0
+c0 = 10**(args.c0)
 rs = args.rs
 
 r = 2 # rank-2 networks
@@ -51,21 +51,22 @@ parameters = [U, V]
 M = Model("Rank2Net", parameters)
 
 # 2. Define the emergent property:
-Js_eig_max_mean = 1.5
 J_eig_realmax_mean = 0.5
-mu = np.array([Js_eig_max_mean, 
-               0.25**2, 
-               J_eig_realmax_mean,
-               0.25**2], dtype=DTYPE)
+Js_eig_max_mean = 1.5
+eig_std = 0.25
+mu = np.array([J_eig_realmax_mean,
+               Js_eig_max_mean,
+               eig_std**2, 
+               eig_std**2], dtype=DTYPE)
 
 W_eigs = get_W_eigs_tf(g, K)
-def stable_amplification_r2(U, V):
+def stable_amp(U, V):
     U = tf.reshape(U, (-1, N, 2))
     V = tf.reshape(V, (-1, N, 2))
     T_x = W_eigs(U, V)
     return T_x
 
-M.set_eps(stable_amplification_r2)
+M.set_eps(stable_amp)
 
 q_theta, opt_data, save_path, failed = M.epi(
     mu, 

### 2021-03-16 scripts/RNN/stab_amp_epi.py
latest stab amp scripts

@@ -48,7 +48,7 @@ V = Parameter("V", D, lb=lb, ub=ub)
 parameters = [U, V]
 
 # Define the model matrix.
-M = Model("Rank2Net", parameters)
+M = Model("Rank2Net_g=%.4f_K=%d" % (g, K), parameters)
 
 # 2. Define the emergent property:
 J_eig_realmax_mean = 0.5
@@ -56,7 +56,7 @@ Js_eig_max_mean = 1.5
 eig_std = 0.25
 mu = np.array([J_eig_realmax_mean,
                Js_eig_max_mean,
-               eig_std**2, 
+	       eig_std**2, 
                eig_std**2], dtype=DTYPE)
 
 W_eigs = get_W_eigs_tf(g, K)

### 2021-03-22 neural_circuits/LRRNN.py
latest LRRNN funcs

@@ -1,8 +1,10 @@
 import os
 import itertools
 import pickle
+import scipy
 import numpy as np
 import tensorflow as tf
+import matplotlib.pyplot as plt
 from epi.normalizing_flows import NormalizingFlow
 from epi.models import Model, Parameter
 from epi.util import get_max_H_dist
@@ -13,118 +15,147 @@ from sbi import analysis as analysis
 from sbi.inference import SNPE, prepare_for_sbi, simulate_for_sbi
 from sbi.utils.get_nn_models import posterior_nn
 
-#EPS = 1e-6
+# EPS = 1e-6
+
 
 def get_W_eigs_np(g, K, feed_noise=False):
     def W_eigs(U, V, noise=None):
-        U, V = U[None,:,:], V[None,:,:]
-        U, V = np.tile(U, [K,1,1]), np.tile(V, [K,1,1])
+        U, V = U[None, :, :], V[None, :, :]
+        U, V = np.tile(U, [K, 1, 1]), np.tile(V, [K, 1, 1])
         if feed_noise:
             U_noise, V_noise = noise
-            U = U + g*U_noise
-            V = V + g*V_noise
+            U = U + g * U_noise
+            V = V + g * V_noise
         else:
-            U = U + g*np.random.normal(0., 1., U.shape)
-            V = V + g*np.random.normal(0., 1., V.shape)
-        J = np.matmul(U, np.transpose(V, [0,2,1]))
-        Js = (J + np.transpose(J, [0,2,1])) / 2.
+            U = U + g * np.random.normal(0.0, 1.0, U.shape)
+            V = V + g * np.random.normal(0.0, 1.0, V.shape)
+        J = np.matmul(U, np.transpose(V, [0, 2, 1]))
+        Js = (J + np.transpose(J, [0, 2, 1])) / 2.0
         Js_eigs = np.linalg.eigvalsh(Js)
         Js_eig_maxs = np.max(Js_eigs, axis=1)
         Js_eig_max = np.mean(Js_eig_maxs)
 
         # Take eig of low rank similar mat
-        Jr = np.matmul(np.transpose(V,[0,2,1]), U) #+ EPS*np.eye(2)[None,:,:]
+        Jr = np.matmul(np.transpose(V, [0, 2, 1]), U)  # + EPS*np.eye(2)[None,:,:]
         Jr_tr = np.trace(Jr, axis1=1, axis2=2)
[…]

### 2021-03-22 neural_circuits/LRRNN.py
refined LRRNN plots

@@ -534,7 +534,7 @@ def eig_scatter(T_xs, colors, ax=None, perm=True):
         T_x, c = T_x[perm], c[perm]
     ax.plot([0.5], [1.5], "*", c=gray, markersize=10)
     ax.scatter(
-        T_x[:, 0], T_x[:, 1], c=c, edgecolors="k", linewidth=0.5, s=50, alpha=alpha
+        T_x[:, 0], T_x[:, 1], c=c, edgecolors="k", linewidth=0.5, s=60, alpha=alpha
     )
     ax.set_yticks([0, 1, 2, 3, 4])
     ax.set_xlim([-1, 3])
@@ -575,9 +575,9 @@ def count_str(n):
 
 def plot_compare_hp(epis, snpes1, snpes2, c_epi, pal, 
                epi_label=None, snpe1_label=None, snpe2_label=None,
-               xlims=None,
+               xlims=None, fontsize=12
                ):
-    fig, axs = plt.subplots(1,2,figsize=(10,5))
+    fig, axs = plt.subplots(1,2,figsize=(10,3))
     
     max_sims, max_time, max_dist = snpe_simplot(snpes1, snpe1_label, axs, pal[0])
     max_sims, max_time, max_dist = snpe_simplot(snpes2, snpe2_label, axs, pal[1], \
@@ -601,11 +601,11 @@ def plot_compare_hp(epis, snpes1, snpes2, c_epi, pal,
         labeled=True
         
     if xlims is not None:
-        if max_sims > xlims[0]:
-            axs[0].set_xlim([0, xlims[0]])
+        _xlim = min(max_sims, xlims[0])
+        axs[0].set_xlim([0, _xlim])
     xticks = axs[0].get_xticks()
     xticklabels = [count_str(count) for count in xticks]
-    axs[0].set_xticklabels(xticklabels)
+    axs[0].set_xticklabels(xticklabels, fontsize=fontsize)
     axs[0].set_ylim([0, 1.1*max_dist])
     axs[0].legend(fontsize=12, loc="upper right")
 
@@ -613,11 +613,14 @@ def plot_compare_hp(epis, snpes1, snpes2, c_epi, pal,
     xticks = hrs*60
     xticklabels = ["0"] + ["%dhr" % hr for hr in hrs[1:]]
     axs[1].set_xticks(xticks)
-    axs[1].set_xticklabels(xticklabels)
+    axs[1].set_xticklabels(xticklabels, fontsize=fontsize)
[…]

### 2021-03-23 written/eLifeRevision/Bittner_eLife2020.tex
incorporate LRRNN figs
latest updates

@@
 \subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
-%In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}a).
-%Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
-%Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
-%We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig.
-\ref{fig:SC}B).
-%The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
-%\begin{equation}
-%x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
-%\end{equation}
-%where $\epsilon = 0.05$ and $\zeta = 0.5$.  
-%The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
-%\begin{equation}
-%\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
-%\end{equation}
-%with gaussian noise of variance $\sigma^2$ = 1.
-%The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-%Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
-%\end{center}
-%\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
-%A. Rapid task switching behavioral paradigm (see text). 
-%B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
-%Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-%Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
-%C.
-The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$.
-Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.
-%D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
-%E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
-%\label{fig:SC}
-%\end{figure}
-%
-%To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
-%At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
-%Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
-%Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
-%\begin{equation}
-%\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
-%\end{equation}
-%Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.
-%
-%We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
-%To make sense of these inferred distributions, we followed the approach of Duan et al.
-by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig.
-\ref{fig:SC}C).
-%These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-%The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
-%
-%We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig.
-\ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
-%Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
-%Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
-%There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
-%Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
-%These hypotheses are confirmed by forward simulation of the SC model (Fig.
-\ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.
-
 In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
@@
 \subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
-Here, we study the scalability of EPI in number of parameters $|\mathbf{z}|$ by inferring the connectivities of recurrent neural networks (RNNs, Fig. \ref{fig:LRRNN}A). 
- We consider a rank-2 RNN with N neurons having connectivity
-\begin{equation}
- W = UV^\top + g\chi
- \end{equation}
- and dynamics
- \begin{equation}
- \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
- \end{equation}
-where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$ and $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$.
-The random component has strength $g=0.01$ and $\chi_{i,j} \sim \mathcal{N}(0, 1)$.  
-We  infer connectivity distributions $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   producing stable amplification.
-RNN's exhibiting stable amplification amplify responses to input along some dimensions, and are stable across all dimensions.
-Two conditions are both necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
-$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$.
-
+Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinisically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+It has been shown that to generate such amplified, yet stabilized responses, the connectivity of RNNs must be non-normal \cite{goldman2009memory, murphy2009balanced}, and satisfy additional constraints \cite{bondanelli2020coding}.
+In theoretical neuroscience, RNNs are optimized and then examined to show how dynamical systems could execute a given computation \cite{sussillo2014neural, barak2017recurrent}, but such biologically realistic constraints on connectivity are  ignored during optimization for practical reasons.
+In general, access to distributions of connectivity adhering to theoretical criteria like stable amplification, chaotic fluctuations \cite{sompolinsky1988chaos}, or low tangling \cite{russo2018motor} would add great scientific value and contextualization to existing research with RNNs.
+Here, we use EPI to learn RNN connectivities producing stable amplification, and demonstrate the superior scalability and efficiency of EPI to alternative approaches.
 \begin{figure}
@@
 \caption{\small 
-\textbf{A}. Recurrent neural network.
-\textbf{B}. EPI scales with $z$ to high dimensions. 
-Convergence definitions: 
-EPI (blue) - satisfies all moment constraints, 
-SNPE (orange)- produces at least 2/$n_{\text{train}}$ parameter samples are in the bounds of emergent property (mean +- 0.5), 
-and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.
-\textbf{C}. Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
-Gray star indicates emergent property mean, and gray dashed lines indicate two standard deviations corresponding to the variance constraint.
-For $N <= 6$ where SMC-ABC converges, samples are not diverse (path degeneracies).  
-For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
+\textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivites producing stable amplification.  
+Each dot shows convergence time for an individual random seed.
+For reference, the mean wall time for EPI to achieve its full constraint convergence (means and variances) is shown (blue line).
+\textbf{B}. Simulation count of each algorithm to achieve convergence. Same conventions as A.
+\textbf{C}. The predictive distributions of connectivities inferred by EPI (blue), SNPE (orange), and SMC-ABC (green), with reference to $\mathbf{x}_0 = \bm{\mu}$ (gray star).
 \textbf{D}.
-Simulations of network parameters resulting from each method ($\tau=100ms$).
-Each trace corresponds to simulation of one $z$.
-(Below) Ratio of obtained samples producing stable amplification.
+Simulations of networks inferred by each method ($\tau=100ms$).
+Each trace (15 per algorithm) corresponds to simulation of one $z$.
+(Below) Ratio of obtained samples producing stable amplification, monotnoic decay, and instability.
 }
@@
 
-In our analysis, we seek to condition rank-2 networks of increasing size on a regime of stable amplification.  
-Networks with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will yield moderate amplification.
+We consider a rank-2 RNN with N neurons having connectivity $W = UV^\top$
+ and dynamics
+ \begin{equation}
+ \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x},
+ \end{equation}
+where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   that produce stable amplification.
+Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
+$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$.
+RNNs  with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will be stable with modest decay rate ($\text{real}(\lambda_1)$ close to its upper bound of 1) and exhibit modest amplification ($\lambda_1^s$ close to its lower bound of 1).
  EPI can naturally condition on this emergent property
@@
 \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
-\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix} .
+\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix},
 \end{split}
 \end{equation}
-
-For comparison, we infer rank-2 RNN connectivites with alternative approaches to likelihood free-inference.
-ABC methods define a tolerance $\epsilon$ from observed data $x_0$ for which we keep sampled parameters.
-To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
-\begin{equation}\label{eq:SNPE_stab_amp_x0}
-x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
-\end{equation}
-located at the mean of our desired emergent property.
-We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the most efficient ABC approach.
-SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
-In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
-Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
-
-As we scale the number of neurons $N$ in the RNN, and thus the dimensionality of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI has superior scaling properties (Fig.
-\ref{fig:LRRNN}B).
-SMC-ABC and SNPE become intractable around 25 and 90 dimensions respectively, while EPI can infer 1000-dimensional distributions in about 1 day.
-No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ of $\mathcal{X}$(Fig.
-\ref{fig:LRRNN}C, blue), and high variation in response profiles \ref{fig:LRRNN}D, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification, are less varied \ref{fig:LRRNN}C, red) and largely produce similar responses \ref{fig:LRRNN}D, red).
-When using SNPE the inferred parameters are widely varied \ref{fig:LRRNN}C, orange), but often produce non-amplified or unstable responses \ref{fig:LRRNN}D, orange).
-In conclusion, we found that deep likelihood-free inference techniques are capable of scaling to higher dimensional inference than SMC-ABC.
-However, only EPI can scale to high dimensions while reproducing the emergent property.
+under the notion that variance constraints with standard deviation 0.25 predicate that the vast majority of samples (those within two standard deviations) are within the specified ranges.
+
+For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative likelihood-free inference approaches.
+We ran sequential monte-carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
+SMC-ABC is a rejection sampling approach that uses sophisticated techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
+Unlike EPI, these statistical inference techniques do not control the mean or variance of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
+$N$ and $g$, Fig.
+\ref{fig:RNN1}).
+To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.25 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
+
+As the number of neurons $N$ in the RNN are scaled, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
+\ref{fig:LRRNN}A).
+It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:LRRNN}B).
+It is well known that ABC techniques struggle mightily in dimensions greater than about 30 \cite{sisson2018handbook}, yet we were careful to assess the scalability of the more comparable approach SNPE.
+Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
+\ref{fig:RNN2}), and tested more aggressive SNPE hyperparameterizations when SNPE failed to converge (Fig.
+\ref{fig:RNN3}).
+
+No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig.
+\ref{fig:LRRNN}C, blue).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification  (Fig.
+\ref{fig:LRRNN}C, green), but are concentrated and offset from $\mathbf{x}_0$.
+When using SNPE the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable amplified activity norms $|r(t)|$, SMC produces a limited variety in respones, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
+
+From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and have the additional benefit of fast sampling after optimization.
+EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration and constantly adapting the approximation $q_{\bm{\theta}}(\mathbf{z})$ to the parameter distribution.
+While EPI can be used for scientific insight into RNNs, this analysis shows that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
 
@@
 \begin{flushleft}
-\caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
-B.
+\caption{\small (LDS1): \textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
+\textbf{B}.
 The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.
 Dashed lines indicate the parameter axes.
-C. Entropy throughout the optimization.  
+\textbf{C}. Entropy throughout the optimization.  
 At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
-D. Emergent property moments throughout optimization.  
+\textbf{D}. Emergent property moments throughout optimization.  
 At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
@@
 \begin{flushleft}
-\caption{\small (LDS2): A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\caption{\small (LDS2): \textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+\textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
 \end{flushleft}
@@
 \caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
-A.
-Each dimension of the simulated trajectories throughout time.
-B  The simulated trajectories in phase space.}
+\textbf{A}. Each dimension of the simulated trajectories throughout time.  
+\textbf{B}.  The simulated trajectories in phase space.}
 \end{flushleft}
@@
 \caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
-A. Entropy throughout optimization. 
-B. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
+\textbf{A}. Entropy throughout optimization. 
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
 \end{flushleft}
@@
 \subsubsection{Superior colliculus}\label{methods_SC}
-%In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
-%They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
-%Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
-%\begin{equation}
-%x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
-%\end{equation}
-%where $\alpha \in \{LP, LA, RA, RP\}$ $\epsilon = 0.05$ and $\zeta = 0.5$ control the position and shape of the nonlinearity, respectively.
-%
-%We order the elements of $x$ and $u$ in the following manner
-%\begin{equation}
-%x = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} u = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix}.
-%\end{equation}
-%
-% The internal variables follow dynamics:
-%\begin{equation}
-%\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
-%\end{equation}
-%with time constant $\tau = 0.09s$ and Gaussian noise $\sigma dB$ controlled by the magnitude of $\sigma=1.0$.
-The weight matrix has 8 parameters $sW_P$, $sW_A$, $vW_{PA}$, $vW_{AP}$, $hW_P$, $hW_A$, $dW_{PA}$, and $dW_{AP}$ (Fig.
-4B):
-%\begin{equation}
-%W = \begin{bmatrix} sW_P & vW_{PA} & hW_P & dW_{PA}  \\ vW_{AP}  & sW_A & dW_{AP}  & hW_A \\ hW_P & dW_{PA}  & sW_P & vW_{PA}  \\ dW_{AP}  & hW_A & vW_{AP}  & sW_A \end{bmatrix}.
-%\end{equation}
-%
-%The system receives five inputs throughout each trial, which has a total length of 1.8s.
-%\begin{equation}
-%h = h_{\text{rule}} + h_{\text{choice-period}} + h_{\text{light}}.
-%\end{equation}
-%
-%There are rule-based inputs depending on the condition,
-%\begin{equation}h_{\text{P,rule}}(t) = \begin{cases}
-%                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
-%                            0,              & \text{otherwise}
-%                         \end{cases}
-%\end{equation}
-%\begin{equation} h_{\text{A,rule}}(t) = \begin{cases}
-%                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
-%                            0,              & \text{otherwise}
-%                         \end{cases}
-%\end{equation}
-%a choice-period input,
-%\begin{equation} h_{\text{choice}}(t) = \begin{cases}
-%                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
-%                            0,              & \text{otherwise}
-%                         \end{cases}
-%\end{equation}
-%and an input to the right or left-side depending on where the light stimulus is delivered.     
-%\begin{equation}  h_{\text{light}}(t) = \begin{cases}
-%                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } t > 1.2s \text{ and Left} \\
-%                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } t > 1.2s \text{ and Right} \\
-%                            0,              & t \leq 1.2s
-%                         \end{cases} .
-%\end{equation}
-%The input parameterization was fixed to $I_{\text{P,rule}} = 10 $,  $I_{\text{A,rule}} = 10$,  $I_{\text{choice}} = 2$,  and $I_{\text{light}} = 1$.
-%
-%To produce an accuracy rate of $p_{LP}$ in the Left, Pro condition, let $\hat{p}_i$ be the empirical average steady state response (final $x_{LP}$ at end of task) over M=500 Gaussian noise draws for a given SC model parameterization $z_i$:
-%\begin{equation}
-% \hat{p}_i = \mathbb{E}_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] = \frac{1}{M}\sum_{j=1}^M x_{LP}(s=L, c=P, z=z_i, \sigma dB_j)
-% \end{equation}
-% where stimulus $s \in \{L, R\}$, cue $c \in \{P, A\}$, and $\sigma dB_j$ is the Gaussian noise on trial $j$.
-%As with the V1 model, we only consider steady state responses of $x$, so $x_\alpha$ is used from here on to denote the steady state activity at the end of the trial.
-%For the first emergent property statistic, the average over EPI samples (from $q_{\bm{\theta}}(z)$) is set to the desired value $p_{LP}$:
-%\begin{equation}
-%\mathbb{E}_{z_i \sim q_\phi} \left[ \mathbb{E}_{\sigma dB} \left[ x_{LP,\text{ss}} \mid s=L, c=P, z=z_i \right] \right] = \mathbb{E}_{z_i \sim q_\phi} \left[ \hat{p}_i \right] = p_{LP}.
-%\end{equation}
-%
-%For the next emergent property statistic, we ask that the variance of the steady state responses across Gaussian draws, is the Bernoulli variance for the empirical rate $\hat{p}_i$:
-%\begin{equation}
-%\mathbb{E}_{z \sim q\phi} \left[ \sigma^2_{err} \right] = 0
-%\end{equation}
-%where the Bernoulli variance error $\sigma^2_{err}$ for the Pro task, left condition is
-%\begin{equation}
-%\sigma^2_{err} = Var_{\sigma dB} \left[ x_{LP} \mid s=L, c=P, z=z_i \right] - \hat{p}_i(1 - \hat{p}_i).
-%\end{equation}
-%
-%We have an additional constraint that the Pro neuron on the opposite hemisphere should have the opposite value (0 and 1).  We can enforce this with another constraint:
-%\begin{equation}
-%\mathbb{E}_{z \sim q\phi} \left[ d_P \right] = 1,
-%\end{equation}
-%where the distance between Pro neuron steady states $d_P$ in the Pro condition is
-%\begin{equation}
-%d_P =  \mathbb{E}_{\sigma dB} \left[ (x_{LP} - x_{RP})^2  \mid s=L, c=P, z=z_i \right]
-%\end{equation}
-%The emergent property statistics only need to be measured during the Left stimulus condition of the Pro and Anti tasks, since the network is symmetrically parameterized.
-%In total, the emergent property of rapid task switching at accuracy level $p$ was defined as
-%\begin{equation}
-%\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
-%\end{equation}
-%
-%Since the maximum variance of a random variable bounded from 0 to 1 is the Bernoulli variance $\hat{p}(1-\hat{p})$, and the maximum squared difference between to variables bounded from 0 to 1 is 1, we do not need to control the second moment of these test statistics.
-%These variables are dynamical system states and can only exponentially decay (or saturate) to 0 (or 1), so the Bernoulli variance error and squared difference constraints cannot be satisfied exactly in simulation.
-%This is important to be mindful of when evaluating the convergence criteria.  
-%Instead of using our usual hypothesis testing criteria for convergence to the emergent property, we set a slack variable threshold only for these technically infeasible emergent property values to 0.05.
-%
-%Using EPI to learn distributions of dynamical systems producing Bernoulli responses at a given rate (with small variance around that rate) was more challenging than expected.  
-%There is a pathology in this optimization setup, where the learned distribution of weights is bimodal attributing a fraction $p$ of the samples to an expansive mode (which always sends $x_{LP}$ to 1), and a fraction $1-p$ to a decaying mode (which always sends $x_{LP}$ to 0).
-%This pathology was avoided using an inequality constraint prohibiting parameter samples that resulted in low variance of responses across noise.
-%
-%
-%\begin{table}[h]
-%\begin{tabular}{l|l|l|l|l}
-%$\lambda$  & $\hat{p}$ & $q_{\bm{\theta}}(z)$ & $r$ & p-value \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $1.24 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $7.56 \times 10^{-01}$ & $p<10^{-4}$ \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $4.59 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $3.76 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%
-%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $4.80 \times 10^{-02}$ & $p<.01$ \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $2.08 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $4.84 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{task}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(90\%))$ & $4.25 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%
-%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(50\%))$ & $-7.57 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(60\%))$ & $-6.73 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(70\%))$ & $-4.86 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(80\%))$ & $-1.43 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_P$ & $q(z \mid \mathcal{B}(90\%))$ & $-1.93 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%
-%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(60\%))$ & $-7.60 \times 10^{-02}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(70\%))$ & $-2.73 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%$\lambda_{\text{side}}$ & $\hat{p}_A$ & $q(z \mid \mathcal{B}(80\%))$ & $-2.74 \times 10^{-01}$ & $p<10^{-4}$  \\ \hline
-%\end{tabular}
-%\caption{\small Table of significant correlation values from Fig. \ref{fig:SC}E.}
-%\end{table}
-%
-%For each accuracy level $p$, we ran EPI for 10 different random seeds  using an architecture of 10 planar flows with a  support of $z \in \mathbb{R}^8$.   
-%We used an augmented Lagrangian coefficient of $c_0 = 10^2$, a batch size $n=300$, and set $\nu = 0.5$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian of zero mean with standard deviation $\sigma_{\text{init}} = 1$.
-%The EPI distributions shown in Fig. \ref{fig:SC} are the converged distributions with maximum entropy across random seeds.
-%
-%We report significant correlations $r$ and their p-values from Figure \ref{fig:SC}E in Table 1.  
-%Correlations were measured from 5,000 samples of $q_{\bm{\theta}}(z \mid \mathcal{B}(p))$ and p-values are reported for one-tailed tests, since we hypothesized a positive correlation between task accuracies $p_P$ or $p_A$ and $\lambda_{\text{task}}$, and a negative correlation between task accuracies $p_P$ and $p_A$ and $\lambda_{\text{side}}$.
-
 In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
@@
 \caption{\small (SC6): 
-A. Entropy throughout optimization. 
-B. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+\textbf{A}. Entropy throughout optimization. 
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
 }
@@
 
-\subsubsection{Rank-2 RNN}\label{methods_RNN}
-Traditional approaches to likelihood-free inference -- approximate Bayesian computation (ABC) methods -- randomly sample parameters $\mathbf{z}$ until a suitable set is obtained.
-State-of-the-art ABC methods leverage sequential Monte Carlo (SMC) sampling techniques to obtain parameter sets more efficiently.
-To obtain more parameter samples, SMC-ABC must be run from scratch again.
-ABC methods do not confer log probabilities of samples.
- Like EPI, sequential neural posterior estimation (SNPE) uses deep learning to produce flexible posterior approximations.
-Like traditional Bayesian inference methods, SNPE conditions directly on the statistics of data.
-This differs from EPI, where posteriors are conditioned on emergent properties (moment constraints on the posterior predictive distribution).
-Peculiarities of SNPE (density estimation approach, two deep networks) make scaling in $\mathbf{z}$ prohibitive.
-
+\subsubsection{Scaling EPI for stable amplification in RNNs}\label{methods_RNN}
+We examine the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
+Rank-2 RNN connectivity is modeled as $W = UV^\top$, where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+This RNN model has dynamics
+\begin{equation}
+\tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}.
+\end{equation}
+In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification.
+
+For this RNN model to be stable, all real eigenvalues of $W$ must be less than 1:  $\text{real}(\lambda_1) < 1$, where $\lambda_1$ denotes the greatest real eigenvalue of $W$.
+For a stable RNN to amplify at least one input pattern, the symmetric connectivity $W^s = \frac{W + W\top}{2}$ must have an eigenvalue greater than 1:
+$\lambda^s_1 > 1$, where  $\lambda^s$ is the maximum eigenvalue of $W^s$.
+These two conditions are necessary and sufficient for stable amplification in RNNs\cite{bondanelli2020coding}.
+We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions and variances ($0.25^2$) about those means that rarely violate the eigenvalue constraints.
+As a maximum entropy distribution, this is written as
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \lambda_1^s(\mathbf{x}; \mathbf{z}) \\ \left(\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) - 0.5 \right)^2 \\ \left( \lambda_1^s(\mathbf{x}; \mathbf{z})  - 1.5 \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} 0.5 \\ 1.5 \\ 0.25^2 \\ 0.25^2 \end{bmatrix}.
+\end{equation}
+For EPI in Fig. \ref{fig:LRRNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
+The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
+We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented LaGrangian epoch.
+
+ABC methods define a tolerance $\epsilon$ from observed data $x_0$ for which we keep sampled parameters.
+To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
+\begin{equation}\label{eq:SNPE_stab_amp_x0}
+x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
+\end{equation}
+located at the mean of our desired emergent property.
+We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the most efficient ABC approach.
+SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
+In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
+Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
+
+\textbf{How we did EPI} \\
+
+\textbf{How we did SNPE} \\
 SMC-ABC has many hyperparameters, of which pyABC selects automatically by running some initial diagnostics upon initialization.
 In concurrence with the literature, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
+
+\textbf{How we did SNPE} \\
 We searched over many SNPE hyperparameter choices: $n_{\text{train}} \in [2,000, 10,000, 100,000]$ is the number of simulations run per training epoch, and $n_{\text{mades}} \in [2, 3]$ is the number of masked autoregressive density estimators in the deep parameter distribution architecture.
@@
 
-%\subsubsection{Rank-1 RNN}\label{methods_LRRNN}
-%Extensive research on random fully-connected recurrent neural networks has resulted in foundational theories of their activity \cite{sompolinsky1988chaos, brunel2000dynamics}.
-%Furthermore, independent research on training these models to perform computations suggests that learning occurs through low-rank perturbations to the connectivity (e.g.
-\cite{jaeger2004harnessing, sussillo2009generating}).
-%Recent theoretical work extends theory for random neural networks \cite{sompolinsky1988chaos} to those with added low-rank structure \cite{mastrogiuseppe2018linking}.
-%In Section \ref{results_RNN}, we used this theory to enable EPI on RNN parameters conditioned on the emergent property of task execution.
-%
-%Such RNNs have the following dynamics: 
-%\begin{equation}
-%\frac{dx}{dt} = -x + W \phi(x) + h,
-%\end{equation}
-%where $x$ is network activity, $W$ is the connectivity weight matrix, $\phi(\cdot) = \tanh(\cdot)$ is the input-output function, and $h$ is the input to the system.
-%In a rank-1 RNN (which was sufficiently complex for the Gaussian posterior conditioning task), $W$ is the sum of a random component with strength $g$ and a structured component determined by the outer product of vectors $m$ and $n$:
-%\begin{equation}
-%W = g\chi + \frac{1}{N}mn^\top,
-%\end{equation}
-%where  $\chi_{ij} \sim \mathcal{N}(0, \frac{1}{N})$, and the entries of $m$ and $n$ are distributed as $m_i \sim \mathcal{N}(M_m, 1)$ and $n_i \sim \mathcal{N}(M_n, 1)$. 
-%For EPI, we consider $z = [g, M_m, M_n]$, which are the parameters governing the connectivity properties of the RNN.
-%
-%From such a parameterization $z$, the theory of Mastrogiuseppe et al. produces solutions for variables describing the low dimensional response properties of the RNN.
-%These ``dynamic mean field" (DMF) variables (e.g.
-the activity along a vector $\kappa_v$, the total variance $\Delta_0$, structured variance $\Delta_\infty$, and the chaotic variance $\Delta_T$) are derived to be functions of one another and connectivity parameters $z$.
-%The collection of these derived functions results in a system of equations, whose solution must be obtained through a nonlinear system of equations solver.
-%The iterative steps of this system of equations solver are differentiable, so we take gradients through this solve process.
-%The DMF variables provide task-relevant information about the RNN's response to task inputs.
-%
-%In the Gaussian posterior conditioning example, $\kappa_r$ and $\Delta_T$ are DMF variables used as task-relevant emergent property statistics $\mu_{\text{post}}$ and $\sigma^2_{\text{post}}$.
-%Specifically, we solve for the DMF variables $\kappa_r$, $\kappa_n$, $\Delta_0$ and $\Delta_\infty$, where the readout is nominally chosen to point in the unit orthant $r = [1, ..., 1]^\top$.  
-%The consistency equations for these variables in the presence of a constant input $h = y r - (n - M_n)$ can be derived following \cite{mastrogiuseppe2018linking}:
-%{\small
-%\begin{equation}
-%\begin{split}
-%\kappa_r = G_1(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_m \kappa_n + y \\
-%\kappa_n = G_2(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = M_n \langle \left[ \phi_i \right] \rangle + \langle \left[ \phi_i' \right] \rangle \\
-%\frac{\Delta_0^2-\Delta_\infty^2}{2} = G_3(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty) = g^2 \left( \int \mathcal{D}z \Phi^2(\kappa_r + \sqrt{\Delta_0}z) - \int \mathcal{D}z \int \mathcal{D}x \Phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z)  \right) \\
-%+ (\kappa_n^2  + 1)(\Delta_0-\Delta_\infty) \\
-%\Delta_\infty = G_4(\kappa_r, \kappa_n, \Delta_0, \Delta_\infty)  = g^2 \int \mathcal{D}z \left[ \int \mathcal{D}x \phi(\kappa_r + \sqrt{\Delta_0 - \Delta_\infty}x + \sqrt{\Delta_\infty}z \right]^2 + \kappa_n^2 + 1
-%\end{split} 
-%\end{equation}
-%}
-%where here $z$ is a gaussian integration variable. 
-%We can solve these equations by simulating the following Langevin dynamical system to a steady state:
-%
-%\begin{equation}
-%\begin{split}
-%l(t) = \frac{\Delta_0(t)^2-\Delta_\infty(t)^2}{2} \\
-%\Delta_0(t) = \sqrt{2l(t) + \Delta_\infty(t)^2} \\
-%\frac{d \kappa_r(t)}{dt}  = -\kappa_r(t) + G_1(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-%\frac{d \kappa_n(t)}{dt} = -\kappa_n + G_2(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-%\frac{d l(t)}{dt}  = -l(t) + G_3(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t)) \\
-%\frac{d \Delta_\infty(t)}{dt}  = -\Delta_\infty(t) + G_4(\kappa_r(t), \kappa_n(t), \Delta_0(t), \Delta_\infty(t))
-%\end{split}.
-%\end{equation}
-%Then, the chaotic variance, which is necessary for the Gaussian posterior conditioning example, is simply calculated via $\Delta_T = \Delta_0 - \Delta_\infty$.
-%
-%We ran EPI using a real NVP architecture of two masks and two layers per mask with 10 units mapped to a support of $z = [g, M_m, M_n] \in [0,5] \times [-5,5] \times[-5,5]$ with no batch normalization.
-%We used an augmented Lagrangian coefficient of $c_0 = 1$, a batch size $n=300$, set $\nu = 0.15$, and initialized $q_{\bm{\theta}}(z)$ to produce an isotropic Gaussian with mean $\mu_{\text{init}} = [2.5, 0, 0]$ with standard deviation $\sigma_{\text{init}} = 2.0$.
-%The EPI distribution shown in Fig. \ref{fig:RNN} is the converged distributions with maximum entropy across five random seeds.
-%
-%To examine the effect of product $M_m M_n$ on the posterior mean, $\mu_{\text{post}}$ we took perturbations in $M_m M_n$ away from two representative parameters  $z_1$ and $z_2$ in 21 equally space increments from -1 to 1.
-%For each perturbation, we sampled 10 2,000-neuron RNNs and measure the calculated posterior means.  
-%In Fig. \ref{fig:RNN}D, we plot the product of $M_m M_n$ in the perturbation versus the average posterior mean across 10 network realizations with standard error bars.  
-%The correlation between perturbation product $M_m M_n$ and $\mu_{\text{post}}$ was measured over all simulations.  
-%For perturbations away from $z_1$ the correlation was 0.995 with $p<10^{-4}$, and for perturbations away from $z_2$ the correlation was 0.983 with $p<10^{-4}$.
-%
-%In addition to the Gaussian posterior conditioning example in Section \ref{results_RNN}, we modeled two tasks from Mastrogiuseppe et al.: noisy detection and context-dependent discrimination.  
-%We used the same theoretical equations and task setups described in their study.
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figRNN1/figRNN1.pdf}
+\end{center}
+\caption{\small (RNN1): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
+\textbf{A}.
+Predictive distribution of EPI (blue) and SNPE (orange) inferred connectivity of RNNs exhibiting stable amplification with $N=2$ (top), $N=10$ (bottom), $g=0.01$ (left), and $g=0.1$ (right).
+\textbf{B}. Entropy of parameter distribution approximations throughout optimization with $N=2$ (top), $N=10$ (bottom), $g=0.1$ (dark shade), and $g=0.01$ (light shade).
+\textbf{C}. Validation log probabilities throughout SNPE optimization. Same conventions as B.
+\textbf{D}. Adherance to EPI constraints. Same conventions as B.
+}
+\label{fig:RNN1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{figures/figRNN2/figRNN2.pdf}
+\end{center}
+\caption{\small (RNN2): 
+Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
+}
+\label{fig:RNN2}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.]{figures/figRNN3/figRNN3.pdf}
+\end{center}
+\caption{\small (RNN3): 
+SNPE convergence was enabled by increasing $n_{\text{round}}$, not $n_{\text{atom}}$.
+\textbf{A}.
+Difference of mean predictions $\mathbf{x}_0$ throughout optimization at $N=50$ with by simulation count (left) and wall time (right) of SNPE with $n_{\text{round}} = 5,000$ (light orange), SNPE with $n_{\text{round}} = 25,000$ (dark orange), and EPI (blue).
+Each line shows an individual random seed.
+\textbf{B}. Same conventions as A at $N=100$ of SNPE with $n_{\text{atom}} = 100$ (light orange) and $n_{\text{atom}} = 1,000$ (dark orange).
+\textbf{C}. Same conventions as A at $N=100$ of SNPE with $n_{\text{round}} = 25,000$ (light orange) and $n_{\text{round}} = 250,000$ (dark orange).
+}
+\label{fig:RNN3}
+\end{figure}
+
 \end{document}

### 2021-03-23 written/eLifeRevision/Bittner_eLife2020.tex
update again

@@
 $\lambda^s_1 > 1$, where  $\lambda^s$ is the maximum eigenvalue of $W^s$.
-These two conditions are necessary and sufficient for stable amplification in RNNs\cite{bondanelli2020coding}.
-We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions and variances ($0.25^2$) about those means that rarely violate the eigenvalue constraints.
+These two conditions are necessary and sufficient for stable amplification in RNNs \cite{bondanelli2020coding}.
+We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions and variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
 As a maximum entropy distribution, this is written as
@@
 \end{equation}
+Gradients of maximum eigenvalues of hermitian matrices like $W^s$ are available with modern automatic differentiation tools.
+To differentiate through the $\text{real}(\lambda_1)$, we solve the following equation for eigenvalues of rank-2 matrices using the rank reduced matrix $W^r = V^{\top} U$
+\begin{equation}
+\lambda_{\pm} = \frac{\text{Tr}(W^r) \pm \sqrt{\text{Tr}(W^r)^2 - 4\text{Det}(W^r)}}{2}.
+\end{equation}
+
 For EPI in Fig. \ref{fig:LRRNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
 The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented LaGrangian epoch.
+We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch.
 

### 2021-03-23 neural_circuits/LRRNN.py
plot update

@@ -609,14 +609,17 @@ def plot_compare_hp(epis, snpes1, snpes2, c_epi, pal,
     axs[0].set_ylim([0, 1.1*max_dist])
     axs[0].legend(fontsize=12, loc="upper right")
 
-    hrs = 4*np.arange(7)
-    xticks = hrs*60
-    xticklabels = ["0"] + ["%dhr" % hr for hr in hrs[1:]]
+    _xlim = min(max_time/60., xlims[1]) if xlims is not None else max_time/60.
+    if _xlim > 24*60:
+        hr_ticks = 12*np.arange(4)
+    else:
+        hr_ticks = 4*np.arange(9)
+    xticks = hr_ticks*60
+    print(xticks)
+    xticklabels = ["0"] + ["%dhr" % hr for hr in hr_ticks[1:]]
     axs[1].set_xticks(xticks)
     axs[1].set_xticklabels(xticklabels, fontsize=fontsize)
-    if xlims is not None:
-        _xlim = min(max_time, xlims[1])
-        axs[1].set_xlim([0, _xlim])
+    axs[1].set_xlim([0, _xlim])
     axs[1].set_ylim([0, 1.1*max_dist])
 
     for i in range(2):

### 2021-03-23 written/eLifeRevision/Bittner_eLife2020.tex
lrrnn method update

@@
 \end{equation}
-In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification.
+In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section \ref{methods_related}).
 
@@
 These two conditions are necessary and sufficient for stable amplification in RNNs \cite{bondanelli2020coding}.
-We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions and variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
-As a maximum entropy distribution, this is written as
+We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions.
+To complete the emergent property definition, we chose variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
+In terms of the EPI optimization variables, this is written as
 \begin{equation} 
@@
 The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch.
-
-ABC methods define a tolerance $\epsilon$ from observed data $x_0$ for which we keep sampled parameters.
-To make this ABC approach as similar as possible to the EPI program defined by Equation \ref{eq:EP_LRRNN}, we chose $\epsilon = 0.5$, an $l$-$2$ distance metric, and 
-\begin{equation}\label{eq:SNPE_stab_amp_x0}
-x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}
-\end{equation}
-located at the mean of our desired emergent property.
-We use sequential Monte Carlo ABC (SMC-ABC), to compare efficiency, since it is considered the most efficient ABC approach.
-SNPE \cite{gonccalves2019training} is another deep likelihood-free inference method that emerged along with this work.
-In contrast to EPI, SNPE cannot condition on the variance of the posterior predictive distribution. 
-Also, there is no tolerance parameter for SNPE like $\epsilon$ in ABC, so the comparative SNPE approach simply conditions on observation $x_0$.
-
-\textbf{How we did EPI} \\
-
-\textbf{How we did SNPE} \\
-SMC-ABC has many hyperparameters, of which pyABC selects automatically by running some initial diagnostics upon initialization.
-In concurrence with the literature, SMC-ABC fails to converge around 25-30 dimensions, since it's proposal samples never get close enough to the target statistics.
-
-\textbf{How we did SNPE} \\
-We searched over many SNPE hyperparameter choices: $n_{\text{train}} \in [2,000, 10,000, 100,000]$ is the number of simulations run per training epoch, and $n_{\text{mades}} \in [2, 3]$ is the number of masked autoregressive density estimators in the deep parameter distribution architecture.
-The greater $n_{\text{train}}$, the longer each epoch will take, but the more likely SNPE may converge during that epoch.
-Greater $n_{\text{mades}}$ increases the flexibility of the deep parameter distribution of SNPE, but slows optimization.
-For the timing plot, we show the fastest among all of these choices, and for the convergence plot, we show the best convergence among all of these choices.
-During optimization, we used $n_{\text{atom}}$=100 atomic proposals as is recommended.
+We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
+\ref{fig:LRRNN}B blue line, and Fig.
+\ref{fig:LRRNN}C-D blue).
+
+We compared EPI to two alternative likelihood-free inference (LFI) techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
+Appoximate Bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parametes $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
+Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) is the state-of-the-art ABC method, which leverages SMC techniques to improve sampling speed.
+We ran SMC-ABC with the pyABC package \cite{klinger2018pyabc} to infer RNNs with stable amplification: connectivities having eigenvalues within an $\epsilon$-defined $l$-$2$ distance of
+\begin{equation}\label{eq:stab_amp_x0}
+x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}.
+\end{equation}
+SMC-ABC was run with a uniform prior over $\mathbf{z} \in \left[-1, 1 \right]^{(4N)}$, a population size of 1,000 particles with simulations parallelized over 32 cores, and a multivariate normal transition model.
 
@@
 \begin{center}
+\includegraphics[scale=0.7]{figures/figRNN2/figRNN2.pdf}
+\end{center}
+\caption{\small (RNN1): 
+Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
+}
+\label{fig:RNN1}
+\end{figure}
+
+SNPE, the next LFI approach in our comparison, is far more similar to EPI.
+Like EPI, SNPE models parameters in mechanistic models with deep probability distributions, yet their learning algorithms are categorically different.
+SNPE uses a two-network architecture to approximate the posterior distribution of the model conditioned on observed data $\mathbf{x}_0$.
+The amortizing network maps observations $\mathbf{x}_i$ to the parameters of the deep probability distribution. 
+The weights and biases of the parameter network are optimized by sequentially augmenting the training data with additional pairs ($\mathbf{z}_i$, $\mathbf{x}_i$) based on the most recent posterior approximation.
+This sequential procedure is important to get training data $\mathbf{z}_i$ to be closer to the true posterior, and $\mathbf{x}_i$ to be closer to the observed data.
+For the deep probability distribution architecture, we chose a masked autoregressive flow with affine couplings (the default choice), three transforms, 50 hidden units, and a normalizing flow mapping to the support as in EPI.
+This architectural choice closely tracked the size of the architecture used by EPI (Fig. \ref{fig:RNN1}).
+As in SMC-ABC, we ran SNPE with $\mathbf{x}_0 = \mu$.
+All SNPE optimizations were run for a limit of 1.5 days on a Tesla V100 GPU, or until two consecutive rounds resulted in a validation log probability lower than the maximum observed.
+
+\begin{figure}
+\begin{center}
 \includegraphics[scale=0.8]{figures/figRNN1/figRNN1.pdf}
 \end{center}
-\caption{\small (RNN1): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
+\caption{\small (RNN2): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
 \textbf{A}.
@@
 }
-\label{fig:RNN1}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{figures/figRNN2/figRNN2.pdf}
-\end{center}
-\caption{\small (RNN2): 
-Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
-}
 \label{fig:RNN2}
 \end{figure}
+
+To clarify the difference in objectives of EPI and SNPE, we show their results on RNN models with different numbers of neurons $N$ and random strength $g$.  
+The parameters inferred by EPI consistently produces the same mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$, while those inferred by SNPE change according to the model definition (Fig.
+\ref{fig:RNN2}A).
+For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater uncertainty (Fig. \ref{fig:RNN2}B top, orange).
+At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Fig. \ref{fig:RNN2}B top).
+However at $N=10$, SNPE results in a predictive distribution of more widely dispersed eigenvalues (Fig.
+\ref{fig:RNN2}A bottom), and an inferred posterior with greater entropy than EPI (Fig.
+\ref{fig:RNN2}B bottom).
+We highlight these differences not to focus on an insightful trend, but to emphasize that these methods optimize different objectives with different implications.
+
+Note that SNPE converges when it's validation log probability has saturated after several rounds of optimization (Fig.
+\ref{fig:RNN2}C), and that EPI converges after several epochs of its own optimization to enforce the emergent property constraints (Fig.
+\ref{fig:RNN2}D blue).
+Importantly, as SNPE optimizes its posterior approximation, the predictive means change, and at convergence may be different than $\mathbf{x}_0$ (Fig. \ref{fig:RNN2}D orange, left).
+It is sensible to assume that predictions of a well-approximated SNPE posterior should closely reflect the data on average (especially given a uniform prior and a low degree of stochasticity), however this is not a given.
+Furthermore, no aspect of the SNPE optimization controls the variance of the predictions (Fig. \ref{fig:RNN2}D orange, right).
+
+To compare the efficiency of these algorithms, we start with the basic principle that you want z to produce x.
+In EPI this is blah with varince.
+In SNPE this is implicitly a goal of posterior inference.
+In SMC this is explicitly a goal, accuracy controlled by epsilon.
+Therefore we measure time to convergence and simulation counts as x
 
@@
 
+The hyperparameters of SNPE have a serous affect.
+Need to check whether more aggressive might lead to converge.
+Nsims
+Natoms
+Nsims at 100.
+
 \end{document}

### 2021-03-23 neural_circuits/LRRNN.py
latest conv nb

@@ -418,7 +418,7 @@ def get_SMC_conv(N, random_seeds):
     smc_sims = []
     for _rs in random_seeds:
         base_path = os.path.join("data", "smc")
-        save_dir = "SMC_RNN_stab_amp_N=%d_rs=%d" % (N, _rs)
+        save_dir = "SMC_RNN_stab_amp_N=%d_eps=%.2f_rs=%d" % (N,0.5, _rs)
         save_path = os.path.join(base_path, save_dir)
     
         try:

### 2021-03-23 written/eLifeRevision/Bittner_eLife2020.tex
rnn3 figs and methods text

@@
 
-To compare the efficiency of these algorithms, we start with the basic principle that you want z to produce x.
-In EPI this is blah with varince.
-In SNPE this is implicitly a goal of posterior inference.
-In SMC this is explicitly a goal, accuracy controlled by epsilon.
-Therefore we measure time to convergence and simulation counts as x
+To compare the efficiency of these algorithms for infering RNN connectivity distributions producing stable amplification, we develop a convergence criteria that can be used across methods.
+While EPI has its own hypothesis testing convergence criteria for the emergent property, it would not make sense to use this criteria on SNPE and SMC-ABC which do not constrain the means and variances of their predictions.
+Instead, we consider EPI and SNPE to have converged after completing its most recent optimization epoch (EPI) or round (SNPE) in which the distance
+\begin{equation}
+d(q_\theta(z)) = |\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}|_2
+\end{equation}
+is less than 0.25.
+We consider SMC-ABC to have converged once the population produces samples within the $\epsilon = 0.5$ ball ensuring stable amplification.
 
@@
 
-The hyperparameters of SNPE have a serous affect.
-Need to check whether more aggressive might lead to converge.
-Nsims
-Natoms
-Nsims at 100.
+When assessing the scalability of SNPE, it is important to check that alternative hyperparamterizations could not yield better performance.
+Key hyperparameters of the SNPE optimization are the number of simulations per round $n_{\text{round}}$, the number of atoms used in the atomic proposals of the SNPE-C algorithm \cite{greenberg2019automatic}, and the batch size $n$.
+To match EPI, we used a batch size of $n = 200$ for $N <= 25$, however we found $n=1,000$ to be helpful for SNPE in higher dimensions.
+While $n_{\text{round}} = 1,000$ yielded SNPE convergence for $N <= 25$, we found that a substantial increase to $n_{\text{round}} = 25,000$ yielded more consistent convergence at $N=50$ (Fig.
+\ref{fig:RNN3}A).
+By increasing $n_{\text{round}}$, we also necessarily increase the duration of each round.
+At $N=100$, we tried two hyperparameter modifications.
+As suggested in \cite{greenberg2019automatic}, we increased $n_{\text{atom}}$ by an order of magnitude to improve gradient quality, but this had little effect on the optimization (much overlap between same random seeds) (Fig.
+\ref{fig:RNN3}B).
+Finally we increased $n_{\text{round}}$ by an order of magnitude, which yielded convergence in one case, but no others.
+We found no way to improve the convergence rate of SNPE without making more aggressive hyperparameter choices requiring high numbers of simulations.
 

### 2021-03-23 written/eLifeRevision/scaling_RNN_inference.tex
rnn3 figs and methods text

--- 2021-03-23 written/eLifeRevision/Bittner_eLife2020.tex
@@
 
-\thispagestyle{plain}
-{\Large Interrogating theoretical models of neural computation with deep inference} \\
-Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3,4}$, Chunyu A. Duan$^{5}$, Carlos D. Brody$^{2,3,6}$, \\
-Kenneth D. Miller$^{1}$, and John P. Cunningham$^{7}$.
+\linenumbers
 
-{\small
-$^{1}$Department of Neuroscience, Columbia University, \\
-$^{2}$Princeton Neuroscience Institute, \\
-$^{3}$Princeton University, \\
-$^{4}$Allen Institute for Brain Science, \\
-$^{5}$Institute of Neuroscience, Chinese Academy of Sciences, \\
-$^{6}$Howard Hughes Medical Institute, \\
-$^{7}$Department of Statistics, Columbia University
-}
-
-\linenumbers
-\section{Abstract}
-%When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
-%We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
-A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
-Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
-The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
-A key process in neuroscientific modeling is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
-While statistical inference has proven effective on a broad variety of neuroscientific datasets, we clarify an important incongruity between theoretical approaches to neuroscience and this probabilistic methodology.
-Theoretical neuroscience is focused on computational properties and how they emerge from biological mechanisms, rather than noisy experimental datasets and their quantified structure.
-In this work, we present a novel technique to directly infer circuit model parameters producing these computational properties.
-This method tailors deep inference, the use of deep neural networks for statistical inference, to the nature of theoretical inverse problems, enabling scaling to high dimensions and considerable technical simplifications.
-With this method, we bring deep inference to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
-First, we emphasize the general applicability of this approach by inferring channel conductance parameters that produce a distinctive spiking frequency in a biophysical model of the stomatogastric ganglion.
-Then, in a model of primary visual cortex with multiple neuron-types, where  analysis becomes untenable as more neuron-types are included, we use this technique to discover how noise properties govern excitatory population variability.
-Next, in a model of superior colliculus, we identify and characterize two distinct regimes of connectivity that facilitate switching between opposite tasks throughout interleaved trials. 
-Finally, we scale inference to 1,000-dimensional parameter spaces of RNN connectivity that exhibit stable, yet amplified responses.
-These analyses illustrate how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
-
-%-------- old sentences -----------
-
-%Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
-%However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-%Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
-%We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
-%More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
-A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
-The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.
-
-Consider the idealized practice: one carefully designs a model and analytically derives how computational properties determine model parameters.
-Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
-Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-Alternatively, we can gain insight into these complex models by identifying the distribution of parameters that produce a specific computational property.
-By solving the inverse problem in this way, scientific analysis of complex biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-%By solving the inverse problem in this way, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-
-%The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-%However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
-%Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-% 
-%Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
-%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
-%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-%Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
-%These works have used deep neural networks to expand the domain of neural data sets amenable to statistical modeling \cite{paninski2018neural}.
-
-One preferred formalism for parameter identification is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-Recent work has used variational autoencoders (VAEs) \cite{kingma2013auto, rezende2014stochastic} to interrogate hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}, thus expanding the domain of neural data sets amenable to statistical modeling.
-However, most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood necessary for statistical inference.
-Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference (LFI) methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-A flourishing new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional LFI approaches.
-However, as we detail, all of these approaches require good datasets for the scientific question at hand.
-
-%
-%Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
-%First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
-%Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
-%Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
-%Sampling-based inference approaches (e.g.
-ABC and Markov chain Monte Carlo (MCMC) \cite{hastings1970monte, metropolis1953equation}) confer flexible approximations, yet scale poorly in number of parameters.
-%While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
-%Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
-%Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
-
-%To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
-%This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
-%First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
-%Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
-%Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
-%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-This work seeks to clarify an important incongruity between theoretical approaches to neuroscience and existing statistical inference methodology.
-In theoretical neuroscience, we are concerned with the computational properties -- the \emph{emergent phenomena} -- of our models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, not noisy observed datasets \cite{paninski2018neural}.
-To use the aforementioned inference paradigm, scientists must shoehorn such mathematical criteria into an artificial dataset compatible with existing statistical approaches.
-Theorists are therefore barred from using the probabilistic modeling toolkit for science, unless they reformulate their inverse problem to fit an evidence accumulation framework.
-
-These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
-As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets  exhibiting some mathematical criteria.
-EPI fixates on the inverse problem by constraining the predictions of the inferred parameter distribution to produce the emergent property exactly.
-Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-By using deep probability distributions EPI flexibly captures parametric structure in mechanistic models.
-This technique is designed to identify the full parameter space producing an emergent property and facilitate the seemless structural analysis of the inferred parameter distribution.
-Thus, EPI provides neuroscientists with an efficient, versatile probabilistic modeling toolkit designed explicitly for theoretical inverse problems.
-
-Equipped with this method, we bring deep inference to bear on theoretical neuroscience to an unprecedented extent.
-Throughout this work, we showcase the capabilities of EPI on four neural circuit models across ranges of biological realism, neural system function, and network scale.
-First, we show EPI's ability to capture subtle, nonlinear parametric structure in a stomatogastric ganglion subcircuit model \cite{gutierrez2013multiple}.
-In a model of primary visual cortex \cite{palmigiano2020structure}, we show how to gain insight from multiple inferred distributions.
-Next, we used EPI to idenitfy and structurally characterize multiple parametric regimes of superior colliculus activity in a model of task switching \cite{duan2018collicular}.
-Finally, we emphasize the superior scalability of EPI compared to other LFI techniques by inferring high-dimensional distributions of RNN connectivities that exhibit amplified, yet stable responses -- a hallmark of cortical sensory systems \cite{more, bondanelli2020coding}.
-
-Most importantly of all in this work, we present novel theories of neural computation borne from EPI analysis.
-We identified an unknown parametric rule of variability with respect to inhibitory neuron type in a V1 model, where analytic techniques became untenable.
-Furthermore, we identified multiple regimes of SC connectivity which confer rapid task switching, and used the structural analytic tools of EPI to gain a mechanistic understanding of circuit responses.
-These valuable theoretical insights illustrate the value of deep inference for the interrogation of neural circuit models.
-
-%-------- old sentences -----------
-
-%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
-
-%This creates an unfavorable tradeoff.
-On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
-%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-
-%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
-
-% now transition to deep ML...
-%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
-%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
-%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-
-%However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.
-%First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.
-%Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
-%Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
-%Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model.
-%Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
-
-%First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
-%Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
-%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-%Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
-%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-%First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-%Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-%Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Results}
-\subsection{Motivating emergent property inference of theoretical models} \label{results_motivating}
-
-%Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
-Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
-First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
-To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
-Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
-%A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
-A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
-Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  
-The hub neuron (hub) couples with either the fast or slow population or both.
-
-%Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
-Second, once the model is selected, one defines the emergent phenomena of scientific interest.
-%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-In the STG example, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model \ref{fig:STG}B.
-%One such emergent property is the phenomenon of \emph{unified intermediacy} -- in certain parameter regimes, the frequency of all neurons match at an intermediate firing rate.
-An interesting emergent property of this stochastic model is when the hub neuron fires at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
-%This emergent property is shown in Figure \ref{fig:STG}D at a frequency of 0.53Hz.
-This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
-
-%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
-%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
-%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
-% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-Third, parameter analyses ensue: brute-force parameter sweeps, ABC sampling \cite{liepe2014framework}, and sensitivity analyses \cite{raue2009structural} are all routinely used to reason about what parameter configurations lead to an emergent property.
-In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-
-%A.
-For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.
-%In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
-%Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-%Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.  
-%B. The EPI distribution of STG model parameters producing network syncing.  
-%Samples are colored by log probability density.  
-%Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
-%Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-%Simulated activity is shown for three samples (stars). 
-%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
-%$v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  
-%C.
-Deep probability distributions map a latent random variable $\mathbf{w}$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = f_{\bm{\theta}}(\mathbf{w})$ distributed as $q_{\bm{\theta}}(\mathbf{z})$.
-%D.
-EPI optimization: To learn the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ of model parameters that produce an emergent property, the emergent property statistics $T(\mathbf{x})$ are set in expectation over model parameter samples $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and model simulations $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ to emergent property values $\bm{\mu}$.
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}
-\end{center}
-\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
-\textbf{A}. Conductance-based biophysical model of the STG subcircuit.
-In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
-Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-Parameters are represented by the vector $\mathbf{z}$ and membrane potentials by the vector $\mathbf{x}$.
-The evolution of this model's activity $\mathbf{x}(t)$ is predicated by differential equations.
-\textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
-In this example, spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
-\textbf{C}. The emergent property of intermediate hub frequency, in which the hub neuron fires at a rate between the fast and slow frequencies.
-This emergent property is defined by a mean and variance on the emergent property statistic.
-Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution (Panel E).
-\textbf{D}. For a choice of model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$.
-In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced.
-The EPI posterior distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$. 
-\textbf{E}. The EPI posterior producing intermediate hub frequency.
-Samples are colored by log probability density.  
-Distribution contours of average hub neuron frequency from mean of .55 Hz are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
-\textbf{F} Simulations from parameters in E.  
-(Top) The predictive distribution of the posterior obeys the emergent property.
-The black and gray dashed lines show the mean and two standard deviations according the emergent property, respectively.
-(Bottom) Simulations at the starred parameter values.
- }
- \label{fig:STG}
-\end{figure}
-
-%Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
-Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
-\ref{fig:STG}E).
-This distribution -- represented for clarity as samples from the distribution -- is then the most random parameter distribution producing the emergent property. 
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) (Fig.
-\ref{fig:STG}E, arrows) agree with error contours (Fig.
-\ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
-\ref{fig:STG}F activity traces) (see Section \ref{methods_STG}).
-
-\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
-Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions.
-%First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
-First, we consider the model as a coupled set of differential equations \cite{gutierrez2013multiple}.  
-In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
-\begin{equation} 
-\begin{split}
-C_m \frac{d\mathbf{x}(t)}{dt} = -h(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
- \end{split}
-\end{equation} 
-where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG}).
-
-%Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
-%Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
--- is close to an intermediate frequency of 0.53Hz.
-%Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz.  
-%Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
-Mathematically, we achieve this with two constraints: by setting the mean hub frequency of model generated activity to 0.55Hz.
-\begin{equation}\label{eq:EP_STG1}
- \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix}
-\end{equation}
-and requiring that the variance around this mean is small
-\begin{equation}\label{eq:EP_STG2}
-\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix}.
-\end{equation}
-This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the compatible parameter space.
-In general, an emergent property
-\begin{equation}\label{eq:EP_STG}
-\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
-\end{equation}
-defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
-The choice of $\bm{\sigma}^2$ predicates the degree of variability in around mean $\bm{\mu}$ that is consistent with the emergent property.
-
-Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}D (see Section \ref{methods_EPI}).
-Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we choose a normative selection principle imposing no additional structure beyond the production of the emergent property \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}, which is the same normative principle of Bayesian inference (see Section \ref{methods_VI}).
-
-%Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
-%This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-%This deep distribution is represented in Figure \ref{fig:STG}E (see Section \ref{methods_EPI}).  
-%Then, mathematically, we must solve the following optimization program: 
-%\begin{equation} \label{eq:EPI}
-%\begin{split}
-%q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) ~~=~~ &\argmax_{q_{\bm{\theta}} \in \mathcal{Q}} H(q_{\bm{\theta}}(\mathbf{z})) \\
-%&\text{  s.t.
-} ~\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2 \\
-%\end{split}
-%\end{equation}
-%where $f(\mathbf{x}, \mathbf{z}), \bm{\mu}$, and $\bm{\sigma}$ are defined as in Equation \ref{eq:EP}.
-%According to the emergent property of interest, $f(\mathbf{x}, \mathbf{z})$ may contain multiple statistics, in which case the mean and variance vectors $\bm{\mu}$ and $\bm{\sigma}^2$ match this dimension.
-%Finally, we recognize that many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select that which has maximum entropy.
-%This principle, captured in Equation \ref{eq:EPI} by the primal objective $H$, identifies parameter distributions with minimal assumptions beyond some chosen structure \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}.
-%Such a normative principle of maximum entropy, which is also that of Bayesian inference, naturally fits with our scientific objective of reasoning about parametric sensitivity and robustness. 
-%The recovered distribution of EPI is as variable as possible along each parametric manifold such that it produces the emergent property.
-
-%EPI optimizes the weights and biases $\bm{\theta}$ of the deep network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-%The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
-%%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
-%Such convergence is evaluated with a hypothesis test that the means and variances of each emergent property statistic are not different than their constrained values (see Section \ref{methods_AL_opt}).
-%Further validation of EPI is available in the supplementary materials, where we analyze a simpler model for which ground-truth statements can be made (Section \ref{methods_2DLDS}).
-
-%In relation to broader methodology, inspection of the EPI objective reveals a natural relationship to posterior inference.
-%Specifically, EPI executes a novel variant of  Bayesian inference with a uniform prior and a gaussian likelihood on the emergent property statistic (see Section \ref{methods_ME_EF}).
-%A key advantage of EPI over established Bayesian inference is that the predictions made by the inferred distribution are constrained to produce the specified emergent property.
-
-The probabilities of the distribution inferred from EPI are the densities of these parameters in the most random distribution producing the emergent property.
-While existing approaches to probabilistic structural identifiability analysis use the lens of evidence accumulation \cite{raue2009structural, gonccalves2019training}, sensitivity and robustness of parameter space dimensions with respect to emergent phenomena can be directly quantified with EPI.
-Sensitivity quantifications are measured by the second order derivative of EPI probability along the parameteric dimension of interest: to what extent is the emergent property maintained or diminished along this dimension?
-Once an EPI distribution has been inferred, this second order derivative requires trivial computation (as long as the correct architecture class is chosen, see Section \ref{methods_NF}).
-Equipped with this method, we may examine structure in the resulting parameter distributions or make comparisons between distributions conditioned at different levels of the same emergent property statistic.
-In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using these techniques to investigate and produce novel scientific insight.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI reveals how neuron-type specific noise governs variability in the stochastic stabilized supralinear network} \label{results_V1}
-
-\begin{figure}
-\vspace{-1cm}
-\begin{center}
-\includegraphics[scale=0.8]{figures/fig2/fig2.pdf}
-\end{center}
-\vspace{-1cm}
-\caption{\small Emergent property inference in the stochastic stabilized supralinear network (SSSN)
-\textbf{A}.
-Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).
-Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
-Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
-Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
-\textbf{B}. Steady-state responses of the SSN model (deterministic, $\bm{\sigma}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
-\textbf{C}.
-Transient network responses of the SSSN model at 50 \% contrast.
-(Left) Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise realization.
-(Right) Mean (solid line) and standard deviation (shading) of responses.
-\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on E-population variability.
-The posterior predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
-\textbf{E}.
-(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posteriors $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
-Each black dot shows the mode at each $\sigma_P$.
-The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
-\textbf{F}.
-The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each posterior $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
-}
- \label{fig:V1}
-\end{figure}
-
-Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
-Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
-Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we use EPI to analyze V1 models of greater complexity in order to characterize properties of slow noise governing circuit variability.
-
-We considered the response properties of a nonlinear dynamical V1 circuit model (Fig.
-\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via supralinear nonlinearity $\phi = \left[\right]^2_+$.
-The input is also comprised of a slow noise component $\bm{\epsilon} \sim OU\left(\tau_{\text{noise}}, \bm{\sigma} \right)$ of time scale $\tau_{\text{noise}} > \tau$ and variance parameters $\bm{\sigma}$ (see Section \ref{methods_V1})
-\begin{equation}
-    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}).
-\end{equation}
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have inhibitory multiplicity, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
-Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
-Here, we consider fixed weights $W$ and input $\mathbf{h}$ according to a fit of the deterministic model to contrast responses \cite{palmigiano2020structure} (Fig.
-\ref{fig:V1}B), and study the effect of noise parameterization $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on fluctuations at 50\% contrast.
-
-For this SSSN, we are interested in how noise variability across neural populations governs stochastic fluctuations in the E-population.
-Here, we quantify different levels $y$ of E-population variability with the emergent property
-\begin{equation}\label{eq:EP_V1}
-\begin{split}
-\mathcal{X}(y) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ y \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2,
-\end{split}
-\end{equation}
-where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-
-We ran EPI to obtain a posterior distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
-From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
-\ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
-Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions of the posterior (Fig.
-\ref{fig:V1}D, bottom-right).
-Together, these observations imply a parametric manifold of degeneracy with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ in the $\sigma_E$-$\sigma_P$ marginal (Fig.
-\ref{fig:V1}E).
-The dimensions of sensitivity conferred by EPI and this plain visual structure suggest a quadratic relationship in the emergent property statistic $s_E(\mathbf{x}; \mathbf{z})$ and parameters $\mathbf{z}$, which is preserved at a greater level of variability $\mathcal{X}(10\text{ Hz})$ (Fig.
-\ref{fig:V1}E).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ (Fig \ref{fig:V1}F, $p=0$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two posteriors (Fig.
-\ref{fig:V1_3}, $p=.402$).
-
-While a quadratic relationship in $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ is potentially derivable by extending the derivation in Section \ref{methods_V1} to the case of $\tau \neq \tau_{\text{noise}}$, the coefficients in front of each quadratic term would be unruly, and likely escape comprehensible analysis.
-This makes EPI an attractive tool for revealing the characteristics of noise governing variability and for answering other questions in this complex model.
-Intriguingly, this circuit exhibited a paradoxical effect in the P-population, and no other inhibitory types at 50\% contrast (Fig. \ref{fig:V1_3}) implying that the E-population is P-stabilized.
-Future work motivated by our analysis here, may uncover a relationship between the neuron-type mediating stability and the factors governing circuit variability.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
-In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}A).
-Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
-Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
-We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig.
-\ref{fig:SC}B).
-The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
-The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.
-The dynamics evolve with timescale $\tau= 90\text{ms}$ governed by connectivity weights $W$
-\begin{equation}
-\begin{split}
-\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
-\mathbf{x} &= \phi(\mathbf{u})
-\end{split}
-\end{equation}
-with white noise of variance $0.2^2$.
-The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching behavior.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
-\end{center}
-\caption{\footnotesize 
-\textbf{A}. Rapid task switching behavioral paradigm (see text). 
-\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
-Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-\textbf{C}. The EPI posterior distribution of rapid task switching networks.  
-Red and purple stars ($\mathbf{z}_1$ and $\mathbf{z}_2$) indicate different connectivity regimes with different sensitivity vectors $\mathbf{v}_1$ and $\mathbf{v}_2$.
-(Middle-left) Posterior predictive distribution of task accuracies.
-(Bottom-left) Task accuracy along dimensions of sensitivity in each connectivity regime.
-\textbf{D}. Means (solid) and standard deviations (shaded) of each population across random simulated trials.
-Top plots show Pro (top) and Anti (bottom) responses for connectivity $\mathbf{z}_1$.
-Bottom rows show the same $\mathbf{z}_2$.
-\textbf{E}. The EPI posterior predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
-\textbf{F}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
-\textbf{G}. Accuracy with delay period inactivation along each connectivity regime's dimension of sensitivity.
-}
-\label{fig:SC}
-\end{figure}
-
-We define rapid task switching behavior as accurate execution of each task.
-Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
-We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
-\begin{equation}\label{eq:SC_EP}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 7.5\%^2 \\ 7.5\%^2  \end{bmatrix}
-\end{split}
-\end{equation}
-A variance of $7.5\%^2$  in each task will confer a posterior producing performances ranging from about $60\%-90\%$, allowing us to examine the properties of connectivity that yield better performance in each task.
-Notably, this is our first example using EPI to condition on multiple emergent property statistics ($|f(\mathbf{x}; \mathbf{z})| = 2$).
-
-The EPI inferred parameters  (Fig.
-\ref{fig:SC}C) generate a distribution of task accuracies (Fig.
-\ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
-The nonlinear patterns of connectivity that govern each task accuracy (Fig. \ref{fig:SC1}A-B) are not fully captured by linear prediction (Fig. \ref{fig:SC1}C).
-For example, the patterns in connectivity increasing Pro accuracy change dramatically after crossing a threshold of $sW$ (Fig. \ref{fig:SC1}A $sW$-$hW$ marginal).
-Not only has EPI captured this complex nonlinear posterior, it offers probabilistic tools for understanding the different regimes of model behavior.
-
-To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
-\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
-Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
-We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
-The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
-\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
-\ref{fig:SC}D, bottom-left, \ref{fig:SC2}).
-In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy tracks monotonic increases in Pro accuracy.
-These responses make intuitive sense, recognizing that $\mathbf{v}_1$ (unlike $\mathbf{v}_2$) points strongly in the direction of connectivity eigenvalue $\lambda_{\text{diag}}$, which is strongly anticorrelated with $p_A$ (Fig.
-\ref{fig:SC3}, \ref{fig:SC4}, see Section \ref{methods_SC}).
-
-In agreement with experimental results from Duan et al., we found optogenetic inactivation during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
-\ref{fig:SC}E)).
-This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
-Similarities across Pro and Anti trials in choice period responses following delay period inactivation (Fig.
-\ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
-\ref{fig:SC}F).
-The strong anticorrelation between $p_P$ and $p_{A, opto}$ across posterior connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
-Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
-\ref{fig:SC}F).
-
-In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This posterior revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
-EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, the model is well-chosen in that regard.
-Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
-
-\subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
+\subsection{Scaling inference in RNNs with EPI} \label{results_LRRNN}
 Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinisically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
@@
 
-%\section{Discussion}
-%\subsection{EPI is a general tool for theoretical neuroscience} 
-%Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
-%We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
-%We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).
-%We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
-%Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.
-%
-%In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
-%These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
-%In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
-%Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
-%Although, theoretical approaches (e.g.
-\cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
-
-%\subsection{Novel hypotheses from EPI} 
-\section{Discussion}
-
-In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{paninski2018neural}.
-Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
-However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
-In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
-The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards careful inference in careful models of neural computation (see Section \ref{methods_related}).
-
-Specifically, approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential} has been the standard approach to parameter inference in neural circuit models lacking tractable likelihoods.
-ABC methods do not confer probabilities on accepted parameters, require an acceptance threshold chosen to trade-off inference quality with tractability, do not scale efficiently to high-dimensional parameter spaces, and require independent techniques to analyze sensitivity for local parameter choices \cite{sisson2018handbook}.
-In contrast, EPI allows probability evaluations at any point in parameter space, conditions posteriors on the natural quantification of emergent properties, scales to high dimensional parameter spaces, and naturally admits sensitivity quantification via fast evaluations of the posterior Hessian.
-
-Technically, EPI is a maximum entropy method, which learns parameter distributions that are as random as possible given that they produce the emergent property.
-Conceptually, maximally random distributions given some constraints are useful for understanding parametric sensitivity.
-This is well understood in Bayesian inference, where maximum entropy is the chosen normative principle.
-This is emphasized by an innovative formalism unifying top-down maximum entropy normative models with bottom-up statistical models \cite{mlynarski2020statistical}.
-Indeed, EPI is an adaptive variational inference program, and may be considered to have a Bayesian uniform prior (see Section \ref{methods_VI}).
-
-%Biologically realistic models of neural circuits often prove formidable to analyze.
-%Two main factors contribute to the difficulty of this endeavor.
-%First, in most neural circuit models, the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space. 
-%Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
-%For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-
-Biologically realistic models of neural circuits often prove formidable to analyze for two main reasons.
-A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
-We see in Section \ref{results_LRRNN} that EPI scales seemlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
-EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
-
-A secondary challenge is that the structure of the parametric regimes governing emergent properties is intricate.  
-For example, even in low dimensional circuits, models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-With EPI, we use deep probabillity distributions to capture the complex nonlinear parameter distributions governing model behavior.
-In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing curcuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
-In Section \ref{results_SC}, we identified two regimes of SC connectivity resulting in rapid task switching, and found that the full distribution of rapid task switching networks reproduced an experimental result.
-
-%In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}.
-%By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out.
-%Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
-%For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
-%This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.
-
-%In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
-%Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
-%This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
-%While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
-%The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.
-
-EPI leverages deep learning technology for neuroscientific inquiry in a categorically different way than approaches focused on training neural networks to execute behavioral tasks \cite{richards2019deep}.
-These works focus on examining optimized deep neural networks while considering the objective function, learning rule, and architecture used.
-This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
-All of these approaches can be used complementarily to enhance the practice of theoretical neuroscience.
-
-\textbf{Acknowledgements}: \\
-This work was funded by NSF Graduate Research Fellowship,  DGE-1644869, McKnight Endowment Fund, NIH NINDS 5R01NS100066, Simons Foundation 542963, NSF NeuroNex Award, DBI-1707398, The Gatsby Charitable Foundation, Simons Collaboration on the Global Brain Postdoctoral Fellowship, Chinese Postdoctoral Science Foundation, and International Exchange Program Fellowship.
-Helpful conversations were had with Francesca Mastrogiuseppe, Srdjan Ostojic, James Fitzgerald, Stephen Baccus, Dhruva Raman, Liam Paninski, and Larry Abbott.
-
-\textbf{Data availability statement}: \\
-The datasets generated during and/or analyzed during the current study are available from the corresponding author upon reasonable request.
-
-\textbf{Code availability statement}: \\
-All software written for the current study is available at https://github.com/cunningham-lab/epi.
-
 \bibliography{eLife2020}
@@
 
-\newpage
-
-\section{Methods}
-
-\subsection{Emergent property inference (EPI)}\label{methods_EPI}
-%Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
-%EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{X}$ defined by the means and variances of emergent property statistics $f(\mathbf{x}; \mathbf{z})$
-% \begin{equation}
-%\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-%\end{equation}
-%Precisely, the emergent property statistics $f(\mathbf{x})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and stochasticity of the data given the parameters defined by the model $p(\mathbf{x} \mid \mathbf{z})$.
-%This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
-%
-%
-%With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
-% In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
-$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-%\begin{equation}
-%\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
-%\end{equation}
-%Given a simulator defined by a theoretical model $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ and some emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
-%\begin{equation} \label{eq:opt}
-%\begin{split}
-%q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
-% \end{split}
-%\end{equation} 
-%Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-%Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0$:
-%\begin{equation}
-%H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-%\end{equation}
-%Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $\mathbf{z}$:
-%\begin{equation}
-%\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-%\end{equation}
-%In EPI, MEFNs are purposed towards variational learning of model parameter distributions.
-
-Determining the combinations of model parameters that can produce observed data or a desired output is a key part of scientific practice.
-Solving inverse problems is especially important in neuroscience, since we require complex models to describe the complex phenomena of neural computations.
-While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models conditioned upon the emergent phenomena they produce.
-Here, we introduce a novel method for statistical inference, which finds distributions of parameter solutions that only produce the desired emergent property.
-This method seamlessly handles neural circuit models with stochastic nonlinear dynamical generative processes, which are predominant in theoretical neuroscience.
-
-Consider model parameterization $\mathbf{z}$, which is a collection of scientifically interesting variables that govern the complex simulation of data $\mathbf{x}$.
-For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
-In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
-Even so, we do not scientifically reason about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
-In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
-With EPI, we learn distributions of $\mathbf{z}$ that results in an average and variance of $f(\mathbf{x}; \mathbf{z})$, denoted $\bm{\mu}$ and $\bm{\sigma}^2$.
-We refer to the collection of these statistical moments as an emergent property.
-Such emergent properties $\mathcal{X}$ are defined through choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), $\bm{\mu}$, and $\bm{\sigma}^2$
- \begin{equation}
-\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}
-Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.
-
-In EPI, deep probability distributions are used as posterior approximations $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
- In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
-$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-\begin{equation} \label{eq:deep_transform}
-\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
-\end{equation}
-Such deep probability distributions embed the posterior distribution in a deep network.
-Once optimized, this deep network representation has remarkably useful properties: immediate posterior sampling, and immediate probability, gradient, and Hessian evaluation at any parameter choice.
-
-Given a choice of model $p(\mathbf{x} \mid \mathbf{z})$ and emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
-\begin{equation} \label{eq:opt}
-\begin{split}
-q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) = q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
- &  \text{s.t.
-} \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\\
- \end{split}
-\end{equation} 
-Entropy is chosen as the normative selection principle, since we want the posterior to only contain structure predicated by the emergent property \cite{jaynes1957information, elsayed2017structure}.
-This choice of selection principle is also that of standard Bayesian inference, and we derive an exact relation between EPI and variational inference (see Section \ref{methods_ME_EF}).
-However, a key difference is that variational inference and other Bayesian methods do not constrain the predictions of their inferred posteriors.
-This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
-
-In the remainder of Section \ref{methods_EPI}, we will explain the finer details and motivation of the EPI method. 
-First, we explain related approaches and what EPI introduces to this domain (Section \ref{methods_related}).
-Second, we describe the special class of deep probability distributions used in EPI called normalizing flows (Section \ref{methods_NF}).  
-Next, we explain the constrained optimization technique used to solve Equation \ref{eq:opt} (Section \ref{methods_AL_opt}).
-Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
-Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain the relation between EPI and variational inference (Section \ref{methods_VI}).
-
- \subsubsection{Related approaches}\label{methods_related}
- When Bayesian inference problems lack conjugacy, scientists use approximate inference methods like variational inference (VI) \cite{saul1998mean} and Markov chain Monte Carlo (MCMC) \cite{metropolis1953equation, hastings1970monte}.
-After optimization, variational methods return a parameterized posterior distribution, which we can analyze.
-Also, the variational approximating distribution class is often chosen such that it permits fast sampling.
-In contrast MCMC methods only produce samples from the approximated posterior distribution.
-No parameterized distribution is estimated, and additional samples are always generated with the same sampling complexity.
-Inference in models defined by systems of differential has been demonstrated with MCMC \cite{girolami2011riemann}, although this approach requires tractable likelihoods.
-Advancements have leveraged structure in stochastic differential equation models to improve likelihood approximations, thus expanding the domain of applicable models \cite{golightly2011bayesian}.
-
-Likelihood-free (or ``simulation-based") inference (LFI) \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
-The most prevalent approach to LFI is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
-The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
-Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
-SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:LRRNN}).
-Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again with a new population of initialized particles to obtain additional samples.
-
-For scientific model analysis, we seek a posterior distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
-For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
-EPI executes likelihood-free inference using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational posterior approximation.
-The deep neural network of EPI defines the parametric form of the posterior approximation.
-Furthermore, the EPI distribution is constrained to produce an emergent property.
-In other words, the summary statistics of the posterior predictive distribution are fixed to have certain first and second moments.
-EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
-The analytic relationship between EPI and variational inference is explained in Secton \ref{methods_VI}.
-
- We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}).%, preprint posted simultaneously with this preprint).
-We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
-
-The method EPI differs from SNPE in some key ways.
-SNPE belongs to a ``sequential" class of recently developed LFI methods in which two neural networks are used for posterior inference.
-This first neural network is a normalizing flow used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
-A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE \cite{hermans2020likelihood}).
-In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
-This contrasts with EPI and SNPE, which afford a normalizing flow approximation to the posterior, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
-The second neural network in this sequential class of methods is the amortizer.
-This network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
-These methods are optimized on a conditional density (or ratio) estimation objective on a sequentially adapting finite sample-based approximation to the posterior.
-
-The approximating fidelity of the first neural network in sequential approaches is optimized to generalize across the entire distribution it is conditioned upon.
-This optimization towards generalization of sequential methods can reduce the accuracy at the singular posterior of interest.
-Whereas in EPI, the entire expressivity of the normalizing flow is dedicated to learning a single distribution as well as possible.
-While amortization is not possible in EPI parameterized by the mean parameter $\mu$ (due to the inverse mapping problem \cite{wainwright2008graphical}), we have shown this two-network amortization approach to be effective in exponential family distributions defined by their natural parameterization \cite{bittner2019approximating}.
-
-Structural identifiability analysis involves the measurement of sensitivity and unidentifiabilities in natural models.
-Around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
-A popular efficient approach for systems of ODEs has been neural ODE adjoint \cite{chen2018neural} and its stochastic adaptation \cite{li2020scalable}.
-Casting identifiability as a statistical estimation problem, the profile likelihood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
-An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point by following degenerate dimensions of the fisher information matrix \cite{raman2017delineating}.
-Global structural non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
-With EPI, we have all the benefits given by a statistical inference method plus the ability to query the gradient or Hessian of the inferred distribution at any chosen parameter value.
-
-%\cite{dhruva/goldman papers}
-
-%A closely related methodology, variational inference, uses optimization to approximate posterior distributions \cite{blei2017variational}.
-%Standard methods like stochastic gradient variational Bayes \cite{kingma2013auto} or black box variational inference \cite{ranganath2014black} simply do not work for inference in theoretical models of neural circuits, since they require tractable likelihoods $p(\mathbf{x} \mid \mathbf{z})$.
-%Work on likelihood-free variational inference (LFVI) \cite{tran2017hierarchical}, which like EPI seeks to do inference in models with intractable likelihoods, employs an additional deep neural network as a ratio estimator, enabling an estimation of the optimization objective for variational inference.
-%Like LFVI, EPI can be framed as variational inference (see Section \ref{methods_VI}).
-%But, unlike LFVI, EPI uses a single deep network to learn a distribution and is optimized to produce an emergent property, rather than condition on data points.  
-%Optimizing the EPI objective is a technological challenge, the details of which we elaborate in Section \ref{methods_AL_opt}.  
-%Before going through those details, we ground this optimization in a toy example.
-% We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}, preprint posted simultaneously with this preprint).
-%We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
-%These distinct methodologies and scientific investigations emphasize the increased importance and timeliness of both works.
-
-\subsubsection{Normalizing flows}\label{methods_NF}
-%Deep probability models typically consist of several layers of fully connected neural networks.  
-Deep probability distributions are comprised of multiple layers of fully connected neural networks (Equation \label{eq:deep_transform}).
-%When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
-For $\mathbf{z}' = g(\mathbf{z})$,
-When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
-For $\mathbf{z}_i = g_i(\mathbf{z_{i-1}})$,
-%\begin{equation}
-%q(\mathbf{z}') = q(f^{-1}(\mathbf{z}')) \left| \det \frac{\partial f^{-1}(\mathbf{z}')}{\partial \mathbf{z}'} \right| = q(\mathbf{z}) \left| \det \frac{\partial f(\mathbf{z})}{\partial \mathbf{z}} \right|^{-1}.
-%\end{equation}
-\begin{equation}
-p(\mathbf{z}_i) = p(g_i^{-1}(\mathbf{z}_i)) \left| \det \frac{\partial g_i^{-1}(\mathbf{z}_i)}{\partial \mathbf{z}_i} \right| = p(\mathbf{z}_{i-1}) \left| \det \frac{\partial g_i(\mathbf{z}_{i-1})}{\partial \mathbf{z}_{i-1}} \right|^{-1}.
-\end{equation}
-
-However, this computation has cubic complexity in dimensionality for fully connected layers.  
-By restricting our layers to normalizing flows \cite{rezende2015variational, papamakarios2019normalizing} -- bijective functions with fast log determinant Jacobian computations, which confer a fast calculation of the sample log probability.
-Fast log probability calculation confers efficient optimization of the maximum entropy objective (see Section \ref{methods_AL_opt}).
-%Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
-%Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
-We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
-Fast probability evaluation in turn facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
-Glow permutations were used in between coupling stages \cite{kingma2018glow}.
-This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
-In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
-We specify the architecture used in each application by the number of Real-NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.
-
-\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented Lagrangian method.  
-The following objective is minimized:
-\begin{equation} \label{eq:AL}
-L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
-\end{equation}
-where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
-The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
-Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
-Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
-
-To take gradients with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, it can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
-\begin{equation}
-H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
-\begin{equation}
-\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-\end{equation}
-
-The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
-A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
-The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
-The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
-Throughout the study, $\gamma = 0.25$, while $\beta$ was chosen to be either 2 or 4.  The batch size of EPI also varied according to application.
-
-The intention is that $c$ and $\bm{\eta}_{\text{opt}}$ start at values encouraging entropic growth early in optimization.  
-With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see example in Section \ref{methods_2DLDS}).
-The momentum parameters of the Adam optimizer are reset at the end of each augmented Lagrangian epoch.
-
-Rather than starting optimization from some $\bm{\theta}$ drawn from a randomized distribution, we found that initializing $q_{\bm{\theta}}(\mathbf{z})$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
-The parameters of the Gaussian initialization were chosen on an application-specific basis.  
-Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some standard deviation $\bm{\sigma}_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
-
-To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we assess whether each individual constraint on the means and variances of $f(\mathbf{x}; \mathbf{z})$ is satisfied.
-We consider the EPI to have converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.
-This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
-The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented Lagrangian epoch.
-
-When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
-First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
-Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
-If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
-Next, one should consider the computational cost of the gradient calculations. 
-In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_RNN}) for the emergent property statistic given the model parameters.  
-On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
-Section \ref{methods_STG}).
-In such cases, backpropagating gradients through the SDE evolution will be expensive.
-
-\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
-\begin{equation} 
-\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
-\end{equation}
-with
-\begin{equation}
-A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
-\end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
- $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
-\begin{equation}
- \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
- \end{equation} 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS1/figLDS1.pdf}
-\end{center}
-\begin{flushleft}
-\caption{\small (LDS1): \textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
-\textbf{B}.
-The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.
-Dashed lines indicate the parameter axes.
-\textbf{C}. Entropy throughout the optimization.  
-At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
-\textbf{D}. Emergent property moments throughout optimization.  
-At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
-\end{flushleft}
-\label{fig:LDS1}
-\end{figure}
-
-Unlike the models we presented in the main text, this model admits an analytical form for the mean emergent property statistics given parameter $\mathbf{z}$, since the eigenvalues can be calculated using the quadratic formula:
-\begin{equation}
-\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
-\end{equation}
-
-Importantly, even though $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}$ directly.
-This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we used EPI to approximate this distribution (Fig.
-S1B).
-We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
-(see Section \ref{methods_NF}).
-
-Even this relatively simple system has nontrivial (though intuitively sensible) structure in the parameter distribution.  
-To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
-In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
-the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). 
-The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
-Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS2/figLDS2.pdf}
-\end{center}
-\begin{flushleft}
-\caption{\small (LDS2): \textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-\textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\end{flushleft}
-\label{fig:LDS2}
-\end{figure}
-
-To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
-\begin{equation}
-\text{imag}(\lambda_1) = \begin{cases}
-                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
-                             0 & \text{otherwise } \\
-                         \end{cases}.
-\end{equation}
-
-When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
-\begin{equation}
-\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
-\end{equation}
-Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
-S2B).
-With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
-This validates the curved structure of the inferred distribution learned through EPI.  
-We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
-Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS3/figLDS3.pdf}
-\end{center}
-\begin{flushleft}
-\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
-\textbf{A}. Each dimension of the simulated trajectories throughout time.  
-\textbf{B}.  The simulated trajectories in phase space.}
-\end{flushleft}
-\label{fig:LDS3}
-\end{figure}
-
-\subsubsection{Maximum entropy distributions and exponential families}\label{methods_ME_EF}
-%As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
-%A maximum entropy distribution of form:
-%\begin{equation} \label{eq:max_ent}
-%\begin{split}
-%p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}. \\
-% \end{split}
-%\end{equation} 
-%will have probability density in the exponential family:
-%\begin{equation}
-%p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-%\end{equation}
-%The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-Maximum entropy distributions have a fundamental link to exponential family distributions. 
-A maximum entropy distribution of form:
-\begin{equation} \label{eq:max_ent}
-\begin{split}
-p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
- &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
- \end{split}
-\end{equation} 
-will have probability density in the exponential family:
-\begin{equation}
-p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-\end{equation}
-The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}
- \begin{equation}
-\mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}
-The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
- \begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
-\end{equation}
-As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
-\begin{equation} \label{eq:moments}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
-\end{equation}
-which are constrained to the chosen means and variances
-\begin{equation} \label{eq:mu_opt}
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
-\end{equation}
-
-\subsubsection{EPI as variational inference}\label{methods_VI}
-%Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
-In Bayesian inference, we obtain a posterior distribution $p(z \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
-\begin{equation}
-p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
-\end{equation}
-The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
-
-%Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid \mathbf{x})$.  
-%We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(\mathbf{x})$.
-%The log normalizing constant $A(\eta(\mathbf{x}))$ remains unchanged:
-%\begin{equation}
-%\begin{split}
-%p(z \mid \mathbf{x}) = b(z) \exp{\left( \eta(\mathbf{x})^\top T(z) - A(\eta(\mathbf{x})) \right)} = \exp{\left( \begin{bmatrix} \eta(\mathbf{x}) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(\mathbf{x})) \right)} \\= \exp{\left(\tilde{\eta(\mathbf{x})}^\top \tilde{T}(z) - A(\eta(\mathbf{x})) \right)}
-%\end{split}.
-%\end{equation}
-%Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})).
-%\end{equation}
-%$q_{\bm{\theta}}(z)$ is the variational approximation to the posterior with variational parameters $\bm{\theta}$.  
-%We can write this KL divergence in terms of entropy of the variational approximation:
-%\begin{equation}
-%KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) = \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(z)) \right] - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (p(z \mid \mathbf{x})) \right]
-%\end{equation}
-%\begin{equation}
-% = -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) - A(\eta(\mathbf{x})) \right].
-%\end{equation}
-%As far as the variational optimization is concerned, the log normalizing constant is independent of $q_{\bm{\theta}}(z)$, so it can be dropped
-%\begin{equation}
-%\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) =  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) \right].
-%\end{equation}
-% Further, we can write the objective in terms of the first moment of the sufficient statistics $\bm{\mu} = \mathbb{E}_{z \sim p(z \mid \mathbf{x})}\left[T(z) \right]$:
-% \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right] + \tilde{\eta}(\mathbf{x})^\top \bm{\mu},
-% \end{equation}
-%which simplifies to
-%  \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right].
-% \end{equation}.
-%
-%In comparison, in emergent property inference (EPI), we solve the following problem:
-%\begin{equation}
-%q_{\bm{\theta}}^*(z) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(z)),   \text{  s.t.
-} \mathbb{E}_{z \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid z)}\left[T(\mathbf{x})\right] \right] = \bm{\mu}.
-%\end{equation}
-%The Lagrangian objective (without augmentation) is
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_{\bm{\theta}}} \left[\tilde{T}(z) \right] - \bm{\mu} \right).
-%\end{equation}
-%Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(\mathbf{x})$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
-%
-%We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
-%Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
-%However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-%Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
-
-In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$
-\begin{equation}
-q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
-\end{equation}
-The KL divergence can be written in terms of entropy of the variational approximation:
-\begin{equation}
-KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
-\end{equation}
-\begin{equation}
-= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
-\end{equation}
-Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
-This is usually framed as maximizing the evidence lower bound (ELBO)
-\begin{equation}
-\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
-\end{equation}
-
-Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
-\begin{equation}
-p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
-\end{equation}
-where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
-The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
-\begin{equation}
-\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
-\end{equation}
-where
-\begin{equation}
-\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
-\end{equation}
-\begin{equation}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
-\end{equation}
-The variational objective is then
-\begin{equation}
-\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
-\end{equation}
-
-Comparing this to the Lagrangian objective (without augmentation) of EPI, we see they are the same
-\begin{equation}
-\begin{split}
-q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
- &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
-\end{split}
-\end{equation}
-where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
-Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
-The data $\mathbf{x}$ used by this implicit variational inference program would be that generated by the adapting variational approximation $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z}) q_{\bm{\theta}}(\mathbf{z})$, and the likelihood parameters $\bm{\eta}_{f}$ of EPI optimization epoch $k$ are predicated by $\bm{\eta}_{\text{opt},k}$ .
-However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
-
-\subsection{Theoretical models}\label{methods_theoretical_models}
-In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  
-Here, we provide the details of each model  and the related analyses.
-
-\subsubsection{Stomatogastric ganglion}\label{methods_STG}
-%We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
-%Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
-%\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(\mathbf{x}; \mathbf{z}) + h_{Ca}(\mathbf{x}; \mathbf{z}) + h_K(\mathbf{x}; \mathbf{z}) + h_{hyp}(\mathbf{x}; \mathbf{z}) + h_{elec}(\mathbf{x}; \mathbf{z}) + h_{syn}(\mathbf{x}; \mathbf{z})\right].
-%\end{equation} 
-%The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $\mathbf{z}$.
-%The capacitance of the cell membrane was set to $C_m = 1nF$.
-Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
-%\begin{equation}  h_{leak}(\mathbf{x}; \mathbf{z}) = g_{leak} (x_\alpha - V_{leak}) 
-%\end{equation} 
-%\begin{equation}  h_{elec}(\mathbf{x}; \mathbf{z}) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
-%\end{equation} 
-%\begin{equation}  h_{syn}(\mathbf{x}; \mathbf{z}) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
-%\begin{equation}  h_{Ca}(\mathbf{x}; \mathbf{z}) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
-%\end{equation} 
-%\begin{equation}  h_K(\mathbf{x}; \mathbf{z}) = g_K N (x_\alpha - V_K) 
-%\end{equation} 
-%\begin{equation}  h_{hyp}(\mathbf{x}; \mathbf{z}) = g_h H(x_\alpha - V_{hyp}).
-%\end{equation} 
-%The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  
-%The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  
-%The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  
-%The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.  
-%Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.
-%
-%Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
-%\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
-%\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
-%\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
-%\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
-%\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
-%\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
-%\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
-% \end{equation}
-%where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
-%$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.  
-%
-%Finally, there is a synaptic gating variable as well:
-%\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
-%\end{equation}
-%When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-%
-%In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.  
-%In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible. 
-%The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
-%Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.  
-%Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$
-%
-%Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.  
-%High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.  
-%Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.  
-%The signal was subsequently mean centered.  
-%After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.
-%
-%Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.  
-%Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-%The ``frequency identification" vector is 
-%\begin{equation}
-%v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
-%\end{equation}
-%The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.
-%
-%Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.  
-%The emergent property statistics are the first and second moments of the firing frequencies. 
-%The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
-%\begin{equation}
-%E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
-%\end{equation}
-%for the EPI distribution shown in Fig. \ref{fig:STG}B.
-%Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  
-%The optimization converged after four epochs (Fig. S4).
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
-%\end{center}
-%\begin{flushleft}
-%{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
-%A. Entropy throughout optimization. 
-%B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
-%Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
-%C. The second moment emergent property statistics converge to the emergent property values.}
-%\end{flushleft}
-%\end{figure}
-%
-%For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-%We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce an isotropic 
-%Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
-%
-%We calculated the Hessian at the mode of the inferred EPI distribution.  
-%The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
-%With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-%In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
-%The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
-%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
-%Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
-%The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
-%In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
-\ref{fig:STG}B, inset).
-%The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
-
-We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of intermediate hub frequency in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
-Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following stochastic differential equation:
-\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(\mathbf{x}; \mathbf{z}) + h_{Ca}(\mathbf{x}; \mathbf{z}) + h_K(\mathbf{x}; \mathbf{z}) + h_{hyp}(\mathbf{x}; \mathbf{z}) + h_{elec}(\mathbf{x}; \mathbf{z}) + h_{syn}(\mathbf{x}; \mathbf{z})\right] + dB.
-\end{equation} 
-The input current of each neuron is the sum of the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents as well as gaussian noise $dB$.
-Each current component is a function of all membrane potentials and the conductance parameters $\mathbf{z}$.
-
-The capacitance of the cell membrane was set to $C_m = 1nF$.
-Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
-\begin{equation}  h_{leak}(\mathbf{x}; \mathbf{z}) = g_{leak} (x_\alpha - V_{leak}) 
-\end{equation} 
-\begin{equation}  h_{elec}(\mathbf{x}; \mathbf{z}) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
-\end{equation} 
-\begin{equation}  h_{syn}(\mathbf{x}; \mathbf{z}) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
-\begin{equation}  h_{Ca}(\mathbf{x}; \mathbf{z}) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
-\end{equation} 
-\begin{equation}  h_K(\mathbf{x}; \mathbf{z}) = g_K N (x_\alpha - V_K) 
-\end{equation} 
-\begin{equation}  h_{hyp}(\mathbf{x}; \mathbf{z}) = g_h H(x_\alpha - V_{hyp}).
-\end{equation} 
-The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  
-The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  
-The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  
-The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.  
-Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.
-
-Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
-\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
-\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
-\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
-\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
-\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
-\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
-\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
- \end{equation}
-where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
-$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.
-
-Finally, there is a synaptic gating variable as well:
-\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
-\end{equation}
-When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-Gaussian noise $d\mathbf{B}$ of variance $(1 \times 10^{-12})^2$ A$^2$ makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
-
-In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 300$ time steps of $dt = 25\text{ms}$.  
-The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
-We used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution to measure frequency from simulated time series
-\begin{equation}
-\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top..
-\end{equation}
-
-To measure spiking frequency, we processed simulated membrane potentials with a relu (spike extraction) and low-pass filter with averaging window of size 20, then took the frequency with the maximum absolute value of the complex exponential basis coefficients of the processed time-series.
-The first 20 temporal samples of the simulation are ignored to account for initial transients.
-
-To differentiate through the maximum frequency identification, we used a soft-argmax
-Let $X_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-The soft-argmax is then calculated using temperature parameter $\beta=100$
-\begin{equation}
-\psi_\alpha = \text{softmax}(\beta|X_\alpha|\odot i),
-\end{equation}
-where $i = [0, 1, ..., 100]$.
-The frequency is then calculated as 
-\begin{equation}
-\omega_\alpha = 0.01\psi_\alpha \text{Hz}.
-\end{equation}
-
-Intermediate hub frequency, like all other emergent properties in this work, is defined by the mean and variance of the emergent property statistics.
-In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ to capture variation in frequency between 0.5Hz and 0.6Hz (Equation \ref{eq:EP_STG}).
-As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \\ \left( \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) - 0.55 \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 0.55 \\ 0.025^2 \end{bmatrix}.
-\end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 5,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after five epochs (Fig. S4).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSTG1/figSTG1.pdf}
-\end{center}
-\begin{flushleft}
-\caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
-\textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
-\end{flushleft}
-\label{fig:STG1}
-\end{figure}
-
-For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
-The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$, initialized to a gaussian approximation of samples returned by a preliminary ABC search.
-We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
-EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
-The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
-
-We calculated the Hessian at the mode of the inferred EPI distribution.  
-The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
-With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (strongly negative eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-In Figure \ref{fig:STG}D, the eigenvectors of the Hessian $v_1$ (solid) and $v_2$ (dashed) are shown evaluated at the mode of the distribution.  
-The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
-Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space are chosen arbitrarily.
-%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
-%Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
-%The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
-%In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
-\ref{fig:STG}B, inset).
-%The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
-
-\subsubsection{Primary visual cortex}\label{methods_V1}
-In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
-\begin{equation}
-    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}),
-\end{equation}
-where the noise is an Ornstein-Uhlenbeck process $\bm{\epsilon} \sim OU(\tau_{\text{noise}}, \bm{\sigma})$
-\begin{equation}
-\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\tilde{\sigma}_\alpha dB
-\end{equation}
-with $\tau_{\text{noise}} = 5\text{ms} > \tau = 1\text{ms}$.
-The noisy process is parameterized as
-\begin{equation}
-\tilde{\sigma}_\alpha = \sigma_\alpha \sqrt{1 + \frac{\tau}{\tau_{\text{noise}}}},
-\end{equation}
-so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
-As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
-Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
-\end{center}
-\caption{\small (V1 1)
-(Left) Simulations for small increases in neuron-type population input.
-Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
-(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
- }
- \label{fig:V1_1}
-\end{figure}
-
-\begin{figure}
-\caption{\small (V1 2)
-Posterior predictive distributions of the sum of squares of each pair of noise parameters.
- }
- \label{fig:V1_2}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
-\end{center}
-\end{figure}
-
-\begin{equation}
-W_{\text{fit}} =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
-W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
-W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
-W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
- \begin{bmatrix} 2.18 & -1.19 & -.594 & -.229 \\
- 1.66 & -.651 & -.680 & -.242 \\
- .895 & -5.22 \times 10^{-3} & -1.51 \times 10^{-4}   & -.761 \\
- 3.34 &  -2.31 & -.254  & -2.52 \times 10^{-4} \\
- \end{bmatrix},
-\end{equation}
-
-\begin{equation}
-\mathbf{h}_{b,\text{fit}} =
- \begin{bmatrix} .416 \\ .429 \\ .491 \\ .486 \end{bmatrix} ,
-\end{equation} 
-and
-\begin{equation} 
-\mathbf{h}_{c,\text{fit}} = 
-\begin{bmatrix} .359 \\ .403 \\ 0 \\ 0 \end{bmatrix}.
-\end{equation}
-
-To obtain rates on a realistic scale (100-fold greater), we map these fitted parameters to an equivalence class
-
-\begin{figure}[h]
-\caption{\small (V1 3)
-EPI posterior for $\mathcal{X}(10\text{ Hz})$.
- }
- \label{fig:V1_3}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
-\end{center}
-\end{figure}
-
-\begin{equation}
-W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
-W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
-W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
-W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
- \begin{bmatrix} .218 & -.119 & -.0594 & -.0229 \\
- .166 & -.0651 & -.068 & -.0242 \\
- .0895 & -5.22 \times 10^{-4} & -1.51 \times 10^{-5}   & -.0761 \\
- .334 &  -.231 & -.0254  & -2.52 \times 10^{-5} \\
- \end{bmatrix},
-\end{equation}
-
-\begin{equation}
-\mathbf{h}_b =  \begin{bmatrix} h_{b,E} \\ h_{b,P} \\ h_{b,S} \\ h_{b,V} \end{bmatrix} =
- \begin{bmatrix} 4.16 \\ 4.29 \\ 4.91 \\ 4.86 \end{bmatrix} ,
-\end{equation} 
-and
-\begin{equation} 
-\mathbf{h}_c = \begin{bmatrix} h_{c,E} \\ h_{c,P} \\ h_{c,S} \\ h_{c,V} \end{bmatrix} = 
-\begin{bmatrix} 3.59 \\ 4.03 \\ 0 \\ 0 \end{bmatrix}.
-\end{equation}
-
-Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{ Hz}, 25\text{ Hz}\right]$.
-Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
-\begin{equation}
-s_E(\mathbf{x}; \mathbf{z}) =\mathbb{E}_{x}\left[\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]}\right]
-\end{equation}
-
-For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
-The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
-EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
-The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
-
-In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
-Specifically, we calculated
-\begin{equation}
-\begin{split}
-\mathbf{z}^*(\sigma_{P,\text{fixed}}) = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } \sigma_P = \sigma_{P,\text{fixed}}\
-\end{split}
-\end{equation}
-At each mode $\mathbf{z}^*$, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
-
-\subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
-TODO Agostina and I are putting this together now.
-
-\subsubsection{Superior colliculus}\label{methods_SC}
-In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
-They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
-Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
-\begin{equation}
-x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
-\end{equation}
-where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
-During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
-\begin{equation}
-x_\alpha = (1-\gamma)\phi(u_\alpha).
-\end{equation}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.9]{figures/figSC1/figSC1.pdf}
-\end{center}
-\caption{\small (SC1): \textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
-\textbf{B}. Same as A colored by Anti task accuracy.  
-\textbf{C}. Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p-value.
-}
-\label{fig:SC1}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
-\end{center}
-\caption{\small (SC2):  
-\textbf{A}.
-Simulations in network regime $\mathbf{z}_1$ (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
-\textbf{B}. Same as A for network regime $\mathbf{z}_2$.
-}
-\label{fig:SC2}
-\end{figure}
-
-We order the neural populations of $x$ and $u$ in the following manner
-\begin{equation}
-\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
-\end{equation}
-
-which evolve according to
-\begin{equation}
-\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
-\end{equation}
-with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.  
-The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
-\begin{equation}
-W = \begin{bmatrix} sW & vW & hW & dW \\ vW & sW & dW  & hW \\ hW & dW & sW & vW \\ dW  & hW & vW  & sW \end{bmatrix}.
-\end{equation}
-
-The circuit receives four different inputs throughout each trial, which has a total length of 1.8s.
-\begin{equation}
-\mathbf{h} = \mathbf{h}_{\text{constant}}  + \mathbf{h}_{\text{P,bias}} + \mathbf{h}_{\text{rule}} +  \mathbf{h}_{\text{choice-period}} + \mathbf{h}_{\text{light}}.
-\end{equation}
-
-There is a constant input to every population,
-\begin{equation}\mathbf{h}_{\text{constant}} = I_{\text{constant}} [1, 1, 1, 1]\top,
-\end{equation}
- a bias to the Pro populations
-\begin{equation}\mathbf{h}_{\text{P,bias}} = I_{\text{P,bias}} [1, 0, 1, 0]\top,
-\end{equation}
-rule-based input depending on the condition
-\begin{equation}\mathbf{h}_{\text{P,rule}}(t) = \begin{cases}
-                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases}
-\end{equation}
-\begin{equation} \mathbf{h}_{\text{A,rule}}(t) = \begin{cases}
-                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases},
-\end{equation}
-a choice-period input
-\begin{equation} \mathbf{h}_{\text{choice}}(t) = \begin{cases}
-                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases},
-\end{equation}
-and an input to the right or left-side depending on where the light stimulus is delivered    
-\begin{equation}  \mathbf{h}_{\text{light}}(t) = \begin{cases}
-                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Left} \\
-                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Right} \\
-                            0,              & \text{otherwise}
-                         \end{cases}.
-\end{equation}
-The input parameterization was fixed to $I_{\text{constant}} = 0.75$, $I_{\text{P,bias}} = 0.5 $, $I_{\text{P,rule}} = 0.6$,  $I_{\text{A,rule}} = 0.6$,  $I_{\text{choice}} = 0.25$,  and $I_{\text{light}} = 0.5$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.9]{figures/figSC3/figSC3.pdf}
-\end{center}
-\caption{\small (SC3):  
-\textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
-\textbf{B}. Eigenvalues of connectivities of EPI distribution versus task accuracies.
-}
-\label{fig:SC3}
-\end{figure}
-
-The accuracies of $p_P$ and $p_A$ are calculated as
-\begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
-\end{equation}
-and 
-\begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
-\end{equation}
-given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function, and the accuracy is averaged over 200 independent trials.  The Heaviside step function is approximated as
-\begin{equation}
-\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
-\end{equation}
-where $\beta = 100$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
-\end{center}
-\caption{\small (SC4):  \textbf{A}.
-Pairplots of eigenvalues of connectivity matrices in EPI distribution colored by Pro task accuracy.
-Red and purple stars and arrows correspond to eigenvalues and sensitivity directions $\mathbf{z}_1$, $\mathbf{z}_2$, $\mathbf{v}_1$, and $\mathbf{v}_2$.
-\textbf{B}.
-Same colored by Anti task accuracy.
-}
-\label{fig:SC4}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC5/figSC5.pdf}
-\end{center}
-\caption{\small (SC5): 
-\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Connectivity eigenvalues versus the task error induced by delay period inactivation.
-\textbf{C}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
-}
-\label{fig:SC5}
-\end{figure}
-
-Writing the EPI posterior as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 7.5\%^2 \\ 7.5\%^2 \end{bmatrix}.
-\end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after six epochs (Fig. \ref{fig:SC5}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figSC6/figSC6.pdf}
-\end{center}
-\caption{\small (SC6): 
-\textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
-}
-\label{fig:SC5}
-\end{figure}
-
-For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
-The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 25$ at greatest entropy across random seeds.
-
-To make sense of this inferred distribution, we identified two modes used to represent the two regimes of connectivity in this posterior:
-\begin{equation}
-\begin{split}
-\mathbf{z}_1 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. }  hw = -1.25\text{, }  sW > 0
-\end{split}
-\end{equation}
-and 
-\begin{equation}
-\begin{split}
-\mathbf{z}_2 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } hw = -1.25\text{, } sW < 0
-\end{split}
-\end{equation}
-
-To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
-\ref{fig:SC3}A).
-These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-We found significant trends across the EPI posterior connectivities: the eigenvalues $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ were correlated with $p_P$, while $\lambda_{\text{all}}$ was anticorrelated with $p_P$.
-$\lambda_{\text{all}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ were all significantly anticorrelated with $p_A$.
-
-Under this decomposition, we can re-visualize the posterior in eigenvalue space (Fig. \ref{fig:SC4}).
-Furthermore, we can project the dimensions of sensitivity into eigenvalue space as well, giving us a more intuitive sense of how connectivity affects computation in each regime.
-We see that sensitivity dimensions $\mathbf{v}_1$ and $\mathbf{v}_2$, which cause $p_P$ to increase and a regime dependent change in $p_A$, both point in the direction of increasing $\lambda_{\text{side}}$ and decreasing $\lambda_{\text{task}}$.
-These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
-As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
-A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
-
-During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.
-We found that $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ were significantly anticorrelated with Anti error during delay period inactivation.
-Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.
-
-\subsubsection{Scaling EPI for stable amplification in RNNs}\label{methods_RNN}
+\subsubsection{Methods}\label{methods_RNN}
 We examine the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.

### 2021-03-23 neural_circuits/LRRNN.py
fig 4 nb

@@ -367,9 +367,9 @@ def get_SNPE_conv(N, g, K, x0, eps,
             print('bad')
         
         assert(num_rounds < args.max_rounds)
-        conv_inds = np.nonzero(distances<eps)[0]
-        conv_time = np.nan if len(conv_inds) == 0 else round_times[conv_inds[0]]
-        _conv_sims = np.nan if len(conv_inds) == 0 else round_sims[conv_inds[0]]
+        conv_inds = np.nonzero(distances[1:]<eps)[0]
+        conv_time = np.nan if len(conv_inds) == 0 else round_times[conv_inds[0]+1]
+        _conv_sims = np.nan if len(conv_inds) == 0 else round_sims[conv_inds[0]+1]
         conv_times.append(conv_time)
         conv_sims.append(_conv_sims)
     
@@ -398,7 +398,7 @@ def get_EPI_conv(N, g, K, random_seeds, eps=None):
         if eps is not None:
             R = np.stack((_epi_df['R1'].to_numpy(), _epi_df['R2'].to_numpy()), axis=1)
             distances = np.linalg.norm(R, axis=1)
-            conv_inds = np.nonzero(distances<eps)[0]
+            conv_inds = np.nonzero(np.logical_and(distances<eps, _epi_df['iteration'].to_numpy()>0))[0]
         else:
             converged = _epi_df['converged'].to_numpy()==1.
             conv_inds = np.nonzero(converged)[0]

### 2021-03-23 written/eLifeRevision/scaling_RNN_inference.tex
latest update to scalingn writeup

@@
 
-\subsection{Scaling inference in RNNs with EPI} \label{results_LRRNN}
-Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinisically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+\section{Scaling inference in RNNs with EPI} \label{results_LRRNN}
+Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinsically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
 It has been shown that to generate such amplified, yet stabilized responses, the connectivity of RNNs must be non-normal \cite{goldman2009memory, murphy2009balanced}, and satisfy additional constraints \cite{bondanelli2020coding}.
@@
 \caption{\small 
-\textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivites producing stable amplification.  
+\textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivities producing stable amplification.  
 Each dot shows convergence time for an individual random seed.
@@
 Each trace (15 per algorithm) corresponds to simulation of one $z$.
-(Below) Ratio of obtained samples producing stable amplification, monotnoic decay, and instability.
+(Below) Ratio of obtained samples producing stable amplification, monotonic decay, and instability.
 }
@@
 For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative likelihood-free inference approaches.
-We ran sequential monte-carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
-SMC-ABC is a rejection sampling approach that uses sophisticated techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
+We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
+SMC-ABC is a rejection sampling approach that SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section Related Methods).
 Unlike EPI, these statistical inference techniques do not control the mean or variance of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
 $N$ and $g$, Fig.
-\ref{fig:RNN1}).
+\ref{fig:RNN2}).
 To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.25 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
@@
 \ref{fig:LRRNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification  (Fig.
-\ref{fig:LRRNN}C, green), but are concentrated and offset from $\mathbf{x}_0$.
-When using SNPE the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable amplified activity norms $|r(t)|$, SMC produces a limited variety in respones, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:LRRNN}C, green).
+When using SNPE the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
 
-From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and have the additional benefit of fast sampling after optimization.
-EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration and constantly adapting the approximation $q_{\bm{\theta}}(\mathbf{z})$ to the parameter distribution.
-While EPI can be used for scientific insight into RNNs, this analysis shows that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
+From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC.
+Another benefit over SMC-ABC is that EPI and SNPE offer fast additional samples after optimization.
+EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration.
+Through this example, we've shown that EPI can be used for insight into RNNs with respect to their theoretical properties, and that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
 
@@
 
-\subsubsection{Methods}\label{methods_RNN}
+\subsection{Methods}\label{methods_RNN}
 We examine the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
@@
 \end{equation}
-In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section \ref{methods_related}).
+In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section Related Methods).
 
@@
 \end{equation}
-Gradients of maximum eigenvalues of hermitian matrices like $W^s$ are available with modern automatic differentiation tools.
+Gradients of maximum eigenvalues of Hermitian matrices like $W^s$ are available with modern automatic differentiation tools.
 To differentiate through the $\text{real}(\lambda_1)$, we solve the following equation for eigenvalues of rank-2 matrices using the rank reduced matrix $W^r = V^{\top} U$
@@
 We compared EPI to two alternative likelihood-free inference (LFI) techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
-Appoximate Bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parametes $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
+Approximate Bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parameters $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
 Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) is the state-of-the-art ABC method, which leverages SMC techniques to improve sampling speed.
@@
 \textbf{C}. Validation log probabilities throughout SNPE optimization. Same conventions as B.
-\textbf{D}. Adherance to EPI constraints. Same conventions as B.
+\textbf{D}. Adherence to EPI constraints. Same conventions as B.
 }
@@
 
-To compare the efficiency of these algorithms for infering RNN connectivity distributions producing stable amplification, we develop a convergence criteria that can be used across methods.
+To compare the efficiency of these algorithms for inferring RNN connectivity distributions producing stable amplification, we develop a convergence criteria that can be used across methods.
 While EPI has its own hypothesis testing convergence criteria for the emergent property, it would not make sense to use this criteria on SNPE and SMC-ABC which do not constrain the means and variances of their predictions.
@@
 \ref{fig:RNN3}B).
-Finally we increased $n_{\text{round}}$ by an order of magnitude, which yielded convergence in one case, but no others.
+Finally, we increased $n_{\text{round}}$ by an order of magnitude, which yielded convergence in one case, but no others.
 We found no way to improve the convergence rate of SNPE without making more aggressive hyperparameter choices requiring high numbers of simulations.
 
+In Figure \ref{fig:LRRNN}C-D, we show samples form the random seed resulting in emergent property convergence at greatest entropy (EPI), the random seed resulting in greatest validation log probability (SNPE), and the result of all converged random seeds (SMC).
+
 \end{document}

### 2021-03-25 written/eLifeRevision/Bittner_eLife2020.tex
typo edits

@@
 \subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
-Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinisically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinsically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
 It has been shown that to generate such amplified, yet stabilized responses, the connectivity of RNNs must be non-normal \cite{goldman2009memory, murphy2009balanced}, and satisfy additional constraints \cite{bondanelli2020coding}.
@@
 \ref{fig:LRRNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters always exhibit stable amplification  (Fig.
-\ref{fig:LRRNN}C, green), but are concentrated and offset from $\mathbf{x}_0$.
-When using SNPE the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable amplified activity norms $|r(t)|$, SMC produces a limited variety in respones, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
-
-From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and have the additional benefit of fast sampling after optimization.
-EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration and constantly adapting the approximation $q_{\bm{\theta}}(\mathbf{z})$ to the parameter distribution.
-While EPI can be used for scientific insight into RNNs, this analysis shows that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:LRRNN}C, green).
+When using SNPE the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
+
+From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC.
+Another benefit over SMC-ABC is that EPI and SNPE offer fast additional samples after optimization.
+EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration.
+Through this example, we've shown that EPI can be used for insight into RNNs with respect to their theoretical properties, and that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
 

### 2021-03-29 written/eLifeRevision/Bittner_eLife2020.tex
latest edits

@@
 The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
-A key process in neuroscientific modeling is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
-While statistical inference has proven effective on a broad variety of neuroscientific datasets, we clarify an important incongruity between theoretical approaches to neuroscience and this probabilistic methodology.
-Theoretical neuroscience is focused on computational properties and how they emerge from biological mechanisms, rather than noisy experimental datasets and their quantified structure.
-In this work, we present a novel technique to directly infer circuit model parameters producing these computational properties.
-This method tailors deep inference, the use of deep neural networks for statistical inference, to the nature of theoretical inverse problems, enabling scaling to high dimensions and considerable technical simplifications.
-With this method, we bring deep inference to bear on important questions in theoretical neuroscience, and demonstrate the broad range of insightful analyses this approach allows.
-First, we emphasize the general applicability of this approach by inferring channel conductance parameters that produce a distinctive spiking frequency in a biophysical model of the stomatogastric ganglion.
-Then, in a model of primary visual cortex with multiple neuron-types, where  analysis becomes untenable as more neuron-types are included, we use this technique to discover how noise properties govern excitatory population variability.
-Next, in a model of superior colliculus, we identify and characterize two distinct regimes of connectivity that facilitate switching between opposite tasks throughout interleaved trials. 
-Finally, we scale inference to 1,000-dimensional parameter spaces of RNN connectivity that exhibit stable, yet amplified responses.
-These analyses illustrate how we can further leverage the power of deep learning towards solving inverse problems in theoretical neuroscience.
+A key process in circuit modeling is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
+To solve challenging inverse problems modeling neural datasets, neuroscientists have used statistical inference techniques to much success. 
+However, most research in theoretical neuroscience focuses on how computation emerges in biologically interpretable circuit models, and how the model parameters govern computation; it is not focused on the latent structure of empirical models of noisy experimental datasets.
+In this work, we present a novel technique that brings the power and versatility of the probabilistic modeling toolkit to theoretical inverse problems.
+Our method uses deep neural networks to learn parameter distributions with rich structure that have specific computational properties in biologically relevant models.
+This methodology is explained through a motivational example inferring conductance parameters in an STG subcircuit model.
+Then, with RNNs of increasing size, we show that only EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
+In the remainder of this work, we explain novel theoretical insights through the examination of intricate parametric structure in complex circuit models.
+In a model of primary visual cortex with multiple neuron-types, where analysis becomes untenable with each additional neuron-type, we discovered how noise distributed across neuron-types governs the excitatory population.
+Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials. 
+We also found that all task-switching connectivities in this model reproduce behaviors from inactivation experiments, further establishing this hypothesized circuit model.
+Beyond its scientific contribution, this work illustrates the variety of analyses possible once deep learning is harnessed towards solving theoretical inverse problems.
 
@@
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-Alternatively, we can gain insight into these complex models by identifying the distribution of parameters that produce a specific computational property.
+Still, we can gain insight into these complex models by identifying the distribution of parameters that produce computations.
 By solving the inverse problem in this way, scientific analysis of complex biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-%By solving the inverse problem in this way, scientists can reason about the sensitivity and robustness of the model with respect to different parameter combinations \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-
-%The preferred formalism for parameter identification in science is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-%However, most neural circuit models in theoretical neuroscience stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood central to the probabilistic modeling toolkit.
-%Therefore, the most popular approaches to the inverse problem have been likelihood-free methods such as approximate Bayesian computation (ABC) \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-% 
-%Of course, the challenge of doing inference in complex models has arisen in many scientific fields.
-%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as powerful inference engines: a flexible function family that can map observations back to probability distributions quantifying the likely parameter configurations.
-%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder (VAE) \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-%Indeed, these tools have been used to great success in neuroscience as well, in particular for interrogating hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}.
-%These works have used deep neural networks to expand the domain of neural data sets amenable to statistical modeling \cite{paninski2018neural}.
-
-One preferred formalism for parameter identification is statistical inference, which has been used to great success in neuroscience through the stipulation of statistical generative models \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring} (see review, \cite{paninski2018neural}).
-Recent work has used variational autoencoders (VAEs) \cite{kingma2013auto, rezende2014stochastic} to interrogate hidden states in models of both cortical population activity \cite{gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring} and animal behavior \cite{wiltschko2015mapping, johnson2016composing, batty2019behavenet}, thus expanding the domain of neural data sets amenable to statistical modeling.
-However, most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood necessary for statistical inference.
+However, this work clarifies an important incongruity between this methodology and theoretical approaches to neuroscience.
+
+A pivotal detail in theoretical neuroscience is the manner in which the computation is specified.
+One common approach is to use data that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
+While a host of suitable algorithms have been developed for inferring parameters in models of neural data \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}), the level of insight gained strongly depends on the quantity and quality of available data.
+Ultimately theoretical neuroscience is concerned with the computational properties -- the \emph{emergent phenomena} -- of our models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, not noisy observed datasets.
+To use the aforementioned inference paradigm, scientists must shoehorn such mathematical criteria into an artificial dataset compatible with existing statistical approaches.
+
+In addition, another crucial challenge in neuroscientific modeling has been the inversion of biologically realistic or ``mechanistic" models.
+Most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood necessary for statistical inference.
 Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference (LFI) methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
 A flourishing new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional LFI approaches.
-However, as we detail, all of these approaches require good datasets for the scientific question at hand.
+However, to use these methods in theoretical neuroscience, we still must represent computation with data in some way.
+Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into an empirical framework.
 
@@
 
-This work seeks to clarify an important incongruity between theoretical approaches to neuroscience and existing statistical inference methodology.
-In theoretical neuroscience, we are concerned with the computational properties -- the \emph{emergent phenomena} -- of our models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, not noisy observed datasets \cite{paninski2018neural}.
-To use the aforementioned inference paradigm, scientists must shoehorn such mathematical criteria into an artificial dataset compatible with existing statistical approaches.
-Theorists are therefore barred from using the probabilistic modeling toolkit for science, unless they reformulate their inverse problem to fit an evidence accumulation framework.
-
 These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
 As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets exhibiting some mathematical criteria.
-EPI fixates on the inverse problem by constraining the predictions of the inferred parameter distribution to produce the emergent property exactly.
-Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-By using deep probability distributions EPI flexibly captures parametric structure in mechanistic models.
-This technique is designed to identify the full parameter space producing an emergent property and facilitate the seemless structural analysis of the inferred parameter distribution.
-Thus, EPI provides neuroscientists with an efficient, versatile probabilistic modeling toolkit designed explicitly for theoretical inverse problems.
-
-Equipped with this method, we bring deep inference to bear on theoretical neuroscience to an unprecedented extent.
-Throughout this work, we showcase the capabilities of EPI on four neural circuit models across ranges of biological realism, neural system function, and network scale.
-First, we show EPI's ability to capture subtle, nonlinear parametric structure in a stomatogastric ganglion subcircuit model \cite{gutierrez2013multiple}.
-In a model of primary visual cortex \cite{palmigiano2020structure}, we show how to gain insight from multiple inferred distributions.
-Next, we used EPI to idenitfy and structurally characterize multiple parametric regimes of superior colliculus activity in a model of task switching \cite{duan2018collicular}.
-Finally, we emphasize the superior scalability of EPI compared to other LFI techniques by inferring high-dimensional distributions of RNN connectivities that exhibit amplified, yet stable responses -- a hallmark of cortical sensory systems \cite{more, bondanelli2020coding}.
-
-Most importantly of all in this work, we present novel theories of neural computation borne from EPI analysis.
-We identified an unknown parametric rule of variability with respect to inhibitory neuron type in a V1 model, where analytic techniques became untenable.
-Furthermore, we identified multiple regimes of SC connectivity which confer rapid task switching, and used the structural analytic tools of EPI to gain a mechanistic understanding of circuit responses.
-These valuable theoretical insights illustrate the value of deep inference for the interrogation of neural circuit models.
+EPI constrains the predictions of the inferred parameter distribution to produce the emergent property, which requires a variant of probabilistic inference methods \cite{loaiza2017maximum}.
+Importantly, EPI uses deep learning to make rich, flexible approximations to the parameter distribution \cite{rezende2015variational} that produces an emergent property.
+The structure captured by these deep probability distributions are scientifically valuable, revealing the sensitivity and robustness of the emergent property to different parameter combinations.
+Perhaps most powerfully, EPI facilitates inference in mechanistic models, allowing theorists to capture rich parametric structure in biologically realistic models that is conditioned upon the emergent phenomena of interest.
+
+Equipped with this method, we prove out the potential of EPI by demonstrating its capabilities and presenting novel theoretical findings borne from its analysis.
+First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion \cite{goldman2001global, gutierrez2013multiple}.
+Then, we show EPI's superior scaling properties by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+In a model of primary visual cortex (V1) with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
+Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
+Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials.
+EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
+We quantified parametric sensitivity and robustness in each regime by simply querying the deep probability distribution learned by EPI, which produced a mechanistic understanding of cortical responses in each regime.
+Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon.
+These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.
 
@@
 To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
-Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and neural activity \cite{goldman2001global, prinz2004similar}.
+Despite full knowledge of STG connectivity and a precise characterization of its rhythmic pattern generation, biophysical models of the STG have complicated relationships between circuit parameters and computation \cite{goldman2001global, prinz2004similar}.
 %A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
-A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
+A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A.
+The jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$, smooth connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.
+Note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
 Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  
-The hub neuron (hub) couples with either the fast or slow population or both.
+The hub neuron (hub) couples with either the fast or slow population, or both.
 
 %Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
-Second, once the model is selected, one defines the emergent phenomena of scientific interest.
+Second, once the model is selected, one must specify what the model should produce.
+In typical statistical inference, this is a dataset -- either collected experimentally or constructed by scientists to fit this emprical paradigm.
+In EPI, a different approach is taken, in which we define an emergent property: a set of mathematical criteria to be obeyed by the datasets predicted by the inferred distribution.
 %To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
@@
 An interesting emergent property of this stochastic model is when the hub neuron fires at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
-%This emergent property is shown in Figure \ref{fig:STG}D at a frequency of 0.53Hz.
 This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
@@
 % Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-Third, parameter analyses ensue: brute-force parameter sweeps, ABC sampling \cite{liepe2014framework}, and sensitivity analyses \cite{raue2009structural} are all routinely used to reason about what parameter configurations lead to an emergent property.
-In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-
-%A.
-For a choice of model (STG) and emergent property (network syncing), emergent property inference (EPI, gray box) learns a distribution of the model parameters $z = \left[g_{\text{el}}, g_{\text{synA}} \right]$ producing network syncing.
-%In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
-%Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-%Network syncing traces are colored by log probability density of their generating parameters (stars) in the EPI-inferred distribution.  
-%B. The EPI distribution of STG model parameters producing network syncing.  
-%Samples are colored by log probability density.  
-%Distribution contours of emergent property value error are shown at levels of $2.5 \times 10^{-5}$, $5 \times 10^{-5}$, $1 \times 10^{-4}$, $2 \times 10^{-4}$, and $4 \times 10^{-4}$ (dark to light gray).
-%Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
-%Simulated activity is shown for three samples (stars). 
-%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
-%$v_1$ is sensitive to network syncing ($p<10^{-4}$), while $v_2$ is not ($p=0.67$) (see Section \ref{methods_STG}).  
-%C.
-Deep probability distributions map a latent random variable $\mathbf{w}$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = f_{\bm{\theta}}(\mathbf{w})$ distributed as $q_{\bm{\theta}}(\mathbf{z})$.
-%D.
-EPI optimization: To learn the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ of model parameters that produce an emergent property, the emergent property statistics $T(\mathbf{x})$ are set in expectation over model parameter samples $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and model simulations $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ to emergent property values $\bm{\mu}$.
+Third, the model parameters producing these outputs are inferred.
+Most often, brute-force parameter sweeps or rejection sampling techniques \cite{liepe2014framework} are used to identify parameters whose model simulations are close to data or some desired feature.
+%Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
+%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
+In this last step lies the opportunity for a paradigmatic shift away from empirical data-oriented representations of model output.
+By precisely quantifying the emergent property of interest as a statistical feature of the model, we can infer a probability distribution over parameter configurations that produce this emergent property.
+This unlocks the deep probabilistic modeling toolkit for treating theoretical inverse problems.
+
 \begin{figure}
@@
 \textbf{A}. Conductance-based biophysical model of the STG subcircuit.
-In the STG model, jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$. 
-Other connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.  
-Parameters are represented by the vector $\mathbf{z}$ and membrane potentials by the vector $\mathbf{x}$.
-The evolution of this model's activity $\mathbf{x}(t)$ is predicated by differential equations.
 \textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
-In this example, spiking frequency is measured from simulated activity of the STG model at parameter choices of $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
-\textbf{C}. The emergent property of intermediate hub frequency, in which the hub neuron fires at a rate between the fast and slow frequencies.
-This emergent property is defined by a mean and variance on the emergent property statistic.
-Simulated activity traces are colored by log probability density of their generating parameters in the EPI-inferred distribution (Panel E).
+Simulated at $g_{\text{el}} = 4.5$nS and $g_{\text{synA}} = 3$nS.
+\textbf{C}. The emergent property of intermediate hub frequency.
+Simulated activity traces are colored by $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ of generating parameters. (Panel E).
 \textbf{D}. For a choice of model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$.
-In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced.
-The EPI posterior distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$. 
 \textbf{E}. The EPI posterior producing intermediate hub frequency.
 Samples are colored by log probability density.  
-Distribution contours of average hub neuron frequency from mean of .55 Hz are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Eigenvectors of the Hessian at the mode of the inferred distribution are indicated as $\mathbf{v}_1$ (solid) and $\mathbf{v}_2$ (dashed) with lengths scaled by the square root of the absolute value of their eigenvalues.
+Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
+Dimension of sensitivity $\mathbf{v}_1$ (solid) and degeneracy $\mathbf{v}_2$.
 %(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
-\textbf{F} Simulations from parameters in E.  
-(Top) The predictive distribution of the posterior obeys the emergent property.
-The black and gray dashed lines show the mean and two standard deviations according the emergent property, respectively.
+\textbf{F} (Top) The predictive distribution of EPI.
+The black and gray dashed lines show the mean and two standard deviations according the emergent property.
 (Bottom) Simulations at the starred parameter values.
@@
 \ref{fig:STG}E).
-This distribution -- represented for clarity as samples from the distribution -- is then the most random parameter distribution producing the emergent property. 
+This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution producing the emergent property. 
 In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
@@
 %First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
-First, we consider the model as a coupled set of differential equations \cite{gutierrez2013multiple}.  
+First, as is typical, we consider the model as a coupled set of differential equations \cite{gutierrez2013multiple}.  
 In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
@@
 \end{equation} 
-where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG}).
+where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG} for more detail).
 
@@
 %Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
-Mathematically, we achieve this with two constraints: by setting the mean hub frequency of model generated activity to 0.55Hz.
+Mathematically, we achieve this with two constraints: by fixing the mean hub frequency over the inferred parameter distribution of $\mathbf{z}$ and its resulting simulations $\mathbf{x}$ to 0.55Hz,
 \begin{equation}\label{eq:EP_STG1}
@@
 \end{equation}
-and requiring that the variance around this mean is small
+and requiring that the variance of the hub frequency over the produced simulations is small
 \begin{equation}\label{eq:EP_STG2}
@@
 \end{equation}
-This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the compatible parameter space.
+This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the inferred parameter distribution.
+By constraining the means and variances of emergent property statistics over $\mathbf{z}$ as well as the stochasticity of $\mathbf{x}$, we can precisely control the behavior that the inferred distribution that EPI infers.
 In general, an emergent property
@@
 defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
-The choice of $\bm{\sigma}^2$ predicates the degree of variability in around mean $\bm{\mu}$ that is consistent with the emergent property.
+The choice of $\bm{\sigma}^2$ predicates the degree of variability around the mean $\bm{\mu}$ that is consistent with the emergent property.
 
 Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep generative distribution of the normalizing flow class \cite{rezende2015variational, dinh2017density, papamakarios2017masked} -- neural networks which transform a simple distribution into a suitably complicated distribution (as is needed here).
-This deep distribution is represented in Figure \ref{fig:STG}D (see Section \ref{methods_EPI}).
-Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we choose a normative selection principle imposing no additional structure beyond the production of the emergent property \cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}, which is the same normative principle of Bayesian inference (see Section \ref{methods_VI}).
+This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, dinh2017density, papamakarios2017masked} (Figure \ref{fig:STG}D, EPI box).
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ to a suitable complicated distribution (see Section \ref{methods_NF} for more details).
+Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (or ``entropic") distribution, which is the same choice made in Bayesian inference (see Section \ref{methods_VI}).
+%\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
+In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI})
+The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, to emphasize that we have conditioned our parameter distribution on emergent property $\mathcal{X}$.
 
@@
 
-The probabilities of the distribution inferred from EPI are the densities of these parameters in the most random distribution producing the emergent property.
-While existing approaches to probabilistic structural identifiability analysis use the lens of evidence accumulation \cite{raue2009structural, gonccalves2019training}, sensitivity and robustness of parameter space dimensions with respect to emergent phenomena can be directly quantified with EPI.
-Sensitivity quantifications are measured by the second order derivative of EPI probability along the parameteric dimension of interest: to what extent is the emergent property maintained or diminished along this dimension?
-Once an EPI distribution has been inferred, this second order derivative requires trivial computation (as long as the correct architecture class is chosen, see Section \ref{methods_NF}).
-Equipped with this method, we may examine structure in the resulting parameter distributions or make comparisons between distributions conditioned at different levels of the same emergent property statistic.
-In Sections \ref{results_V1} and \ref{results_SC}, we prove out the value of EPI by using these techniques to investigate and produce novel scientific insight.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI reveals how neuron-type specific noise governs variability in the stochastic stabilized supralinear network} \label{results_V1}
-
-\begin{figure}
-\vspace{-1cm}
-\begin{center}
-\includegraphics[scale=0.8]{figures/fig2/fig2.pdf}
-\end{center}
-\vspace{-1cm}
-\caption{\small Emergent property inference in the stochastic stabilized supralinear network (SSSN)
-\textbf{A}.
-Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).
-Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
-Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
-Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
-\textbf{B}. Steady-state responses of the SSN model (deterministic, $\bm{\sigma}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
-\textbf{C}.
-Transient network responses of the SSSN model at 50 \% contrast.
-(Left) Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise realization.
-(Right) Mean (solid line) and standard deviation (shading) of responses.
-\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on E-population variability.
-The posterior predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
-\textbf{E}.
-(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posteriors $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
-Each black dot shows the mode at each $\sigma_P$.
-The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
-\textbf{F}.
-The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each posterior $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
-}
- \label{fig:V1}
-\end{figure}
-
-Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
-Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
-Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-Here, we use EPI to analyze V1 models of greater complexity in order to characterize properties of slow noise governing circuit variability.
-
-We considered the response properties of a nonlinear dynamical V1 circuit model (Fig.
-\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$ from synaptic projections of effective connectivity $W$ and an external input $\mathbf{h}$, which determine the population rate via supralinear nonlinearity $\phi = \left[\right]^2_+$.
-The input is also comprised of a slow noise component $\bm{\epsilon} \sim OU\left(\tau_{\text{noise}}, \bm{\sigma} \right)$ of time scale $\tau_{\text{noise}} > \tau$ and variance parameters $\bm{\sigma}$ (see Section \ref{methods_V1})
-\begin{equation}
-    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}).
-\end{equation}
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have inhibitory multiplicity, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
-Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
-Here, we consider fixed weights $W$ and input $\mathbf{h}$ according to a fit of the deterministic model to contrast responses \cite{palmigiano2020structure} (Fig.
-\ref{fig:V1}B), and study the effect of noise parameterization $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on fluctuations at 50\% contrast.
-
-For this SSSN, we are interested in how noise variability across neural populations governs stochastic fluctuations in the E-population.
-Here, we quantify different levels $y$ of E-population variability with the emergent property
-\begin{equation}\label{eq:EP_V1}
-\begin{split}
-\mathcal{X}(y) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ y \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2,
-\end{split}
-\end{equation}
-where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-
-We ran EPI to obtain a posterior distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
-From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
-\ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
-Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions of the posterior (Fig.
-\ref{fig:V1}D, bottom-right).
-Together, these observations imply a parametric manifold of degeneracy with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ in the $\sigma_E$-$\sigma_P$ marginal (Fig.
-\ref{fig:V1}E).
-The dimensions of sensitivity conferred by EPI and this plain visual structure suggest a quadratic relationship in the emergent property statistic $s_E(\mathbf{x}; \mathbf{z})$ and parameters $\mathbf{z}$, which is preserved at a greater level of variability $\mathcal{X}(10\text{ Hz})$ (Fig.
-\ref{fig:V1}E).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ (Fig \ref{fig:V1}F, $p=0$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two posteriors (Fig.
-\ref{fig:V1_3}, $p=.402$).
-
-While a quadratic relationship in $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ is potentially derivable by extending the derivation in Section \ref{methods_V1} to the case of $\tau \neq \tau_{\text{noise}}$, the coefficients in front of each quadratic term would be unruly, and likely escape comprehensible analysis.
-This makes EPI an attractive tool for revealing the characteristics of noise governing variability and for answering other questions in this complex model.
-Intriguingly, this circuit exhibited a paradoxical effect in the P-population, and no other inhibitory types at 50\% contrast (Fig. \ref{fig:V1_3}) implying that the E-population is P-stabilized.
-Future work motivated by our analysis here, may uncover a relationship between the neuron-type mediating stability and the factors governing circuit variability.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
-In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}A).
-Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
-Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
-We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig.
-\ref{fig:SC}B).
-The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$.
-The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.
-The dynamics evolve with timescale $\tau= 90\text{ms}$ governed by connectivity weights $W$
-\begin{equation}
-\begin{split}
-\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
-\mathbf{x} &= \phi(\mathbf{u})
-\end{split}
-\end{equation}
-with white noise of variance $0.2^2$.
-The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching behavior.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
-\end{center}
-\caption{\footnotesize 
-\textbf{A}. Rapid task switching behavioral paradigm (see text). 
-\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
-Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
-\textbf{C}. The EPI posterior distribution of rapid task switching networks.  
-Red and purple stars ($\mathbf{z}_1$ and $\mathbf{z}_2$) indicate different connectivity regimes with different sensitivity vectors $\mathbf{v}_1$ and $\mathbf{v}_2$.
-(Middle-left) Posterior predictive distribution of task accuracies.
-(Bottom-left) Task accuracy along dimensions of sensitivity in each connectivity regime.
-\textbf{D}. Means (solid) and standard deviations (shaded) of each population across random simulated trials.
-Top plots show Pro (top) and Anti (bottom) responses for connectivity $\mathbf{z}_1$.
-Bottom rows show the same $\mathbf{z}_2$.
-\textbf{E}. The EPI posterior predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
-\textbf{F}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
-\textbf{G}. Accuracy with delay period inactivation along each connectivity regime's dimension of sensitivity.
-}
-\label{fig:SC}
-\end{figure}
-
-We define rapid task switching behavior as accurate execution of each task.
-Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
-We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
-\begin{equation}\label{eq:SC_EP}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 7.5\%^2 \\ 7.5\%^2  \end{bmatrix}
-\end{split}
-\end{equation}
-A variance of $7.5\%^2$  in each task will confer a posterior producing performances ranging from about $60\%-90\%$, allowing us to examine the properties of connectivity that yield better performance in each task.
-Notably, this is our first example using EPI to condition on multiple emergent property statistics ($|f(\mathbf{x}; \mathbf{z})| = 2$).
-
-The EPI inferred parameters  (Fig.
-\ref{fig:SC}C) generate a distribution of task accuracies (Fig.
-\ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
-The nonlinear patterns of connectivity that govern each task accuracy (Fig. \ref{fig:SC1}A-B) are not fully captured by linear prediction (Fig. \ref{fig:SC1}C).
-For example, the patterns in connectivity increasing Pro accuracy change dramatically after crossing a threshold of $sW$ (Fig. \ref{fig:SC1}A $sW$-$hW$ marginal).
-Not only has EPI captured this complex nonlinear posterior, it offers probabilistic tools for understanding the different regimes of model behavior.
-
-To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
-\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
-Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
-We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
-The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
-\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
-\ref{fig:SC}D, bottom-left, \ref{fig:SC2}).
-In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy tracks monotonic increases in Pro accuracy.
-These responses make intuitive sense, recognizing that $\mathbf{v}_1$ (unlike $\mathbf{v}_2$) points strongly in the direction of connectivity eigenvalue $\lambda_{\text{diag}}$, which is strongly anticorrelated with $p_A$ (Fig.
-\ref{fig:SC3}, \ref{fig:SC4}, see Section \ref{methods_SC}).
-
-In agreement with experimental results from Duan et al., we found optogenetic inactivation during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
-\ref{fig:SC}E)).
-This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
-Similarities across Pro and Anti trials in choice period responses following delay period inactivation (Fig.
-\ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
-\ref{fig:SC}F).
-The strong anticorrelation between $p_P$ and $p_{A, opto}$ across posterior connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
-Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
-\ref{fig:SC}F).
-
-In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This posterior revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
-EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, the model is well-chosen in that regard.
-Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
-
-\subsection{EPI scales well to high-dimensional parameter spaces} \label{results_LRRNN}
+The major scientific value of EPI is in the rich, queryable structure of these deep probability distributions.
+The probabilities of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ are the densities of these parameters in the distribution producing the emergent property.
+The greatest probabilities (the modes) indicate prototypical parameter configurations, and the manner in which probabilities change away from the modes shows how different parameter combinations preserve or diminish the emergent property.
+The dimensions of greatest sensitivity (e.g.
+Fig.
+1E solid arrow) or degeneracy (e.g.
+Fig.
+1E dashed arrow) can be measured directly from the second order derivative of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ called the ``Hessian."
+Around the mode, eigenvalues of the Hessian are negative; probabilities decrease locally in all directions away from the mode.
+The eigenvector with most negative eigenvalue is the parameter combination causing probability to decrease the fastest, making it the most sensitive dimension.
+Likewise, the flattest eigenvector, corresponding to the least negative eigenvalue, points in the most degenerate dimension.
+Once an EPI distribution has been inferred, this second order derivative requires trivial computation (when correct architecture class is chosen, see Section \ref{methods_NF}).
+
+In the following sections, we showcase the versatility of EPI for scientific analysis on three neural circuit models across ranges of biological realism, neural system function, and network scale.
+First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of RNN connectivities that exhibit amplified, yet stable responses.
+Also in this RNN example, we emphasize that EPI is the only technique that controls the predictions made by the inferred parameter distribution.
+Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how to gain insight by comparing multiple inferred distributions.
+Finally, we used EPI to capture subtle parametric structure allowing the mechanistic characterization of multiple parametric regimes of superior colliculus activity in a model of task switching \cite{duan2018collicular}.
+This work is the first to produce this level of theoretical insight via the quantification and examination of the intricate structure captured by deep probability distributions.
+
+\subsection{Scaling inference of RNN connectivity with EPI} \label{results_LRRNN}
 Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinsically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
@@
  \end{equation}
-where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
 We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   that produce stable amplification.
@@
 We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
-SMC-ABC is a rejection sampling approach that SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section Related Methods).
+SMC-ABC is a rejection sampling approach that SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
 Unlike EPI, these statistical inference techniques do not control the mean or variance of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
@@
 \ref{fig:RNN2}).
-To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.25 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
+To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
 
@@
 Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
-\ref{fig:RNN2}), and tested more aggressive SNPE hyperparameterizations when SNPE failed to converge (Fig.
+\ref{fig:RNN1}), and tested more aggressive SNPE hyperparameterizations when SNPE failed to converge (Fig.
 \ref{fig:RNN3}).
+From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both criteria in high-dimensions.
 
@@
 When using SNPE the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
-
-From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC.
-Another benefit over SMC-ABC is that EPI and SNPE offer fast additional samples after optimization.
-EPI outperforms SNPE in high dimensions by leveraging gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration.
-Through this example, we've shown that EPI can be used for insight into RNNs with respect to their theoretical properties, and that when $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z})$ is tractable, EPI can be used to efficiently infer high dimensional parameter distributions in mechanistic models of neural computation.
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
+
+Through this example, we have shown that EPI can be used for well-controlled insight into RNNs with respect to their theoretical properties.
+EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration.
+This agrees with recent speculation that such gradient information could improve the efficiency of LFI techniques \cite{cranmer2020frontier}.
+While scaling to high dimensions is important, we show in the next two sections how insight can be gained by inspecting structure in lower dimensional parameter distributions.
 
@@
 \cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{EPI reveals how neuron-type specific noise governs variability in the stochastic stabilized supralinear network} \label{results_V1}
+
+\begin{figure}
+\vspace{-1cm}
+\begin{center}
+\includegraphics[scale=0.8]{figures/fig2/fig2.pdf}
+\end{center}
+\vspace{-1cm}
+\caption{\small Emergent property inference in the stochastic stabilized supralinear network (SSSN)
+\textbf{A}.
+Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).
+Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
+Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
+Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
+\textbf{B}. Steady-state responses of the SSN model (deterministic, $\bm{\sigma}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
+\textbf{C}.
+Transient network responses of the SSSN model at 50 \% contrast.
+(Left) Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise realization.
+(Right) Mean (solid line) and standard deviation (shading) of responses.
+\textbf{D}. EPI distribution of noise parameters $\mathbf{z}$ conditioned on E-population variability.
+The EPI predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
+\textbf{E}.
+(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+Each black dot shows the mode at each $\sigma_P$.
+The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
+\textbf{F}.
+The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+}
+ \label{fig:V1}
+\end{figure}
+
+Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
+However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
+%Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
+While research has shown that V1 only shares specific dimensions of neuronal variability with downstream areas \cite{semedo2019cortical}, the role played by recurrent dynamics and the connectivity across neuron-type populations is not understood.
+Here, in a model of V1 with biologically realistic connectivity, we use EPI to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
+
+We considered response variability of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity estimated from post-synaptic potential and connectivity rate measurements (see Section \ref{methods_V1}).
+Each population also experiences an external input $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
+To build on previous work, we model visual contrast-dependent input to the E- and P-populations $\mathbf{h} = \mathbf{b} + c\mathbf{h}_c$.
+There is also an additive noisy input $\bm{\epsilon}$ parameterized by variances for each neuron type population $\mathbf{z} = \bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
+This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ then the population rate, allowing fluctuations around a stimulus-dependent steady-state
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}).
+\end{equation}
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have multiply inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
+Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
+Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure} (Fig.
+\ref{fig:V1}B), and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability at 50\% contrast.
+
+We quantify different levels $y$ of E-population variability with the emergent property
+\begin{equation}\label{eq:EP_V1}
+\begin{split}
+\mathcal{X}(y) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ y \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2,
+\end{split}
+\end{equation}
+where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
+
+We ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
+From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
+\ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
+Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions (Fig. \ref{fig:V1}D, bottom-right).
+Together, these observations imply a curved path of parametric degeneracy with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
+The dimensions of sensitivity conferred by EPI and this plain visual structure suggest a quadratic relationship in the emergent property statistic $s_E(\mathbf{x}; \mathbf{z})$ and parameters $\mathbf{z}$, which is preserved at a greater level of variability $\mathcal{X}(10\text{ Hz})$ (Fig.
+\ref{fig:V1}E).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p=0$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
+\ref{fig:V1_3}, $p=.402$).
+The strong compensatory influence of the E- and P-population input variability on excitatory varability is intriguing, since this circuit exhibited a paradoxical effect in the P-population (and no other inhibitory types) at 50\% contrast (Fig.
+\ref{fig:V1_3}) meaning that the E-population is P-stabilized.
+Future research may uncover a link between the populationsof stabilizations and compensatory interactions governing excitatory variability.
+
+By defining a clear theoretical question and executing the EPI optimization, we used EPI to reveal the quadratic relationship between $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ .
+While this property is ultimately derivable, we show that with each additional neuron-type population, the formula becomes quite unruly and likely escapes comprehensible analysis in our case (see Section \ref{methods_V1_complexity}).
+This emphasizes the need for streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes prohibitive.
+
+\textbf{Methods:} \\
+Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
+It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity allow this ability.
+In this experimental setup, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
+\ref{fig:SC}A).
+Neural recordings in superior colliculus (SC) exhibited two populations of neurons that represented task context (Pro or Anti).
+Furthermore, Pro/Anti neurons in each hemisphere were strongly correlated with the animal's decision \cite{duan2018collicular}.
+These results motivated a model of SC that is a four-population dynamical system with functionally-defined neuron-types.
+Here, our goal is to understand how connectivity in this circuit model governs the ability to switch tasks rapidly.
+
+In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$.
+The connectivity of these populations is parameterized by self- $sW$, vertical- $vW$, diagonal- $dW$ and horizontal-connections $hW$  (Fig. \ref{fig:SC}B).
+The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
+Model responses are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$
+\begin{equation}
+\begin{split}
+\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
+\mathbf{x} &= \phi(\mathbf{u})
+\end{split}
+\end{equation}
+where $\tau= 90\text{ms}$ and there is white noise of variance $0.2^2$.
+The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.  
+Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
+\end{center}
+\caption{\footnotesize 
+\textbf{A}. Rapid task switching behavioral paradigm (see text). 
+\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
+Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
+\textbf{C}. The EPI inferred distribution of rapid task switching networks.  
+Red and purple stars ($\mathbf{z}_1$ and $\mathbf{z}_2$) indicate different connectivity regimes with different sensitivity vectors $\mathbf{v}_1$ and $\mathbf{v}_2$.
+(Middle-left) EPI predictive distribution of task accuracies.
+(Bottom-left) Task accuracy along dimensions of sensitivity in each connectivity regime.
+\textbf{D}. Means (solid) and standard deviations (shaded) of each population across random simulated trials.
+Top plots show Pro (top) and Anti (bottom) responses for connectivity $\mathbf{z}_1$.
+Bottom rows show the same $\mathbf{z}_2$.
+\textbf{E}. The EPI distribution predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
+\textbf{F}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
+\textbf{G}. Accuracy with delay period inactivation along each connectivity regime's dimension of sensitivity.
+}
+\label{fig:SC}
+\end{figure}
+
+We define the computation of rapid task switching as accurate execution of each task.
+Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
+We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
+\begin{equation}\label{eq:SC_EP}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 7.5\%^2 \\ 7.5\%^2  \end{bmatrix}
+\end{split}
+\end{equation}
+
+The EPI inferred distribution  (Fig.
+\ref{fig:SC}C) produces task accuracies (Fig.
+\ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
+The patterns of connectivity that govern each task accuracy are nonlinear (Fig. \ref{fig:SC1}A-B); they are not captured well by linear prediction (Fig. \ref{fig:SC1}C).
+For example, the patterns in connectivity increasing Pro accuracy change dramatically after crossing a threshold of $sW$ (Fig. \ref{fig:SC1}A $sW$-$hW$ marginal distribution).
+Not only has EPI captured this intricate, nonlinear posterior, it offers probabilistic tools for understanding the different regimes of model behavior.
+
+To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
+\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
+Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
+We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
+The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
+\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
+\ref{fig:SC}D, bottom-left, Fig.
+\ref{fig:SC2}).
+In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy matches monotonic increases in Pro accuracy.
+
+To understand why these distinct connectivity regimes have different failure modes, we can analyze the properties of connectivity in each regime.
+By taking the eigendecomposition of $W$, we can quantify how strongly different modes of processing are reflected in the connectivity matrix (see Section \ref{methods_SC}).  
+\textbf{Note SB: I need to alter these plots a tad to confirm and support the mechanistic explanation.}
+
+During the delay period of this task, the circuit must prepare to execute the correct task based on the cue input.
+Experimental results from Duan et al.
+found that optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
+\ref{fig:SC}E)).
+All network connectivities inferred by EPI exhibited this same effect, when network activities were silenced during the delay period (see Section \ref{methods_SC}).
+Notably, EPI inferred connectivities were only conditioned upon the emergent property of rapid task switching, not on Anti task failure during delay period silencing.
+
+Similarities across Pro and Anti trials in choice period responses following delay period inactivation (Fig.
+\ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
+\ref{fig:SC}F).
+The strong anticorrelation between $p_P$ and $p_{A, opto}$ across EPI inferred connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
+Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
+\ref{fig:SC}F).
+Thus, the altered state caused by delay period silencing makes the $p_P$ connectivity better than the $p_A$ down connectivity.
+
+In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
+This EPI distribution revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
+EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, the model is well-chosen in that regard.
+Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
 
@@
 
-\subsubsection{Primary visual cortex}\label{methods_V1}
-In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
-\begin{equation}
-    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}),
-\end{equation}
-where the noise is an Ornstein-Uhlenbeck process $\bm{\epsilon} \sim OU(\tau_{\text{noise}}, \bm{\sigma})$
-\begin{equation}
-\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\tilde{\sigma}_\alpha dB
-\end{equation}
-with $\tau_{\text{noise}} = 5\text{ms} > \tau = 1\text{ms}$.
-The noisy process is parameterized as
-\begin{equation}
-\tilde{\sigma}_\alpha = \sigma_\alpha \sqrt{1 + \frac{\tau}{\tau_{\text{noise}}}},
-\end{equation}
-so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
-As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
-Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
-\end{center}
-\caption{\small (V1 1)
-(Left) Simulations for small increases in neuron-type population input.
-Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
-(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
- }
- \label{fig:V1_1}
-\end{figure}
-
-\begin{figure}
-\caption{\small (V1 2)
-Posterior predictive distributions of the sum of squares of each pair of noise parameters.
- }
- \label{fig:V1_2}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
-\end{center}
-\end{figure}
-
-\begin{equation}
-W_{\text{fit}} =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
-W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
-W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
-W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
- \begin{bmatrix} 2.18 & -1.19 & -.594 & -.229 \\
- 1.66 & -.651 & -.680 & -.242 \\
- .895 & -5.22 \times 10^{-3} & -1.51 \times 10^{-4}   & -.761 \\
- 3.34 &  -2.31 & -.254  & -2.52 \times 10^{-4} \\
- \end{bmatrix},
-\end{equation}
-
-\begin{equation}
-\mathbf{h}_{b,\text{fit}} =
- \begin{bmatrix} .416 \\ .429 \\ .491 \\ .486 \end{bmatrix} ,
-\end{equation} 
-and
-\begin{equation} 
-\mathbf{h}_{c,\text{fit}} = 
-\begin{bmatrix} .359 \\ .403 \\ 0 \\ 0 \end{bmatrix}.
-\end{equation}
-
-To obtain rates on a realistic scale (100-fold greater), we map these fitted parameters to an equivalence class
-
-\begin{figure}[h]
-\caption{\small (V1 3)
-EPI posterior for $\mathcal{X}(10\text{ Hz})$.
- }
- \label{fig:V1_3}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
-\end{center}
-\end{figure}
-
-\begin{equation}
-W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
-W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
-W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
-W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
- \begin{bmatrix} .218 & -.119 & -.0594 & -.0229 \\
- .166 & -.0651 & -.068 & -.0242 \\
- .0895 & -5.22 \times 10^{-4} & -1.51 \times 10^{-5}   & -.0761 \\
- .334 &  -.231 & -.0254  & -2.52 \times 10^{-5} \\
- \end{bmatrix},
-\end{equation}
-
-\begin{equation}
-\mathbf{h}_b =  \begin{bmatrix} h_{b,E} \\ h_{b,P} \\ h_{b,S} \\ h_{b,V} \end{bmatrix} =
- \begin{bmatrix} 4.16 \\ 4.29 \\ 4.91 \\ 4.86 \end{bmatrix} ,
-\end{equation} 
-and
-\begin{equation} 
-\mathbf{h}_c = \begin{bmatrix} h_{c,E} \\ h_{c,P} \\ h_{c,S} \\ h_{c,V} \end{bmatrix} = 
-\begin{bmatrix} 3.59 \\ 4.03 \\ 0 \\ 0 \end{bmatrix}.
-\end{equation}
-
-Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{ Hz}, 25\text{ Hz}\right]$.
-Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
-\begin{equation}
-s_E(\mathbf{x}; \mathbf{z}) =\mathbb{E}_{x}\left[\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]}\right]
-\end{equation}
-
-For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
-The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
-EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
-The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
-
-In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
-Specifically, we calculated
-\begin{equation}
-\begin{split}
-\mathbf{z}^*(\sigma_{P,\text{fixed}}) = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } \sigma_P = \sigma_{P,\text{fixed}}\
-\end{split}
-\end{equation}
-At each mode $\mathbf{z}^*$, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
-
-\subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
-TODO Agostina and I are putting this together now.
-
-\subsubsection{Superior colliculus}\label{methods_SC}
-In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
-They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
-Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
-\begin{equation}
-x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
-\end{equation}
-where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
-During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
-\begin{equation}
-x_\alpha = (1-\gamma)\phi(u_\alpha).
-\end{equation}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.9]{figures/figSC1/figSC1.pdf}
-\end{center}
-\caption{\small (SC1): \textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
-\textbf{B}. Same as A colored by Anti task accuracy.  
-\textbf{C}. Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p-value.
-}
-\label{fig:SC1}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
-\end{center}
-\caption{\small (SC2):  
-\textbf{A}.
-Simulations in network regime $\mathbf{z}_1$ (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
-\textbf{B}. Same as A for network regime $\mathbf{z}_2$.
-}
-\label{fig:SC2}
-\end{figure}
-
-We order the neural populations of $x$ and $u$ in the following manner
-\begin{equation}
-\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
-\end{equation}
-
-which evolve according to
-\begin{equation}
-\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
-\end{equation}
-with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.  
-The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
-\begin{equation}
-W = \begin{bmatrix} sW & vW & hW & dW \\ vW & sW & dW  & hW \\ hW & dW & sW & vW \\ dW  & hW & vW  & sW \end{bmatrix}.
-\end{equation}
-
-The circuit receives four different inputs throughout each trial, which has a total length of 1.8s.
-\begin{equation}
-\mathbf{h} = \mathbf{h}_{\text{constant}}  + \mathbf{h}_{\text{P,bias}} + \mathbf{h}_{\text{rule}} +  \mathbf{h}_{\text{choice-period}} + \mathbf{h}_{\text{light}}.
-\end{equation}
-
-There is a constant input to every population,
-\begin{equation}\mathbf{h}_{\text{constant}} = I_{\text{constant}} [1, 1, 1, 1]\top,
-\end{equation}
- a bias to the Pro populations
-\begin{equation}\mathbf{h}_{\text{P,bias}} = I_{\text{P,bias}} [1, 0, 1, 0]\top,
-\end{equation}
-rule-based input depending on the condition
-\begin{equation}\mathbf{h}_{\text{P,rule}}(t) = \begin{cases}
-                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases}
-\end{equation}
-\begin{equation} \mathbf{h}_{\text{A,rule}}(t) = \begin{cases}
-                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases},
-\end{equation}
-a choice-period input
-\begin{equation} \mathbf{h}_{\text{choice}}(t) = \begin{cases}
-                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
-                            0,              & \text{otherwise}
-                         \end{cases},
-\end{equation}
-and an input to the right or left-side depending on where the light stimulus is delivered    
-\begin{equation}  \mathbf{h}_{\text{light}}(t) = \begin{cases}
-                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Left} \\
-                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Right} \\
-                            0,              & \text{otherwise}
-                         \end{cases}.
-\end{equation}
-The input parameterization was fixed to $I_{\text{constant}} = 0.75$, $I_{\text{P,bias}} = 0.5 $, $I_{\text{P,rule}} = 0.6$,  $I_{\text{A,rule}} = 0.6$,  $I_{\text{choice}} = 0.25$,  and $I_{\text{light}} = 0.5$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.9]{figures/figSC3/figSC3.pdf}
-\end{center}
-\caption{\small (SC3):  
-\textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
-\textbf{B}. Eigenvalues of connectivities of EPI distribution versus task accuracies.
-}
-\label{fig:SC3}
-\end{figure}
-
-The accuracies of $p_P$ and $p_A$ are calculated as
-\begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
-\end{equation}
-and 
-\begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
-\end{equation}
-given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function, and the accuracy is averaged over 200 independent trials.  The Heaviside step function is approximated as
-\begin{equation}
-\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
-\end{equation}
-where $\beta = 100$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
-\end{center}
-\caption{\small (SC4):  \textbf{A}.
-Pairplots of eigenvalues of connectivity matrices in EPI distribution colored by Pro task accuracy.
-Red and purple stars and arrows correspond to eigenvalues and sensitivity directions $\mathbf{z}_1$, $\mathbf{z}_2$, $\mathbf{v}_1$, and $\mathbf{v}_2$.
-\textbf{B}.
-Same colored by Anti task accuracy.
-}
-\label{fig:SC4}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC5/figSC5.pdf}
-\end{center}
-\caption{\small (SC5): 
-\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Connectivity eigenvalues versus the task error induced by delay period inactivation.
-\textbf{C}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
-}
-\label{fig:SC5}
-\end{figure}
-
-Writing the EPI posterior as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 7.5\%^2 \\ 7.5\%^2 \end{bmatrix}.
-\end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after six epochs (Fig. \ref{fig:SC5}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figSC6/figSC6.pdf}
-\end{center}
-\caption{\small (SC6): 
-\textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
-}
-\label{fig:SC5}
-\end{figure}
-
-For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
-The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 25$ at greatest entropy across random seeds.
-
-To make sense of this inferred distribution, we identified two modes used to represent the two regimes of connectivity in this posterior:
-\begin{equation}
-\begin{split}
-\mathbf{z}_1 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. }  hw = -1.25\text{, }  sW > 0
-\end{split}
-\end{equation}
-and 
-\begin{equation}
-\begin{split}
-\mathbf{z}_2 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } hw = -1.25\text{, } sW < 0
-\end{split}
-\end{equation}
-
-To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
-\ref{fig:SC3}A).
-These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-We found significant trends across the EPI posterior connectivities: the eigenvalues $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ were correlated with $p_P$, while $\lambda_{\text{all}}$ was anticorrelated with $p_P$.
-$\lambda_{\text{all}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ were all significantly anticorrelated with $p_A$.
-
-Under this decomposition, we can re-visualize the posterior in eigenvalue space (Fig. \ref{fig:SC4}).
-Furthermore, we can project the dimensions of sensitivity into eigenvalue space as well, giving us a more intuitive sense of how connectivity affects computation in each regime.
-We see that sensitivity dimensions $\mathbf{v}_1$ and $\mathbf{v}_2$, which cause $p_P$ to increase and a regime dependent change in $p_A$, both point in the direction of increasing $\lambda_{\text{side}}$ and decreasing $\lambda_{\text{task}}$.
-These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
-As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
-A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
-
-During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.
-We found that $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ were significantly anticorrelated with Anti error during delay period inactivation.
-Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.
-
 \subsubsection{Scaling EPI for stable amplification in RNNs}\label{methods_RNN}
-We examine the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
-Rank-2 RNN connectivity is modeled as $W = UV^\top$, where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+We examined the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
+Rank-2 RNN connectivity was modeled as $W = UV^\top$, where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
 This RNN model has dynamics
@@
 \end{equation}
-In this analysis, we infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produce stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section Related Methods).
+In this analysis, we inferred connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produced stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section Related Methods).
 
@@
 These two conditions are necessary and sufficient for stable amplification in RNNs \cite{bondanelli2020coding}.
-We define the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions.
+We defined the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions.
 To complete the emergent property definition, we chose variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
@@
 Gradients of maximum eigenvalues of Hermitian matrices like $W^s$ are available with modern automatic differentiation tools.
-To differentiate through the $\text{real}(\lambda_1)$, we solve the following equation for eigenvalues of rank-2 matrices using the rank reduced matrix $W^r = V^{\top} U$
+To differentiate through the $\text{real}(\lambda_1)$, we solved the following equation for eigenvalues of rank-2 matrices using the rank reduced matrix $W^r = V^{\top} U$
 \begin{equation}
@@
 SNPE, the next LFI approach in our comparison, is far more similar to EPI.
-Like EPI, SNPE models parameters in mechanistic models with deep probability distributions, yet their learning algorithms are categorically different.
+Like EPI, SNPE treats parameters in mechanistic models with deep probability distributions, yet the two learning algorithms are categorically different.
 SNPE uses a two-network architecture to approximate the posterior distribution of the model conditioned on observed data $\mathbf{x}_0$.
@@
 As in SMC-ABC, we ran SNPE with $\mathbf{x}_0 = \mu$.
-All SNPE optimizations were run for a limit of 1.5 days on a Tesla V100 GPU, or until two consecutive rounds resulted in a validation log probability lower than the maximum observed.
+All SNPE optimizations were run for a limit of 1.5 days on a Tesla V100 GPU, or until two consecutive rounds resulted in a validation log probability lower than the maximum observed for that random seed.
 
@@
 \ref{fig:RNN2}A).
-For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater uncertainty (Fig. \ref{fig:RNN2}B top, orange).
+For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater randomness (Fig. \ref{fig:RNN2}B top, orange).
 At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Fig. \ref{fig:RNN2}B top).
@@
 \end{equation}
-is less than 0.25.
+is less than 0.5.
 We consider SMC-ABC to have converged once the population produces samples within the $\epsilon = 0.5$ ball ensuring stable amplification.
@@
 
+\subsubsection{Primary visual cortex}\label{methods_V1}
+In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}),
+\end{equation}
+where the noise is an Ornstein-Uhlenbeck process $\bm{\epsilon} \sim OU(\tau_{\text{noise}}, \bm{\sigma})$
+\begin{equation}
+\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\tilde{\sigma}_\alpha dB
+\end{equation}
+with $\tau_{\text{noise}} = 5\text{ms} > \tau = 1\text{ms}$.
+The noisy process is parameterized as
+\begin{equation}
+\tilde{\sigma}_\alpha = \sigma_\alpha \sqrt{1 + \frac{\tau}{\tau_{\text{noise}}}},
+\end{equation}
+so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
+As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
+Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
+\end{center}
+\caption{\small (V1 1)
+(Left) Simulations for small increases in neuron-type population input.
+Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
+(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
+ }
+ \label{fig:V1_1}
+\end{figure}
+
+\begin{figure}
+\caption{\small (V1 2)
+Posterior predictive distributions of the sum of squares of each pair of noise parameters.
+ }
+ \label{fig:V1_2}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
+\end{center}
+\end{figure}
+
+\begin{equation}
+W_{\text{fit}} =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
+W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
+W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
+W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
+ \begin{bmatrix} 2.18 & -1.19 & -.594 & -.229 \\
+ 1.66 & -.651 & -.680 & -.242 \\
+ .895 & -5.22 \times 10^{-3} & -1.51 \times 10^{-4}   & -.761 \\
+ 3.34 &  -2.31 & -.254  & -2.52 \times 10^{-4} \\
+ \end{bmatrix},
+\end{equation}
+
+\begin{equation}
+\mathbf{h}_{b,\text{fit}} =
+ \begin{bmatrix} .416 \\ .429 \\ .491 \\ .486 \end{bmatrix} ,
+\end{equation} 
+and
+\begin{equation} 
+\mathbf{h}_{c,\text{fit}} = 
+\begin{bmatrix} .359 \\ .403 \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+To obtain rates on a realistic scale (100-fold greater), we map these fitted parameters to an equivalence class
+
+\begin{figure}[h]
+\caption{\small (V1 3)
+EPI posterior for $\mathcal{X}(10\text{ Hz})$.
+ }
+ \label{fig:V1_3}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
+\end{center}
+\end{figure}
+
+\begin{equation}
+W =  \begin{bmatrix} W_{EE} & W_{EP} & W_{ES} & W_{EV} \\
+W_{PE} & W_{PP} & W_{PS} & W_{PV} \\
+W_{SE} & W_{SP} & W_{SS} & W_{SV} \\
+W_{VE} & W_{VP} & W_{VS} & W_{VV}  \end{bmatrix} = 
+ \begin{bmatrix} .218 & -.119 & -.0594 & -.0229 \\
+ .166 & -.0651 & -.068 & -.0242 \\
+ .0895 & -5.22 \times 10^{-4} & -1.51 \times 10^{-5}   & -.0761 \\
+ .334 &  -.231 & -.0254  & -2.52 \times 10^{-5} \\
+ \end{bmatrix},
+\end{equation}
+
+\begin{equation}
+\mathbf{h}_b =  \begin{bmatrix} h_{b,E} \\ h_{b,P} \\ h_{b,S} \\ h_{b,V} \end{bmatrix} =
+ \begin{bmatrix} 4.16 \\ 4.29 \\ 4.91 \\ 4.86 \end{bmatrix} ,
+\end{equation} 
+and
+\begin{equation} 
+\mathbf{h}_c = \begin{bmatrix} h_{c,E} \\ h_{c,P} \\ h_{c,S} \\ h_{c,V} \end{bmatrix} = 
+\begin{bmatrix} 3.59 \\ 4.03 \\ 0 \\ 0 \end{bmatrix}.
+\end{equation}
+
+Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{ Hz}, 25\text{ Hz}\right]$.
+Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
+\begin{equation}
+s_E(\mathbf{x}; \mathbf{z}) =\mathbb{E}_{x}\left[\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]}\right]
+\end{equation}
+
+For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
+The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
+EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
+The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
+
+In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
+Specifically, we calculated
+\begin{equation}
+\begin{split}
+\mathbf{z}^*(\sigma_{P,\text{fixed}}) = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. } \sigma_P = \sigma_{P,\text{fixed}}\
+\end{split}
+\end{equation}
+At each mode $\mathbf{z}^*$, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
+
+\begin{figure}[h]
+\caption{\small (V1 4)
+Optimization for V1
+ }
+ \label{fig:V1_4}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_4/figV1_4.pdf}
+\end{center}
+\end{figure}
+
+\subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
+TODO Agostina and I are putting this together now.
+
+\subsubsection{Superior colliculus}\label{methods_SC}
+In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
+They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
+Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
+\begin{equation}
+x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
+\end{equation}
+where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
+During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
+\begin{equation}
+x_\alpha = (1-\gamma)\phi(u_\alpha).
+\end{equation}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.9]{figures/figSC1/figSC1.pdf}
+\end{center}
+\caption{\small (SC1): \textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
+\textbf{B}. Same as A colored by Anti task accuracy.  
+\textbf{C}. Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p-value.
+}
+\label{fig:SC1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
+\end{center}
+\caption{\small (SC2):  
+\textbf{A}.
+Simulations in network regime $\mathbf{z}_1$ (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
+\textbf{B}. Same as A for network regime $\mathbf{z}_2$.
+}
+\label{fig:SC2}
+\end{figure}
+
+We order the neural populations of $x$ and $u$ in the following manner
+\begin{equation}
+\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
+\end{equation}
+
+which evolve according to
+\begin{equation}
+\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
+\end{equation}
+with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.  
+The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
+\begin{equation}
+W = \begin{bmatrix} sW & vW & hW & dW \\ vW & sW & dW  & hW \\ hW & dW & sW & vW \\ dW  & hW & vW  & sW \end{bmatrix}.
+\end{equation}
+
+The circuit receives four different inputs throughout each trial, which has a total length of 1.8s.
+\begin{equation}
+\mathbf{h} = \mathbf{h}_{\text{constant}}  + \mathbf{h}_{\text{P,bias}} + \mathbf{h}_{\text{rule}} +  \mathbf{h}_{\text{choice-period}} + \mathbf{h}_{\text{light}}.
+\end{equation}
+
+There is a constant input to every population,
+\begin{equation}\mathbf{h}_{\text{constant}} = I_{\text{constant}} [1, 1, 1, 1]\top,
+\end{equation}
+ a bias to the Pro populations
+\begin{equation}\mathbf{h}_{\text{P,bias}} = I_{\text{P,bias}} [1, 0, 1, 0]\top,
+\end{equation}
+rule-based input depending on the condition
+\begin{equation}\mathbf{h}_{\text{P,rule}}(t) = \begin{cases}
+                           I_{\text{P,rule}} [1, 0, 1, 0]^\top,& \text{if } t\leq 1.2s \\
+                            0,              & \text{otherwise}
+                         \end{cases}
+\end{equation}
+\begin{equation} \mathbf{h}_{\text{A,rule}}(t) = \begin{cases}
+                           I_{\text{A,rule}} [0, 1, 0, 1]^\top,& \text{if } t\leq 1.2s \\
+                            0,              & \text{otherwise}
+                         \end{cases},
+\end{equation}
+a choice-period input
+\begin{equation} \mathbf{h}_{\text{choice}}(t) = \begin{cases}
+                           I_{\text{choice}} [1, 1, 1, 1]^\top,& \text{if } t > 1.2s \\
+                            0,              & \text{otherwise}
+                         \end{cases},
+\end{equation}
+and an input to the right or left-side depending on where the light stimulus is delivered    
+\begin{equation}  \mathbf{h}_{\text{light}}(t) = \begin{cases}
+                           I_{\text{light}} [1, 1, 0, 0]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Left} \\
+                           I_{\text{light}} [0, 0, 1, 1]^\top,& \text{if } 1.2s < t < 1.5s \text{ and Right} \\
+                            0,              & \text{otherwise}
+                         \end{cases}.
+\end{equation}
+The input parameterization was fixed to $I_{\text{constant}} = 0.75$, $I_{\text{P,bias}} = 0.5 $, $I_{\text{P,rule}} = 0.6$,  $I_{\text{A,rule}} = 0.6$,  $I_{\text{choice}} = 0.25$,  and $I_{\text{light}} = 0.5$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.9]{figures/figSC3/figSC3.pdf}
+\end{center}
+\caption{\small (SC3):  
+\textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
+\textbf{B}. Eigenvalues of connectivities of EPI distribution versus task accuracies.
+}
+\label{fig:SC3}
+\end{figure}
+
+The accuracies of $p_P$ and $p_A$ are calculated as
+\begin{equation}
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+\end{equation}
+and 
+\begin{equation}
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+\end{equation}
+given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function, and the accuracy is averaged over 200 independent trials.  The Heaviside step function is approximated as
+\begin{equation}
+\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
+\end{equation}
+where $\beta = 100$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
+\end{center}
+\caption{\small (SC4):  \textbf{A}.
+Pairplots of eigenvalues of connectivity matrices in EPI distribution colored by Pro task accuracy.
+Red and purple stars and arrows correspond to eigenvalues and sensitivity directions $\mathbf{z}_1$, $\mathbf{z}_2$, $\mathbf{v}_1$, and $\mathbf{v}_2$.
+\textbf{B}.
+Same colored by Anti task accuracy.
+}
+\label{fig:SC4}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC5/figSC5.pdf}
+\end{center}
+\caption{\small (SC5): 
+\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
+\textbf{B}. Connectivity eigenvalues versus the task error induced by delay period inactivation.
+\textbf{C}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
+}
+\label{fig:SC5}
+\end{figure}
+
+Writing the EPI posterior as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 7.5\%^2 \\ 7.5\%^2 \end{bmatrix}.
+\end{equation}
+Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
+The optimization converged after six epochs (Fig. \ref{fig:SC5}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{figures/figSC6/figSC6.pdf}
+\end{center}
+\caption{\small (SC6): 
+\textbf{A}. Entropy throughout optimization. 
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+}
+\label{fig:SC5}
+\end{figure}
+
+For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
+The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
+We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
+The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 25$ at greatest entropy across random seeds.
+
+To make sense of this inferred distribution, we identified two modes used to represent the two regimes of connectivity in this posterior:
+\begin{equation}
+\begin{split}
+\mathbf{z}_1 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. }  hw = -1.25\text{, }  sW > 0
+\end{split}
+\end{equation}
+and 
+\begin{equation}
+\begin{split}
+\mathbf{z}_2 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
+&\text{s.t. } hw = -1.25\text{, } sW < 0
+\end{split}
+\end{equation}
+
+To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
+\ref{fig:SC3}A).
+These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
+We found significant trends across the EPI posterior connectivities: the eigenvalues $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ were correlated with $p_P$, while $\lambda_{\text{all}}$ was anticorrelated with $p_P$.
+$\lambda_{\text{all}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ were all significantly anticorrelated with $p_A$.
+
+Under this decomposition, we can re-visualize the posterior in eigenvalue space (Fig. \ref{fig:SC4}).
+Furthermore, we can project the dimensions of sensitivity into eigenvalue space as well, giving us a more intuitive sense of how connectivity affects computation in each regime.
+We see that sensitivity dimensions $\mathbf{v}_1$ and $\mathbf{v}_2$, which cause $p_P$ to increase and a regime dependent change in $p_A$, both point in the direction of increasing $\lambda_{\text{side}}$ and decreasing $\lambda_{\text{task}}$.
+These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
+As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
+A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
+
+During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.
+We found that $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ were significantly anticorrelated with Anti error during delay period inactivation.
+Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.
+
 \end{document}

### 2021-03-29 written/eLifeRevision/Bittner_eLife2020.tex
latest updates

@@
 This emphasizes the need for streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes prohibitive.
-
-\textbf{Methods:} \\
-Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
 
@@
 In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
+Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
 They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  

### 2021-03-31 neural_circuits/SC_Circuit_4.py
latest SC plots

@@ -4,6 +4,7 @@ import scipy
 from scipy.special import expit
 import matplotlib.pyplot as plt
 import os
+from epi.util import get_conditional_mode
 
 DTYPE = tf.float32
 
@@ -24,8 +25,7 @@ E_choice = 0.25
 E_light = 0.5
 
 # set constant parameters
-C = 2
-
+C = 2 
 theta = 0.05
 beta = 0.5
 tau = 0.09
@@ -111,7 +111,7 @@ def unwrap(z):
 
 
 def SC_acc(sW, vW, dW, hW):
-    N = 500
+    N = 200
     Wrow1 = tf.stack([sW, vW, dW, hW], axis=2)
     Wrow2 = tf.stack([vW, sW, hW, dW], axis=2)
     Wrow3 = tf.stack([dW, hW, sW, vW], axis=2)
@@ -175,7 +175,6 @@ C_opto = 4
 I_opto = tf.concat((I_LP, I_LA, I_LP, I_LA), axis=2)
 
 def SC_sim_opto(strength, period):
-    print('str', strength)
     eta = np.ones((T, 1, C_opto, 1, 1), dtype=np.float32)
     if period == 'delay':
         eta[np.logical_and(0.8 <= t, t <= 1.2), :, 2:, :, :] = strength
@@ -226,7 +225,7 @@ def SC_acc_opto(strength, period):
         eta[:, :, 2:, :, :] = strength
 
     def _SC_acc_opto(sW, vW, dW, hW):
-        N = 500
+        N = 200
         Wrow1 = tf.stack([sW, vW, dW, hW], axis=2)
         Wrow2 = tf.stack([vW, sW, hW, dW], axis=2)
         Wrow3 = tf.stack([dW, hW, sW, vW], axis=2)
@@ -275,16 +274,18 @@ def z_to_W(z):
     W = tf.stack([Wrow1, Wrow2, Wrow3, Wrow4], axis=1)
     return W
 
+MODES = np.array([[1.0, 1.0, 1.0, 1.0],   # all mode
+                  [-1.0, -1.0, 1.0, 1.0], # side mode
+                  [1.0, -1.0, -1.0, 1.0], # task mode
+                  [-1.0, 1.0, -1.0, 1.0]]) # diag mode
+
 def get_schur_eigs(W):
     # returns 
     T, Z = scipy.linalg.schur(W)
     b = Z.copy()
     b[b<0.0] = -1
     b[b>0.0] = 1
[…]

### 2021-03-31 written/eLifeRevision/Bittner_eLife2020.tex
section title change'
sc methods update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI reveals how neuron-type specific noise governs variability in the stochastic stabilized supralinear network} \label{results_V1}
+\subsection{EPI reveals how noisy input across neuron-types governs excitatory variability in a V1 model} \label{results_V1}
 
@@
 In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$.
-The connectivity of these populations is parameterized by self- $sW$, vertical- $vW$, diagonal- $dW$ and horizontal-connections $hW$  (Fig. \ref{fig:SC}B).
+The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
 The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
@@
 \subsubsection{Superior colliculus}\label{methods_SC}
-In the model of Duan et al \cite{duan2018collicular}, there are four total units: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
-Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
-They are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
+The ability to switch between two separate tasks throughout randomly interleaved trials, or ``rapid task switching," has been studied in rats, and midbrain superior colliculus (SC) has been show to play an important in this computation \cite{duan2015requirement}.
+Neural recordings in SC exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side), which led to the distinction of two functional classes: the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
+Given this evidence, Duan et al. proposed a model with four functionally-defined neuron-type populations: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
+We study how the connectivity of this neural circuit governs rapid task switching ability.
+
+The four populations of this model are denoted as left Pro (LP), left Anti (LA), right Pro (RP) and right Anti (RA).  
 Each unit has an activity ($x_\alpha$) and internal variable ($u_\alpha$) related by
 \begin{equation}
-x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right)
-\end{equation}
-where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity, respectively.
-During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
-\begin{equation}
-x_\alpha = (1-\gamma)\phi(u_\alpha).
-\end{equation}
+x_\alpha = \phi(u_\alpha) = \left(\frac{1}{2}\tanh\left(\frac{u_\alpha - a}{b}\right)+ \frac{1}{2} \right),
+\end{equation}
+where $\alpha \in \{LP, LA, RA, RP\}$, $a = 0.05$ and $b = 0.5$ control the position and shape of the nonlinearity.
+We order the neural populations of $x$ and $u$ in the following manner
+\begin{equation}
+\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
+\end{equation}
+which evolve according to
+\begin{equation}
+\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
+\end{equation}
+with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.
+These hyperparameter values are motivated by modeling choices and results from \cite{duan2018collicular}.
 
@@
 
-We order the neural populations of $x$ and $u$ in the following manner
-\begin{equation}
-\mathbf{x} = \begin{bmatrix} x_{LP} \\ x_{LA} \\ x_{RP} \\ x_{RA} \end{bmatrix} \hspace{2cm} \mathbf{u} = \begin{bmatrix} u_{LP} \\ u_{LA} \\ u_{RP} \\ u_{RA} \end{bmatrix},
-\end{equation}
-
-which evolve according to
-\begin{equation}
-\tau \frac{d\mathbf{u}}{dt} = -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B}.
-\end{equation}
-with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.  
-The weight matrix has 4 parameters $sW$, $vW$, $hW$, and $dW$:
+The weight matrix has 4 parameters for self $sW$, vertical $vW$, horizontal $hW$, and diagonal $dW$ connections:
 \begin{equation}
@@
 \end{equation}
+We study the role of parameters $\mathbf{z} = [sW, vW, hW, dW]^\top$ in rapid task switching.
 
@@
 \end{equation}
-
 There is a constant input to every population,
@@
 
-The accuracies of $p_P$ and $p_A$ are calculated as
-\begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+The accuracies of each task $p_P$ and $p_A$ are calculated as
+\begin{equation}
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x}}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
 \end{equation}
@@
 \begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x}}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
 \end{equation}
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
-\end{center}
-\caption{\small (SC4):  \textbf{A}.
-Pairplots of eigenvalues of connectivity matrices in EPI distribution colored by Pro task accuracy.
-Red and purple stars and arrows correspond to eigenvalues and sensitivity directions $\mathbf{z}_1$, $\mathbf{z}_2$, $\mathbf{v}_1$, and $\mathbf{v}_2$.
-\textbf{B}.
-Same colored by Anti task accuracy.
+%\begin{center}
+%\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
+%\end{center}
+\caption{\small (SC4):  \textbf{A}. Something showing differences in regimes.
 }
@@
 
-Writing the EPI posterior as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p(\mathbf{x}; \mathbf{z})_P \\ p(\mathbf{x}; \mathbf{z})_A \\ \left(p(\mathbf{x}; \mathbf{z})_P - 75\% \right)^2 \\ \left(p(\mathbf{x}; \mathbf{z})_A - 75\% \right)^2 \end{bmatrix},
+Writing the EPI distribution as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+\begin{equation} 
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \\ \left(p_P(\mathbf{x}; \mathbf{z}) - 75\% \right)^2 \\ \left(p_A(\mathbf{x}; \mathbf{z}) - 75\% \right)^2 \end{bmatrix},
 \end{equation}
@@
 
+\textbf{OPTO} \\
+During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
+\begin{equation}
+x_\alpha = (1-\gamma)\phi(u_\alpha).
+\end{equation}
+
 During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.

### 2021-03-31 neural_circuits/SC_Circuit_4.py
latest sc nb

@@ -396,7 +396,7 @@ def perturbed_acc_plots(facs, pPs, pAs, c_stars, fontsize=12, label="", figdir=N
                          np.mean(_ps, axis=0), 
                          np.std(_ps, axis=0)/np.sqrt(N_sim), 
                          c=c_stars[i], 
-                         lw=3)
+                         lw=2)
         ax.set_yticks(yticks)
         ax.set_yticklabels(["%d%%" % tick for tick in yticks], fontsize=fontsize)
         ax.set_xticks(xticks)

### 2021-03-31 written/eLifeRevision/Bittner_eLife2020.tex
more sc updates

@@
 \textbf{C}. The EPI inferred distribution of rapid task switching networks.  
-Red and purple stars ($\mathbf{z}_1$ and $\mathbf{z}_2$) indicate different connectivity regimes with different sensitivity vectors $\mathbf{v}_1$ and $\mathbf{v}_2$.
-(Middle-left) EPI predictive distribution of task accuracies.
-(Bottom-left) Task accuracy along dimensions of sensitivity in each connectivity regime.
-\textbf{D}. Means (solid) and standard deviations (shaded) of each population across random simulated trials.
-Top plots show Pro (top) and Anti (bottom) responses for connectivity $\mathbf{z}_1$.
-Bottom rows show the same $\mathbf{z}_2$.
-\textbf{E}. The EPI distribution predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
-\textbf{F}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
-\textbf{G}. Accuracy with delay period inactivation along each connectivity regime's dimension of sensitivity.
+Red and purple stars indicate modes $\mathbf{z}^*$ of each connectivity regime.
+Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
+(Bottom-left) EPI predictive distribution of task accuracies.
+\textbf{D}. The connectivity regimes have different responses to perturbation.
+(Top) Mean and standard error ($N_{\text{test}}$ = 25) of accuracy with respect to perturbation along the sensitivity dimension of each mode $\mathbf{z}^*$.
+(Middle) Same with perturbation in the dimension of increasing $\lambda_{\text{task}}$ ($\mathbf{v}_{\text{task}}$).
+(Bottom) Same with perturbation in the dimension of increasing $\lambda_{\text{diag}}$ ($\mathbf{v}_{\text{diag}}$).
 }
@@
 Thus, the altered state caused by delay period silencing makes the $p_P$ connectivity better than the $p_A$ down connectivity.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/fig5/fig5.pdf}
+\end{center}
+\caption{\footnotesize 
+\textbf{A}. The EPI distribution predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
+\textbf{B}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
+\textbf{C}. Mean and standard error ($N_{\text{test}}$ = 25) of accuracy with respect to perturbation along the sensitivity dimension of each mode $\mathbf{z}^*$.
+}
+\label{fig:SC_opt}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/figSC2/figSC2.pdf}
+\includegraphics[scale=0.7]{figures/figSC2/figSC2.pdf}
 \end{center}
@@
 \textbf{A}.
-Simulations in network regime $\mathbf{z}_1$ (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
-\textbf{B}. Same as A for network regime $\mathbf{z}_2$.
+Simulations in network regime 1 ($hW_{\text{fixed}} = -1.2$) (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
+\textbf{B}. Same as A for network regime 2.
 }
@@
 \textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
-\textbf{B}. Eigenvalues of connectivities of EPI distribution versus task accuracies.
+\textbf{B}. Accuracies for connectivity perturbations for increasing $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ (rest shown in Fig. \ref{fig:SC}D).
 }
@@
 \begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
-%\end{center}
-\caption{\small (SC4):  \textbf{A}. Something showing differences in regimes.
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
+\end{center}
+\caption{\small (SC4):  \textbf{A}. Shows how eigvals change throughout blah
 }
@@
 \end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after six epochs (Fig. \ref{fig:SC5}).
+Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations (see Section \ref{methods_AL_opt}).  
+The optimization converged after ten epochs (Fig. \ref{fig:SC6}).
 
@@
 }
-\label{fig:SC5}
+\label{fig:SC6}
 \end{figure}
@@
 We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
-The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 25$ at greatest entropy across random seeds.
+The distribution converged with criteria $N_{\text{test}} = 25$.
 

### 2021-03-31 written/eLifeRevision/Bittner_eLife2020.tex
finished methods section

@@
 \end{center}
-\caption{\small (SC4):  \textbf{A}. Shows how eigvals change throughout blah
+\caption{\small (SC4):  \textbf{A}. The individual parameters of each mode throughout the two regimes.
+\textbf{B}. The individual sensitivites of parameters of each mode throughout the two regimes.
 }
@@
 \textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Connectivity eigenvalues versus the task error induced by delay period inactivation.
-\textbf{C}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
+\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
 }
@@
 
-To make sense of this inferred distribution, we identified two modes used to represent the two regimes of connectivity in this posterior:
-\begin{equation}
-\begin{split}
-\mathbf{z}_1 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. }  hw = -1.25\text{, }  sW > 0
-\end{split}
-\end{equation}
-and 
-\begin{equation}
-\begin{split}
-\mathbf{z}_2 = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } hw = -1.25\text{, } sW < 0
-\end{split}
-\end{equation}
-
-To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $\mathbf{v}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
+The EPI distribution of SC model connectivites producing rapid task switching has interesting structure.
+Throughout $q_{\bm{theta}}(\mathbf{z} \mid \mathcal{X})$, we see that the probability distribution is narrow in $hW$ (Fig. \ref{fig:SC}C).
+This suggests that rapid task switching is sensitive to changes in $hW$, but this is only a single parameter.
+The local structure of the distribution varies across parameter space, and thus the nature in which parameter combinations affect rapid task switching.
+From visual inspection, we may hypothesize that there are two distinct regimes, most easily visualized in the $sW$-$hW$ marginal distribution: one where $sW$ and $hW$ are correlated for greater $sW$ and one where $sW$ and $hW$ are anticorrelated for lesser $sW$.
+
+We sought two sets of parameters in this distribution representative of each regime, so that we could assess their implications on computation.
+For fixed values of $hW$, we hypothesized that there are two modes: one in each regime of greater and lesser $sW$.
+To begin, we found one mode for each regime at $hW_{\text{fixed}} = -1.5$ using 200 steps of gradient ascent of the deep probability distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+In regime 1, the initialization had positive $sW$, and the initialization had negative $sW$ in regime 2, which led to disparate modes (Fig. \ref{fig:SC4} top).
+These modes were then used as the initialization to find the next mode at $hW_{\text{fixed}} = -1.4$ and so on.
+200 steps of gradient ascent were always taken, and learning rates of $2.5 \times 10^-4$ and $5 \times 10^-4$ were used for regimes 1 and 2, respectively.
+Each of these modes is denoted $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for regime $r \in \{1, 2\}$.
+
+At each mode, we measure the sensitivity dimension (that of most negative eigenvalue in the Hessian of the EPI distribution) $\mathbf{v}_1(\mathbf{z}^*)$.
+To resolve sign degeneracy in eigenvectors, we chose $\mathbf{v}_1(\mathbf{z}^*)$ to have negative element in $hW$.
+This tells us what parameter combination rapid task switching is most sensitive to at this parameter choice in the regime.
+We see that while the modes of each regime gradually converge to similar connectivites at $hW_{\text{fixed}} = -1.05$ (Fig.
+\ref{fig:SC4} top), the sensitivity dimensions remain categorically different throughout the two regimes (Fig.
+\ref{fig:SC4} bottom)
+Only at $hW_{\text{fixed}} = -1.05$ is there a flip in sensitivity from regime 2 to regime 1 (in $\mathbf{v}_1(\mathbf{z}^*)_{sW}$ and $\mathbf{v}_1(\mathbf{z}^*)_{hW}$).
+There is thus some ambiguity regarding the ``regime" of $\mathbf{z}^*(-1.05, 2)$, since the mode is derived from an initialization in regime 2, but has sensitivity like regime 1.
+We can consider this as an intermediate tranisitionary region of parameter space between the two regimes.
+To emphasize this, $\mathbf{z}^*(-1.05, 1)$ and $\mathbf{z}^*(-1.05, 2)$ have the same color.
+
+To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same basis vectors $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
 \ref{fig:SC3}A).
-These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
-We found significant trends across the EPI posterior connectivities: the eigenvalues $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ were correlated with $p_P$, while $\lambda_{\text{all}}$ was anticorrelated with $p_P$.
-$\lambda_{\text{all}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ were all significantly anticorrelated with $p_A$.
-
-Under this decomposition, we can re-visualize the posterior in eigenvalue space (Fig. \ref{fig:SC4}).
-Furthermore, we can project the dimensions of sensitivity into eigenvalue space as well, giving us a more intuitive sense of how connectivity affects computation in each regime.
-We see that sensitivity dimensions $\mathbf{v}_1$ and $\mathbf{v}_2$, which cause $p_P$ to increase and a regime dependent change in $p_A$, both point in the direction of increasing $\lambda_{\text{side}}$ and decreasing $\lambda_{\text{task}}$.
-These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
-As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
-A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
-
-\textbf{OPTO} \\
-During periods of optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
+These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} eigenmode - all neurons co-fluctuate, \textit{side} eigenmode - one side dominates the other, \textit{task} eigenmode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
+Due to the parametric structure of the connectivity matrix, the parameters $\mathbf{z}$ are a linear function of the eigenvalues $\bm{\lambda} = [\lambda_{\text{all}}, \lambda_{\text{side}}, \lambda_{\text{task}}\lambda_{\text{diag}}]^\top$ associated with these eigenmodes.
+\begin{equation}
+\mathbf{z} = A\bm{\lambda}
+\end{equation}
+\begin{equation}
+A = \frac{1}{4} \begin{bmatrix} 1 & 1 & 1 & 1 \\ 1 & -1 & -1 & 1 \\ 1 & 1 & -1 & -1 \\ 1 & -1 & 1 & -1 \end{bmatrix}.
+\end{equation}
+
+We are interested in the effect of raising or lowering the amplification of each eigenmode in the connectivity matrix.
+To test this, we calculate the the unit vector of changes in the connectivity $\mathbf{z}$ that result form a change in the associated eigenvalues
+\begin{equation}
+\mathbf{v}_{\text{a}} = \frac{\frac{\partial \mathbf{z}}{\partial \lambda_{\text{a}}}}{|\frac{\partial \mathbf{z}}{\partial \lambda_{\text{a}}}|_2},
+\end{equation}
+where
+\begin{equation} \label{eq:dzdlambda}
+\frac{\partial \mathbf{z}}{\partial \lambda_{\text{a}}} = A\mathbf{e}_{\text{a}},
+\end{equation}
+and e.g. $\mathbf{e}_{\text{all}} = [1, 0, 0, 0]^\top$.
+So $\mathbf{v}_{\text{a}}$ is the normalized column of A corresponding to eigenmode $a$.
+While perturbations in the sensitivity dimension $\mathbf{v}_1(\mathbf{z}^*)$ adapt with the mode $\mathbf{z}^*$ chosen, perturbations in $\mathbf{v}_a$ for $a \in \{\text{all}, \text{side}, \text{text}, \text{diag}\}$ are invariant to $\mathbf{z}$ (Equation \ref{eq:dzdlambda}).
+%
+%These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
+%As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
+%A key differentiating factor is that $\mathbf{v}_1$ substantially increases $\lambda_{\text{diag}}$, while $\mathbf{v}_2$ does not.
+
+We tested whether the inferred SC model connectivites could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.
+During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
 \begin{equation}
@@
 \end{equation}
-
-During optogenetic silencing simulations, activations $x_\alpha(t)$ were set to a fraction of their values ($1 - \gamma$), where $\gamma$ is the optogenetic perturbation strength.
-We found that $\lambda_{\text{all}}$ and $\lambda_{\text{diag}}$ were significantly anticorrelated with Anti error during delay period inactivation.
 Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.

### 2021-03-31 written/eLifeRevision/Bittner_eLife2020.tex
SC section update

@@
 \tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
-\mathbf{x} &= \phi(\mathbf{u})
+\mathbf{x} &= \phi(\mathbf{u}).
 \end{split}
 \end{equation}
-where $\tau= 90\text{ms}$ and there is white noise of variance $0.2^2$.
 The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.  
@@
 The patterns of connectivity that govern each task accuracy are nonlinear (Fig. \ref{fig:SC1}A-B); they are not captured well by linear prediction (Fig. \ref{fig:SC1}C).
-For example, the patterns in connectivity increasing Pro accuracy change dramatically after crossing a threshold of $sW$ (Fig. \ref{fig:SC1}A $sW$-$hW$ marginal distribution).
-Not only has EPI captured this intricate, nonlinear posterior, it offers probabilistic tools for understanding the different regimes of model behavior.
-
-To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
-\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
-Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
-We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
-The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
-\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
-\ref{fig:SC}D, bottom-left, Fig.
-\ref{fig:SC2}).
-In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy matches monotonic increases in Pro accuracy.
-
-To understand why these distinct connectivity regimes have different failure modes, we can analyze the properties of connectivity in each regime.
-By taking the eigendecomposition of $W$, we can quantify how strongly different modes of processing are reflected in the connectivity matrix (see Section \ref{methods_SC}).  
-\textbf{Note SB: I need to alter these plots a tad to confirm and support the mechanistic explanation.}
+For example, there appear to be two regimes of connectivity: the local structure of the EPI distribution changes dramatically after crossing a threshold of $sW$ (Fig.
+\ref{fig:SC1}A $sW$-$hW$ marginal distribution).
+Not only has EPI captured this intricate, nonlinear distribution, we can use the probabilistic toolkit on the deep probability distribution $q_{\bm{theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.
+
+To identify the parameters of these two parametric regimes, we point out that for many fixed values of parameter $hW$, there are two modes in the EPI distribution.  
+Thus, by fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ in the parameter spaces proximal to each regime, we can identify modes $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for each regime $r \in \{1, 2\}$ (see Section \ref{methods_SC}).
+As $hW_{\text{fixed}}$ increases, the modes eventually coalesce to intermediate parameters reflecting a transition between the two regimes (Fig. \ref{fig:SC4} top).
+However the sensitivity dimensions of these modes $\mathbf{v}_1$ (refer to Section \ref{results_dgm}), which reflect the structure of the EPI distribution around those modes, are very similar throughout the entirety of each regime independent of $hW_{\text{fixed}}$.
+This suggests that there are categorically different ways in which connectivity governs computation in each regime.
+
+To understand how SC connectivity governs computation in each regime, we can examine how perturbations along $\mathbf{v}_1(\mathbf{z}^*)$ affect task performance in each regime.
+While the increase in Pro accuracy with $\mathbf{v}_1$ perturbation is largely unnaffected by regime (Fig.
+\ref{fig:SC}D, top-left), we see a stark difference in Anti accuracy: Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotically increases in regime 2 (Fig.
+\ref{fig:SC}D, top-right).
+Therefore, the sharp change in structure of the EPI distribution can be attributed to this changing pathology in rapid task switching behavior.
+
+The relation of biologically interpretable variables to regime and sensitivity dimension are salient from the pairplot in (Fig.
+\ref{fig:SC}C), but we may seek to gain a more computation-driven understanding of each regime.
+A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$ (see Section \ref{methods_SC}).
+\textbf{explain eigenmode intuition}.
+\textbf{Summarize with concise description and inuition from Fig. \ref{fig:SC}D.}
+
+\subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiements} \label{results_SC}
 
@@
 found that optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
-\ref{fig:SC}E)).
+\ref{fig:SC_opto}A).
 All network connectivities inferred by EPI exhibited this same effect, when network activities were silenced during the delay period (see Section \ref{methods_SC}).
@@
 \ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
-\ref{fig:SC}F).
+\ref{fig:SC_opto}B).
 The strong anticorrelation between $p_P$ and $p_{A, opto}$ across EPI inferred connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
 Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
-\ref{fig:SC}F).
+\ref{fig:SC_opto}C).
 Thus, the altered state caused by delay period silencing makes the $p_P$ connectivity better than the $p_A$ down connectivity.
@@
 }
-\label{fig:SC_opt}
+\label{fig:SC_opto}
 \end{figure}
@@
 Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
+
+%To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
+\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
+%Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
+%We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
+%The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
+\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
+\ref{fig:SC}D, bottom-left, Fig.
+\ref{fig:SC2}).
+%In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy matches monotonic increases in Pro accuracy.
+%
+%To understand why these distinct connectivity regimes have different failure modes, we can analyze the properties of connectivity in each regime.
+%By taking the eigendecomposition of $W$, we can quantify how strongly different modes of processing are reflected in the connectivity matrix (see Section \ref{methods_SC}).
 

### 2021-04-01 written/eLifeRevision/Bittner_eLife2020.tex
SC refine and spell check

@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI identifies multiple regimes of rapid task switching} \label{results_SC}
+\subsection{EPI identifies two regimes of rapid task switching} \label{results_SC}
 It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity allow this ability.
@@
 The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
-Model responses are bounded from 0 to 1 as a function $\phi$ of a dynamically evolving internal variable $\mathbf{u}$
+Model responses are bounded from 0 to 1 as a function $\phi$ of an internal variable $\mathbf{u}$
 \begin{equation}
@@
 \ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
-The patterns of connectivity that govern each task accuracy are nonlinear (Fig. \ref{fig:SC1}A-B); they are not captured well by linear prediction (Fig. \ref{fig:SC1}C).
+The patterns of connectivity that govern each task accuracy are nonlinear (Fig. \ref{fig:SC1}A-B); they are not captured well by linear correlations (Fig. \ref{fig:SC1}C).
 For example, there appear to be two regimes of connectivity: the local structure of the EPI distribution changes dramatically after crossing a threshold of $sW$ (Fig.
 \ref{fig:SC1}A $sW$-$hW$ marginal distribution).
-Not only has EPI captured this intricate, nonlinear distribution, we can use the probabilistic toolkit on the deep probability distribution $q_{\bm{theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.
-
-To identify the parameters of these two parametric regimes, we point out that for many fixed values of parameter $hW$, there are two modes in the EPI distribution.  
-Thus, by fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ in the parameter spaces proximal to each regime, we can identify modes $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for each regime $r \in \{1, 2\}$ (see Section \ref{methods_SC}).
-As $hW_{\text{fixed}}$ increases, the modes eventually coalesce to intermediate parameters reflecting a transition between the two regimes (Fig. \ref{fig:SC4} top).
-However the sensitivity dimensions of these modes $\mathbf{v}_1$ (refer to Section \ref{results_dgm}), which reflect the structure of the EPI distribution around those modes, are very similar throughout the entirety of each regime independent of $hW_{\text{fixed}}$.
-This suggests that there are categorically different ways in which connectivity governs computation in each regime.
-
-To understand how SC connectivity governs computation in each regime, we can examine how perturbations along $\mathbf{v}_1(\mathbf{z}^*)$ affect task performance in each regime.
-While the increase in Pro accuracy with $\mathbf{v}_1$ perturbation is largely unnaffected by regime (Fig.
-\ref{fig:SC}D, top-left), we see a stark difference in Anti accuracy: Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotically increases in regime 2 (Fig.
+Not only has EPI captured this intricate, nonlinear distribution, we can use the distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ returned by EPI to understand these two parametric regimes of SC connectivity.
+
+To distinguish these two parts of the distribution, we point out that for many fixed values of parameter $hW$, there are two modes in the EPI distribution.  
+Thus, by fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ in the parameter spaces proximal to each mode, we can identify a set of modes $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for each putative regime $r \in \{1, 2\}$ (see Section \ref{methods_SC}).
+As $hW_{\text{fixed}}$ increases, the modes coalesce to intermediate parameters reflecting a transition between the two sets of modes (Fig. \ref{fig:SC4} top).
+However, the sensitivity dimensions of these modes $\mathbf{v}_1$ (refer to Section \ref{results_dgm}), which reflect the structure of the EPI distribution around each mode, are different across putative regime, yet consistent across $hW_{\text{fixed}}$.
+This categorical difference in sensitivity dimension across the two sets of modes shows that they indeed represent two different regimes of computation in which connectivity governs computation in different ways.
+
+To understand how SC connectivity governs computation in each regime, we can examine how perturbations along $\mathbf{v}_1(\mathbf{z}^*)$ affect task performance in each regime; we measure task accuracy for connectivity changes in the dimension that rapid task switching is most sensitive.
+While the monotonic increase in Pro accuracy with $\mathbf{v}_1$ perturbation is largely unaffected by regime (Fig.
+\ref{fig:SC}D, top-left), we see a stark difference in Anti accuracy: Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig.
 \ref{fig:SC}D, top-right).
-Therefore, the sharp change in structure of the EPI distribution can be attributed to this changing pathology in rapid task switching behavior.
-
-The relation of biologically interpretable variables to regime and sensitivity dimension are salient from the pairplot in (Fig.
-\ref{fig:SC}C), but we may seek to gain a more computation-driven understanding of each regime.
-A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$ (see Section \ref{methods_SC}).
-\textbf{explain eigenmode intuition}.
-\textbf{Summarize with concise description and inuition from Fig. \ref{fig:SC}D.}
-
-\subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiements} \label{results_SC}
+These two rapid task switching pathologies are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(\cdot, 1))$ vs $\mathbf{v}_1(\mathbf{z}^*(\cdot, 2))$) and explain the sharp change in local structure of the EPI distribution.
+
+To understand the connectivity mechanisms that distinguish these two regimes, we perturb connectivity at each mode in dimensions that have well defined roles in processing for the Pro and Anti tasks.
+A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$.
+These eigenmodes have intuitive roles in processing in each task, and are accordingly named the \textit{all}, \textit{side}, \textit{task}, and \textit{diag} eigenmodes (see Section \ref{methods_SC}).
+Furthermore, the parameter dimension $\mathbf{v}_a$ ($a \in \{\text{all}, \text{side}, \text{task}, \text{ and diag}\}$) that increases the eigenvalue of connectivity $\lambda_a$ is $\mathbf{z}$-invariant (unlike the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$) and $\mathbf{v}_a \perp \mathbf{v}_{b \neq a}$.
+Thus, by changing the degree of amplification of each processing mode by  perturbing $\mathbf{z}$ along $\mathbf{v}_a$, we can elicit the differentiating properties of the two regimes.
+
+Through these connectivity perturbation analyses, we found that increasing $\lambda_{\text{task}}$ strongly reduced Pro accuracy in regime 1, yet strongly reduced Anti accuracy in regime 2.
+This suggests that stronger task representations can inhibit both Pro and Anti task performance in different contexts.
+Furthermore, changing $\lambda_{\text{task}}$ in either direction decreases Anti performance in regime 1, showing that Anti task performance in regime 1 is dependent on a specific level of task representation.
+We also found that with increasing $\lambda_{\text{diag}}$, Pro accuracy increased in both regimes, but there were opposite effects on Anti accuracy.
+In regime 1, stronger amplification of diagonal population patterns decreased Anti accuracy, while in regime 2 accuracy increased.
+These findings give us an understanding of the mechanistic differences in computation enabling rapid task switching in each regime.
+
+\subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiments} \label{results_SC}
 
@@
 \ref{fig:SC_opto}C).
-Thus, the altered state caused by delay period silencing makes the $p_P$ connectivity better than the $p_A$ down connectivity.
+Thus, the altered state caused by delay period silencing makes the connectivity governing $p_P$ more influential than the connectivity governing $p_A$.
 
@@
 A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
-We see in Section \ref{results_LRRNN} that EPI scales seemlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
+We see in Section \ref{results_LRRNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
 EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
@@
 \caption{\small (SC4):  \textbf{A}. The individual parameters of each mode throughout the two regimes.
-\textbf{B}. The individual sensitivites of parameters of each mode throughout the two regimes.
+\textbf{B}. The individual sensitivities of parameters of each mode throughout the two regimes.
 }
@@
 
-The EPI distribution of SC model connectivites producing rapid task switching has interesting structure.
+The EPI distribution of SC model connectivities producing rapid task switching has interesting structure.
 Throughout $q_{\bm{theta}}(\mathbf{z} \mid \mathcal{X})$, we see that the probability distribution is narrow in $hW$ (Fig. \ref{fig:SC}C).
@@
 This tells us what parameter combination rapid task switching is most sensitive to at this parameter choice in the regime.
-We see that while the modes of each regime gradually converge to similar connectivites at $hW_{\text{fixed}} = -1.05$ (Fig.
+We see that while the modes of each regime gradually converge to similar connectivities at $hW_{\text{fixed}} = -1.05$ (Fig.
 \ref{fig:SC4} top), the sensitivity dimensions remain categorically different throughout the two regimes (Fig.
@@
 There is thus some ambiguity regarding the ``regime" of $\mathbf{z}^*(-1.05, 2)$, since the mode is derived from an initialization in regime 2, but has sensitivity like regime 1.
-We can consider this as an intermediate tranisitionary region of parameter space between the two regimes.
+We can consider this as an intermediate transitionary region of parameter space between the two regimes.
 To emphasize this, $\mathbf{z}^*(-1.05, 1)$ and $\mathbf{z}^*(-1.05, 2)$ have the same color.
@@
 We are interested in the effect of raising or lowering the amplification of each eigenmode in the connectivity matrix.
-To test this, we calculate the the unit vector of changes in the connectivity $\mathbf{z}$ that result form a change in the associated eigenvalues
+To test this, we calculate the unit vector of changes in the connectivity $\mathbf{z}$ that result from a change in the associated eigenvalues
 \begin{equation}
@@
 
-We tested whether the inferred SC model connectivites could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.
+We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.
 During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$

### 2021-04-12 written/eLifeRevision/Bittner_eLife2020.tex

@@
 \textbf{D}. For a choice of model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
-\textbf{E}. The EPI posterior producing intermediate hub frequency.
+\textbf{E}. The EPI distribution producing intermediate hub frequency.
 Samples are colored by log probability density.  
@@
 
-\subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiments} \label{results_SC}
+\subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiments} \label{results_SC_opt}
 
@@
 \ref{fig:SC_opto}C).
-Thus, the altered state caused by delay period silencing makes the connectivity governing $p_P$ more influential than the connectivity governing $p_A$.
+Thus, the altered state caused by delay period silencing makes the connectivity governing $p_P$ more influential on Anti accuracy than the connectivity governing $p_A$.
 
@@
 In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This EPI distribution revealed multiple regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
-EPI allowed us to conclude that since \textit{all} parameters of this model producing rapid task switching make an experimentally verified prediction, the model is well-chosen in that regard.
-Finally, we used our knowledge about how $\mathbf{z}$ governs $p_{A,opto}$ to make accurate predictions about each identified regime of connectivity.
+This EPI distribution revealed two regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
+We found that both of these parametric regimes identified by EPI reproduce results from optogenetic inactivation experiments: when activity is silenced during the delay period, only Anti accuracy suffers.
+We then identified the connectivity mechanisms governing Anti accuracy during delay period silencing, and showed that they are regime invariant.
 
@@
 \section{Discussion} 
-
 In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{paninski2018neural}.
-Such careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
-However, these statistical models lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it. 
-In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
+Careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
+However, these statistical models often lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it.
+In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation, rather than measuring structure in some noisy observed dataset.
 The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards careful inference in careful models of neural computation (see Section \ref{methods_related}).
-
-Specifically, approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential} has been the standard approach to parameter inference in neural circuit models lacking tractable likelihoods.
-ABC methods do not confer probabilities on accepted parameters, require an acceptance threshold chosen to trade-off inference quality with tractability, do not scale efficiently to high-dimensional parameter spaces, and require independent techniques to analyze sensitivity for local parameter choices \cite{sisson2018handbook}.
-In contrast, EPI allows probability evaluations at any point in parameter space, conditions posteriors on the natural quantification of emergent properties, scales to high dimensional parameter spaces, and naturally admits sensitivity quantification via fast evaluations of the posterior Hessian.
-
-Technically, EPI is a maximum entropy method, which learns parameter distributions that are as random as possible given that they produce the emergent property.
-Conceptually, maximally random distributions given some constraints are useful for understanding parametric sensitivity.
-This is well understood in Bayesian inference, where maximum entropy is the chosen normative principle.
-This is emphasized by an innovative formalism unifying top-down maximum entropy normative models with bottom-up statistical models \cite{mlynarski2020statistical}.
-Indeed, EPI is an adaptive variational inference program, and may be considered to have a Bayesian uniform prior (see Section \ref{methods_VI}).
-
-%Biologically realistic models of neural circuits often prove formidable to analyze.
-%Two main factors contribute to the difficulty of this endeavor.
-%First, in most neural circuit models, the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space. 
-%Second, even in low dimensional circuits, the structure of the parametric regimes governing emergent properties is intricate.  
-%For example, these circuit models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-
-Biologically realistic models of neural circuits often prove formidable to analyze for two main reasons.
-A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
-We see in Section \ref{results_LRRNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
-EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
-
-A secondary challenge is that the structure of the parametric regimes governing emergent properties is intricate.  
-For example, even in low dimensional circuits, models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-With EPI, we use deep probabillity distributions to capture the complex nonlinear parameter distributions governing model behavior.
-In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing curcuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
-In Section \ref{results_SC}, we identified two regimes of SC connectivity resulting in rapid task switching, and found that the full distribution of rapid task switching networks reproduced an experimental result.
-
-%In Section \ref{results_V1}, we advanced the tractability of low-dimensional neural circuit models by showing that EPI offers insights about cell-type specific input-responsivity that cannot be afforded through the available linear analytical methods  \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019}.
-%By flexibly conditioning this V1 model on different emergent properties, we performed an exploratory analysis of a \emph{model} rather than a dataset, generating a set of testable hypotheses, which were proved out.
-%Furthermore, exploratory analyses can be directed towards formulating hypotheses of a specific form. 
-%For example, model parameter dependencies on behavioral performance can be assessed by using EPI to condition on various levels of task accuracy (See Section \ref{results_SC}). 
-%This analysis identified experimentally testable predictions (proved out \textit{in-silico}) of patterns of effective connectivity in SC that should be correlated with increased performance.
-
-%In our final analysis, we presented a novel procedure for doing statistical inference on interpretable parameterizations of RNNs executing simple tasks. 
-%Specifically, we analyzed RNNs solving a posterior conditioning problem in the spirit of \cite{pitkow2017inference, echeveste2019cortical}. 
-%This methodology relies on recently extended theory of responses in random neural networks with low-rank structure \cite{mastrogiuseppe2018linking}. 
-%While we focused on rank-1 RNNs, which were sufficient for solving this task, this inference procedure generalizes to RNNs of greater rank necessary for more complex tasks.
-%The ability to apply the probabilistic model selection toolkit to RNNs should prove invaluable as their use in neuroscience increases.
-
-EPI leverages deep learning technology for neuroscientific inquiry in a categorically different way than approaches focused on training neural networks to execute behavioral tasks \cite{richards2019deep}.
-These works focus on examining optimized deep neural networks while considering the objective function, learning rule, and architecture used.
-This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
-All of these approaches can be used complementarily to enhance the practice of theoretical neuroscience.
+In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to emergent phenomena in  interpretable models of neural computation (see Section \ref{methods_related}).
+
+Methodology for statistical inference in mechanistic models of neural circuits has evolved considerably in recent years.
+Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but more recently developed methodology employs deep learning to improve efficiency or provide flexible distribution approximations.
+SNPE \cite{gonccalves2019training} and other sequential techniques for inference in mechanistic models developed along with EPI (see Section \ref{methods_related}) have been used for posterior inference with noisy experimental datasets.
+On the other hand, EPI is a deep inference technique designed to condition upon mathematical criteria, such that the parameter distribution only produces  the specified \textit{emergent properties} of computation.
+EPI is thus ideally suited for questions in theoretical neuroscience, and we show that it has superior scaling properties to these other inference techniques (see Section \ref{results_LRRNN}).
+
+In this work, we prove out the utility of deep probability distributions for theoretical neuroscience.
+While previous work has used SNPE to obtain flexible posterior approximations in mechanistic models conditioned on experimental datasets, we use the rich structure captured by deep probability distributions in EPI to gain new theoretical insights.
+This is first done in a complex model of V1, where we combine the modeling advancements \cite{litwin2016inhibitory, hennequin2018dynamical}, which make analytic characterization of excitatory variability very complicated.
+There, EPI clearly and simply revealed the parametric structure of input variability across neuron-type populations that governed excitatory variability, which has implications on the dimensionality and nature information transmission in visual cortex.
+
+Finally, we uesd EPI to identify two distinct regimes of SC connectivity that enabled rapid task switching.
+By systematically characterizing the local structure of the inferred distribution using the analytic capabilities deep probability distributions, we discerned a mechanistic understanding of each computational regime.
+Each of these regimes reproduced effects from optogenetic experiments \cite{duan2015requirement}, suggesting that both are biologically plausible.
+These analyses of the V1 and SC models serve as examples of how to leverage the probabilistic toolkit for theoretical insight into models of neural computation.
+
+%In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{paninski2018neural}.
+%Careful inference procedures are developed for these statistical models allowing precise, quantitative reasoning, which clarifies the way data informs beliefs about the model parameters.  
+%However, these statistical models often lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it.
+%In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation.  
+%The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
+%In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards careful inference in careful models of neural computation (see Section \ref{methods_related}).
+%
+%Specifically, approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential} has been the standard approach to parameter inference in neural circuit models lacking tractable likelihoods.
+%ABC methods do not confer probabilities on accepted parameters, require an acceptance threshold chosen to trade-off inference quality with tractability, do not scale efficiently to high-dimensional parameter spaces, and require independent techniques to analyze sensitivity for local parameter choices \cite{sisson2018handbook}.
+%In contrast, EPI allows probability evaluations at any point in parameter space, conditions distributions on the natural quantification of emergent properties, scales to high dimensional parameter spaces, and admits fast sensitivity calculations.
+%
+%Technically, EPI is a maximum entropy method, which learns parameter distributions that are as random as possible given that they produce the emergent property.
+%Conceptually, maximally random distributions given some constraints are useful for understanding parametric sensitivity.
+%This is well understood in Bayesian inference, where maximum entropy is the chosen normative principle.
+%This is emphasized by an innovative formalism unifying top-down maximum entropy normative models with bottom-up statistical models \cite{mlynarski2020statistical}.
+%Indeed, EPI is an adaptive variational inference program, and may be considered to have a Bayesian uniform prior (see Section \ref{methods_VI}).
+%
+%Biologically realistic models of neural circuits often prove formidable to analyze for two main reasons.
+%A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
+%We see in Section \ref{results_LRRNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
+%EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
+%
+%A secondary challenge is that the structure of the parametric regimes governing emergent properties is intricate.  
+%For example, even in low dimensional circuits, models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
+%With EPI, we use deep probabillity distributions to capture the complex nonlinear parameter distributions governing model behavior.
+%In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing curcuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
+%In Section \ref{results_SC}, we identified two regimes of SC connectivity resulting in rapid task switching, and found that the full distribution of rapid task switching networks reproduced an experimental result.
+%
+%EPI leverages deep learning technology for neuroscientific inquiry in a categorically different way than approaches focused on training neural networks to execute behavioral tasks \cite{richards2019deep}.
+%These works focus on examining optimized deep neural networks while considering the objective function, learning rule, and architecture used.
+%This endeavor efficiently obtains sets of parameters that can be reasoned about with respect to such considerations, but lacks the careful probabilistic treatment of parameter inference in EPI.
+%All of these approaches can be used complementarily to enhance the practice of theoretical neuroscience.
 
@@
 Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.
-
-In EPI, deep probability distributions are used as posterior approximations $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
+By defining these means and variances over both levels of stochasticity -- the inferred distribution and that of the model -- there is a fine degree of control over predictions made by the inferred parameters.
+
+In EPI, deep probability distributions are optimized to learn the inferred distribution.
  In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
 $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-\begin{equation} \label{eq:deep_transform}
+\begin{equation}
+\label{eq:deep_transform}
 \mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
 \end{equation}
-Such deep probability distributions embed the posterior distribution in a deep network.
-Once optimized, this deep network representation has remarkably useful properties: immediate posterior sampling, and immediate probability, gradient, and Hessian evaluation at any parameter choice.
-
-Given a choice of model $p(\mathbf{x} \mid \mathbf{z})$ and emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
+Such deep probability distributions embed the inferred distribution in a deep network.
+Once optimized, this deep network representation has remarkably useful properties: fast sampling, probability evaluations, and also first- and second-order probability gradient evaluations.
+
+By choosing a neural circuit model, often represented as a system of differential equations, we implictly definge a model likelihood $p(\mathbf{x} \mid \mathbf{z})$, which may be unknown or intractable for our purposes.
+Given this model choice and that of an emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
 \begin{equation} \label{eq:opt}
@@
 \end{equation} 
-Entropy is chosen as the normative selection principle, since we want the posterior to only contain structure predicated by the emergent property \cite{jaynes1957information, elsayed2017structure}.
-This choice of selection principle is also that of standard Bayesian inference, and we derive an exact relation between EPI and variational inference (see Section \ref{methods_ME_EF}).
-However, a key difference is that variational inference and other Bayesian methods do not constrain the predictions of their inferred posteriors.
+Entropy is chosen as the normative selection principle to match that of Bayesian inference (see Section \ref{methods_ME_EF}).
+However, a key difference is that variational inference and other Bayesian methods do not constrain the predictions of their inferred parameter distribution.
 This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
@@
 Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
-Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain the relation between EPI and variational inference (Section \ref{methods_VI}).
+Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain how EPI can be viewed as a form of variational inference (Section \ref{methods_VI}).
 
@@
 SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:LRRNN}).
-Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again with a new population of initialized particles to obtain additional samples.
-
-For scientific model analysis, we seek a posterior distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
+Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again from scratch with a new population of initialized particles to obtain additional samples.
+
+For scientific model analysis, we seek a parameter distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
 For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
-EPI executes likelihood-free inference using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational posterior approximation.
-The deep neural network of EPI defines the parametric form of the posterior approximation.
+EPI infers parameters in likelihood-free models using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational approximation.
+The deep neural network of EPI defines the parametric form of the distribution approximation.
 Furthermore, the EPI distribution is constrained to produce an emergent property.
@@
 SNPE belongs to a ``sequential" class of recently developed LFI methods in which two neural networks are used for posterior inference.
-This first neural network is a normalizing flow used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
+This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
 A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE \cite{hermans2020likelihood}).
 In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
-This contrasts with EPI and SNPE, which afford a normalizing flow approximation to the posterior, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
+This contrasts with EPI and SNPE, which use deep probability distributions to model parameters, which facilitates immediate measurements of sample probability, gradient, or Hessian for system analysis.
 The second neural network in this sequential class of methods is the amortizer.
-This network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
-These methods are optimized on a conditional density (or ratio) estimation objective on a sequentially adapting finite sample-based approximation to the posterior.
-
-The approximating fidelity of the first neural network in sequential approaches is optimized to generalize across the entire distribution it is conditioned upon.
-This optimization towards generalization of sequential methods can reduce the accuracy at the singular posterior of interest.
-Whereas in EPI, the entire expressivity of the normalizing flow is dedicated to learning a single distribution as well as possible.
-While amortization is not possible in EPI parameterized by the mean parameter $\mu$ (due to the inverse mapping problem \cite{wainwright2008graphical}), we have shown this two-network amortization approach to be effective in exponential family distributions defined by their natural parameterization \cite{bittner2019approximating}.
+This unconstrained deep network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
+These methods are optimized on a conditional density (or ratio) estimation objective.
+The data used to optimize this objective are generated via an adaptive procedure, in which training data pairs ($\mathbf{x}_i$, $\mathbf{z}_i$) become sequentiall closer to the true data and posterior.
+
+The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
+This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
+Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
+Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section \ref{methods_ME_EF}).
+Since EPI distributions are defined by the mean $\bm{\mu}$ of their statistics, there is the well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} that prohibits an amortization based approach.
+However, we have shown that the same two-network architecture of the sequential LFI methods can be used for amortized inference in intractable exponential family posteriors using their natural parameterization \cite{bittner2019approximating}.
+
+Finally, one important differentiating factor between EPI and sequential LFI methods is that EPI leverages gradients $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ during optimization.
+These gradients can improve convergence time and scalability, as we have shown on an example conditioning low-rank RNN connectivity on the property of stable amplification (see Section \ref{results_LRRNN}).
+With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these model gradients when they are tractable.
+Sequential LFI techniques may be better suited for sceintific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable: when there is a nondifferentiable model or it requires lengthy simulations.
+However, the sequential LFI techniques cannot constrain the predictions of the inferred distribution in the manner of EPI.
 
@@
 Global structural non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
-With EPI, we have all the benefits given by a statistical inference method plus the ability to query the gradient or Hessian of the inferred distribution at any chosen parameter value.
+With EPI, we have all the benefits given by a statistical inference method plus the ability to query the first- or second-order gradient of the probability of the inferred distribution at any chosen parameter value.
+The second-order gradient of the log probability (the Hessian), which is directly afforded by EPI distributions, produces salient information about parametric sensitivty of the emergent property.
+For example, the eigenvector with most negative eigenvalue of the Hessian shows parametric combinations away from a parameter choice that decrease the in EPI distribution probability the fastest.
+We refer to this eigenvector as the sensitivity dimension, and it is used to generate scientific insight about a model of superior colliculus connectivity (see Section \ref{results_SC}).
 
@@
 
-\subsubsection{Normalizing flows}\label{methods_NF}
-%Deep probability models typically consist of several layers of fully connected neural networks.  
-Deep probability distributions are comprised of multiple layers of fully connected neural networks (Equation \label{eq:deep_transform}).
-%When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
-For $\mathbf{z}' = g(\mathbf{z})$,
+\subsubsection{Deep probability distributions and normalizing flows}\label{methods_NF}
+Deep probability distributions are comprised of multiple layers of fully connected neural networks (Equation \ref{eq:deep_transform}).
 When each neural network layer is restricted to be a bijective function, the sample density can be calculated using the change of variables formula at each layer of the network.
@@
 Glow permutations were used in between coupling stages \cite{kingma2018glow}.
-This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
+This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only one of the forward or backward passes can be efficient.
 In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
@@
 where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
-The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
-Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
+The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt} (see Section \ref{methods_VI}).
+Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_ME_EF}).
 Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
 
-To take gradients with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, it can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
+The gradient with respect to entropy $H(q_{\bm{\theta}}(\mathbf{z}))$ can be expressed using the reparameterization trick as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0(\mathbf{z}_0$):
 \begin{equation}
@@
 %The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-Maximum entropy distributions have a fundamental link to exponential family distributions. 
+EPI is a maximum entropy distribution, which have fundamental links to exponential family distributions. 
 A maximum entropy distribution of form:
@@
 \end{equation}
-The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}
- \begin{equation}
-\mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}
+The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify except in special cases \cite{wainwright2008graphical}.
+
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}.
 The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
@@
 In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
-In Bayesian inference, we obtain a posterior distribution $p(z \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+In Bayesian inference, we obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
 \begin{equation}
@@
 The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{blei2017variational}, are utilized.
+However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{saul1998mean}, are utilized.
 
@@
 \begin{equation} 
+\label{eq:eta_f}
 \bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
@@
 \begin{equation}
+\label{eq:VI_obj}
 \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
@@
 Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
-The data $\mathbf{x}$ used by this implicit variational inference program would be that generated by the adapting variational approximation $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z}) q_{\bm{\theta}}(\mathbf{z})$, and the likelihood parameters $\bm{\eta}_{f}$ of EPI optimization epoch $k$ are predicated by $\bm{\eta}_{\text{opt},k}$ .
-However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
+The mean and variances of the mean-field gaussian likelihood are predicated by $\bm{\eta}_{\text{opt}}$ (Equations \ref{eq:eta_f} and \ref{eq:VI_obj}), which is adapted after each EPI optimization epoch based on $\mathcal{X}$ (see Section \ref{methods_AL_opt}).
+In EPI, the inferred distribution is not conditioned on a finite dataset as in variational inference, but rather the emergent property $\mathcal{X}$ dictates the likelihood parameterization such that the inferred distribution will produce the emergent property.
+As a note, we could not simply choose $\bm{\mu}_f$ and $\bm{\sigma}_f$ directly from the outset, since we do not know which of these choices will produce the emergent property $\mathcal{X}$, which necessitates the EPI optimization routine that adapts $\bm{\eta}_{\text{opt}}$.
 Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
@@
 \caption{\small (V1 2)
-Posterior predictive distributions of the sum of squares of each pair of noise parameters.
+EPI predictive distributions of the sum of squares of each pair of noise parameters.
  }
@@
 \caption{\small (V1 3)
-EPI posterior for $\mathcal{X}(10\text{ Hz})$.
+EPI inferred distribution for $\mathcal{X}(10\text{ Hz})$.
  }
@@
 \textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI posterior.
+\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI distribution.
 }

### 2021-04-15 written/eLifeRevision/Bittner_eLife2020.tex
latest draft

@@
 \usepackage{bm}
+\usepackage{cite}
 
@@
 The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
-A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
+A neural circuit is systematized with a set of equations -- the mechanistic model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
 The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
@@
 Consider the idealized practice: one carefully designs a model and analytically derives how computational properties determine model parameters.
-Seminal examples of this gold standard (which often adopt approaches from statistical physics) include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
+Seminal examples of this gold standard include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
 Still, we can gain insight into these complex models by identifying the distribution of parameters that produce computations.
-By solving the inverse problem in this way, scientific analysis of complex biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-However, this work clarifies an important incongruity between this methodology and theoretical approaches to neuroscience.
-
-A pivotal detail in theoretical neuroscience is the manner in which the computation is specified.
-One common approach is to use data that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
-While a host of suitable algorithms have been developed for inferring parameters in models of neural data \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}), the level of insight gained strongly depends on the quantity and quality of available data.
-Ultimately theoretical neuroscience is concerned with the computational properties -- the \emph{emergent phenomena} -- of our models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, not noisy observed datasets.
-To use the aforementioned inference paradigm, scientists must shoehorn such mathematical criteria into an artificial dataset compatible with existing statistical approaches.
-
-In addition, another crucial challenge in neuroscientific modeling has been the inversion of biologically realistic or ``mechanistic" models.
-Most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack the explicit likelihood necessary for statistical inference.
-Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference (LFI) methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-A flourishing new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional LFI approaches.
-However, to use these methods in theoretical neuroscience, we still must represent computation with data in some way.
-Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into an empirical framework.
+By solving the inverse problem in this way, scientific analysis of biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+
+As models of computation become more complex, two problems arise with existing approaches.
+First, the level of insight into computation afforded by parameter inference techniques \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}) is limited by the quantity and quality of available data.
+Even with vast amounts of data recorded with state-of-the-art equipment, neural activity in non-movement generating brain areas have been shown to strongly encode uninstructed behavior \cite{niell2010modulation, saleem2013integration, musall2019single}.
+Thus, the observational datasets we use to constrain our models of computation represent computations embedded in a sea of task-irrelevant factors like behavioral idiosyncrasies and likely other factors like hunger and thirst state, and abstract internal thoughts.
+Thus, when examining parameters inferred from neural datasets, rather than the specific computational properties -- the \emph{emergent phenomena} -- of neural circuit models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, it’s impossible to identify which aspect of the neural data is determining the inferred parameters: the computation or extraneous factors.
+
+%A common alternative is to synthesize data that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
+%In this case, mathematical criteria are being shoehorned into artificial datasets for the purpose of methodological compatability.
+
+Second, most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack a tractable likelihood function, which is necessary for statistical inference.
+Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional likelihood-free inference approaches.
+However, to use these methods in theoretical neuroscience, we still must represent computation with an explicit dataset in some way.
+Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into a framework for observational datasets.
 
@@
 First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion \cite{goldman2001global, gutierrez2013multiple}.
-Then, we show EPI's superior scaling properties by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
-In a model of primary visual cortex (V1) with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
+Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+In a model of primary visual cortex (V1) \cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
 Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
-Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials.
+Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials \cite{duan2018collicular}.
 EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
-We quantified parametric sensitivity and robustness in each regime by simply querying the deep probability distribution learned by EPI, which produced a mechanistic understanding of cortical responses in each regime.
+We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
 Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon.
@@
 In theoretical neuroscience, RNNs are optimized and then examined to show how dynamical systems could execute a given computation \cite{sussillo2014neural, barak2017recurrent}, but such biologically realistic constraints on connectivity are  ignored during optimization for practical reasons.
-In general, access to distributions of connectivity adhering to theoretical criteria like stable amplification, chaotic fluctuations \cite{sompolinsky1988chaos}, or low tangling \cite{russo2018motor} would add great scientific value and contextualization to existing research with RNNs.
+In general, access to distributions of connectivity adhering to theoretical criteria like stable amplification, chaotic fluctuations \cite{sompolinsky1988chaos}, or low tangling \cite{russo2018motor} would add scientific value and context to existing research with RNNs.
 Here, we use EPI to learn RNN connectivities producing stable amplification, and demonstrate the superior scalability and efficiency of EPI to alternative approaches.
@@
 \begin{center}
-\includegraphics[scale=.8]{figures/fig4/fig4.pdf}
+\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
 \end{center}
@@
 
-We consider a rank-2 RNN with N neurons having connectivity $W = UV^\top$
+We consider a rank-2 RNN with $N$ neurons having connectivity $W = UV^\top$
  and dynamics
@@
 Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
-$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W\top}{2}$.
+$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W^\top}{2}$.
 RNNs  with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will be stable with modest decay rate ($\text{real}(\lambda_1)$ close to its upper bound of 1) and exhibit modest amplification ($\lambda_1^s$ close to its lower bound of 1).
@@
 
-As the number of neurons $N$ in the RNN are scaled, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
+As the number of neurons $N$ in the RNN is scaled, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
 \ref{fig:LRRNN}A).
 It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:LRRNN}B).
-It is well known that ABC techniques struggle mightily in dimensions greater than about 30 \cite{sisson2018handbook}, yet we were careful to assess the scalability of the more comparable approach SNPE.
+It is well known that ABC techniques struggle in dimensions greater than about 30 \cite{sisson2018handbook}, yet we were careful to assess the scalability of the more comparable approach SNPE.
 Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
@@
 \ref{fig:RNN3}).
-From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both criteria in high-dimensions.
+From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both criteria in high dimensions.
 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig2/fig2.pdf}
+\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
 \end{center}
@@
 \end{equation}
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have multiply inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have multiple inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
 Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
@@
 \ref{fig:V1}E).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p=0$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
-\ref{fig:V1_3}, $p=.402$).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
+\ref{fig:V1_3}, $p=.40$).
 The strong compensatory influence of the E- and P-population input variability on excitatory varability is intriguing, since this circuit exhibited a paradoxical effect in the P-population (and no other inhibitory types) at 50\% contrast (Fig.
 \ref{fig:V1_3}) meaning that the E-population is P-stabilized.
-Future research may uncover a link between the populationsof stabilizations and compensatory interactions governing excitatory variability.
-
-By defining a clear theoretical question and executing the EPI optimization, we used EPI to reveal the quadratic relationship between $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ .
+Future research may uncover a link between the populations of stabilizations and compensatory interactions governing excitatory variability.
+
+EPI revealed the quadratic relationship between $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ .
 While this property is ultimately derivable, we show that with each additional neuron-type population, the formula becomes quite unruly and likely escapes comprehensible analysis in our case (see Section \ref{methods_V1_complexity}).
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
+\includegraphics[scale=0.8]{figures/fig4/fig4.pdf}
 \end{center}
@@
 \subsubsection{Stomatogastric ganglion}\label{methods_STG}
-%We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of network syncing in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
-%Each neuron's membrane potential $x_\alpha(t)$ for $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$ is the solution of the following differential equation:
-%\begin{equation} C_m \frac{dx_\alpha}{dt} = - \left[ h_{leak}(\mathbf{x}; \mathbf{z}) + h_{Ca}(\mathbf{x}; \mathbf{z}) + h_K(\mathbf{x}; \mathbf{z}) + h_{hyp}(\mathbf{x}; \mathbf{z}) + h_{elec}(\mathbf{x}; \mathbf{z}) + h_{syn}(\mathbf{x}; \mathbf{z})\right].
-%\end{equation} 
-%The membrane potential of each neuron is affected by the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents, respectively, which are functions of  all membrane potentials and the conductance parameters $\mathbf{z}$.
-%The capacitance of the cell membrane was set to $C_m = 1nF$.
-Specifically, the currents are the difference in the neuron's membrane potential and that current type's reversal potential multiplied by a conductance:
-%\begin{equation}  h_{leak}(\mathbf{x}; \mathbf{z}) = g_{leak} (x_\alpha - V_{leak}) 
-%\end{equation} 
-%\begin{equation}  h_{elec}(\mathbf{x}; \mathbf{z}) = g_{\text{el}} (x_\alpha^{post} - x_\alpha^{pre})
-%\end{equation} 
-%\begin{equation}  h_{syn}(\mathbf{x}; \mathbf{z}) = g_{syn} S_\infty^{pre} (x_\alpha^{post} - V_{syn}) \end{equation} 
-%\begin{equation}  h_{Ca}(\mathbf{x}; \mathbf{z}) = g_{Ca} M_\infty (x_\alpha - V_{Ca}) 
-%\end{equation} 
-%\begin{equation}  h_K(\mathbf{x}; \mathbf{z}) = g_K N (x_\alpha - V_K) 
-%\end{equation} 
-%\begin{equation}  h_{hyp}(\mathbf{x}; \mathbf{z}) = g_h H(x_\alpha - V_{hyp}).
-%\end{equation} 
-%The reversal potentials were set to $V_{leak} = -40mV$, $V_{Ca} = 100mV$, $V_K = -80mV$, $V_{hyp} = -20mV$, and $V_{syn} = -75mV$.  
-%The other conductance parameters were fixed to $g_{leak} = 1 \times 10^{-4} \mu S$. $g_{Ca}$, $g_{K}$, and $g_{hyp}$ had different values based on fast, intermediate (hub) or slow neuron.  
-%The fast conductances had values $g_{Ca} = 1.9 \times 10^{-2}$, $ g_K = 3.9 \times 10^{-2} $, and $ g_{hyp} = 2.5 \times 10^{-2} $.  
-%The intermediate conductances had values $g_{Ca} = 1.7 \times 10^{-2}$, $ g_K = 1.9 \times 10^{-2} $, and $ g_{hyp} = 8.0 \times 10^{-3} $.  
-%Finally, the slow conductances had values $g_{Ca} = 8.5 \times 10^{-3}$, $ g_K = 1.5 \times 10^{-2} $, and $ g_{hyp} = 1.0 \times 10^{-2} $.
-%
-%Furthermore, the Calcium, Potassium, and hyperpolarization channels have time-dependent gating dynamics dependent on steady-state gating variables $M_\infty$, $N_\infty$ and $H_\infty$, respectively:
-%\begin{equation}  M_{\infty} = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_1}{v_2} \right) \right) \end{equation}
-%\begin{equation}  \frac{dN}{dt} = \lambda_N (N_\infty - N)  \end{equation}
-%\begin{equation}  N_\infty = 0.5 \left( 1 + \tanh \left( \frac{x_\alpha - v_3}{v_4} \right) \right) \end{equation}
-%\begin{equation}  \lambda_N = \phi_N \cosh \left( \frac{x_\alpha - v_3}{2 v_4} \right) \end{equation}
-%\begin{equation}  \frac{dH}{dt} = \frac{\left( H_\infty - H \right)}{\tau_h} \end{equation}
-%\begin{equation}  H_\infty = \frac{1}{1 + \exp \left( \frac{x_\alpha + v_5}{v_6} \right)} \end{equation}
-%\begin{equation}  \tau_h = 272 - \left( \frac{-1499}{1 + \exp \left( \frac{-x_\alpha + v_7}{v_8} \right)} \right).
-% \end{equation}
-%where we set $v_1 = 0mV$, $v_2  = 20mV$, $v_3 = 0mV$, $v_4 = 15mV$, $v_5 = 78.3mV$,
-%$v_6 = 10.5mV$, $v_7 = -42.2mV$, $v_8 = 87.3mV$, $v_9 = 5mV$, and $v_{th} = -25mV$.  
-%
-%Finally, there is a synaptic gating variable as well:
-%\begin{equation} S_\infty = \frac{1}{1 + \exp \left( \frac{v_{th} - x_\alpha}{v_9} \right)} .
-%\end{equation}
-%When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.
-%
-%In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 200$ time steps of $dt = 25ms$.  
-%In EPI, since gradients are taken through the simulation process, the number of time steps are kept modest if possible. 
-%The chosen $dt$ and $T$ were the most computationally convenient choices yielding accurate frequency measurement.  
-%Poor resolution afforded by the discrete Fourier transform motivated the use of an alternative basis of complex exponentials to measure spiking frequency.  
-%Instead, we used a basis of complex exponentials with frequencies from 0.0-1.0 Hz at 0.01Hz resolution, $\Phi = \left[ 0.0, 0.01, ..., 1.0 \right]^\top$
-%
-%Another consideration was that the frequency spectra of the neuron membrane potentials had several peaks.  
-%High-frequency sub-threshold activity obscured the maximum frequency measurement in the complex exponential basis.  
-%Accordingly, subthreshold activity was set to zero, and the whole signal was low-pass filtered with a moving average window of length 20.  
-%The signal was subsequently mean centered.  
-%After this preprocessing, the maximum frequency in the filter bank accurately reflected the firing frequency.
-%
-%Finally, to differentiate through the maximum frequency identification, we used a sum-of-powers normalization.  
-%Let $\mathcal{X}_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-%The ``frequency identification" vector is 
-%\begin{equation}
-%v_\alpha = \frac{|\mathcal{X}_\alpha|^\beta}{\sum_{k=1}^N |\mathcal{X}_\alpha(k)|^\beta}.
-%\end{equation}
-%The frequency is then calculated as $\omega_\alpha = v_\alpha^\top \Phi$ with $\beta = 100$.
-%
-%Network syncing, like all other emergent properties in this work, are defined by the emergent property statistics and values.  
-%The emergent property statistics are the first and second moments of the firing frequencies. 
-%The first moments were set to 0.53Hz, and the second moments were set to 0.025Hz$^2$:
-%\begin{equation}
-%E \begin{bmatrix} \omega_{\text{f1}} \\ \omega_{\text{f2}} \\ \omega_{\text{hub}} \\ \omega_{\text{s1}} \\ \omega_{\text{s2}} \\ (\omega_{\text{f1}} - 0.53)^2 \\ (\omega_{\text{f2}} - 0.53)^2 \\ (\omega_{\text{hub}} - 0.53)^2 \\ (\omega_{\text{s1}} - 0.53)^2 \\ (\omega_{\text{s2}} - 0.53)^2  \end{bmatrix} = \begin{bmatrix} 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.53 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \\ 0.025^2 \end{bmatrix}
-%\end{equation}
-%for the EPI distribution shown in Fig. \ref{fig:STG}B.
-%Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,500 iterations (see Section \ref{methods_AL_opt}).  
-%The optimization converged after four epochs (Fig. S4).
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
-%\end{center}
-%\begin{flushleft}
-%{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
-%A. Entropy throughout optimization. 
-%B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
-%Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.
-%C. The second moment emergent property statistics converge to the emergent property values.}
-%\end{flushleft}
-%\end{figure}
-%
-%For EPI in Fig 2C, we used a real NVP architecture with four masks and two layers of 10 units per mask, and batch normalization momentum of 0.99 mapped onto a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0,4]$.
-%We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=300$, set $\nu = 0.1$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce an isotropic 
-%Gaussian with mean $\mu_{\text{init}} = [6, 2]$ with standard deviation $\sigma_{\text{init}} = 0.5$.
-%
-%We calculated the Hessian at the mode of the inferred EPI distribution.  
-%The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
-%With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (high magnitude eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-%In Figure \ref{fig:STG}B, the eigenvectors of the Hessian $v_1$ and $v_2$ are shown evaluated at the mode of the distribution.  
-%The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.8$ and $\lambda_2 = -2.27$.
-%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
-%Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
-%The sensitivity slope coefficient of eigenvector $v_1$ with respect to network syncing was significant ($\beta = 4.82 \times 10^{-2}$, $p<10^{-4}$).
-%In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
-\ref{fig:STG}B, inset).
-%The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
-
 We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of intermediate hub frequency in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figRNN2/figRNN2.pdf}
+\includegraphics[scale=0.7]{figures/figRNN1/figRNN1.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/figRNN1/figRNN1.pdf}
+\includegraphics[scale=0.8]{figures/figRNN2/figRNN2.pdf}
 \end{center}

### 2021-04-15 written/eLifeRevision/Bittner_eLife2020.tex
intro edits

@@
 
-As models of computation become more complex, two problems arise with existing approaches.
-First, the level of insight into computation afforded by parameter inference techniques \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}) is limited by the quantity and quality of available data.
-Even with vast amounts of data recorded with state-of-the-art equipment, neural activity in non-movement generating brain areas have been shown to strongly encode uninstructed behavior \cite{niell2010modulation, saleem2013integration, musall2019single}.
-Thus, the observational datasets we use to constrain our models of computation represent computations embedded in a sea of task-irrelevant factors like behavioral idiosyncrasies and likely other factors like hunger and thirst state, and abstract internal thoughts.
-Thus, when examining parameters inferred from neural datasets, rather than the specific computational properties -- the \emph{emergent phenomena} -- of neural circuit models \cite{hopfield1982neural, sompolinsky1988chaos, tsodyks1997paradoxical, wong2006recurrent}, it’s impossible to identify which aspect of the neural data is determining the inferred parameters: the computation or extraneous factors.
-
-%A common alternative is to synthesize data that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
-%In this case, mathematical criteria are being shoehorned into artificial datasets for the purpose of methodological compatability.
-
-Second, most neural circuit models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation; they lack a tractable likelihood function, which is necessary for statistical inference.
-Therefore, the most popular approaches to theoretical inverse problems have been likelihood-free inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+While theoretical neuroscience is concerned with how model parameters govern  computational properties, existing methodology for statistical inference in neuroscience \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}) requires that parameters be conditioned on an explicit dataset.
+The scientific insight for a model of computation is then limited by the quantity and quality of available neural data.
+Even with a vast amount of high-quality recordings, neural data often reflect uninstructed behaviors \cite{niell2010modulation, saleem2013integration, musall2019single}, and thus may only reflect the computation of interest amidst a sea of task-irrelevant factors.
+A common alternative is to synthesize an explicit dataset that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
+In this case, well-defined computational properties are being shoehorned into artificial datasets for the purpose of methodological compatibility.
+
+Another key challenge is that as models of computation become more complex, statistical inference becomes intractable.
+Such mechanistic models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation \cite{dayan2003theoretical, izhikevich2007dynamical}; they lack a tractable likelihood function, which is necessary for statistical inference.
+Therefore, the most popular approaches to parameter inference in mechanistic models have been likelihood-free inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
 A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional likelihood-free inference approaches.
-However, to use these methods in theoretical neuroscience, we still must represent computation with an explicit dataset in some way.
+However, to use these methods in theoretical neuroscience, we must represent computation with an explicit dataset in some way.
 Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into a framework for observational datasets.
@@
 %We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
-As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets exhibiting some mathematical criteria.
-EPI constrains the predictions of the inferred parameter distribution to produce the emergent property, which requires a variant of probabilistic inference methods \cite{loaiza2017maximum}.
-Importantly, EPI uses deep learning to make rich, flexible approximations to the parameter distribution \cite{rezende2015variational} that produces an emergent property.
-The structure captured by these deep probability distributions are scientifically valuable, revealing the sensitivity and robustness of the emergent property to different parameter combinations.
-Perhaps most powerfully, EPI facilitates inference in mechanistic models, allowing theorists to capture rich parametric structure in biologically realistic models that is conditioned upon the emergent phenomena of interest.
+%
+%These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
+%As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets exhibiting some mathematical criteria.
+%EPI constrains the predictions of the inferred parameter distribution to produce the emergent property, which requires a variant of probabilistic inference methods \cite{loaiza2017maximum}.
+%Importantly, EPI uses deep learning to make rich, flexible approximations to the parameter distribution \cite{rezende2015variational} that produces an emergent property.
+%The structure captured by these deep probability distributions are scientifically valuable, revealing the sensitivity and robustness of the emergent property to different parameter combinations.
+%Perhaps most powerfully, EPI facilitates inference in mechanistic models, allowing theorists to capture rich parametric structure in biologically realistic models that is conditioned upon the emergent phenomena of interest.
+
+To address the methodological incongruity between explicit datasets and emergent properties, we present a statistical inference method for conditioning parameters of neural circuit models directly on computation.
+In this work, we define computation by an emergent property, which is a statistical description of the phenomena to be produced by the neural circuit model.
+In emergent property inference (EPI), we infer the distribution of model parameters that produce this emergent property.
+With EPI, parameters are conditioned directly on an implicit dataset defined by the computation of interest.
+By using recent optimization techniques \cite{loaiza2017maximum}, EPI uses deep learning to make rich, flexible approximations to the parameter distributions \cite{rezende2015variational}, the structure of which reveals scientific insight about how parameters govern the emergent property.
 

### 2021-04-16 written/eLifeRevision/Bittner_eLife2020.tex
STG edits

@@
 
-%Second, once the model is selected, one defines the emergent property, the measurable signal of scientific interest.  
 Second, once the model is selected, one must specify what the model should produce.
-In typical statistical inference, this is a dataset -- either collected experimentally or constructed by scientists to fit this emprical paradigm.
-In EPI, a different approach is taken, in which we define an emergent property: a set of mathematical criteria to be obeyed by the datasets predicted by the inferred distribution.
+In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model \ref{fig:STG}B.
+An emergent property studied by Guttierez et al. of this stochastic model is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
+This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
+Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and and slow (.68Hz) frequencies, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
 %To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-In the STG example, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model \ref{fig:STG}B.
-%One such emergent property is the phenomenon of \emph{unified intermediacy} -- in certain parameter regimes, the frequency of all neurons match at an intermediate firing rate.
-An interesting emergent property of this stochastic model is when the hub neuron fires at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
-This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
 
@@
 % Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-Third, the model parameters producing these outputs are inferred.
-Most often, brute-force parameter sweeps or rejection sampling techniques \cite{liepe2014framework} are used to identify parameters whose model simulations are close to data or some desired feature.
 %Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
 %In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-In this last step lies the opportunity for a paradigmatic shift away from empirical data-oriented representations of model output.
-By precisely quantifying the emergent property of interest as a statistical feature of the model, we can infer a probability distribution over parameter configurations that produce this emergent property.
-This unlocks the deep probabilistic modeling toolkit for treating theoretical inverse problems.
+Third, the model parameters producing these outputs are inferred.
+To infer the STG parameters of intermediate hub frequency with existing methodology, we  need an explicit dataset: experimentally recorded or synthesized.
+By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
+EPI learns a probability distribution of model parameters constrained to produce the emergent property.
+In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
+%Most often, brute-force parameter sweeps or rejection sampling techniques \cite{liepe2014framework} are used to identify parameters whose model simulations are close to data or some desired feature.
 
@@
 \ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
-Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig.
-\ref{fig:STG}E).
-This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution producing the emergent property. 
-In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) (Fig.
+Before presenting technical details (in the following section), let us understand emergent property inference schematically.
+EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig. \ref{fig:STG}E).  
+This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution that produces the emergent property. 
+%In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
+%Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) (Fig.
 \ref{fig:STG}E, arrows) agree with error contours (Fig.
@@
 %First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
-First, as is typical, we consider the model as a coupled set of differential equations \cite{gutierrez2013multiple}.  
-In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
+First, as is typical, we consider the model as a coupled set of differential equations.  
+In this STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
 \begin{equation} 
@@
 \end{equation} 
-where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG} for more detail).
+where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{morris1981voltage, gutierrez2013multiple} (see Section \ref{methods_STG} for more detail).
 
 %Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
 %Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
@@
 %Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz.  
 %Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
-Mathematically, we achieve this with two constraints: by fixing the mean hub frequency over the inferred parameter distribution of $\mathbf{z}$ and its resulting simulations $\mathbf{x}$ to 0.55Hz,
+
+%Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
+%Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz.  
+%Mathematically, we achieve this with two constraints: by fixing the mean hub frequency over the inferred parameter distribution of $\mathbf{z}$ and its resulting simulations $\mathbf{x}$ to 0.55Hz,
+%\begin{equation}\label{eq:EP_STG1}
+% \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix}
+%\end{equation}
+%and requiring that the variance of the hub frequency over the produced simulations is small
+%\begin{equation}\label{eq:EP_STG2}
+%\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix}.
+%\end{equation}
+%This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the inferred parameter distribution.
+%By constraining the means and variances of emergent property statistics over $\mathbf{z}$ as well as the stochasticity of $\mathbf{x}$, we can precisely control the behavior that the inferred distribution that EPI infers.
+%In general, an emergent property
+%\begin{equation}\label{eq:EP_STG}
+%\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
+%\end{equation}
+%defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
+%The choice of $\bm{\sigma}^2$ predicates the degree of variability around the mean $\bm{\mu}$ that is consistent with the emergent property.
+
+Second, we stipulate that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
+We stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
+Mathematically, we define this emergent property with two statistical constraints: that the mean hub frequency is 0.55Hz,
 \begin{equation}\label{eq:EP_STG1}
- \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix}
-\end{equation}
-and requiring that the variance of the hub frequency over the produced simulations is small
+ \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55 
+\end{equation}
+and that the variance of the hub frequency is moderate
 \begin{equation}\label{eq:EP_STG2}
-\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix}.
-\end{equation}
-This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the inferred parameter distribution.
-By constraining the means and variances of emergent property statistics over $\mathbf{z}$ as well as the stochasticity of $\mathbf{x}$, we can precisely control the behavior that the inferred distribution that EPI infers.
-In general, an emergent property
+\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2.
+\end{equation}
+The hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$  and the distribution of the data $\mathbf{x}$ that those parameters produce.
+Formally, the emergent property is the collection of these two constraints
 \begin{equation}\label{eq:EP_STG}
-\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
-\end{equation}
-defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
-The choice of $\bm{\sigma}^2$ predicates the degree of variability around the mean $\bm{\mu}$ that is consistent with the emergent property.
-
-Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
+\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] = 0.55,~~\text{Var}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] =0.025^2.
+\end{equation}
+In general, an emergent property is a collection of first-, second- and higher moments of  statistics that together define the phenomena.
+
+Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$ that produces the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
 This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, dinh2017density, papamakarios2017masked} (Figure \ref{fig:STG}D, EPI box).
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ to a suitable complicated distribution (see Section \ref{methods_NF} for more details).
-Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (or ``entropic") distribution, which is the same choice made in Bayesian inference (see Section \ref{methods_VI}).
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ to a suitably complicated distribution (see Section \ref{methods_NF} for more details).
+Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in Bayesian posterior inference (see Section \ref{methods_VI}).
 %\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
 In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI})
-The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, to emphasize that we have conditioned our parameter distribution on emergent property $\mathcal{X}$.
+The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since it is conditioned upon emergent property $\mathcal{X}$. 
+This is meant to share the same notation as a posterior distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathbf{x})$ that is conditioned upon an explicit dataset.
 

### 2021-04-16 written/eLifeRevision/Bittner_eLife2020.tex
added para on SNPE vs EPI in the RNN section

@@
 Such mechanistic models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation \cite{dayan2003theoretical, izhikevich2007dynamical}; they lack a tractable likelihood function, which is necessary for statistical inference.
-Therefore, the most popular approaches to parameter inference in mechanistic models have been likelihood-free inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional likelihood-free inference approaches.
+Therefore, the most popular approaches to parameter inference in mechanistic models have been simulation-based inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional simulation-based inference approaches.
 However, to use these methods in theoretical neuroscience, we must represent computation with an explicit dataset in some way.
@@
 Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Dimension of sensitivity $\mathbf{v}_1$ (solid) and degeneracy $\mathbf{v}_2$.
+Dimension of sensitivity $\mathbf{v}_1$ (solid) and degeneracy $\mathbf{v}_2$ (dashed).
 %(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
@@
 
+EPI produces parameter distributions that can be queried for scientific insight.
+The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices exemplary of the emergent property (Fig. \ref{fig:STG}E yellow star).
+As probability in the EPI inferred distribution decreases, the emergent property deteriorates.
+Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ does not change will not disturb the emergent property, making this parameter combination \emph{degenerate} with respect to the emergent property.
+In contrast, if $\mathbf{z}$ is perturbed along a dimension that strongly decreases $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we call that parameter combination \emph{sensitive}.
+By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative the eigenvalue, the more sensitive.
+Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$, smaller eigenvalue) and robustness ($v_2$, greater eigenvalue) (Fig.
+\ref{fig:STG}E, arrows) agree with error contours (Fig.
+\ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
+\ref{fig:STG}F activity traces).
+Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (when the correct architecture class is chosen, see Section \ref{methods_NF}).
+
+In the following sections, we demonstrate EPI on three neural circuit models across ranges of biological realism, neural system function, and network scale.
+First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of recurrent neural network (RNN) connectivities that exhibit amplified, yet stable responses.
+Also in this RNN example, we emphasize that EPI is the only technique that controls the predictions made by the inferred parameter distribution.
+Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI captures a curved manifold of parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
+Finally, in a model of superior colliculus \cite{duan2018collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of sensitivity ($\mathbf{v}_1(\mathbf{z})$) to mechanistically characterize each regime.
+
 %Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$, and insist that samples from this distribution produce the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
@@
 
-The major scientific value of EPI is in the rich, queryable structure of these deep probability distributions.
-The probabilities of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ are the densities of these parameters in the distribution producing the emergent property.
-The greatest probabilities (the modes) indicate prototypical parameter configurations, and the manner in which probabilities change away from the modes shows how different parameter combinations preserve or diminish the emergent property.
-The dimensions of greatest sensitivity (e.g.
+%The major scientific value of EPI is in the rich, queryable structure of these deep probability distributions.
+%The probabilities of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ are the densities of these parameters in the distribution producing the emergent property.
+%The greatest probabilities (the modes) indicate prototypical parameter configurations, and the manner in which probabilities change away from the modes shows how different parameter combinations preserve or diminish the emergent property.
+%The dimensions of greatest sensitivity (e.g.
 Fig.
@@
 1E dashed arrow) can be measured directly from the second order derivative of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ called the ``Hessian."
-Around the mode, eigenvalues of the Hessian are negative; probabilities decrease locally in all directions away from the mode.
-The eigenvector with most negative eigenvalue is the parameter combination causing probability to decrease the fastest, making it the most sensitive dimension.
-Likewise, the flattest eigenvector, corresponding to the least negative eigenvalue, points in the most degenerate dimension.
-Once an EPI distribution has been inferred, this second order derivative requires trivial computation (when correct architecture class is chosen, see Section \ref{methods_NF}).
-
-In the following sections, we showcase the versatility of EPI for scientific analysis on three neural circuit models across ranges of biological realism, neural system function, and network scale.
-First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of RNN connectivities that exhibit amplified, yet stable responses.
-Also in this RNN example, we emphasize that EPI is the only technique that controls the predictions made by the inferred parameter distribution.
-Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how to gain insight by comparing multiple inferred distributions.
-Finally, we used EPI to capture subtle parametric structure allowing the mechanistic characterization of multiple parametric regimes of superior colliculus activity in a model of task switching \cite{duan2018collicular}.
-This work is the first to produce this level of theoretical insight via the quantification and examination of the intricate structure captured by deep probability distributions.
+%Around the mode, eigenvalues of the Hessian are negative; probabilities decrease locally in all directions away from the mode.
+%The eigenvector with most negative eigenvalue is the parameter combination causing probability to decrease the fastest, making it the most sensitive dimension.
+%Likewise, the flattest eigenvector, corresponding to the least negative eigenvalue, points in the most degenerate dimension.
+%Once an EPI distribution has been inferred, this second order derivative requires trivial computation (when correct architecture class is chosen, see Section \ref{methods_NF}).
 
@@
 
-For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative likelihood-free inference approaches.
+For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
 We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
 SMC-ABC is a rejection sampling approach that SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
-Unlike EPI, these statistical inference techniques do not control the mean or variance of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
+Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
 $N$ and $g$, Fig.
@@
 
-Through this example, we have shown that EPI can be used for well-controlled insight into RNNs with respect to their theoretical properties.
-EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each optimization iteration.
-This agrees with recent speculation that such gradient information could improve the efficiency of LFI techniques \cite{cranmer2020frontier}.
-While scaling to high dimensions is important, we show in the next two sections how insight can be gained by inspecting structure in lower dimensional parameter distributions.
+EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each iteration of optimization.
+This agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}.
+Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
+Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
+This can be invaluable for understanding how RNNs produce complex emergent phenomena.
+Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
+However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
+The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
+burst duration and phase gap) are not easily calculated or approximated with differentiable functions.
+In such cases, gradient-free approaches like SNPE have proved to be a powerful option \cite{gonccalves2019training}.
+In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.
 
@@
 
-Likelihood-free (or ``simulation-based") inference (LFI) \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
-The most prevalent approach to LFI is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
+Simulation-based inference \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
+The most prevalent approach to simulation-based inference is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
 The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
@@
 For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
-EPI infers parameters in likelihood-free models using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational approximation.
+EPI infers parameters in mechanistic models using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational approximation.
 The deep neural network of EPI defines the parametric form of the distribution approximation.
@@
 The method EPI differs from SNPE in some key ways.
-SNPE belongs to a ``sequential" class of recently developed LFI methods in which two neural networks are used for posterior inference.
+SNPE belongs to a ``sequential" class of recently developed simulation-based inference methods in which two neural networks are used for posterior inference.
 This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
@@
 Since EPI distributions are defined by the mean $\bm{\mu}$ of their statistics, there is the well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} that prohibits an amortization based approach.
-However, we have shown that the same two-network architecture of the sequential LFI methods can be used for amortized inference in intractable exponential family posteriors using their natural parameterization \cite{bittner2019approximating}.
-
-Finally, one important differentiating factor between EPI and sequential LFI methods is that EPI leverages gradients $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ during optimization.
+However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors using their natural parameterization \cite{bittner2019approximating}.
+
+Finally, one important differentiating factor between EPI and sequential simulation-based inference methods is that EPI leverages gradients $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ during optimization.
 These gradients can improve convergence time and scalability, as we have shown on an example conditioning low-rank RNN connectivity on the property of stable amplification (see Section \ref{results_LRRNN}).
 With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these model gradients when they are tractable.
-Sequential LFI techniques may be better suited for sceintific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable: when there is a nondifferentiable model or it requires lengthy simulations.
-However, the sequential LFI techniques cannot constrain the predictions of the inferred distribution in the manner of EPI.
+Sequential simulation-based inference techniques may be better suited for sceintific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable: when there is a nondifferentiable model or it requires lengthy simulations.
+However, the sequential simulation-based inference techniques cannot constrain the predictions of the inferred distribution in the manner of EPI.
 
@@
 
-We compared EPI to two alternative likelihood-free inference (LFI) techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
+We compared EPI to two alternative simulation-based inference techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
 Approximate Bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parameters $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
@@
 
-SNPE, the next LFI approach in our comparison, is far more similar to EPI.
+SNPE, the next approach in our comparison, is far more similar to EPI.
 Like EPI, SNPE treats parameters in mechanistic models with deep probability distributions, yet the two learning algorithms are categorically different.

### 2021-04-17 written/eLifeRevision/Bittner_eLife2020.tex

@@
 The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not easily calculated or approximated with differentiable functions.
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
 In such cases, gradient-free approaches like SNPE have proved to be a powerful option \cite{gonccalves2019training}.
@@
 Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
-\textbf{B}. Steady-state responses of the SSN model (deterministic, $\bm{\sigma}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
-\textbf{C}.
-Transient network responses of the SSSN model at 50 \% contrast.
-(Left) Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise realization.
-(Right) Mean (solid line) and standard deviation (shading) of responses.
+\textbf{B}. Transient network responses of the SSSN model. Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise $\bm{\epsilon}$. 
+\textbf{C}. Mean (solid line) and standard deviation $s_E(\mathbf{x}; \mathbf{z})$ (shading)  across 100 trials.
 \textbf{D}. EPI distribution of noise parameters $\mathbf{z}$ conditioned on E-population variability.
@@
 \textbf{F}.
-The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
 }
@@
 Each population also experiences an external input $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
-To build on previous work, we model visual contrast-dependent input to the E- and P-populations $\mathbf{h} = \mathbf{b} + c\mathbf{h}_c$.
-There is also an additive noisy input $\bm{\epsilon}$ parameterized by variances for each neuron type population $\mathbf{z} = \bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
-This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ then the population rate, allowing fluctuations around a stimulus-dependent steady-state
+There is also an additive noisy input $\bm{\epsilon}$ parameterized by variances for each neuron type population $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
+This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Fig. \ref{fig:V1}B).
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} 
 \begin{equation}
@@
 \end{equation}
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} generalized to have multiple inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
+generalized to have multiple inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
 Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
-Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure} (Fig.
-\ref{fig:V1}B), and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability at 50\% contrast.
-
-We quantify different levels $y$ of E-population variability with the emergent property
+Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure}, and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability.
+
+We quantify levels $y$ of E-population variability with the emergent property
 \begin{equation}\label{eq:EP_V1}
@@
 where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-
-We ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
+In the following analyses, we compare levels of 5Hz and 10Hz, and select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.
+
+First, we ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
 From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
@@
 Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions (Fig. \ref{fig:V1}D, bottom-right).
-Together, these observations imply a curved path of parametric degeneracy with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
-The dimensions of sensitivity conferred by EPI and this plain visual structure suggest a quadratic relationship in the emergent property statistic $s_E(\mathbf{x}; \mathbf{z})$ and parameters $\mathbf{z}$, which is preserved at a greater level of variability $\mathcal{X}(10\text{ Hz})$ (Fig.
-\ref{fig:V1}E).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
-\ref{fig:V1_3}, $p=.40$).
-The strong compensatory influence of the E- and P-population input variability on excitatory varability is intriguing, since this circuit exhibited a paradoxical effect in the P-population (and no other inhibitory types) at 50\% contrast (Fig.
-\ref{fig:V1_3}) meaning that the E-population is P-stabilized.
-Future research may uncover a link between the populations of stabilizations and compensatory interactions governing excitatory variability.
-
-EPI revealed the quadratic relationship between $s_E(\mathbf{x}; \mathbf{z})$ and $\mathbf{z}$ .
-While this property is ultimately derivable, we show that with each additional neuron-type population, the formula becomes quite unruly and likely escapes comprehensible analysis in our case (see Section \ref{methods_V1_complexity}).
-This emphasizes the need for streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes prohibitive.
+Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
+
+Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensatorily decreases to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
+This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
+\ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
+The strong interactive influence of E- and P-population input variability on excitatory varability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
+\ref{fig:V1_3}), meaning that the E-population is P-stabilized.
+Future research may uncover a link between the population of network stabilization and compensatory interactions governing excitatory variability.
+
+EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
+We show that with each neuron-type population added to this E-I model, calculations of excitatory variability with respect to noise parameters become unruly and challenging to work with (see Section \ref{methods_V1_complexity}).
+This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.
+While EPI can be used to investigate fundamental aspects of sensory processing, in the next section, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to  behavioral experiments.
 
@@
 These results motivated a model of SC that is a four-population dynamical system with functionally-defined neuron-types.
-Here, our goal is to understand how connectivity in this circuit model governs the ability to switch tasks rapidly.
+Here, our goal is to understand how connectivity in this circuit model governs the ability to perform rapid task switching: to respond with satisfactory accuracy in both tasks on randomly interleaved trials.
 
@@
 
-We define the computation of rapid task switching as accurate execution of each task.
-Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
-We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$ be $75\%$ with variance $7.5\%^2$.
+Rapid task switching is formalized mathematically as an emergent property with two statistics: accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$.
+We stipulate that accuracy be on average $.75$ in each task with variance $.075^2$
 \begin{equation}\label{eq:SC_EP}
 \begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 7.5\%^2 \\ 7.5\%^2  \end{bmatrix}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} .75 \\ .75 \end{bmatrix}  \\ 
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} .075^2 \\ .075^2  \end{bmatrix}.
 \end{split}
 \end{equation}
-
-The EPI inferred distribution  (Fig.
-\ref{fig:SC}C) produces task accuracies (Fig.
-\ref{fig:SC}C, middle-left) according to our mathematical definition of rapid task switching (Equation \ref{eq:SC_EP}).
-The patterns of connectivity that govern each task accuracy are nonlinear (Fig. \ref{fig:SC1}A-B); they are not captured well by linear correlations (Fig. \ref{fig:SC1}C).
-For example, there appear to be two regimes of connectivity: the local structure of the EPI distribution changes dramatically after crossing a threshold of $sW$ (Fig.
-\ref{fig:SC1}A $sW$-$hW$ marginal distribution).
-Not only has EPI captured this intricate, nonlinear distribution, we can use the distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ returned by EPI to understand these two parametric regimes of SC connectivity.
-
-To distinguish these two parts of the distribution, we point out that for many fixed values of parameter $hW$, there are two modes in the EPI distribution.  
-Thus, by fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ in the parameter spaces proximal to each mode, we can identify a set of modes $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for each putative regime $r \in \{1, 2\}$ (see Section \ref{methods_SC}).
+75\% accuracy is a realistic level of performance in each task, and with the chosen variance, inferred models will not exhibit fully random responses (50\%), nor perfect performance (100\%).
+
+The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, middle-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
+This parameter distribution has intricate structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
+Specifically, the shape of the EPI distribution changes dramatically on different sides of parameter space.
+This is most saliently pointed out in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
+Not only has EPI captured this complicated distribution of connectivities producing rapid task switching, we can query the EPI distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.
+
+To distinguish these two regimes, we use the EPI distribution to identify two sets of modes.
+By fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we arrive at two solutions $\mathbf{z}^*(hW_{\text{fixed}}, r)$ where regime $r \in [1,2]$, and regime 1 is that of greater $sW$ (see Section \ref{methods_SC}).
 As $hW_{\text{fixed}}$ increases, the modes coalesce to intermediate parameters reflecting a transition between the two sets of modes (Fig. \ref{fig:SC4} top).
-However, the sensitivity dimensions of these modes $\mathbf{v}_1$ (refer to Section \ref{results_dgm}), which reflect the structure of the EPI distribution around each mode, are different across putative regime, yet consistent across $hW_{\text{fixed}}$.
-This categorical difference in sensitivity dimension across the two sets of modes shows that they indeed represent two different regimes of computation in which connectivity governs computation in different ways.
-
-To understand how SC connectivity governs computation in each regime, we can examine how perturbations along $\mathbf{v}_1(\mathbf{z}^*)$ affect task performance in each regime; we measure task accuracy for connectivity changes in the dimension that rapid task switching is most sensitive.
-While the monotonic increase in Pro accuracy with $\mathbf{v}_1$ perturbation is largely unaffected by regime (Fig.
-\ref{fig:SC}D, top-left), we see a stark difference in Anti accuracy: Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig.
-\ref{fig:SC}D, top-right).
-These two rapid task switching pathologies are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(\cdot, 1))$ vs $\mathbf{v}_1(\mathbf{z}^*(\cdot, 2))$) and explain the sharp change in local structure of the EPI distribution.
-
-To understand the connectivity mechanisms that distinguish these two regimes, we perturb connectivity at each mode in dimensions that have well defined roles in processing for the Pro and Anti tasks.
-A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$.
-These eigenmodes have intuitive roles in processing in each task, and are accordingly named the \textit{all}, \textit{side}, \textit{task}, and \textit{diag} eigenmodes (see Section \ref{methods_SC}).
-Furthermore, the parameter dimension $\mathbf{v}_a$ ($a \in \{\text{all}, \text{side}, \text{task}, \text{ and diag}\}$) that increases the eigenvalue of connectivity $\lambda_a$ is $\mathbf{z}$-invariant (unlike the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$) and $\mathbf{v}_a \perp \mathbf{v}_{b \neq a}$.
-Thus, by changing the degree of amplification of each processing mode by  perturbing $\mathbf{z}$ along $\mathbf{v}_a$, we can elicit the differentiating properties of the two regimes.
-
-Through these connectivity perturbation analyses, we found that increasing $\lambda_{\text{task}}$ strongly reduced Pro accuracy in regime 1, yet strongly reduced Anti accuracy in regime 2.
-This suggests that stronger task representations can inhibit both Pro and Anti task performance in different contexts.
-Furthermore, changing $\lambda_{\text{task}}$ in either direction decreases Anti performance in regime 1, showing that Anti task performance in regime 1 is dependent on a specific level of task representation.
-We also found that with increasing $\lambda_{\text{diag}}$, Pro accuracy increased in both regimes, but there were opposite effects on Anti accuracy.
-In regime 1, stronger amplification of diagonal population patterns decreased Anti accuracy, while in regime 2 accuracy increased.
-These findings give us an understanding of the mechanistic differences in computation enabling rapid task switching in each regime.
+By using EPI to connect these two regimes through this transitionary region of parameter space, we can explore what distinguishes the two regimes by stepping from the prototypical connectivity of regime 1 to that of regime 2.
+
+While the connectivities gradually coalesce to the transitionary part of parameter space, the sensitivity dimensions $\mathbf{v}_1(\mathbf{z})$ are categorically different across regimes (Fig.
+\ref{fig:SC4} bottom).
+The sensitivity dimension identifies the parameter combination which causes the emergent property to diminish with the shortest perturbation.
+Since the two regimes have different $\mathbf{v}_1(\mathbf{z})$, this suggests they have different pathologies in their connectivity.
+By perturbing connectivity in each regime along the sensitivity dimension, we can get a sense of the differing nature of these pathologies.
+
+When perturbing connectivity along the sensitivity dimension, Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
+However, there is a stark difference between regimes in Anti accuracy. 
+Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, top-right).
+These distinct pathologies of rapid task switching are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(r=1))$ vs $\mathbf{v}_1(\mathbf{z}^*(r=2))$) and explain the sharp change in local structure of the EPI distribution.
+With further perturbation analyses along dimensions of connectivity having established effects on processing, we are able to distinguish these two regimes with mechanistic insights (Fig.
+\ref{fig:SC}D, middle, bottom) (Section \ref{methods_SC}).
 
@@
 
-Similarities across Pro and Anti trials in choice period responses following delay period inactivation (Fig.
-\ref{fig:SC5}A) suggested that connectivity patterns inducing greater Pro task accuracy increase error in delay period inactivated Anti trials (Fig.
+Following delay period inactivation, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC5}A).
+We interpreted these similarities to suggest that delay period inactivation flips the internal representation of task in the model.
+Connectivity patterns inducing greater Pro task accuracy would antagonistically reduce accuracy in Anti trials subject to delay period inactivation in which this flip in task representation occurs.
+In fact, across connectivities in the EPI inferred distribution, there was strong anti-correlation between Pro task accuracy $p_P$ and Anti task accuracy during delay period silencing $p_{A, opto}$ (Fig.
 \ref{fig:SC_opto}B).
-The strong anticorrelation between $p_P$ and $p_{A, opto}$ across EPI inferred connectivities led to the following hypothesis about each connectivity regime: the sensitivity dimension of each regime decreases $p_{A,opto}$ irrespective of its effect on $p_A$, since both $\mathbf{v}_1$ and $\mathbf{v}_2$ increase $p_P$.
-Indeed, in regimes 1 and 2 where sensitivity dimensions elicit different responses in $p_A$, $p_{A,opto}$ decreases since the connectivity changes enhancing $p_P$ exacerbate Anti trial error (Fig.
-\ref{fig:SC_opto}C).
-Thus, the altered state caused by delay period silencing makes the connectivity governing $p_P$ more influential on Anti accuracy than the connectivity governing $p_A$.
+
+We explored whether this antagonism between $p_P$ and $p_{A, opto}$ was present in each connectivity regime.
+As in the previous section, we tested perturbations in the dimension of sensitivity $\mathbf{v}_1(\mathbf{z})$ from regime 1 to regime 2.
+In both regimes, we recall that perturbations along $\mathbf{v}_1(\mathbf{z})$ yield increasing $p_P$.
+We found that only in regime 2, does $p_{A,\text{opto}}$ decrease as $p_P$ increases(Fig. \ref{fig:SC_opto}C).
+This context further distinguishes the two regimes discovered with EPI.
 
@@
 In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This EPI distribution revealed two regimes of rapid task switching, which we characterized using the probabilistic toolkit EPI seemlessly affords.
+This EPI distribution revealed two regimes of rapid task switching, which we characterized using the probabilistic toolkit afforded by EPI.
 We found that both of these parametric regimes identified by EPI reproduce results from optogenetic inactivation experiments: when activity is silenced during the delay period, only Anti accuracy suffers.
-We then identified the connectivity mechanisms governing Anti accuracy during delay period silencing, and showed that they are regime invariant.
+We then identified connectivity mechanisms governing Anti accuracy during delay period silencing, and showed the parametric regime they describe.
 
@@
 In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation, rather than measuring structure in some noisy observed dataset.
-The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
+%The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
 In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to emergent phenomena in  interpretable models of neural computation (see Section \ref{methods_related}).
@@
 SNPE \cite{gonccalves2019training} and other sequential techniques for inference in mechanistic models developed along with EPI (see Section \ref{methods_related}) have been used for posterior inference with noisy experimental datasets.
-On the other hand, EPI is a deep inference technique designed to condition upon mathematical criteria, such that the parameter distribution only produces  the specified \textit{emergent properties} of computation.
+On the other hand, EPI is a deep inference technique designed to condition directly on emergent properties, such that the parameter distribution only produces the computation of interest.
 EPI is thus ideally suited for questions in theoretical neuroscience, and we show that it has superior scaling properties to these other inference techniques (see Section \ref{results_LRRNN}).
-
-In this work, we prove out the utility of deep probability distributions for theoretical neuroscience.
-While previous work has used SNPE to obtain flexible posterior approximations in mechanistic models conditioned on experimental datasets, we use the rich structure captured by deep probability distributions in EPI to gain new theoretical insights.
-This is first done in a complex model of V1, where we combine the modeling advancements \cite{litwin2016inhibitory, hennequin2018dynamical}, which make analytic characterization of excitatory variability very complicated.
-There, EPI clearly and simply revealed the parametric structure of input variability across neuron-type populations that governed excitatory variability, which has implications on the dimensionality and nature information transmission in visual cortex.
-
-Finally, we uesd EPI to identify two distinct regimes of SC connectivity that enabled rapid task switching.
-By systematically characterizing the local structure of the inferred distribution using the analytic capabilities deep probability distributions, we discerned a mechanistic understanding of each computational regime.
-Each of these regimes reproduced effects from optogenetic experiments \cite{duan2015requirement}, suggesting that both are biologically plausible.
-These analyses of the V1 and SC models serve as examples of how to leverage the probabilistic toolkit for theoretical insight into models of neural computation.
 
@@
 so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
-As contrast increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
+As contrast $c \in [0, 1]$ increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
 Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
@@
 While perturbations in the sensitivity dimension $\mathbf{v}_1(\mathbf{z}^*)$ adapt with the mode $\mathbf{z}^*$ chosen, perturbations in $\mathbf{v}_a$ for $a \in \{\text{all}, \text{side}, \text{text}, \text{diag}\}$ are invariant to $\mathbf{z}$ (Equation \ref{eq:dzdlambda}).
+
+To understand the connectivity mechanisms that distinguish these two regimes, we perturb connectivity at each mode in dimensions that have well defined roles in processing for the Pro and Anti tasks.
+A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$.
+These eigenmodes have intuitive roles in processing in each task, and are accordingly named the \textit{all}, \textit{side}, \textit{task}, and \textit{diag} eigenmodes (see Section \ref{methods_SC}).
+Furthermore, the parameter dimension $\mathbf{v}_a$ ($a \in \{\text{all}, \text{side}, \text{task}, \text{ and diag}\}$) that increases the eigenvalue of connectivity $\lambda_a$ is $\mathbf{z}$-invariant (unlike the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$) and $\mathbf{v}_a \perp \mathbf{v}_{b \neq a}$.
+Thus, by changing the degree of amplification of each processing mode by  perturbing $\mathbf{z}$ along $\mathbf{v}_a$, we can elicit the differentiating properties of the two regimes.
+
+Through these connectivity perturbation analyses, we found that increasing $\lambda_{\text{task}}$ strongly reduced Pro accuracy in regime 1, yet strongly reduced Anti accuracy in regime 2.
+This suggests that stronger task representations can inhibit both Pro and Anti task performance in different contexts.
+Furthermore, changing $\lambda_{\text{task}}$ in either direction decreases Anti performance in regime 1, showing that Anti task performance in regime 1 is dependent on a specific level of task representation.
+We also found that with increasing $\lambda_{\text{diag}}$, Pro accuracy increased in both regimes, but there were opposite effects on Anti accuracy.
+In regime 1, stronger amplification of diagonal population patterns decreased Anti accuracy, while in regime 2 accuracy increased.
+These findings give us an understanding of the mechanistic differences in computation enabling rapid task switching in each regime.
+
 %

### 2021-04-17 written/eLifeRevision/Bittner_eLife2020.tex

@@
 \subsection{Theoretical models}\label{methods_theoretical_models}
-In this study, we used emergent property inference to examine several models relevant to theoretical neuroscience.  
+In this study, we used emergent property inference to examine several theoretical models of computation.
 Here, we provide the details of each model and the related analyses.
@@
 Intermediate hub frequency, like all other emergent properties in this work, is defined by the mean and variance of the emergent property statistics.
-In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ to capture variation in frequency between 0.5Hz and 0.6Hz (Equation \ref{eq:EP_STG}).
+In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ (Equation \ref{eq:EP_STG}).
 As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
@@
 EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
-The distribution shown is that of the architecture converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
-
-We calculated the Hessian at the mode of the inferred EPI distribution.  
-The Hessian of a probability model is the second order gradient of the log probability density $\log q_{\bm{\theta}}(\mathbf{z})$ with respect to the parameters $\mathbf{z}$:  $\frac{\partial^2 \log q_{\bm{\theta}}(\mathbf{z})}{\partial \mathbf{z} \partial \mathbf{z}^\top}$.
-With EPI, we can examine the Hessian, which is analytically available throughout distribution, to indicate the dimensions of parameter space that are sensitive (strongly negative eigenvalue), and which are degenerate (low magnitude eigenvalue) with respect to the emergent property produced.
-In Figure \ref{fig:STG}D, the eigenvectors of the Hessian $v_1$ (solid) and $v_2$ (dashed) are shown evaluated at the mode of the distribution.  
-The length of the arrows is inversely proportional to the square root of absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
-Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space are chosen arbitrarily.
+The architecture converged with criteria $N_{\text{test}} = 100$.
+
+In Figure \ref{fig:STG}D, the sensitivity dimension $v_1$ (solid) and the second eigenvector of the Hessian $v_2$ (dashed) are shown evaluated at the mode of the distribution.  
+The length of the arrows is inversely proportional to the square root of the absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
+Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space were chosen to have positive $g_{\text{synA}}$.$  $
 %We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).

### 2021-04-19 written/eLifeRevision/Bittner_eLife2020.tex
revisit section 3.5

@@
 This is most saliently pointed out in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
+The two regimes produce different types of responses in the Pro and Anti tasks (Fig. SC2).
 Not only has EPI captured this complicated distribution of connectivities producing rapid task switching, we can query the EPI distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.
@@
 These distinct pathologies of rapid task switching are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(r=1))$ vs $\mathbf{v}_1(\mathbf{z}^*(r=2))$) and explain the sharp change in local structure of the EPI distribution.
-With further perturbation analyses along dimensions of connectivity having established effects on processing, we are able to distinguish these two regimes with mechanistic insights (Fig.
-\ref{fig:SC}D, middle, bottom) (Section \ref{methods_SC}).
+
+To further examine the two regimes, we can perturb $\mathbf{z}$ in the same way along dimensions that independently change the eigenvalues of the connectivity matrix (which has constant eigenvectors with respect to $\mathbf{z}$).
+These eigenvalues $\lambda_{\text{all}}$, $\lambda_{\text{side}}$, $\lambda_{\text{task}}$, and $\lambda_{\text{diag}}$ correspond to connectivity eigenmodes with intuitive roles in processing in this task (Fig.
+\ref{fig:SC3}A).
+For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_SC}).
+Perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has close to the same effect on Anti accuracy as perturbations along the sensitivtity dimension (Fig. \ref{fig:SC}D, middle).
+This suggests that there is a carefully tuned strength of task representation in regime 1, which if disturbed results in random Anti trial responses.
+Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, bottom).
+In the next section, we relate these mechanistic characterizations of each regime to how connectivity executes each task in the presence of optogenetic silencing.
 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig5/fig5.pdf}
+\includegraphics[scale=1.]{figures/fig5/fig5.pdf}
 \end{center}
@@
 For EPI in Fig. \ref{fig:LRRNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
+The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
 We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
@@
 EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
-The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across random seeds.
+The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across three random seeds.
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figSC2/figSC2.pdf}
+\includegraphics[scale=1.]{figures/figSC2/figSC2.pdf}
 \end{center}
 \caption{\small (SC2):  
-\textbf{A}.
-Simulations in network regime 1 ($hW_{\text{fixed}} = -1.2$) (center) with simulations given connectivity perturbations in the negative direction of the sensitivity vector $\mathbf{v}_1$ (left) and positive direction (right).
-\textbf{B}. Same as A for network regime 2.
+\textbf{A}. Simulations in network regime 1 ($hW_{\text{fixed}} = -1.5$).
+\textbf{B}. Simulations in network regime 2 ($hW_{\text{fixed}} = -1.5$) .
 }
@@
 \begin{center}
-\includegraphics[scale=0.9]{figures/figSC3/figSC3.pdf}
+\includegraphics[scale=1.1]{figures/figSC3/figSC3.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/figSC5/figSC5.pdf}
+\includegraphics[scale=1.2]{figures/figSC5/figSC5.pdf}
 \end{center}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.2]{figures/figSC6/figSC6.pdf}
+\end{center}
+\caption{\small (SC6): 
+\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
+\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI distribution.
+}
+\label{fig:SC6}
+\end{figure}
+
 Writing the EPI distribution as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
@@
 \begin{center}
-\includegraphics[scale=0.5]{figures/figSC6/figSC6.pdf}
-\end{center}
-\caption{\small (SC6): 
+\includegraphics[scale=0.5]{figures/figSC7/figSC7.pdf}
+\end{center}
+\caption{\small (SC7): 
 \textbf{A}. Entropy throughout optimization. 
@@
 }
-\label{fig:SC6}
+\label{fig:SC7}
 \end{figure}
@@
 For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
-The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
+The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
 We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
-The distribution converged with criteria $N_{\text{test}} = 25$.
+The distribution was the greatest EPI distribution to converge across 5 random seeds with criteria $N_{\text{test}} = 25$.
 
 The EPI distribution of SC model connectivities producing rapid task switching has interesting structure.
-Throughout $q_{\bm{theta}}(\mathbf{z} \mid \mathcal{X})$, we see that the probability distribution is narrow in $hW$ (Fig. \ref{fig:SC}C).
+Throughout $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we see that the probability distribution is narrow in $hW$ (Fig. \ref{fig:SC}C).
 This suggests that rapid task switching is sensitive to changes in $hW$, but this is only a single parameter.

### 2021-04-19 written/eLifeRevision/Bittner_eLife2020.tex

@@
 Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, bottom).
-In the next section, we relate these mechanistic characterizations of each regime to how connectivity executes each task in the presence of optogenetic silencing.
+In the next section, we build on these mechanistic characterizations of each regime by examining their resilience to optogenetic silencing.
 
@@
 \ref{fig:SC_opto}A).
-All network connectivities inferred by EPI exhibited this same effect, when network activities were silenced during the delay period (see Section \ref{methods_SC}).
-Notably, EPI inferred connectivities were only conditioned upon the emergent property of rapid task switching, not on Anti task failure during delay period silencing.
-
-Following delay period inactivation, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC5}A).
-We interpreted these similarities to suggest that delay period inactivation flips the internal representation of task in the model.
-Connectivity patterns inducing greater Pro task accuracy would antagonistically reduce accuracy in Anti trials subject to delay period inactivation in which this flip in task representation occurs.
-In fact, across connectivities in the EPI inferred distribution, there was strong anti-correlation between Pro task accuracy $p_P$ and Anti task accuracy during delay period silencing $p_{A, opto}$ (Fig.
-\ref{fig:SC_opto}B).
-
-We explored whether this antagonism between $p_P$ and $p_{A, opto}$ was present in each connectivity regime.
-As in the previous section, we tested perturbations in the dimension of sensitivity $\mathbf{v}_1(\mathbf{z})$ from regime 1 to regime 2.
-In both regimes, we recall that perturbations along $\mathbf{v}_1(\mathbf{z})$ yield increasing $p_P$.
-We found that only in regime 2, does $p_{A,\text{opto}}$ decrease as $p_P$ increases(Fig. \ref{fig:SC_opto}C).
-This context further distinguishes the two regimes discovered with EPI.
+This suggests that SC maintains a representation of task throughout the delay period, which is important for correct execution of the Anti task.
+Network connectivities inferred by EPI exhibited this same effect in simulation at high  optogenetic strength $\gamma \in [0, 1]$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_SC}).
+To emphasize, EPI inferred connectivities were only conditioned upon the emergent property of rapid task switching, not on Anti task failure during delay period silencing.
 
@@
 \caption{\footnotesize 
-\textbf{A}. The EPI distribution predicts experimental results (left) showing no change in the Pro task, but larger error in the Anti task (right).
-\textbf{B}. Accuracy in the Anti task during delay period optogenetic inactivation $p_{A,\text{opto}}$ is strongly anticorrelated with accuracy in the Pro task.
-\textbf{C}. Mean and standard error ($N_{\text{test}}$ = 25) of accuracy with respect to perturbation along the sensitivity dimension of each mode $\mathbf{z}^*$.
+\textbf{A}. Experimental effect of delay period silencing on Pro and Anti task accuracy in rats.
+\textbf{B}. Mean and standard error (bars) across EPI distribution of Pro and Anti error induced by delay period inactivation of varying optogenetic strength.
+\textbf{C}. (Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.675$.
+(Right) Correlations of connectivity eigenvalues with Anti error from regime 1 to regime 2 at $\gamma = 0.675$.
+\textbf{D}.
+(Left) Responses of the SC model to delay period inactivation at $\gamma = 0.85$.
+(Right) Anti accuracy following delay period inactivation at $\gamma = 0.85$ versus accuracy in the Pro task across connectivities in the EPI distribution.
 }
@@
 
-In summary, we used EPI to obtain the full distribution of connectivities that execute rapid task switching.
-This EPI distribution revealed two regimes of rapid task switching, which we characterized using the probabilistic toolkit afforded by EPI.
-We found that both of these parametric regimes identified by EPI reproduce results from optogenetic inactivation experiments: when activity is silenced during the delay period, only Anti accuracy suffers.
-We then identified connectivity mechanisms governing Anti accuracy during delay period silencing, and showed the parametric regime they describe.
+The mean increase in Anti error is closest to the experimentally measured value of 7\% at $\gamma = 0.675$ (Fig. \ref{fig:SC_opto}B, black dot).  
+At this level of optogenetic strength, only regime 1 exhibits an increase in Anti error with delay period silencing (Fig. \ref{fig:SC_opto}C, left).
+The connectivities in regime 2 are thus more resilient in the Anti task to delay period silencing than regime 1.
+In regime 1, greater $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ decrease Anti error (Fig. \ref{fig:SC_opto}C, right).
+In other words, these anticorrelations show that stronger task representations and diagonal amplification make the SC model more resilient to delay period silencing in the Anti task.
+All correlations of connectivity eigenvalue with Anti error degrade in regime 2, where there is no effect of delay period silencing on Anti error (Fig. \ref{fig:SC_opto}C, right).
+
+At about $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero
+Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC_opto}D, left).
+We interpreted these similarities to suggest that delay period inactivation at this saturated level flips the internal representation of task (from Anti to Pro) in the model.
+This would explain why the Anti error saturates at 50\%: the average Anti accuracy in EPI inferred connectivities is $75\%$, but is 25\% when the internal representation is flipped during delay period silencing.
+This hypothesis prescribes a model of Anti accuracy during delay period silencing of $p_{A,\text{opto}} = 100\% - p_P$, which is fit closely across both regimes of the EPI inferred connectivities (Fig.
+\ref{fig:SC_opto}D, right).
+Similarities between Pro and Anti trial responses were not present at the experiment-matching level of $\gamma=0.675$ (Fig.
+\ref{fig:SC6} left) and neither was anti-correlation in $p_P$ and $p_{A,\text{opto}}$ (Fig.
+\ref{fig:SC6} right).
+
+In summary, the connectivity inferred by EPI to perform rapid task switching replicated  results from optogenetic silencing experiments.
+We found that at levels of optogenetic strength matching experimental levels of Anti error, only one regime actually exhibited the effect.  
+This suggests that one regime is less resilient to optogenetic perturbation, and perhaps more biologically realistic.
+Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
+This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (10 times faster than brute force methods, see Section \ref{methods_SC}).
 
@@
 \caption{\small (SC5): 
-\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI distribution.
+(Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
+(Right) Correlations of connectivity eigenvalues with Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
 }
@@
 \caption{\small (SC6): 
-\textbf{A}. Response of each parameter regime to optogenetic silencing during the delay period.
-\textbf{B}. Error induced by delay period inactivation with increasing optogenetic strength.  Means and standard deviations are calculated across the entire EPI distribution.
+(Left) Responses of the SC model to delay period inactivation at $\gamma = 0.675$.
+(Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.
 }
@@
 Each of these modes is denoted $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for regime $r \in \{1, 2\}$.
+
+For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampled from the EPI distribution.
+Each sample was assigned to the closest mode $\mathbf{z}^*(hW_{\text{fixed}}, r)$.
+Sampling continued until 500 samples were assigned to each mode, which took 7.36 seconds.
+To obtain this many samples for each mode with brute force sampling over the chosen prior, this would take 90 days.
 
@@
 \end{equation}
-Delay period inactivation was from $0.8 < t < 1.2$, choice period inactivation was for $t > 1.2$ and total inactivation was for the entire trial.
+Delay period inactivation was from $0.8 < t < 1.2$.
 

### 2021-04-19 written/eLifeRevision/Bittner_eLife2020.tex

@@
 \textbf{D}.
-(Left) Responses of the SC model to delay period inactivation at $\gamma = 0.85$.
+(Left) Responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.85$.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.85$ versus accuracy in the Pro task across connectivities in the EPI distribution.
@@
 Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
-This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (10 times faster than brute force methods, see Section \ref{methods_SC}).
+This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (X times faster than brute force methods, see Section \ref{methods_SC}).
 
@@
 \caption{\small (SC6): 
-(Left) Responses of the SC model to delay period inactivation at $\gamma = 0.675$.
+(Left) Responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.

### 2021-04-19 written/eLifeRevision/Response.tex
response doc

%--------------------
% Packages
% -------------------
\documentclass[11pt,a4paper]{article}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage{gentium}
\usepackage{mathptmx} % Use Times Font

\usepackage[pdftex]{graphicx} % Required for including pictures
\usepackage[swedish]{babel} % Swedish translations
\usepackage[pdftex,linkcolor=black,pdfborder={0 0 0}]{hyperref} % Format links for pdf
\usepackage{calc} % To reset the counter in the document after title page
\usepackage{enumitem} % Includes lists

\frenchspacing % No double spacing between sentences
\linespread{1.2} % Set linespace
\usepackage[a4paper, lmargin=0.1666\paperwidth, rmargin=0.1666\paperwidth, tmargin=0.1111\paperheight, bmargin=0.1111\paperheight]{geometry} %margins
%\usepackage{parskip}

\usepackage[all]{nowidow} % Tries to remove widows
\usepackage[protrusion=true,expansion=true]{microtype} % Improves typography, load after fontpackage is selected

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

%-----------------------
% Set pdf information and add title, fill in the fields
%-----------------------
\hypersetup{ 	
pdfsubject = {},
pdftitle = {},
pdfauthor = {}
}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

%-----------------------
% Begin document
%-----------------------
\begin{document}

{\Large \textbf{Response to reviewers}} \\
(original comments in bold)

\textbf{Bittner and colleagues introduce a machine learning framework for maximum entropy inference of model parameter distributions that are consistent with certain emergent model properties, specified by the investigator. The approach is illustrated on several models of potential interest.}

\textbf{Reviewers were broadly enthusiastic about the potential usefulness of this methodology. However, the reviews and ensuing discussion revealed several points of concern about the manuscript and the approach. The full reviewer comments are included below.}

We thank the reviewers for the very constructive feedback.  To address these points of concern, we have made extensive improvements to the manuscript.  We have made significant changes to our writing and explanation of the inference technique, made an extensive comparison to other parameter inference methods, and increased the quality and depth of our scientific analyses, which yield strong theoretical results.  Our model analyses now focus on analyzing the rich structure of parameter distributions, which the deep probability distributions of EPI make possible.  Here, we provide a list of key manuscript updates, and below we explain how these changes address each of the reviewer’s concerns.

List of key manuscript improvements:
\begin{itemize}
\item We significantly changed how we explain the parameter inference technique (EPI) to clarify the methodological contribution of this research.  We now emphasize the novelty and scientific value of conditioning probability distributions on theoretical criteria of interest (emergent properties), in contrast to noisy observed neural datasets.
\item We show EPI’s superior scalability in parameter dimension to alternative likelihood-free inference techniques SMC-ABC and SNPE when inferring RNN connectivities exhibiting stable amplification.
\item We show how derivations of excitatory variability in a model of primary visual cortex become analytically intractable as additional neuron types are added.  We then use EPI to produce a clear insight that mathematical analysis did not produce.
\item We discover multiple parametric regimes of computation in a model of superior colliculus from the fine distributional structure captured by EPI.  We then use the deep probability distribution to mechanistically characterize each regime, and also find that both of the discovered regimes reproduce an important result from inactivation experiments.
\end{itemize}

\textbf{The main concerns are summarized as follows:}

\textbf{1. The methodology is not adequately explained. Both the body text and methods section present a somewhat selective description that is very hard to follow in places and should be checked and rewritten for clarity, completeness, notational consistency and correctness.}

We thank the reviewers for pointing out that our explanation of the method was too narrowly focused and hard to follow. In the introduction, we now frame the method within the more general context of parameter inference techniques in neuroscience, rather than the context of recent advancements in machine learning.  We have made concentrated efforts to simplify language, relate to all relevant existing methodology (Section 5.1.1 Related approaches), and precisely explain the novelty and utility of this deep inference technique.

Primarily, we explain the incongruity between parameter inference techniques and the practice of theoretical neuroscience that motivates EPI: modern parameter inference techniques are designed to condition on noisy datasets rather than the mathematically defined criteria.  This mismatch has consequences – primarily that data predicted from the inferred distribution are left unconstrained and often violate the desired conditioning criteria.  We then explain how formalizing the mathematical criteria of interest into a statistical feature of the model – the emergent property – we enable deep inference of mechanistic model parameters with respect to this property.

In Section 3.1 and 3.2, we present EPI through a motivational example of the STG subcircuit.  The figure has been updated to reflect the modified presentation, and we more clearly emphasize the scientific tools afforded once a distribution is fit with EPI.

\textbf{2. The computational resources required to use this method are not adequately benchmarked. For example, the cosubmission (Macke et al) reported wall clock time, required hardware and iterations required to produce results by directly comparing to existing methods (approximate bayesian computation, naive sampling, etc.) Without transparent benchmarks it is not possible to assess the advance offered by this method.}

We thank the reviewers for emphasizing the importance of this methodological comparison.  In Section 3.3, we provide a direct comparison of EPI to alternative likelihood-free inference techniques SMC-ABC and SNPE by inferring RNN connectivities that exhibit stable amplification.  These comparisons evaluate both wall time and simulation count, and we explain how each algorithm was run in its ideal hardware setting.

In this analysis, we demonstrate the superior scalability of deep inference techniques (EPI and SNPE) to the state-of-the-art random sampling technique (SMC-ABC) .  We then show that SNPE does not constrain the properties of inferred parameters, resulting in many unstable and nonamplified models.  Furthermore, we push the limits of SNPE through targeted hyperparameter modifications, and show that for this model EPI scales to a higher dimensional parameter space than SNPE.

\textbf{3. The extent to which this method is generally/straightforwardly applicable was in doubt. It seemed as though a significant amount of computation was required to do inference on one specified property and that the computation would need to be run afresh to query a new property. The methodology in the cosubmission (Macke) made clear that computation required for successive inferences is 'amortized' during training on random parameters. Moreover, EPI seemed less flexible than the cosubmission's approach in that it required a differentiable loss function. The complementarity and advantages of this approach as opposed to the cosubmission's approach are therefore unclear. }

The reviewers are right to point out these characteristics of EPI: it does not amortize across emergent properties, and it requires differentiability of the emergent properties of the model. Indeed SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.  However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a trade-off between likelihood-free inference approaches.  Unlike SNPE, EPI leverages gradients of the model to improve efficiency (Section 3.3).  The emergent properties of many models in neuroscience are tractably differentiable, four of which we analyze in this manuscript ranging across levels of biological realism, network size, and computational function.   Furthermore, EPI focuses the entire expressivity of the approximating deep probability distribution on a single distribution, rather than spreading this expressivity to some degree across the chosen training distribution of amortized posteriors in SNPE (Section 5.1.1 Related Approaches).

Finally, we emphasize that EPI does something fundamental that SNPE and other inference techniques cannot.  EPI learns parameter distributions whose predictions are constrained to produce the emergent property.  We show in Supplementary Figure RNN2 that SNPE does not control the statistical properties of its predictions, resulting in the inference of many parameters that are not consistent with the desired emergent property.

\textbf{4. Some examples lack depth in their treatment (see reviewer comments) and in some cases the presentation is somewhat misleading. The STG example is not in fact a model of the STG. The cosubmission (Macke) uses a model close to the original Prinz et al model, which is a model of the pyloric subnetwork. It would be instructive to benchmark against this same model, including computation time/resources required. Secondly, the subsequent example (input-responsivity in a nonlinear sensory system) appeared to imply that EPI permits 'generation' and testing of hypotheses in a way that other methods do not. All the method really does is estimate a joint distribution of feasible parameters in a specific model which is manually inspected to propose hypotheses. Any other method (including brute force sampling) could be used in a similar way, so any claim that this is an integral advantage of EPI would be spurious. Indeed, one reviewer was confused about the origin of these hypotheses. While it is helpful to illustrate how EPI (and other methods) could be used, the writing needs to be far clearer in general and should clarify that EPI does not offer any new specific means of generating or evaluating hypotheses.}

We thank the reviewers for explaining how they found some of the presentation misleading.  We have taken serious care in this manuscript to clarify a.) what is novel, appreciable scientific insight provided by EPI, as well as b.) which scientific analyses are made possible by EPI.

a.)	In the revised manuscript we clarify that novel theoretical insights are not being made into the STG subcircuit model or the recurrent neural network models.  The STG subcircuit serves as a motivational example to explain how EPI works, and we use RNNs exhibiting stable amplification as a substrate for scalability analyses.  We do produce strong, theoretical insights into a model of primary visual cortex (Section 3.4) and superior colliculus (Section 3.5).  These analyses have substantially more depth than the previous manuscript.

b.)	The ability to infer a flexible approximation to a probability distribution constrained to produce an emergent property is novel in its own right (Figure 2).  The deep probability distribution fit by EPI facilitates the mode identification (via gradient ascent of the parameter log probability) and sensitivity measurement (via the measurement of the eigenvector of the Hessian at a parameter value). These mode identifications and sensitivity measurements are done in Sections 3.1, 3.4, and 3.5.  By using this mode identification technique along the ridges of high parameter probability in the SC model, we identify the parameters transitioning between the two regimes.  Finally, the sensitivity dimensions identified by the SC model facilitated regime characterization through perturbation analyses.

Importantly, we do not claim that these theoretical insights were necessary with the techniques in b.).  One could have come to these conclusions via brute force techniques, although less efficiently.  The point is that we have brought deep inference to bear on theoretical neuroscience in a way that is principled, efficient, and scientifically compelling.

As the reviewers indicate, the STG model analyzed in our manuscript is not that of Prinz et al. 2004, and thus not the model analyzed by the cosubmission.  We found the Prinz et al. model prohibitive to infer with EPI, since the gradients of spiking frequency from its simulation are quite computationally intensive.  However, we found it critical to show that it’s very possible to do inference in such biophysically realistic Morris-Lecar models when gradients are tractable, which is the case of the 5-neuron STG model we analyzed from Gutierrez et al. 2013.  This 5-neuron model represents the IC neuron (hub) and its coupling to the pyloric (fast) or gastric mill (slow) subcircuit rhythms.  In the introductory text, we refer to this model as the “STG subcircuit” model (rather than “STG model”), and we clarify what is being modeled in Results Section 3.1.

\textbf{5. There is a substantial literature on parameter sensitivity analysis and inference in systems biology, applied dynamical systems and control that has been neglected in this manuscript. The manuscript needs to acknowledge, draw parallels and explain distinctions from current methods (ABC, profile likelihood, deep learning approaches, gaussian processes, etc). The under-referencing of this literature deepened concerns about whether this approach represented an advance. DOIs for a small subset of potentially relevant papers include:}

\textbf{https://doi.org/10.1038/nprot.2014.025 \\
http://doi.org/10.1085/jgp.201311116 \\
http://doi.org/10.1016/j.tcs.2008.07.005 \\
http://doi.org/10.3182/20120711-3-BE-2027.00381 \\
http://doi.org/10.1093/bioinformatics/btm382 \\
http://doi.org/10.1111/j.1467-9868.2010.00765.x \\
http://doi.org/10.1098/rsfs.2011.0051 \\
https://doi.org/10.1098/rsfs.2011.0047 \\
http://doi.org/10.1214/16-BA1017 \\
https://doi.org/10.1039/C0MB00107D}

Thank you for pointing us to these great references on sensitivity analyses and applied dynamical systems.  We have incorporated them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.

\textbf{6. One of the reviewers expressed concern that the work might have had significant input from a senior colleague during its early stages, and that that it might be worth discussing with the senior colleague whether their contribution was worthy of authorship. The authors may take this concern into account in revising the manuscript.}

We have reached out to Woods Hole Course Project mentors where this work was discussed and explored in its early stages.  We are confident that there are no authorship issues.

\textbf{7. Finally, please also address specific points raised by the reviewers, included below. } \\

{\Large \textbf{Reviewer \#1: }}

\textbf{General Assessment: The authors introduce a machine learning framework for maximum entropy inference of model parameters which are consistent with certain emergent properties, which they call 'emergent property inference'. I think this is an interesting and direction, and this looks like a useful step towards this program. I think the paper could be improved with a more thorough discussion both of the broad principles their black box approach seeks to optimize, as well as the details of its implementation. I also think the detailed examples should be more self-contained. Finally I find this work to be somewhat misrepresented as a key to all of theoretical neuroscience. This approach may have some things to offer to the interesting problem of finding parameter regions of models, but this is not the entirety of, nor really a major part of theoretical neuroscience as I see it. }

\textbf{Other concerns: }

\textbf{(1) Maximizing the entropy of the distribution is not a reparameterization invariant exercise. That is, results depend on whether the model parameters contains rates or time constants, for example. I wonder if this approach is attempting to use a 'flat prior' in some sense which has the same reparameterization issue? Can the authors comment? }

The reviewer is correct to point out that maximum entropy solutions are not reparameterization invariant, and indeed the units matter.  The reviewer’s suggestion that the method is in some sense using a flat prior is also correct.  To clarify, EPI does not execute posterior inference, because there is no empirical dataset or specified prior belief in EPI framework.  However, we derive the relation of EPI to Bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.

In our examples, we only infer distributions of parameters with the same units.  As suggested, the EPI solution will differ according to relative scaling of parameter values under the maximum entropy selection principle. Thus, an important clarification is that sensitivity quantifications are made in the context of the chosen parameter scalings. A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.  It makes most sense to make sure EPI is exploring through parameter space through a standard range of values, and that sensitivity measurements are interpretable.

\textbf{(2) I don't think this is a criticism of the work, but instead of the writing about it: I find the introductory paragraphs to give a rather limited overview of theory as finding parameters of models which contain the right phenomenology. }

(*Do this?) We have edited the introduction to more appropriately characterize the practice of theoretical neuroscience.

\textbf{(3) I am somewhat familiar with the stomatognathic circuit model, and so that is where I think I understand what they have done best. I don't understand what I should take away from their paper with regards to this model. Are there any findings that hadn't been appreciated before? What does this method tell us about the system and or its model? }

asdfasdfa

\textbf{(4) I don't follow the other examples. Ideally more details should be given so that readers like myself who don't already know these systems can understand what's been done. }

Thank you for the feedback.  We have taken care to give more general context, and motivation for each neural circuit model.

\textbf{(5) In figure 2C, the difference between the confidence interval between linear and nonlinear predictions is huge! How much of this is due to nonlinearities, and how much is due to differences in the way these models are being evaluated? }

In the current manuscript, we do not examine the difference between linear and nonlinear predictions of the V1 model. \\

{\Large \textbf{Reviewer \#2: }}

\textbf{General assessment }

\textbf{This is a very interesting approach to an extremely important question in theoretical neuroscience, and the mathematics and algorithms appear to be very rigorous. The complexities in using this in practice make me wonder if it will find wide application though: setting up the objective to be differentiable, tweaking hyperparameters for training, and interpreting the results; all seem to require a lot from the user. On the other hand, the authors are to be congratulated on providing high quality open source code including clear tutorials on how to use it. }

\textbf{Major points }

\textbf{1. Training deep networks is hard. Indeed the authors devote a substantial amount of the manuscript to techniques for training them, and note that different hyperparameters were necessary for each of the different studies. Can the authors be confident that they have found the network which gives maximum entropy or close to it? If so, how. If not, how does that affect the conclusions? }

\textbf{2. Interpreting the results still seems to require quite a lot of work. For example, from inspecting Fig 2 the authors extract four hypotheses. Why these four? Are there other hypotheses that could be extracted and if not how do we know there aren't? Could something systematic be said here? }

\textbf{3. Scalability. The authors state that the method should in principle be scalable, but does that apply to interpreting the results? For example, for the V1 model it seems that you need to look at 48 figures for 4 variables, and I believe this would scale as O(n$^2$) with n variables. This seems to require an unsustainable amount of manual work? }

\textbf{4. There are some very particular choices made in the applications and I wonder how general the conclusions are as a consequence. For example, in equation (5) the authors choose an arbitrary amount of variance 0.01$^2$ - why? In the same example, why look at y=0.1 and 0.5? }

\textbf{Minor points }

\textbf{The introduction and discussion are very clearly written but the results section is hard going. Partly this is unavoidable given the subject matter, but a few sentences here and there might help the reader along. Things like "x is the internal state of the model, z are the parameters we will change, ...". When introducing entropy in equation (3), H isn't previously defined, and again it might help to give the reader a hand here, e.g. "max entropy means the distribution is as spread out as possible" (you can surely find a better thing to say than this, but just to give an idea). The other point which is quite hard to follow is interpreting e.g. Fig 2C. Perhaps for Hypothesis 1 you could write a couple of sentences explaining slightly more clearly why seeing small blobs or horizontal/vertical lines in these distribution plots means that it's mainly determined by the direct input. }

Things I will say \\

{\Large \textbf{Reviewer \#3:}}

\textbf{This paper addresses a major issue in fitting neuroscience models: how to identify the often degenerate, or nearly degenerate, set of parameters that can underlie a set of experimental observations. Whereas previous techniques often depended upon brute force explorations or special parametric forms or local linearizations to generate sets of parameters consistent with measured properties, the authors take advantage of deep generative networks to address this problem. Overall, I think this paper and the complementary submission have the potential to be transformative contributions to model fitting efforts in neuroscience. That being said, since the primary contribution is the methodology, I think the paper requires more systematic comparisons to ground truth examples to demonstrate potential strengths and weaknesses, and more focus on methodology rather than applications. }

\textbf{Substantive concerns: \\
1) The authors only have a single ground-truth example where they compare to a known result (a 2x2 linear dynamical system). It would be good to show how well this method compares to results from, for example, a direct brute force grid search of a system with a strongly non-elliptical (e.g. sharply bent) shaped parameter regime and a reasonably large (e.g. 5?) number of parameters corresponding to a particular property, to see how well the derived probability distribution overlaps the brute force grid search parameters (perhaps shown via several 2-D projections). }

\textbf{2) It was not obvious whether EPI actually scales well to higher dimensions and how much computation it would take (there is one claim that it 'should scale reasonably'). While I agree that examples with a small number of parameters is nice for illustration, a major issue is how to develop techniques that can handle large numbers of parameters (brute force, while inelegant, inefficient, and not producing an explicit probability distribution can do a reasonable job for small \#'s of parameters). The authors should show some example of extending to larger number of parameters and do some checks to show that it appears to work. As a methodological contribution, the authors should also give some sense of how computationally intensive the method is and some sense of how it scales with size. This seems particularly relevant to, for example, trying to infer uncertainties in a large weight matrix or a non-parametric description of spatial or temporal responses or a sensory neuron (which I'm assuming this technique is not appropriate for? See point\#4 below). }

\textbf{3) For the STG-like example, this was done for a very simple model that was motivated by the STG but isn't based on experimental recordings. Most of the brute force models of the STG seek to fit various waveform properties of neurons and relative phases. Could the model handle these types of analyses, or would it run into problems due to either needing to specify too many properties or because properties like "number of spikes per burst" are discrete rather than continuous? This isn't fatal, but would be good to consider and/or note explicitly. }

\textbf{4) The discussion should be expanded to be more specific about what problems the authors think the model is, or is not, appropriate for. Comparisons to the Goncalves article would also be helpful since users will want to know the comparative advantages/disadvantages of each method. (if the authors could coordinate running their methods on a common illustrative example, that would be cool, but not required). }

\textbf{5) Given that the paper is heavily a (very valuable!) methods paper for a general audience, the method should be better explained both in the main text and the supplement. Some specific ones are below, but the authors should more generally send the paper to naïve readers to check what is/is not well explained. 
-Figure 1 is somewhat opaque and also has notational issues (e.g. omega is the frequency but also appears to be the random input sample). 
-For the general audience of eLife, panels C and D are not well described individually or well connected to each other and don't illustrate or describe all of the relevant variables (including what q0 is and what x is). 
-In equation 2 (and also in the same equation in the supplement), it was not immediately obvious what the expectation was taken over. 
-The authors don't specific the distribution of w (it's referred to only as 'a simple random variable', which is not clear). 
-It was also sometimes hard to quickly find in the text basic, important quantities like what z was for a given simulation. 
-The augmented Lagrangian optimization was not well explained or motivated. There is a reference to m=absolute value(mu) but I didn't see m in the above equation. 
-Using mu to describe a vector that includes means and variances is confusing notation since mu often denotes means 
-It would be helpful to have a pseudo-code 'Algorithm' figure or section of the text }

\textbf{Minor Comments: }

\textbf{1) I'm not sure if the authors are referring to a particular constrained form of the Schur decomposition, but the general statement in the Figure caption that the Schur decomposition is unique is not true. Also, one does not need to refer to "Schur eigenvalues" since the diagonal elements of the Schur decomposition are the (usual) eigenvalues. }

\textbf{2) p. 31: usually one reserves the variable omega for angular frequencies: omega = 2*pi*f where f is frequency. }

\textbf{3) Some references for other approaches and work that might be worth listing for scholarship: Sloppy models and information geometry (including MCMC approaches, e.g. Mannakkee, Ragsdale, Transtrum, Gutenkunst); higher dimensional sloppy models in neuroscience (O'Leary, Sutton, \& Marder 2015, Fisher, Olasagasti, et al., 2013); Compensatory parameter combinations through the implicit function theorem (Olypher and Calabrese, J. Neurophys. 2007). }

\textbf{Additional data files and statistical comments:}

\textbf{Code should be made available in well-documented form if it isn't already. }

\end{document}

### 2021-04-19 written/eLifeRevision/Bittner_eLife2020.tex
edits'

@@
 \ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity estimated from post-synaptic potential and connectivity rate measurements (see Section \ref{methods_V1}).
+Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}).
 Each population also experiences an external input $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
@@
 Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
-This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (X times faster than brute force methods, see Section \ref{methods_SC}).
+This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).
 
@@
 Sampling continued until 500 samples were assigned to each mode, which took 7.36 seconds.
-To obtain this many samples for each mode with brute force sampling over the chosen prior, this would take 90 days.
+To obtain this many samples for each mode with brute force sampling over the chosen prior, this would take 4.20 days.
 

### 2021-04-19 epi/util.py
small util changes

@@ -125,7 +125,7 @@ def np_column_vec(x):
 
 
 def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1., check_last_k=None, by_df=False):
-    paths = epi_df['path'].unique()
+    paths = sorted(epi_df['path'].unique())
     best_Hs = []
     best_ks = []
     df_rows = []
@@ -134,7 +134,7 @@ def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1., check_last_k=None, by_d
         epi_df2 = epi_df[epi_df['path'] == path]
         _D = epi_df2.iloc[0]['arch']['D']
         _rs = epi_df2.iloc[0]['arch']['random_seed']
-        print("Processing EPI: D=%d, rs=%d.\n" % (_D, _rs), end="")
+        print("Processing EPI: D=%d, rs=%d." % (_D, _rs))
         if by_df:
             df_row = epi_df2.iloc[-1]
             converged = df_row['converged']
@@ -161,6 +161,7 @@ def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1., check_last_k=None, by_d
                 nu=nu,
                 start_k=start_k
             )
+            print('k', best_k, 'H', best_H)
             tf.keras.backend.clear_session()
         if (not any_converged) and converged:
             any_converged = True

### 2021-04-20 written/eLifeRevision/Bittner_eLife2020.tex
methods sub sections, STG language

@@
 %A model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A, and note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $z = [g_{\text{el}}, g_{\text{synA}}]$.
+
 A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A.
+The fast population ($f1$ and $f2$) represents the subnetwork generating the pyloric rhythm and the slow population ($s1$ and $s2$) represents the subnetwork of the gastric mill rhythm.
+The two fast neurons mutually inhibit one another, and spike at a greater frequency than the mutually inhibiting slow neurons.
+The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions.
 The jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$, smooth connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.
 Note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
-Specifically, the two fast neurons ($f1$ and $f2$) mutually inhibit one another, and oscillate at a faster frequency than the mutually inhibiting slow neurons ($s1$ and $s2$).  
-The hub neuron (hub) couples with either the fast or slow population, or both.
 
@@
 This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution that produces the emergent property. 
+Scientifically, we can use this parameter distribution to efficiently generate many parameters producing the emergent property or analyze the structure of the distribution which informs how model parameters govern the emergent property.
 %In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
@@
 \end{equation} 
-where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{morris1981voltage, gutierrez2013multiple} (see Section \ref{methods_STG} for more detail).
-
-%Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
-%Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc.
--- is close to an intermediate frequency of 0.53Hz.
-%Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
-%Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.
-
-%Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
-%Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz.  
-%Mathematically, we achieve this with two constraints: by fixing the mean hub frequency over the inferred parameter distribution of $\mathbf{z}$ and its resulting simulations $\mathbf{x}$ to 0.55Hz,
-%\begin{equation}\label{eq:EP_STG1}
-% \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.55 \end{bmatrix}
-%\end{equation}
-%and requiring that the variance of the hub frequency over the produced simulations is small
-%\begin{equation}\label{eq:EP_STG2}
-%\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  \begin{bmatrix} 0.025^2 \end{bmatrix}.
-%\end{equation}
-%This level of variance was chosen to be low enough to exclude the fast and slow frequencies of the two populations, but large enough to allow structural examination of the inferred parameter distribution.
-%By constraining the means and variances of emergent property statistics over $\mathbf{z}$ as well as the stochasticity of $\mathbf{x}$, we can precisely control the behavior that the inferred distribution that EPI infers.
-%In general, an emergent property
-%\begin{equation}\label{eq:EP_STG}
-%\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
-%\end{equation}
-%defines a collection of datasets with a statistic $f(\mathbf{x}; \mathbf{z})$ (which may be comprised of multiple statistics) and the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of those statistics over the datasets.
-%The choice of $\bm{\sigma}^2$ predicates the degree of variability around the mean $\bm{\mu}$ that is consistent with the emergent property.
+where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{morris1981voltage, gutierrez2013multiple} (see Section \ref{methods_stg_model} for more detail).
 
@@
 Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ to a suitably complicated distribution (see Section \ref{methods_NF} for more details).
-Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in Bayesian posterior inference (see Section \ref{methods_VI}).
+Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in variational bayesian methods (see Section \ref{methods_VI}).
 %\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
@@
 \ref{fig:STG}F activity traces).
-Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (when the correct architecture class is chosen, see Section \ref{methods_NF}).
+Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (see Section \ref{methods_stg_hessian}).
 
@@
 The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
-The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
+The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_sc_model}).
 Model responses are bounded from 0 to 1 as a function $\phi$ of an internal variable $\mathbf{u}$
@@
 To distinguish these two regimes, we use the EPI distribution to identify two sets of modes.
-By fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we arrive at two solutions $\mathbf{z}^*(hW_{\text{fixed}}, r)$ where regime $r \in [1,2]$, and regime 1 is that of greater $sW$ (see Section \ref{methods_SC}).
+By fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we arrive at two solutions $\mathbf{z}^*(hW_{\text{fixed}}, r)$ where regime $r \in [1,2]$, and regime 1 is that of greater $sW$ (see Section \ref{methods_sc_regime}).
 As $hW_{\text{fixed}}$ increases, the modes coalesce to intermediate parameters reflecting a transition between the two sets of modes (Fig. \ref{fig:SC4} top).
@@
 \ref{fig:SC3}A).
-For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_SC}).
+For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_sc_eig}).
 Perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has close to the same effect on Anti accuracy as perturbations along the sensitivtity dimension (Fig. \ref{fig:SC}D, middle).
@@
 This suggests that SC maintains a representation of task throughout the delay period, which is important for correct execution of the Anti task.
-Network connectivities inferred by EPI exhibited this same effect in simulation at high  optogenetic strength $\gamma \in [0, 1]$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_SC}).
+Network connectivities inferred by EPI exhibited this same effect in simulation at high  optogenetic strength $\gamma \in [0, 1]$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
 To emphasize, EPI inferred connectivities were only conditioned upon the emergent property of rapid task switching, not on Anti task failure during delay period silencing.
@@
 \end{equation} 
-Entropy is chosen as the normative selection principle to match that of Bayesian inference (see Section \ref{methods_ME_EF}).
-However, a key difference is that variational inference and other Bayesian methods do not constrain the predictions of their inferred parameter distribution.
+Entropy is chosen as the normative selection principle to match that of variational Bayesian methods (see Section \ref{methods_ME_EF}).
+However, a key difference is that variational Bayesian methods do not constrain the predictions of their inferred parameter distribution.
 This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
@@
 %Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
+
 We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
-Fast probability evaluation in turn facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
+Fast probability evaluation  facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
 Glow permutations were used in between coupling stages \cite{kingma2018glow}.
 This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only one of the forward or backward passes can be efficient.
-In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
+In this work, normalizing flows are used as flexible parameter distribution approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
 We specify the architecture used in each application by the number of Real-NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.
+
+When calculating Hessians of log probabilities in deep probability distributions, it is important to consider the normalizing flow architecture.
+With autoregressive architectures \cite{kingma2016improved, papamakarios2017masked}, fast sampling and fast log probability evaluations are mutually exclusive.
+That makes these architectures undesirable for EPI, where efficient sampling is important for optimization, and log probability evaluation speed predicates the efficiency of gradient and Hessian calculations.
+With Real NVP coupling architectures, we get both fast sampling and fast Hessians making both optimization and scientific analysis efficient.
 
@@
 
-\subsection{Theoretical models}\label{methods_theoretical_models}
-In this study, we used emergent property inference to examine several theoretical models of computation.
-Here, we provide the details of each model and the related analyses.
-
-\subsubsection{Stomatogastric ganglion}\label{methods_STG}
+\subsection{Stomatogastric ganglion}\label{methods_STG}
+In Section \ref{results_motivating} and \ref{results_dgm}, we used EPI to infer conductance parameters in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple}.
+This 5-neuron circuit model represents two subcircuits: that generating the pyloric rhythm (fast population) and that generating the gastric mill rhythm (slow population).
+The additional neuron (the IC neuron of the STG) receives inhibitory synaptic input from both subcircuits, and can couple to either rhythm dependent on  modulatory conditions.
+There is also a parametric regime in which this neuron fires at an intermediate frequency between that of the fast and slow populations \cite{gutierrez2013multiple}, which we infer with EPI as a motivational example.
+This model is not to be confused with an STG subcircuit model of the pyloric rhythm \cite{marder1992dynamic}, which has been statistically inferred in other studies \cite{prinz2004similar, gonccalves2019training}.
+
+\subsubsection{STG model} \label{methods_stg_model}
 We analyze how the parameters $\mathbf{z} = [ g_{\text{el}}, g_{\text{synA}}]$ govern the emergent phenomena of intermediate hub frequency in a model of the stomatogastric ganglion (STG) \cite{gutierrez2013multiple} shown in Figure \ref{fig:STG}A with activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$, using the same hyperparameter choices as Gutierrez et al.
@@
 \end{equation} 
-The input current of each neuron is the sum of the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents as well as gaussian noise $dB$.
+The input current of each neuron is the sum of the leak, calcium, potassium, hyperpolarization, electrical and synaptic currents.
 Each current component is a function of all membrane potentials and the conductance parameters $\mathbf{z}$.
+Finally, we include gaussian noise $dB$ to the model of Gutierrez et al. so that the model stochastic, although this is not required by EPI.
 
@@
 When the dynamic gating variables are considered, this is actually a 15-dimensional nonlinear dynamical system.  
-Gaussian noise $d\mathbf{B}$ of variance $(1 \times 10^{-12})^2$ A$^2$ makes the model stochastic, and introduces variability in frequency at each parameterization $\mathbf{z}$.
-
+The gaussian noise $d\mathbf{B}$ has variance $(1 \times 10^{-12})^2$ A$^2$, and introduces variability in frequency at each parameterization $\mathbf{z}$.
+
+\subsubsection{Hub frequency calculation} \label{methods_stg_hub}
 In order to measure the frequency of the hub neuron during EPI, the STG model was simulated for $T = 300$ time steps of $dt = 25\text{ms}$.  
@@
 In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ (Equation \ref{eq:EP_STG}).
+
+\subsubsection{EPI details for the STG model} \label{methods_stg_epi}
 As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
@@
 For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
-The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$, initialized to a gaussian approximation of samples returned by a preliminary ABC search.
+The normalizing flow architecture mapped $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$, initialized to a gaussian approximation of samples returned by a preliminary ABC search.
 We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
@@
 
+\subsubsection{Hessian sensitivity vectors} \label{methods_stg_hessian}
+To quantify the second-order structure of the EPI distribution, we evaluated the Hessian of the log probability $\frac{\partial^2 \log q(\mathbf{z} \mid \mathcal{X})}{\partial \mathbf{z} \mathbf{z}^\top}$.
+The eigenvector of this Hessian with most negative eigenvalue is defined as the sensitivity dimension $\mathbf{v}_1$, and all subsequent eigenvectors are ordered by increasing eigenvalue.
+These eigenvalues are quantifications of how fast the emergent property deteriorates via the parameter combination of their associated eigenvector.
 In Figure \ref{fig:STG}D, the sensitivity dimension $v_1$ (solid) and the second eigenvector of the Hessian $v_2$ (dashed) are shown evaluated at the mode of the distribution.  
+Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space were chosen to have positive $g_{\text{synA}}$.$  $
 The length of the arrows is inversely proportional to the square root of the absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
-Since the Hessian eigenvectors have sign degeneracy, the visualized directions in 2-D parameter space were chosen to have positive $g_{\text{synA}}$.$  $
+For the same magnitude perturbation away from the mode, intermediate hub frequency only diminishes along the sensitivity dimension $\mathbf{v}_1$ (Fig. \ref{fig:STG}E-F).
 %We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
@@
 
-\subsubsection{Scaling EPI for stable amplification in RNNs}\label{methods_RNN}
+\subsection{Scaling EPI for stable amplification in RNNs}\label{methods_RNN}
+\subsubsection{Rank-2 RNN model}
 We examined the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
@@
 
+\subsubsection{Stable amplification}
 For this RNN model to be stable, all real eigenvalues of $W$ must be less than 1:  $\text{real}(\lambda_1) < 1$, where $\lambda_1$ denotes the greatest real eigenvalue of $W$.
@@
 These two conditions are necessary and sufficient for stable amplification in RNNs \cite{bondanelli2020coding}.
+
+\subsubsection{EPI details for RNNs}
 We defined the emergent property of stable amplification with means of these eigenvalues (0.5 and 1.5, respectively) that satisfy these conditions.
@@
 
+\subsubsection{Methodological comparison}
 We compared EPI to two alternative simulation-based inference techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
@@
 
-\subsubsection{Primary visual cortex}\label{methods_V1}
+\subsection{Primary visual cortex}\label{methods_V1}
+\subsubsection{V1 model}
 In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
@@
 
+\subsubsection{EPI details for the V1 model}
 For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
@@
 
+\subsubsection{Sensitivity analyses}
 In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
@@
 
-\subsubsection{Superior colliculus}\label{methods_SC}
+\subsection{Superior colliculus}\label{methods_SC}
+\subsubsection{SC model} \label{methods_sc_model}
 The ability to switch between two separate tasks throughout randomly interleaved trials, or ``rapid task switching," has been studied in rats, and midbrain superior colliculus (SC) has been show to play an important in this computation \cite{duan2015requirement}.
@@
 
+\subsubsection{Task accuracy calculation} \label{methods_sc_acc}
 The accuracies of each task $p_P$ and $p_A$ are calculated as
@@
 
+\subsubsection{EPI details for the SC model} \label{methods_sc_epi}
 Writing the EPI distribution as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
@@
 
-The EPI distribution of SC model connectivities producing rapid task switching has interesting structure.
-Throughout $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we see that the probability distribution is narrow in $hW$ (Fig. \ref{fig:SC}C).
-This suggests that rapid task switching is sensitive to changes in $hW$, but this is only a single parameter.
-The local structure of the distribution varies across parameter space, and thus the nature in which parameter combinations affect rapid task switching.
-From visual inspection, we may hypothesize that there are two distinct regimes, most easily visualized in the $sW$-$hW$ marginal distribution: one where $sW$ and $hW$ are correlated for greater $sW$ and one where $sW$ and $hW$ are anticorrelated for lesser $sW$.
-
-We sought two sets of parameters in this distribution representative of each regime, so that we could assess their implications on computation.
+\subsubsection{Regime identification with EPI} \label{methods_sc_regime}
+We sought two sets of parameters from $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ that were representative of each regime, so that we could assess their implications on computation.
 For fixed values of $hW$, we hypothesized that there are two modes: one in each regime of greater and lesser $sW$.
@@
 
-For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampled from the EPI distribution.
+For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampling from the EPI distribution.
 Each sample was assigned to the closest mode $\mathbf{z}^*(hW_{\text{fixed}}, r)$.
@@
 
+\subsubsection{Sensitivity analysis} \label{methods_sc_hessian}
 At each mode, we measure the sensitivity dimension (that of most negative eigenvalue in the Hessian of the EPI distribution) $\mathbf{v}_1(\mathbf{z}^*)$.
@@
 
+\subsubsection{Connectivity eigendecomposition and processing modes} \label{methods_sc_eig}
 To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same basis vectors $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
@@
 
-Through these connectivity perturbation analyses, we found that increasing $\lambda_{\text{task}}$ strongly reduced Pro accuracy in regime 1, yet strongly reduced Anti accuracy in regime 2.
-This suggests that stronger task representations can inhibit both Pro and Anti task performance in different contexts.
-Furthermore, changing $\lambda_{\text{task}}$ in either direction decreases Anti performance in regime 1, showing that Anti task performance in regime 1 is dependent on a specific level of task representation.
-We also found that with increasing $\lambda_{\text{diag}}$, Pro accuracy increased in both regimes, but there were opposite effects on Anti accuracy.
-In regime 1, stronger amplification of diagonal population patterns decreased Anti accuracy, while in regime 2 accuracy increased.
-These findings give us an understanding of the mechanistic differences in computation enabling rapid task switching in each regime.
-
 %
@@
 
+\subsubsection{Optogenetic silencing modeling} \label{methods_sc_opto}
 We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.

### 2021-04-20 written/eLifeRevision/Response.tex
response and manuscript updates

@@
 %\usepackage{gentium}
-\usepackage{mathptmx} % Use Times Font
 
@@
 \linespread{1.2} % Set linespace
-\usepackage[a4paper, lmargin=0.1666\paperwidth, rmargin=0.1666\paperwidth, tmargin=0.1111\paperheight, bmargin=0.1111\paperheight]{geometry} %margins
+\usepackage[a4paper, lmargin=1in, rmargin=1in, tmargin=1in, bmargin=1in]{geometry} %margins
 %\usepackage{parskip}
@@
 \usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
+\usepackage{enumitem}
+\usepackage{amsmath}
+\usepackage{csquotes}
 
@@
 
-We thank the reviewers for the very constructive feedback.
-To address these points of concern, we have made extensive improvements to the manuscript.
-We have made significant changes to our writing and explanation of the inference technique, made an extensive comparison to other parameter inference methods, and increased the quality and depth of our scientific analyses, which yield strong theoretical results.
+We thank the reviewers for the excellent and constructive feedback.
+These valid criticisms of the original manuscript impressed upon us the importance of making key improvements to this research project.
+We spent more than half of a year advancing this work according to the reviewer requests and suggestions, and we are excited to present them to you in a dramatically improved revision.
+
+We have made significant changes to our writing and explanation of emergent property inference (EPI), done extensive comparisons of EPI to other parameter inference methods, and have seriously increased the quality and depth of our scientific analyses, which now yield strong theoretical results.
 Our model analyses now focus on analyzing the rich structure of parameter distributions, which the deep probability distributions of EPI make possible.
@@
 \begin{itemize}
-\item We significantly changed how we explain the parameter inference technique (EPI) to clarify the methodological contribution of this research.
-We now emphasize the novelty and scientific value of conditioning probability distributions on theoretical criteria of interest (emergent properties), in contrast to noisy observed neural datasets.
-\item We show EPI’s superior scalability in parameter dimension to alternative likelihood-free inference techniques SMC-ABC and SNPE when inferring RNN connectivities exhibiting stable amplification.
-\item We show how derivations of excitatory variability in a model of primary visual cortex become analytically intractable as additional neuron types are added.
-We then use EPI to produce a clear insight that mathematical analysis did not produce.
-\item We discover multiple parametric regimes of computation in a model of superior colliculus from the fine distributional structure captured by EPI.
-We then use the deep probability distribution to mechanistically characterize each regime, and also find that both of the discovered regimes reproduce an important result from inactivation experiments.
+\item We completely overhauled our presentation of the parameter inference technique emergent property inference (EPI). See response to main concern \#1.
+\item We added an entirely new results section to compare EPI to alternative parameter inference techniques (SMC-ABC and SNPE) as parameter count increases, which demonstrates the advantages of EPI in high dimensional parameter spaces.
+See response to main concern \#2.
+\item EPI is used to understand how noise across neuron type populations governs excitatory variability in a model of primary visual cortex.  
+Here, EPI yields visually striking, novel insight where the conventional analtyic approach became infeasible with increasing neuron types.  See response to main concern \#4.
+\item EPI is used to discover multiple parametric regimes of ``rapid task switching" in a model of superior colliculus.  
+These regimes are efficiently identified and mechanistically characterized using the probabilistic modeling tools afforded by EPI, and we relate the inferred connectivities to results from optogenetic silencing experiments in rats.
+See response to main concern \#4.
 \end{itemize}
@@
 We then explain how formalizing the mathematical criteria of interest into a statistical feature of the model – the emergent property – we enable deep inference of mechanistic model parameters with respect to this property.
+We take care to make clear in our explanation that the role of deep networks in the method is to capture rich structure in parameter distributions:
 
-In Section 3.1 and 3.2, we present EPI through a motivational example of the STG subcircuit.
-The figure has been updated to reflect the modified presentation, and we more clearly emphasize the scientific tools afforded once a distribution is fit with EPI.
+Quote <Third and fourth paragraph of Section 3.2>
+
+Figure 1 has been completely redone to reflect the modified presentation, and simplify the pictorial representation of the method.
+
+We clarify the methodological contribution of this approach and now properly contextualize the method within the broader literature of biological systems modeling and statistical inference.
+We motivate EPI by identifying an incongruity between statistical inference techniques and the focus of theoretical neuroscience:
+The mismatch between empirical datasets and emergent properties comes at a cost.
+Primarily that scientists cannot fix what their inferred distributions produce.
+We now emphasize the novelty and scientific value of conditioning probability distributions on theoretical criteria of interest (emergent properties), in contrast to noisy observed neural datasets in an empirical evidence integration framework.
+
+Quote <early part of paragraph 5 from Introduction>
+
+And, we show that EPI is the only method that has explicit control over the predictions of its inferred distribution (Section 3.3, Figure 2C-D).
+
+Finally, we deemphasize the role of maximum entropy in the EPI algorithm.
+We show in Section 5.1.6 that EPI is a special case of variational inference, since maximum entropy is the same normative selection principle as Bayesian inference.
+Therefore, it makes sense to simply present EPI as a statistical inference technique that constrains the predictions of the interred parameters to be an emergent property.
+In the main text, we mention that EPI has a direct link to variational inference, and the details of the maximum entropy method are left for the technically proficient in the methods section.
 
@@
 We thank the reviewers for emphasizing the importance of this methodological comparison.
-In Section 3.3, we provide a direct comparison of EPI to alternative likelihood-free inference techniques SMC-ABC and SNPE by inferring RNN connectivities that exhibit stable amplification.
-These comparisons evaluate both wall time and simulation count, and we explain how each algorithm was run in its ideal hardware setting.
+In Section 3.3, we provide a direct comparison of EPI to alternative simulation-based inference techniques SMC-ABC and SNPE by inferring RNN connectivities that exhibit stable amplification.
+These comparisons evaluate both wall time (Fig.
+2A) and simulation count (Fig.
+2B), and we explain how each algorithm was run in its preferred hardware setting in Section 5.3.4.
 
-In this analysis, we demonstrate the superior scalability of deep inference techniques (EPI and SNPE) to the state-of-the-art random sampling technique (SMC-ABC) .
-We then show that SNPE does not constrain the properties of inferred parameters, resulting in many unstable and nonamplified models.
-Furthermore, we push the limits of SNPE through targeted hyperparameter modifications, and show that for this model EPI scales to a higher dimensional parameter space than SNPE.
+In this analysis, we demonstrate the improved scalability of deep inference techniques (EPI and SNPE) with respect to the state-of-the-art approximate Bayesian computation technique (SMC-ABC).
+While controlling for architecture size (Fig.
+RNN1), we push the limits of SNPE through targeted hyperparameter modifications (Fig.
+RNN3), and show that EPI scales to higher dimensional RNN connectivites producing stable amplification than SNPE.
+Furthermore, we emphasize that SNPE does not constrain the properties of the inferred parameters; many connectivities inferred by SNPE result in unstable or nonamplified models (Fig. 2C, Fig. RNN2).
 
@@
 Indeed SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.  
-However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a trade-off between likelihood-free inference approaches.
-Unlike SNPE, EPI leverages gradients of the model to improve efficiency (Section 3.3).
+However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a trade-off between simulation-based inference approaches.
+
+Unlike SNPE, EPI leverages gradients of the emergent property throughout optimization, which may lead to better efficiency and scalability (Section 3.3).  
 The emergent properties of many models in neuroscience are tractably differentiable, four of which we analyze in this manuscript ranging across levels of biological realism, network size, and computational function.
-Furthermore, EPI focuses the entire expressivity of the approximating deep probability distribution on a single distribution, rather than spreading this expressivity to some degree across the chosen training distribution of amortized posteriors in SNPE (Section 5.1.1 Related Approaches).
+This trade-off is explained at the end of Section 3.3:
+
+\begin{displayquote}
+Lines X-Y\\
+``Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
+Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
+This can be invaluable for understanding how RNNs produce complex emergent phenomena.
+Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section ref{results\_motivating}).
+However, conditioning on the pyloric rhythm cite{marder1992dynamic} in a model of the pyloric subnetwork model cite{prinz2004similar} proved to be prohibitive with EPI.
+The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
+In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach cite{gonccalves2019training}."
+\end{displayquote}
+
+Furthermore, EPI focuses the entire expressivity of the approximating deep probability distribution on a single distribution, rather than spreading this expressivity to some uncharacterized degree across the chosen training distribution of amortized posteriors in SNPE (Section 5.1.1 Related Approaches).
+
+\begin{displayquote}
+Lines X-Y\\
+``The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
+This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
+Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
+Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section ref{methods\_ME\_EF})."
+\end{displayquote}
 
@@
 EPI learns parameter distributions whose predictions are constrained to produce the emergent property.
-We show in Supplementary Figure RNN2 that SNPE does not control the statistical properties of its predictions, resulting in the inference of many parameters that are not consistent with the desired emergent property.
+We show in Figure 2 and Supplementary Figure RNN2 that SNPE does not control the statistical properties of its predictions, resulting in the inference of many parameters that are not consistent with the desired emergent property.
+
+\begin{displayquote}
+Lines X-Y\\
+``No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig. 2C, blue).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. 2C, green).
+When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. 2C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions."
+\end{displayquote}
 
@@
 
-We thank the reviewers for explaining how they found some of the presentation misleading.
-We have taken serious care in this manuscript to clarify a.) what is novel, appreciable scientific insight provided by EPI, as well as b.) which scientific analyses are made possible by EPI.
+We thank the reviewers for explaining how they found some of the
+presentation misleading.  We have taken serious care in this manuscript to clarify a.)
+what is novel, appreciable scientific insight provided by EPI, as well as b.) which scientific
+analyses are made possible by EPI.
 
-a.)	In the revised manuscript we clarify that novel theoretical insights are not being made into the STG subcircuit model or the recurrent neural network models.
+\begin{enumerate}[label=(\alph*)]
+\item  In the revised manuscript we clarify that novel theoretical insights are not being made into the STG subcircuit model or the recurrent neural network models.
 The STG subcircuit serves as a motivational example to explain how EPI works, and we use RNNs exhibiting stable amplification as a substrate for scalability analyses.
-We do produce strong, theoretical insights into a model of primary visual cortex (Section 3.4) and superior colliculus (Section 3.5).
+
+\begin{displayquote}
+Lines X-Y \\
+First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion cite{goldman2001global, gutierrez2013multiple}.
+Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
+``"
+\end{displayquote}
+
+We do produce strong theoretical insights into a model of primary visual cortex (Section 3.4) and superior colliculus (Section 3.5).
 These analyses have substantially more depth than the previous manuscript.
 
-b.)	The ability to infer a flexible approximation to a probability distribution constrained to produce an emergent property is novel in its own right (Figure 2).
+\begin{displayquote}
+Lines X-Y \\
+``In a model of primary visual cortex (V1) cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
+Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
+Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials cite{duan2018collicular}.
+EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
+We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
+Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon."
+\end{displayquote}
+
+\item The ability to infer a flexible approximation to a probability distribution constrained to produce an emergent property is novel in its own right (Figure 2).
 The deep probability distribution fit by EPI facilitates the mode identification (via gradient ascent of the parameter log probability) and sensitivity measurement (via the measurement of the eigenvector of the Hessian at a parameter value).
-These mode identifications and sensitivity measurements are done in Sections 3.1, 3.4, and 3.5.
+These mode identifications and sensitivity measurements are done in Sections 3.1 (Fig.
+1E), 3.4 (Fig.
+3E), and 3.5 (Fig.
+4C).
 By using this mode identification technique along the ridges of high parameter probability in the SC model, we identify the parameters transitioning between the two regimes.
-Finally, the sensitivity dimensions identified by the SC model facilitated regime characterization through perturbation analyses.
+Finally, the sensitivity dimensions of the SC model identified by EPI facilitated regime characterization through perturbation analyses (Fig.
+4D, Fig.
+5C).
 
-Importantly, we do not claim that these theoretical insights were necessary with the techniques in b.).
-One could have come to these conclusions via brute force techniques, although less efficiently.
-The point is that we have brought deep inference to bear on theoretical neuroscience in a way that is principled, efficient, and scientifically compelling.
+\end{enumerate}
+
+Importantly, we do not claim that these theoretical insights were necessarily dependent on using the techniques in b.).  
+One could have come to these conclusions via various combinations of techniques mentioned in Section 5.1.1 Related Methods.
+In the case of the V1 model inference, the main point is to indicate that such insight can be afforded by EPI and its related methods, in contrast to the analytic derivations emblematic of practice in theoretical neuroscience.
+\begin{displayquote}
+Lines X-Y \\
+``EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
+We show that with each neuron-type population added to this E-I model, calculations of excitatory variability with respect to noise parameters become unruly and challenging to work with (see Section ref{methods\_V1\_complexity}).
+This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical."
+\end{displayquote}
+In the case of the SC model inference, random sampling would have taken prohibitively long, and it is unclear how the continuum between the two connectivity regimes would have been identified with alternative techniques:
+\begin{displayquote}
+Lines X-Y \\
+``The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section ref{methods\_SC})."
+\end{displayquote}
 
@@
 This 5-neuron model represents the IC neuron (hub) and its coupling to the pyloric (fast) or gastric mill (slow) subcircuit rhythms.
-In the introductory text, we refer to this model as the “STG subcircuit” model (rather than “STG model”), and we clarify what is being modeled in Results Section 3.1.
+In the introductory text, we refer to this model as the “STG subcircuit” model (rather than “STG model”), and we better clarify what aspect of the STG is being modeled in Results Section 3.1.
+
+\begin{displayquote}
+Lines X-Y \\
+``A subcircuit model of the STG (Guttierez et al. 2013) is shown schematically in Figure 1A.
+The fast population ($f1$ and $f2$) represents the subnetwork generating the pyloric rhythm and the slow population ($s1$ and $s2$) represents the subnetwork of the gastric mill rhythm.
+The two fast neurons mutually inhibit one another, and spike at a greater frequency than the mutually inhibiting slow neurons.
+The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions."
+\end{displayquote}
+
+The difference between this model and the STG model of the pyloric subnetwork is emphasized in Section 3.3:
+
+\begin{displayquote}
+Lines X-Y \\
+``However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
+The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions."
+\end{displayquote}
 
@@
 We have incorporated them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.
+
+\begin{displayquote}
+Lines X-Y
+``"
+\end{displayquote}
 
@@
 
-In our examples, we only infer distributions of parameters with the same units.
+In our examples, we only infer distributions of parameters with the same units, so issue should not draw concern over the validity of our model analyses.
 As suggested, the EPI solution will differ according to relative scaling of parameter values under the maximum entropy selection principle.
@@
 A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.
-It makes most sense to make sure EPI is exploring through parameter space through a standard range of values, and that sensitivity measurements are interpretable.
+It makes most sense to make sure EPI is exploring through parameter space through an industry-recommended range of values, and that sensitivity measurements are interpretable.
+
+\textbf{SB: Should we add text about this in the paper?}
 
@@
 
-(*Do this?) We have edited the introduction to more appropriately characterize the practice of theoretical neuroscience.
+(*Do this?) We have edited the introduction to more appropriately characterize the
+practice of theoretical neuroscience.
 
@@
 
-asdfasdfa
+We clarify in point 4 above that we are not claiming to produce novel, appreciable scientific insight about the STG subcircuit model, which is used as a motivation example.
+The takeaway is that the conductance parameters producing intermediate hub frequency belong to a complex 2-D distribution, which EPI captures accurately, and that EPI can tell us the parameter changes away from the prototypical configuration that change hub frequency the most or least.
+For example, for increases in $g_{\text{el}}$ and $g_{\text{synA}}$ according to the proportions of the degenerate dimension of parameter space, intermediate hub neuron frequency will be preserved in this model.
+This suggests that in the real STG neural circuit, the IC neuron will remain at an intermediate frequency between the pyloric and gastric mill rhythms if parameter changes are made along such
+a dimension.
+
+I should probably add some text with a big picture takeaway at the end of Section 3.2.
 
@@
 
+We thank the reviewer for their positive comments.
+Training deep networks for visual processing used to be impossible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.
+Similarly, training deep networks via EPI to learn parameter distributions became easier throughout this research project as we learned through trial and error what works well.
+In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge while capturing complex structure much more regularly.
+(*Make empirical statements about regularity in the paper?
+Cite GLOW or Real NVP?) A big part of the value of this manuscript is its explanation of hyperparameter choices, and the extensive set of examples in the online code github.
+Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: augmented Lagrangian constant, normalizing flow architecture.
+
+In general, we cannot know if we have arrived at the global maximum entropy distribution for a given emergent property.
+The reviewer is correct to point out that multiple distributions may satisfy the emergent property and have different levels of entropy.
+In the new manuscript, we present the method as an inference technique without focusing very greatly on maximum entropy, since it tends to distract and confuse the reader.
+We derive an analytic equivalence to variational inference (Section 5.1.6) showing that a.) EPI is validly conceptualized as an inference method, and b.) to emphasize that maximum entropy is the normative selection principle of Bayesian inference methods in general.
+Thus, the concern of not having the globally optimal inferred distribution is the same that applies to all other inference techniques.
+
+Practically, this has scientific implications.
+It means that we may be missing important structure in the inferred distribution, or we may be missing additional modes in parameter space.
+To handle this methodologically, we always run EPI with multiple random seeds, and select the distribution that has converged with the greatest entropy for scientific analysis.
+
 \textbf{2.
@@
 
+This analysis is no longer in the manuscript.
+
 \textbf{3.
@@
 For example, for the V1 model it seems that you need to look at 48 figures for 4 variables, and I believe this would scale as O(n$^2$) with n variables.
-This seems to require an unsustainable amount of manual work?
-}
+This seems to require an unsustainable amount of manual work?}
+
+We refer the reviewer to Figure 2 and Section 3.3 for scalability analysis.
+The scaling analysis treats the issue of discovering parameters that produce an emergent property in high-dimensional parameter spaces, which is an important question by itself.
+Another important question the reviewer brings up is how well can we handle the high-dimensional parameter distributions that EPI produces?
+Indeed, these distributions become more challenging to understand and visualize in high dimensions.
+This is where the sensitivity measurements appearing in sections 3.1, 3.4, and 3.5 can be extra useful.
+Even in high dimensions, trained deep probability distributions offer tractable quantitative assessments of how parametric combinations affect the emergent property that was conditioned upon.
 
@@
 }
+
+In the current manuscript, we make sure to explain all choices of the emergent property means and variances, and why our conclusions are not spuriously affected by ambiguity in these choices.
+
+Quote <All instances of emergent property explanation>
 
@@
 
-Things I will say \\
+Thank you for the comments… (*refer to exact lines of examples of this.) \\
 
@@
 
+We thank the reviewer for their positive comments.
+We have taken this feedback into serious account and its requests are reflected in multiple sections of the paper.
+In our updated model analyses, our novel scientific findings come precisely from the rich, fine structure (Fig.
+3E, 4C) of the inferred parameter distributions captured by EPI.
+These are “validated” in the sense that they pass a hypothesis testing criteria for emergent property convergence, and are the greatest entropy distributions across all random seeds for EPI optimization.
+See response to R1 point 1 concerning global maximum entropy solutions.
+
+We find the 2x2 linear dynamical system example to be a good ground truth example.
+It has a multimodal, intricately structured posterior, with analytically derived contours matching the EPI distribution.
+It’s also a digestible example for a broad audience.
+Further ground truth of the underlying maximum entropy flow network (MEFN) algorithm can be found for the Dirichlet exponential family distribution in Loaiza Ganem et al.
+2017.
+Comparisons to random-sampling techniques are made in Section 3.3.
+
 \textbf{2) It was not obvious whether EPI actually scales well to higher dimensions and how much computation it would take (there is one claim that it 'should scale reasonably').
@@
 
+The reviewer is right to point out the importance of a scaling analysis.  Please see the new Section 3.3.
+
 \textbf{3) For the STG-like example, this was done for a very simple model that was motivated by the STG but isn't based on experimental recordings.
@@
 
+(* Looks like we need to really add a para about limitations.)
+
 \textbf{4) The discussion should be expanded to be more specific about what problems the authors think the model is, or is not, appropriate for.
@@
 }
+
+(* I think we should add a para to discussion about this.)
 
@@
 
+We do not use a Schur decomposition to analyze the SC model in the current manuscript.
+
 \textbf{2) p. 31: usually one reserves the variable omega for angular frequencies: omega = 2*pi*f where f is frequency. }
+
+(* Will change).
 
@@
 
+Thank you for these references, we have incorporated them where appropriate. (TODO)
+
 \textbf{Additional data files and statistical comments:}

### 2021-04-20 written/eLifeRevision/Bittner_eLife2020.tex
response and manuscript updates

@@
 
-\subsection{Scaling inference of RNN connectivity with EPI} \label{results_LRRNN}
+\subsection{Scaling inference of RNN connectivity with EPI} \label{results_RNN}
 Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinsically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
@@
 For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:LRRNN}C, green).
-When using SNPE the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
+When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
 We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
@@
 burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, gradient-free approaches like SNPE have proved to be a powerful option \cite{gonccalves2019training}.
+In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach \cite{gonccalves2019training}.
 In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.
@@
 Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
-This insight was made possible by using the probabilistic tools afforded by EPI: we identified two regimes and the continuum of connectivities between them, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).
+The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).
 
@@
 On the other hand, EPI is a deep inference technique designed to condition directly on emergent properties, such that the parameter distribution only produces the computation of interest.
-EPI is thus ideally suited for questions in theoretical neuroscience, and we show that it has superior scaling properties to these other inference techniques (see Section \ref{results_LRRNN}).
+EPI is thus ideally suited for questions in theoretical neuroscience, and we show that it has superior scaling properties to these other inference techniques (see Section \ref{results_RNN}).
 
@@
 %A primary challenge is that the number of parameters scales dramatically with the number of neurons, limiting analysis of its parameter space.
-%We see in Section \ref{results_LRRNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
+%We see in Section \ref{results_RNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
 %EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
@@
 Finally, one important differentiating factor between EPI and sequential simulation-based inference methods is that EPI leverages gradients $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ during optimization.
-These gradients can improve convergence time and scalability, as we have shown on an example conditioning low-rank RNN connectivity on the property of stable amplification (see Section \ref{results_LRRNN}).
+These gradients can improve convergence time and scalability, as we have shown on an example conditioning low-rank RNN connectivity on the property of stable amplification (see Section \ref{results_RNN}).
 With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these model gradients when they are tractable.

### 2021-04-21 written/eLifeRevision/Response.tex
automated quoting and referencing between response and manuscript

@@
 \usepackage[swedish]{babel} % Swedish translations
-\usepackage[pdftex,linkcolor=black,pdfborder={0 0 0}]{hyperref} % Format links for pdf
 \usepackage{calc} % To reset the counter in the document after title page
@@
 \usepackage{csquotes}
+\usepackage{xr}
+\externaldocument{Bittner_eLife2020}
+\usepackage{xcite}
+\externalcitedocument{Bittner_eLife2020}
+\usepackage{qting}
+\quotefrom{Bittner_eLife2020}
 
-%-----------------------
-% Set pdf information and add title, fill in the fields
-%-----------------------
-\hypersetup{ 	
-pdfsubject = {},
-pdftitle = {},
-pdfauthor = {}
-}
 \setlength{\parindent}{0pt}
@@
 In the introduction, we now frame the method within the more general context of parameter inference techniques in neuroscience, rather than the context of recent advancements in machine learning.
-We have made concentrated efforts to simplify language, relate to all relevant existing methodology (Section 5.1.1 Related approaches), and precisely explain the novelty and utility of this deep inference technique.
+We have made concentrated efforts to simplify language, relate to all relevant existing methodology (Section \ref{methods_related} Related approaches), and precisely explain the novelty and utility of this deep inference technique.
 
@@
 
-And, we show that EPI is the only method that has explicit control over the predictions of its inferred distribution (Section 3.3, Figure 2C-D).
+And, we show that EPI is the only method that has explicit control over the predictions of its inferred distribution (Section \ref{results_RNN}, Figure \ref{fig:RNN}C-D).
 
@@
 Lines X-Y\\
-``Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
-Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
-This can be invaluable for understanding how RNNs produce complex emergent phenomena.
-Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section ref{results\_motivating}).
-However, conditioning on the pyloric rhythm cite{marder1992dynamic} in a model of the pyloric subnetwork model cite{prinz2004similar} proved to be prohibitive with EPI.
-The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach cite{gonccalves2019training}."
+``\quoting{rnn_gradients}"
 \end{displayquote}
@@
 Lines X-Y\\
-``The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
-This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
-Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
-Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section ref{methods\_ME\_EF})."
+``\quoting{amortization}"
 \end{displayquote}
@@
 Lines X-Y\\
-``No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig. 2C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. 2C, green).
-When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. 2C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions."
+``\quoting{rnn_constrained}"
 \end{displayquote}
@@
 Lines X-Y \\
-First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion cite{goldman2001global, gutierrez2013multiple}.
-Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
-``"
+``\quoting{intro_results1}"
 \end{displayquote}
@@
 Lines X-Y \\
-``In a model of primary visual cortex (V1) cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
-Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
-Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials cite{duan2018collicular}.
-EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
-We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
-Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon."
+``\quoting{intro_results2}"
 \end{displayquote}
@@
 Lines X-Y \\
-``EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
-We show that with each neuron-type population added to this E-I model, calculations of excitatory variability with respect to noise parameters become unruly and challenging to work with (see Section ref{methods\_V1\_complexity}).
-This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical."
+``\quoting{v1_summary}"
 \end{displayquote}
@@
 Lines X-Y \\
-``The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section ref{methods\_SC})."
+``\quoting{sc_summary}"s
 \end{displayquote}
@@
 Lines X-Y \\
-``A subcircuit model of the STG (Guttierez et al. 2013) is shown schematically in Figure 1A.
-The fast population ($f1$ and $f2$) represents the subnetwork generating the pyloric rhythm and the slow population ($s1$ and $s2$) represents the subnetwork of the gastric mill rhythm.
-The two fast neurons mutually inhibit one another, and spike at a greater frequency than the mutually inhibiting slow neurons.
-The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions."
+``\quoting{stg_subcircuit}"
 \end{displayquote}
 
-The difference between this model and the STG model of the pyloric subnetwork is emphasized in Section 3.3:
+The difference between this model and the STG model of the pyloric subnetwork is emphasized in Section \ref{results_RNN}:
 
+% Within another quote so need to copy.
 \begin{displayquote}
@@
 
-Thank you for pointing us to these great references on sensitivity analyses and applied dynamical systems.
+Thank you for pointing us to these references on sensitivity analyses and applied dynamical systems.  
+We acknowledge that our previous manuscript ...
 We have incorporated them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.
@@
 
-We have reached out to Woods Hole Course Project mentors where this work was discussed and explored in its early stages.  We are confident that there are no authorship issues.
+We have reached out to Woods Hole Course Project mentors where this work was discussed and explored in its early stages.
+James Fitzgerald and Dhruva Rhaman are happy with their acknowledgement, and do not insist on deeper involvement and authorship on this paper.
+Stephen Baccus has requested that we acknowledge the summer course, which we have agreed to do.
 
@@
 }
+
+\begin{itemize}
+\item Talk about new presentation of EPI.
+\item Talk about how its VI
+\item Refer to 2D LDS example.
+\item All examples are explained at a high level in main text and full details for implementation (and all code with interactive notebooks are online).
+\item Theoretical neuroscience is model evaluation.  Contrast that with philosophy. Ref Larry's Neuron perspective article.
+\end{itemize}
 
@@
 (*Make empirical statements about regularity in the paper?
-Cite GLOW or Real NVP?) A big part of the value of this manuscript is its explanation of hyperparameter choices, and the extensive set of examples in the online code github.
+Cite GLOW or Real NVP?) This manuscript has much value in its explanation of hyperparameter choices, and the extensive set of examples in the online code github.
 Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: augmented Lagrangian constant, normalizing flow architecture.
@@
 It means that we may be missing important structure in the inferred distribution, or we may be missing additional modes in parameter space.
-To handle this methodologically, we always run EPI with multiple random seeds, and select the distribution that has converged with the greatest entropy for scientific analysis.
+To handle this methodologically, we run EPI with multiple random seeds, and select the distribution that has converged with the greatest entropy for scientific analysis.
 
@@
 We refer the reviewer to Figure 2 and Section 3.3 for scalability analysis.
-The scaling analysis treats the issue of discovering parameters that produce an emergent property in high-dimensional parameter spaces, which is an important question by itself.
-Another important question the reviewer brings up is how well can we handle the high-dimensional parameter distributions that EPI produces?
+The scaling analysis addresses the question of the issue of parameter discovery with EPI in high-dimensional parameter spaces.
+
+Another important question the reviewer brings up is how well one can analyze the high-dimensional parameter distributions that EPI produces?
 Indeed, these distributions become more challenging to understand and visualize in high dimensions.
@@
 
-In the current manuscript, we make sure to explain all choices of the emergent property means and variances, and why our conclusions are not spuriously affected by ambiguity in these choices.
+In the current manuscript, we make sure to explain all choices of the emergent property constraints  Here, we show the description of each emergent property with equations omitted.
 
-Quote <All instances of emergent property explanation>
+\begin{displayquote}
+Section \ref{results_dgm}, Lines X-Y\\
+``\quoting{stg_ep_text1}~\quoting{stg_ep_text2}."
+\end{displayquote}
 
+\begin{displayquote}
+Section \ref{results_RNN}, Lines X-Y\\
+``\quoting{rnn_ep_text1}~... \quoting{rnn_ep_text2}~\quoting{rnn_ep_text3}"
+\end{displayquote}
+
+\begin{displayquote}
+Section \ref{results_V1}, Lines X-Y\\
+``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}."
+\end{displayquote}
+
+\begin{displayquote}
+Section \ref{results_SC}, Lines X-Y\\
+``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}"
+\end{displayquote}
 \textbf{Minor points }
@@
 
-Thank you for the comments… (*refer to exact lines of examples of this.) \\
+\begin{itemize}
+\item call out x as internal state and reference.
+\item ref to how we handle entropy
+\item talk about how we don't do this anymore.  ref hypothesis in sc
+\end{itemize}
 

### 2021-04-21 written/eLifeRevision/Bittner_eLife2020.tex
automated quoting and referencing between response and manuscript

@@
 \usepackage{cite}
+\usepackage{qting}
 
@@
 Equipped with this method, we prove out the potential of EPI by demonstrating its capabilities and presenting novel theoretical findings borne from its analysis.
-First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion \cite{goldman2001global, gutierrez2013multiple}.
-Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
-In a model of primary visual cortex (V1) \cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
+\q{intro_results1}{First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion \cite{goldman2001global, gutierrez2013multiple}.
+Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
+\q{intro_results2}{In a model of primary visual cortex (V1) \cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
 Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
@@
 We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
-Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon.
+Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon.}
 These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.
@@
 
-A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A.
+\q{stg_subcircuit}{A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A.
 The fast population ($f1$ and $f2$) represents the subnetwork generating the pyloric rhythm and the slow population ($s1$ and $s2$) represents the subnetwork of the gastric mill rhythm.
 The two fast neurons mutually inhibit one another, and spike at a greater frequency than the mutually inhibiting slow neurons.
-The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions.
+The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions.}
 The jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$, smooth connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.
@@
 Second, we stipulate that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
-We stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
-Mathematically, we define this emergent property with two statistical constraints: that the mean hub frequency is 0.55Hz,
-\begin{equation}\label{eq:EP_STG1}
+\q{stg_ep_text1}{We stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
+Mathematically, we define this emergent property with two statistical constraints: that the mean hub frequency is 0.55Hz,}
+\begin{equation}
  \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55 
 \end{equation}
-and that the variance of the hub frequency is moderate
-\begin{equation}\label{eq:EP_STG2}
+\q{stg_ep_text2}{and that the variance of the hub frequency is moderate}
+\begin{equation}
 \text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2.
@@
 }
-\label{fig:LRRNN}
+\label{fig:RNN}
 \end{figure}
@@
 We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   that produce stable amplification.
-Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
-$\lambda^s_1 > 1$, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W^\top}{2}$.
+\q{rnn_ep_text1}{Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
+$\lambda^s_1 > 1$}, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W^\top}{2}$.
 RNNs  with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will be stable with modest decay rate ($\text{real}(\lambda_1)$ close to its upper bound of 1) and exhibit modest amplification ($\lambda_1^s$ close to its lower bound of 1).
- EPI can naturally condition on this emergent property
+\q{rnn_ep_text2}{EPI can naturally condition on this emergent property}
 \begin{equation}\label{eq:EP_LRRNN}
@@
 \end{equation}
-under the notion that variance constraints with standard deviation 0.25 predicate that the vast majority of samples (those within two standard deviations) are within the specified ranges.
+\q{rnn_ep_text3}{under the notion that variance constraints with standard deviation 0.25 predicate that the vast majority of samples (those within two standard deviations) are within the specified ranges.}
 
@@
 As the number of neurons $N$ in the RNN is scaled, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
-\ref{fig:LRRNN}A).
-It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:LRRNN}B).
+\ref{fig:RNN}A).
+It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
 It is well known that ABC techniques struggle in dimensions greater than about 30 \cite{sisson2018handbook}, yet we were careful to assess the scalability of the more comparable approach SNPE.
@@
 
-No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig.
-\ref{fig:LRRNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:LRRNN}C, green).
-When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:LRRNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.
+\q{rnn_constrained}{No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig.
+\ref{fig:RNN}C, blue).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
+When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:RNN}C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.}
 
@@
 This agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}.
-Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
+\q{rnn_gradients}{Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
 Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
@@
 burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach \cite{gonccalves2019training}.
+In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach \cite{gonccalves2019training}.}
 In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.
-
-%\section{Discussion}
-%\subsection{EPI is a general tool for theoretical neuroscience} 
-%Biologically realistic models of neural circuits are comprised of complex nonlinear differential equations, making traditional theoretical analysis and statistical inference intractable. 
-%We advance the capabilities of statistical inference in theoretical neuroscience by presenting EPI, a deep inference methodology for learning parameter distributions of theoretical models performing neural computation.
-%We have demonstrated the utility of EPI on biological models (STG), intermediate-level models of interacting genetically- and functionally-defined neuron-types (V1, SC), and the most abstract of models (RNNs).
-%We are able to condition both deterministic and stochastic models on low-level emergent properties like spiking frequency of membrane potentials, as well as high-level cognitive function like posterior conditioning.
-%Technically, EPI is tractable when the emergent property statistics are continuously differentiable with respect to the model parameters, which is very often the case; this emphasizes the general applicability of EPI.
-%
-%In this study, we have focused on applying EPI to low dimensional parameter spaces of models with low dimensional dynamical states.
-%These choices were made to present the reader with a series of  interpretable conclusions, which is more challenging in high dimensional spaces.
-%In fact, EPI should scale reasonably to high dimensional parameter spaces, as the underlying technology has produced state-of-the-art performance on high-dimensional tasks such as texture generation \cite{loaiza2017maximum}.
-%Of course, increasing the dimensionality of the dynamical state of the model makes optimization more expensive, and there is a practical limit there as with any machine learning approach.
-%Although, theoretical approaches (e.g.
-\cite{mastrogiuseppe2018linking}) can be used to reason about the wholistic activity of such high dimensional systems by introducing some degree of additional structure into the model.
 
@@
 
-We quantify levels $y$ of E-population variability with the emergent property
+\q{v1_ep_text1}{We quantify levels $y$ of E-population variability with the emergent property}
 \begin{equation}\label{eq:EP_V1}
@@
 \end{equation}
-where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-In the following analyses, we compare levels of 5Hz and 10Hz, and select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.
+\q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
+In the following analyses, we compare levels of 5Hz and 10Hz, and select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
 
@@
 
-EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
+\q{v1_summary}{EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
 We show that with each neuron-type population added to this E-I model, calculations of excitatory variability with respect to noise parameters become unruly and challenging to work with (see Section \ref{methods_V1_complexity}).
-This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.
+This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
 While EPI can be used to investigate fundamental aspects of sensory processing, in the next section, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to  behavioral experiments.
@@
 Rapid task switching is formalized mathematically as an emergent property with two statistics: accuracy in the Pro task $p_P(\mathbf{x}; \mathbf{z})$ and Anti task $p_A(\mathbf{x}; \mathbf{z})$.
-We stipulate that accuracy be on average $.75$ in each task with variance $.075^2$
+\q{sc_ep_text1}{We stipulate that accuracy be on average $.75$ in each task with variance $.075^2$}
 \begin{equation}\label{eq:SC_EP}
@@
 \end{equation}
-75\% accuracy is a realistic level of performance in each task, and with the chosen variance, inferred models will not exhibit fully random responses (50\%), nor perfect performance (100\%).
+\q{sc_ep_text2}{75\% accuracy is a realistic level of performance in each task, and with the chosen variance, inferred models will not exhibit fully random responses (50\%), nor perfect performance (100\%).}
 
@@
 Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
-The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).
+\q{sc_summary}{The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).}
 
@@
 Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
-SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:LRRNN}).
+SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:RNN}).
 Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again from scratch with a new population of initialized particles to obtain additional samples.
@@
 
-The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
+\q{amortization}{The approximating fidelity of the deep probability distribution in sequential approaches is optimized to generalize across the training distribution of the conditioning variable.
 This generalization property of the sequential methods can reduce the accuracy at the singular posterior of interest.
 Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
-Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section \ref{methods_ME_EF}).
+Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section \ref{methods_ME_EF}).}
 Since EPI distributions are defined by the mean $\bm{\mu}$ of their statistics, there is the well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} that prohibits an amortization based approach.
@@
 
-For EPI in Fig. \ref{fig:LRRNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
+For EPI in Fig. \ref{fig:RNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
 We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
-\ref{fig:LRRNN}B blue line, and Fig.
-\ref{fig:LRRNN}C-D blue).
+\ref{fig:RNN}B blue line, and Fig.
+\ref{fig:RNN}C-D blue).
 
@@
 
-In Figure \ref{fig:LRRNN}C-D, we show samples form the random seed resulting in emergent property convergence at greatest entropy (EPI), the random seed resulting in greatest validation log probability (SNPE), and the result of all converged random seeds (SMC).
+In Figure \ref{fig:RNN}C-D, we show samples form the random seed resulting in emergent property convergence at greatest entropy (EPI), the random seed resulting in greatest validation log probability (SNPE), and the result of all converged random seeds (SMC).
 

### 2021-04-21 written/eLifeRevision/Bittner_eLife2020.tex
additional SC supp fig

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figSC8/figSC8.pdf}
+\end{center}
+\caption{\small (SC8): 
+\textbf{A}. Entropy throughout optimization. 
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+}
+\label{fig:SC8}
+\end{figure}
+
+The bend in the EPI distribution is not a spurious result of the EPI optimization.
+The structure discovered by EPI matches the shape of the set of points returned from brute-force random sampling (Fig. \ref{fig:SC8}A) 
+These connectivities were sampled from a uniform distribution over the range of each connectivity parameter, and all parameters producing accuracy in each task within the range of 60\% to 90\% were kept.
+This set of connectivities will not match the distribution of EPI exactly, since it is not conditioned on the emergent property.
+For example the parameter set returned by the brute-force search is biased towards lower accuraciess (Fig. \ref{fig:SC8}B).
+
 \subsubsection{Regime identification with EPI} \label{methods_sc_regime}
@@
 
-\subsubsection{Optogenetic silencing modeling} \label{methods_sc_opto}
+\subsubsection{Modeling optogenetic silencing.} \label{methods_sc_opto}
 We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.

### 2021-04-22 written/eLifeRevision/Bittner_eLife2020.tex
abstract update

@@
 Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
-The operation of these mechanistic circuits, like all models, critically depends on the choices of model parameters.
-A key process in circuit modeling is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
+The operation of these circuits, like all models, critically depends on the choices of model parameters.
+A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
 To solve challenging inverse problems modeling neural datasets, neuroscientists have used statistical inference techniques to much success. 
-However, most research in theoretical neuroscience focuses on how computation emerges in biologically interpretable circuit models, and how the model parameters govern computation; it is not focused on the latent structure of empirical models of noisy experimental datasets.
-In this work, we present a novel technique that brings the power and versatility of the probabilistic modeling toolkit to theoretical inverse problems.
-Our method uses deep neural networks to learn parameter distributions with rich structure that have specific computational properties in biologically relevant models.
-This methodology is explained through a motivational example inferring conductance parameters in an STG subcircuit model.
-Then, with RNNs of increasing size, we show that only EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
-In the remainder of this work, we explain novel theoretical insights through the examination of intricate parametric structure in complex circuit models.
-In a model of primary visual cortex with multiple neuron-types, where analysis becomes untenable with each additional neuron-type, we discovered how noise distributed across neuron-types governs the excitatory population.
-Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials. 
-We also found that all task-switching connectivities in this model reproduce behaviors from inactivation experiments, further establishing this hypothesized circuit model.
+However, when theorizing circuit models, scientists predominantly focus on reproducing established computational properties rather than explaining a given neural dataset well.
+In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
+Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
+This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion (STG).
+Then, with recurrent neural networks of increasing size, we show that only EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
+In the remainder of this work, we present novel theoretical findings gained through  the examination of complex parametric structure captured by EPI.
+In a model of primary visual cortex, we discovered how connectivity with multiple inhibitory subtypes shapes variability in the excitatory population.
+Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials, mechanistically characterized each regime using probabilistic tools afforded by EPI, and found conditions where these circuit models reproduce results from optogenetic silencing experiments.
 Beyond its scientific contribution, this work illustrates the variety of analyses possible once deep learning is harnessed towards solving theoretical inverse problems.
@@
 
-Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
+Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena in primary visual cortex (V1).
 In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
-Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
-However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+Recent theoretical work \cite{hennequin2018dynamical} shows that stabilized E-I models reproduce the effect of variability suppression \cite{Churchland2010}.
+However, the way in which E-I recurrence structures variability into shared and private dimensions \cite{semedo2019cortical} is not understood.
+Furthermore, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
 composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
 \ref{fig:V1}A) \cite{pfeffer2013inhibition}.
-%Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools.
-While research has shown that V1 only shares specific dimensions of neuronal variability with downstream areas \cite{semedo2019cortical}, the role played by recurrent dynamics and the connectivity across neuron-type populations is not understood.
-Here, in a model of V1 with biologically realistic connectivity, we use EPI to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
+Here, we use EPI on a model of V1 with biologically realistic connectivity to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
 
@@
 \ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}).
-Each population also experiences an external input $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
-There is also an additive noisy input $\bm{\epsilon}$ parameterized by variances for each neuron type population $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
+Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}) and
+an external input with mean $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
+The external input has an additive noisy component $\bm{\epsilon}$ with variance $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
 This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Fig. \ref{fig:V1}B).
@@
 generalized to have multiple inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
-Stochasticity and inhibitory multiplicity introduce substantial complexity to mathematical derivations (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
+Stochasticity and inhibitory multiplicity introduce substantial complexity to the mathematical treatment of this problem (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
 Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure}, and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability.
@@
 \q{v1_summary}{EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
-We show that with each neuron-type population added to this E-I model, calculations of excitatory variability with respect to noise parameters become unruly and challenging to work with (see Section \ref{methods_V1_complexity}).
+In a simplified model ($\tau = \tau_{\text{noise}}$), it can be shown that surfaces of equal variance are ellipsoids as a function of $\bm{\sigma}$.
+Nevertheless, the sensitive and degenerate parameters are challenging to predict mathematically, since the covariance matrix depends on the steady-state solution of the network \cite{hennequin2018dynamical,Gardiner2009}, and terms in the covariance expression increase quadratically with each additional neuron-type population (see also Section \ref{methods_V1_complexity}).
 This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
@@
 \subsubsection{V1 model}
-In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
+ E-I circuit models, rely on the assumption that inhibition can be studied as an indivisible unit, despite ample experimental evidence showing that inhibition is instead composed of distinct elements \cite{tremblay2016}.
+In particular three types of genetically identified inhibitory cell-types -- parvalbumin (P), somatostatin (S), VIP (V) --
+compose 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and follow specific connectivity patterns (Fig.
+\ref{fig:V1}A) \cite{pfeffer2013inhibition}, which lead to cell-type specific computations \cite{Mossing2021,palmigiano2020structure}.
+Currently, how the subdivision of inhibitory cell-types, shapes correlated variability by reconfiguring  recurrent network dynamics is not understood.
+
+In the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, population rate responses $\mathbf{x}$ to mean input $\mathbf{h}$, recurrent input $W\mathbf{x}$ and slow noise $\bm{\epsilon}$ are governed by
 \begin{equation}
@@
 The noisy process is parameterized as
-\begin{equation}
+\begin{equation} \label{sigma_reparam}
 \tilde{\sigma}_\alpha = \sigma_\alpha \sqrt{1 + \frac{\tau}{\tau_{\text{noise}}}},
@@
 
-\subsubsection{Primary visual cortex: challenges to analysis}\label{methods_V1_complexity}
-TODO Agostina and I are putting this together now.
+\subsubsection{Primary visual cortex: Mathematical intuition and challenges}\label{methods_V1_complexity}
+
+The dynamical system that we are working with can be written as
+
+\begin{align}\label{SSSN_agos}
+\begin{split}
+ dx=\frac{1}{\tau}(-x+f(W x +h+\epsilon)) dt\\
+d \epsilon=-\frac{dt}{\tau_{\text{noise}}}\epsilon+\frac{\sqrt{2}}{\sqrt{\tau_{\text{noise}}}}\Sigma_\epsilon dW
+\end{split}
+\end{align}
+
+Where in this paper we chose 
+\begin{equation}
+\Sigma_\epsilon=\tau_{\text{noise}} 
+\begin{bmatrix} 
+\tilde{\sigma}_E & 0 & 0 & 0 \\
+0 & \tilde{\sigma}_P & 0 & 0 \\
+0 & 0 & \tilde{\sigma}_S & 0 \\
+0 & 0 & 0 & \tilde{\sigma}_V \\
+\end{bmatrix}
+\end{equation}
+where $\tilde{\sigma}_\alpha$ is the reparameterized standard deviation of the noise for population $\alpha$ from Equation \ref{sigma_reparam}.
+
+In order to compute this covariance, we define $v=\omega x +h+\epsilon$ and $S=I-\omega f'(v))$, to re-write Eq. \eqref{SSSN_agos} as an 8-dimensional system:
+
+\begin{align}
+\begin{split}
+d\begin{pmatrix}
+\delta v \\
+ \epsilon
+\end{pmatrix}= -
+\begin{pmatrix}
+	S &- \frac{\tau_{\text{noise}}-\tau}{\tau \tau_{\text{noise}}}I\\
+	 0&\frac{1}{\tau_{\text{noise}}}I\\
+\end{pmatrix}
+\begin{pmatrix}
+	\delta v\\
+	\epsilon 
+\end{pmatrix} 
+dt +
+\begin{pmatrix}
+	0&\frac{\sqrt{2}}{\sqrt{\tau_{\text{noise}}}} \Sigma_\epsilon\\
+	0&\frac{\sqrt{2}}{\sqrt{\tau_{\text{noise}}}} \Sigma_\epsilon \\
+\end{pmatrix}    d\mathbf{W} 
+\end{split}
+\end{align}
+
+Where $d\mathbf{W}$ is a vector with the private noise of each variable.
+The $d\mathbf{W}$ term is multiplied by a non-diagonal matrix is because the noise that the voltage receives is the exact same than the one that comes from the OU process and not another process.
+The solution of this problem is given by the Lyapunov Equation \cite{hennequin2018dynamical,Gardiner2009}:
+
+\begin{align}
+\begin{split}
+\begin{pmatrix}
+	S & -\frac{\tau_{\text{noise}}-\tau}{\tau \tau_{\text{noise}}}I\\
+	 0&\frac{1}{\tau_{\text{noise}}}I\\
+\end{pmatrix}
+\begin{pmatrix}
+	\Lambda_v & \Lambda_c\\
+	 \Lambda_c^T&\Lambda_\epsilon\\
+\end{pmatrix}
++
+\begin{pmatrix}
+	\Lambda_v & \Lambda_c\\
+	 \Lambda_c ^T &\Lambda_\epsilon\\
+\end{pmatrix}
+\begin{pmatrix}
+	S^T & 0\\
+	 -\frac{\tau_{\text{noise}}-\tau}{\tau \tau_{\text{noise}}}I&\frac{1}{\tau_{\text{noise}}}I\\
+\end{pmatrix}
+=
+\begin{pmatrix}
+	\frac{2}{\tau_{\text{noise}}} \Lambda_\epsilon&\frac{2}{\tau_{\text{noise}}} \Lambda_\epsilon\\
+	\frac{2}{\tau_{\text{noise}}} \Lambda_\epsilon&\frac{2}{\tau_{\text{noise}}} \Lambda_\epsilon \\
+	\end{pmatrix}
+\end{split}
+\end{align}
+
+To obtain an equation for $\Lambda_v$, we solve this block matrix multiplication:
+
+\begin{align}
+\begin{split}
+S\Lambda_v +\Lambda_v S^T =\frac{2 \Lambda_\epsilon }{\tau_{\text{noise}} }+\frac{\tau_{\text{noise}}^2 -\tau^2 }{(\tau  \tau_{\text{noise}} )^2}\left( (\frac{1}{\tau_{\text{noise}}}I + S)^{-1} \Lambda_\epsilon+   \Lambda_\epsilon(\frac{1}{\tau_{\text{noise}}}I + S^T)^{-1}\right)  \\
+\end{split}
+\end{align}
+
+Which is another Lyapunov Equation, now in 4 dimensions.
+In the simplest case in which $\tau_{\text{noise}}=\tau$, the volatge is directly driven by white noise, and $\Lambda_v$ can be expressed in powers of $S$ and $S^T$.
+Because $S$ satisfies its own polynomial equation (Cayley Hamilton theorem), there will be 4 coefficients for the expansion of $S$ and 4 for $S^T$, resulting in 16 coefficients that define $\Lambda_v$ for a given $S$.
+Due to symmetry arguments\cite{Gardiner2009}, in this case the diagonal elements of the covariance matrix of the voltage will have the form:
+\begin{flalign} \label{GeneralEllipsoid}
+{\Lambda_{v_{ii}}}=\sum_{i=\{E,P,S,V\}} g_i(S)\sigma^2_{ii}
+\end{flalign}
+These coefficients $g_i(S)$ are complicated functions of the Jacobian of the system.
+Although expressions for these coefficients can be found explicitly, only numerical evaluation of those expressions determine which components of the  noisy input are going to strongly influence the variability of excitatory population.
+Showing the generality of this dependence in more complicated noise scenarios (e.g. $\tau_{\text{noise}} > \tau$ as in Section \ref{results_V1}), is the focus of current research.
 

### 2021-04-22 written/eLifeRevision/Response.tex
response to ground truth asks

@@
 \usepackage{csquotes}
+\usepackage{bm}
 \usepackage{xr}
@@
 
+We thank the reviewer for their positive comments and thoughtful feedback.
+We have made great efforts to provide several clear comparisons to ground truth (see response to concern \#1), and now provide an extensive methodological comparison to SMC-ABC and SNPE (see response to concern \#2).
+
 \textbf{Substantive concerns: \\
@@
 
-We thank the reviewer for their positive comments.
-We have taken this feedback into serious account and its requests are reflected in multiple sections of the paper.
-In our updated model analyses, our novel scientific findings come precisely from the rich, fine structure (Fig.
-3E, 4C) of the inferred parameter distributions captured by EPI.
-These are “validated” in the sense that they pass a hypothesis testing criteria for emergent property convergence, and are the greatest entropy distributions across all random seeds for EPI optimization.
-See response to R1 point 1 concerning global maximum entropy solutions.
+We thank the reviewer for pointing out the importance of ground truth comparisons in this manuscript.
+In this revision, we make ground truth comparisons via analytic derivations, empirical error contours, and brute-force sampling.
 
-We find the 2x2 linear dynamical system example to be a good ground truth example.
-It has a multimodal, intricately structured posterior, with analytically derived contours matching the EPI distribution.
-It’s also a digestible example for a broad audience.
-Further ground truth of the underlying maximum entropy flow network (MEFN) algorithm can be found for the Dirichlet exponential family distribution in Loaiza Ganem et al.
-2017.
-Comparisons to random-sampling techniques are made in Section 3.3.
+\textit{Analytic comparisons}:
+The 2x2 linear dynamical system is chosen as a worked example because it has multi-modal non-elliptical structure (Fig.
+\ref{fig:LDS1}), and its contours can be derived analytically (Fig.
+\ref{fig:LDS2}).
+Similarly, in Section \ref{methods_V1_complexity}, we derive the quadratic relationship between excitatory variability and input noise variability (in a simplified model) suggesting that the quadratic relationship uncovered by EPI (see Section \ref{results_V1}) is correct.
+
+\textit{Error contours}:
+In the motivation example, we compare the EPI inferred distribution of STG conductances to hub frequency contours (Figure \ref{fig:STG}E), which show that the non-elliptical parametric structure captured by EPI is in agreement with these contours.
+This general region of parameter space was labeled following grid search analyses in a previous study (Gutierrez et al. 2013, Figure 2, parameter regime G).
+
+\textit{Brute-force}:
+The EPI inferred distribution for rapid task switching in the SC model is sharply bent (Fig. \ref{fig:SC}), and matches the parameter set returned from random sampling (Figure \ref{fig:SC8}A).  
+We note that the brute-force parameter set is actually not the ground-truth solution, because it does not obey the constraints of the emergent property as the EPI dsitribution does (Figure \ref{fig:SC8}B).
+This can explain the spurious samples in the brute-force set that are not in the EPI inferred distribution.
+
+All EPI distributions shown in this manuscript are “validated” in the sense that they pass a hypothesis testing criteria for emergent property convergence; all EPI distributions produce their emergent properties.
+Finally, the underlying maximum entropy flow network (MEFN) algorithm is compared to a ground truth solution (Loaiza Ganem et al.
+2017, Figure 2) by deriving ground truth from the duality of maximum entropy distributions and exponential famillies (see Section \ref{methods_ME_EF}).
 

### 2021-04-24 written/eLifeRevision/Response.tex
first complete draft with response

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{csquotes}
@@
 \item EPI is used to understand how noise across neuron type populations governs excitatory variability in a model of primary visual cortex.  
-Here, EPI yields visually striking, novel insight where the conventional analtyic approach became infeasible with increasing neuron types.  See response to main concern \#4.
+Here, EPI yields visually striking, novel insight where the conventional analytic approach became infeasible with increasing neuron types.  See response to main concern \#4.
 \item EPI is used to discover multiple parametric regimes of ``rapid task switching" in a model of superior colliculus.  
@@
 We thank the reviewers for pointing out that our explanation of the method was too narrowly focused and hard to follow. 
-In the introduction, we now frame the method within the more general context of parameter inference techniques in neuroscience, rather than the context of recent advancements in machine learning.
-We have made concentrated efforts to simplify language, relate to all relevant existing methodology (Section \ref{methods_related} Related approaches), and precisely explain the novelty and utility of this deep inference technique.
+In the introduction, we present our method by focusing on the key aspect of EPI that differentiates it from other approaches to inverse problems.
+Specifically, we explain how current methodology infers the parameters producing computation by conditioning on \textit{exemplar datasets} (real or simulated), whereas in EPI we condition directly on the \textit{emergent properties} that define the computation.
 
-Primarily, we explain the incongruity between parameter inference techniques and the practice of theoretical neuroscience that motivates EPI: modern parameter inference techniques are designed to condition on noisy datasets rather than the mathematically defined criteria.
-This mismatch has consequences – primarily that data predicted from the inferred distribution are left unconstrained and often violate the desired conditioning criteria.
-We then explain how formalizing the mathematical criteria of interest into a statistical feature of the model – the emergent property – we enable deep inference of mechanistic model parameters with respect to this property.
-We take care to make clear in our explanation that the role of deep networks in the method is to capture rich structure in parameter distributions:
+\begin{displayquote}
+Lines 45-56\\
+``\quoting{intro_data_ep}"
+\end{displayquote}
 
-Quote <Third and fourth paragraph of Section 3.2>
+We now frame the method within the more general context of parameter inference techniques in neuroscience, rather than the context of recent advancements in machine learning.
+We have made concentrated efforts to simplify language and to relate to all relevant existing methodology.
 
-Figure 1 has been completely redone to reflect the modified presentation, and simplify the pictorial representation of the method.
+\begin{displayquote}
+Lines 57-67\\
+``\quoting{intro_epi}"
+\end{displayquote}
 
-We clarify the methodological contribution of this approach and now properly contextualize the method within the broader literature of biological systems modeling and statistical inference.
-We motivate EPI by identifying an incongruity between statistical inference techniques and the focus of theoretical neuroscience:
-The mismatch between empirical datasets and emergent properties comes at a cost.
-Primarily that scientists cannot fix what their inferred distributions produce.
-We now emphasize the novelty and scientific value of conditioning probability distributions on theoretical criteria of interest (emergent properties), in contrast to noisy observed neural datasets in an empirical evidence integration framework.
+In Section \ref{results_dgm} and \ref{methods_EPI}, we precisely explain the details of EPI, and in Section \ref{methods_AL_opt} we show how it is optimized.
+To improve clarity, we have changed the notation and presentation of emergent properties.
+Emergent properties are now denoted with $\mathcal{X}$ to signify an implicit dataset predicated by the emergent property.
+And rather than a vector of first- and second-moment constraints, we present the emergent property more readably as mean and variance constraints on emergent property statistics:
 
-Quote <early part of paragraph 5 from Introduction>
+\begin{displayquote}
+Line 705 \\
+``
+\begin{equation} \tag{11}
+\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\end{equation}"
+\end{displayquote}
 
-And, we show that EPI is the only method that has explicit control over the predictions of its inferred distribution (Section \ref{results_RNN}, Figure \ref{fig:RNN}C-D).
+In our revised writing, we deemphasize the role of maximum entropy in the EPI algorithm, because this has largely served as a distraction in our experience.
+We show in Section \ref{methods_VI} that EPI is a special case of variational inference, since maximum entropy is the same normative selection principle as variational bayesian methods.  
+Therefore, it makes sense to present EPI in the main text as a statistical inference technique that constrains the predictions of the inferred parameters to be an emergent property, and leave the details of the maximum entropy to the technically proficient in Section \ref{methods_EPI}.
 
-Finally, we deemphasize the role of maximum entropy in the EPI algorithm.
-We show in Section 5.1.6 that EPI is a special case of variational inference, since maximum entropy is the same normative selection principle as Bayesian inference.
-Therefore, it makes sense to simply present EPI as a statistical inference technique that constrains the predictions of the interred parameters to be an emergent property.
-In the main text, we mention that EPI has a direct link to variational inference, and the details of the maximum entropy method are left for the technically proficient in the methods section.
+\begin{displayquote}
+Lines 152-154\\
+``\quoting{results_ME}"
+\end{displayquote}
+
+Figure \ref{fig:STG} has been largely redone to reflect the modified presentation, and improve the pictorial representation of the method.
+In Section \ref{methods_RNN}, we now show that EPI is the only simulation-based inference method that controls the predictions of its inferred distribution (Fig. \ref{fig:RNN}C-D).
+
+Finally, we emphasize the utility of this deep inference technique for scientific inquiry in a new paragraph at the end of Section \ref{results_dgm}.
+
+\begin{displayquote}
+Lines 159-174\\
+``\quoting{results_hess}"
+\end{displayquote}
 
@@
 
-In this analysis, we demonstrate the improved scalability of deep inference techniques (EPI and SNPE) with respect to the state-of-the-art approximate Bayesian computation technique (SMC-ABC).
+In this analysis, we demonstrate the improved scalability of deep inference techniques (EPI and SNPE) with respect to the state-of-the-art approximate bayesian computation technique (SMC-ABC).
 While controlling for architecture size (Fig.
 RNN1), we push the limits of SNPE through targeted hyperparameter modifications (Fig.
-RNN3), and show that EPI scales to higher dimensional RNN connectivites producing stable amplification than SNPE.
+RNN3), and show that EPI scales to higher dimensional RNN connectivities producing stable amplification than SNPE.
 Furthermore, we emphasize that SNPE does not constrain the properties of the inferred parameters; many connectivities inferred by SNPE result in unstable or nonamplified models (Fig. 2C, Fig. RNN2).
@@
 The reviewers are right to point out these characteristics of EPI: it does not amortize across emergent properties, and it requires differentiability of the emergent properties of the model. 
-Indeed SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.  
+Indeed, SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.  
 However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a trade-off between simulation-based inference approaches.
@@
 \begin{displayquote}
-Lines X-Y\\
+Lines 240-251\\
 ``\quoting{rnn_gradients}"
@@
 \begin{displayquote}
-Lines X-Y\\
+Lines 788-793\\
 ``\quoting{amortization}"
@@
 \begin{displayquote}
-Lines X-Y\\
+Lines 229-236\\
 ``\quoting{rnn_constrained}"
@@
 \begin{displayquote}
-Lines X-Y \\
+Lines 69-74 \\
 ``\quoting{intro_results1}"
@@
 \begin{displayquote}
-Lines X-Y \\
+Lines 74-81 \\
 ``\quoting{intro_results2}"
@@
 \begin{displayquote}
-Lines X-Y \\
+Lines 302-310 \\
 ``\quoting{v1_summary}"
@@
 \begin{displayquote}
-Lines X-Y \\
-``\quoting{sc_summary}"s
+Lines 414-419 \\
+``\quoting{sc_summary}"
 \end{displayquote}
@@
 \begin{displayquote}
-Lines X-Y \\
+Lines 92-97 \\
 ``\quoting{stg_subcircuit}"
@@
 \begin{displayquote}
-Lines X-Y \\
+Lines 246-250 \\
 ``However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
@@
 Thank you for pointing us to these references on sensitivity analyses and applied dynamical systems.  
-We acknowledge that our previous manuscript ...
-We have incorporated them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.
+We have incorporated many of them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.
 
 \begin{displayquote}
-Lines X-Y
-``"
+Lines 809-822 \\
+``\quoting{structural_id}"
 \end{displayquote}
@@
 We have reached out to Woods Hole Course Project mentors where this work was discussed and explored in its early stages.
-James Fitzgerald and Dhruva Rhaman are happy with their acknowledgement, and do not insist on deeper involvement and authorship on this paper.
+James Fitzgerald and Dhruva Raman are happy with their acknowledgement, and do not insist on deeper involvement and authorship on this paper.
 Stephen Baccus has requested that we acknowledge the summer course, which we have agreed to do.
@@
 
-\begin{itemize}
-\item Talk about new presentation of EPI.
-\item Talk about how its VI
-\item Refer to 2D LDS example.
-\item All examples are explained at a high level in main text and full details for implementation (and all code with interactive notebooks are online).
-\item Theoretical neuroscience is model evaluation.  Contrast that with philosophy. Ref Larry's Neuron perspective article.
-\end{itemize}
+We thank the reviewer for their positive comments and thoughtful feedback.
+We have made serious effort to improve our presentation and explanation of EPI (see response to main concern \#1).
+Furthermore, we have focused on clearly motivating and describing each neural circuit model studied in this manuscript.
+Sufficient mathematical detail is written in each results section, while full details are presented in Methods.
+All code for training EPI on these models and their analysis are available in well-documented scripts and notebooks in our github repository.
+
+We modify our writing to clarify that EPI is not a key to all of theoretical neuroscience, but rather a powerful solution to inverse problems in neural circuit modeling.
+Inverse problems are indeed a major part of theoretical neuroscience.
+Their solution is necessary for the scientific evaluation of mathematical representations of neural circuits, and ultimately for progress in the field.
+We clarify this point in the introduction:
+
+\begin{displayquote}
+Lines 28-36 \\
+``\quoting{intro_first}"
+\end{displayquote}
 
@@
 To clarify, EPI does not execute posterior inference, because there is no empirical dataset or specified prior belief in EPI framework.
-However, we derive the relation of EPI to Bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.
+However, we derive the relation of EPI to bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.
+
+\begin{displayquote}
+Lines 986-989 \\
+``\quoting{ME_prior}"
+\end{displayquote}
 
@@
 
-(*Do this?) We have edited the introduction to more appropriately characterize the
-practice of theoretical neuroscience.
+We appreciate this feedback.  We have adapted the introduction to clarify that we are focusing on solving inverse problems in theoretical neuroscience.
 
-\textbf{(3) I am somewhat familiar with the stomatognathic circuit model, and so that is where I think I understand what they have done best.
+\textbf{(3) I am somewhat familiar with the stomatogastric circuit model, and so that is where I think I understand what they have done best.
 I don't understand what I should take away from their paper with regards to this model.
@@
 For example, for increases in $g_{\text{el}}$ and $g_{\text{synA}}$ according to the proportions of the degenerate dimension of parameter space, intermediate hub neuron frequency will be preserved in this model.
-This suggests that in the real STG neural circuit, the IC neuron will remain at an intermediate frequency between the pyloric and gastric mill rhythms if parameter changes are made along such
-a dimension.
+This suggests that in the real STG neural circuit, the IC neuron will remain at an intermediate frequency between the pyloric and gastric mill rhythms if parameter changes are made along such a dimension.
 
-I should probably add some text with a big picture takeaway at the end of Section 3.2.
+\begin{displayquote}
+Lines 171 -173 \\
+``This suggests that changes in conductance along the parameter combination described by $\mathbf{v}_2$ will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms."
+\end{displayquote}
 
@@
 We thank the reviewer for their positive comments.  
-Training deep networks for visual processing used to be impossible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.
+To draw a parallel, training deep networks for visual processing used to be considered infeasible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.
 Similarly, training deep networks via EPI to learn parameter distributions became easier throughout this research project as we learned through trial and error what works well.  
-In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge while capturing complex structure much more regularly.
-(*Make empirical statements about regularity in the paper?
-Cite GLOW or Real NVP?) This manuscript has much value in its explanation of hyperparameter choices, and the extensive set of examples in the online code github.
+In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge while capturing complex structure much more regularly (e.g.
+Dinh et al.
+2017 and Kingma et al.
+2018).
+This manuscript has much value in its explanation of hyperparameter choices, and the extensive set of examples in the online code github. 
 Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: augmented Lagrangian constant, normalizing flow architecture.
@@
 In the new manuscript, we present the method as an inference technique without focusing very greatly on maximum entropy, since it tends to distract and confuse the reader.  
-We derive an analytic equivalence to variational inference (Section 5.1.6) showing that a.) EPI is validly conceptualized as an inference method, and b.) to emphasize that maximum entropy is the normative selection principle of Bayesian inference methods in general.
+We derive an analytic equivalence to variational inference (Section 5.1.6) showing that a.) EPI is a valid inference method, and b.) to emphasize that maximum entropy is the normative selection principle of bayesian inference methods in general.
 Thus, the concern of not having the globally optimal inferred distribution is the same that applies to all other inference techniques.
@@
 To handle this methodologically, we run EPI with multiple random seeds, and select the distribution that has converged with the greatest entropy for scientific analysis.
+Throughout the manuscript, we compare to analytic, error contour, and brute-force ground truth to ensure we are capturing the correct distribution with EPI (see response to R3 concern 1).
 
@@
 Indeed, these distributions become more challenging to understand and visualize in high dimensions.  
-This is where the sensitivity measurements appearing in sections 3.1, 3.4, and 3.5 can be extra useful.
+This is where the sensitivity measurements appearing in sections 3.1, 3.4, and 3.5 can be particularly useful.  
 Even in high dimensions, trained deep probability distributions offer tractable quantitative assessments of how parametric combinations affect the emergent property that was conditioned upon.
@@
 \begin{displayquote}
-Section \ref{results_dgm}, Lines X-Y\\
+Section \ref{results_dgm}, Lines 136-140\\
 ``\quoting{stg_ep_text1}~\quoting{stg_ep_text2}."
@@
 \begin{displayquote}
-Section \ref{results_RNN}, Lines X-Y\\
+Section \ref{results_RNN}, Lines 200-207\\
 ``\quoting{rnn_ep_text1}~... \quoting{rnn_ep_text2}~\quoting{rnn_ep_text3}"
@@
 \begin{displayquote}
-Section \ref{results_V1}, Lines X-Y\\
-``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}."
+Section \ref{results_V1}, Lines 281-284\\
+``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}"
 \end{displayquote}
@@
 \begin{displayquote}
-Section \ref{results_SC}, Lines X-Y\\
+Section \ref{results_SC}, Lines 336-339\\
 ``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}"
@@
 
-\begin{itemize}
-\item call out x as internal state and reference.
-\item ref to how we handle entropy
-\item talk about how we don't do this anymore.  ref hypothesis in sc
-\end{itemize}
+Thank you for the detailed suggestions on how to improve writing in the results sections.
+ We have taken the specific suggestion of explicitly calling out $\mathbf{x}$ as network state, and we have moved complicated discussion of the role of entropy to the Methods section.
 
@@
 We thank the reviewer for their positive comments and thoughtful feedback.
-We have made great efforts to provide several clear comparisons to ground truth (see response to concern \#1), and now provide an extensive methodological comparison to SMC-ABC and SNPE (see response to concern \#2).
+We have made great efforts to provide several clear comparisons to ground truth (see response to concern \#1), and now provide an extensive methodological comparison to SMC-ABC and SNPE (see response to Main concern \#2).
 
@@
 The EPI inferred distribution for rapid task switching in the SC model is sharply bent (Fig. \ref{fig:SC}), and matches the parameter set returned from random sampling (Figure \ref{fig:SC8}A).  
-We note that the brute-force parameter set is actually not the ground-truth solution, because it does not obey the constraints of the emergent property as the EPI dsitribution does (Figure \ref{fig:SC8}B).
+We note that the brute-force parameter set is actually not the ground-truth solution, because it does not obey the constraints of the emergent property as the EPI distribution does (Figure \ref{fig:SC8}B).
 This can explain the spurious samples in the brute-force set that are not in the EPI inferred distribution.
@@
 Finally, the underlying maximum entropy flow network (MEFN) algorithm is compared to a ground truth solution (Loaiza Ganem et al.
-2017, Figure 2) by deriving ground truth from the duality of maximum entropy distributions and exponential famillies (see Section \ref{methods_ME_EF}).
+2017, Figure 2) by deriving ground truth from the duality of maximum entropy distributions and exponential families (see Section \ref{methods_ME_EF}).
 
@@
 
-The reviewer is right to point out the importance of a scaling analysis.  Please see the new Section 3.3.
+The reviewer is right to point out the importance of a scaling analysis. Please see response to Main concern \#2.
 
@@
 
-(* Looks like we need to really add a para about limitations.)
+The STG subcircuit model of Gutierrez et al. 2013 is certainly less complex than other models of the STG, yet 5 connected Morris-Lecar neurons is certainly a nontrivial system.
+We clarify why this model is analyzed in Section \ref{results_motivating} instead of more complex STG models when discussing the differences between EPI and SNPE:
+
+\begin{displayquote}
+Lines 240-251\\
+``\quoting{rnn_gradients}"
+\end{displayquote}
 
@@
 
-(* I think we should add a para to discussion about this.)
+Thank you for this recommendation.
+We now include a substantial text in discussion devoted to this topic.
+
+\begin{displayquote}
+Lines 431-454\\
+``\quoting{discussion}"
+\end{displayquote}
 
@@
 
+Thank you for this detailed list of improvements to be made when describing the method.
+We have taken care to address each item in this list.
+
 \textbf{Minor Comments: }
@@
 
-(* Will change).
+We have removed the variable $\omega$ as it was unnecessary, and avoided introducing another variable $f$ to be conflated with $f$'s use as emergent property statistic.
 
@@
 
-Thank you for these references, we have incorporated them where appropriate. (TODO)
+Thank you for these references, we have incorporated some where appropriate.
 

### 2021-04-24 written/eLifeRevision/Bittner_eLife2020.tex
first complete draft with response

@@
 \section{Abstract}
-%When analytic derivation of the relationship between model parameters and computational properties is intractable, approximate inference and simulation-based techniques are relied upon for scientific insight.
-%We bring the use of deep generative models for probabilistic inference to bear on this problem, learning complex distributions of parameters that produce the specified properties of computation.
 A cornerstone of theoretical neuroscience is the circuit model: a system of equations that captures a hypothesized neural mechanism.  
@@
 A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
-To solve challenging inverse problems modeling neural datasets, neuroscientists have used statistical inference techniques to much success. 
-However, when theorizing circuit models, scientists predominantly focus on reproducing established computational properties rather than explaining a given neural dataset well.
+To solve challenging inverse problems concerning neural datasets, neuroscientists have used statistical inference techniques to much success. 
+However, when theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than explaining a given neural dataset well.
 In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
 Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
-This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion (STG).
-Then, with recurrent neural networks of increasing size, we show that only EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
+This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion.
+Then, with recurrent neural networks of increasing size, we show that EPI allows precise control over the behavior of inferred parameters, and that EPI scales better in parameter dimension than alternative techniques.
 In the remainder of this work, we present novel theoretical findings gained through the examination of complex parametric structure captured by EPI.
@@
 
-%-------- old sentences -----------
-
-%Historically, the gold standard has been to analytically derive the relationship between model parameters and computational properties.  
-%However, this enterprise quickly becomes infeasible as biologically realistic constraints are included into the model increasing its complexity, often resulting in \emph{ad hoc} approaches to understanding the relationship between model and computation.
-%Importantly, the techniques we introduce offer a principled means to understand the implications of model parameter choices on computational properties of interest.  
-%We then use it to go beyond linear theory of neuron-type input-responsivity in a model of primary visual cortex, gain a mechanistic understanding of rapid task switching in superior colliculus models, and attribute error to connectivity properties in recurrent neural networks solving a simple mathematical task.
-%More generally, this work suggests a departure from realism vs tractability considerations, towards the use of modern machine learning for sophisticated interrogation of biologically relevant models.
+\section{Introduction}
+\q{intro_first}{The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
+A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
+The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
+The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
+The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.}
+
+Ideally, one carefully designs a model and analytically derives how computational properties determine model parameters.
+Seminal examples of this gold standard include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
+Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
+Absent this analysis, statistical inference offers a toolkit by which to solve the inverse problem by identifying, at least approximately, the distribution of parameters that produce computations in a biologically realistic model \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+
+\q{intro_data_ep}{Statistical inference, of course, requires quantification of the vague term \emph{computation}.  
+In neuroscience, two perspectives are dominant.  
+First, often we directly use an \emph{exemplar dataset}: a collection of samples that express the computation of interest, this data being gathered either experimentally in the lab or from a computer simulation.
+While in some sense the best choice given its connection to experiment \cite{paninski2018neural} , some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
+Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or evidence, rather than the computation itself.
+Alternatively, researchers often quantify an \emph{emergent property}: a statistic of data that directly quantifies the computation of interest, wherein the dataset is only implicit.  
+While such a choice may seem esoteric, it is not: the above ``gold standard" examples all quantify and focus on some derived feature of the data, rather than the data drawn from the model.}
+
+\q{intro_epi}{An emergent property (EP) is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of datasets drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.
+This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.
+However, the adaptation of these techniques to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for scalable, flexible generative modeling \cite{dinh2017density, kingma2018glow}.
+We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.}
+
+Equipped with this method, we prove out EPI by demonstrating its capabilities and presenting novel theoretical findings from its analysis.
+\q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
+Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
+\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals the how recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
+Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
+EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
+We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
+Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
+These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Introduction}
-The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
-A neural circuit is systematized with a set of equations -- the mechanistic model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
-The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.
-
-Consider the idealized practice: one carefully designs a model and analytically derives how computational properties determine model parameters.
-Seminal examples of this gold standard include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
-Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-Still, we can gain insight into these complex models by identifying the distribution of parameters that produce computations.
-By solving the inverse problem in this way, scientific analysis of biologically realistic models is made possible \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
-
-While theoretical neuroscience is concerned with how model parameters govern  computational properties, existing methodology for statistical inference in neuroscience \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}) requires that parameters be conditioned on an explicit dataset.
-The scientific insight for a model of computation is then limited by the quantity and quality of available neural data.
-Even with a vast amount of high-quality recordings, neural data often reflect uninstructed behaviors \cite{niell2010modulation, saleem2013integration, musall2019single}, and thus may only reflect the computation of interest amidst a sea of task-irrelevant factors.
-A common alternative is to synthesize an explicit dataset that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
-In this case, well-defined computational properties are being shoehorned into artificial datasets for the purpose of methodological compatibility.
-
-Another key challenge is that as models of computation become more complex, statistical inference becomes intractable.
-Such mechanistic models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation \cite{dayan2003theoretical, izhikevich2007dynamical}; they lack a tractable likelihood function, which is necessary for statistical inference.
-Therefore, the most popular approaches to parameter inference in mechanistic models have been simulation-based inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
-A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional simulation-based inference approaches.
-However, to use these methods in theoretical neuroscience, we must represent computation with an explicit dataset in some way.
-Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into a framework for observational datasets.
-
-%
-%Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
-%First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
-%Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
-%Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
-%Sampling-based inference approaches (e.g.
-ABC and Markov chain Monte Carlo (MCMC) \cite{hastings1970monte, metropolis1953equation}) confer flexible approximations, yet scale poorly in number of parameters.
-%While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
-%Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
-%Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
-
-%To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
-%This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
-%First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
-%Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
-%Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
-%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-%
-%These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
-%As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets exhibiting some mathematical criteria.
-%EPI constrains the predictions of the inferred parameter distribution to produce the emergent property, which requires a variant of probabilistic inference methods \cite{loaiza2017maximum}.
-%Importantly, EPI uses deep learning to make rich, flexible approximations to the parameter distribution \cite{rezende2015variational} that produces an emergent property.
-%The structure captured by these deep probability distributions are scientifically valuable, revealing the sensitivity and robustness of the emergent property to different parameter combinations.
-%Perhaps most powerfully, EPI facilitates inference in mechanistic models, allowing theorists to capture rich parametric structure in biologically realistic models that is conditioned upon the emergent phenomena of interest.
-
-To address the methodological incongruity between explicit datasets and emergent properties, we present a statistical inference method for conditioning parameters of neural circuit models directly on computation.
-In this work, we define computation by an emergent property, which is a statistical description of the phenomena to be produced by the neural circuit model.
-In emergent property inference (EPI), we infer the distribution of model parameters that produce this emergent property.
-With EPI, parameters are conditioned directly on an implicit dataset defined by the computation of interest.
-By using recent optimization techniques \cite{loaiza2017maximum}, EPI uses deep learning to make rich, flexible approximations to the parameter distributions \cite{rezende2015variational}, the structure of which reveals scientific insight about how parameters govern the emergent property.
-
-Equipped with this method, we prove out the potential of EPI by demonstrating its capabilities and presenting novel theoretical findings borne from its analysis.
-\q{intro_results1}{First, we show EPI's ability to handle mechanistic models using a classic model of parametric degeneracy in biology: the stomatogastric ganglion \cite{goldman2001global, gutierrez2013multiple}.
-Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
-\q{intro_results2}{In a model of primary visual cortex (V1) \cite{litwin2016inhibitory, palmigiano2020structure} with different neuron-types, we show that the equation for excitatory variability become analytically intractable as more populations are added.
-Strikingly, the way in which noisy inputs across neuron-types governs excitatory variability is salient in the visualized structure of the EPI inferred parameter distribution.
-Finally, we investigated the possible connectivities of superior colliculus (SC) that allow execution of different tasks on intervleaved trials \cite{duan2018collicular}.
-EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
-We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
-Intriguingly, all inferred connectivities reproduced results from optogenetic inactivation experiments in this behavioral paradigm -- emergent phenomena that EPI was not conditioned upon.}
-These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.
-
-%-------- old sentences -----------
-
-%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
-
-%This creates an unfavorable tradeoff.
-On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
-%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
-
-%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
-
-% now transition to deep ML...
-%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
-%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
-%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
-
-%However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.
-%First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.
-%Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
-%Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
-%Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model.
-%Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
-
-%First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
-%Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
-%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
-%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
-%Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
-%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
-%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
-
-%First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
-%Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
-%Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Results}
@@
 Consideration of the typical workflow of theoretical modeling clarifies the need for emergent property inference.  
-First, one designs or chooses an existing model that, it is hypothesized, captures the computation of interest. 
+First, one designs or chooses an existing circuit model that, it is hypothesized, captures the computation of interest. 
 To ground this process in a well-known example, consider the stomatogastric ganglion (STG) of crustaceans, a small neural circuit which generates multiple rhythmic muscle activation patterns for digestion \cite{marder2002cellular}.
@@
 \q{stg_subcircuit}{A subcircuit model of the STG \cite{gutierrez2013multiple} is shown schematically in Figure \ref{fig:STG}A.
-The fast population ($f1$ and $f2$) represents the subnetwork generating the pyloric rhythm and the slow population ($s1$ and $s2$) represents the subnetwork of the gastric mill rhythm.
+The fast population (f1 and f2) represents the subnetwork generating the pyloric rhythm and the slow population (s1 and s2) represents the subnetwork of the gastric mill rhythm.
 The two fast neurons mutually inhibit one another, and spike at a greater frequency than the mutually inhibiting slow neurons.
-The hub neuron (hub) couples with either the fast or slow population, or both depending on modulatory conditions.}
+The hub neuron couples with either the fast or slow population, or both depending on modulatory conditions.}
 The jagged connections indicate electrical coupling having electrical conductance $g_{\text{el}}$, smooth connections in the diagram are inhibitory synaptic projections having strength $g_{\text{synA}}$ onto the hub neuron, and $g_{\text{synB}}=5$nS for mutual inhibitory connections.
@@
 Second, once the model is selected, one must specify what the model should produce.
-In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model \ref{fig:STG}B.
+In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
 An emergent property studied by Guttierez et al. of this stochastic model is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
 This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
-Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and and slow (.68Hz) frequencies, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
+Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
 %To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
@@
 %In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-Third, the model parameters producing these outputs are inferred.
-To infer the STG parameters of intermediate hub frequency with existing methodology, we  need an explicit dataset: experimentally recorded or synthesized.
+Third, the model parameters producing the emergent property are inferred.
+To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
 By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
@@
 In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
-%Most often, brute-force parameter sweeps or rejection sampling techniques \cite{liepe2014framework} are used to identify parameters whose model simulations are close to data or some desired feature.
 
@@
 \end{center}
-\caption{\small Emergent property inference (EPI) in the stomatogastric ganglion.  
-\textbf{A}. Conductance-based biophysical model of the STG subcircuit.
+\caption{Emergent property inference (EPI) in the stomatogastric ganglion.  
+\textbf{A}. Conductance-based subcircuit model of the STG.
 \textbf{B}. Spiking frequency $\omega(\mathbf{x}; \mathbf{z})$ is an emergent property statistic. 
@@
 Simulated activity traces are colored by $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ of generating parameters. (Panel E).
-\textbf{D}. For a choice of model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
+\textbf{D}. For a choice of circuit model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
 \textbf{E}. The EPI distribution producing intermediate hub frequency.
@@
 Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Dimension of sensitivity $\mathbf{v}_1$ (solid) and degeneracy $\mathbf{v}_2$ (dashed).
-%(Inset) Sensitivity of the system with respect to network syncing along all dimensions of parameter space away from the mode.  
+Dimension of sensitivity $\mathbf{v}_1$ (solid) and robustness $\mathbf{v}_2$ (dashed).
 \textbf{F} (Top) The predictive distribution of EPI.
@@
 
-%Before presenting technical details (in the following section), let us understand emergent property inference schematically:  EPI (Fig.
-\ref{fig:STG}A gray box) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution shown in Figure \ref{fig:STG}B.
 Before presenting technical details (in the following section), let us understand emergent property inference schematically.
@@
 This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution that produces the emergent property. 
-Scientifically, we can use this parameter distribution to efficiently generate many parameters producing the emergent property or analyze the structure of the distribution which informs how model parameters govern the emergent property.
-%In the STG model, this distribution can be specifically queried to reveal the prototypical parameter configuration for intermediate hub frequency (the mode; Figure \ref{fig:STG}E yellow star), and how it decays based on changes away from the mode.
-%Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$) and degeneracy ($v_2$) (Fig.
-\ref{fig:STG}E, arrows) agree with error contours (Fig.
-\ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
-\ref{fig:STG}F activity traces) (see Section \ref{methods_STG}).
+We can then use this parameter distribution to efficiently generate many parameters producing the emergent property.
+Most importantly, the resulting parameter distribution can be efficiently queried to quantify its own structure, which informs the central question of this inverse problem: how do model parameters govern the emergent property?
 
@@
 Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions.
-%First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
 First, as is typical, we consider the model as a coupled set of differential equations.  
-In this STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
+In this STG example, the model activity (or state) $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
 \begin{equation} 
@@
 \end{equation} 
-where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{morris1981voltage, gutierrez2013multiple} (see Section \ref{methods_stg_model} for more detail).
-
-Second, we stipulate that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
-\q{stg_ep_text1}{We stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$ is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
-Mathematically, we define this emergent property with two statistical constraints: that the mean hub frequency is 0.55Hz,}
-\begin{equation}
- \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55 
+where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on activity $\mathbf{x}$ and parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise \cite{gutierrez2013multiple} (see Section \ref{methods_stg_model} for more detail).
+
+Second, we determine that our model should produce the emergent property of ``intermediate hub frequency" (Figure \ref{fig:STG}C).
+\q{stg_ep_text1}{We stipulate that the hub neuron's spiking frequency -- denoted by statistic $\omega_{\text{hub}}(\mathbf{x})$ -- is close to a frequency of 0.55Hz, between that of the slow and fast frequencies.
+Mathematically, we define this emergent property with two constraints: that the mean hub frequency is 0.55Hz,}
+\begin{equation}
+ \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55\text{Hz}
 \end{equation}
@@
 \begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2.
-\end{equation}
-The hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$  and the distribution of the data $\mathbf{x}$ that those parameters produce.
+\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2 \text{Hz}^2.
+\end{equation}
+In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
 Formally, the emergent property is the collection of these two constraints
 \begin{equation}\label{eq:EP_STG}
-\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] = 0.55,~~\text{Var}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] =0.025^2.
-\end{equation}
-In general, an emergent property is a collection of first-, second- and higher moments of  statistics that together define the phenomena.
+\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] = 0.55\text{Hz},~~\text{Var}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] =0.025^2\text{Hz}^2.
+\end{equation}
+In general, an emergent property is a collection of constraints on first-, second- and higher moments of one or more statistics (see Sections \ref{results_RNN}, \ref{results_SC}, and \ref{methods_2DLDS}) that together define the computation.
 
@@
 This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, dinh2017density, papamakarios2017masked} (Figure \ref{fig:STG}D, EPI box).
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ to a suitably complicated distribution (see Section \ref{methods_NF} for more details).
-Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in variational bayesian methods (see Section \ref{methods_VI}).
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ (drawn from an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
+\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in variational bayesian methods (see Section \ref{methods_VI}).}
 %\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
-In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI})
-The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since it is conditioned upon emergent property $\mathcal{X}$. 
-This is meant to share the same notation as a posterior distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathbf{x})$ that is conditioned upon an explicit dataset.
-
-EPI produces parameter distributions that can be queried for scientific insight.
-The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices exemplary of the emergent property (Fig. \ref{fig:STG}E yellow star).
-As probability in the EPI inferred distribution decreases, the emergent property deteriorates.
-Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ does not change will not disturb the emergent property, making this parameter combination \emph{degenerate} with respect to the emergent property.
-In contrast, if $\mathbf{z}$ is perturbed along a dimension that strongly decreases $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we call that parameter combination \emph{sensitive}.
-By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative the eigenvalue, the more sensitive.
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($v_1$, smaller eigenvalue) and robustness ($v_2$, greater eigenvalue) (Fig.
+In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI}).
+The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since it is conditioned upon the implicit datasets of emergent property $\mathcal{X}$ in a manner methodologically equivalent to exemplar data when using variational bayesian approaches.
+
+\q{results_hess}{EPI produces parameter distributions that can be queried for scientific insight.
+The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices emblematic of the emergent property (Fig. \ref{fig:STG}E yellow star).
+As probability in the EPI distribution decreases, the emergent property deteriorates.
+Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ does not change will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
+In contrast, if $\mathbf{z}$ is perturbed along a dimension that strongly decreases $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive}.
+By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative, the more sensitive and the closer to zero, the more robust (see Section \ref{methods_stg_hessian}).
+Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig.
 \ref{fig:STG}E, arrows) agree with error contours (Fig.
@@
 \ref{fig:STG}F activity traces).
-Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (see Section \ref{methods_stg_hessian}).
+This suggests that changes in conductance along the parameter combination described by $\mathbf{v}_2$ will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
+Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (see Section \ref{methods_NF}).}
 
@@
 \end{center}
-\caption{\small 
+\caption{ 
 \textbf{A}. Wall time of EPI (blue), SNPE (orange), and SMC-ABC (green) to converge on RNN connectivities producing stable amplification.  
@@
  \end{equation}
-where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{u}_1 \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
-We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top$   that produce stable amplification.
+where $U = \begin{bmatrix} \mathbf{U}_1 & \mathbf{U}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{V}_1 & \mathbf{V}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{U}_1 \mathbf{U}_2, \mathbf{V}_1, \mathbf{V}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{U}_1^\top, \mathbf{U}_2^\top, \mathbf{V}_1^\top, \mathbf{V}_2^\top \right]^\top$   that produce stable amplification.
 \q{rnn_ep_text1}{Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
@@
 For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
-We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with observation $\mathbf{x}_0 = \bm{\mu}$.
-SMC-ABC is a rejection sampling approach that SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
+We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with exemplary dataset $\mathbf{x}_0 = \bm{\mu}$.
+SMC-ABC is a rejection sampling approach that uses SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
 Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
@@
 It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
-It is well known that ABC techniques struggle in dimensions greater than about 30 \cite{sisson2018handbook}, yet we were careful to assess the scalability of the more comparable approach SNPE.
+It is well-known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
 Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
@@
 \ref{fig:RNN3}).
-From this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional parameter distributions than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both criteria in high dimensions.
+In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time and simulation count.
 
@@
 \ref{fig:RNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from the exemplary dataset $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
 When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:RNN}C, orange).
@@
 This agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}.
-\q{rnn_gradients}{Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractabillity is a key criteria when determining the suitability of a simulation-based inference technique.
+\q{rnn_gradients}{Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
 Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
-This can be invaluable for understanding how RNNs produce complex emergent phenomena.
+This can be invaluable for understanding how RNNs produce complex computations.
 Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
@@
 burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, approaches that do not require differentiability of the emergent property like SNPE have proved to be a powerful approach \cite{gonccalves2019training}.}
+In such cases, SNPE, which does not require differentiability of the emergent property has proved to be a powerful approach \cite{gonccalves2019training}.}
 In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{EPI reveals how noisy input across neuron-types governs excitatory variability in a V1 model} \label{results_V1}
+\subsection{EPI reveals how recurrence with multiple inhibitory subtypes governs excitatory variability in a V1 model} \label{results_V1}
 
@@
 \vspace{-1cm}
-\caption{\small Emergent property inference in the stochastic stabilized supralinear network (SSSN)
+\caption{Emergent property inference in the stochastic stabilized supralinear network (SSSN)
 \textbf{A}.
@@
 \ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
-The strong interactive influence of E- and P-population input variability on excitatory varability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
+The strong interactive influence of E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
 \ref{fig:V1_3}), meaning that the E-population is P-stabilized.
@@
 This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
-While EPI can be used to investigate fundamental aspects of sensory processing, in the next section, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to  behavioral experiments.
+While we have just shown that EPI can be used to investigate fundamental aspects of sensory computation, in the next two sections, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to behavioral experiments.
 
@@
 \subsection{EPI identifies two regimes of rapid task switching} \label{results_SC}
-It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity allow this ability.
-In this experimental setup, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig.
+It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity produce this computation.
+In this experimental setup, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from the stimulus in the Anti (A) task (Fig.
 \ref{fig:SC}A).
@@
 \end{center}
-\caption{\footnotesize 
+\caption{ 
 \textbf{A}. Rapid task switching behavioral paradigm (see text). 
-\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
+\textbf{B}. Model of superior colliculus (SC). Neurons: LP - Left Pro, RP - Right Pro, LA - Left Anti, RA - Right Anti. 
 Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
 \textbf{C}. The EPI inferred distribution of rapid task switching networks.  
-Red and purple stars indicate modes $\mathbf{z}^*$ of each connectivity regime.
+Stars indicate modes $\mathbf{z}^*$ whose color indicates connectivity regime (see legend Fig \ref{fig:SC}D).
 Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
@@
 The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, middle-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
-This parameter distribution has intricate structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
-Specifically, the shape of the EPI distribution changes dramatically on different sides of parameter space.
+This parameter distribution has rich structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
+Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8})
 This is most saliently pointed out in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
@@
 
-While the connectivities gradually coalesce to the transitionary part of parameter space, the sensitivity dimensions $\mathbf{v}_1(\mathbf{z})$ are categorically different across regimes (Fig.
+Although the connectivities gradually coalesce to the transitionary part of parameter space, the sensitivity dimensions $\mathbf{v}_1(\mathbf{z})$ are categorically different across regimes (Fig.
 \ref{fig:SC4} bottom).
@@
 For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_sc_eig}).
-Perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has close to the same effect on Anti accuracy as perturbations along the sensitivtity dimension (Fig. \ref{fig:SC}D, middle).
-This suggests that there is a carefully tuned strength of task representation in regime 1, which if disturbed results in random Anti trial responses.
+Perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has a very similar effect on Anti accuracy as perturbations along the sensitivity dimension (Fig. \ref{fig:SC}D, middle).
+This suggests that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
 Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, bottom).
@@
 
-During the delay period of this task, the circuit must prepare to execute the correct task based on the cue input.
+During the delay period of this task, the circuit must prepare to execute the correct task according to the presented cue.
 Experimental results from Duan et al.
-found that optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
+found that bilateral optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
 \ref{fig:SC_opto}A).
 This suggests that SC maintains a representation of task throughout the delay period, which is important for correct execution of the Anti task.
-Network connectivities inferred by EPI exhibited this same effect in simulation at high  optogenetic strength $\gamma \in [0, 1]$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
-To emphasize, EPI inferred connectivities were only conditioned upon the emergent property of rapid task switching, not on Anti task failure during delay period silencing.
+Network connectivities inferred by EPI exhibited this same effect in simulation at high optogenetic strengths $\gamma$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
 
@@
 \end{center}
-\caption{\footnotesize 
+\caption{ 
 \textbf{A}. Experimental effect of delay period silencing on Pro and Anti task accuracy in rats.
-\textbf{B}. Mean and standard error (bars) across EPI distribution of Pro and Anti error induced by delay period inactivation of varying optogenetic strength.
+\textbf{B}. Mean and standard error (bars) of task error induced by delay period inactivation of varying optogenetic strength across the EPI distribution.
 \textbf{C}. (Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.675$.
@@
 \textbf{D}.
-(Left) Responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.85$.
+(Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.85$.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.85$ versus accuracy in the Pro task across connectivities in the EPI distribution.
@@
 At this level of optogenetic strength, only regime 1 exhibits an increase in Anti error with delay period silencing (Fig. \ref{fig:SC_opto}C, left).
-The connectivities in regime 2 are thus more resilient in the Anti task to delay period silencing than regime 1.
+The connectivities in regime 2 are thus more resilient to delay period silencing during Anti trials than regime 1.
 In regime 1, greater $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ decrease Anti error (Fig. \ref{fig:SC_opto}C, right).
@@
 Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC_opto}D, left).
-We interpreted these similarities to suggest that delay period inactivation at this saturated level flips the internal representation of task (from Anti to Pro) in the model.
+We interpreted these similarities to suggest that delay period inactivation at this saturated level flips the internal representation of task (from Anti to Pro) in the circuit model.
 This would explain why the Anti error saturates at 50\%: the average Anti accuracy in EPI inferred connectivities is $75\%$, but is 25\% when the internal representation is flipped during delay period silencing.
@@
 
-%To establish these two regimes of connectivity, we took gradient steps along $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to produce modes $\mathbf{z}_1$ and $\mathbf{z}_2$  (Fig.
-\ref{fig:SC}C red and purple stars, Section \ref{methods_SC}).
-%Simulations from these two regimes reveal different responses in each task (Fig. \ref{fig:SC}D).
-%We charcaterized these regimes by identifying the dimensions of connectivity that rapid task switching is most sensitive to.
-%The sensitivity dimensions $\mathbf{v}_1$ and  $\mathbf{v}_2$ (Fig.
-\ref{fig:SC}C, red and purple arrows) point in different directions, resulting in different changes to task accuracy (Fig.
-\ref{fig:SC}D, bottom-left, Fig.
-\ref{fig:SC2}).
-%In regime 1, Anti accuracy diminishes in either direction of sensitivity away from the mode, while in regime 2, Anti accuracy matches monotonic increases in Pro accuracy.
-%
-%To understand why these distinct connectivity regimes have different failure modes, we can analyze the properties of connectivity in each regime.
-%By taking the eigendecomposition of $W$, we can quantify how strongly different modes of processing are reflected in the connectivity matrix (see Section \ref{methods_SC}).
-
 %\subsection{Novel hypotheses from EPI} 
@@
 However, these statistical models often lack resemblance to the underlying biology, making it unclear how to go from the structure revealed by these methods, to the neural mechanisms giving rise to it.
-In contrast, theoretical neuroscience has focused on careful mechanistic modeling and the production of emergent properties of computation, rather than measuring structure in some noisy observed dataset.
+In contrast, theoretical neuroscience has primarily focused on careful models of neural circuits and the production of emergent properties of computation, rather than measuring structure in neural datasets.
 %The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to emergent phenomena in  interpretable models of neural computation (see Section \ref{methods_related}).
-
-Methodology for statistical inference in mechanistic models of neural circuits has evolved considerably in recent years.
-Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but more recently developed methodology employs deep learning to improve efficiency or provide flexible distribution approximations.
-SNPE \cite{gonccalves2019training} and other sequential techniques for inference in mechanistic models developed along with EPI (see Section \ref{methods_related}) have been used for posterior inference with noisy experimental datasets.
-On the other hand, EPI is a deep inference technique designed to condition directly on emergent properties, such that the parameter distribution only produces the computation of interest.
-EPI is thus ideally suited for questions in theoretical neuroscience, and we show that it has superior scaling properties to these other inference techniques (see Section \ref{results_RNN}).
+In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to computation in interpretable models of neural circuits (see Section \ref{methods_related}).
+
+\q{discussion}{Methodology for statistical inference in circuit models has evolved considerably in recent years.
+Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but more recently developed methodology employs deep learning to improve efficiency or provide deep, flexible distribution approximations.
+SNPE \cite{gonccalves2019training}, developed along with EPI (see Section \ref{methods_related}), has been used for posterior inference of parameters in circuit models conditioned upon exemplar data used to represent computation.
+Like SNPE, EPI is a deep inference technique, but it infers parameter distributions that only produce the computation of interest  (see Section \ref{results_RNN}).
+
+Exemplary data versus emergent properties aside, EPI has better scaling properties than SNPE when emergent property gradients are tractable (Section \ref{results_RNN}).  
+However, SNPE has its own relative advantages.
+SNPE is effective when circuit model simulations are lengthy or nondifferentiable.
+For example, SNPE can infer the STG parameters that produce the pyloric rhythm \cite{gonccalves2019training}, while EPI cannot.
+Thus, while it is nice to infer parameter distributions with constrained emergent properties with EPI,  SNPE is most appropriate when emergent property gradients are intractable.
+
+The scientific analyses of Sections \ref{results_V1} and \ref{results_SC} derived theoretical findings by querying the structure of the inferred distribution of EPI.
+By measuring the dimensions in which probability decreases fastest, the dimensions of \emph{sensitivity}, we gain valuable understanding of how model parameters govern the emergent property.
+A rich literature on parameter sensitivity analyses in biological models presents several methods towards this scientific approach \cite{saccomani2003parameter, raue2009structural, karlsson2012efficient, raman2017delineating}.
+The value offered by EPI (and other deep inference methods like SNPE), is that the once the flexible deep probability distributions are fit to the parameter distribution, this distribution's structure can be quantified at any parameter choice, offering instantly available sensitivity measurements.
+Together, the ability to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a powerful new method for addressing inverse problems in theoretical neuroscience.}
 
@@
 %We see in Section \ref{results_RNN} that EPI scales seamlessly to high dimensional parameter spaces of RNN connectivities, while maintaining the production of the specified emergent property.
-%EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of it's ability to leverage the gradient information of the emergent property statistics and to adapt it's paramter sampling distribution at every step of gradient descent.
+%EPI strongly outperforms the standard likelihood-free inference technique (SMC-ABC \cite{sisson2007sequential}), and a recently developed deep likelihood-free inference technique (SNPE \cite{gonccalves2019training}), most likely because of its ability to leverage the gradient information of the emergent property statistics and to adapt its parameter sampling distribution at every step of gradient descent.
 %
@@
 %For example, even in low dimensional circuits, models can support more than one steady state \cite{kraynyukova2018stabilized} and non-trivial dynamics on strange attractors \cite{morrison2016diversity}.
-%With EPI, we use deep probabillity distributions to capture the complex nonlinear parameter distributions governing model behavior.
-%In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing curcuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
+%With EPI, we use deep probability distributions to capture the complex nonlinear parameter distributions governing model behavior.
+%In Section \ref{results_V1}, we used EPI to reveal a curved parametric manifolds governing circuit variability in the stochastic stabilized supralinear network, and used hypothesis testing techniques to validate our findings.
 %In Section \ref{results_SC}, we identified two regimes of SC connectivity resulting in rapid task switching, and found that the full distribution of rapid task switching networks reproduced an experimental result.
@@
 \subsection{Emergent property inference (EPI)}\label{methods_EPI}
-%Consider model parameterization $\mathbf{z}$ and data $\mathbf{x}$ which has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$ defined by a model simulator of which samples are available $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$.
-%EPI optimizes a distribution $q_{\bm{\theta}}(\mathbf{z})$ (itself parameterized by $\bm{\theta}$) of model parameters $\mathbf{z}$ to produce an emergent property of interest $\mathcal{X}$ defined by the means and variances of emergent property statistics $f(\mathbf{x}; \mathbf{z})$
-% \begin{equation}
-%\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-%\end{equation}
-%Precisely, the emergent property statistics $f(\mathbf{x})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters $q_{\bm{\theta}}(\mathbf{z})$ and stochasticity of the data given the parameters defined by the model $p(\mathbf{x} \mid \mathbf{z})$.
-%This is a viable way to represent emergent properties in theoretical models, as we have demonstrated in the main text, and enables the EPI optimization.
-%
-%
-%With EPI, we use deep probability distributions to learn flexible approximations to model parameter distributions $q_{\bm{\theta}}(\mathbf{z})$.
-% In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
-$g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
-%\begin{equation}
-%\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0) = g_l(..g_1(\mathbf{z}_0)) \sim q_{\bm{\theta}}(\mathbf{z}).
-%\end{equation}
-%Given a simulator defined by a theoretical model $\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$ and some emergent property of interest $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property:
-%\begin{equation} \label{eq:opt}
-%\begin{split}
-%q_{\bm{\theta}}^*(\mathbf{z}) &= \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2. \\
-% \end{split}
-%\end{equation} 
-%Since we are optimizing parameters $\bm{\theta}$ of our deep probability distribution with respect to the entropy $H(q_{\bm{\theta}}(\mathbf{z}))$, we must take gradients with respect to the log probability density of samples from the deep probability distribution.
-%Entropy of $q_{\bm{\theta}}(\mathbf{z})$ can be expressed as an expectation of the negative log density of parameter samples $\mathbf{z}$ over the randomness in the parameterless initial distribution $q_0$:
-%\begin{equation}
-%H(q_{\bm{\theta}}(\mathbf{z})) = \int - q_{\bm{\theta}}(\mathbf{z}) \log(q_{\bm{\theta}}(\mathbf{z})) d\mathbf{z} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[-\log(q_{\bm{\theta}}(\mathbf{z})) \right] = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[-\log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-%\end{equation}
-%Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients of the log density of samples $\mathbf{z}$:
-%\begin{equation}
-%\nabla_{\bm{\theta}} H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}_0 \sim q_0}\left[- \nabla_{\bm{\theta}} \log(q_{\bm{\theta}}(g_{\bm{\theta}}(\mathbf{z}_0))) \right].
-%\end{equation}
-%In EPI, MEFNs are purposed towards variational learning of model parameter distributions.
-
-Determining the combinations of model parameters that can produce observed data or a desired output is a key part of scientific practice.
-Solving inverse problems is especially important in neuroscience, since we require complex models to describe the complex phenomena of neural computations.
-While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models conditioned upon the emergent phenomena they produce.
-Here, we introduce a novel method for statistical inference, which finds distributions of parameter solutions that only produce the desired emergent property.
+Determining the combinations of model parameters that can produce a desired output is a key part of scientific practice.
+Solving inverse problems is especially important in neuroscience, since we require detailed circuit models to produce computation of varying levels of complexity.
+While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models conditioned upon the emergent properties of computation.
+Here, we introduce a novel method for statistical inference, which finds distributions of parameter solutions that are constrained to produce the desired emergent property.
 This method seamlessly handles neural circuit models with stochastic nonlinear dynamical generative processes, which are predominant in theoretical neuroscience.
 
-Consider model parameterization $\mathbf{z}$, which is a collection of scientifically interesting variables that govern the complex simulation of data $\mathbf{x}$.
-For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
+Consider model parameterization $\mathbf{z}$, which is a collection of scientifically meaningful variables that govern the complex simulation of data $\mathbf{x}$.
+For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials (the state) of the five neurons.
 In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
-Even so, we do not scientifically reason about how $\mathbf{z}$ governs all of $\mathbf{x}$, but rather specific phenomena that are a function of the data $f(\mathbf{x}; \mathbf{z})$.
-In the STG example, $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by $\mathbf{z}$.
-With EPI, we learn distributions of $\mathbf{z}$ that results in an average and variance of $f(\mathbf{x}; \mathbf{z})$, denoted $\bm{\mu}$ and $\bm{\sigma}^2$.
-We refer to the collection of these statistical moments as an emergent property.
-Such emergent properties $\mathcal{X}$ are defined through choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), $\bm{\mu}$, and $\bm{\sigma}^2$
- \begin{equation}
+From a theoretical perspective, we are less concerned about the likelihood of an exemplary dataset $\mathbf{x}$, but rather the emergent property of intermediate hub frequency (which implies a consistent dataset $\mathbf{x}$).
+
+In the STG example, the statistic $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by parameters $\mathbf{z}$.
+With EPI, we learn distributions of $\mathbf{z}$ constrained to produce intermediate hub frequency: to obey the constraints placed on the mean and variance of $f(\mathbf{x}; \mathbf{z})$.
+In general, an emergent property $\mathcal{X}$ is defined through the choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), and its means $\bm{\mu}$, and variances $\bm{\sigma}^2$:
+ \begin{equation} \label{eq:methods_ep_general}
 \mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
 \end{equation}
-Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and stochasticity of the data given the parameters.
-By defining these means and variances over both levels of stochasticity -- the inferred distribution and that of the model -- there is a fine degree of control over predictions made by the inferred parameters.
+Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and the data produced by those parameters.
+Technically, an emergent property may be a combination of first-, second-, or higher-order moments, but this study focuses on the case written in Equation \ref{eq:methods_ep_general}.
 
 In EPI, deep probability distributions are optimized to learn the inferred distribution.
- In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
+ In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ (we choose an isotropic gaussian) is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
 $g_l$) parameterized by weights and biases $\bm{\theta}$ to the support of the distribution of interest:
@@
 Such deep probability distributions embed the inferred distribution in a deep network.
-Once optimized, this deep network representation has remarkably useful properties: fast sampling, probability evaluations, and also first- and second-order probability gradient evaluations.
-
-By choosing a neural circuit model, often represented as a system of differential equations, we implictly definge a model likelihood $p(\mathbf{x} \mid \mathbf{z})$, which may be unknown or intractable for our purposes.
-Given this model choice and that of an emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
+Once optimized, this deep network representation has remarkably useful properties: fast sampling and probability evaluations
+Importantly, fast probability evaluations confer fast gradient and Hessian calculations as well.
+
+Given this choice of circuit model and emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
 \begin{equation} \label{eq:opt}
@@
 Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
-Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain how EPI can be viewed as a form of variational inference (Section \ref{methods_VI}).
+Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain how EPI is a form of variational inference (Section \ref{methods_VI}).
 
  \subsubsection{Related approaches}\label{methods_related}
- When Bayesian inference problems lack conjugacy, scientists use approximate inference methods like variational inference (VI) \cite{saul1998mean} and Markov chain Monte Carlo (MCMC) \cite{metropolis1953equation, hastings1970monte}.
+ When bayesian inference problems lack conjugacy, scientists use approximate inference methods like variational inference (VI) \cite{saul1998mean} and Markov chain Monte Carlo (MCMC) \cite{metropolis1953equation, hastings1970monte}.
 After optimization, variational methods return a parameterized posterior distribution, which we can analyze.
@@
 Simulation-based inference \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
-The most prevalent approach to simulation-based inference is approximate Bayesian computation \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
+The most prevalent approach to simulation-based inference is approximate Bayesian computation (ABC) \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
 The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
@@
 
-For scientific model analysis, we seek a parameter distribution exhibiting the properties of a well-chosen variational approximation: a parametric form conferring analytic calculations, and trivial sampling time.
-For this reason, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples and have unchanging sampling rate.
-EPI infers parameters in mechanistic models using the MEFN \cite{loaiza2017maximum} algorithm using a deep variational approximation.
-The deep neural network of EPI defines the parametric form of the distribution approximation.
-Furthermore, the EPI distribution is constrained to produce an emergent property.
-In other words, the summary statistics of the posterior predictive distribution are fixed to have certain first and second moments.
+For scientific model analysis, we seek a parameter distribution represented by an approximating distribution as in variational inference \cite{saul1998mean} : a variational approximation that once optimized yields fast analytic calculations and samples.
+For the reasons described above, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples lacking probabilities and have unchanging sampling rate.
+EPI infers parameters in circuit models using the MEFN \cite{loaiza2017maximum} algorithm with a deep variational approximation.
+The deep neural network of EPI (Fig. \ref{fig:STG}E) defines the parametric form (with variational parameters $\bm{\theta}$) of the deep variational approximation of circuit parameters $\mathbf{z}$.
+
+Since EPI is not conditioning upon exemplary data as in variational bayesian methodology, EPI is not doing established variational inference.
+In contrast, the EPI distribution is constrained to produce an emergent property.
 EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
-The analytic relationship between EPI and variational inference is explained in Secton \ref{methods_VI}.
+The analytic relationship between EPI and variational inference is explained in Section \ref{methods_VI}.
 
@@
 These methods are optimized on a conditional density (or ratio) estimation objective.
-The data used to optimize this objective are generated via an adaptive procedure, in which training data pairs ($\mathbf{x}_i$, $\mathbf{z}_i$) become sequentiall closer to the true data and posterior.
+The data used to optimize this objective are generated via an adaptive procedure, in which training data pairs ($\mathbf{x}_i$, $\mathbf{z}_i$) become sequentially closer to the true data and posterior.
 
@@
 With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these model gradients when they are tractable.
-Sequential simulation-based inference techniques may be better suited for sceintific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable: when there is a nondifferentiable model or it requires lengthy simulations.
+Sequential simulation-based inference techniques may be better suited for scientific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable, like when there is a nondifferentiable model or it requires lengthy simulations.
 However, the sequential simulation-based inference techniques cannot constrain the predictions of the inferred distribution in the manner of EPI.
 
-Structural identifiability analysis involves the measurement of sensitivity and unidentifiabilities in natural models.
-Around a point, one can measure the Jacobian. One approach that scales well is EAR \cite{karlsson2012efficient}.
+\q{structural_id}{Structural identifiability analysis involves the measurement of sensitivity and unidentifiabilities in scientific models.
+Around a single parameter choice, one can measure the Jacobian. 
+One approach for this calculation that scales well is EAR \cite{karlsson2012efficient}.
 A popular efficient approach for systems of ODEs has been neural ODE adjoint \cite{chen2018neural} and its stochastic adaptation \cite{li2020scalable}.
-Casting identifiability as a statistical estimation problem, the profile likelihood can assess via iterated optimization while holding parameters fixed \cite{raue2009structural}.
+Casting identifiability as a statistical estimation problem, the profile likelihood works via iterated optimization while holding parameters fixed \cite{raue2009structural}.
 An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point by following degenerate dimensions of the fisher information matrix \cite{raman2017delineating}.
@@
 With EPI, we have all the benefits given by a statistical inference method plus the ability to query the first- or second-order gradient of the probability of the inferred distribution at any chosen parameter value.
-The second-order gradient of the log probability (the Hessian), which is directly afforded by EPI distributions, produces salient information about parametric sensitivty of the emergent property.
-For example, the eigenvector with most negative eigenvalue of the Hessian shows parametric combinations away from a parameter choice that decrease the in EPI distribution probability the fastest.
-We refer to this eigenvector as the sensitivity dimension, and it is used to generate scientific insight about a model of superior colliculus connectivity (see Section \ref{results_SC}).
+The second-order gradient of the log probability (the Hessian), which is directly afforded by EPI distributions, produces quantified information about parametric sensitivity of the emergent property in parameter space (see Section \ref{results_dgm}).}
 
@@
 
+\subsubsection{Maximum entropy distributions and exponential families}\label{methods_ME_EF}
+EPI is a maximum entropy distribution, which have fundamental links to exponential family distributions. 
+A maximum entropy distribution of form:
+\begin{equation} \label{eq:max_ent}
+\begin{split}
+p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
+ &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
+ \end{split}
+\end{equation} 
+will have probability density in the exponential family:
+\begin{equation}
+p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
+\end{equation}
+The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify except in special cases \cite{wainwright2008graphical}.
+
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}.
+The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
+ \begin{equation}
+\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
+\end{equation}
+As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
+\begin{equation} \label{eq:moments}
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
+\end{equation}
+which are constrained to the chosen means and variances
+\begin{equation} \label{eq:mu_opt}
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
+\end{equation}
+
 \subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
@@
 \end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
+To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$s), the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ were chosen to contain the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
  $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
-\begin{equation}
- \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
+To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean 1Hz with variance (0.1Hz)$^2$:
+\begin{equation}
+ \mathbb{E}\left[T(\mathbf{x}) \right]_{\mathbf{z}, \mathbf{x}} ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ (\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z})-0)^2  \\ (\text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z})-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}_{\text{opt}}.
  \end{equation} 
@@
 \end{center}
-\begin{flushleft}
 \caption{\small (LDS1): \textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
@@
 At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
-\end{flushleft}
 \label{fig:LDS1}
@@
 \end{center}
-\begin{flushleft}
 \caption{\small (LDS2): \textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
 \textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\end{flushleft}
 \label{fig:LDS2}
@@
 \end{equation}
-Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$ (with $\omega=1$), we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
+Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$, we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
 S2B).
@@
 \end{figure}
-
-\subsubsection{Maximum entropy distributions and exponential families}\label{methods_ME_EF}
-%As we compare EPI to variational inference, it is important to consider that EPI is a maximum entropy method, and that maximum entropy methods have a fundamental relationship with exponential family distributions.
-%A maximum entropy distribution of form:
-%\begin{equation} \label{eq:max_ent}
-%\begin{split}
-%p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
-% &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}. \\
-% \end{split}
-%\end{equation} 
-%will have probability density in the exponential family:
-%\begin{equation}
-%p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-%\end{equation}
-%The mappings between the mean parameterization $\bm{\mu}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify \cite{wainwright2008graphical}.
-EPI is a maximum entropy distribution, which have fundamental links to exponential family distributions. 
-A maximum entropy distribution of form:
-\begin{equation} \label{eq:max_ent}
-\begin{split}
-p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
- &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
- \end{split}
-\end{equation} 
-will have probability density in the exponential family:
-\begin{equation}
-p^*(\mathbf{z}) \propto \exp(\bm{\eta}^\top T(\textbf{z})).
-\end{equation}
-The mappings between the mean parameterization $\bm{\mu}_{\text{opt}}$ and the natural parameterization $\bm{\eta}$ are formally hard to identify except in special cases \cite{wainwright2008graphical}.
-
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}.
-The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
- \begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
-\end{equation}
-As a general maximum entropy distribution (Equation \ref{eq:max_ent}), the sufficient statistics vector contains both first and second order moments of $f(\mathbf{x}; \mathbf{z})$
-\begin{equation} \label{eq:moments}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
-\end{equation}
-which are constrained to the chosen means and variances
-\begin{equation} \label{eq:mu_opt}
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} \bm{\mu} \\ \bm{\sigma}^2 \end{bmatrix}.
-\end{equation}
 
@@
 where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
-Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
-The mean and variances of the mean-field gaussian likelihood are predicated by $\bm{\eta}_{\text{opt}}$ (Equations \ref{eq:eta_f} and \ref{eq:VI_obj}), which is adapted after each EPI optimization epoch based on $\mathcal{X}$ (see Section \ref{methods_AL_opt}).
+\q{ME_prior}{Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
+The mean and variances of the mean-field gaussian likelihood are predicated by $\bm{\eta}_{\text{opt}}$ (Equations \ref{eq:eta_f} and \ref{eq:VI_obj}), which is adapted after each EPI optimization epoch based on $\mathcal{X}$ (see Section \ref{methods_AL_opt}).}
 In EPI, the inferred distribution is not conditioned on a finite dataset as in variational inference, but rather the emergent property $\mathcal{X}$ dictates the likelihood parameterization such that the inferred distribution will produce the emergent property.
@@
 We examined the scaling properties of EPI by learning connectivities of RNNs of increasing size that exhibit stable amplification.
-Rank-2 RNN connectivity was modeled as $W = UV^\top$, where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix} + g\chi^{(V)}$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+Rank-2 RNN connectivity was modeled as $W = UV^\top$, where $U = \begin{bmatrix} \mathbf{U}_1 & \mathbf{U}_2 \end{bmatrix} + g \chi^{(W)}$, $V = \begin{bmatrix} \mathbf{V}_1 & \mathbf{V}_2 \end{bmatrix} + g\chi^{(V)}$, and $\chi^{(W)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
 This RNN model has dynamics
@@
 \end{equation}
-In this analysis, we inferred connectivity parameterizations $\mathbf{z} = \left[\mathbf{u}_1^\top, \mathbf{u}_2^\top, \mathbf{v}_1^\top, \mathbf{v}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produced stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section Related Methods).
+In this analysis, we inferred connectivity parameterizations $\mathbf{z} = \left[\mathbf{U}_1^\top, \mathbf{U}_2^\top, \mathbf{V}_1^\top, \mathbf{V}_2^\top \right]^\top \in \left[-1, 1 \right]^{(4N)}$   that produced stable amplification using EPI, SMC-ABC \cite{sisson2007sequential}, and SNPE \cite{gonccalves2019training} (see Section Related Methods).
 
@@
 \end{equation}
-so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the abscence of recurrent connectivity ($W = \bm{0}$).
+so that $\bm{\sigma}$ parameterizes the variance of the noisy input in the absence of recurrent connectivity ($W = \bm{0}$).
 As contrast $c \in [0, 1]$ increases, input to the E- and P-populations increases  relative to a baseline input $\mathbf{h} = \mathbf{h}_b + c\mathbf{h}_c$.
@@
 Which is another Lyapunov Equation, now in 4 dimensions.
-In the simplest case in which $\tau_{\text{noise}}=\tau$, the volatge is directly driven by white noise, and $\Lambda_v$ can be expressed in powers of $S$ and $S^T$.
+In the simplest case in which $\tau_{\text{noise}}=\tau$, the voltage is directly driven by white noise, and $\Lambda_v$ can be expressed in powers of $S$ and $S^T$.
 Because $S$ satisfies its own polynomial equation (Cayley Hamilton theorem), there will be 4 coefficients for the expansion of $S$ and 4 for $S^T$, resulting in 16 coefficients that define $\Lambda_v$ for a given $S$.
@@
 \caption{\small (SC6): 
-(Left) Responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
+(Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.
@@
 \begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \\ \left(p_P(\mathbf{x}; \mathbf{z}) - 75\% \right)^2 \\ \left(p_A(\mathbf{x}; \mathbf{z}) - 75\% \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 75\% \\ 75\% \\ 7.5\%^2 \\ 7.5\%^2 \end{bmatrix}.
+T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \\ \left(p_P(\mathbf{x}; \mathbf{z}) - .75 \right)^2 \\ \left(p_A(\mathbf{x}; \mathbf{z}) - .75 \right)^2 \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu}_{\text{opt}} = \begin{bmatrix} .75 \\ .75 \\ .075^2 \\ .075^2 \end{bmatrix}.
 \end{equation}
@@
 This set of connectivities will not match the distribution of EPI exactly, since it is not conditioned on the emergent property.
-For example the parameter set returned by the brute-force search is biased towards lower accuraciess (Fig. \ref{fig:SC8}B).
+For example the parameter set returned by the brute-force search is biased towards lower accuracies (Fig. \ref{fig:SC8}B).
 
@@
 We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.
-During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma$
+During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma \in [0, 1]$
 \begin{equation}
@@
 \end{document}
+
+%-------- old intro sentences -----------
+
+%While theoretical neuroscience is concerned with how model parameters govern  computational properties, existing methodology for statistical inference in neuroscience \cite{kass2001spike, brown1998statistical, paninski2004maximum, truccolo2005point, schneidman2006weak, druckmann2007novel, turner2007maximum, byron2009gaussian, macke2011empirical, park2011bayesian, granot2013stimulus, latimer2015single, lakshminarasimhan2018dynamic, duncker2019learning, ladenbauer2019inferring, gao2016linear, zhao2017recursive, barello2018sparse, pandarinath2018inferring, wiltschko2015mapping, johnson2016composing, batty2019behavenet} (see review, \cite{paninski2018neural}) requires that parameters be conditioned on an explicit dataset.
+%The scientific insight for a model of computation is then limited by the quantity and quality of available neural data.
+%Even with a vast amount of high-quality recordings, neural data often reflect uninstructed behaviors \cite{niell2010modulation, saleem2013integration, musall2019single}, and thus may only reflect the computation of interest amidst a sea of task-irrelevant factors.
+%A common alternative is to synthesize an explicit dataset that is exemplary of that computation, so that the framework of statistical inference can be applied for parameter identification.
+%In this case, well-defined computational properties are being shoehorned into artificial datasets for the purpose of methodological compatibility.
+%
+%Another key challenge is that as models of computation become more complex, statistical inference becomes intractable.
+%Such mechanistic models in theoretical neuroscience are noisy systems of differential equations that can only be sampled or realized through forward simulation \cite{dayan2003theoretical, izhikevich2007dynamical}; they lack a tractable likelihood function, which is necessary for statistical inference.
+%Therefore, the most popular approaches to parameter inference in mechanistic models have been simulation-based inference methods \cite{sisson2007sequential, liepe2014framework}, in which reasonable parameters are obtained via simulation and rejection.
+%A new class of techniques \cite{gonccalves2019training, papamakarios2019sequential, hermans2020likelihood} use deep learning to improve upon traditional simulation-based inference approaches.
+%However, to use these methods in theoretical neuroscience, we must represent computation with an explicit dataset in some way.
+%Theorists are therefore barred from using the probabilistic modeling toolkit for science with circuit models, unless they reformulate their inverse problem into a framework for observational datasets.
+
+%
+%Existing approaches to the inverse problem in theoretical neuroscience fall short in three key ways. 
+%First, theoretical models of neural computation aim to reflect a complex biological reality, and as a result, such models lack tractable likelihoods.
+%Without an efficient calculation of the probability of model properties given model parameters, neuroscientists resort to approximate Bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, which requires a rejection heuristic, scales poorly, and only produces sets of accepted parameters lacking probabilities.
+%Second, there is an undesirable trade-off between the flexibility and sampling speed of approximated posterior distributions.
+%Sampling-based inference approaches (e.g.
+ABC and Markov chain Monte Carlo (MCMC) \cite{hastings1970monte, metropolis1953equation}) confer flexible approximations, yet scale poorly in number of parameters.
+%While variational inference (VI) \cite{saul1998mean} often results in fast posterior sampling, existing practice relies heavily on simplified classes of distributions \cite{rezende2015variational}.  
+%Third, such parameter inference methods are designed to operate on experimentally collected data-sets.
+%Ultimately, the objects of interest in theoretical neuroscience are phenomena or features of the model rather than singular data-sets.
+%
+%To address the methodological incongruity between explicit datasets and emergent properties, we present a statistical inference method for conditioning parameters of neural circuit models directly on computation.
+%In this work, we define computation by an emergent property, which is a statistical description of the phenomena to be produced by the neural circuit model.
+%In emergent property inference (EPI), we infer the distribution of model parameters that produce this emergent property.
+%With EPI, parameters are conditioned directly on an implicit dataset defined by the computation of interest.
+%By using recent optimization techniques \cite{loaiza2017maximum}, EPI uses deep learning to make rich, flexible approximations to the parameter distributions \cite{rezende2015variational}, the structure of which reveals scientific insight about how parameters govern the emergent property.
+
+%To address these three challenges, we developed an inference methodology -- `emergent property inference' -- which learns a distribution over parameter configurations in a theoretical model.  
+%This distribution has two critical properties: \emph{(i)} it is chosen such that draws from the distribution (parameter configurations) correspond to systems of equations that give rise to a specified emergent property (a set of constraints); and \emph{(ii)} it is chosen to have maximum entropy given those constraints, such that we identify all likely parameters and can use the distribution to reason about parametric sensitivity and degeneracies \cite{transtrum2015perspective}.
+%First, we use stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical} to enable inference in likelihood-free models of neural computation.
+%Second, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2017density, papamakarios2017masked}, which confers fast sampling and sensitivity measurements.
+%Third, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model.
+%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
+%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
+%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
+%
+%These challenges motivate the development of a novel inference framework called emergent property inference (EPI).
+%As an adaption of variational inference \cite{saul1998mean}, EPI infers parameter distributions that produce an emergent property: not a singular dataset, but a collection of datasets exhibiting some mathematical criteria.
+%EPI constrains the predictions of the inferred parameter distribution to produce the emergent property, which requires a variant of probabilistic inference methods \cite{loaiza2017maximum}.
+%Importantly, EPI uses deep learning to make rich, flexible approximations to the parameter distribution \cite{rezende2015variational} that produces an emergent property.
+%The structure captured by these deep probability distributions are scientifically valuable, revealing the sensitivity and robustness of the emergent property to different parameter combinations.
+%Perhaps most powerfully, EPI facilitates inference in mechanistic models, allowing theorists to capture rich parametric structure in biologically realistic models that is conditioned upon the emergent phenomena of interest.
+
+%The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, attractor states and phase transitions, and predictions made by the model.
+
+%This creates an unfavorable tradeoff.
+On the one hand, one may tractably analyze systems of equations with unrealistic assumptions (for example symmetry or gaussianity), mathematically formalizing how parameters affect computation in a too-simple model.
+%On the other hand, one may choose a more biologically accurate, scientifically relevant model at the cost of \emph{ad hoc} approaches to analysis (such as simply examining simulated activity), potentially resulting in bad inference of parameters and thus erroneous scientific predictions or conclusions.
+
+%Exact Bayesian inference is rarely tractable, so Markov chain Monte Carlo (MCMC) and variation inference (VI) techniques are used in practice.
+
+% now transition to deep ML...
+%Of course, this same tradeoff has been confronted in many scientific fields characterized by the need to do inference in complex models.
+%In response, the machine learning community has made remarkable progress in recent years, via the use of deep neural networks as a powerful inference engine: a flexible function family that can map observed phenomena (in this case the measurable signal of some computation) back to probability distributions quantifying the likely parameter configurations.
+%One celebrated example of this approach from machine learning, of which we draw key inspiration for this work, is the variational autoencoder \cite{kingma2013auto, rezende2014stochastic}, which uses a deep neural network to induce an (approximate) posterior distribution on hidden variables in a latent variable model, given data.
+
+%However, these inference tools have not significantly influenced the study of theoretical neuroscience models, for at least three reasons.
+%First, at a practical level, the nonlinearities and dynamics of many theoretical models are such that conventional inference tools typically produce a narrow set of insights into these models.
+%Indeed, only in the last few years has deep learning research advanced to a point of relevance to this class of problem.
+%Second, the object of interest from a theoretical model is not typically data itself, but rather a qualitative phenomenon -- inspection of model behavior, or better, a measurable signature of some computation -- an \emph{emergent property} of the model.
+%Third, because theoreticians work carefully to construct a model that has biological relevance, such a model as a result often does not fit cleanly into the framing of a statistical model.
+%Technically, because many such models stipulate a noisy system of differential equations that can only be sampled or realized through forward simulation, they lack the explicit likelihood and priors central to the probabilistic modeling toolkit.
+
+%First, we stipulate a bijective deep neural network that induces a flexible family of probability distributions over model parameterizations with a probability density we can calculate \cite{rezende2015variational, dinh2016density, papamakarios2017masked}.
+%Second, we quantify the notion of emergent properties as a set of moment constraints on datasets generated by the model.  
+%Thus, an emergent property is not a single data realization, but a phenomenon or a feature of the model, which is ultimately the object of interest in theoretical neuroscience.
+%Conditioning on an emergent property requires a variant of deep probabilistic inference methods, which we have previously introduced \cite{loaiza2017maximum}.
+%Third,  because we cannot assume the theoretical model has explicit likelihood on data or the emergent property of interest, we use stochastic gradient techniques in the spirit of likelihood free variational inference \cite{tran2017hierarchical}.
+%Taken together, emergent property inference (EPI) provides a methodology for inferring parameter configurations consistent with a particular emergent phenomena in theoretical models.
+%We use a classic example of parametric degeneracy in a biological system, the stomatogastric ganglion \cite{goldman2001global}, to motivate and clarify the technical details of EPI.
+
+%First, we use EPI to produce a set of verifiable hypotheses of input-responsivity in a four neuron-type dynamical model of primary visual cortex; we then validate these hypotheses in the model.
+%Second, we demonstrated how the systematic application of EPI to levels of task performance can generate experimentally testable hypotheses regarding connectivity in superior colliculus.  
+%Third, we use EPI to uncover the sources of error in a low-rank recurrent neural network executing a simple mathematical task.

### 2021-04-27 written/eLifeRevision/Bittner_eLife2020.tex
citation review

@@
 \q{intro_first}{The fundamental practice of theoretical neuroscience is to use a mathematical model to understand neural computation, whether that computation enables perception, action, or some intermediate processing.
-A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological}.
+A neural circuit is systematized with a set of equations -- the model -- and these equations are motivated by biophysics, neurophysiology, and other conceptual considerations \cite{kopell1988coupled,  marder1998biophysics, abbott2008theoretical, wang2010neurophysiological, o2015computational}.
 The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
 The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
-The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, o2014cell}.}
+The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, erguler2011practical, mannakee2016sloppiness}.}
 
 Ideally, one carefully designs a model and analytically derives how computational properties determine model parameters.
-Seminal examples of this gold standard include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
+Seminal examples of this gold standard include our field's understanding of memory capacity in associative neural networks \cite{hopfield1982neural}, chaos and autocorrelation timescales in random neural networks \cite{sompolinsky1988chaos}, central pattern generation \cite{olypher2007using}, the paradoxical effect \cite{tsodyks1997paradoxical}, and decision making \cite{wong2006recurrent}.
 Unfortunately, as circuit models include more biological realism, theory via analytical derivation becomes intractable.
-Absent this analysis, statistical inference offers a toolkit by which to solve the inverse problem by identifying, at least approximately, the distribution of parameters that produce computations in a biologically realistic model \cite{foster1993significance, prinz2004similar, achard2006complex, o2014cell, alonso2019visualization}.
+Absent this analysis, statistical inference offers a toolkit by which to solve the inverse problem by identifying, at least approximately, the distribution of parameters that produce computations in a biologically realistic model \cite{foster1993significance, prinz2004similar, achard2006complex, fisher2013modeling, o2014cell, alonso2019visualization}.
 
@@
 First, often we directly use an \emph{exemplar dataset}: a collection of samples that express the computation of interest, this data being gathered either experimentally in the lab or from a computer simulation.
-While in some sense the best choice given its connection to experiment \cite{paninski2018neural} , some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
+While in some sense the best choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
 Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or evidence, rather than the computation itself.
@@
 \q{intro_epi}{An emergent property (EP) is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of datasets drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.
-This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.
-However, the adaptation of these techniques to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for scalable, flexible generative modeling \cite{dinh2017density, kingma2018glow}.
+This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, karlsson2012efficient, hines2014determination, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.
+However, the adaptation of modern techniques to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for scalable, flexible generative modeling \cite{dinh2017density, kingma2018glow}.
 We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.}
@@
 %The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to computation in interpretable models of neural circuits (see Section \ref{methods_related}).
+In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to computation in neural circuit models (see Section \ref{methods_related}).
 

### 2021-04-28 written/eLifeRevision/Bittner_eLife2020.tex
ref import and discussion

@@
 \q{discussion}{Methodology for statistical inference in circuit models has evolved considerably in recent years.
-Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but more recently developed methodology employs deep learning to improve efficiency or provide deep, flexible distribution approximations.
-SNPE \cite{gonccalves2019training}, developed along with EPI (see Section \ref{methods_related}), has been used for posterior inference of parameters in circuit models conditioned upon exemplar data used to represent computation.
-Like SNPE, EPI is a deep inference technique, but it infers parameter distributions that only produce the computation of interest  (see Section \ref{results_RNN}).
-
-Exemplary data versus emergent properties aside, EPI has better scaling properties than SNPE when emergent property gradients are tractable (Section \ref{results_RNN}).  
-However, SNPE has its own relative advantages.
-SNPE is effective when circuit model simulations are lengthy or nondifferentiable.
+Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but EPI and other recently developed methodology (SNPE \cite{gonccalves2019training}) employ deep learning to improve efficiency and provide deep, flexible distribution approximations.
+SNPE has been used for posterior inference of parameters in circuit models conditioned upon exemplar data used to represent computation, but it does not infer parameter distributions that only produce the computation of interest like EPI  (see Section \ref{results_RNN}).
+However, SNPE is more effective when circuit model simulations are lengthy or nondifferentiable.
 For example, SNPE can infer the STG parameters that produce the pyloric rhythm \cite{gonccalves2019training}, while EPI cannot.
-Thus, while it is nice to infer parameter distributions with constrained emergent properties with EPI,  SNPE is most appropriate when emergent property gradients are intractable.
-
-The scientific analyses of Sections \ref{results_V1} and \ref{results_SC} derived theoretical findings by querying the structure of the inferred distribution of EPI.
-By measuring the dimensions in which probability decreases fastest, the dimensions of \emph{sensitivity}, we gain valuable understanding of how model parameters govern the emergent property.
-A rich literature on parameter sensitivity analyses in biological models presents several methods towards this scientific approach \cite{saccomani2003parameter, raue2009structural, karlsson2012efficient, raman2017delineating}.
-The value offered by EPI (and other deep inference methods like SNPE), is that the once the flexible deep probability distributions are fit to the parameter distribution, this distribution's structure can be quantified at any parameter choice, offering instantly available sensitivity measurements.
-Together, the ability to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a powerful new method for addressing inverse problems in theoretical neuroscience.}
+Finally, we show that EPI has better scaling properties than SNPE when emergent property gradients are tractable (Section \ref{results_RNN}).
+In summary, choice of deep inference technique should consider emergent property complexity and differentiability, dimensionality of parameter space, and the importance of constraining the model behavior predicted by the inferred parameter distribution.
+
+In this paper, we prove out the value of deep inference for parameter sensitivity analyses at both the local and global level.
+With these techniques, flexible deep probability distributions are optimized to capture global structure by approximating the full distribution of suitable parameters.
+Importantly, the local structure of this deep probability distribution can be quantified at any parameter choice, offering instant sensitivity measurements after fitting.
+For example, the global structure captured by EPI revealed two distinct parameter regimes, which had different local structure quantified by the deep probability distribution (see Section \ref{methods_SC}).
+In comparison, Bayesian MCMC is considered a popular approach for capturing global parameter structure \cite{girolami2011riemann}, but there is no variational approximation (the deep probability distribution in EPI), so sensitivity information is not queryable and sampling remains slow after convergence.
+Local sensitivity analyses (e.g.
+\cite{raue2009structural}) may be performed independently at individual parameter samples, but these methods alone do not capture the full picture in nonlinear, complex distributions.
+Therefore, deep inference is the only approach yielding an object -- the deep probability distribution -- which produces a wholistic assessment of parameter sensitivity at the local and global level, which were used in this study to make novel insights into theoretical models of neural computation.
+Together, the abilities to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a powerful new method for addressing inverse problems in theoretical neuroscience.}
 
@@
 Inference in models defined by systems of differential has been demonstrated with MCMC \cite{girolami2011riemann}, although this approach requires tractable likelihoods.
-Advancements have leveraged structure in stochastic differential equation models to improve likelihood approximations, thus expanding the domain of applicable models \cite{golightly2011bayesian}.
+Advancements have introduced sampling \cite{calderhead2011statistical}, likelihood approximation \cite{golightly2011bayesian}, and uncertainty quantification techniques \cite{chkrebtii2016bayesian} to make MCMC approaches more efficient and expand the class of applicable models.
 
@@
 Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
-SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality (cf. Fig. \ref{fig:RNN}).
+SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality \cite{sisson2018handbook} (cf. Fig. \ref{fig:RNN}).
+Still, this method has enjoyed much success in systems biology \cite{liepe2014framework}.
 Furthermore, once a parameter set has been obtained by SMC-ABC from a finite set of particles, the SMC-ABC algorithm must be run again from scratch with a new population of initialized particles to obtain additional samples.
 
-For scientific model analysis, we seek a parameter distribution represented by an approximating distribution as in variational inference \cite{saul1998mean} : a variational approximation that once optimized yields fast analytic calculations and samples.
+For scientific model analysis, we seek a parameter distribution represented by an approximating distribution as in variational inference \cite{saul1998mean}: a variational approximation that once optimized yields fast analytic calculations and samples.
 For the reasons described above, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples lacking probabilities and have unchanging sampling rate.
@@
 An exciting recent method is capable of recovering the functional form of such unidentifiabilities away from a point by following degenerate dimensions of the fisher information matrix \cite{raman2017delineating}.
-Global structural non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}.
+Global structural non-identifiabilities can be found for models with polynomial or rational dynamics equations using DAISY \cite{saccomani2003parameter}, or through mean optimal transformations \cite{hengl2007data}.
 With EPI, we have all the benefits given by a statistical inference method plus the ability to query the first- or second-order gradient of the probability of the inferred distribution at any chosen parameter value.

### 2021-05-04 written/eLifeRevision/Response.tex
citation review
through 3.4

@@
 Here, EPI yields visually striking, novel insight where the conventional analytic approach became infeasible with increasing neuron types.  See response to main concern \#4.
-\item EPI is used to discover multiple parametric regimes of ``rapid task switching" in a model of superior colliculus.  
+\item EPI is used to discover multiple parametric regimes of rapid task switching in a model of superior colliculus.  
 These regimes are efficiently identified and mechanistically characterized using the probabilistic modeling tools afforded by EPI, and we relate the inferred connectivities to results from optogenetic silencing experiments in rats.
@@
 The reviewer’s suggestion that the method is in some sense using a flat prior is also correct.
-To clarify, EPI does not execute posterior inference, because there is no empirical dataset or specified prior belief in EPI framework.
+To clarify, EPI does not execute posterior inference, because there is no explicit dataset or specified prior belief in the EPI framework.
 However, we derive the relation of EPI to bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.
@@
 
-In our examples, we only infer distributions of parameters with the same units, so issue should not draw concern over the validity of our model analyses.
+In our examples, we only infer distributions of parameters with the same units, so this issue should not draw concern over the validity of our model analyses.
 As suggested, the EPI solution will differ according to relative scaling of parameter values under the maximum entropy selection principle.
 Thus, an important clarification is that sensitivity quantifications are made in the context of the chosen parameter scalings.
-A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.
-It makes most sense to make sure EPI is exploring through parameter space through an industry-recommended range of values, and that sensitivity measurements are interpretable.
+%A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.
+%It makes most sense to make sure EPI is exploring through parameter space through an industry-recommended range of values, and that sensitivity measurements are interpretable.
 
@@
 Similarly, training deep networks via EPI to learn parameter distributions became easier throughout this research project as we learned through trial and error what works well.  
-In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge while capturing complex structure much more regularly (e.g.
+In fact, there has been extraordinary progress in the field of deep probability distributions (specifically normalizing flows), that have allowed EPI to converge regularly while capturing complex structure (e.g.
 Dinh et al.
@@
 2018).
-This manuscript has much value in its explanation of hyperparameter choices, and the extensive set of examples in the online code github. 
-Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: augmented Lagrangian constant, normalizing flow architecture.
+This manuscript's explanation of hyperparameter choices and the extensive set of examples in the online code will serve as valuable guidelines for future research using this method.
+Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: the augmented Lagrangian constant and deep probability distribution architecture.
 
-In general, we cannot know if we have arrived at the global maximum entropy distribution for a given emergent property.  
+In general, we cannot know if we have obtained the global maximum entropy distribution for a given emergent property.  
 The reviewer is correct to point out that multiple distributions may satisfy the emergent property and have different levels of entropy.  
@@
 We derive an analytic equivalence to variational inference (Section 5.1.6) showing that a.) EPI is a valid inference method, and b.) to emphasize that maximum entropy is the normative selection principle of bayesian inference methods in general.
-Thus, the concern of not having the globally optimal inferred distribution is the same that applies to all other inference techniques.
+Thus, the concern of not having the globally optimal inferred distribution is the same that applies to all other statistical inference techniques.
 

### 2021-05-04 written/eLifeRevision/Bittner_eLife2020.tex
through 3.4

@@
 A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
-To solve challenging inverse problems concerning neural datasets, neuroscientists have used statistical inference techniques to much success. 
-However, when theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than explaining a given neural dataset well.
+To solve challenging inverse problems in neuroscience, statistical inference techniques have been used to infer parameters distributions most likely to have produced neural datasets.
 In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
+When theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than a particular dataset.
 Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
@@
 The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about particular parameter configurations?  
+The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about the space, manifold, or distribution of parameters that give rise to it?
 The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, erguler2011practical, mannakee2016sloppiness}.}
@@
 While in some sense the best choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
-Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or evidence, rather than the computation itself.
-Alternatively, researchers often quantify an \emph{emergent property}: a statistic of data that directly quantifies the computation of interest, wherein the dataset is only implicit.  
-While such a choice may seem esoteric, it is not: the above ``gold standard" examples all quantify and focus on some derived feature of the data, rather than the data drawn from the model.}
-
-\q{intro_epi}{An emergent property (EP) is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of datasets drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.
-This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, karlsson2012efficient, hines2014determination, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.
-However, the adaptation of modern techniques to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for scalable, flexible generative modeling \cite{dinh2017density, kingma2018glow}.
-We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.}
-
-Equipped with this method, we prove out EPI by demonstrating its capabilities and presenting novel theoretical findings from its analysis.
+Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or marginal evidence, rather than the computation itself.}
+
+\q{intro_epi}{Alternatively, researchers often quantify an \emph{emergent property} (EP): a statistic of data that directly quantifies the computation of interest, wherein the dataset is implicit.  
+While such a choice may seem esoteric, it is not: the above ``gold standard" examples  \cite{hopfield1982neural, sompolinsky1988chaos, olypher2007using, tsodyks1997paradoxical, wong2006recurrent}
+all quantify and focus on some derived feature of the data, rather than the data drawn from the model.
+An emergent property is of course a dataset by another name, but it suggests different approach to solving the same inverse problem:  here we directly specify the desired emergent property -- a statistic of data drawn from the model -- and the value we wish that property to have, and we set up an optimization program to find the distribution of parameters that produce this computation.
+This statistical framework is not new: it is intimately connected to the literature on approximate bayesian computation \cite{beaumont2002approximate, marjoram2003markov, sisson2007sequential}, parameter sensitivity analyses \cite{raue2009structural, karlsson2012efficient, hines2014determination, raman2017delineating}, maximum entropy modeling \cite{elsayed2017structure, savin2017maximum, mlynarski2020statistical}, and approximate bayesian inference \cite{tran2017hierarchical, gonccalves2019training}; we detail these connections in Section \ref{methods_related}.}
+
+The parameter distributions producing a computation may be thin, curved, bent, or multimodal along various parameter axes and combinations.
+It is by capturing and quantifying this complex structure that EPI offers scientific insight.
+Traditional approximation families (e.g. mean-field or mixture of gaussians) are limited in the distributional structure they may learn.
+To address such restrictions on expressivity, major advancements in machine learning have enabled the use of deep probability distributions as flexible approximating families for such complicated distributions \cite{rezende2015variational, papamakarios2019normalizing} (see Section \ref{methods_NF}).
+However, the adaptation of deep probability distributions to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for efficient and expressive deep generative modeling \cite{dinh2017density, kingma2018glow}.
+We detail our method, which we call emergent property inference (EPI) in Section \ref{results_dgm}.
+
+Equipped with this method, we demonstrate the capabilities of EPI and present novel theoretical findings from its analysis.
 \q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
 Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
-\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals the how recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
+\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
 Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
 EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
-We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of cortical responses in each regime. 
+We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of neural responses in each regime. 
 Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
@@
 In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
-An emergent property studied by Guttierez et al. of this stochastic model is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
+An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
 This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
-Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
+Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies.
+%, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
 %To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
@@
 %In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
+
+%To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
+%In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
 Third, the model parameters producing the emergent property are inferred.
-To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
 By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
-EPI learns a probability distribution of model parameters constrained to produce the emergent property.
-In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
+Before presenting technical details (in the following section), let us understand emergent property inference schematically.
+EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Fig. \ref{fig:STG}E).  
+This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution constrained such that the circuit model produces the emergent property. 
+Once EPI is run, the returned distribution can be used to efficiently generate additional parameter samples.
+Most importantly, the inferred distribution can be efficiently queried to quantify the parametric structure that it captures.
+By quantifying the parametric structure governing the emergent property, EPI  informs the central question of this inverse problem: what aspects or combinations of model parameters have the desired emergent property?
 
@@
 \textbf{C}. The emergent property of intermediate hub frequency.
-Simulated activity traces are colored by $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ of generating parameters. (Panel E).
+Simulated activity traces are colored by log probability of generating parameters in the EPI distribution (Panel E).
 \textbf{D}. For a choice of circuit model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
@@
 Contours of hub neuron frequency error are shown at levels of $.525$, $.53$, ... $.575$ Hz (dark to light gray away from mean).
-Dimension of sensitivity $\mathbf{v}_1$ (solid) and robustness $\mathbf{v}_2$ (dashed).
-\textbf{F} (Top) The predictive distribution of EPI.
+Dimension of sensitivity $\mathbf{v}_1$ (solid arrow) and robustness $\mathbf{v}_2$ (dashed arrow).
+\textbf{F} (Top) The predictions of the EPI distribution.
 The black and gray dashed lines show the mean and two standard deviations according the emergent property.
@@
 
-Before presenting technical details (in the following section), let us understand emergent property inference schematically.
-EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, produces the parameter distribution  EPI (Fig. \ref{fig:STG}E).  
-This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution that produces the emergent property. 
-We can then use this parameter distribution to efficiently generate many parameters producing the emergent property.
-Most importantly, the resulting parameter distribution can be efficiently queried to quantify its own structure, which informs the central question of this inverse problem: how do model parameters govern the emergent property?
-
 \subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
-Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions.
-First, as is typical, we consider the model as a coupled set of differential equations.  
+Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions \cite{rezende2015variational, papamakarios2019normalizing}.
+First, as is typical, we consider the model as a coupled set of noisy differential equations.  
 In this STG example, the model activity (or state) $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
-\begin{equation} 
+\begin{equation} \label{eq:STG_model1}
 \begin{split}
@@
 Mathematically, we define this emergent property with two constraints: that the mean hub frequency is 0.55Hz,}
-\begin{equation}
- \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55\text{Hz}
+\begin{equation}\label{eq:EP_STG1}
+ \mathbb{E}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  0.55
 \end{equation}
 \q{stg_ep_text2}{and that the variance of the hub frequency is moderate}
-\begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2 \text{Hz}^2.
+\begin{equation}\label{eq:EP_STG2}
+\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2
 \end{equation}
 In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
-Formally, the emergent property is the collection of these two constraints
-\begin{equation}\label{eq:EP_STG}
-\mathcal{X}:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] = 0.55\text{Hz},~~\text{Var}_{\mathbf{z},\mathbf{x}}\left[\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})\right] =0.025^2\text{Hz}^2.
-\end{equation}
-In general, an emergent property is a collection of constraints on first-, second- and higher moments of one or more statistics (see Sections \ref{results_RNN}, \ref{results_SC}, and \ref{methods_2DLDS}) that together define the computation.
-
-Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$ that produces the emergent property; in other words, they obey the constraints introduced in Equation \ref{eq:EP_STG}.
-This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, dinh2017density, papamakarios2017masked} (Figure \ref{fig:STG}D, EPI box).
-Deep probability distributions map a simple random variable $\mathbf{z}_0$ (drawn from an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
-\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice made in variational bayesian methods (see Section \ref{methods_VI}).}
+This expectation over $\mathbf{x}$ is taken over any randomness in initial state ($\mathbf{x}(t = 0)$) and noise (e.g. $d\mathbf{B}$ in Equation \ref{eq:STG_model1}).
+Formally, the emergent property is the collection of these two constraints.
+In general, an emergent property is a collection of constraints on statistical moments that together define the computation.
+
+Third, we perform emergent property inference: we find a distribution over parameter configurations $\mathbf{z}$ of models that produce the emergent property; in other words, they satisfy the constraints introduced in Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}.
+This distribution will be chosen from a family of probability distributions $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$, defined by a deep neural network \cite{rezende2015variational, papamakarios2019normalizing} (Figure \ref{fig:STG}D, EPI box).
+Deep probability distributions map a simple random variable $\mathbf{z}_0$ (e.g.
+an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
+\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice commonly made in variational bayesian methods (see Section \ref{methods_VI}).}
 %\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
 In EPI optimization, stochastic gradient steps in $\bm{\theta}$ are taken such that entropy is maximized, and the emergent property $\mathcal{X}$ is produced (see Section \ref{methods_EPI}).
-The inferred EPI distribution is denoted $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since it is conditioned upon the implicit datasets of emergent property $\mathcal{X}$ in a manner methodologically equivalent to exemplar data when using variational bayesian approaches.
-
-\q{results_hess}{EPI produces parameter distributions that can be queried for scientific insight.
+We then denote the inferred EPI distribution as $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, since the structure of the learned parameter distribution is determined by weights and biases $\bm{\theta}$, and this distribution is conditioned upon emergent property $\mathcal{X}$.
+
+\q{results_hess}{The structure of the inferred parameter distributions of EPI can be analyzed to reveal key information about how the circuit model produces the emergent property.
 The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices emblematic of the emergent property (Fig. \ref{fig:STG}E yellow star).
 As probability in the EPI distribution decreases, the emergent property deteriorates.
-Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ does not change will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
-In contrast, if $\mathbf{z}$ is perturbed along a dimension that strongly decreases $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive}.
+Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ change little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
+In contrast, if $\mathbf{z}$ is perturbed along a dimension with strongly decreasing $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive} \cite{raue2009structural, raman2017delineating}.
 By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative, the more sensitive and the closer to zero, the more robust (see Section \ref{methods_stg_hessian}).
-Indeed, samples equidistant from the mode along these EPI-identified dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig.
+Indeed, samples equidistant from the mode along these dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig.
 \ref{fig:STG}E, arrows) agree with error contours (Fig.
@@
 \ref{fig:STG}F activity traces).
-This suggests that changes in conductance along the parameter combination described by $\mathbf{v}_2$ will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
-Once an EPI distribution has been inferred, this Hessian calculation requires trivial computation (see Section \ref{methods_NF}).}
+The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
+Importantly, once an EPI distribution has been learned, the modes and Hessians of the distribution can be measured with trivial computation (see Section \ref{methods_NF}).}
 
 In the following sections, we demonstrate EPI on three neural circuit models across ranges of biological realism, neural system function, and network scale.
-First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of recurrent neural network (RNN) connectivities that exhibit amplified, yet stable responses.
-Also in this RNN example, we emphasize that EPI is the only technique that controls the predictions made by the inferred parameter distribution.
-Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI captures a curved manifold of parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
-Finally, in a model of superior colliculus \cite{duan2018collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of sensitivity ($\mathbf{v}_1(\mathbf{z})$) to mechanistically characterize each regime.
+First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of recurrent neural network connectivities that exhibit amplified, yet stable responses.
+Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI discovers parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
+Finally, in a model of superior colliculus \cite{duan2018collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.
 
@@
 
-\subsection{Scaling inference of RNN connectivity with EPI} \label{results_RNN}
+\subsection{Scaling inference of recurrent neural network connectivity with EPI} \label{results_RNN}
+To understand how EPI scales in comparison to existing techniques, we consider recurrent neural networks (RNNs).
 Transient amplification is a hallmark of neural activity throughout cortex, and is often thought to be intrinsically generated by recurrent connectivity in the responding cortical area \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.
 It has been shown that to generate such amplified, yet stabilized responses, the connectivity of RNNs must be non-normal \cite{goldman2009memory, murphy2009balanced}, and satisfy additional constraints \cite{bondanelli2020coding}.
-In theoretical neuroscience, RNNs are optimized and then examined to show how dynamical systems could execute a given computation \cite{sussillo2014neural, barak2017recurrent}, but such biologically realistic constraints on connectivity are  ignored during optimization for practical reasons.
-In general, access to distributions of connectivity adhering to theoretical criteria like stable amplification, chaotic fluctuations \cite{sompolinsky1988chaos}, or low tangling \cite{russo2018motor} would add scientific value and context to existing research with RNNs.
+In theoretical neuroscience, RNNs are optimized and then examined to show how dynamical systems could execute a given computation \cite{sussillo2014neural, barak2017recurrent}, but such biologically realistic constraints on connectivity \cite{goldman2009memory, murphy2009balanced, bondanelli2020coding} are ignored for simplicity or because constrained optimization is difficult.
+In general, access to distributions of connectivity that produce theoretical criteria like stable amplification, chaotic fluctuations \cite{sompolinsky1988chaos}, or low tangling \cite{russo2018motor} would add scientific value to existing research with RNNs.
 Here, we use EPI to learn RNN connectivities producing stable amplification, and demonstrate the superior scalability and efficiency of EPI to alternative approaches.
+
+We consider a rank-2 RNN with $N$ neurons having connectivity $W = UV^\top$
+ and dynamics
+ \begin{equation}
+ \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x},
+ \end{equation}
+where $U = \begin{bmatrix} \mathbf{U}_1 & \mathbf{U}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{V}_1 & \mathbf{V}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{U}_1 \mathbf{U}_2, \mathbf{V}_1, \mathbf{V}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
+We infer connectivity parameters $\mathbf{z} = \left[\mathbf{U}_1^\top, \mathbf{U}_2^\top, \mathbf{V}_1^\top, \mathbf{V}_2^\top \right]^\top$   that produce stable amplification.
+\q{rnn_ep_text1}{Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
+$\lambda^s_1 > 1$}, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W^\top}{2}$.
+RNNs  with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will be stable with modest decay rate ($\text{real}(\lambda_1)$ close to its upper bound of 1) and exhibit modest amplification ($\lambda_1^s$ close to its lower bound of 1).
+\q{rnn_ep_text2}{EPI can naturally condition on this emergent property}
+\begin{equation}\label{eq:EP_LRRNN}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
+\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix}.
+\end{split}
+\end{equation}
+\q{rnn_ep_text3}{Variance constraints predicate that the majority of the distribution (within two standard deviations) are within the specified ranges.}
+
+For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
+Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
+Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplary dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
+To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
+
 \begin{figure}
@@
 
-We consider a rank-2 RNN with $N$ neurons having connectivity $W = UV^\top$
- and dynamics
- \begin{equation}
- \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x},
- \end{equation}
-where $U = \begin{bmatrix} \mathbf{U}_1 & \mathbf{U}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{V}_1 & \mathbf{V}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{U}_1 \mathbf{U}_2, \mathbf{V}_1, \mathbf{V}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
-We infer connectivity parameterizations $\mathbf{z} = \left[\mathbf{U}_1^\top, \mathbf{U}_2^\top, \mathbf{V}_1^\top, \mathbf{V}_2^\top \right]^\top$   that produce stable amplification.
-\q{rnn_ep_text1}{Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
-$\lambda^s_1 > 1$}, where $\lambda_1$ is the eigenvalue of $W$ with greatest real part and $\lambda^s$ is the maximum eigenvalue of $W^s = \frac{W + W^\top}{2}$.
-RNNs  with $\text{real}(\lambda_1) = 0.5 \pm 0.5$ and $\lambda_1^s = 1.5 \pm 0.5$ will be stable with modest decay rate ($\text{real}(\lambda_1)$ close to its upper bound of 1) and exhibit modest amplification ($\lambda_1^s$ close to its lower bound of 1).
-\q{rnn_ep_text2}{EPI can naturally condition on this emergent property}
-\begin{equation}\label{eq:EP_LRRNN}
-\begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix} \\
-\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix},
-\end{split}
-\end{equation}
-\q{rnn_ep_text3}{under the notion that variance constraints with standard deviation 0.25 predicate that the vast majority of samples (those within two standard deviations) are within the specified ranges.}
-
-For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
-We ran sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} with exemplary dataset $\mathbf{x}_0 = \bm{\mu}$.
-SMC-ABC is a rejection sampling approach that uses SMC techniques to improve efficiency, and SNPE approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
-Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, and these predictions of the inferred posteriors are typically affected by model characteristics (e.g.
-$N$ and $g$, Fig.
-\ref{fig:RNN2}).
-To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
-
-As the number of neurons $N$ in the RNN is scaled, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
+As the number of neurons $N$ in the RNN, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, is scaled, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
 \ref{fig:RNN}A).
 It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
-It is well-known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
+It is well known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
 Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
@@
 \ref{fig:RNN3}).
-In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time and simulation count.
+In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time (elapsed real time) and simulation count.
 
@@
 When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:RNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|r(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions.}
-
-EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$) on each iteration of optimization.
-This agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}.
-\q{rnn_gradients}{Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
-Evidenced by this analysis, EPI is a clear choice for inferring high dimensional parameter distributions when the emergent property gradient is efficiently calculated.
-This can be invaluable for understanding how RNNs produce complex computations.
-Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
-However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
-The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, SNPE, which does not require differentiability of the emergent property has proved to be a powerful approach \cite{gonccalves2019training}.}
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|\mathbf{x}(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions (Fig.
+\ref{fig:RNN}D).
+Even for moderate neuron counts, the predictions of the inferred distribution of SNPE are highly dependent on $N$ and $g$, while EPI maintains the emergent property across choices of RNN (see Section \ref{methods_RNN_params}).}
+
+To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
+This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient and sampling efficiency (scaling and speed versus generality).
+Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
+If the emergent property gradient is efficiently calculated, EPI is a clear choice for inferring high dimensional parameter distributions. 
 In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.
@@
 \subsection{EPI reveals how recurrence with multiple inhibitory subtypes governs excitatory variability in a V1 model} \label{results_V1}
+
+Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena in primary visual cortex (V1).
+In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
+Recent theoretical work \cite{hennequin2018dynamical} shows that stabilized E-I models reproduce the effect of variability suppression \cite{Churchland2010}.
+%However, the way in which E-I recurrence structures variability into shared and private dimensions \cite{semedo2019cortical} is not understood.
+Furthermore, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
+\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
+Here, we use EPI on a model of V1 with biologically realistic connectivity to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
+
+We considered response variability of a nonlinear dynamical V1 circuit model (Fig.
+\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
+Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}) and
+an external input with mean $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
+The external input has an additive noisy component $\bm{\epsilon}$ with variance $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
+This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Fig. \ref{fig:V1}B).
+This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} 
+\begin{equation}
+    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}),
+\end{equation}
+generalized to have multiple inhibitory neuron types.  It introduces stochasticity to four neuron-type models of V1 \cite{litwin2016inhibitory}.
+Stochasticity and inhibitory multiplicity introduce substantial complexity to the mathematical treatment of this problem (see Section \ref{methods_V1_complexity}) motivating the analysis of this model with EPI.
+Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure}, and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability.
 
@@
 
-Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena in primary visual cortex (V1).
-In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}.
-Recent theoretical work \cite{hennequin2018dynamical} shows that stabilized E-I models reproduce the effect of variability suppression \cite{Churchland2010}.
-However, the way in which E-I recurrence structures variability into shared and private dimensions \cite{semedo2019cortical} is not understood.
-Furthermore, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
-Here, we use EPI on a model of V1 with biologically realistic connectivity to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
-
-We considered response variability of a nonlinear dynamical V1 circuit model (Fig.
-\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
-Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}) and
-an external input with mean $\mathbf{h}$, which determines population rate via supralinear nonlinearity $\phi(\cdot) = \left[\cdot \right]^2_+$.
-The external input has an additive noisy component $\bm{\epsilon}$ with variance $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
-This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Fig. \ref{fig:V1}B).
-This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} 
-\begin{equation}
-    \tau \frac{d\mathbf{x}}{dt} = -\mathbf{x} +\phi(W\mathbf{x} + \mathbf{h} + \bm{\epsilon}).
-\end{equation}
-generalized to have multiple inhibitory neuron types, and introduces stochasticity to previous four neuron-type models of V1 \cite{litwin2016inhibitory}.
-Stochasticity and inhibitory multiplicity introduce substantial complexity to the mathematical treatment of this problem (see Section \ref{methods_V1_complexity}) motivating the treatment of this model with EPI.
-Here, we consider fixed weights $W$ and input $\mathbf{h}$ \cite{palmigiano2020structure}, and study the effect of input variability $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V]^\top$ on excitatory variability.
-
-\q{v1_ep_text1}{We quantify levels $y$ of E-population variability with the emergent property}
+\q{v1_ep_text1}{We quantify levels of E-population variability by studying two emergent properties}
 \begin{equation}\label{eq:EP_V1}
 \begin{split}
-\mathcal{X}(y) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ y \\ 
- \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2,
+\mathcal{X}(5\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{ Hz}~~~~~~~\mathcal{X}(10\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{ Hz}  \\
+ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
 \end{split}
@@
 \q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-In the following analyses, we compare levels of 5Hz and 10Hz, and select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
+In the following analyses, we select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
 
@@
 \ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
-Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the high variability in those dimensions (Fig. \ref{fig:V1}D, bottom-right).
+Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the, unexpectedly, high variability in those dimensions (Fig.
+\ref{fig:V1}D, bottom-right).
 Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
 
-Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensatorily decreases to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
+Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
 This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E).
@@
 \ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
-The strong interactive influence of E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
+The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
 \ref{fig:V1_3}), meaning that the E-population is P-stabilized.
@@
 \q{v1_summary}{EPI revealed the quadratic dependence of excitatory variability on input variability to the E- and P-populations, as well as its independence to input from the other two inhibitory populations.
-In a simplified model ($\tau = \tau_{\text{noise}}$), it can be shown that surfaces of equal variance are ellipsoids as a function of $\bm{\sigma}$.
+In a simplified model ($\tau = \tau_{\text{noise}}$), it can be shown that surfaces of equal variance are ellipsoids as a function of $\bm{\sigma}$ (see Section \ref{methods_V1_complexity}).
 Nevertheless, the sensitive and degenerate parameters are challenging to predict mathematically, since the covariance matrix depends on the steady-state solution of the network \cite{hennequin2018dynamical,Gardiner2009}, and terms in the covariance expression increase quadratically with each additional neuron-type population (see also Section \ref{methods_V1_complexity}).
-This emphasizes the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
+By pointing out this mathematical complexity, we emphasize the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
 While we have just shown that EPI can be used to investigate fundamental aspects of sensory computation, in the next two sections, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to behavioral experiments.
@@
 SNPE has been used for posterior inference of parameters in circuit models conditioned upon exemplar data used to represent computation, but it does not infer parameter distributions that only produce the computation of interest like EPI  (see Section \ref{results_RNN}).
-However, SNPE is more effective when circuit model simulations are lengthy or nondifferentiable.
-For example, SNPE can infer the STG parameters that produce the pyloric rhythm \cite{gonccalves2019training}, while EPI cannot.
 Finally, we show that EPI has better scaling properties than SNPE when emergent property gradients are tractable (Section \ref{results_RNN}).
@@
 Together, the abilities to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a powerful new method for addressing inverse problems in theoretical neuroscience.}
+
+Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
+However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
+The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
+In such cases, SNPE, which does not require differentiability of the emergent property has proved to be a powerful approach \cite{gonccalves2019training}.
 
@@
 
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equation \ref{eq:EP_STG}.
+In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}.
 The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
@@
 Intermediate hub frequency, like all other emergent properties in this work, is defined by the mean and variance of the emergent property statistics.
-In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz, and variance was chosen to be (0.025Hz)$^2$ (Equation \ref{eq:EP_STG}).
+In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz,(Equation \ref{eq:EP_STG1}) and variance was chosen to be (0.025Hz)$^2$ (Equation \ref{eq:EP_STG2}).
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figRNN2/figRNN2.pdf}
-\end{center}
-\caption{\small (RNN2): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
-\textbf{A}.
-Predictive distribution of EPI (blue) and SNPE (orange) inferred connectivity of RNNs exhibiting stable amplification with $N=2$ (top), $N=10$ (bottom), $g=0.01$ (left), and $g=0.1$ (right).
-\textbf{B}. Entropy of parameter distribution approximations throughout optimization with $N=2$ (top), $N=10$ (bottom), $g=0.1$ (dark shade), and $g=0.01$ (light shade).
-\textbf{C}. Validation log probabilities throughout SNPE optimization. Same conventions as B.
-\textbf{D}. Adherence to EPI constraints. Same conventions as B.
-}
-\label{fig:RNN2}
-\end{figure}
-
-To clarify the difference in objectives of EPI and SNPE, we show their results on RNN models with different numbers of neurons $N$ and random strength $g$.  
-The parameters inferred by EPI consistently produces the same mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$, while those inferred by SNPE change according to the model definition (Fig.
-\ref{fig:RNN2}A).
-For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater randomness (Fig. \ref{fig:RNN2}B top, orange).
-At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Fig. \ref{fig:RNN2}B top).
-However at $N=10$, SNPE results in a predictive distribution of more widely dispersed eigenvalues (Fig.
-\ref{fig:RNN2}A bottom), and an inferred posterior with greater entropy than EPI (Fig.
-\ref{fig:RNN2}B bottom).
-We highlight these differences not to focus on an insightful trend, but to emphasize that these methods optimize different objectives with different implications.
-
-Note that SNPE converges when it's validation log probability has saturated after several rounds of optimization (Fig.
-\ref{fig:RNN2}C), and that EPI converges after several epochs of its own optimization to enforce the emergent property constraints (Fig.
-\ref{fig:RNN2}D blue).
-Importantly, as SNPE optimizes its posterior approximation, the predictive means change, and at convergence may be different than $\mathbf{x}_0$ (Fig. \ref{fig:RNN2}D orange, left).
-It is sensible to assume that predictions of a well-approximated SNPE posterior should closely reflect the data on average (especially given a uniform prior and a low degree of stochasticity), however this is not a given.
-Furthermore, no aspect of the SNPE optimization controls the variance of the predictions (Fig. \ref{fig:RNN2}D orange, right).
-
 To compare the efficiency of these algorithms for inferring RNN connectivity distributions producing stable amplification, we develop a convergence criteria that can be used across methods.
@@
 We found no way to improve the convergence rate of SNPE without making more aggressive hyperparameter choices requiring high numbers of simulations.
-
-In Figure \ref{fig:RNN}C-D, we show samples form the random seed resulting in emergent property convergence at greatest entropy (EPI), the random seed resulting in greatest validation log probability (SNPE), and the result of all converged random seeds (SMC).
+In Figure \ref{fig:RNN}C-D, we show samples from the random seed resulting in emergent property convergence at greatest entropy (EPI), the random seed resulting in greatest validation log probability (SNPE), and the result of all converged random seeds (SMC).
+
+\subsubsection{Effect of RNN parameters on EPI and SNPE inferred distributions} \label{methods_RNN_params}
+To clarify the difference in objectives of EPI and SNPE, we show their results on RNN models with different numbers of neurons $N$ and random strength $g$.  
+The parameters inferred by EPI consistently produces the same mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$, while those inferred by SNPE change according to the model definition (Fig.
+\ref{fig:RNN2}A).
+For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater randomness (Fig. \ref{fig:RNN2}B top, orange).
+At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Fig. \ref{fig:RNN2}B top).
+However at $N=10$, SNPE results in a predictive distribution of more widely dispersed eigenvalues (Fig.
+\ref{fig:RNN2}A bottom), and an inferred posterior with greater entropy than EPI (Fig.
+\ref{fig:RNN2}B bottom).
+We highlight these differences not to focus on an insightful trend, but to emphasize that these methods optimize different objectives with different implications.
+
+Note that SNPE converges when it's validation log probability has saturated after several rounds of optimization (Fig.
+\ref{fig:RNN2}C), and that EPI converges after several epochs of its own optimization to enforce the emergent property constraints (Fig.
+\ref{fig:RNN2}D blue).
+Importantly, as SNPE optimizes its posterior approximation, the predictive means change, and at convergence may be different than $\mathbf{x}_0$ (Fig. \ref{fig:RNN2}D orange, left).
+It is sensible to assume that predictions of a well-approximated SNPE posterior should closely reflect the data on average (especially given a uniform prior and a low degree of stochasticity), however this is not a given.
+Furthermore, no aspect of the SNPE optimization controls the variance of the predictions (Fig. \ref{fig:RNN2}D orange, right).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figRNN2/figRNN2.pdf}
+\end{center}
+\caption{\small (RNN2): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
+\textbf{A}.
+Predictive distribution of EPI (blue) and SNPE (orange) inferred connectivity of RNNs exhibiting stable amplification with $N=2$ (top), $N=10$ (bottom), $g=0.01$ (left), and $g=0.1$ (right).
+\textbf{B}. Entropy of parameter distribution approximations throughout optimization with $N=2$ (top), $N=10$ (bottom), $g=0.1$ (dark shade), and $g=0.01$ (light shade).
+\textbf{C}. Validation log probabilities throughout SNPE optimization. Same conventions as B.
+\textbf{D}. Adherence to EPI constraints. Same conventions as B.
+}
+\label{fig:RNN2}
+\end{figure}
 

### 2021-05-06 written/eLifeRevision/Bittner_eLife2020.tex
latest revision

@@
 \usepackage{qting}
+
+\newcommand\startsupplement{%
+    \makeatletter 
+       \setcounter{table}{0}
+       \renewcommand{\thetable}{S\arabic\c@table}
+       \setcounter{figure}{0}
+       \renewcommand{\thefigure}{S\arabic{figure}}
+    \makeatother}
 
@@
 \q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
-Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
+Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2019collicular}.
 EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
@@
 Note that the behavior of this model will be critically dependent on its parameterization -- the choices of conductance parameters $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$.
-
-Second, once the model is selected, one must specify what the model should produce.
-In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
-An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
-This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
-Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies.
-%, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
-%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
-
-%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
-%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
-%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
-% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
-%Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
-%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
-
-%To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
-%In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
-Third, the model parameters producing the emergent property are inferred.
-By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
-Before presenting technical details (in the following section), let us understand emergent property inference schematically.
-EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Fig. \ref{fig:STG}E).  
-This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution constrained such that the circuit model produces the emergent property. 
-Once EPI is run, the returned distribution can be used to efficiently generate additional parameter samples.
-Most importantly, the inferred distribution can be efficiently queried to quantify the parametric structure that it captures.
-By quantifying the parametric structure governing the emergent property, EPI  informs the central question of this inverse problem: what aspects or combinations of model parameters have the desired emergent property?
 
@@
 
+Second, once the model is selected, one must specify what the model should produce.
+In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
+An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
+This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
+Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies.
+%, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
+%To continue our running STG example, one such emergent property is the phenomenon of \emph{network syncing} -- in certain parameter regimes, the frequency of the hub neuron matches that of the fast and slow populations at an intermediate frequency.
+
+%Third, qualitative parameter analysis ensues: since precise mathematical analysis is intractable in this model, a brute force sweep of parameters is done \cite{gutierrez2013multiple}.
+%Subsequently, a qualitative description is formulated to describe the different parameter configurations that lead to the emergent property.  
+%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model. 
+% Once we have such a methodology, we can infer a probability distribution over parameter configurations that produce this emergent property.
+%Sensitivity analyses \cite{raue2009structural} are also used to identify which parameter contributions have the greatest effect.
+%In this last step lies the opportunity for a precise quantification of the emergent property as a statistical feature of the model.
+
+%To infer the STG parameters of intermediate hub frequency with existing methodology, we need an exemplar dataset: experimentally recorded or synthesized.
+%In this last step lies the opportunity for a shift away from a dataset-oriented representation of model output towards that of an implicit dataset, where the only structure is the emergent property of interest.
+Third, the model parameters producing the emergent property are inferred.
+By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
+Before presenting technical details (in the following section), let us understand emergent property inference schematically.
+EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Fig. \ref{fig:STG}E).  
+This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution constrained such that the circuit model produces the emergent property. 
+Once EPI is run, the returned distribution can be used to efficiently generate additional parameter samples.
+Most importantly, the inferred distribution can be efficiently queried to quantify the parametric structure that it captures.
+By quantifying the parametric structure governing the emergent property, EPI  informs the central question of this inverse problem: what aspects or combinations of model parameters have the desired emergent property?
+
 \subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
@@
 \q{results_hess}{The structure of the inferred parameter distributions of EPI can be analyzed to reveal key information about how the circuit model produces the emergent property.
-The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices emblematic of the emergent property (Fig. \ref{fig:STG}E yellow star).
-As probability in the EPI distribution decreases, the emergent property deteriorates.
+As probability in the EPI distribution decreases away from the mode of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ (Fig. \ref{fig:STG}E yellow star), the emergent property deteriorates.
 Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ change little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
@@
 Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI discovers parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
-Finally, in a model of superior colliculus \cite{duan2018collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.
+Finally, in a model of superior colliculus \cite{duan2019collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.
 
@@
 For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
-Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
+Sequential Monte Carlo approximate bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
 Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplary dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
@@
 Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
-\ref{fig:RNN1}), and tested more aggressive SNPE hyperparameterizations when SNPE failed to converge (Fig.
+\ref{fig:RNN1}), and tested more aggressive SNPE hyperparameter choices when SNPE failed to converge (Fig.
 \ref{fig:RNN3}).
@@
 To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
-This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient and sampling efficiency (scaling and speed versus generality).
+This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient-based and sampling-based estimators (scaling and speed versus generality).
 Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
@@
 \end{center}
-\vspace{-1cm}
 \caption{Emergent property inference in the stochastic stabilized supralinear network (SSSN)
@@
 Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
-This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E).
+This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E and \ref{fig:V1_1}).
 Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
@@
 The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
-\ref{fig:V1_3}), meaning that the E-population is P-stabilized.
+\ref{fig:V1_4}), meaning that the E-population is P-stabilized.
 Future research may uncover a link between the population of network stabilization and compensatory interactions governing excitatory variability.
@@
 \subsection{EPI identifies two regimes of rapid task switching} \label{results_SC}
-It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity produce this computation.
-In this experimental setup, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from the stimulus in the Anti (A) task (Fig.
-\ref{fig:SC}A).
-Neural recordings in superior colliculus (SC) exhibited two populations of neurons that represented task context (Pro or Anti).
-Furthermore, Pro/Anti neurons in each hemisphere were strongly correlated with the animal's decision \cite{duan2018collicular}.
-These results motivated a model of SC that is a four-population dynamical system with functionally-defined neuron-types.
-Here, our goal is to understand how connectivity in this circuit model governs the ability to perform rapid task switching: to respond with satisfactory accuracy in both tasks on randomly interleaved trials.
-
-In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$.
+It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what neural mechanisms produce this computation.
+In this experimental setup, rats were given an explicit task cue on each trial, either Pro or Anti. 
+After a delay period, rats were shown a stimulus, and made a context  (task) dependent response (Fig. \ref{fig:SC}A). 
+In the Pro task, rats were required to orient towards the stimulus, while in the Anti task, rats were required to orient away from the stimulus.
+Pharmacological inactivation of the SC impaired rat performance, and time-specific optogenetic inactivation revealed a crucial role for the SC on the cognitively demanding Anti trials \cite{duan2019collicular}.
+These results motivated a nonlinear dynamical model of the SC containing four functionally-defined neuron-type populations.
+In Duan et al. 2019, a computationally intensive procedure was used to obtain a set of 373 connectivity parameters that qualitatively reproduced these optogenetic inactivation results.
+To build upon the insights of this previous work, we use the probabilistic tools afforded by EPI to identify and mechanistically characterize two linked, yet distinct regimes of rapid task switching connectivity.
+
+In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ \cite{duan2019collicular}.
 The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig4/fig4.pdf}
-\end{center}
-\caption{ 
+\includegraphics[scale=0.75]{figures/fig4/fig4.pdf}
+\end{center}
+\vspace{-.25cm}
+\caption{\small
 \textbf{A}. Rapid task switching behavioral paradigm (see text). 
@@
 \textbf{C}. The EPI inferred distribution of rapid task switching networks.  
-Stars indicate modes $\mathbf{z}^*$ whose color indicates connectivity regime (see legend Fig \ref{fig:SC}D).
+Red/purple parameters indicate modes $\mathbf{z}^*(sW)$ colored by $sW$.
 Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
 (Bottom-left) EPI predictive distribution of task accuracies.
-\textbf{D}. The connectivity regimes have different responses to perturbation.
-(Top) Mean and standard error ($N_{\text{test}}$ = 25) of accuracy with respect to perturbation along the sensitivity dimension of each mode $\mathbf{z}^*$.
-(Middle) Same with perturbation in the dimension of increasing $\lambda_{\text{task}}$ ($\mathbf{v}_{\text{task}}$).
-(Bottom) Same with perturbation in the dimension of increasing $\lambda_{\text{diag}}$ ($\mathbf{v}_{\text{diag}}$).
+\textbf{D}.
+Mean and standard error ($N_{\text{test}}$ = 25) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
 }
@@
 \end{figure}
+
+%\begin{figure}
+%\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=5cm}}]{figure}[\FBwidth]
+%{\caption{\small
+%\textbf{A}. Rapid task switching behavioral paradigm (see text). 
+%\textbf{B}. Model of superior colliculus (SC). Neurons: LP - Left Pro, RP - Right Pro, LA - Left Anti, RA - Right Anti. 
+%Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
+%\textbf{C}. The EPI inferred distribution of rapid task switching networks.  
+%Red/purple parameters indicate modes $\mathbf{z}^*(sW)$ colored by $sW$.
+%Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
+%(Bottom-left) EPI predictive distribution of task accuracies.
+%\textbf{D}.
+Mean and standard error ($N_{\text{test}}$ = 25) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode $\mathbf{z}^*(sW)$ by distance $\delta$ along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
+%}
+%\label{fig:test}}
+%{\hspace{-.5in}\includegraphics[scale=.73]{figures/fig4/fig4.pdf}}
+%\end{figure}
 
@@
 
-The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, middle-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
+The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, bottom-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
 This parameter distribution has rich structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
-Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8})
-This is most saliently pointed out in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
-The two regimes produce different types of responses in the Pro and Anti tasks (Fig. SC2).
-Not only has EPI captured this complicated distribution of connectivities producing rapid task switching, we can query the EPI distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.
-
-To distinguish these two regimes, we use the EPI distribution to identify two sets of modes.
-By fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we arrive at two solutions $\mathbf{z}^*(hW_{\text{fixed}}, r)$ where regime $r \in [1,2]$, and regime 1 is that of greater $sW$ (see Section \ref{methods_sc_regime}).
-As $hW_{\text{fixed}}$ increases, the modes coalesce to intermediate parameters reflecting a transition between the two sets of modes (Fig. \ref{fig:SC4} top).
-By using EPI to connect these two regimes through this transitionary region of parameter space, we can explore what distinguishes the two regimes by stepping from the prototypical connectivity of regime 1 to that of regime 2.
-
-Although the connectivities gradually coalesce to the transitionary part of parameter space, the sensitivity dimensions $\mathbf{v}_1(\mathbf{z})$ are categorically different across regimes (Fig.
-\ref{fig:SC4} bottom).
-The sensitivity dimension identifies the parameter combination which causes the emergent property to diminish with the shortest perturbation.
-Since the two regimes have different $\mathbf{v}_1(\mathbf{z})$, this suggests they have different pathologies in their connectivity.
-By perturbing connectivity in each regime along the sensitivity dimension, we can get a sense of the differing nature of these pathologies.
-
-When perturbing connectivity along the sensitivity dimension, Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
+Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8}).
+This is most saliently observed in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
+By identifying the modes of the EPI distribution $\mathbf{z}^*(sW)$ at different values of $sW$ (Fig.
+\ref{fig:SC}C red/purple dots), we can quantify this change in distributional structure with the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$ (Fig.
+\ref{fig:SC}C red/purple arrows).
+Note that the directionality of these sensitivity dimensions at $\mathbf{z}^*(sW)$ changes distinctly with $sW$, and are perpendicular to the flat dimensions of the EPI distribution that preserve rapid task switching.
+These two directionalities of sensitivity motivate the distinction of connectivity into two regimes, which produce different types of responses in the Pro and Anti tasks (Fig. \ref{fig:SC2}).
+
+When perturbing connectivity along the sensitivity dimension away from the modes
+\begin{equation}
+\mathbf{z} = \mathbf{z}^*(sW) + \delta \mathbf{v}_1(\mathbf{z}^*(sW)),
+\end{equation}
+Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
 However, there is a stark difference between regimes in Anti accuracy. 
-Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, top-right).
-These distinct pathologies of rapid task switching are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(r=1))$ vs $\mathbf{v}_1(\mathbf{z}^*(r=2))$) and explain the sharp change in local structure of the EPI distribution.
-
-To further examine the two regimes, we can perturb $\mathbf{z}$ in the same way along dimensions that independently change the eigenvalues of the connectivity matrix (which has constant eigenvectors with respect to $\mathbf{z}$).
+Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, bottom-left).
+The sharp change in local structure of the EPI distribution is therefore explained by distinct sensitivities: Anti accuracy diminishes in only one or both directions of the sensitivity perturbation.
+
+To understand the mechanisms differentiating the two regimes, we can make connectivity perturbations along dimensions that only modify a single eigenvalue of the connectivity matrix.
 These eigenvalues $\lambda_{\text{all}}$, $\lambda_{\text{side}}$, $\lambda_{\text{task}}$, and $\lambda_{\text{diag}}$ correspond to connectivity eigenmodes with intuitive roles in processing in this task (Fig.
@@
 For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_sc_eig}).
-Perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has a very similar effect on Anti accuracy as perturbations along the sensitivity dimension (Fig. \ref{fig:SC}D, middle).
-This suggests that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
-Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, bottom).
-In the next section, we build on these mechanistic characterizations of each regime by examining their resilience to optogenetic silencing.
+Unlike the sensitivity dimension, these perturbations of connectivity eigenvalue are independent of $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_eig}), e.g.
+\begin{equation}
+\mathbf{z} = \mathbf{z}^*(sW) + \delta \mathbf{v}_{\text{task}}.
+\end{equation}
+Connectivity perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has a very similar effect on Anti accuracy as perturbations along the sensitivity dimension (Fig.
+\ref{fig:SC}D, middle).
+The similar effects of $\mathbf{v}_1(\mathbf{z}^*)$ and $-\mathbf{v}_{\text{task}}$ suggest that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
+Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, right).
+In the next section, we build on these mechanistic characterizations of each regime by examining their resilience to optogenetic inactivation.
 
 \subsection{EPI inferred SC connectivities reproduce results from optogenetic inactivation experiments} \label{results_SC_opt}
-
 During the delay period of this task, the circuit must prepare to execute the correct task according to the presented cue.
-Experimental results from Duan et al.
+The circuit must then maintain a representation of task throughout the delay period, which is important for correct execution of the Anti task.
+Duan et al.
 found that bilateral optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
-\ref{fig:SC_opto}A).
-This suggests that SC maintains a representation of task throughout the delay period, which is important for correct execution of the Anti task.
-Network connectivities inferred by EPI exhibited this same effect in simulation at high optogenetic strengths $\gamma$ (Fig. \ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1.]{figures/fig5/fig5.pdf}
+\ref{fig:SC_opto}A) \cite{duan2019collicular}.
+The distribution of connectivities inferred by EPI exhibited this same effect in simulation at high optogenetic strengths $\gamma$, which reduce the network activities $\mathbf{x}(t)$ by a factor $1 - \gamma$ (Fig.
+\ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.1]{figures/fig5/fig5.pdf}
 \end{center}
 \caption{ 
-\textbf{A}. Experimental effect of delay period silencing on Pro and Anti task accuracy in rats.
-\textbf{B}. Mean and standard error (bars) of task error induced by delay period inactivation of varying optogenetic strength across the EPI distribution.
+\textbf{A}. Mean and standard error (bars) across recording sessions of task error following delay period optogenetic inactivation in rats.
+\textbf{B}. Mean and standard deviation (bars) of task error induced by delay period inactivation of varying optogenetic strength $\gamma$ across the EPI distribution.
 \textbf{C}. (Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.675$.
@@
 (Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.85$.
+Accuracy in Pro (top) and Anti (bottom) task is shown as a percentage.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.85$ versus accuracy in the Pro task across connectivities in the EPI distribution.
@@
 
-The mean increase in Anti error is closest to the experimentally measured value of 7\% at $\gamma = 0.675$ (Fig. \ref{fig:SC_opto}B, black dot).  
-At this level of optogenetic strength, only regime 1 exhibits an increase in Anti error with delay period silencing (Fig. \ref{fig:SC_opto}C, left).
-The connectivities in regime 2 are thus more resilient to delay period silencing during Anti trials than regime 1.
+To examine how connectivity affects response to delay period inactivation, we grouped connectivities of the EPI distribution along the continuum linking regimes 1 and 2 of Section \ref{results_SC}:
+\begin{equation}\label{eq:mode_group}
+Z(sW)  = \{\mathbf{z} ~~\text{if}~\argmin_{y \in Y} |\mathbf{z} - \mathbf{z}^*(y)|_2 = sW, ~~\text{for}~z \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})\},
+\end{equation}
+where $Y = \{-1., -0.75, ..., 1.\}$ are the values of $sW$ for which we calculated the mode.
+$Z(sW)$ is then the set of EPI samples for which the closest mode was $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_regime}).
+In the following analyses, we examine how error, and the influence of connectivity eigenvalue on Anti error change along this continuum of connectivities.
+Obtaining the parameter samples for these analysis with the learned EPI distribution was more than 20,000 times faster than a brute force approach (see Section \ref{methods_sc_group}).
+
+The mean increase in Anti error of the EPI distribution is closest to the experimentally measured value of 7\% at $\gamma = 0.675$ (Fig. \ref{fig:SC_opto}B, black dot).
+At this level of optogenetic strength, regime 1 exhibits an increase in Anti error with delay period silencing (Fig. \ref{fig:SC_opto}C, left), while regime 2 does not.
 In regime 1, greater $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ decrease Anti error (Fig. \ref{fig:SC_opto}C, right).
-In other words, these anticorrelations show that stronger task representations and diagonal amplification make the SC model more resilient to delay period silencing in the Anti task.
-All correlations of connectivity eigenvalue with Anti error degrade in regime 2, where there is no effect of delay period silencing on Anti error (Fig. \ref{fig:SC_opto}C, right).
-
-At about $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero
+In other words, stronger task representations and diagonal amplification make the SC model more resilient to delay period silencing in the Anti task.
+This complements the finding from Duan et al. 2019 \cite{duan2019collicular} that $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ improve Anti accuracy.
+
+At about $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero.
 Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC_opto}D, left).
@@
 Similarities between Pro and Anti trial responses were not present at the experiment-matching level of $\gamma=0.675$ (Fig.
-\ref{fig:SC6} left) and neither was anti-correlation in $p_P$ and $p_{A,\text{opto}}$ (Fig.
+\ref{fig:SC6} left) and neither was anticorrelation in $p_P$ and $p_{A,\text{opto}}$ (Fig.
 \ref{fig:SC6} right).
@@
 We found that at levels of optogenetic strength matching experimental levels of Anti error, only one regime actually exhibited the effect.  
-This suggests that one regime is less resilient to optogenetic perturbation, and perhaps more biologically realistic.
-Finally, we mechanistically characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels.
-\q{sc_summary}{The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum (in 7.36 seconds with the EPI distribution rather than 4.2 days with brute force methods, see Section \ref{methods_SC}).}
-
-%\subsection{Novel hypotheses from EPI} 
-\section{Discussion} 
+This connectivity regime is less resilient to optogenetic perturbation, and perhaps more biologically realistic.
+Finally, we characterized the pathology in Anti error that occurs in both regimes when optogenetic strength is increased to high levels, leading to a mechanistic hypothesis that is experimentally testable.
+\q{sc_summary}{The probabilistic tools afforded by EPI yielded this insight: we identified two regimes and the continuum of connectivities between them by taking gradients of parameter probabilities in the EPI distribution, we identified sensitivity dimensions by measuring the Hessian of the EPI distribution, and we obtained many parameter samples at each step along the continuum at an efficient rate.}
+
+\section{Discussion} \label{discussion}
 In neuroscience, machine learning has primarily been used to reveal structure in neural datasets \cite{paninski2018neural}.
@@
 %The careful steps of \emph{i}.) model design and \emph{ii}.) emergent property definition, are followed by \emph{iii}.) practical inference methods resulting in an opaque characterization of the way model parameters govern computation.
-In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference with respect to computation in neural circuit models (see Section \ref{methods_related}).
-
-\q{discussion}{Methodology for statistical inference in circuit models has evolved considerably in recent years.
-Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but EPI and other recently developed methodology (SNPE \cite{gonccalves2019training}) employ deep learning to improve efficiency and provide deep, flexible distribution approximations.
+In this work, we improve upon parameter inference techniques in theoretical neuroscience with emergent property inference, harnessing deep learning towards parameter inference in neural circuit models (see Section \ref{methods_related}).
+
+\q{discussion_methods}{Methodology for statistical inference in circuit models has evolved considerably in recent years.
+Early work used rejection sampling techniques \cite{beaumont2002approximate, marjoram2003markov,sisson2007sequential}, but EPI and another recently developed methodology \cite{gonccalves2019training} employ deep learning to improve efficiency and provide flexible approximations.
 SNPE has been used for posterior inference of parameters in circuit models conditioned upon exemplar data used to represent computation, but it does not infer parameter distributions that only produce the computation of interest like EPI  (see Section \ref{results_RNN}).
-Finally, we show that EPI has better scaling properties than SNPE when emergent property gradients are tractable (Section \ref{results_RNN}).
-In summary, choice of deep inference technique should consider emergent property complexity and differentiability, dimensionality of parameter space, and the importance of constraining the model behavior predicted by the inferred parameter distribution.
+When strict control over the predictions of the inferred parameters is necessary, EPI uses a constrained optimization technique \cite{loaiza2017maximum} (see Section \ref{methods_AL_opt}) to make inference conditioned on the emergent property possible.}
+
+\q{discussion_gradients}{A key difference between EPI and SNPE, is that EPI uses gradients of the emergent property throughout optimization.
+In Section \ref{results_RNN}, we showed that such gradients confer beneficial scaling properties, but a concern remains that emergent property gradients may be too computationally intensive.
+Even in a case of close biophysical realism with an expensive emergent property gradient, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
+However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
+The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
+In such cases, SNPE, which does not require differentiability of the emergent property, has proved to be a powerful approach \cite{gonccalves2019training}.
+In summary, choice of deep inference technique should consider emergent property complexity and differentiability, dimensionality of parameter space, and the importance of constraining the model behavior predicted by the inferred parameter distribution.}
 
@@
 For example, the global structure captured by EPI revealed two distinct parameter regimes, which had different local structure quantified by the deep probability distribution (see Section \ref{methods_SC}).
-In comparison, Bayesian MCMC is considered a popular approach for capturing global parameter structure \cite{girolami2011riemann}, but there is no variational approximation (the deep probability distribution in EPI), so sensitivity information is not queryable and sampling remains slow after convergence.
+In comparison, bayesian MCMC is considered a popular approach for capturing global parameter structure \cite{girolami2011riemann}, but there is no variational approximation (the deep probability distribution in EPI), so sensitivity information is not queryable and sampling remains slow after convergence.
 Local sensitivity analyses (e.g.
 \cite{raue2009structural}) may be performed independently at individual parameter samples, but these methods alone do not capture the full picture in nonlinear, complex distributions.
-Therefore, deep inference is the only approach yielding an object -- the deep probability distribution -- which produces a wholistic assessment of parameter sensitivity at the local and global level, which were used in this study to make novel insights into theoretical models of neural computation.
-Together, the abilities to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a powerful new method for addressing inverse problems in theoretical neuroscience.}
-
-Even with a high degree of biophysical realism and expensive emergent property gradients, EPI was run successfully on intermediate hub frequency in a 5-neuron subcircuit model of the STG (Section \ref{results_motivating}).
-However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
-The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, SNPE, which does not require differentiability of the emergent property has proved to be a powerful approach \cite{gonccalves2019training}.
+In contrast, deep inference yields a probability distribution that produces a wholistic assessment of parameter sensitivity at the local and global level, which we used in this study to make novel insights into a range of theoretical models.
+Together, the abilities to condition upon emergent properties, the efficient inference algorithm, and the capacity for parameter sensitivity analyses make EPI a useful method for addressing inverse problems in theoretical neuroscience.
 
@@
 \newpage 
-
+\startsupplement
 \section{Methods}
@@
 \end{equation} 
-Entropy is chosen as the normative selection principle to match that of variational Bayesian methods (see Section \ref{methods_ME_EF}).
-However, a key difference is that variational Bayesian methods do not constrain the predictions of their inferred parameter distribution.
+Entropy is chosen as the normative selection principle to match that of variational bayesian methods (see Section \ref{methods_ME_EF}).
+However, a key difference is that variational bayesian methods do not constrain the predictions of their inferred parameter distribution.
 This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
@@
 Second, we describe the special class of deep probability distributions used in EPI called normalizing flows (Section \ref{methods_NF}).
+Then, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}).
 Next, we explain the constrained optimization technique used to solve Equation \ref{eq:opt} (Section \ref{methods_AL_opt}).
 Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
-Finally, we establish the known relationship between maximum entropy distributions and exponential families (Section \ref{methods_ME_EF}), which is used to explain how EPI is a form of variational inference (Section \ref{methods_VI}).
+Finally, we explain how EPI is a form of variational inference (Section \ref{methods_VI}).
 
@@
 Simulation-based inference \cite{cranmer2020frontier} is model parameter inference in the absence of a tractable likelihood function.
-The most prevalent approach to simulation-based inference is approximate Bayesian computation (ABC) \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
+The most prevalent approach to simulation-based inference is approximate bayesian computation (ABC) \cite{beaumont2002approximate}, in which satisfactory parameter samples are kept from random prior sampling according to a rejection heuristic.
 The obtained set of parameters do not have a probabilities, and further insight about the model must be gained from examination of the parameter set and their generated activity.
@@
 
-\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
-To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented Lagrangian method.  
+\subsubsection{Augmented lagrangian optimization}\label{methods_AL_opt}
+To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented lagrangian method.  
 The following objective is minimized:
@@
 
-The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
+The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
 A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
@@
 This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see example in Section \ref{methods_2DLDS}).
-The momentum parameters of the Adam optimizer are reset at the end of each augmented Lagrangian epoch.
+The momentum parameters of the Adam optimizer are reset at the end of each augmented lagrangian epoch.
 
@@
 This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
-The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented Lagrangian epoch.
+The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented lagrangian epoch.
 
@@
 \subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
+To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. \ref{fig:LDS1}A):
 \begin{equation} 
@@
 \end{center}
-\caption{\small (LDS1): \textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
+\caption{\textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
 \textbf{B}.
@@
 \textbf{C}. Entropy throughout the optimization.  
-At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
+At the beginning of each augmented lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
 \textbf{D}. Emergent property moments throughout optimization.  
-At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
+At the beginning of each augmented lagrangian epoch, the emergent property moments adjust closer to their constraints.}
 \label{fig:LDS1}
@@
 Instead, we used EPI to approximate this distribution (Fig.
-S1B).
+\ref{fig:LDS1}B).
 We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
@@
 the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. S2A). 
+$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
 The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
 Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS2/figLDS2.pdf}
-\end{center}
-\caption{\small (LDS2): \textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-\textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\label{fig:LDS2}
-\end{figure}
 
@@
 Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$, we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
-S2B).
+\ref{fig:LDS2}B).
 With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
@@
 Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. S3).
+As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. \ref{fig:LDS3}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{figures/figLDS2/figLDS2.pdf}
+\end{center}
+\caption{\textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+\textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\label{fig:LDS2}
+\end{figure}
 
@@
 \end{center}
-\begin{flushleft}
-\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+\caption{Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(t=0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
 \textbf{A}. Each dimension of the simulated trajectories throughout time.  
 \textbf{B}.  The simulated trajectories in phase space.}
-\end{flushleft}
 \label{fig:LDS3}
@@
 %Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In Bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
-In Bayesian inference, we obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+In bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
+In bayesian inference, we obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
 \begin{equation}
@@
 \label{eq:VI_obj}
-\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right]
-\end{equation}
-
-Comparing this to the Lagrangian objective (without augmentation) of EPI, we see they are the same
+\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right].
+\end{equation}
+
+Comparing this to the lagrangian objective (without augmentation) of EPI, we see they are the same
 \begin{equation}
@@
 Let $X_\alpha \in \mathcal{C}^{|\Phi|}$ be the complex exponential filter bank dot products with the signal $x_\alpha \in \mathbb{R}^{N}$, where $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$.
-The soft-argmax is then calculated using temperature parameter $\beta=100$
-\begin{equation}
-\psi_\alpha = \text{softmax}(\beta|X_\alpha|\odot i),
+The soft-argmax is then calculated using temperature parameter $\beta_{\psi}=100$
+\begin{equation}
+\psi_\alpha = \text{softmax}(\beta_{\psi}|X_\alpha|\odot i),
 \end{equation}
@@
 Intermediate hub frequency, like all other emergent properties in this work, is defined by the mean and variance of the emergent property statistics.
-In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz,(Equation \ref{eq:EP_STG1}) and variance was chosen to be (0.025Hz)$^2$ (Equation \ref{eq:EP_STG2}).
+In this case, we have one statistic, hub neuron frequency, where the mean was chosen to be 0.55Hz,(Equation \ref{eq:EP_STG1}) and variance was chosen to be 0.025$^2$ Hz$^2$ (Equation \ref{eq:EP_STG2}).
 
@@
 \end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 5,000 iterations(see Section \ref{methods_AL_opt}).  
-The optimization converged after five epochs (Fig. S4).
+Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of 5,000 iterations(see Section \ref{methods_AL_opt}).  
+The optimization converged after five epochs (Fig. \ref{fig:STG1}).
 
@@
 \end{center}
-\begin{flushleft}
-\caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
+\caption{EPI optimization of the STG model producing network syncing. 
 \textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
-\end{flushleft}
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented lagrangian epoch.}
 \label{fig:STG1}
@@
 We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
-EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
+EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
 The architecture converged with criteria $N_{\text{test}} = 100$.
@@
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented Lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
+We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
 \ref{fig:RNN}B blue line, and Fig.
@@
 We compared EPI to two alternative simulation-based inference techniques, since the likelihood of these eigenvalues given $\mathbf{z}$ is not available.
-Approximate Bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parameters $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
-Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) is the state-of-the-art ABC method, which leverages SMC techniques to improve sampling speed.
+Approximate bayesian computation (ABC) \cite{beaumont2002approximate} is a rejection sampling technique for obtaining sets of parameters $\mathbf{z}$ that produce activity $\mathbf{x}$ close to some observed data $\mathbf{x}_0$.
+Sequential Monte Carlo approximate bayesian computation (SMC-ABC) is the state-of-the-art ABC method, which leverages SMC techniques to improve sampling speed.
 We ran SMC-ABC with the pyABC package \cite{klinger2018pyabc} to infer RNNs with stable amplification: connectivities having eigenvalues within an $\epsilon$-defined $l$-$2$ distance of
 \begin{equation}\label{eq:stab_amp_x0}
-x_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}.
+\mathbf{x}_0 = \begin{bmatrix} \text{real}(\lambda_1) \\ \lambda^s_1 \end{bmatrix} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix}.
 \end{equation}
 SMC-ABC was run with a uniform prior over $\mathbf{z} \in \left[-1, 1 \right]^{(4N)}$, a population size of 1,000 particles with simulations parallelized over 32 cores, and a multivariate normal transition model.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{figures/figRNN1/figRNN1.pdf}
-\end{center}
-\caption{\small (RNN1): 
-Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
-}
-\label{fig:RNN1}
-\end{figure}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{figures/figRNN1/figRNN1.pdf}
+\end{center}
+\caption{Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
+}
+\label{fig:RNN1}
+\end{figure}
+
 To compare the efficiency of these algorithms for inferring RNN connectivity distributions producing stable amplification, we develop a convergence criteria that can be used across methods.
@@
 \begin{equation}
-d(q_\theta(z)) = |\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}|_2
+d(q_{\bm{\theta}}(\mathbf{z})) = |\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}|_2
 \end{equation}
@@
 \end{center}
-\caption{\small (RNN3): 
-SNPE convergence was enabled by increasing $n_{\text{round}}$, not $n_{\text{atom}}$.
+\caption{SNPE convergence was enabled by increasing $n_{\text{round}}$, not $n_{\text{atom}}$.
 \textbf{A}.
@@
 \end{center}
-\caption{\small (RNN2): Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
+\caption{Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
 \textbf{A}.
@@
 Connectivity ($W_{\text{fit}}$) and input ($\mathbf{h}_{b,\text{fit}}$ and $\mathbf{h}_{c,\text{fit}}$) parameters were fit using the deterministic V1 circuit model \cite{palmigiano2020structure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
-\end{center}
-\caption{\small (V1 1)
-(Left) Simulations for small increases in neuron-type population input.
-Input magnitudes are chosen so that effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
-(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
- }
- \label{fig:V1_1}
-\end{figure}
-
-\begin{figure}
-\caption{\small (V1 2)
-EPI predictive distributions of the sum of squares of each pair of noise parameters.
- }
- \label{fig:V1_2}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
-\end{center}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-\caption{\small (V1 3)
-EPI inferred distribution for $\mathcal{X}(10\text{ Hz})$.
- }
- \label{fig:V1_3}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
-\end{center}
-\end{figure}
-
 \begin{equation}
@@
 \subsubsection{EPI details for the V1 model}
-For EPI in Fig \ref{fig:V1}D-E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
+For EPI in Figures \ref{fig:V1}D-E and \ref{fig:V1_1}, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
 The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
-EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented Lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
+EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
 The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across three random seeds.
+Optimization details are shown in Figure \ref{fig:V1_2}.
+The sums of squares of each pair of parameters are shown for each EPI distribution in Figure \ref{fig:V1_3}.
+
+\begin{figure}[h]
+\caption{EPI inferred distribution for $\mathcal{X}(10\text{ Hz})$.
+ }
+ \label{fig:V1_1}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\caption{EPI optimization $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$
+\textbf{A}. Entropy throughout optimization. 
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 8,000 iterations following the fourth augmented lagrangian epoch.
+ }
+ \label{fig:V1_2}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_4/figV1_4.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\caption{EPI predictive distributions of the sum of squares of each pair of noise parameters.
+ }
+ \label{fig:V1_3}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
+\end{center}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-\caption{\small (V1 4)
-Optimization for V1
+\subsubsection{Testing for the paradoxical effect}
+The paradoxical effect occurs when a populations steady state rate is decreased (or increased) when an increase (decrease) in current is applied to that population \cite{tsodyks1997paradoxical}.
+To see which, if any, populations exhibited a paradoxical effect, we examined responses to changes in input (Fig. \ref{fig:V1_4}).
+Input magnitudes were chosen so that the effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
+Only the P-population exhibited the paradoxical effect at this connectivity $W$ and input $\mathbf{h}$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
+\end{center}
+\caption{(Left) SSSN simulations for small increases in neuron-type population input.
+(Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
  }
  \label{fig:V1_4}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_4/figV1_4.pdf}
-\end{center}
 \end{figure}
@@
 The ability to switch between two separate tasks throughout randomly interleaved trials, or ``rapid task switching," has been studied in rats, and midbrain superior colliculus (SC) has been show to play an important in this computation \cite{duan2015requirement}.
-Neural recordings in SC exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side), which led to the distinction of two functional classes: the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
+Neural recordings in SC exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side), which led to the distinction of two functional classes: the Pro/Contra and Anti/Ipsi neurons \cite{duan2019collicular}.
 Given this evidence, Duan et al. proposed a model with four functionally-defined neuron-type populations: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
@@
 with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.
-These hyperparameter values are motivated by modeling choices and results from \cite{duan2018collicular}.
+These hyperparameter values are motivated by modeling choices and results from \cite{duan2019collicular}.
 
@@
 \end{center}
-\caption{\small (SC1): \textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
+\caption{\textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
 \textbf{B}. Same as A colored by Anti task accuracy.  
@@
 \end{center}
-\caption{\small (SC2):  
+\caption{
 \textbf{A}. Simulations in network regime 1 ($hW_{\text{fixed}} = -1.5$).
@@
 \begin{center}
-\includegraphics[scale=1.1]{figures/figSC3/figSC3.pdf}
-\end{center}
-\caption{\small (SC3):  
+\includegraphics[scale=.9]{figures/figSC3/figSC3.pdf}
+\end{center}
+\caption{
 \textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
-\textbf{B}. Accuracies for connectivity perturbations for increasing $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ (rest shown in Fig. \ref{fig:SC}D).
+\textbf{B}.
+Accuracies for connectivity perturbations when changing $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ ($\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ shown in Fig.
+\ref{fig:SC}D).
 }
@@
 \subsubsection{Task accuracy calculation} \label{methods_sc_acc}
-The accuracies of each task $p_P$ and $p_A$ are calculated as
+The accuracies of each Pro and Anti tasks are calculated as
 \begin{equation}
@@
 \end{equation}
-given that the stimulus is on the left side, where $\Theta$ is the Heaviside step function, and the accuracy is averaged over 200 independent trials.  The Heaviside step function is approximated as
-\begin{equation}
-\Theta(\mathbf{x}) = \text{sigmoid}(\beta \mathbf{x}),
-\end{equation}
-where $\beta = 100$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
-\end{center}
-\caption{\small (SC4):  \textbf{A}. The individual parameters of each mode throughout the two regimes.
-\textbf{B}. The individual sensitivities of parameters of each mode throughout the two regimes.
-}
-\label{fig:SC4}
-\end{figure}
+given that the stimulus is on the left side and $\Theta$ approximates the Heaviside step function.
+Our accuracy calculation only considers one stimulus presentation (Left), since the model is left-right symmetric.
+The accuracy is averaged over 200 independent trials, and the Heaviside step function is approximated as
+\begin{equation}
+\Theta(\mathbf{x}) = \text{sigmoid}(\beta_{\Theta} \mathbf{x}),
+\end{equation}
+where $\beta_{\Theta} = 100$.
+
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.8]{figures/figSC4/figSC4.pdf}
+%\end{center}
+%\caption{\small (SC4):  \textbf{A}. The individual parameters of each mode throughout the two regimes.
+%\textbf{B}. The individual sensitivities of parameters of each mode throughout the two regimes.
+%}
+%\label{fig:SC4}
+%\end{figure}
 
@@
 \end{center}
-\caption{\small (SC5): 
+\caption{
 (Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
@@
 \end{center}
-\caption{\small (SC6): 
+\caption{
 (Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
+Accuracy in Pro (top) and Anti (bottom) task is shown as a percentage.
 (Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.
@@
 \end{equation}
-Throughout optimization, the augmented Lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations (see Section \ref{methods_AL_opt}).  
+Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations (see Section \ref{methods_AL_opt}).  
 The optimization converged after ten epochs (Fig. \ref{fig:SC6}).
@@
 \end{center}
-\caption{\small (SC7): 
+\caption{
 \textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented lagrangian epoch.
 }
@@
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
-We used an augmented Lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
+We used an augmented lagrangian coefficient of $c_0 = 10^{2}$, a batch size $n=100$, and $\beta=2$.
 The distribution was the greatest EPI distribution to converge across 5 random seeds with criteria $N_{\text{test}} = 25$.
@@
 \end{center}
-\caption{\small (SC8): 
+\caption{
 \textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented Lagrangian epoch.
+\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented lagrangian epoch.
 }
@@
 
-\subsubsection{Regime identification with EPI} \label{methods_sc_regime}
-We sought two sets of parameters from $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ that were representative of each regime, so that we could assess their implications on computation.
-For fixed values of $hW$, we hypothesized that there are two modes: one in each regime of greater and lesser $sW$.
-To begin, we found one mode for each regime at $hW_{\text{fixed}} = -1.5$ using 200 steps of gradient ascent of the deep probability distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
-In regime 1, the initialization had positive $sW$, and the initialization had negative $sW$ in regime 2, which led to disparate modes (Fig. \ref{fig:SC4} top).
-These modes were then used as the initialization to find the next mode at $hW_{\text{fixed}} = -1.4$ and so on.
-200 steps of gradient ascent were always taken, and learning rates of $2.5 \times 10^-4$ and $5 \times 10^-4$ were used for regimes 1 and 2, respectively.
-Each of these modes is denoted $\mathbf{z}^*(hW_{\text{fixed}}, r)$ for regime $r \in \{1, 2\}$.
-
+\subsubsection{Mode identification with EPI} \label{methods_sc_regime}
+We found one mode of the EPI distribution for fixed values of $sW$ from 1 to -1 in steps of 0.25.
+To begin, we chose an initial parameter value from 500 parameter samples $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ that had closest $sW$ value to 1.
+We then optimized this estimate of the mode (for fixed $sW$) using probability gradients of the deep probability distribution for 500 steps of gradient ascent with a learning rate of $5 \times 10^{-3}$.
+The next mode (at $sW = 0.75$) was found using the previous mode as the initialization.
+This and all subsequent optimizations used 200 steps of gradient ascent with a learning rate of $1 \times 10^{-3}$, except at $sW = -1$ where a learning rate of $5 \times 10^{-4}$ was used.
+During all mode identification optimizations, the learning rate was reduced by half (decay = 0.5) after every 100 iterations.
+
+\subsubsection{Sample grouping by mode} \label{methods_sc_group}
 For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampling from the EPI distribution.
-Each sample was assigned to the closest mode $\mathbf{z}^*(hW_{\text{fixed}}, r)$.
-Sampling continued until 500 samples were assigned to each mode, which took 7.36 seconds.
-To obtain this many samples for each mode with brute force sampling over the chosen prior, this would take 4.20 days.
+Each sample was assigned to the closest mode $\mathbf{z}^*(sW)$ (Equation \ref{eq:mode_group}).
+Sampling continued until 500 samples were assigned to each mode, which took 2.67 seconds (5.34ms/sample-per-mode).
+It took 9.59 minutes to obtain just 5 samples for each mode with brute force sampling requiring accuracies between 60\% and 90\% in each task (115s/sample-per-mode).
+This corresponds to a sampling speed increase of roughly 21,500 once the EPI distribution has been learned.
 
@@
 This tells us what parameter combination rapid task switching is most sensitive to at this parameter choice in the regime.
-We see that while the modes of each regime gradually converge to similar connectivities at $hW_{\text{fixed}} = -1.05$ (Fig.
-\ref{fig:SC4} top), the sensitivity dimensions remain categorically different throughout the two regimes (Fig.
-\ref{fig:SC4} bottom)
-Only at $hW_{\text{fixed}} = -1.05$ is there a flip in sensitivity from regime 2 to regime 1 (in $\mathbf{v}_1(\mathbf{z}^*)_{sW}$ and $\mathbf{v}_1(\mathbf{z}^*)_{hW}$).
-There is thus some ambiguity regarding the ``regime" of $\mathbf{z}^*(-1.05, 2)$, since the mode is derived from an initialization in regime 2, but has sensitivity like regime 1.
-We can consider this as an intermediate transitionary region of parameter space between the two regimes.
-To emphasize this, $\mathbf{z}^*(-1.05, 1)$ and $\mathbf{z}^*(-1.05, 2)$ have the same color.
 
 \subsubsection{Connectivity eigendecomposition and processing modes} \label{methods_sc_eig}
-To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the symmetric connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same basis vectors $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
+To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same eigenmodes $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
 \ref{fig:SC3}A).
+These eigenvectors are always the same, because the connectivity matrix is symmetric and the model also assumes symmetry across hemispheres, but the eigenvalues of connectivity (or degree of eigenmode amplification) change with $\mathbf{z}$.
 These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} eigenmode - all neurons co-fluctuate, \textit{side} eigenmode - one side dominates the other, \textit{task} eigenmode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.
@@
 
-We are interested in the effect of raising or lowering the amplification of each eigenmode in the connectivity matrix.
+We are interested in the effect of raising or lowering the amplification of each eigenmode in the connectivity matrix by perturbing individual eigenvalues $\lambda$.
 To test this, we calculate the unit vector of changes in the connectivity $\mathbf{z}$ that result from a change in the associated eigenvalues
@@
 So $\mathbf{v}_{\text{a}}$ is the normalized column of A corresponding to eigenmode $a$.
-While perturbations in the sensitivity dimension $\mathbf{v}_1(\mathbf{z}^*)$ adapt with the mode $\mathbf{z}^*$ chosen, perturbations in $\mathbf{v}_a$ for $a \in \{\text{all}, \text{side}, \text{text}, \text{diag}\}$ are invariant to $\mathbf{z}$ (Equation \ref{eq:dzdlambda}).
-
-To understand the connectivity mechanisms that distinguish these two regimes, we perturb connectivity at each mode in dimensions that have well defined roles in processing for the Pro and Anti tasks.
-A convenient property of this connectivity parameterization is that there are $\mathbf{z}$-invariant eigenmodes of connectivity, whose eigenvalues (or degree of amplification) change with $\mathbf{z}$.
-These eigenmodes have intuitive roles in processing in each task, and are accordingly named the \textit{all}, \textit{side}, \textit{task}, and \textit{diag} eigenmodes (see Section \ref{methods_SC}).
-Furthermore, the parameter dimension $\mathbf{v}_a$ ($a \in \{\text{all}, \text{side}, \text{task}, \text{ and diag}\}$) that increases the eigenvalue of connectivity $\lambda_a$ is $\mathbf{z}$-invariant (unlike the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$) and $\mathbf{v}_a \perp \mathbf{v}_{b \neq a}$.
-Thus, by changing the degree of amplification of each processing mode by  perturbing $\mathbf{z}$ along $\mathbf{v}_a$, we can elicit the differentiating properties of the two regimes.
+The parameter dimension $\mathbf{v}_a$ ($a \in \{\text{all}, \text{side}, \text{task}, \text{ and diag}\}$) that increases the eigenvalue of connectivity $\lambda_a$ is $\mathbf{z}$-invariant (Equation \ref{eq:dzdlambda}) and $\mathbf{v}_a \perp \mathbf{v}_{b \neq a}$.
+By perturbing $\mathbf{z}$ along $\mathbf{v}_a$, we can examine how model function changes by directly modulating the connectivity amplification of specific eigenmodes, which having interpretable roles in processing in each task.
 
@@
 \subsubsection{Modeling optogenetic silencing.} \label{methods_sc_opto}
-We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2015requirement}.
+We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2019collicular}.
 During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma \in [0, 1]$

### 2021-05-06 epi/util.py
latest notebook and SC code updates

@@ -188,7 +188,7 @@ def get_max_H_dist(model, epi_df, mu, alpha=0.05, nu=1., check_last_k=None, by_d
     
     return dist, path, best_k
 
-def get_conditional_mode(dist, ind, val, z0=None, lr=1e-6, num_steps=100):
+def get_conditional_mode(dist, ind, val, z0=None, lr=1e-6, num_steps=100, decay=1., decay_steps=100):
     if z0 is None:
         z0 = (dist.nf.lb + dist.nf.ub) / 2.
         z0[ind] = val
@@ -200,6 +200,8 @@ def get_conditional_mode(dist, ind, val, z0=None, lr=1e-6, num_steps=100):
     log_q_zs = [log_q_z]
     
     for k in range(num_steps):
+        if (k!=0 and (np.mod(k, decay_steps)==0)):
+            lr = lr*decay
         print('Finding mode %d/%d.\r' %(k+1, num_steps), end="")
         grad_z = dist._gradient(z).numpy()
         z_np = z.numpy()
@@ -562,6 +564,7 @@ def pairplot(
     s_star=100,
     starred=None,
     c_starred = None,
+    star_marker = '*',
     traj = None,
     fontsize=12,
     figsize=(12, 12),
@@ -649,12 +652,12 @@ def pairplot(
                     if c_starred is None:
                         ax.scatter(
                             starred[:, dim_j], starred[:, dim_i], s=s_star, c='k', 
-                            marker='*', edgecolors="k", linewidths=1.,
+                            marker=star_marker, edgecolors="k", linewidths=1.,
                         )
                     else:
                         ax.scatter(
                             starred[:, dim_j], starred[:, dim_i], s=s_star, c=c_starred, 
-                            marker='*', edgecolors="k", linewidths=1.5,
+                            marker=star_marker, edgecolors="k", linewidths=1.5,
                         )
 
                 if traj is not None:

### 2021-05-10 epi/normalizing_flows.py
indent issue

@@ -534,7 +534,7 @@ class NormalizingFlow(tf.keras.Model):
                             "%.2E s" % ts_time,
                             flush=True,
                         )
-                        else:
+                    else:
                         print(i, "H", H, "loss", loss, "%.2E s" % ts_time, flush=True)
         init_time = time.time() - t1
         opt_df = pd.concat(opt_it_dfs, ignore_index=True)

### 2021-05-10 epi/normalizing_flows.py
time save print

@@ -539,8 +539,10 @@ class NormalizingFlow(tf.keras.Model):
         init_time = time.time() - t1
         opt_df = pd.concat(opt_it_dfs, ignore_index=True)
         opt_df.to_csv(init_path + "opt_data.csv")
+        print("Saving timing...")
         np.savez(os.path.join(init_path, "timing.npz"),
                  init_time=init_time)
+        print("saved!")
         checkpoint.save(file_prefix=init_file)
         return opt_df
 

### 2021-05-10 neural_circuits/LRRNN.py

@@ -232,9 +232,12 @@ def load_ME_EPI_LRRNN(
 
 
 def load_best_SNPE_LRRNN(
-    N, g, K, x0, num_sims=1000, num_batch=200, num_atoms=100, random_seeds=[1, 2, 3]
+    N, g, K, x0, num_sims=1000, num_batch=200, num_atoms=100, random_seeds=[1, 2, 3], gpu=False,
 ):
-    snpe_base_path = os.path.join("data", "snpe")
+    if gpu:
+        snpe_base_path = os.path.join("data", "snpe_gpu")
+    else:
+        snpe_base_path = os.path.join("data", "snpe")
     num_transforms = 3
     # Choose best SNPE
     best_val_prob = None
@@ -284,9 +287,12 @@ def get_simulator(N, g, K):
     return simulator, prior
 
 
-def get_epi_times(optim):
+def get_epi_times(init, optim):
+    init_time = init['init_time']
     iteration = optim["iteration"]
-    return iteration * optim["time_per_it"]
+    iteration = optim["iteration"]
+    times = iteration * optim["time_per_it"]
+    return init_time + times
 
 
 def get_snpe_times(optim, num_sims=None):
@@ -328,11 +334,15 @@ def get_SNPE_conv(N, g, K, x0, eps,
                         num_sims=1000, 
                         num_batch=200, 
                         num_atoms=100,
-                        random_seeds=None):
+                        random_seeds=None,
+                        gpu=False):
    
     conv_times = []
     conv_sims = []
-    snpe_base_path = os.path.join("data", "snpe")
+    if gpu:
+        snpe_base_path = os.path.join("data", "snpe_gpu")
+    else:
+        snpe_base_path = os.path.join("data", "snpe")
     num_transforms = 3
     # Choose best SNPE
     for _i, rs in enumerate(random_seeds):
@@ -368,6 +378,7 @@ def get_SNPE_conv(N, g, K, x0, eps,
         
         assert(num_rounds < args.max_rounds)
         conv_inds = np.nonzero(distances[1:]<eps)[0]
+        print(conv_inds[0])
         conv_time = np.nan if len(conv_inds) == 0 else round_times[conv_inds[0]+1]
         _conv_sims = np.nan if len(conv_inds) == 0 else round_sims[conv_inds[0]+1]
         conv_times.append(conv_time)
@@ -393,8 +404,13 @@ def get_EPI_conv(N, g, K, random_seeds, eps=None):
     conv_sims = []
[…]

### 2021-05-10 neural_circuits/LRRNN.py
lrrnn edits

@@ -16,8 +16,6 @@ from sbi.inference import SNPE, prepare_for_sbi, simulate_for_sbi
 from sbi.utils.get_nn_models import posterior_nn
 
 # EPS = 1e-6
-
-
 def get_W_eigs_np(g, K, feed_noise=False):
     def W_eigs(U, V, noise=None):
         U, V = U[None, :, :], V[None, :, :]
@@ -392,6 +390,7 @@ def get_EPI_conv(N, g, K, random_seeds, eps=None):
     model = LRRNN_setup(N, g, K)
     # Choose max entropy
     epi_df = model.get_epi_df()
+    print(epi_df['path'].unique())
     epi_df["arch_D"] = [row["arch"]["D"] for i, row in epi_df.iterrows()]
     epi_df["c0"] = [row["AL_hps"]["c0"] for i, row in epi_df.iterrows()]
     epi_df["rs"] = [row["arch"]["random_seed"] for i, row in epi_df.iterrows()]
@@ -403,11 +402,11 @@ def get_EPI_conv(N, g, K, random_seeds, eps=None):
     conv_times = []
     conv_sims = []
     for path in paths:
+        print(path)
         _epi_df = epi_df[epi_df['path']==path]
         _row = _epi_df.iloc[0]
         nf = model._df_row_to_nf(_row)
         init_path = nf.get_init_path(_row['init']['mu'], _row['init']['Sigma'])
-        print('init', init_path)
         epi_init = np.load(os.path.join(init_path, "timing.npz"))
         epi_optim = get_epi_optim(model, None, _epi_df, path)
         epi_times = get_epi_times(epi_init, epi_optim)
@@ -426,6 +425,7 @@ def get_EPI_conv(N, g, K, random_seeds, eps=None):
         _conv_sims = np.nan if len(conv_inds) == 0 else epi_batch*iterations[conv_inds[0]]
         conv_times.append(conv_time)
         conv_sims.append(_conv_sims)
+        print(conv_times)
    
     return conv_times, conv_sims
 

### 2021-05-10 written/eLifeRevision/Response.tex

@@
 We thank the reviewers for the excellent and constructive feedback.
-These valid criticisms of the original manuscript impressed upon us the importance of making key improvements to this research project.
-We spent more than half of a year advancing this work according to the reviewer requests and suggestions, and we are excited to present them to you in a dramatically improved revision.
 
 We have made significant changes to our writing and explanation of emergent property inference (EPI), done extensive comparisons of EPI to other parameter inference methods, and have seriously increased the quality and depth of our scientific analyses, which now yield strong theoretical results.
-Our model analyses now focus on analyzing the rich structure of parameter distributions, which the deep probability distributions of EPI make possible.
+Our model analyses now focus on analyzing the structure of parameter distributions, which EPI captures using flexible approximations enabled by deep learning.
 Here, we provide a list of key manuscript updates, and below we explain how these changes address each of the reviewer’s concerns.
@@
 \begin{itemize}
-\item We completely overhauled our presentation of the parameter inference technique emergent property inference (EPI). See response to main concern \#1.
+\item We completely overhauled our presentation of the parameter inference technique emergent property inference (EPI). See response to main concern \#1, especially noted by R1 and R3.
 \item We added an entirely new results section to compare EPI to alternative parameter inference techniques (SMC-ABC and SNPE) as parameter count increases, which demonstrates the advantages of EPI in high dimensional parameter spaces.
-See response to main concern \#2.
+See response to main concern \#2, especially noted by R2 and R3.
 \item EPI is used to understand how noise across neuron type populations governs excitatory variability in a model of primary visual cortex.  
@@
 
-We thank the reviewers for pointing out that our explanation of the method was too narrowly focused and hard to follow. 
+Agreed.
 In the introduction, we present our method by focusing on the key aspect of EPI that differentiates it from other approaches to inverse problems.
@@
 \begin{displayquote}
-Lines 45-56\\
-``\quoting{intro_data_ep}"
+Lines 46-54\\
+``\quoting{intro_data_ep}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 57-67\\
-``\quoting{intro_epi}"
+Lines 55-65\\
+``\quoting{intro_epi}''
 \end{displayquote}
 
-In Section \ref{results_dgm} and \ref{methods_EPI}, we precisely explain the details of EPI, and in Section \ref{methods_AL_opt} we show how it is optimized.
-To improve clarity, we have changed the notation and presentation of emergent properties.
+To improve clarity, we have overhauled the notation and presentation of emergent properties.
 Emergent properties are now denoted with $\mathcal{X}$ to signify an implicit dataset predicated by the emergent property.
-And rather than a vector of first- and second-moment constraints, we present the emergent property more readably as mean and variance constraints on emergent property statistics:
+And rather than a vector of first- and second-moment constraints, we present the emergent property more readably (and equivalently) as mean and variance constraints on emergent property statistics:
 
 \begin{displayquote}
-Line 705 \\
+Line 759 \\
 ``
@@
 \mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\end{equation}"
+\end{equation}''
 \end{displayquote}
+
+Emergent properties and EPI are introduced and explained with this revised formalism in Section \ref{results_dgm}, and we have completely redone the Methods Section \ref{methods_EPI} to improve clarity, precisely explain EPI optimization (Section \ref{methods_AL_opt}), and derive equivalence to established bayesian inference techniques (Section \ref{methods_VI}).
 
@@
 \begin{displayquote}
-Lines 152-154\\
-``\quoting{results_ME}"
+Lines 156-158\\
+``\quoting{results_ME}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 159-174\\
-``\quoting{results_hess}"
+Lines 163-179\\
+``\quoting{results_hess}''
 \end{displayquote}
@@
 The emergent properties of many models in neuroscience are tractably differentiable, four of which we analyze in this manuscript ranging across levels of biological realism, network size, and computational function.
-This trade-off is explained at the end of Section 3.3:
+This trade-off is now explained in the Discussion (Section \ref{discussion}).
 
 \begin{displayquote}
-Lines 240-251\\
-``\quoting{rnn_gradients}"
+Lines 439-451\\
+``\quoting{discussion_gradients}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 788-793\\
-``\quoting{amortization}"
+Lines 828-833\\
+``\quoting{amortization}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 229-236\\
-``\quoting{rnn_constrained}"
+Lines 232-242\\
+``\quoting{rnn_constrained}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 69-74 \\
-``\quoting{intro_results1}"
+Lines 77-81 \\
+``\quoting{intro_results1}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 74-81 \\
-``\quoting{intro_results2}"
+Lines 81-89 \\
+``\quoting{intro_results2}''
 \end{displayquote}
@@
 Finally, the sensitivity dimensions of the SC model identified by EPI facilitated regime characterization through perturbation analyses (Fig.
-4D, Fig.
-5C).
+4D).
 
@@
 \begin{displayquote}
-Lines 302-310 \\
-``\quoting{v1_summary}"
+Lines 301-309 \\
+``\quoting{v1_summary}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 414-419 \\
-``\quoting{sc_summary}"
+Lines 416-419 \\
+``\quoting{sc_summary}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 92-97 \\
-``\quoting{stg_subcircuit}"
+Lines 100-105 \\
+``\quoting{stg_subcircuit}''
 \end{displayquote}
 
-The difference between this model and the STG model of the pyloric subnetwork is emphasized in Section \ref{results_RNN}:
+The difference between this model and the STG model of the pyloric subnetwork is emphasized in Discussion:
 
@@
 \begin{displayquote}
-Lines 246-250 \\
+Lines 444-447 \\
 ``However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
 The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not calculable or easily approximated with differentiable functions."
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.\\
+''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 809-822 \\
-``\quoting{structural_id}"
+Lines 849-862 \\
+``\quoting{structural_id}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 28-36 \\
-``\quoting{intro_first}"
+Lines 28-37 \\
+``\quoting{intro_first}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 986-989 \\
-``\quoting{ME_prior}"
+Lines 1026-1029 \\
+``\quoting{ME_prior}''
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 171 -173 \\
-``This suggests that changes in conductance along the parameter combination described by $\mathbf{v}_2$ will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms."
+Lines 175 -177 \\
+``The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.''
 \end{displayquote}
@@
 \begin{displayquote}
-Section \ref{results_dgm}, Lines 136-140\\
-``\quoting{stg_ep_text1}~\quoting{stg_ep_text2}."
+Section \ref{results_dgm}, Lines 139-143\\
+``\quoting{stg_ep_text1}~\quoting{stg_ep_text2}.''
 \end{displayquote}
@@
 \begin{displayquote}
-Section \ref{results_RNN}, Lines 200-207\\
-``\quoting{rnn_ep_text1}~... \quoting{rnn_ep_text2}~\quoting{rnn_ep_text3}"
+Section \ref{results_RNN}, Lines 204-210\\
+``\quoting{rnn_ep_text1}~... \quoting{rnn_ep_text2}~\quoting{rnn_ep_text3}''
 \end{displayquote}
@@
 \begin{displayquote}
-Section \ref{results_V1}, Lines 281-284\\
-``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}"
+Section \ref{results_V1}, Lines 280-283\\
+``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}''
 \end{displayquote}
@@
 \begin{displayquote}
-Section \ref{results_SC}, Lines 336-339\\
-``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}"
+Section \ref{results_SC}, Lines 338-341\\
+``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}''
 \end{displayquote}
@@
 Most of the brute force models of the STG seek to fit various waveform properties of neurons and relative phases.
-Could the model handle these types of analyses, or would it run into problems due to either needing to specify too many properties or because properties like "number of spikes per burst" are discrete rather than continuous?
+Could the model handle these types of analyses, or would it run into problems due to either needing to specify too many properties or because properties like "number of spikes per burst'' are discrete rather than continuous?
 This isn't fatal, but would be good to consider and/or note explicitly.
@@
 The STG subcircuit model of Gutierrez et al. 2013 is certainly less complex than other models of the STG, yet 5 connected Morris-Lecar neurons is certainly a nontrivial system.
-We clarify why this model is analyzed in Section \ref{results_motivating} instead of more complex STG models when discussing the differences between EPI and SNPE:
+We clarify why this model is analyzed in Section \ref{results_motivating} instead of more complex STG models when discussing the differences between EPI and SNPE in Discussion:
 
 \begin{displayquote}
-Lines 240-251\\
-``\quoting{rnn_gradients}"
+Lines 439-451\\
+``\quoting{discussion_gradients}''
+
 \end{displayquote}
@@
 Thank you for this recommendation.
-We now include a substantial text in discussion devoted to this topic.
+We now include substantial text in discussion devoted to this topic in addition to that quoted in the previous response (R3 concern \#3).
 
 \begin{displayquote}
-Lines 431-454\\
-``\quoting{discussion}"
+Lines 431-438\\
+``\quoting{discussion_methods}''
 \end{displayquote}
@@
 
-Thank you for these references, we have incorporated some where appropriate.
+Thank you for these references. We have incorporated most of them where appropriate.
 
@@
 
+Please see the github repository for this manuscript, which includes code for the EPI method and the analyses that went into each figure.
+
 \end{document}

### 2021-05-10 neural_circuits/LRRNN.py
black style for neural circuit python files

@@ -172,7 +172,6 @@ def LRRNN_setup(N, g, K):
     return model
 
 
-
 def load_ME_EPI_LRRNN(
     N, g, K, mu, random_seeds=None, return_df=False, by_df=False, nu=1.0
 ):
@@ -203,7 +202,15 @@ def load_ME_EPI_LRRNN(
 
 
 def load_best_SNPE_LRRNN(
-    N, g, K, x0, num_sims=1000, num_batch=200, num_atoms=100, random_seeds=[1, 2, 3], gpu=False,
+    N,
+    g,
+    K,
+    x0,
+    num_sims=1000,
+    num_batch=200,
+    num_atoms=100,
+    random_seeds=[1, 2, 3],
+    gpu=False,
 ):
     if gpu:
         snpe_base_path = os.path.join("data", "snpe_gpu")
@@ -281,7 +288,8 @@ def get_snpe_times(optim, num_sims=None):
     round_times = np.concatenate((np.array([0.0]), round_times))
     if num_sims is not None:
         epoch_sims = [[0]] + [
-            num_epochs * [(i + 1) * num_sims] for i, num_epochs in enumerate(epochs_per_round)
+            num_epochs * [(i + 1) * num_sims]
+            for i, num_epochs in enumerate(epochs_per_round)
         ]
         epoch_sims = list(itertools.chain.from_iterable(epoch_sims))
         round_sims = np.concatenate(
@@ -293,13 +301,20 @@ def get_snpe_times(optim, num_sims=None):
     else:
         return epoch_times, round_times
 
-def get_SNPE_conv(N, g, K, x0, eps, 
-                        num_sims=1000, 
-                        num_batch=200, 
-                        num_atoms=100,
-                        random_seeds=None,
-                        gpu=False):
-   
+
+def get_SNPE_conv(
+    N,
+    g,
+    K,
+    x0,
+    eps,
+    num_sims=1000,
+    num_batch=200,
+    num_atoms=100,
+    random_seeds=None,
+    gpu=False,
+):
+
     conv_times = []
     conv_sims = []
     if gpu:
@@ -326,34 +341,37 @@ def get_SNPE_conv(N, g, K, x0, eps,
         else:
             print("Error: no save path %s." % save_path)
             continue
-        
-        epoch_times, round_times, epoch_sims, round_sims = get_snpe_times(optim, num_sims)
-        distances, args = optim['distances'], optim['args']
+
+        epoch_times, round_times, epoch_sims, round_sims = get_snpe_times(
+            optim, num_sims
[…]

### 2021-05-10 neural_circuits/SC_Circuit_4.py

@@ -25,7 +25,7 @@ E_choice = 0.25
 E_light = 0.5
 
 # set constant parameters
-C = 2 
+C = 2
 theta = 0.05
 beta = 0.5
 tau = 0.09
@@ -36,38 +36,39 @@ I_constant = E_constant * tf.ones((T, 1, 1, 4, 1), dtype=DTYPE)
 
 I_Pbias = np.zeros((T, 4))
 I_Pbias[t < T * dt] = np.array([1, 0, 0, 1])
-I_Pbias = I_Pbias[:,None,None,:,None]
+I_Pbias = I_Pbias[:, None, None, :, None]
 I_Pbias = E_Pbias * tf.constant(I_Pbias, dtype=DTYPE)
 
 I_Prule = np.zeros((T, 4))
 I_Prule[t < 1.2] = np.array([1, 0, 0, 1])
-I_Prule = I_Prule[:,None,None,:,None]
+I_Prule = I_Prule[:, None, None, :, None]
 I_Prule = E_Prule * tf.constant(I_Prule, dtype=DTYPE)
 
 I_Arule = np.zeros((T, 4))
 I_Arule[t < 1.2] = np.array([0, 1, 1, 0])
-I_Arule = I_Arule[:,None,None,:,None]
+I_Arule = I_Arule[:, None, None, :, None]
 I_Arule = E_Arule * tf.constant(I_Arule, dtype=DTYPE)
 
 I_choice = np.zeros((T, 4))
 I_choice[t > 1.2] = np.array([1, 1, 1, 1])
-I_choice = I_choice[:,None,None,:,None]
+I_choice = I_choice[:, None, None, :, None]
 I_choice = E_choice * tf.constant(I_choice, dtype=DTYPE)
 
 I_lightL = np.zeros((T, 4))
 I_lightL[np.logical_and(1.2 < t, t < 1.5)] = np.array([1, 1, 0, 0])
-I_lightL = I_lightL[:,None,None,:,None]
+I_lightL = I_lightL[:, None, None, :, None]
 I_lightL = E_light * tf.constant(I_lightL, dtype=DTYPE)
 
 I_lightR = np.zeros((T, 4))
 I_lightR[np.logical_and(1.2 < t, t < 1.5)] = np.array([0, 0, 1, 1])
-I_lightR = I_lightR[:,None,None,:,None]
+I_lightR = I_lightR[:, None, None, :, None]
[…]

### 2021-05-10 epi/batch_norm.py
black style for epi main files

@@ -27,18 +27,14 @@ from tensorflow_probability.python.bijectors import bijector
 
 
 __all__ = [
-    'BatchNormalization',
+    "BatchNormalization",
 ]
 
 
-def _undo_batch_normalization(x,
-                              mean,
-                              variance,
-                              offset,
-                              scale,
-                              variance_epsilon,
-                              name=None):
-  r"""Inverse of tf.nn.batch_normalization.
+def _undo_batch_normalization(
+    x, mean, variance, offset, scale, variance_epsilon, name=None
+):
+    r"""Inverse of tf.nn.batch_normalization.
 
   Args:
     x: Input `Tensor` of arbitrary dimensionality.
@@ -55,22 +51,23 @@ def _undo_batch_normalization(x,
   Returns:
     batch_unnormalized: The de-normalized, de-scaled, de-offset `Tensor`.
   """
-  with tf.name_scope(name or 'undo_batch_normalization'):
-    # inv = tf.rsqrt(variance + variance_epsilon)
-    # if scale is not None:
-    #   inv *= scale
-    # return x * inv + (
-    #     offset - mean * inv if offset is not None else -mean * inv)
-    rescale = tf.sqrt(variance + variance_epsilon)
-    if scale is not None:
-      rescale = rescale / scale
-    batch_unnormalized = x * rescale + (
-        mean - offset * rescale if offset is not None else mean)
-    return batch_unnormalized
+    with tf.name_scope(name or "undo_batch_normalization"):
+        # inv = tf.rsqrt(variance + variance_epsilon)
+        # if scale is not None:
+        #   inv *= scale
+        # return x * inv + (
+        #     offset - mean * inv if offset is not None else -mean * inv)
+        rescale = tf.sqrt(variance + variance_epsilon)
+        if scale is not None:
+            rescale = rescale / scale
+        batch_unnormalized = x * rescale + (
+            mean - offset * rescale if offset is not None else mean
+        )
+        return batch_unnormalized
 
 
 class BatchNormalization(bijector.Bijector):
-  """Compute `Y = g(X) s.t. X = g^-1(Y) = (Y - mean(Y)) / std(Y)`.
+    """Compute `Y = g(X) s.t. X = g^-1(Y) = (Y - mean(Y)) / std(Y)`.
 
   Applies Batch Normalization [(Ioffe and Szegedy, 2015)][1] to samples from a
   data distribution. This can be used to stabilize training of normalizing
[…]

### 2021-05-10 epi/example_eps.py

@@ -98,4 +98,3 @@ def linear2D_freq_np(a11, a12, a21, a22):
         imag_lambda = eig1_i
 
     return real_lambda, imag_lambda
-

### 2021-05-10 epi/models.py

@@ -18,7 +18,7 @@ from epi.util import (
     get_hash,
     set_dir_index,
     get_dir_index,
-    dbg_check, 
+    dbg_check,
 )
 import matplotlib.pyplot as plt
 from matplotlib import animation
@@ -29,6 +29,7 @@ import time
 import os
 
 REAL_NUMERIC_TYPES = (int, float)
+
 
 class Parameter(object):
     """Univariate parameter of a model.
@@ -209,7 +210,7 @@ class Model(object):
         num_units=50,
         elemwise_fn="affine",
         batch_norm=False,
-        bn_momentum=0.,
+        bn_momentum=0.0,
         post_affine=True,
         random_seed=1,
         init_type=None,  # "iso_gauss",
@@ -304,7 +305,7 @@ class Model(object):
 
         # Initialize architecture to gaussian.
         print("Initializing %s architecture." % nf.to_string(), flush=True)
-        if init_type is None or init_type == 'gaussian':
+        if init_type is None or init_type == "gaussian":
             if init_params is None:
                 mu_init = np.zeros((self.D))
                 Sigma = np.zeros((self.D, self.D))
@@ -322,44 +323,56 @@ class Model(object):
                         mu_init[i] = (nf.lb[i] + nf.ub[i]) / 2.0
                         Sigma[i, i] = np.square((nf.ub[i] - nf.lb[i]) / 4)
                 init_params = {"mu": mu_init, "Sigma": Sigma}
-        elif init_type == 'abc':
-            if 'num_keep' in init_params.keys():
-                num_keep = init_params['num_keep']
+        elif init_type == "abc":
+            if "num_keep" in init_params.keys():
+                num_keep = init_params["num_keep"]
             else:
                 num_keep = 200
-            if 'means' in init_params.keys():
-                means = init_params['means']
+            if "means" in init_params.keys():
+                means = init_params["means"]
             else:
-                means = mu[:len(mu)//2]
-            if 'stds' in init_params.keys():
-                stds = init_params['stds']
+                means = mu[: len(mu) // 2]
+            if "stds" in init_params.keys():
+                stds = init_params["stds"]
             else:
-                stds = np.sqrt(mu[len(mu)//2:])
+                stds = np.sqrt(mu[len(mu) // 2 :])
 
             hash_str = get_hash([nf.lb, nf.ub])
             abc_dir = os.path.join("data", "abc")
[…]

### 2021-05-10 epi/normalizing_flows.py

@@ -117,8 +117,10 @@ class NormalizingFlow(tf.keras.Model):
         self._set_num_layers(num_layers)
         self._set_num_units(num_units)
         self._set_elemwise_fn(elemwise_fn)
-        if (self.arch_type == "autoregressive" and self.elemwise_fn == "spline"):
-            raise NotImplementedError("Error: MAF flows with splines are not implemented yet.")
+        if self.arch_type == "autoregressive" and self.elemwise_fn == "spline":
+            raise NotImplementedError(
+                "Error: MAF flows with splines are not implemented yet."
+            )
         self._set_num_bins(num_bins)
         self._set_batch_norm(batch_norm)
         if not self.batch_norm:
@@ -150,8 +152,9 @@ class NormalizingFlow(tf.keras.Model):
                     )
                     self.bijector_fns.append(bijector_fn)
                 elif elemwise_fn == "spline":
-                    bijector_fn = SplineParams(D - num_masked, num_layers, num_units,
-                                               self.num_bins, B=4.)
+                    bijector_fn = SplineParams(
+                        D - num_masked, num_layers, num_units, self.num_bins, B=4.0
+                    )
                     stage = tfb.RealNVP(num_masked=num_masked, bijector_fn=bijector_fn)
                     # make parameters visible to autograd
                     self.bijector_fns.append(bijector_fn._bin_widths)
@@ -163,14 +166,12 @@ class NormalizingFlow(tf.keras.Model):
                 bijector_fn = tfb.AutoregressiveNetwork(
                     params=2, hidden_units=num_layers * [num_units]
                 )
-                stage = tfb.MaskedAutoregressiveFlow(
-                    shift_and_log_scale_fn=bijector_fn
-                )
+                stage = tfb.MaskedAutoregressiveFlow(shift_and_log_scale_fn=bijector_fn)
                 self.bijector_fns.append(bijector_fn)
 
             self.stages.append(stage)
             bijectors.append(stage)
-            #self.shift_and_log_scale_fns.append(shift_and_log_scale_fn)
+            # self.shift_and_log_scale_fns.append(shift_and_log_scale_fn)
             self.bijector_fns.append(bijector_fn)
 
             if i < self.num_stages - 1:
@@ -540,8 +541,7 @@ class NormalizingFlow(tf.keras.Model):
         opt_df = pd.concat(opt_it_dfs, ignore_index=True)
         opt_df.to_csv(init_path + "opt_data.csv")
         print("Saving timing...")
[…]

### 2021-05-10 scripts/LDS_2D/lds_epi.py
black style for script files

@@ -8,15 +8,15 @@ from epi.example_eps import linear2D_freq
 
 # Get random seed.
 parser = argparse.ArgumentParser()
-parser.add_argument('--seed', type=int, default=1)
+parser.add_argument("--seed", type=int, default=1)
 args = parser.parse_args()
 
-print('Running epi on 2D-LDS with hyper parameter random seed %d.' % args.seed)
+print("Running epi on 2D-LDS with hyper parameter random seed %d." % args.seed)
 
-# Define the 2D LDS model parameters.  
+# Define the 2D LDS model parameters.
 # The four entries of the dynamics matrix will be bounded.
-lb = -10.
-ub = 10.
+lb = -10.0
+ub = 10.0
 a11 = Parameter("a11", 1, lb=lb, ub=ub)
 a12 = Parameter("a12", 1, lb=lb, ub=ub)
 a21 = Parameter("a21", 1, lb=lb, ub=ub)
@@ -28,7 +28,7 @@ M = Model("lds_2D", params)
 M.set_eps(linear2D_freq)
 
 # Set the mergent property values
-mu = np.array([0.0, 0.5**2, 2 * np.pi, (0.1 * 2 * np.pi)**2])
+mu = np.array([0.0, 0.5 ** 2, 2 * np.pi, (0.1 * 2 * np.pi) ** 2])
 
 np.random.seed(args.seed)
 num_stages = 3
@@ -36,17 +36,17 @@ num_layers = 2
 num_units = 50
 
 q_theta, opt_data, save_path, failed = M.epi(
-    mu, 
-    arch_type='coupling', 
+    mu,
+    arch_type="coupling",
     num_stages=num_stages,
     num_layers=num_layers,
     num_units=num_units,
     post_affine=True,
-    K = 10, 
-    num_iters=2500, 
+    K=10,
+    num_iters=2500,
     N=500,
-    lr=1e-3, 
-    c0=1e-3, 
+    lr=1e-3,
+    c0=1e-3,
     verbose=True,
     stop_early=True,
     log_rate=100,

### 2021-05-10 scripts/RNN/stab_amp_smc.py

@@ -7,7 +7,8 @@ import pickle
 import tempfile
 import pandas as pd
 import time
-#import matplotlib.pyplot as plt
+
+# import matplotlib.pyplot as plt
 import argparse
 
 import pyabc
@@ -17,11 +18,11 @@ DTYPE = np.float32
 
 # Get random seed.
 parser = argparse.ArgumentParser()
-parser.add_argument('--N', type=int)
-parser.add_argument('--g', type=float, default=0.01)
-parser.add_argument('--K', type=int, default=1)
-parser.add_argument('--eps', type=float, default=0.5)
-parser.add_argument('--rs', type=int, default=1)
+parser.add_argument("--N", type=int)
+parser.add_argument("--g", type=float, default=0.01)
+parser.add_argument("--K", type=int, default=1)
+parser.add_argument("--eps", type=float, default=0.5)
+parser.add_argument("--rs", type=int, default=1)
 args = parser.parse_args()
 
 N = args.N
@@ -30,11 +31,11 @@ K = args.K
 eps = args.eps
 rs = args.rs
 
-sleep_time = N*0.5 + rs*0.05
+sleep_time = N * 0.5 + rs * 0.05
 time.sleep(sleep_time)
 
-print('Running SNPE on RNN conditioned on stable amplification with:')
-print('N = %d, seed=%d' % (N, rs))
+print("Running SNPE on RNN conditioned on stable amplification with:")
+print("N = %d, seed=%d" % (N, rs))
 
 base_path = os.path.join("data", "smc")
 save_dir = "SMC_RNN_stab_amp_N=%d_eps=%.2f_rs=%d" % (N, eps, rs)
@@ -45,14 +46,16 @@ if not os.path.exists(save_path):
 
 if os.path.exists(os.path.join(base_path, save_dir, "optim.pkl")):
     print("SMC optimization already run. Exitting.")
-    exit()    
+    exit()
 
 W_eigs = get_W_eigs_np(g, K)
+
+
 def model(parameter):
-    u1 = np.array([parameter["U%d1" % i] for i in range(1, N+1)])
-    u2 = np.array([parameter["U%d2" % i] for i in range(1, N+1)])
-    v1 = np.array([parameter["V%d1" % i] for i in range(1, N+1)])
-    v2 = np.array([parameter["V%d2" % i] for i in range(1, N+1)])
[…]

### 2021-05-10 scripts/RNN/stab_amp_snpe.py

@@ -16,18 +16,18 @@ DTYPE = np.float32
 
 # Get random seed.
 parser = argparse.ArgumentParser()
-parser.add_argument('--N', type=int)
-parser.add_argument('--num_sims', type=int, default=1000)
-parser.add_argument('--num_batch', type=int, default=50)
-parser.add_argument('--num_transforms', type=int, default=3)
-parser.add_argument('--num_atoms', type=int, default=10)
-parser.add_argument('--g', type=float, default=0.01)
-parser.add_argument('--K', type=int, default=1)
-parser.add_argument('--max_rounds', type=int, default=50)
-parser.add_argument('--persist_rounds', type=int, default=2)
-parser.add_argument('--min_epochs', type=int, default=10)
-parser.add_argument('--stop_distance', type=float, default=None)
-parser.add_argument('--rs', type=int, default=1)
+parser.add_argument("--N", type=int)
+parser.add_argument("--num_sims", type=int, default=1000)
+parser.add_argument("--num_batch", type=int, default=50)
+parser.add_argument("--num_transforms", type=int, default=3)
+parser.add_argument("--num_atoms", type=int, default=10)
+parser.add_argument("--g", type=float, default=0.01)
+parser.add_argument("--K", type=int, default=1)
+parser.add_argument("--max_rounds", type=int, default=50)
+parser.add_argument("--persist_rounds", type=int, default=2)
+parser.add_argument("--min_epochs", type=int, default=10)
+parser.add_argument("--stop_distance", type=float, default=None)
+parser.add_argument("--rs", type=int, default=1)
 args = parser.parse_args()
 
 N = args.N
@@ -45,13 +45,18 @@ rs = args.rs
 np.random.seed(rs)
 torch.manual_seed(rs)
 
-print('\n\nRunning SNPE on RNN conditioned on stable amplification with:')
-print("N = %d, \nnum_sims = %d, \nnum_batch = %d, \nnum_transforms = %d, \nnum_atoms = %d, \ng=%.4f, \nK=%d, \nseed=%d\n\n"  \
-      % (N, num_sims, num_batch, num_transforms, num_atoms, g, K, rs), flush=True)
+print("\n\nRunning SNPE on RNN conditioned on stable amplification with:")
+print(
+    "N = %d, \nnum_sims = %d, \nnum_batch = %d, \nnum_transforms = %d, \nnum_atoms = %d, \ng=%.4f, \nK=%d, \nseed=%d\n\n"
[…]

### 2021-05-10 scripts/SSSN/SSSN_epi.py

@@ -3,18 +3,23 @@ import argparse
 import numpy as np
 import tensorflow as tf
 from epi.models import Parameter, Model
-from neural_circuits.SSSN import SSSN_sim, SSSN_sim, load_SSSN_variable, get_stddev_sigma
+from neural_circuits.SSSN import (
+    SSSN_sim,
+    SSSN_sim,
+    load_SSSN_variable,
+    get_stddev_sigma,
+)
 
 # Parse script command-line parameters.
 parser = argparse.ArgumentParser()
-parser.add_argument('--alpha', type=str, default='E') # neuron type
-parser.add_argument('--ind', type=int, default=62) # neuron type
-parser.add_argument('--lim', type=float, default=0.005) # neuron type
-parser.add_argument('--sE_mean', type=float, default=5.) # neuron type
-parser.add_argument('--sE_std', type=float, default=0.25) # neuron type
-parser.add_argument('--beta', type=float, default=4.) # aug lag hp
-parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
-parser.add_argument('--random_seed', type=int, default=1)
+parser.add_argument("--alpha", type=str, default="E")  # neuron type
+parser.add_argument("--ind", type=int, default=62)  # neuron type
+parser.add_argument("--lim", type=float, default=0.005)  # neuron type
+parser.add_argument("--sE_mean", type=float, default=5.0)  # neuron type
+parser.add_argument("--sE_std", type=float, default=0.25)  # neuron type
+parser.add_argument("--beta", type=float, default=4.0)  # aug lag hp
+parser.add_argument("--logc0", type=float, default=0.0)  # log10 of c_0
+parser.add_argument("--random_seed", type=int, default=1)
 args = parser.parse_args()
 
 alpha = args.alpha
@@ -23,7 +28,7 @@ lim = args.lim
 sE_mean = args.sE_mean
 sE_std = args.sE_std
 beta = args.beta
-c0 = 10.**args.logc0
+c0 = 10.0 ** args.logc0
 random_seed = args.random_seed
 
 if lim is not None:
@@ -35,19 +40,19 @@ if lim is not None:
         raise NotImplementedError()
 
[…]

### 2021-05-10 scripts/STG_Circuit/stg_epi.py

@@ -10,30 +10,30 @@ DTYPE = tf.float32
 
 # Parse script command-line parameters.
 parser = argparse.ArgumentParser()
-parser.add_argument('--freq', type=float, default=0.55) # frequency for mu
-parser.add_argument('--mu_std', type=float, default=0.05) # std in mu constraint
-parser.add_argument('--beta', type=float, default=4.) # aug lag hp
-parser.add_argument('--logc0', type=float, default=0.) # log10 of c_0
-parser.add_argument('--random_seed', type=int, default=1)
+parser.add_argument("--freq", type=float, default=0.55)  # frequency for mu
+parser.add_argument("--mu_std", type=float, default=0.05)  # std in mu constraint
+parser.add_argument("--beta", type=float, default=4.0)  # aug lag hp
+parser.add_argument("--logc0", type=float, default=0.0)  # log10 of c_0
+parser.add_argument("--random_seed", type=int, default=1)
 args = parser.parse_args()
 
 freq = args.freq
 mu_std = args.mu_std
 beta = args.beta
-c0 = 10.**args.logc0
+c0 = 10.0 ** args.logc0
 random_seed = args.random_seed
 
-g_el_lb = 4.
+g_el_lb = 4.0
 
-sigma_I = 1.e-12
-#sleep_dur = np.abs(args.logc0) + random_seed/5. + beta/3.
-#print('short stagger sleep of', sleep_dur, flush=True)
-#time.sleep(sleep_dur)
+sigma_I = 1.0e-12
+# sleep_dur = np.abs(args.logc0) + random_seed/5. + beta/3.
+# print('short stagger sleep of', sleep_dur, flush=True)
+# time.sleep(sleep_dur)
 
 # 1. Specify the V1 model for EPI.
-D = 2 
-g_el = Parameter("g_el", 1, lb=g_el_lb, ub=8.)
-g_synA = Parameter("g_synA", 1, lb=0.01, ub=4.)
+D = 2
+g_el = Parameter("g_el", 1, lb=g_el_lb, ub=8.0)
+g_synA = Parameter("g_synA", 1, lb=0.01, ub=4.0)
 
 # Define model
 name = "STG_sigmaI=%.2E" % sigma_I
@@ -41,13 +41,11 @@ parameters = [g_el, g_synA]
 model = Model(name, parameters)
 
 # Emergent property values.
-mu = np.array([freq, mu_std**2])
+mu = np.array([freq, mu_std ** 2])
 
[…]

### 2021-05-10 tests/test_models.py
black style for test files

@@ -107,7 +107,7 @@ def test_epi():
     M = Model("lds_2D", params)
     M.set_eps(linear2D_freq)
     q_theta, opt_data, epi_path, failed = M.epi(
-        mu, num_iters=100, K=1, save_movie_data=True, log_rate=10, 
+        mu, num_iters=100, K=1, save_movie_data=True, log_rate=10,
     )
     z = q_theta(50)
     g = q_theta.plot_dist(z)
@@ -121,10 +121,10 @@ def test_epi():
         mu, num_iters=100, K=1, save_movie_data=True
     )
 
-    print('epi_path', epi_path)
+    print("epi_path", epi_path)
 
     epi_df = M.get_epi_df()
-    epi_df_row = epi_df[epi_df['iteration']==100].iloc[0]
+    epi_df_row = epi_df[epi_df["iteration"] == 100].iloc[0]
     q_theta = M.get_epi_dist(epi_df_row)
 
     opt_data_filename = os.path.join(epi_path, "opt_data.csv")
@@ -156,8 +156,13 @@ def test_epi():
     M = Model("lds", params)
     M.set_eps(linear2D_freq)
     q_theta, opt_data, epi_path, _ = M.epi(
-        mu, K=2, num_iters=100, stop_early=True, verbose=True, 
-        save_movie_data=True, log_rate=10,
+        mu,
+        K=2,
+        num_iters=100,
+        stop_early=True,
+        verbose=True,
+        save_movie_data=True,
+        log_rate=10,
     )
     M.epi_opt_movie(epi_path)
 
@@ -171,32 +176,41 @@ def test_epi():
     assert np.sum(z[:, 3] > 0.0) == 0
     assert np.sum(1 - np.isfinite(z)) == 0
 
-    print('DOING ABC NOW')
+    print("DOING ABC NOW")
     # Need finite support for ABC
-    a11 = Parameter("a11", 1, -10., 10.)
-    a12 = Parameter("a12", 1, -10., 10.)
-    a21 = Parameter("a21", 1, -10., 10.)
-    a22 = Parameter("a22", 1, -10., 10.)
+    a11 = Parameter("a11", 1, -10.0, 10.0)
+    a12 = Parameter("a12", 1, -10.0, 10.0)
+    a21 = Parameter("a21", 1, -10.0, 10.0)
+    a22 = Parameter("a22", 1, -10.0, 10.0)
     params = [a11, a12, a21, a22]
[…]

### 2021-05-19 written/eLifeRevision/Bittner_eLife2020.tex
John's edits incorporated

@@
 \thispagestyle{plain}
-{\Large Interrogating theoretical models of neural computation with deep inference} \\
+{\Large Interrogating theoretical models of neural computation with emergent property inference} \\
 Sean R. Bittner$^{1}$, Agostina Palmigiano$^{1}$, Alex T. Piet$^{2,3,4}$, Chunyu A. Duan$^{5}$, Carlos D. Brody$^{2,3,6}$, \\
@@
 Such models are valuable when they give rise to an experimentally observed phenomenon -- whether behavioral or a pattern of neural activity -- and thus can offer insights into neural computation.
-The operation of these circuits, like all models, critically depends on the choices of model parameters.
+The operation of these circuits, like all models, critically depends on the choice of model parameters.
 A key step is then to identify the model parameters consistent with observed phenomena: to solve the inverse problem.
-To solve challenging inverse problems in neuroscience, statistical inference techniques have been used to infer parameters distributions most likely to have produced neural datasets.
-In this work, we present a novel technique, emergent property inference (EPI), that brings the power and versatility of the modern probabilistic modeling toolkit to theoretical neuroscience.
-When theorizing circuit models, scientists predominantly focus on reproducing computational properties rather than a particular dataset.
-Our method uses deep neural networks to learn parameter distributions with complex structure that produce specific computational properties in circuit models.
+In this work, we present a novel technique, emergent property inference (EPI), that brings the modern probabilistic modeling toolkit to theoretical neuroscience.
+When theorizing circuit models, theoreticians predominantly focus on reproducing computational properties rather than a particular dataset.
+Our method uses deep neural networks to learn parameter distributions with these computational properties.
 This methodology is introduced through a motivational example inferring conductance parameters in a circuit model of the stomatogastric ganglion.
@@
 In a model of primary visual cortex, we discovered how connectivity with multiple inhibitory subtypes shapes variability in the excitatory population.
-Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials, mechanistically characterized each regime using probabilistic tools afforded by EPI, and found conditions where these circuit models reproduce results from optogenetic silencing experiments.
+Finally, in a model of superior colliculus, we identified and characterized two distinct regimes of connectivity that facilitate switching between opposite tasks amidst interleaved trials,  characterized each regime via insights afforded by EPI, and found conditions where these circuit models reproduce results from optogenetic silencing experiments.
 Beyond its scientific contribution, this work illustrates the variety of analyses possible once deep learning is harnessed towards solving theoretical inverse problems.
@@
 The function of this system is governed by the choice of model \emph{parameters}, which when configured in a particular way, give rise to a measurable signature of a computation.   
-The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about the space, manifold, or distribution of parameters that give rise to it?
+The work of analyzing a model then requires solving the inverse problem: given a computation of interest, how can we reason about the distribution of parameters that give rise to it?  
 The inverse problem is crucial for reasoning about likely parameter values, uniquenesses and degeneracies, and predictions made by the model \cite{gutenkunst2007universally, erguler2011practical, mannakee2016sloppiness}.}
@@
 First, often we directly use an \emph{exemplar dataset}: a collection of samples that express the computation of interest, this data being gathered either experimentally in the lab or from a computer simulation.
-While in some sense the best choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
+Though a natural choice given its connection to experiment \cite{paninski2018neural}, some drawbacks exist: these data are well known to have features irrelevant to the computation of interest \cite{niell2010modulation, saleem2013integration, musall2019single}, confounding inferences made on such data.
 Related to this point, use of a conventional dataset encourages conventional data likelihoods or loss functions, which focus on some global metric like squared error or marginal evidence, rather than the computation itself.}
@@
 
-The parameter distributions producing a computation may be thin, curved, bent, or multimodal along various parameter axes and combinations.
-It is by capturing and quantifying this complex structure that EPI offers scientific insight.
+The parameter distributions producing a computation may be curved or multimodal along various parameter axes and combinations.
+It is by quantifying this complex structure that EPI offers scientific insight.
 Traditional approximation families (e.g. mean-field or mixture of gaussians) are limited in the distributional structure they may learn.
-To address such restrictions on expressivity, major advancements in machine learning have enabled the use of deep probability distributions as flexible approximating families for such complicated distributions \cite{rezende2015variational, papamakarios2019normalizing} (see Section \ref{methods_NF}).
+To address such restrictions on expressivity, advances in machine learning have used deep probability distributions as flexible approximating families for such complicated distributions \cite{rezende2015variational, papamakarios2019normalizing} (see Section \ref{methods_NF}).
 However, the adaptation of deep probability distributions to the problem of theoretical circuit analysis requires recent developments in deep learning for  constrained optimization \cite{loaiza2017maximum}, and architectural  choices for efficient and expressive deep generative modeling \cite{dinh2017density, kingma2018glow}.
@@
 \q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
-Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2019collicular}.
+Finally, we investigated the possible connectivities of a superior colliculus model that allow execution of different tasks on interleaved trials \cite{duan2019collicular}.
 EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
@@
 An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
-This emergent property is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
-Our notion of intermediate hub frequency is not strictly 0.55Hz, but also moderate deviations of this frequency between the fast (.35Hz) and slow (.68Hz) frequencies.
+This emergent property (EP) is shown in Figure \ref{fig:STG}C at an average frequency of 0.55Hz.
+To be precise, we define intermediate hub frequency not strictly as 0.55Hz, but frequencies of moderate deviation from 0.55Hz between the fast (.35Hz) and slow (.68Hz) frequencies.
 %, which are quantified in the emergent property with variance {0.025$^2$Hz$^2$}.
@@
 
-\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
+\subsection{Emergent property inference via deep generative models} \label{results_dgm}
 Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions \cite{rezende2015variational, papamakarios2019normalizing}.
@@
 \end{equation}
-In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is constrained over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
-This expectation over $\mathbf{x}$ is taken over any randomness in initial state ($\mathbf{x}(t = 0)$) and noise (e.g. $d\mathbf{B}$ in Equation \ref{eq:STG_model1}).
-Formally, the emergent property is the collection of these two constraints.
-In general, an emergent property is a collection of constraints on statistical moments that together define the computation.
+In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is an expectation over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
+We define the EP as the collection of these two constraints.
+In general, an emergent property is a collection of constraints on statistical moments that together define the computation of interest.
 
@@
 an isotropic gaussian) through a deep neural network with weights and biases $\bm{\theta}$ to parameters $\mathbf{z} = g_{\bm{\theta}}(\mathbf{z}_0)$ of a suitably complicated distribution (see Section \ref{methods_NF} for more details).
-\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which is the same choice commonly made in variational bayesian methods (see Section \ref{methods_VI}).}
+\q{results_ME}{Many distributions in $\mathcal{Q}$ will respect the emergent property constraints, so we select the most random (highest entropy) distribution, which also means this approach is equivalent to bayesian variational inference (see Section \ref{methods_VI}).}
 %\cite{jaynes1957information, elsayed2017structure, loaiza2017maximum, savin2017maximum}
@@
 As probability in the EPI distribution decreases away from the mode of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ (Fig. \ref{fig:STG}E yellow star), the emergent property deteriorates.
-Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ change little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
+Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ changes little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
 In contrast, if $\mathbf{z}$ is perturbed along a dimension with strongly decreasing $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive} \cite{raue2009structural, raman2017delineating}.
@@
 The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
-Importantly, once an EPI distribution has been learned, the modes and Hessians of the distribution can be measured with trivial computation (see Section \ref{methods_NF}).}
+Importantly and unlike alternative techniques, once an EPI distribution has been learned, the modes and Hessians of the distribution can be measured with trivial computation (see Section \ref{methods_NF}).}
 
@@
 where $U = \begin{bmatrix} \mathbf{U}_1 & \mathbf{U}_2 \end{bmatrix} + g \chi^{(U)}$, $V = \begin{bmatrix} \mathbf{V}_1 & \mathbf{V}_2 \end{bmatrix} + g\chi^{(V)}$, $\mathbf{U}_1 \mathbf{U}_2, \mathbf{V}_1, \mathbf{V}_2 \in \left[-1, 1 \right]^N$, and $\chi^{(U)}_{i,j}, \chi^{(V)}_{i,j} \sim \mathcal{N}(0, 1)$.
-We infer connectivity parameters $\mathbf{z} = \left[\mathbf{U}_1^\top, \mathbf{U}_2^\top, \mathbf{V}_1^\top, \mathbf{V}_2^\top \right]^\top$   that produce stable amplification.
+We infer connectivity parameters $\mathbf{z} = \left[\mathbf{U}_1, \mathbf{U}_2, \mathbf{V}_1, \mathbf{V}_2 \right]$   that produce stable amplification.
 \q{rnn_ep_text1}{Two conditions are necessary and sufficient for RNNs to exhibit stable amplification \cite{bondanelli2020coding}:  $\text{real}(\lambda_1) < 1$ and
@@
 For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
-Sequential Monte Carlo approximate bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
-Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplary dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
+Sequential Monte Carlo approximate bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions (see Section \ref{methods_related}).
+Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplar dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
 To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
@@
 This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient-based and sampling-based estimators (scaling and speed versus generality).
-Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
+Since gradients of the emergent property are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
 If the emergent property gradient is efficiently calculated, EPI is a clear choice for inferring high dimensional parameter distributions. 
@@
 \textbf{E}.
-(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz})$.
 Each black dot shows the mode at each $\sigma_P$.
@@
 \textbf{F}.
-The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
+The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz})$.
 }
@@
 \begin{split}
-\mathcal{X}(5\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{ Hz}~~~~~~~\mathcal{X}(10\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{ Hz}  \\
+\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\
  \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
@@
 \q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
-In the following analyses, we select 1 Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
-
-First, we ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ producing E-population variability around 5 Hz (Fig. \ref{fig:V1}D).
+In the following analyses, we select 1Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
+
+First, we ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ producing E-population variability around 5Hz (Fig. \ref{fig:V1}D).
 From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
 \ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
-Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the, unexpectedly, high variability in those dimensions (Fig.
+Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the unexpectedly high variability in those dimensions (Fig.
 \ref{fig:V1}D, bottom-right).
-Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5 Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
+Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
 
 Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
-This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E and \ref{fig:V1_1}).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
+This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{Hz})$ (Fig. \ref{fig:V1}E and \ref{fig:V1_1}).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
 \ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
@@
 In a simplified model ($\tau = \tau_{\text{noise}}$), it can be shown that surfaces of equal variance are ellipsoids as a function of $\bm{\sigma}$ (see Section \ref{methods_V1_complexity}).
-Nevertheless, the sensitive and degenerate parameters are challenging to predict mathematically, since the covariance matrix depends on the steady-state solution of the network \cite{hennequin2018dynamical,Gardiner2009}, and terms in the covariance expression increase quadratically with each additional neuron-type population (see also Section \ref{methods_V1_complexity}).
-By pointing out this mathematical complexity, we emphasize the value of streamlined methods for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
-While we have just shown that EPI can be used to investigate fundamental aspects of sensory computation, in the next two sections, we use the probabilistic tools of EPI to identify and characterize two distinct parametric regimes of a neural circuit executing a computation, and then relate these insights to behavioral experiments.
+Nevertheless, the sensitive and degenerate parameters are intractable to predict mathematically, since the covariance matrix depends on the steady-state solution of the network \cite{hennequin2018dynamical,Gardiner2009}, and terms in the covariance expression increase quadratically with each additional neuron-type population (see also Section \ref{methods_V1_complexity}).
+By pointing out this mathematical complexity, we emphasize the value of EPI for gaining understanding about theoretical models when mathematical analysis becomes onerous or impractical.}
 
@@
 In Duan et al. 2019, a computationally intensive procedure was used to obtain a set of 373 connectivity parameters that qualitatively reproduced these optogenetic inactivation results.
-To build upon the insights of this previous work, we use the probabilistic tools afforded by EPI to identify and mechanistically characterize two linked, yet distinct regimes of rapid task switching connectivity.
+To build upon the insights of this previous work, we use the probabilistic tools afforded by EPI to identify and characterize two linked, yet distinct regimes of rapid task switching connectivity.
 
@@
 The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, bottom-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
-This parameter distribution has rich structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
+This parameter distribution has rich structure that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
 Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8}).
@@
 \ref{fig:SC}C red/purple arrows).
-Note that the directionality of these sensitivity dimensions at $\mathbf{z}^*(sW)$ changes distinctly with $sW$, and are perpendicular to the flat dimensions of the EPI distribution that preserve rapid task switching.
+Note that the directionality of these sensitivity dimensions at $\mathbf{z}^*(sW)$ changes distinctly with $sW$, and are perpendicular to the robust dimensions of the EPI distribution that preserve rapid task switching.
 These two directionalities of sensitivity motivate the distinction of connectivity into two regimes, which produce different types of responses in the Pro and Anti tasks (Fig. \ref{fig:SC2}).
@@
 For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_sc_eig}).
-Unlike the sensitivity dimension, these perturbations of connectivity eigenvalue are independent of $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_eig}), e.g.
+Unlike the sensitivity dimension, the dimensions $\mathbf{v}_a$ that perturb isolated connectivity eigenvalues $\lambda_a$ for $a \in \{\text{all}, \text{side}, \text{task}, \text{diag}\}$ are independent of $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_eig}), e.g.
 \begin{equation}
@@
 \ref{fig:SC}D, middle).
-The similar effects of $\mathbf{v}_1(\mathbf{z}^*)$ and $-\mathbf{v}_{\text{task}}$ suggest that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
+The similar effects of perturbations along the sensitivity dimension $\mathbf{v}_1(\mathbf{z}^*)$ and reduction of task eigenvalue (via perturbations along $-\mathbf{v}_{\text{task}}$) suggest that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
 Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, right).
@@
 
-At about $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero.
+At roughly $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero.
 Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC_opto}D, left).
 We interpreted these similarities to suggest that delay period inactivation at this saturated level flips the internal representation of task (from Anti to Pro) in the circuit model.
-This would explain why the Anti error saturates at 50\%: the average Anti accuracy in EPI inferred connectivities is $75\%$, but is 25\% when the internal representation is flipped during delay period silencing.
+A flipped task representation would explain why the Anti error saturates at 50\%: the average Anti accuracy in EPI inferred connectivities is $75\%$, but is 25\% when the internal representation is flipped during delay period silencing.
 This hypothesis prescribes a model of Anti accuracy during delay period silencing of $p_{A,\text{opto}} = 100\% - p_P$, which is fit closely across both regimes of the EPI inferred connectivities (Fig.
@@
 burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
-In such cases, SNPE, which does not require differentiability of the emergent property, has proved to be a powerful approach \cite{gonccalves2019training}.
+In such cases, SNPE, which does not require differentiability of the emergent property, has proven useful \cite{gonccalves2019training}.
 In summary, choice of deep inference technique should consider emergent property complexity and differentiability, dimensionality of parameter space, and the importance of constraining the model behavior predicted by the inferred parameter distribution.}
 
-In this paper, we prove out the value of deep inference for parameter sensitivity analyses at both the local and global level.
+In this paper, we demonstrate the value of deep inference for parameter sensitivity analyses at both the local and global level.
 With these techniques, flexible deep probability distributions are optimized to capture global structure by approximating the full distribution of suitable parameters.

### 2021-05-20 written/eLifeRevision/Bittner_eLife2020.tex
V1 fig and other captions

@@
 Sequential Monte Carlo approximate bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions (see Section \ref{methods_related}).
-Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplar dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
+Unlike EPI, these statistical inference techniques do not constrain the predictions of the inferred distribution, so they were run by conditioning on an exemplar dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
 To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).
@@
 \begin{split}
-\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{ Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\
+\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\
  \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
@@
 \textbf{D}.
-Mean and standard error ($N_{\text{test}}$ = 25) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
+Mean and standard error ($N_{\text{test}}$ = 25, bars not visible) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
 }
@@
 
-Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{ Hz}, 25\text{ Hz}\right]$.
+Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{Hz}, 25\text{Hz}\right]$.
 Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
@@
 \begin{figure}[h]
-\caption{EPI inferred distribution for $\mathcal{X}(10\text{ Hz})$.
+\caption{EPI inferred distribution for $\mathcal{X}(10\text{Hz})$.
  }
@@
 \begin{figure}[h]
-\caption{EPI optimization $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$
+\caption{EPI optimization $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$
 \textbf{A}. Entropy throughout optimization. 
@@
 \caption{
-\textbf{A}. Simulations in network regime 1 ($hW_{\text{fixed}} = -1.5$).
-\textbf{B}. Simulations in network regime 2 ($hW_{\text{fixed}} = -1.5$) .
+\textbf{A}. Simulations in network regime 1: $\mathbf{z}^*(sW=-0.75)$.
+\textbf{B}. Simulations in network regime 2: $\mathbf{z}^*(sW=0.75)$.
 }
@@
 \end{center}
-\caption{
+\caption{EPI optimization of the SC model producing rapid task switching.
 \textbf{A}. Entropy throughout optimization. 
@@
 \caption{
-\textbf{A}. Entropy throughout optimization. 
-\textbf{B}. The emergent property statistic means and variances converge to their constraints at 20,000 iterations following the tenth augmented lagrangian epoch.
+\textbf{A}.Rapid task switching SC connectivities obtained from random sampling.
+\textbf{B}. Task accuracies of the inferred distributions from random sampling (top) and EPI (bottom).
 }

### 2021-05-20 written/eLifeRevision/Bittner_eLife2020.tex
RNN update

@@
 \ref{fig:RNN}C-D blue).
+It was fastest to initialize the EPI distribution on a Tesla V100 GPU, and then subsequently optimize it on a CPU with 32 cores.
+EPI timing measurements accounted for this initialization period.
 
@@
 As in SMC-ABC, we ran SNPE with $\mathbf{x}_0 = \mu$.
-All SNPE optimizations were run for a limit of 1.5 days on a Tesla V100 GPU, or until two consecutive rounds resulted in a validation log probability lower than the maximum observed for that random seed.
+All SNPE optimizations were run for a limit of 1.5 days, or until two consecutive rounds resulted in a validation log probability lower than the maximum observed for that random seed.
+It was always faster to run SNPE on a CPU with 32 cores rather than on a Tesla V100 GPU.
 

### 2021-05-20 written/eLifeRevision/Response.tex

@@
 We thank the reviewers for the excellent and constructive feedback.
-
-We have made significant changes to our writing and explanation of emergent property inference (EPI), done extensive comparisons of EPI to other parameter inference methods, and have seriously increased the quality and depth of our scientific analyses, which now yield strong theoretical results.
-Our model analyses now focus on analyzing the structure of parameter distributions, which EPI captures using flexible approximations enabled by deep learning.
-Here, we provide a list of key manuscript updates, and below we explain how these changes address each of the reviewer’s concerns.
-
+The reviews contained broad support for this work, but with significant calls for improvements, especially around clarity of exposition and the thoroughness of the evaluation (comparison to existing methods, more depth in scientific analyses, etc).
+We strongly agreed and thus committed a significant period of time to improving these areas.
 List of key manuscript improvements:
@@
 \item EPI is used to understand how noise across neuron type populations governs excitatory variability in a model of primary visual cortex.
-Here, EPI yields visually striking, novel insight where the conventional analytic approach became infeasible with increasing neuron types.  See response to main concern \#4.
+Our model analyses in general now focus on analyzing the structure of parameter distributions, which EPI captures using flexible approximations enabled by deep learning.
+Here, EPI yields novel insight where the conventional analytic approach became infeasible with increasing neuron types.  See response to main concern \#4.
 \item EPI is used to discover multiple parametric regimes of rapid task switching in a model of superior colliculus.  
-These regimes are efficiently identified and mechanistically characterized using the probabilistic modeling tools afforded by EPI, and we relate the inferred connectivities to results from optogenetic silencing experiments in rats.
+These regimes are efficiently identified and characterized using the probabilistic modeling tools afforded by EPI, and we relate the inferred connectivities to results from optogenetic silencing experiments in rats.
 See response to main concern \#4.
@@
 Specifically, we explain how current methodology infers the parameters producing computation by conditioning on \textit{exemplar datasets} (real or simulated), whereas in EPI we condition directly on the \textit{emergent properties} that define the computation.
+We also note (and have an appendix detailing) how EPI is in fact doing variational bayesian inference, but that the parameterization of the problem as we solve it is more natural given the goal of reproducing a computation.
+In other words, the methodology is mathematically sound and directly connected to well-worn techniques, but the specific form we solve is the appropriate choice for the motivating problem.
 
@@
 
-To improve clarity, we have overhauled the notation and presentation of emergent properties.
-Emergent properties are now denoted with $\mathcal{X}$ to signify an implicit dataset predicated by the emergent property.
-And rather than a vector of first- and second-moment constraints, we present the emergent property more readably (and equivalently) as mean and variance constraints on emergent property statistics:
+To improve clarity, we have overhauled and simplified the notation and presentation of emergent properties.
+Emergent properties are now denoted with $\mathcal{X}$: rather than a vector of first- and second-moment constraints, we present the emergent property more readably (and equivalently) as mean and variance constraints on emergent property statistics:
 
@@
 
-Emergent properties and EPI are introduced and explained with this revised formalism in Section \ref{results_dgm}, and we have completely redone the Methods Section \ref{methods_EPI} to improve clarity, precisely explain EPI optimization (Section \ref{methods_AL_opt}), and derive equivalence to established bayesian inference techniques (Section \ref{methods_VI}).
+Emergent properties and EPI are introduced and explained with this revised notation in Section \ref{results_dgm}, and we have completely redone the Methods Section \ref{methods_EPI} to improve clarity, precisely explain EPI optimization (Section \ref{methods_AL_opt}), and derive equivalence to established bayesian inference techniques (Section \ref{methods_VI}).
 
 In our revised writing, we deemphasize the role of maximum entropy in the EPI algorithm, because this has largely served as a distraction in our experience.
-We show in Section \ref{methods_VI} that EPI is a special case of variational inference, since maximum entropy is the same normative selection principle as variational bayesian methods.  
+We show in Section \ref{methods_VI} that EPI is exactly variational inference (this was mentioned in the previous draft but not as a point of connection and focus, as it should be and now is).
 Therefore, it makes sense to present EPI in the main text as a statistical inference technique that constrains the predictions of the inferred parameters to be an emergent property, and leave the details of the maximum entropy to the technically proficient in Section \ref{methods_EPI}.
@@
 
-The reviewers are right to point out these characteristics of EPI: it does not amortize across emergent properties, and it requires differentiability of the emergent properties of the model. 
-Indeed, SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.  
-However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a trade-off between simulation-based inference approaches.
+We respectfully disagree with this characterization, but we take responsibility here: we certainly needed to improve our exposition and comparisons to clarify the general applicability of EPI.
+We have done so, and what results is a meaningful comparison with the cosubmission (SNPE, Macke et al, now published) and other methods.
+Of course there is a tradeoff (no free lunch, as usual), and these new analyses clarify when EPI is preferable to SNPE, and vice versa.
 
-Unlike SNPE, EPI leverages gradients of the emergent property throughout optimization, which may lead to better efficiency and scalability (Section 3.3).  
+First, unlike SNPE, EPI leverages gradients of the emergent property throughout optimization, which leads to better efficiency and scalability (Section 3.3).
+This is the usual tradeoff of gradient-based vs sampling-based methods.
 The emergent properties of many models in neuroscience are tractably differentiable, four of which we analyze in this manuscript ranging across levels of biological realism, network size, and computational function.
-This trade-off is now explained in the Discussion (Section \ref{discussion}).
+This tradeoff is now explained in the Discussion (Section \ref{discussion}).
+
+Second, of course the reviewers are right to point out EPI does not amortize across emergent properties, and it requires differentiability of the emergent properties of the model.
+SNPE is more suitable for inference with nondifferentiable mechanistic models and scientific problems requiring many inferred parameter distributions.
+However, these relative drawbacks of EPI with respect to SNPE can be considered choices made in a tradeoff between simulation-based inference approaches.
+
+On the balance, these two methods occupy different areas of use, and both are equivalently ``generally/straightforwardly applicable''; we believe the new analyses and discussion in the manuscript clarify that, as it is a key point for practitioners going forward.
 
@@
 As the reviewers indicate, the STG model analyzed in our manuscript is not that of Prinz et al.
-2004, and thus not the model analyzed by the cosubmission.
-We found the Prinz et al.
-model prohibitive to infer with EPI, since the gradients of spiking frequency from its simulation are quite computationally intensive.
-However, we found it critical to show that it’s very possible to do inference in such biophysically realistic Morris-Lecar models when gradients are tractable, which is the case of the 5-neuron STG model we analyzed from Gutierrez et al.
+2004, and thus not the model analyzed by the cosubmission (Macke et al).
+We chose an alternative model, believing it important to demonstrate inference in biophysically realistic Morris-Lecar models when gradients are tractable, which is the case of the 5-neuron STG model we analyzed from Gutierrez et al.
 2013.
@@
 
-We have reached out to Woods Hole Course Project mentors where this work was discussed and explored in its early stages.
-James Fitzgerald and Dhruva Raman are happy with their acknowledgement, and do not insist on deeper involvement and authorship on this paper.
-Stephen Baccus has requested that we acknowledge the summer course, which we have agreed to do.
+Thank you for pointing this out.
+We take authorship and scientific contribution very seriously (it is always part of our manuscript submission process, as it was here).
+We have gone back and discussed every researcher who was in any way involved in this work, and to our best estimation, we think this comment references Woods Hole Course Project mentors where this work was discussed and explored in its early stages.
+We have had explicit followup conversations with James Fitzgerald and Dhruva Raman and noted that of course we would be happy to have them involved at any level (great colleagues!).
+They both reported they are happy with their acknowledgement in the paper and don't feel that there is a justification for authorship (again we were very welcoming of this possibility and maintain positive enthusiasm for both).
+Furthermore, Stephen Baccus has requested that we acknowledge the summer course, which we have done.
+If there is any other ``senior colleague'' the reviewer had in mind, please let us know and we will of course gladly pursue the matter.
 

### 2021-05-25 written/eLifeRevision/Bittner_eLife2020.tex
All the way through 5.1.5

@@
 \usepackage{qting}
+\usepackage[ruled,linesnumbered]{algorithm2e}
 
@@
 \begin{equation}\label{eq:EP_STG2}
-\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2
+\text{Var}_{\mathbf{z},\mathbf{x}}\begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~   0.025^2.
 \end{equation}
 In the emergent property of intermediate hub frequency, the statistic of hub neuron frequency is an expectation over the distribution of parameters $\mathbf{z}$ and the distribution of the data $\mathbf{x}$ that those parameters produce.
-We define the EP as the collection of these two constraints.
+We define the emergent property $\mathcal{X}$ as the collection of these two constraints.
 In general, an emergent property is a collection of constraints on statistical moments that together define the computation of interest.
@@
 \ref{fig:RNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from the exemplary dataset $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from the exemplar dataset $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
 When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:RNN}C, orange).
@@
 
-To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
+To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
 This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient-based and sampling-based estimators (scaling and speed versus generality).
@@
 This work was funded by NSF Graduate Research Fellowship,  DGE-1644869, McKnight Endowment Fund, NIH NINDS 5R01NS100066, Simons Foundation 542963, NSF NeuroNex Award, DBI-1707398, The Gatsby Charitable Foundation, Simons Collaboration on the Global Brain Postdoctoral Fellowship, Chinese Postdoctoral Science Foundation, and International Exchange Program Fellowship.
-Helpful conversations were had with Francesca Mastrogiuseppe, Srdjan Ostojic, James Fitzgerald, Stephen Baccus, Dhruva Raman, Liam Paninski, and Larry Abbott.
+We also acknowledge the Marine Biological Laboratory Methods in Computational Neuroscience Course, where this work was discussed and explored in its early stages.
+Helpful conversations were had with Larry Abbott, Stephen Baccus,  James Fitzgerald, Gabrielle Gutierrez,  Francesca Mastrogiuseppe, Srdjan Ostojic, Liam Paninski, and Dhruva Raman.
 
@@
 \subsection{Emergent property inference (EPI)}\label{methods_EPI}
-Determining the combinations of model parameters that can produce a desired output is a key part of scientific practice.
-Solving inverse problems is especially important in neuroscience, since we require detailed circuit models to produce computation of varying levels of complexity.
-While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models conditioned upon the emergent properties of computation.
-Here, we introduce a novel method for statistical inference, which finds distributions of parameter solutions that are constrained to produce the desired emergent property.
-This method seamlessly handles neural circuit models with stochastic nonlinear dynamical generative processes, which are predominant in theoretical neuroscience.
+Solving inverse problems is an important part of theoretical neuroscience, since we must understand how neural circuit models and their parameter choices can produce computations of varying levels of complexity.
+While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models with respect to their computational properties.
+Here, we introduce a novel method for statistical inference, which uses deep networks to learn parameter distributions constrained to produce emergent properties of computation.
 
 Consider model parameterization $\mathbf{z}$, which is a collection of scientifically meaningful variables that govern the complex simulation of data $\mathbf{x}$.
-For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials (the state) of the five neurons.
+For example (see Section \ref{results_motivating}), $\mathbf{z}$ may be the electrical conductance parameters of an STG subcircuit, and $\mathbf{x}$ the evolving membrane potentials of the five neurons.
 In terms of statistical modeling, this circuit model has an intractable likelihood $p(\mathbf{x} \mid \mathbf{z})$, which is predicated by the stochastic differential equations that define the model.
-From a theoretical perspective, we are less concerned about the likelihood of an exemplary dataset $\mathbf{x}$, but rather the emergent property of intermediate hub frequency (which implies a consistent dataset $\mathbf{x}$).
-
-In the STG example, the statistic $f(\mathbf{x}; \mathbf{z})$ measures hub neuron frequency from the evolution of $\mathbf{x}$ governed by parameters $\mathbf{z}$.
-With EPI, we learn distributions of $\mathbf{z}$ constrained to produce intermediate hub frequency: to obey the constraints placed on the mean and variance of $f(\mathbf{x}; \mathbf{z})$.
-In general, an emergent property $\mathcal{X}$ is defined through the choice of $f(\mathbf{x}; \mathbf{z})$ (which may be one or multiple statistics), and its means $\bm{\mu}$, and variances $\bm{\sigma}^2$:
+From a theoretical perspective, we are less concerned about the likelihood of an exemplar dataset $\mathbf{x}$, but rather the emergent property of intermediate hub frequency (which implies a consistent dataset $\mathbf{x}$).
+
+In this work, emergent properties $\mathcal{X}$ are defined through the choice of emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (which is a vector of one or more statistics), and its means $\bm{\mu}$, and variances $\bm{\sigma}^2$:
  \begin{equation} \label{eq:methods_ep_general}
@@
 \end{equation}
-Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$  over the EPI distribution of parameters and the data produced by those parameters.
-Technically, an emergent property may be a combination of first-, second-, or higher-order moments, but this study focuses on the case written in Equation \ref{eq:methods_ep_general}.
-
-In EPI, deep probability distributions are optimized to learn the inferred distribution.
+In general, an emergent property may be a collection of first-, second-, or higher-order moments of a group of statistics, but this study focuses on the case written in Equation \ref{eq:methods_ep_general}.
+In the STG example, intermediate hub frequency is defined by mean and variance constraints on the statistic of hub neuron frequency $\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ (Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}).
+Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$ over the EPI distribution of parameters ($\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$) and the data produced by those parameters ($\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$), where the learned parameter distribution $q_{\bm{\theta}}(z)$ is determined by deep network weights and biases $\bm{\theta}$.
+
+In EPI, deep probability distributions are optimized to approximate the inferred distribution.
  In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ (we choose an isotropic gaussian) is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
@@
 Such deep probability distributions embed the inferred distribution in a deep network.
-Once optimized, this deep network representation has remarkably useful properties: fast sampling and probability evaluations
+Once optimized, this deep network representation of a distribution has remarkably useful properties: fast sampling and probability evaluations.
 Importantly, fast probability evaluations confer fast gradient and Hessian calculations as well.
 
-Given this choice of circuit model and emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q}$ producing the emergent property $\mathcal{X}$:
+Given this choice of circuit model and emergent property $\mathcal{X}$, $q_{\bm{\theta}}(\mathbf{z})$ is optimized via the neural network parameters $\bm{\theta}$ to find a maximally entropic distribution $q_{\bm{\theta}}^*$ within the deep variational family $\mathcal{Q} = \left\{ q_{\bm{\theta}}(\mathbf{z}) : \bm{\theta} \in \Theta \right\}$ that produces the emergent property $\mathcal{X}$:
 \begin{equation} \label{eq:opt}
@@
  &  \text{s.t.
-} \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
-\\
+} \mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2, \\
  \end{split}
 \end{equation} 
-Entropy is chosen as the normative selection principle to match that of variational bayesian methods (see Section \ref{methods_ME_EF}).
-However, a key difference is that variational bayesian methods do not constrain the predictions of their inferred parameter distribution.
-This optimization is executed using the algorithm of Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
+where $H(q_{\bm{\theta}}(\mathbf{z})) = \mathbb{E}_{\mathbf{z}}\left[- \log q_{\bm{\theta}}(z) \right]$ is entropy.
+By maximizing the entropy of the inferred distribution $q_{\bm{\theta}}$, we select the most random distribution in family $\mathcal{Q}$ that satisfies the constraints of the emergent property.
+Because entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}).
+To run this constrained optimization, we use an augmented lagrangian objective, which is the standard approach for constrained optimization \cite{bertsekas2014constrained}, and the approach taken to fit Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
+This procedure is detailed in Section \ref{methods_AL_opt} and the pseudocode in Algorithm \ref{alg:epi}.
 
@@
 Then, we demonstrate the details of this optimization in a toy example (Section \ref{methods_2DLDS}).
-Finally, we explain how EPI is a form of variational inference (Section \ref{methods_VI}).
+Finally, we explain how EPI is equivalent to variational inference (Section \ref{methods_VI}).
 
@@
 After optimization, variational methods return a parameterized posterior distribution, which we can analyze.
-Also, the variational approximating distribution class is often chosen such that it permits fast sampling.
+Also, the variational approximation is often chosen such that it permits fast sampling.
 In contrast MCMC methods only produce samples from the approximated posterior distribution.
@@
 For scientific model analysis, we seek a parameter distribution represented by an approximating distribution as in variational inference \cite{saul1998mean}: a variational approximation that once optimized yields fast analytic calculations and samples.
-For the reasons described above, ABC and MCMC techniques are unattractive, since they only produce a set of parameter samples lacking probabilities and have unchanging sampling rate.
+For the reasons described above, ABC and MCMC techniques are not suitable, since they only produce a set of parameter samples lacking probabilities and have unchanging sampling rate.
 EPI infers parameters in circuit models using the MEFN \cite{loaiza2017maximum} algorithm with a deep variational approximation.
-The deep neural network of EPI (Fig. \ref{fig:STG}E) defines the parametric form (with variational parameters $\bm{\theta}$) of the deep variational approximation of circuit parameters $\mathbf{z}$.
-
-Since EPI is not conditioning upon exemplary data as in variational bayesian methodology, EPI is not doing established variational inference.
-In contrast, the EPI distribution is constrained to produce an emergent property.
-EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
+The deep neural network of EPI (Fig.
+\ref{fig:STG}E) defines the parametric form (with weights and biases as variational parameters $\bm{\theta}$) of the variational approximation of the inferred parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathbf{x})$.
+The EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
 The analytic relationship between EPI and variational inference is explained in Section \ref{methods_VI}.
 
- We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits (\cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}).%, preprint posted simultaneously with this preprint).
+ We note that, during our preparation and early presentation of this work \cite{bittner2019degenerate, bittner2019examining}, another work has arisen with broadly similar goals: bringing statistical inference to mechanistic models of neural circuits \cite{nonnenmacher2018sbi, desitler2019statistical, gonccalves2019training}.
+ %, preprint posted simultaneously with this preprint).
 We are encouraged by this general problem being recognized by others in the community, and we emphasize that these works offer complementary neuroscientific contributions (different theoretical models of focus) and use different technical methodologies (ours is built on our prior work \cite{loaiza2017maximum}, theirs similarly \cite{LueckmannGoncalves_17}).
@@
 SNPE belongs to a ``sequential" class of recently developed simulation-based inference methods in which two neural networks are used for posterior inference.
-This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL \cite{papamakarios2019sequential})).
-A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE \cite{hermans2020likelihood}).
+This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL) \cite{papamakarios2019sequential}).
+A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE) \cite{hermans2020likelihood}).
 In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
@@
 Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
-Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (see Section \ref{methods_ME_EF}).}
-Since EPI distributions are defined by the mean $\bm{\mu}$ of their statistics, there is the well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} that prohibits an amortization based approach.
-However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors using their natural parameterization \cite{bittner2019approximating}.
+Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (in contrast to its natural parameter, see Section \ref{methods_ME_EF}).}
+The well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} prohibits an amortization based approach in EPI, since emergent properties are defined via the mean parameter of the exponential family distribution.
+However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors when using their natural parameterization \cite{bittner2019approximating}.
 
@@
 These gradients can improve convergence time and scalability, as we have shown on an example conditioning low-rank RNN connectivity on the property of stable amplification (see Section \ref{results_RNN}).
-With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these model gradients when they are tractable.
-Sequential simulation-based inference techniques may be better suited for scientific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable, like when there is a nondifferentiable model or it requires lengthy simulations.
+With EPI, we prove out the suggestion that a deep inference technique can improve efficiency by leveraging these emergent property gradients when they are tractable.
+Sequential simulation-based inference techniques may be better suited for scientific problems where $\nabla_{\mathbf{z}} f(\mathbf{x}; \mathbf{z})$ is intractable or unavailable, like when there is a nondifferentiable emergent property.
 However, the sequential simulation-based inference techniques cannot constrain the predictions of the inferred distribution in the manner of EPI.
@@
 The second-order gradient of the log probability (the Hessian), which is directly afforded by EPI distributions, produces quantified information about parametric sensitivity of the emergent property in parameter space (see Section \ref{results_dgm}).}
-
-%\cite{dhruva/goldman papers}
 
@@
 
-We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
+We use the real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
 Fast probability evaluation  facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
@@
 In this work, normalizing flows are used as flexible parameter distribution approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. 
-We specify the architecture used in each application by the number of Real-NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.
+We specify the architecture used in each application by the number of real NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.
 
@@
 That makes these architectures undesirable for EPI, where efficient sampling is important for optimization, and log probability evaluation speed predicates the efficiency of gradient and Hessian calculations.
-With Real NVP coupling architectures, we get both fast sampling and fast Hessians making both optimization and scientific analysis efficient.
+With real NVP coupling architectures, we get both fast sampling and fast Hessians making both optimization and scientific analysis efficient.
 
 \subsubsection{Maximum entropy distributions and exponential families}\label{methods_ME_EF}
-EPI is a maximum entropy distribution, which have fundamental links to exponential family distributions. 
+The inferred distribution of EPI is a maximum entropy distribution, which have fundamental links to exponential family distributions. 
 A maximum entropy distribution of form:
@@
 p^*(\mathbf{z}) &= \argmax_{p \in \mathcal{P}} H(p(\mathbf{z})) \\
- &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}. \\
+ &  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim p}\left[T(\mathbf{z})\right] = \bm{\mu}_{\text{opt}}, \\
  \end{split}
 \end{equation} 
-will have probability density in the exponential family:
+where $T(\mathbf{z})$ is the sufficient statistics vector and $\bm{\mu}_{\text{opt}}$ a vector of their mean values, will have probability density in the exponential family:
 \begin{equation}
@@
 
-In EPI, emergent properties are defined as statistics having a fixed mean and variance as in Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}.
-The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$
- \begin{equation}
-\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right]
+In this manuscript, emergent properties are defined by statistics $f(\mathbf{x}; \mathbf{z})$ having a fixed mean $\bm{\mu}$ and variance $\bm{\sigma}^2$ as in Equation \ref{eq:methods_ep_general}.
+The variance constraint is a second moment constraint on $f(\mathbf{x}; \mathbf{z})$:
+ \begin{equation}
+\text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[\left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}\right) ^2\right].
 \end{equation}
@@
 \end{equation}
+
+Thus, $\bm{\mu}_{\text{opt}}$ is used to denote the mean parameter of the maximum entropy distribution defined by the emergent property (all constraints), while $\bm{\mu}$ is only the mean of $f(\mathbf{x}; \mathbf{z})$.
+The subscript ``$\text{opt}$" of $\bm{\mu}_{\text{opt}}$ is chosen since it contains all of the contraint values to which the optimization algorithm must adhere.
 
@@
 \end{equation}
-where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
-The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt} (see Section \ref{methods_VI}).
-Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_ME_EF}).
+where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
+The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{x}; \mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
+Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments (Equation \ref{eq:moments}) and $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (Equation \ref{eq:mu_opt}).
+(Although, note that this algorithm is written for general $T(\mathbf{x}; \mathbf{z})$ and $\bm{\mu}$ to satisfy the more general class of emergent properties.)
+The lagrange multipliers $\bm{\eta}_{\text{opt}}$ are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
 Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.
@@
 
+The full EPI optimization algorithm is detailed in Algorithm \ref{alg:epi}.
 The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
@@
 
-The intention is that $c$ and $\bm{\eta}_{\text{opt}}$ start at values encouraging entropic growth early in optimization.  
+\begin{algorithm}\label{alg:epi}
+\SetAlgoLined
+ initialize $\bm{\theta}$ by fitting $q_{\bm{\theta}}$ to an isotropic gaussian of mean $\bm{\mu}_{\text{init}}$ and variance $\bm{\sigma}_{\text{init}}^2$ \\
+ initialize $c_0 > 0$ and $\bm{\eta}_{\text{opt},0} = \mathbf{0}$. \\
+ \For{\upshape Augmented lagrangian epoch $k = 1, ..., k_{\text{max}}$}{
+  \For{\upshape SGD iteration $i = 1, ..., i_{\text{max}}$}{
+   Sample $\mathbf{z}_0^{(1)}, ..., \mathbf{z}_0^{(n)} \sim q_0$, get transformed variable $\mathbf{z}^{(j)} = g_{\bm{\theta}}(\mathbf{z}_0^{(j)})$, $j = 1, ..., n$\\
+   Update $\bm{\theta}$ by descending its stochastic gradient (using ADAM optimizer \cite{kingma2014adam}).
+   {
+   \begin{equation*}
+    \begin{split}
+    \nabla_{\bm{\theta}} L(\bm{\theta}; \bm{\eta}, c) &= \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} \log q_{\bm{\theta}}(\mathbf{z}^{(j)}) + \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)}) \bm{\eta}_{\text{opt}} \\ &+ c_k \frac{2}{n} \sum_{j=1}^{\frac{n}{2}} \nabla_{\bm{\theta}} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)}) \cdot \frac{2}{n} \sum_{j=1 + \frac{n}{2}}^{n} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)})
+    \end{split}
+    \end{equation*}
+   }
+  }
+ Sample $\mathbf{z}_0^{(1)}, ..., \mathbf{z}_0^{(n)} \sim q_0$, get transformed variable $\mathbf{z}^{(j)} = g_{\bm{\theta}}(\mathbf{z}_0^{(j)})$, $j = 1, ..., n$\\
+ Update $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{j=1}^n T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)})$. \\
+ Update $c_{k+1} > c_k$ (see text for detail). \\
+ }
+ \caption{Emergent property inference}
+\end{algorithm}
+
+In general, $c$ and $\bm{\eta}_{\text{opt}}$ should start at values encouraging entropic growth early in optimization.  
 With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see example in Section \ref{methods_2DLDS}).
-The momentum parameters of the Adam optimizer are reset at the end of each augmented lagrangian epoch.
-
-Rather than starting optimization from some $\bm{\theta}$ drawn from a randomized distribution, we found that initializing $q_{\bm{\theta}}(\mathbf{z})$ to approximate an isotropic Gaussian distribution conferred more stable, consistent optimization.
-The parameters of the Gaussian initialization were chosen on an application-specific basis.  
-Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some standard deviation $\bm{\sigma}_{\text{init}}$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
+This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see Figure \ref{fig:LDS1}).
+The momentum parameters of the Adam optimizer are reset at the end of each augmented lagrangian epoch, which proceeds for $i_{\text{max}}$ iterations.
+In this work, we used a maximum number of augmented lagrangian epochs $k_{\text{max}} >= 5$.
+
+Rather than starting optimization from some $\bm{\theta}$ drawn from a randomized distribution, we found that initializing $q_{\bm{\theta}}(\mathbf{z})$ to approximate an isotropic gaussian distribution conferred more stable, consistent optimization.
+The parameters of the gaussian initialization were chosen on an application-specific basis.  
+Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some variance $\bm{\sigma}_{\text{init}}^2$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
+Deep probability distributions were fit to these gaussian initializations using 10,000 iterations of stochastic gradient descent on the evidence lower bound (as in \cite{bittner2019approximating}) with Adam optimizer and a learning rate of $10^{-3}$.
 
@@
 The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented lagrangian epoch.
+Of all $k_{\text{max}}$ augmented lagrangian epochs, we select the EPI inferred distribution as that which satisfies the convergence criteria and has greatest entropy.
 
@@
 If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
-Next, one should consider the computational cost of the gradient calculations. 
-In the best circumstance, there is a simple, closed form expression (e.g. Section \ref{methods_RNN}) for the emergent property statistic given the model parameters.  
-On the other end of the spectrum, many forward simulation iterations may be required before a high quality measurement of the emergent property statistic is available  (e.g.
-Section \ref{methods_STG}).
-In such cases, backpropagating gradients through the SDE evolution will be expensive.
 
@@
 \begin{equation}
-A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
-\end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$s), the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ were chosen to contain the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system.
- $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean 1Hz with variance (0.1Hz)$^2$:
-\begin{equation}
- \mathbb{E}\left[T(\mathbf{x}) \right]_{\mathbf{z}, \mathbf{x}} ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ (\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z})-0)^2  \\ (\text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z})-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}_{\text{opt}}.
+A = \begin{bmatrix} a_{1,1} & a_{1,2} \\ a_{2,1} & a_{2,2} \end{bmatrix}.
+\end{equation}
+To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_{1,1}, a_{1,2}, a_{2,1}, a_{2,2}]$ (fixing $\tau=1$s), the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ were chosen to contain parts of the primary eigenvalue of $A$, which predicate frequency, $\text{imag}(\lambda_1)$, and the growth/decay, $\text{real}(\lambda_1)$, of the system.
+$\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively that of positive imaginary component when the eigenvalues are complex conjugate pairs.  
+To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean 1Hz with variance 0.1Hz$^2$:
+\begin{equation}
+\begin{split}
+\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0 \\ 2\pi \end{bmatrix} \\
+\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix}.
+\end{split}
+\end{equation}
+
+To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), we concatenate these first and second moment constraints into a vector of sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and constraint values $\bm{\mu}_{\text{opt}}$.
+
+\begin{equation}
+ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] ~~ \triangleq ~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}}  \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ (\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z})-0)^2  \\ (\text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z})-2 \pi)^2 \end{bmatrix} = \begin{bmatrix} 0 \\ 2 \pi \\ 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}_{\text{opt}}.
  \end{equation} 
@@
 \textbf{C}. Entropy throughout the optimization.  
-At the beginning of each augmented lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
+At the beginning of each augmented lagrangian epoch ($i_{\text{max}} = 2,000$ iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.
 \textbf{D}. Emergent property moments throughout optimization.  
@@
 \begin{equation}
-\lambda = \frac{(\frac{a_1 + a_4}{\tau}) \pm \sqrt{(\frac{a_1+a_4}{\tau})^2 + 4(\frac{a_2 a_3 - a_1 a_4}{\tau})}}{2}.
-\end{equation}
-
-Importantly, even though $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x})\right]$ is calculable directly via a closed form function and does not require simulation, we cannot derive the distribution $q^*_{\bm{\theta}}$ directly.
+\lambda = \frac{(\frac{a_{1,1} + a_{2,2}}{\tau}) \pm \sqrt{(\frac{a_{1,1}+a_{2,2}}{\tau})^2 + 4(\frac{a_{1,2} a_{2,1} - a_{1,1} a_{2,2}}{\tau})}}{2}.
+\end{equation}
+We study this example, because the inferred distribution is curved and multimodal, and we can compare the result of EPI to analytically derived contours of the emergent property statistics.
+
+Despite the simple analytic form of the emergent property statistics, the EPI distribution in this example is not simply determined.
+Although $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z})\right]$ is calculable directly via a closed form function, the distribution $q^*_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ cannot be derived directly.
 This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we used EPI to approximate this distribution (Fig.
-\ref{fig:LDS1}B).
-We used a real-NVP normalizing flow architecture with four masks, two neural network layers of 15 units per mask, with batch normalization momentum 0.99, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
+Instead, we used EPI to approximate this distribution (Fig. \ref{fig:LDS1}B). 
+We used a real NVP normalizing flow architecture three coupling layers and two-layer neural networks of 50 units per layer, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
 (see Section \ref{methods_NF}).
@@
 To validate our method, we analytically derived the contours of the probability density from the emergent property statistics and values.
-In the $a_1$-$a_4$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, dotted black line at
-the standard deviation $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.25$, and the dotted gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
-The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_1$ and $a_4$.
+In the $a_{1,1}$-$a_{2,2}$ plane, the black line at $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} = 0$, dashed black line at
+the standard deviation $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.25$, and the dashed gray line at twice the standard deviation
+$\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
+The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_{1,1}$ and $a_{2,2}$.
 Intuitively, the parameters equivalent with respect to emergent property statistic $\text{real}(\lambda_1)$ have similar log densities.
 
-To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  When $\text{real}(\lambda_1) = \frac{a_1 + a_4}{2} = 0$, we have
+To explain the bimodality of the EPI distribution, we examined the imaginary component of $\lambda_1$.  
+When $\text{real}(\lambda_1) = a_{1,1} + a_{2,2} = 0$ (which is the case on average in $\mathcal{X}$), we have
 \begin{equation}
 \text{imag}(\lambda_1) = \begin{cases}
-                             \sqrt{\frac{a_1 a_4 - a_2 a_3}{\tau}},  & \text{if } a_1 a_4 < a_2 a_3 \\
+                             \sqrt{\frac{a_{1,1} a_{2,2} - a_{1,2} a_{2,1}}{\tau}},  & \text{if } a_{1,1} a_{2,2} < a_{1,2} a_{2,1} \\
                              0 & \text{otherwise } \\
@@
 \end{equation}
-
-When $\tau=1$ and $a_1 a_4 > a_2 a_3$ (center of distribution above), we have the following equation for the other two dimensions:
-\begin{equation}
-\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3
-\end{equation}
-Since we constrained $\mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[\text{imag}(\lambda)\right] = 2 \pi$, we can plot contours of the equation $\text{imag}(\lambda_1)^2 = a_1 a_4 - a_2 a_3 = (2 \pi)^2$ for various $a_1 a_4$ (Fig.
-\ref{fig:LDS2}B).
-With $\sigma_{1,4} = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}(|a_1 a_4 - E_{q_{\bm{\theta}}}[a_1 a_4]|)$, we show the contours as $a_1 a_4 = 0$ (black), $a_1 a_4 = -\sigma_{1,4}$ (black dotted), and $a_1 a_4 = -2\sigma_{1,4}$ (grey dotted).
-This validates the curved structure of the inferred distribution learned through EPI.  
-We took steps in negative standard deviation of $a_1 a_4$ (dotted and gray lines), since there are few positive values $a_1 a_4$ in the learned distribution.  
+In Figure \ref{fig:LDS2}B, we plot the contours of $\text{imag}(\lambda_1)$ where $a_{1,1} a_{2,2}$ is fixed to 0 at one standard devation ($\frac{\pi}{5}$, black dashed) and two standard deviations ($\frac{2\pi}{5}$, gray dashed) from the mean of $2 \pi$.
+This validates the curved multimodal structure of the inferred distribution learned through EPI.  
 Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
@@
 \end{center}
-\caption{\textbf{A}. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
-\textbf{B}. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
+\caption{\textbf{A}. Probability contours in the $a_{1,1}$-$a_{2,2}$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+\textbf{B}. Probability contours in the $a_{1,2}$-$a_{2,1}$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
 \label{fig:LDS2}
@@
 \end{center}
-\caption{Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(t=0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+\caption{Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ and their simulated activity from $\mathbf{x}(t=0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
 \textbf{A}. Each dimension of the simulated trajectories throughout time.  
@@
 \subsubsection{EPI details for the STG model} \label{methods_stg_epi}
-As a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
 \begin{equation} 
@@
 \end{equation}
-Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of 5,000 iterations(see Section \ref{methods_AL_opt}).  
+Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 5,000$ iterations (see Section \ref{methods_AL_opt}).  
 The optimization converged after five epochs (Fig. \ref{fig:STG1}).
@@
 
-For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
+For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three coupling layers and two-layer neural networks of 25 units per layer.
 The normalizing flow architecture mapped $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$, initialized to a gaussian approximation of samples returned by a preliminary ABC search.
@@
 To complete the emergent property definition, we chose variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
-In terms of the EPI optimization variables, this is written as
+To write the emergent property of Equation \ref{eq:EP_LRRNN} in terms of the EPI augmented lagrangian optimization, this is written as
 \begin{equation} 
@@
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
-We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
+We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $i_{\text{max}} = 500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
 \ref{fig:RNN}B blue line, and Fig.
@@
 \subsubsection{EPI details for the V1 model}
-For EPI in Figures \ref{fig:V1}D-E and \ref{fig:V1_1}, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 50 units per layer.
+For EPI in Figures \ref{fig:V1}D-E and \ref{fig:V1_1}, we used a real NVP architecture with three coupling layers and two-layer neural networks of 50 units per layer.
 The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
-EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, and $\beta = 2$.
+EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, $\beta = 2$, and $i_{\text{max}} = 2,000$ iterations per epoch.
 The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across three random seeds.
@@
 \subsubsection{EPI details for the SC model} \label{methods_sc_epi}
-Writing the EPI distribution as a maximum entropy distribution, $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
+To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
 \begin{equation} 
@@
 \end{equation}
-Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of 2,000 iterations (see Section \ref{methods_AL_opt}).  
+Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 2,000$ iterations (see Section \ref{methods_AL_opt}).  
 The optimization converged after ten epochs (Fig. \ref{fig:SC6}).

### 2021-05-25 written/eLifeRevision/Response.tex
updated response

@@
 Figure \ref{fig:STG} has been largely redone to reflect the modified presentation, and improve the pictorial representation of the method.
-In Section \ref{methods_RNN}, we now show that EPI is the only simulation-based inference method that controls the predictions of its inferred distribution (Fig. \ref{fig:RNN}C-D).
+In Section \ref{results_RNN}, we now show that EPI is the only simulation-based inference method that controls the predictions of its inferred distribution (Fig. \ref{fig:RNN}C-D).
 
@@
 
-Thank you for pointing us to these references on sensitivity analyses and applied dynamical systems.  
-We have incorporated many of them into the current manuscript and explain EPI’s relation to each class of techniques in Section 5.1.1 Related approaches.
-
-\begin{displayquote}
-Lines 849-862 \\
-``\quoting{structural_id}''
-\end{displayquote}
+Thank you for pointing us to these references on sensitivity analyses, MCMC inference, and applied dynamical systems.  
+We have incorporated most of them into the current manuscript and explain EPI’s relation to each class of these techniques throughout Section \ref{methods_related} Related approaches.
 
@@
 We modify our writing to clarify that EPI is not a key to all of theoretical neuroscience, but rather a powerful solution to inverse problems in neural circuit modeling.
-Inverse problems are indeed a major part of theoretical neuroscience.
-Their solution is necessary for the scientific evaluation of mathematical representations of neural circuits, and ultimately for progress in the field.
+Inverse problems are indeed a major part of theoretical neuroscience, and their solutions are important for the evaluation  of neural circuit models.
 We clarify this point in the introduction:
@@
 To clarify, EPI does not execute posterior inference, because there is no explicit dataset or specified prior belief in the EPI framework.
-However, we derive the relation of EPI to bayesian inference in Section 5.1.6, which shows EPI uses a uniform prior when framed as variational inference.
+However, we derive the relation of EPI to bayesian variational inference in Section \ref{methods_VI}, which shows EPI uses a uniform prior when framed as variational inference.
 
@@
 Lines 1026-1029 \\
-``\quoting{ME_prior}''
+``\quoting{EPI_VI}''
 \end{displayquote}
@@
 Thus, an important clarification is that sensitivity quantifications are made in the context of the chosen parameter scalings.
-%A final concern here is numerical: if the inferred distribution is in a thin region of parameter space, that is not well represented by the precision of the numerical format, there can be issues in optimization.
-%It makes most sense to make sure EPI is exploring through parameter space through an industry-recommended range of values, and that sensitivity measurements are interpretable.
-
-\textbf{SB: Should we add text about this in the paper?}
 
@@
 
-Thank you for the feedback.  We have taken care to give more general context, and motivation for each neural circuit model.
+Thank you for the feedback.  We have taken care to give more general context and motivation for each neural circuit model.
 
@@
 
+We thank the reviewer for their positive comments.
+
 \textbf{Major points }
@@
 
-We thank the reviewer for their positive comments.  
+We agree with the reviewer that hyperparameter sensitivity and global optima are important considerations of any optimization algorithm using deep neural networks.
 To draw a parallel, training deep networks for visual processing used to be considered infeasible, but became easier through iterative improvements in architectural and hyperparameter choices that spread across the field.
@@
 Lines 431-438\\
-``\quoting{discussion_methods}''
+``\quoting{discussion_methods} \\ \quoting{discussion_gradients}''
 \end{displayquote}
@@
 Some specific ones are below, but the authors should more generally send the paper to naïve readers to check what is/is not well explained.
--Figure 1 is somewhat opaque and also has notational issues (e.g. omega is the frequency but also appears to be the random input sample). 
--For the general audience of eLife, panels C and D are not well described individually or well connected to each other and don't illustrate or describe all of the relevant variables (including what q0 is and what x is).
--In equation 2 (and also in the same equation in the supplement), it was not immediately obvious what the expectation was taken over. 
--The authors don't specific the distribution of w (it's referred to only as 'a simple random variable', which is not clear). 
--It was also sometimes hard to quickly find in the text basic, important quantities like what z was for a given simulation. 
--The augmented Lagrangian optimization was not well explained or motivated. There is a reference to m=absolute value(mu) but I didn't see m in the above equation. 
--Using mu to describe a vector that includes means and variances is confusing notation since mu often denotes means 
--It would be helpful to have a pseudo-code 'Algorithm' figure or section of the text }
+}
 
 Thank you for this detailed list of improvements to be made when describing the method.
-We have taken care to address each item in this list.
+
+\begin{itemize}
+\item \textbf{Figure 1 is somewhat opaque and also has notational issues (e.g. omega is the frequency but also appears to be the random input sample).}
+
+Figure 1 has been completely revised.
+
+\item \textbf{For the general audience of eLife, panels C and D are not well described individually or well connected to each other and don't illustrate or describe all of the relevant variables (including what q0 is and what x is).}
+
+Agreed.  In the new figure, we clearly depict $\mathbf{z}$ as parameters and $\mathbf{x}$ as circuit activity.  
+We keep the vertical directionality of panel D exclusively for the deep generative process of the deep probability distribution (where $q_0$ is replaced and shown explicitly as an isotropic gaussian input to the deep network).
+The horizontal directionality shared between panels D and E reflect the procedure of theoretical neuroscience described in Section \ref{results_motivating}.
+
+\item \textbf{In equation 2 (and also in the same equation in the supplement), it was not immediately obvious what the expectation was taken over.}
+
+We take care to always indicate the variables over which expectations are taken in the updated manuscript.
+
+\item \textbf{The authors don't specific the distribution of w (it's referred to only as 'a simple random variable', which is not clear).}
+
+It is clarified in the text that this simple initial distribution transformed by the deep neural network is an isotropic gaussian.
+
+\item \textbf{It was also sometimes hard to quickly find in the text basic, important quantities like what z was for a given simulation.}
+
+We have made sure to make this explicit in each section of the paper.
+
+\item \textbf{The augmented Lagrangian optimization was not well explained or motivated. There is a reference to m=absolute value(mu) but I didn't see m in the above equation.}
+
+Lines X-Y \\
+\begin{displayquote}
+``\quoting{AL_standard}"
+\end{displayquote}
+
+\item \textbf{Using mu to describe a vector that includes means and variances is confusing notation since mu often denotes means }
+
+Agreed. Throughout the updated main text, we only describe emergent properties through explicit mean and variance constraints as in Equation 11.
+We reserve the mean parameterization $\bm{\mu}_{\text{opt}}$ of the maximum entropy solution of the EPI optimization for technical details in methods.
+The difference between $\bm{\mu}$ and $\bm{\mu}_{\text{opt}}$ is described in Section \ref{methods_ME_EF}.
+
+\item \textbf{It would be helpful to have a pseudo-code 'Algorithm' figure or section of the text}
+
+We provide pseudocode for the EPI optimization in Algorithm \ref{alg:epi}, which mirrors the pseudocode found in the paper describing the underlying algorithm for MEFN \cite{loaiza2017maximum}.
+
+\end{itemize}
 
@@
 
-We have removed the variable $\omega$ as it was unnecessary, and avoided introducing another variable $f$ to be conflated with $f$'s use as emergent property statistic.
+We have removed the variable $\omega$ as it was unnecessary, and avoided introducing another variable $f$ to be conflated with $f$'s use as the emergent property statistics.
 

### 2021-05-25 written/eLifeRevision/Bittner_eLife2020.tex
updated response

@@
 
-To examine how connectivity affects response to delay period inactivation, we grouped connectivities of the EPI distribution along the continuum linking regimes 1 and 2 of Section \ref{results_SC}:
-\begin{equation}\label{eq:mode_group}
-Z(sW)  = \{\mathbf{z} ~~\text{if}~\argmin_{y \in Y} |\mathbf{z} - \mathbf{z}^*(y)|_2 = sW, ~~\text{for}~z \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})\},
-\end{equation}
-where $Y = \{-1., -0.75, ..., 1.\}$ are the values of $sW$ for which we calculated the mode.
-$Z(sW)$ is then the set of EPI samples for which the closest mode was $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_regime}).
+To examine how connectivity affects response to delay period inactivation, we grouped connectivities of the EPI distribution along the continuum linking regimes 1 and 2 of Section \ref{results_SC}.
+$Z(sW)$ is the set of EPI samples for which the closest mode was $\mathbf{z}^*(sW)$ (see Section \ref{methods_sc_regime}).
 In the following analyses, we examine how error, and the influence of connectivity eigenvalue on Anti error change along this continuum of connectivities.
@@
 Because entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}).
-To run this constrained optimization, we use an augmented lagrangian objective, which is the standard approach for constrained optimization \cite{bertsekas2014constrained}, and the approach taken to fit Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.
+\q{AL_standard}{
+To run this constrained optimization, we use an augmented lagrangian objective, which is the standard approach for constrained optimization \cite{bertsekas2014constrained}, and the approach taken to fit Maximum Entropy Flow Networks (MEFNs) \cite{loaiza2017maximum}.}
 This procedure is detailed in Section \ref{methods_AL_opt} and the pseudocode in Algorithm \ref{alg:epi}.
@@
 \end{equation}
-where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient.
+where there are average constraint violations 
+\begin{equation}
+R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right],
+\end{equation}
+ $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the lagrange multipliers where $m$ is the number of total constraints
+\begin{equation}
+m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|,
+\end{equation}
+and $c$ is the penalty coefficient. 
 The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{x}; \mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
@@
 
-When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
+\q{ME_technical}{When assessing the suitability of EPI for a particular modeling question, there are some important technical considerations. 
 First and foremost, as in any optimization problem, the defined emergent property should always be appropriately conditioned (constraints should not have wildly different units).  
 Furthermore, if the program is underconstrained (not enough constraints), the distribution grows (in entropy) unstably unless mapped to a finite support.  
-If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).
+If overconstrained, there is no parameter set producing the emergent property, and EPI optimization will fail (appropriately).}
 
@@
 %Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood distribution $p(\mathbf{x} \mid \mathbf{z})$.
-In bayesian inference, we obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+In bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood $p(\mathbf{x} \mid \mathbf{z})$.
+We obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
 \begin{equation}
@@
 The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods, such as variational inference \cite{saul1998mean}, are utilized.
-
-%Now, consider the goal of doing variational inference with an exponential family posterior distribution $p(z \mid \mathbf{x})$.  
-%We use the following abbreviated notation to collect the base measure $b(z)$ and sufficient statistics $T(z)$ into $\tilde{T}(z)$ and likewise concatenate a 1 onto the end of the natural parameter $\tilde{\eta}(\mathbf{x})$.
-%The log normalizing constant $A(\eta(\mathbf{x}))$ remains unchanged:
+However, conjugacy is rare in practice, and alternative methods such as variational inference \cite{saul1998mean} are utilized.
+
+In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$ to be as close as possible to the posterior under the KL divergence criteria
+\begin{equation}
+q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
+\end{equation}
+This KL divergence can be written in terms of entropy of the variational approximation:
+\begin{equation}
+KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
+\end{equation}
+\begin{equation}
+= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
+\end{equation}
+Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
+This is usually framed as maximizing the evidence lower bound (ELBO)
+\begin{equation} \label{eq:VI_opt}
+\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
+\end{equation}
+
+Now, we will show how the maximum entropy problem of EPI is equivalent to variational inference.
+In general, a maximum entropy problem (as in Equation \ref{eq:max_ent}) has an equivalent lagrange dual form:
+\begin{equation}
+\begin{split}
+\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) &\Longleftrightarrow \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \bm{\eta}^{*\top} \mathbb{E}_{\mathbf{z} \sim q} \left[ \phi(\mathbf{z}) \right], \\
+  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim q}\left[\phi(\mathbf{z})\right] = \mathbf{0}~& \\
+ \end{split}
+\end{equation} 
+with sufficient statistics $\phi(\cdot)$ and lagrange multipliers $\bm{\eta}^*$.
+By moving the lagrange multipliers within the expectation
+\begin{equation}q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\bm{\eta}^{*\top} \phi(\mathbf{z}) \right],
+\end{equation} 
+inserting a $\log \exp(\cdot)$ within the expectation, 
+\begin{equation} q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} \phi(\mathbf{z}) \right) \right],
+\end{equation}
+and finally replacing $\phi(\cdot)$ with the sufficient statistics of the EPI optimization
+\begin{equation} \label{eq:epi_vi}
+q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T_1(\mathbf{x}; \mathbf{z}) \right] \\ ...
+\\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T_m(\mathbf{x}; \mathbf{z}) \right] \end{bmatrix} \right) \right],
+\end{equation}
+we can then compare to the objective used in variational inference  (Equation \ref{eq:VI_opt}).
+\q{EPI_VI}{
+We see that EPI is exactly variational inference with an exponential family likelihood defined by sufficient statistics $\phi(z) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T(\mathbf{x}; \mathbf{z}) \right]$, and where the natural parameter $\bm{\eta}^*$ is predicated by the mean parameter $\bm{\mu}_{\text{opt}}$.
+Equation \ref{eq:epi_vi} implies that EPI uses an improper (or uniform) prior, which is easily changed.}
+
+This derivation of the equivalence between EPI and variational inference emphasizes why defining a statistical inference program by its mean parameterization $\bm{\mu}_{\text{opt}}$ is so useful.
+With EPI, one can clearly define the emergent property $\mathcal{X}$ that the model of interest should produce through intuitive selection of $\bm{\mu}_{\text{opt}}$ for a given $T(\mathbf{x}; \mathbf{z})$.
+Alternatively, figuring out the correct natural parameters $\bm{\eta}^*$ for the same $T(\mathbf{x}; \mathbf{z})$ that produces $\mathcal{X}$ will be generally unclear.
+
+%Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
+%\begin{equation}
+%p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
+%\end{equation}
+%where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
+%The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
+%\begin{equation}
+%\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
+%\end{equation}
+%where
+%\begin{equation} 
+%\label{eq:eta_f}
+%\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
+%\end{equation}
+%\begin{equation}
+%T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
+%\end{equation}
+%The variational objective is then
+%\begin{equation}
+%\label{eq:VI_obj}
+%\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right].
+%\end{equation}
+
+%Comparing this to the lagrangian objective (without augmentation) of EPI, we see they are the same
 %\begin{equation}
 %\begin{split}
-%p(z \mid \mathbf{x}) = b(z) \exp{\left( \eta(\mathbf{x})^\top T(z) - A(\eta(\mathbf{x})) \right)} = \exp{\left( \begin{bmatrix} \eta(\mathbf{x}) \\ 1 \end{bmatrix}^\top \begin{bmatrix} T(z) \\ b(z) \end{bmatrix} - A(\eta(\mathbf{x})) \right)} \\= \exp{\left(\tilde{\eta(\mathbf{x})}^\top \tilde{T}(z) - A(\eta(\mathbf{x})) \right)}
-%\end{split}.
+%q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
+% &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
+%\end{split}
 %\end{equation}
-%Variational inference with an exponential family posterior distribution uses optimization to minimize the following divergence \cite{blei2017variational}:
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})).
-%\end{equation}
-%$q_{\bm{\theta}}(z)$ is the variational approximation to the posterior with variational parameters $\bm{\theta}$.  
-%We can write this KL divergence in terms of entropy of the variational approximation:
-%\begin{equation}
-%KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) = \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(z)) \right] - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \log (p(z \mid \mathbf{x})) \right]
-%\end{equation}
-%\begin{equation}
-% = -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) - A(\eta(\mathbf{x})) \right].
-%\end{equation}
-%As far as the variational optimization is concerned, the log normalizing constant is independent of $q_{\bm{\theta}}(z)$, so it can be dropped
-%\begin{equation}
-%\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(z \mid \mathbf{x})) =  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top  \tilde{T}(z) \right].
-%\end{equation}
-% Further, we can write the objective in terms of the first moment of the sufficient statistics $\bm{\mu} = \mathbb{E}_{z \sim p(z \mid \mathbf{x})}\left[T(z) \right]$:
-% \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right] + \tilde{\eta}(\mathbf{x})^\top \bm{\mu},
-% \end{equation}
-%which simplifies to
-%  \begin{equation}
-%=  \argmin_{q_{\bm{\theta}} \in Q} -H(q_{\bm{\theta}}) - \mathbb{E}_{z \sim q_{\bm{\theta}}} \left[ \tilde{\eta}(\mathbf{x})^\top \left(  \tilde{T}(z) -\bm{\mu} \right) \right].
-% \end{equation}.
-%
-%In comparison, in emergent property inference (EPI), we solve the following problem:
-%\begin{equation}
-%q_{\bm{\theta}}^*(z) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}(z)),   \text{  s.t.
-} \mathbb{E}_{z \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid z)}\left[T(\mathbf{x})\right] \right] = \bm{\mu}.
-%\end{equation}
-%The Lagrangian objective (without augmentation) is
-%\begin{equation}
-%q_{\bm{\theta}}^* = \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \eta_{\text{opt}}^\top \left(\mathbb{E}_{z \sim q_{\bm{\theta}}} \left[\tilde{T}(z) \right] - \bm{\mu} \right).
-%\end{equation}
-%Thus, as the optimization proceeds, $\eta_{\text{opt}}^\top$ should converge to the natural parameter $\tilde{\eta}(\mathbf{x})$ through its adaptations in each epoch (see Section \ref{methods_AL_opt}).
-%
-%We have shown that there is indeed a clear relationship between Bayesian inference and EPI.
-%Specifically, EPI is executing variational inference in an exponential family posterior,  whose sufficient statistics are the emergent property statistics and mean parameterization are the emergent property values.
-%However, in EPI we  have not specified a prior distribution, or collected data, which can inform us about model parameters.
-%Instead we have a mathematical specification of an emergent property, which the model must produce, and a maximum entropy selection principle.
-%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{B})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{B}$ (see Section \ref{methods_EPI}).
-
-In variational inference, a posterior approximation $q_{\bm{\theta}}^*$ is chosen from within some variational family $\mathcal{Q}$
-\begin{equation}
-q_{\bm{\theta}}^*(\mathbf{z}) = \argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}}(\mathbf{z}) \mid \mid p(\mathbf{z} \mid \mathbf{x})).
-\end{equation}
-The KL divergence can be written in terms of entropy of the variational approximation:
-\begin{equation}
-KL(q_{\bm{\theta}}(\mathbf{z})  \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (q_{\bm{\theta}}(\mathbf{z})) \right] - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{z} \mid \mathbf{x})) \right]
-\end{equation}
-\begin{equation}
-= -H(q_{\bm{\theta}}) - \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z})) - \log(p(\mathbf{x}))\right]
-\end{equation}
-Since the marginal distribution of the data $p(\mathbf{x})$ (or ``evidence") is independent of $\bm{\theta}$, variational inference is executed by optimizing the remaining expression.
-This is usually framed as maximizing the evidence lower bound (ELBO)
-\begin{equation}
-\argmin_{q_{\bm{\theta}} \in Q} KL(q_{\bm{\theta}} \mid \mid p(\mathbf{z} \mid \mathbf{x})) = \argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}} \left[ \log (p(\mathbf{x} \mid \mathbf{z}))  + \log(p(\mathbf{z}))\right].
-\end{equation}
-
-Now, consider the setting where we have chosen a uniform prior, and stipulate a mean-field gaussian likelihood on a chosen statistic of the data $f(\mathbf{x}; \mathbf{z})$
-\begin{equation}
-p(\mathbf{x} \mid \mathbf{z}) = \mathcal{N}(f(\mathbf{x}; \mathbf{z}) \mid \bm{\mu}_f, \Sigma_f),
-\end{equation}
-where $\Sigma_f = \text{diag}(\bm{\sigma}_f^2)$.
-The log likelihood is then proportional to a dot product of the natural parameter of this mean-field gaussian distribution and the first and second moment statistics.
-\begin{equation}
-\log p(\mathbf{x} \mid \mathbf{z}) \propto \bm{\eta}_f^\top T(\mathbf{x}, \mathbf{z}),
-\end{equation}
-where
-\begin{equation} 
-\label{eq:eta_f}
-\bm{\eta}_{f} = \begin{bmatrix} \frac{\bm{\mu}_f}{\bm{\sigma}_f^2} \\ \frac{-1}{2\bm{\sigma}_f^2} \end{bmatrix}, \text{ and}
-\end{equation}
-\begin{equation}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu}_f \right)^2 \end{bmatrix}.
-\end{equation}
-The variational objective is then
-\begin{equation}
-\label{eq:VI_obj}
-\argmax_{q_{\bm{\theta}} \in Q} H(q_{\bm{\theta}}) + \bm{\eta}_f^\top \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ T(\mathbf{x}; \mathbf{z}) \right].
-\end{equation}
-
-Comparing this to the lagrangian objective (without augmentation) of EPI, we see they are the same
-\begin{equation}
-\begin{split}
-q_{\bm{\theta}}^*(\mathbf{z}) &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \left(\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}_{\text{opt}} \right) \\
- &= \argmin_{q_{\bm{\theta}} \in Q} - H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right].
-\end{split}
-\end{equation}
-where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
-\q{ME_prior}{Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
-The mean and variances of the mean-field gaussian likelihood are predicated by $\bm{\eta}_{\text{opt}}$ (Equations \ref{eq:eta_f} and \ref{eq:VI_obj}), which is adapted after each EPI optimization epoch based on $\mathcal{X}$ (see Section \ref{methods_AL_opt}).}
-In EPI, the inferred distribution is not conditioned on a finite dataset as in variational inference, but rather the emergent property $\mathcal{X}$ dictates the likelihood parameterization such that the inferred distribution will produce the emergent property.
-As a note, we could not simply choose $\bm{\mu}_f$ and $\bm{\sigma}_f$ directly from the outset, since we do not know which of these choices will produce the emergent property $\mathcal{X}$, which necessitates the EPI optimization routine that adapts $\bm{\eta}_{\text{opt}}$.
-Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
+%where $T(\mathbf{x}; \mathbf{z})$ consists of the first and second moments of the emergent property statistic $f(\mathbf{x}; \mathbf{z})$ (Equation \ref{eq:moments}).
+%\q{ME_prior}{Thus, EPI is implicitly executing variational inference with a uniform prior and a mean-field gaussian likelihood on the emergent property statistics.
+%The mean and variances of the mean-field gaussian likelihood are predicated by $\bm{\eta}_{\text{opt}}$ (Equations \ref{eq:eta_f} and \ref{eq:VI_obj}), which is adapted after each EPI optimization epoch based on $\mathcal{X}$ (see Section \ref{methods_AL_opt}).}
+%In EPI, the inferred distribution is not conditioned on a finite dataset as in variational inference, but rather the emergent property $\mathcal{X}$ dictates the likelihood parameterization such that the inferred distribution will produce the emergent property.
+%As a note, we could not simply choose $\bm{\mu}_f$ and $\bm{\sigma}_f$ directly from the outset, since we do not know which of these choices will produce the emergent property $\mathcal{X}$, which necessitates the EPI optimization routine that adapts $\bm{\eta}_{\text{opt}}$.
+%Accordingly, we replace the notation of $p(\mathbf{z} \mid \mathbf{x})$ with $p(\mathbf{z} \mid \mathcal{X})$ conceptualizing an inferred distribution that obeys emergent property $\mathcal{X}$ (see Section \ref{methods_EPI}).
 
@@
 In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
-Specifically, we calculated
-\begin{equation}
-\begin{split}
-\mathbf{z}^*(\sigma_{P,\text{fixed}}) = &\argmax_{\mathbf{z}} \log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}) \\
-&\text{s.t. } \sigma_P = \sigma_{P,\text{fixed}}\
-\end{split}
-\end{equation}
-At each mode $\mathbf{z}^*$, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
+At each local mode $\mathbf{z}^*(\sigma_P)$, where $\sigma_P$ is fixed, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
 
@@
 For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampling from the EPI distribution.
-Each sample was assigned to the closest mode $\mathbf{z}^*(sW)$ (Equation \ref{eq:mode_group}).
+Each sample was assigned to the closest mode $\mathbf{z}^*(sW)$.
 Sampling continued until 500 samples were assigned to each mode, which took 2.67 seconds (5.34ms/sample-per-mode).

### 2021-05-27 written/eLifeRevision/Response.tex
update algo and pass through paper.

@@
 We have done so, and what results is a meaningful comparison with the cosubmission (SNPE, Macke et al, now published) and other methods.
-Of course there is a tradeoff (no free lunch, as usual), and these new analyses clarify when EPI is preferable to SNPE, and vice versa.
+Of course, there is a tradeoff (no free lunch, as usual), and these new analyses clarify when EPI is preferable to SNPE, and vice versa.
 
@@
 This manuscript's explanation of hyperparameter choices and the extensive set of examples in the online code will serve as valuable guidelines for future research using this method.
-Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: the augmented Lagrangian constant and deep probability distribution architecture.
+Every figure of this paper is reproducible with the jupyter notebooks, and there are several tutorials for understanding the most consequential hyperparameters: the augmented lagrangian constant and deep probability distribution architecture.
 

### 2021-05-27 written/eLifeRevision/Bittner_eLife2020.tex
update algo and pass through paper.

@@
 Each trace (15 per algorithm) corresponds to simulation of one $z$.
-(Below) Ratio of obtained samples producing stable amplification, monotonic decay, and instability.
+(Below) Ratio of obtained samples producing stable amplification, stable monotonic decay, and instability.
 }
@@
 \begin{split}
-\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\
- \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
+\mathcal{X}(5\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}, 
+\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~ 5\text{Hz}~~~~~~~\mathcal{X}(10\text{Hz}) ~~:~~  \mathbb{E}_{\mathbf{z}, 
+\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~ 10\text{Hz}  \\ 
+ \text{Var}_{\mathbf{z}, 
+\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  1 \text{Hz}^2~~~~~~~~~~~~~~~~~~~~~~~ \text{Var}_{\mathbf{z}, 
+\mathbf{x}}\begin{bmatrix} s_E(\mathbf{x}; \mathbf{z}) \end{bmatrix}  ~~=~~  1 \text{Hz}^2,
 \end{split}
@@
 \subsection{Emergent property inference (EPI)}\label{methods_EPI}
-Solving inverse problems is an important part of theoretical neuroscience, since we must understand how neural circuit models and their parameter choices can produce computations of varying levels of complexity.
-While much machine learning research has focused on how to find latent structure in large-scale neural datasets, less has focused on inverting theoretical circuit models with respect to their computational properties.
+Solving inverse problems is an important part of theoretical neuroscience, since we must understand how neural circuit models and their parameter choices produce computations.
+Recently, research on machine learning methodology for neuroscience has focused on finding latent structure in large-scale neural datasets, while research in theoretical neuroscience generally focuses on developing precise neural circuit models that can produce computations of interest.
+By quantifying computation into an \emph{emergent property} through statistics of the emergent activity of neural circuit models, we can adapt the modern technique of deep probabilistic inference towards solving inverse problems in theoretical neuroscience.
 Here, we introduce a novel method for statistical inference, which uses deep networks to learn parameter distributions constrained to produce emergent properties of computation.
@@
  \begin{equation} \label{eq:methods_ep_general}
-\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\mathcal{X}:~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu},~~ \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
 \end{equation}
@@
 In the STG example, intermediate hub frequency is defined by mean and variance constraints on the statistic of hub neuron frequency $\omega_{\text{hub}}(\mathbf{x}; \mathbf{z})$ (Equations \ref{eq:EP_STG1} and \ref{eq:EP_STG2}).
-Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$ over the EPI distribution of parameters ($\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$) and the data produced by those parameters ($\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$), where the learned parameter distribution $q_{\bm{\theta}}(z)$ is determined by deep network weights and biases $\bm{\theta}$.
-
-In EPI, deep probability distributions are optimized to approximate the inferred distribution.
+Precisely, the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ must have means $\bm{\mu}$ and variances $\bm{\sigma}^2$ over the EPI distribution of parameters ($\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$) and the data produced by those parameters ($\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})$), where the inferred parameter distribution $q_{\bm{\theta}}(\mathbf{z})$ itself is parameterized by deep network weights and biases $\bm{\theta}$.
+
+In EPI, a deep probability distribution $q_{\bm{\theta}}(\mathbf{z})$ is optimized to approximate the parameter distribution producing the emergent property $\mathcal{X}$.
+In contrast to simpler classes of distributions like the gaussian or mixture of gaussians, deep probability distributions are far more flexible and capable of fitting rich structure \cite{rezende2015variational, papamakarios2019normalizing}.
  In deep probability distributions, a simple random variable $\mathbf{z}_0 \sim q_0(\mathbf{z}_0)$ (we choose an isotropic gaussian) is mapped deterministically via a sequence of deep neural network layers ($g_1$, ..
@@
 By maximizing the entropy of the inferred distribution $q_{\bm{\theta}}$, we select the most random distribution in family $\mathcal{Q}$ that satisfies the constraints of the emergent property.
-Because entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}).
+Since entropy is maximized in Equation \ref{eq:opt}, EPI is equivalent to bayesian variational inference (see Section \ref{methods_VI}), which is why we specify the inferred distribution of EPI as conditioned upon emergent property $\mathcal{X}$ with the notation $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$.
 \q{AL_standard}{
@@
 This first neural network is a deep probability distribution (normalizing flow) used to estimate the posterior $p(\mathbf{z} \mid \mathbf{x})$ (SNPE) or the likelihood  $p(\mathbf{x} \mid \mathbf{z})$ (sequential neural likelihood (SNL) \cite{papamakarios2019sequential}).
-A recent advance uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE) \cite{hermans2020likelihood}).
+A recent approach uses an unconstrained neural network to estimate the likelihood ratio (sequential neural ratio estimation (SNRE) \cite{hermans2020likelihood}).
 In SNL and SNRE, MCMC sampling techniques are used to obtain samples from the approximated posterior.
@@
 The second neural network in this sequential class of methods is the amortizer.
-This unconstrained deep network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$ to the weights and biases of the first neural network.
+This unconstrained deep network maps data $\mathbf{x}$ (or statistics $f(\mathbf{x}; \mathbf{z})$ or model parameters $\mathbf{z}$) to the weights and biases of the first neural network.
 These methods are optimized on a conditional density (or ratio) estimation objective.
@@
 Whereas in EPI, the entire expressivity of the deep probability distribution is dedicated to learning a single distribution as well as possible.
-Amortization is not possible in EPI, since EPI learns an exponential family distribution parameterized by its mean (in contrast to its natural parameter, see Section \ref{methods_ME_EF}).}
-The well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} prohibits an amortization based approach in EPI, since emergent properties are defined via the mean parameter of the exponential family distribution.
+The well-known inverse mapping problem of exponential families \cite{wainwright2008graphical} prohibits an amortization-based approach in EPI, since EPI learns an exponential family distribution parameterized by its mean (in contrast to its natural parameter, see Section \ref{methods_ME_EF}).}
 However, we have shown that the same two-network architecture of the sequential simulation-based inference methods can be used for amortized inference in intractable exponential family posteriors when using their natural parameterization \cite{bittner2019approximating}.
@@
 \begin{equation} \label{eq:moments}
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} f(\mathbf{x}; \mathbf{z}) \\ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2 \end{bmatrix},
+T(\mathbf{z}) = \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ f(\mathbf{x}; \mathbf{z}) \right] \\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \left( f(\mathbf{x}; \mathbf{z}) - \bm{\mu} \right)^2\right] \end{bmatrix},
 \end{equation}
@@
 Thus, $\bm{\mu}_{\text{opt}}$ is used to denote the mean parameter of the maximum entropy distribution defined by the emergent property (all constraints), while $\bm{\mu}$ is only the mean of $f(\mathbf{x}; \mathbf{z})$.
-The subscript ``$\text{opt}$" of $\bm{\mu}_{\text{opt}}$ is chosen since it contains all of the contraint values to which the optimization algorithm must adhere.
+The subscript ``$\text{opt}$" of $\bm{\mu}_{\text{opt}}$ is chosen since it contains all of the constraint values to which the EPI optimization algorithm must adhere.
 
@@
 \begin{equation}
-R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right],
+R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})}\left[T(\mathbf{z}) - \bm{\mu}_{\text{opt}} \right],
 \end{equation}
@@
 \begin{equation}
-m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|,
+m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|,
 \end{equation}
 and $c$ is the penalty coefficient. 
-The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{x}; \mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
-Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments (Equation \ref{eq:moments}) and $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (Equation \ref{eq:mu_opt}).
-(Although, note that this algorithm is written for general $T(\mathbf{x}; \mathbf{z})$ and $\bm{\mu}$ to satisfy the more general class of emergent properties.)
+The mean parameter $\bm{\mu}_{\text{opt}}$ and sufficient statistics $T(\mathbf{z})$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
+Specifically, $T(\mathbf{z})$ is a concatenation of the first and second moments (Equation \ref{eq:moments}) and $\bm{\mu}_{\text{opt}}$ is a concatenation of their constraints $\bm{\mu}$ and $\bm{\sigma}^2$ (Equation \ref{eq:mu_opt}).
+(Although, note that this algorithm is written for general $T(\mathbf{z})$ and $\bm{\mu}_{\text{opt}}$ to satisfy the more general class of emergent properties.)
 The lagrange multipliers $\bm{\eta}_{\text{opt}}$ are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
@@
 \end{equation}
-Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average with respect to the base distribution $\mathbf{z}_0$:
+Thus, the gradient of the entropy of the deep probability distribution can be estimated as an average of gradients with respect to the base distribution $\mathbf{z}_0$:
 \begin{equation}
@@
 \end{equation}
+The gradients of the log density of the deep probability distribution are tractable through the use of normalizing flows (see Section \ref{methods_NF}).
 
 The full EPI optimization algorithm is detailed in Algorithm \ref{alg:epi}.
-The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations.
+The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic gradient descent (SGD) iterations.
 A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
@@
 The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
-The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
+The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{z}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$ and $\mathbf{z}^{(i)} \sim q_{\bm{\theta}}(\mathbf{z})$.
 Throughout the study, $\gamma = 0.25$, while $\beta$ was chosen to be either 2 or 4.  The batch size of EPI also varied according to application.
@@
     \begin{split}
-    \nabla_{\bm{\theta}} L(\bm{\theta}; \bm{\eta}, c) &= \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} \log q_{\bm{\theta}}(\mathbf{z}^{(j)}) + \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)}) \bm{\eta}_{\text{opt}} \\ &+ c_k \frac{2}{n} \sum_{j=1}^{\frac{n}{2}} \nabla_{\bm{\theta}} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)}) \cdot \frac{2}{n} \sum_{j=1 + \frac{n}{2}}^{n} T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)})
+    \nabla_{\bm{\theta}} L(\bm{\theta}; \bm{\eta}_{\text{opt}, k}, c) &= \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} \log q_{\bm{\theta}}(\mathbf{z}^{(j)}) 
+    + \frac{1}{n}\sum_{j=1}^n \nabla_{\bm{\theta}} \left(T\left(\mathbf{z}^{(j)}\right) - \bm{\mu}_{\text{opt}} \right) \bm{\eta}_{\text{opt}, k} \\ 
+    &+ c_k \frac{2}{n} \sum_{j=1}^{\frac{n}{2}} \nabla_{\bm{\theta}} \left(T\left(\mathbf{z}^{(j)}\right) - \bm{\mu}_{\text{opt}} \right) \cdot \frac{2}{n} \sum_{j=\frac{n}{2}+1}^{n} \left(T\left(\mathbf{z}^{(j)}\right) - \bm{\mu}_{\text{opt}} \right)
     \end{split}
@@
  Sample $\mathbf{z}_0^{(1)}, ..., \mathbf{z}_0^{(n)} \sim q_0$, get transformed variable $\mathbf{z}^{(j)} = g_{\bm{\theta}}(\mathbf{z}_0^{(j)})$, $j = 1, ..., n$\\
- Update $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{j=1}^n T(\mathbf{x}^{(j)}; \mathbf{z}^{(j)})$. \\
+ Update $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{j=1}^n \left(T\left(\mathbf{z}^{(j)}\right) - \bm{\mu}_{\text{opt}} \right)$. \\
  Update $c_{k+1} > c_k$ (see text for detail). \\
@@
 In general, $c$ and $\bm{\eta}_{\text{opt}}$ should start at values encouraging entropic growth early in optimization.  
-With each training epoch in which the update rule for $c$ is invoked by unsatisfactory constraint error reduction, the constraint satisfaction terms are increasingly weighted, resulting in a decreased entropy.
-This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property (see Figure \ref{fig:LDS1}).
+With each training epoch in which the update rule for $c$ is invoked, the constraint satisfaction terms are increasingly weighted, which generally results in decreased entropy (e.g.
+see Figure \ref{fig:LDS1}C).
+This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
 The momentum parameters of the Adam optimizer are reset at the end of each augmented lagrangian epoch, which proceeds for $i_{\text{max}}$ iterations.
@@
 The parameters of the gaussian initialization were chosen on an application-specific basis.  
-Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the distribution support and some variance $\bm{\sigma}_{\text{init}}^2$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
+Throughout the study, we chose isotropic Gaussian initializations with mean $\bm{\mu}_{\text{init}}$ at the center of the support of the distribution and some variance $\bm{\sigma}_{\text{init}}^2$, except for one case, where an initialization informed by random search was used (see Section \ref{methods_STG}).
 Deep probability distributions were fit to these gaussian initializations using 10,000 iterations of stochastic gradient descent on the evidence lower bound (as in \cite{bittner2019approximating}) with Adam optimizer and a learning rate of $10^{-3}$.
@@
 \end{equation}
-To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_{1,1}, a_{1,2}, a_{2,1}, a_{2,2}]$ (fixing $\tau=1$s), the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ were chosen to contain parts of the primary eigenvalue of $A$, which predicate frequency, $\text{imag}(\lambda_1)$, and the growth/decay, $\text{real}(\lambda_1)$, of the system.
+To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_{1,1}, a_{1,2}, a_{2,1}, a_{2,2}]$ (fixing $\tau=1$s), the emergent property statistics $f(\mathbf{x}; \mathbf{z})$ were chosen to contain parts of the primary eigenvalue of $A$, which predicate frequency, $\text{imag}(\lambda_1)$, and the growth/decay, $\text{real}(\lambda_1)$, of the system
+\begin{equation}
+f(\mathbf{x}; \mathbf{z}) \triangleq \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix}
+\end{equation}
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively that of positive imaginary component when the eigenvalues are complex conjugate pairs.  
-To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean 1Hz with variance 0.1Hz$^2$:
+To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $\frac{\text{imag}(\lambda_1)}{2 \pi}$ having mean 1Hz with variance 0.1Hz$^2$:
 \begin{equation}
 \begin{split}
-\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0 \\ 2\pi \end{bmatrix} \\
-\text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix}.
+\mathcal{X} ~~:~~\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] \triangleq \mathbb{E}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0 \\ 2\pi \end{bmatrix} \triangleq \bm{\mu} \\
+\text{Var}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] \triangleq \text{Var}_{\mathbf{z}, \mathbf{x}} \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \end{bmatrix} &= \begin{bmatrix} 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix} \triangleq \bm{\sigma}^2.
 \end{split}
@@
 
-To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), we concatenate these first and second moment constraints into a vector of sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and constraint values $\bm{\mu}_{\text{opt}}$.
-
-\begin{equation}
- \mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[T(\mathbf{x}; \mathbf{z}) \right] ~~ \triangleq ~~ \mathbb{E}_{\mathbf{z}, \mathbf{x}}  \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ (\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z})-0)^2  \\ (\text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z})-2 \pi)^2 \end{bmatrix} = \begin{bmatrix} 0 \\ 2 \pi \\ 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}_{\text{opt}}.
- \end{equation} 
+To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), we concatenate these first and second moment constraints into a vector of sufficient statistics $T(\mathbf{z})$ and constraint values $\bm{\mu}_{\text{opt}}$.
+\begin{equation}
+ \mathbb{E}_{\mathbf{z}} \left[T(\mathbf{z}) \right] ~~ \triangleq ~~ \mathbb{E}_{\mathbf{z}}  \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \right] \\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z}) \right] \\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ (\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z})-0)^2 \right]  \\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ (\text{imag}(\lambda_1)(\mathbf{x}; \mathbf{z})-2 \pi)^2 \right] \end{bmatrix} = \begin{bmatrix} 0 \\ 2 \pi \\ 0.25^2 \\ (\frac{\pi}{5})^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}_{\text{opt}}.
+ \end{equation} 
+ From now on in all scientific applications (Sections \ref{methods_STG}-\ref{methods_SC}, we specify how the EPI optimization was setup by specifying $f(\mathbf{x}; \mathbf{z})$, $\bm{\mu}$, and $\bm{\sigma}^2$.
+
 \begin{figure}
@@
 Despite the simple analytic form of the emergent property statistics, the EPI distribution in this example is not simply determined.
-Although $\mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z})\right]$ is calculable directly via a closed form function, the distribution $q^*_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ cannot be derived directly.
-This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\eta$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
+Although $\mathbb{E}_{\mathbf{z}}\left[T(\mathbf{z})\right]$ is calculable directly via a closed form function, the distribution $q^*_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ cannot be derived directly.
+This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\bm{\eta}$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
 Instead, we used EPI to approximate this distribution (Fig. \ref{fig:LDS1}B). 
@@
 \end{equation}
-In Figure \ref{fig:LDS2}B, we plot the contours of $\text{imag}(\lambda_1)$ where $a_{1,1} a_{2,2}$ is fixed to 0 at one standard devation ($\frac{\pi}{5}$, black dashed) and two standard deviations ($\frac{2\pi}{5}$, gray dashed) from the mean of $2 \pi$.
+In Figure \ref{fig:LDS2}B, we plot the contours of $\text{imag}(\lambda_1)$ where $a_{1,1} a_{2,2}$ is fixed to 0 at one standard deviation ($\frac{\pi}{5}$, black dashed) and two standard deviations ($\frac{2\pi}{5}$, gray dashed) from the mean of $2 \pi$.
 This validates the curved multimodal structure of the inferred distribution learned through EPI.  
@@
 \begin{split}
-\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) &\Longleftrightarrow \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \bm{\eta}^{*\top} \mathbb{E}_{\mathbf{z} \sim q} \left[ \phi(\mathbf{z}) \right], \\
-  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim q}\left[\phi(\mathbf{z})\right] = \mathbf{0}~& \\
+\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) &\Longleftrightarrow \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \bm{\eta}^{*\top} \mathbb{E}_{\mathbf{z} \sim q} \left[ T(\mathbf{z}) \right], \\
+  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim q}\left[T(\mathbf{z})\right] = \mathbf{0}~& \\
  \end{split}
 \end{equation} 
-with sufficient statistics $\phi(\cdot)$ and lagrange multipliers $\bm{\eta}^*$.
+with sufficient statistics $T(\cdot)$ and lagrange multipliers $\bm{\eta}^*$.
 By moving the lagrange multipliers within the expectation
-\begin{equation}q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\bm{\eta}^{*\top} \phi(\mathbf{z}) \right],
+\begin{equation}q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\bm{\eta}^{*\top} T(\mathbf{z}) \right],
 \end{equation} 
 inserting a $\log \exp(\cdot)$ within the expectation, 
-\begin{equation} q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} \phi(\mathbf{z}) \right) \right],
-\end{equation}
-and finally replacing $\phi(\cdot)$ with the sufficient statistics of the EPI optimization
+\begin{equation} q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} T(\mathbf{z}) \right) \right],
+\end{equation}
+and finally choosing $T(\cdot)$ to be likelihood averaged statistics as in EPI
 \begin{equation} \label{eq:epi_vi}
-q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T_1(\mathbf{x}; \mathbf{z}) \right] \\ ...
-\\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T_m(\mathbf{x}; \mathbf{z}) \right] \end{bmatrix} \right) \right],
-\end{equation}
-we can then compare to the objective used in variational inference  (Equation \ref{eq:VI_opt}).
+q^* = \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \mathbb{E}_{\mathbf{z} \sim q} \left[\log \exp \left( \bm{\eta}^{*\top} \begin{bmatrix} \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \phi_1(\mathbf{x}; \mathbf{z}) \right] \\ ...
+\\ \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \phi_m(\mathbf{x}; \mathbf{z}) \right] \end{bmatrix} \right) \right],
+\end{equation}
+we can compare directly to the objective used in variational inference  (Equation \ref{eq:VI_opt}).
 \q{EPI_VI}{
-We see that EPI is exactly variational inference with an exponential family likelihood defined by sufficient statistics $\phi(z) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ T(\mathbf{x}; \mathbf{z}) \right]$, and where the natural parameter $\bm{\eta}^*$ is predicated by the mean parameter $\bm{\mu}_{\text{opt}}$.
+We see that EPI is exactly variational inference with an exponential family likelihood defined by sufficient statistics $T(\mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})} \left[ \phi(\mathbf{x}; \mathbf{z}) \right]$, and where the natural parameter $\bm{\eta}^*$ is predicated by the mean parameter $\bm{\mu}_{\text{opt}}$.
 Equation \ref{eq:epi_vi} implies that EPI uses an improper (or uniform) prior, which is easily changed.}
@@
 This derivation of the equivalence between EPI and variational inference emphasizes why defining a statistical inference program by its mean parameterization $\bm{\mu}_{\text{opt}}$ is so useful.
-With EPI, one can clearly define the emergent property $\mathcal{X}$ that the model of interest should produce through intuitive selection of $\bm{\mu}_{\text{opt}}$ for a given $T(\mathbf{x}; \mathbf{z})$.
-Alternatively, figuring out the correct natural parameters $\bm{\eta}^*$ for the same $T(\mathbf{x}; \mathbf{z})$ that produces $\mathcal{X}$ will be generally unclear.
+With EPI, one can clearly define the emergent property $\mathcal{X}$ that the model of interest should produce through intuitive selection of $\bm{\mu}_{\text{opt}}$ for a given $T(\mathbf{z})$.
+Alternatively, figuring out the correct natural parameters $\bm{\eta}^*$ for the same $T(\mathbf{z})$ that produces $\mathcal{X}$ will be generally unclear.
 
@@
 \subsubsection{EPI details for the STG model} \label{methods_stg_epi}
-To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the hub neuron frequency (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) \\ \left( \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}) - 0.55 \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 0.55 \\ 0.025^2 \end{bmatrix}.
-\end{equation}
+EPI was run for the STG model using
+\begin{equation}
+f(\mathbf{x}; \mathbf{z}) = \omega_{\text{hub}}(\mathbf{x}; \mathbf{z}),
+\end{equation}
+\begin{equation}
+\bm{\mu} = \begin{bmatrix} 0.55 \end{bmatrix},
+\end{equation}
+and
+\begin{equation}
+\bm{\sigma}^2 = \begin{bmatrix} 0.025^2 \end{bmatrix}
+\end{equation}
+(see Sections \ref{methods_ME_EF}-\ref{methods_AL_opt}, and example in Section \ref{methods_2DLDS}).
 Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 5,000$ iterations (see Section \ref{methods_AL_opt}).  
@@
 We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
-EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, and $\beta = 2$.
+EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{5}$, $\beta = 2$, a batch size $n=400$, and we simulated one $\mathbf{x}^{(i)}$ per $\mathbf{z}^{(i)}$.
 The architecture converged with criteria $N_{\text{test}} = 100$.
@@
 To complete the emergent property definition, we chose variances ($0.25^2$) about those means such that samples rarely violate the eigenvalue constraints.
-To write the emergent property of Equation \ref{eq:EP_LRRNN} in terms of the EPI augmented lagrangian optimization, this is written as
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \lambda_1^s(\mathbf{x}; \mathbf{z}) \\ \left(\text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) - 0.5 \right)^2 \\ \left( \lambda_1^s(\mathbf{x}; \mathbf{z})  - 1.5 \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} 0.5 \\ 1.5 \\ 0.25^2 \\ 0.25^2 \end{bmatrix}.
-\end{equation}
+To write the emergent property of Equation \ref{eq:EP_LRRNN} in terms of the EPI optimization, we have
+\begin{equation} 
+f(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} \text{real}(\lambda_1)(\mathbf{x}; \mathbf{z}) \\ \lambda_1^s(\mathbf{x}; \mathbf{z})  \end{bmatrix},
+\end{equation}
+\begin{equation} 
+\bm{\mu} = \begin{bmatrix} 0.5 \\ 1.5 \end{bmatrix},
+\end{equation}
+and
+\begin{equation} 
+\bm{\sigma}^2 = \begin{bmatrix} 0.25^2 \\ 0.25^2 \end{bmatrix}
+\end{equation}
+(see Sections \ref{methods_ME_EF}-\ref{methods_AL_opt}, and example in Section \ref{methods_2DLDS}).
 Gradients of maximum eigenvalues of Hermitian matrices like $W^s$ are available with modern automatic differentiation tools.
@@
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
-We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and chose to use $i_{\text{max}} = 500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
+We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and we simulated one $\mathbf{W}^{(i)}$ per $\mathbf{z}^{(i)}$.
+We chose to use $i_{\text{max}} = 500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
 \ref{fig:RNN}B blue line, and Fig.
@@
 While EPI has its own hypothesis testing convergence criteria for the emergent property, it would not make sense to use this criteria on SNPE and SMC-ABC which do not constrain the means and variances of their predictions.
-Instead, we consider EPI and SNPE to have converged after completing its most recent optimization epoch (EPI) or round (SNPE) in which the distance
-\begin{equation}
-d(q_{\bm{\theta}}(\mathbf{z})) = |\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}|_2
-\end{equation}
-is less than 0.5.
+Instead, we consider EPI and SNPE to have converged after completing its most recent optimization epoch (EPI) or round (SNPE) in which the distance $|\mathbb{E}_{\mathbf{z}, \mathbf{x}} \left[f(\mathbf{x}; \mathbf{z}) \right] - \bm{\mu}|_2$ is less than 0.5.
 We consider SMC-ABC to have converged once the population produces samples within the $\epsilon = 0.5$ ball ensuring stable amplification.
@@
 Circuit responses are simulated using $T = 200$ time steps at $dt = 0.5\text{ms}$ from an initial condition drawn from $\mathbf{x}(0) \sim U\left[10\text{Hz}, 25\text{Hz}\right]$.
-Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$ averaged over 100 independent trials.
-\begin{equation}
-s_E(\mathbf{x}; \mathbf{z}) =\mathbb{E}_{x}\left[\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]}\right]
+Standard deviation of the E-population $s_E(\mathbf{x}; \mathbf{z})$ is calculated as the square root of the temporal variance from $t_{ss} = 75\text{ms}$ to $Tdt = 100\text{ms}$
+\begin{equation}
+s_E(\mathbf{x}; \mathbf{z}) =\sqrt{\mathbb{E}_{t > t_{ss}}\left[\left(x_E(t) -\mathbb{E}_{t > t_{ss}}\left[ x_E(t) \right]\right)^2 \right]} .
 \end{equation}
@@
 \subsubsection{EPI details for the V1 model}
+To write the emergent properties of Equation \ref{eq:EP_V1} in terms of the EPI optimization, we have
+\begin{equation} 
+f(\mathbf{x}; \mathbf{z}) = s_E(\mathbf{x}; \mathbf{z}),
+\end{equation}
+\begin{equation} 
+\bm{\mu} = \begin{bmatrix} 5 \end{bmatrix}
+\end{equation}
+(or $\bm{\mu} = \begin{bmatrix} 10 \end{bmatrix}$), and
+\begin{equation} 
+\bm{\sigma}^2 = \begin{bmatrix} 1^2 \end{bmatrix}
+\end{equation}
+(see Sections \ref{methods_ME_EF}-\ref{methods_AL_opt}, and example in Section \ref{methods_2DLDS}).
+
 For EPI in Figures \ref{fig:V1}D-E and \ref{fig:V1_1}, we used a real NVP architecture with three coupling layers and two-layer neural networks of 50 units per layer.
 The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
-EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{-1}$, a batch size $n=100$, $\beta = 2$, and $i_{\text{max}} = 2,000$ iterations per epoch.
+EPI optimization was run using three different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{-1}$, $\beta = 2$, a batch size $n=100$, and simulated 100 trials to calculate average $s_E(\mathbf{x}; \mathbf{z})$ for each $\mathbf{z}^{(i)}$.
+We used $i_{\text{max}} = 2,000$ iterations per epoch.
 The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across three random seeds.
@@
 \subsubsection{Task accuracy calculation} \label{methods_sc_acc}
-The accuracies of each Pro and Anti tasks are calculated as
-\begin{equation}
-p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x}}\left[\Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]\right]
+The accuracies of the Pro and Anti tasks are calculated as
+\begin{equation}
+p_P(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x} \sim p(\mathbf{x} \mid \mathbf{z})}\left[d_P(\mathbf{x}; \mathbf{z})\right]
 \end{equation}
@@
 \begin{equation}
-p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x}}\left[\Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]\right]
-\end{equation}
-given that the stimulus is on the left side and $\Theta$ approximates the Heaviside step function.
+p_A(\mathbf{x}; \mathbf{z}) = \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[d_A(\mathbf{x}; \mathbf{z}) \right]
+\end{equation}
+where $d_P(\mathbf{x}; \mathbf{z})$ and $d_A(\mathbf{x}; \mathbf{z})$ calculate the decision made in each trial (approximately 1 for correct and 0 for incorrect choices).
+Specifically, 
+\begin{equation}
+d_P(\mathbf{x}; \mathbf{z}) = \Theta[x_{LP}(t=1.8s) - x_{RP}(t=1.8s)]
+\end{equation}
+in Pro trials where the stimulus is on the left side, and $\Theta$ approximates the Heaviside step function.
+Similarly, 
+\begin{equation}
+d_A(\mathbf{x}; \mathbf{z}) = \Theta[x_{RP}(t=1.8s) - x_{LP}(t=1.8s)]
+\end{equation}
+in Anti trials where the stimulus was on the left side.
 Our accuracy calculation only considers one stimulus presentation (Left), since the model is left-right symmetric.
@@
 \subsubsection{EPI details for the SC model} \label{methods_sc_epi}
-To write the emergent property $\mathcal{X}$ in the form required for the augmented lagrangian optimization (Section \ref{methods_AL_opt}), $T(\mathbf{x}, \mathbf{z})$ is comprised of both these first and second moments of the accuracy in each task (as in Equations \ref{eq:moments} and \ref{eq:mu_opt})
-\begin{equation} 
-T(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \\ \left(p_P(\mathbf{x}; \mathbf{z}) - .75 \right)^2 \\ \left(p_A(\mathbf{x}; \mathbf{z}) - .75 \right)^2 \end{bmatrix},
-\end{equation}
-\begin{equation} 
-\bm{\mu}_{\text{opt}} = \begin{bmatrix} .75 \\ .75 \\ .075^2 \\ .075^2 \end{bmatrix}.
-\end{equation}
+To write the emergent properties of Equation \ref{eq:SC_EP} in terms of the EPI optimization, we have
+\begin{equation} 
+f(\mathbf{x}; \mathbf{z}) = \begin{bmatrix} d_P(\mathbf{x}; \mathbf{z}) \\ d_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}
+\end{equation}
+\begin{equation} 
+\bm{\mu} = \begin{bmatrix} .75 \\ .75 \end{bmatrix},
+\end{equation}
+and
+\begin{equation} 
+\bm{\sigma}^2 = \begin{bmatrix} .075^2 \\ .075^2 \end{bmatrix}
+\end{equation}
+(see Sections \ref{methods_ME_EF}-\ref{methods_AL_opt}, and example in Section \ref{methods_2DLDS}).
+
 Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 2,000$ iterations (see Section \ref{methods_AL_opt}).  

### 2021-06-04 written/eLifeRevision/Bittner_eLife2020.tex
final edits

@@
 
-\q{intro_data_ep}{Statistical inference, of course, requires quantification of the vague term \emph{computation}.  
+\q{intro_data_ep}{Statistical inference, of course, requires quantification of the sometimes vague term \emph{computation}.  
 In neuroscience, two perspectives are dominant.  
@@
 The parameter distributions producing a computation may be curved or multimodal along various parameter axes and combinations.
-It is by quantifying this complex structure that EPI offers scientific insight.
+It is by quantifying this complex structure that emergent property inference offers scientific insight.
 Traditional approximation families (e.g. mean-field or mixture of gaussians) are limited in the distributional structure they may learn.
@@
 \q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
-Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
+Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
 \q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
@@
 \end{center}
-\caption{Emergent property inference (EPI) in the stomatogastric ganglion.  
+\caption{Emergent property inference in the stomatogastric ganglion.  
 \textbf{A}. Conductance-based subcircuit model of the STG.
@@
 Simulated activity traces are colored by log probability of generating parameters in the EPI distribution (Panel E).
-\textbf{D}. For a choice of circuit model and emergent property, emergent property inference (EPI) learns a deep probability distribution of parameters $\mathbf{z}$.
+\textbf{D}. For a choice of circuit model and emergent property, EPI learns a deep probability distribution of parameters $\mathbf{z}$.
 \textbf{E}. The EPI distribution producing intermediate hub frequency.
@@
 Third, the model parameters producing the emergent property are inferred.
-By precisely quantifying the emergent property of interest as a statistical feature of the model, we use EPI to condition directly on this emergent property.
+By precisely quantifying the emergent property of interest as a statistical feature of the model, we use emergent property inference (EPI) to condition directly on this emergent property.
 Before presenting technical details (in the following section), let us understand emergent property inference schematically.
@@
 \subsection{Emergent property inference via deep generative models} \label{results_dgm}
-Emergent property inference (EPI) formalizes the three-step procedure of the previous section with deep probability distributions \cite{rezende2015variational, papamakarios2019normalizing}.
+EPI formalizes the three-step procedure of the previous section with deep probability distributions \cite{rezende2015variational, papamakarios2019normalizing}.
 First, as is typical, we consider the model as a coupled set of noisy differential equations.  
@@
 \textbf{E}.
-(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz})$.
+(Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz}))$.
 Each black dot shows the mode at each $\sigma_P$.
@@
 \textbf{F}.
-The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz})$.
+The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz}))$.
 }
@@
 \subsubsection{EPI as variational inference}\label{methods_VI}
-%Now that we have fully described the EPI method, we  consider its broader contextualization as a statistical method and its relation to Bayesian inference. 
-In bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood $p(\mathbf{x} \mid \mathbf{z})$.
-We obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
-\begin{equation}
-p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
-\end{equation}
-The posterior distribution is analytically available when the prior is conjugate with the likelihood.
-However, conjugacy is rare in practice, and alternative methods such as variational inference \cite{saul1998mean} are utilized.
+%In bayesian inference a prior belief about model parameters $\mathbf{z}$ is stated in a prior distribution $p(\mathbf{z})$, and the statistical model capturing the effect of $\mathbf{z}$ on observed data points $\mathbf{x}$ is formalized in the likelihood $p(\mathbf{x} \mid \mathbf{z})$.
+%We obtain a posterior distribution $p(\mathbf{z} \mid \mathbf{x})$, which captures how the data inform our knowledge of model parameters using Bayes' rule:
+%\begin{equation}
+%p(\mathbf{z} \mid \mathbf{x}) = \frac{p(\mathbf{x} \mid \mathbf{z}) p(\mathbf{z})}{p(\mathbf{x})}.
+%\end{equation}
+%The posterior distribution is analytically available when the prior is conjugate with the likelihood.
+%However, conjugacy is rare in practice, and alternative methods such as variational inference \cite{saul1998mean} are utilized.
 
@@
 \begin{split}
-\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) &\Longleftrightarrow \argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \bm{\eta}^{*\top} \mathbb{E}_{\mathbf{z} \sim q} \left[ T(\mathbf{z}) \right], \\
-  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim q}\left[T(\mathbf{z})\right] = \mathbf{0}~& \\
+\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) ~~~~~&\Longleftrightarrow ~~~~~\argmax_{q \in \mathcal{Q}} H(q(\mathbf{z})) + \bm{\eta}^{*\top} \mathbb{E}_{\mathbf{z} \sim q} \left[ T(\mathbf{z}) \right], \\
+  \text{s.t.  } \mathbb{E}_{\mathbf{z} \sim q}\left[T(\mathbf{z})\right] = \mathbf{0}~~~~~& \\
  \end{split}
 \end{equation} 
-with sufficient statistics $T(\cdot)$ and lagrange multipliers $\bm{\eta}^*$.
+with lagrange multipliers $\bm{\eta}^*$.
 By moving the lagrange multipliers within the expectation
@@
 With EPI, one can clearly define the emergent property $\mathcal{X}$ that the model of interest should produce through intuitive selection of $\bm{\mu}_{\text{opt}}$ for a given $T(\mathbf{z})$.
-Alternatively, figuring out the correct natural parameters $\bm{\eta}^*$ for the same $T(\mathbf{z})$ that produces $\mathcal{X}$ will be generally unclear.
+Alternatively, figuring out the correct natural parameters $\bm{\eta}^*$ for the same $T(\mathbf{z})$ that produces $\mathcal{X}$ is a formally hard problem.
 

### 2021-06-04 written/eLifeRevision/Response.tex
final edits
updated line numbers

@@
 \begin{displayquote}
-Lines 46-54\\
+Lines 44-52\\
 ``\quoting{intro_data_ep}''
@@
 \begin{displayquote}
-Lines 55-65\\
+Lines 53-63\\
 ``\quoting{intro_epi}''
@@
 \begin{displayquote}
-Line 759 \\
+Line 739 \\
 ``
-\begin{equation} \tag{11}
-\mathcal{X}: \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu}, \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2.
+\begin{equation} \tag{12}
+\mathcal{X}:~ \mathbb{E}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\mu},~~ \text{Var}_{\mathbf{z},\mathbf{x}}\left[f(\mathbf{x}; \mathbf{z})\right] = \bm{\sigma}^2
 \end{equation}''
@@
 \begin{displayquote}
-Lines 156-158\\
+Lines 152-154\\
 ``\quoting{results_ME}''
@@
 \begin{displayquote}
-Lines 439-451\\
+Lines 432-444\\
 ``\quoting{discussion_gradients}''
@@
 \begin{displayquote}
-Lines 828-833\\
+Lines 829-835\\
 ``\quoting{amortization}''
@@
 \begin{displayquote}
-Lines 232-242\\
+Lines 227-237\\
 ``\quoting{rnn_constrained}''
@@
 \begin{displayquote}
-Lines 77-81 \\
+Lines 75-79 \\
 ``\quoting{intro_results1}''
@@
 \begin{displayquote}
-Lines 81-89 \\
+Lines 79-87 \\
 ``\quoting{intro_results2}''
@@
 \begin{displayquote}
-Lines 301-309 \\
+Lines 295-303 \\
 ``\quoting{v1_summary}''
@@
 \begin{displayquote}
-Lines 416-419 \\
+Lines 409-412 \\
 ``\quoting{sc_summary}''
@@
 \begin{displayquote}
-Lines 100-105 \\
+Lines 98-103 \\
 ``\quoting{stg_subcircuit}''
@@
 \begin{displayquote}
-Lines 444-447 \\
+Lines 437-440 \\
 ``However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
 The pyloric subnetwork requires many time steps for simulation and many key emergent property statistics (e.g.
-burst duration and phase gap) are not calculable or easily approximated with differentiable functions.\\
+burst duration and phase gap) are not calculable or easily approximated with differentiable functions.
 ''
@@
 All code for training EPI on these models and their analysis are available in well-documented scripts and notebooks in our github repository.
+\begin{verbatim} https://github.com/cunningham-lab/epi \end{verbatim}
 
@@
 \begin{displayquote}
-Lines 28-37 \\
+Lines 26-35 \\
 ``\quoting{intro_first}''
@@
 \begin{displayquote}
-Lines 1026-1029 \\
+Lines 1021-1025 \\
 ``\quoting{EPI_VI}''
@@
 \begin{displayquote}
-Lines 175 -177 \\
+Lines 171 -173 \\
 ``The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.''
@@
 \begin{displayquote}
-Section \ref{results_dgm}, Lines 139-143\\
+Section \ref{results_dgm}, Lines 136-140\\
 ``\quoting{stg_ep_text1}~\quoting{stg_ep_text2}.''
@@
 \begin{displayquote}
-Section \ref{results_RNN}, Lines 204-210\\
+Section \ref{results_RNN}, Lines 200-206 \\
 ``\quoting{rnn_ep_text1}~... \quoting{rnn_ep_text2}~\quoting{rnn_ep_text3}''
@@
 \begin{displayquote}
-Section \ref{results_V1}, Lines 280-283\\
+Section \ref{results_V1}, Lines 274-277\\
 ``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}''
@@
 \begin{displayquote}
-Section \ref{results_SC}, Lines 338-341\\
+Section \ref{results_SC}, Lines 329-332\\
 ``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}''
@@
 \begin{displayquote}
-Lines 439-451\\
+Lines 432-444\\
 ``\quoting{discussion_gradients}''
-
 \end{displayquote}
@@
 \begin{displayquote}
-Lines 431-438\\
+Lines 424-444\\
 ``\quoting{discussion_methods} \\ \quoting{discussion_gradients}''
@@
 
-Lines X-Y \\
+Lines 764-766 \\
 \begin{displayquote}
@@
 Please see the github repository for this manuscript, which includes code for the EPI method and the analyses that went into each figure.
+\begin{verbatim} https://github.com/cunningham-lab/epi \end{verbatim}
 

### 2021-06-06 written/eLifeRevision/figures/rename_figs.sh
figure file rename

@@ -1,0 +1,31 @@
+cp fig1/fig1.pdf fig1.pdf
+cp figLDS1/figLDS1.pdf fig1supp1.pdf
+cp figLDS2/figLDS2.pdf fig1supp2.pdf
+cp figLDS3/figLDS3.pdf fig1supp3.pdf
+cp figSTG1/figSTG1.pdf fig1supp4.pdf
+
+cp fig2/fig2.pdf fig2.pdf
+cp figRNN1/figRNN1.pdf fig2supp1.pdf
+cp figRNN2/figRNN2.pdf fig2supp3.pdf
+cp figRNN3/figRNN3.pdf fig2supp2.pdf
+
+cp fig3/fig3.pdf fig3.pdf
+cp figV1_1/figV1_1.pdf fig3supp4.pdf
+cp figV1_2/figV1_2.pdf fig3supp3.pdf
+cp figV1_3/figV1_3.pdf fig3supp1.pdf
+cp figV1_4/figV1_4.pdf fig3supp2.pdf
+
+cp fig4/fig4.pdf fig4.pdf
+cp figSC1/figSC1.pdf fig4supp1.pdf
+cp figSC2/figSC2.pdf fig4supp2.pdf
+cp figSC3/figSC3.pdf fig4supp3.pdf
+
+cp figSC5/figSC5.pdf fig5supp1.pdf
+cp figSC6/figSC6.pdf fig5supp2.pdf
+
+cp figSC7/figSC7.pdf fig4supp4.pdf
+cp figSC8/figSC8.pdf fig4supp5.pdf
+
+git add *.pdf
+#git commit -m "figure filename updates"
+

### 2021-06-08 written/eLifeRevision/Response.tex
rename from 2020 to 2021
Duan 2021 ref update

@@
 \usepackage{xr}
-\externaldocument{Bittner_eLife2020}
+\externaldocument{Bittner_eLife2021}
 \usepackage{xcite}
-\externalcitedocument{Bittner_eLife2020}
+\externalcitedocument{Bittner_eLife2021}
 \usepackage{qting}
-\quotefrom{Bittner_eLife2020}
+\quotefrom{Bittner_eLife2021}
 
@@
 \begin{displayquote}
-Line 739 \\
+Line 742 \\
 ``
@@
 Figure \ref{fig:STG} has been largely redone to reflect the modified presentation, and improve the pictorial representation of the method.
-In Section \ref{results_RNN}, we now show that EPI is the only simulation-based inference method that controls the predictions of its inferred distribution (Fig. \ref{fig:RNN}C-D).
+In Section \ref{results_RNN}, we now show that EPI is the only simulation-based inference method that controls the predictions of its inferred distribution (Figure \ref{fig:RNN}C-D).
 
@@
 \begin{displayquote}
-Lines 163-179\\
+Lines 159-175\\
 ``\quoting{results_hess}''
@@
 In Section 3.3, we provide a direct comparison of EPI to alternative simulation-based inference techniques SMC-ABC and SNPE by inferring RNN connectivities that exhibit stable amplification.
-These comparisons evaluate both wall time (Fig.
-2A) and simulation count (Fig.
-2B), and we explain how each algorithm was run in its preferred hardware setting in Section 5.3.4.
+These comparisons evaluate both wall time (Figure 2A) and simulation count (Figure 2B), and we explain how each algorithm was run in its preferred hardware setting in Section 5.3.4.
 
 In this analysis, we demonstrate the improved scalability of deep inference techniques (EPI and SNPE) with respect to the state-of-the-art approximate bayesian computation technique (SMC-ABC).
-While controlling for architecture size (Fig.
-RNN1), we push the limits of SNPE through targeted hyperparameter modifications (Fig.
-RNN3), and show that EPI scales to higher dimensional RNN connectivities producing stable amplification than SNPE.
-Furthermore, we emphasize that SNPE does not constrain the properties of the inferred parameters; many connectivities inferred by SNPE result in unstable or nonamplified models (Fig. 2C, Fig. RNN2).
+While controlling for architecture size (Figure 2-figure supplement 1), we push the limits of SNPE through targeted hyperparameter modifications (Figure 2-figure supplement 2), and show that EPI scales to higher dimensional RNN connectivities producing stable amplification than SNPE.
+Furthermore, we emphasize that SNPE does not constrain the properties of the inferred parameters; many connectivities inferred by SNPE result in unstable or nonamplified models (Figure 2C, Figure 2-figure supplement 3).
 
@@
 \begin{displayquote}
-Lines 432-444\\
+Lines 435-447\\
 ``\quoting{discussion_gradients}''
@@
 \begin{displayquote}
-Lines 829-835\\
+Lines 832-838\\
 ``\quoting{amortization}''
@@
 EPI learns parameter distributions whose predictions are constrained to produce the emergent property.
-We show in Figure 2 and Supplementary Figure RNN2 that SNPE does not control the statistical properties of its predictions, resulting in the inference of many parameters that are not consistent with the desired emergent property.
+We show in Figure 2 and Supplementary Figure 2-figure supplement 3 that SNPE does not control the statistical properties of its predictions, resulting in the inference of many parameters that are not consistent with the desired emergent property.
 
 \begin{displayquote}
-Lines 227-237\\
+Lines 228-238\\
 ``\quoting{rnn_constrained}''
@@
 The deep probability distribution fit by EPI facilitates the mode identification (via gradient ascent of the parameter log probability) and sensitivity measurement (via the measurement of the eigenvector of the Hessian at a parameter value).
-These mode identifications and sensitivity measurements are done in Sections 3.1 (Fig.
-1E), 3.4 (Fig.
-3E), and 3.5 (Fig.
-4C).
+These mode identifications and sensitivity measurements are done in Sections 3.1 (Figure 1E), 3.4 (Figure 3E), and 3.5 (Figure 4C).
 By using this mode identification technique along the ridges of high parameter probability in the SC model, we identify the parameters transitioning between the two regimes.
-Finally, the sensitivity dimensions of the SC model identified by EPI facilitated regime characterization through perturbation analyses (Fig.
-4D).
+Finally, the sensitivity dimensions of the SC model identified by EPI facilitated regime characterization through perturbation analyses (Figure 4D).
 
@@
 \begin{displayquote}
-Lines 295-303 \\
+Lines 297-305 \\
 ``\quoting{v1_summary}''
@@
 \begin{displayquote}
-Lines 409-412 \\
+Lines 412-415 \\
 ``\quoting{sc_summary}''
@@
 \begin{displayquote}
-Lines 437-440 \\
+Lines 440-443 \\
 ``However, conditioning on the pyloric rhythm \cite{marder1992dynamic} in a model of the pyloric subnetwork model \cite{prinz2004similar} proved to be prohibitive with EPI.
@@
 \begin{displayquote}
-Lines 1021-1025 \\
+Lines 1027-1031 \\
 ``\quoting{EPI_VI}''
@@
 \begin{displayquote}
-Section \ref{results_V1}, Lines 274-277\\
+Section \ref{results_V1}, Lines 275-278\\
 ``\quoting{v1_ep_text1}~\quoting{v1_ep_text2}''
@@
 \begin{displayquote}
-Section \ref{results_SC}, Lines 329-332\\
+Section \ref{results_SC}, Lines 331-334\\
 ``\quoting{sc_ep_text1}.~\quoting{sc_ep_text2}''
@@
 \textit{Analytic comparisons}:
-The 2x2 linear dynamical system is chosen as a worked example because it has multi-modal non-elliptical structure (Fig.
-\ref{fig:LDS1}), and its contours can be derived analytically (Fig.
-\ref{fig:LDS2}).
+The 2x2 linear dynamical system is chosen as a worked example because it has multi-modal non-elliptical structure (Figure 1-figure supplement 1), and its contours can be derived analytically (Figure 1-figure supplement 2).
 Similarly, in Section \ref{methods_V1_complexity}, we derive the quadratic relationship between excitatory variability and input noise variability (in a simplified model) suggesting that the quadratic relationship uncovered by EPI (see Section \ref{results_V1}) is correct.
@@
 \textit{Brute-force}:
-The EPI inferred distribution for rapid task switching in the SC model is sharply bent (Fig. \ref{fig:SC}), and matches the parameter set returned from random sampling (Figure \ref{fig:SC8}A).  
-We note that the brute-force parameter set is actually not the ground-truth solution, because it does not obey the constraints of the emergent property as the EPI distribution does (Figure \ref{fig:SC8}B).
+The EPI inferred distribution for rapid task switching in the SC model is sharply bent (Figure \ref{fig:SC}), and matches the parameter set returned from random sampling (Figure 4-figure supplement 5A).
+We note that the brute-force parameter set is actually not the ground-truth solution, because it does not obey the constraints of the emergent property as the EPI distribution does (Figure 4-figure supplement 5B).
 This can explain the spurious samples in the brute-force set that are not in the EPI inferred distribution.
@@
 \begin{displayquote}
-Lines 424-444\\
+Lines 427-447\\
 ``\quoting{discussion_methods} \\ \quoting{discussion_gradients}''
@@
 
-Lines 764-766 \\
+Lines 767-769 \\
 \begin{displayquote}

### 2021-06-08 written/eLifeRevision/Bittner_eLife2021.tex
rename from 2020 to 2021
Duan 2021 ref update

--- 2021-06-04 written/eLifeRevision/Bittner_eLife2020.tex
@@
 \usepackage[ruled,linesnumbered]{algorithm2e}
+\usepackage{caption}
 
@@
 \q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
-Finally, we investigated the possible connectivities of a superior colliculus model that allow execution of different tasks on interleaved trials \cite{duan2019collicular}.
+Finally, we investigated the possible connectivities of a superior colliculus model that allow execution of different tasks on interleaved trials \cite{duan2021collicular}.
 EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}
+\includegraphics[scale=0.8]{fig1.pdf}
 \end{center}
@@
 Second, once the model is selected, one must specify what the model should produce.
-In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Fig. \ref{fig:STG}B).
+In this STG model, we are concerned with neural spiking frequency, which emerges from the dynamics of the circuit model (Figure \ref{fig:STG}B).
 An emergent property studied by Gutierrez et al. is the hub neuron firing at an intermediate frequency between the intrinsic spiking rates of the fast and slow populations.
@@
 Before presenting technical details (in the following section), let us understand emergent property inference schematically.
-EPI (Fig. \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Fig. \ref{fig:STG}E).  
+EPI (Figure \ref{fig:STG}D) takes, as input, the model and the specified emergent property, and as its output, returns the parameter distribution (Figure \ref{fig:STG}E).  
 This distribution -- represented for clarity as samples from the distribution -- is a parameter distribution constrained such that the circuit model produces the emergent property. 
@@
 \q{results_hess}{The structure of the inferred parameter distributions of EPI can be analyzed to reveal key information about how the circuit model produces the emergent property.
-As probability in the EPI distribution decreases away from the mode of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ (Fig. \ref{fig:STG}E yellow star), the emergent property deteriorates.
+As probability in the EPI distribution decreases away from the mode of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ (Figure \ref{fig:STG}E yellow star), the emergent property deteriorates.
 Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ changes little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
@@
 By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative, the more sensitive and the closer to zero, the more robust (see Section \ref{methods_stg_hessian}).
-Indeed, samples equidistant from the mode along these dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig.
-\ref{fig:STG}E, arrows) agree with error contours (Fig.
-\ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig.
-\ref{fig:STG}F activity traces).
+Indeed, samples equidistant from the mode along these dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Figure \ref{fig:STG}E, arrows) agree with error contours (Figure \ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Figure \ref{fig:STG}F activity traces).
 The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
@@
 Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI discovers parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
-Finally, in a model of superior colliculus \cite{duan2019collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.
+Finally, in a model of superior colliculus \cite{duan2021collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.
 
@@
 %EPI optimizes the weights and biases $\bm{\theta}$ of the deep network (which induces the probability distribution) by iteratively solving Equation \ref{eq:EPI}. 
-%The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Fig. S4).
+%The optimization is complete when the sampled models with parameters $\mathbf{z} \sim q_{\bm{\theta}}(z \mid \mathcal{X})$ produce activity consistent with the specified emergent property (Figure S4).
 %%Such convergence is evaluated with a hypothesis test that the mean of each emergent property statistic is not different than its emergent property value (see Section \ref{methods_AL_opt}). 
@@
 %The dimensions of greatest sensitivity (e.g.
-Fig.
-1E solid arrow) or degeneracy (e.g.
-Fig.
-1E dashed arrow) can be measured directly from the second order derivative of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ called the ``Hessian."
+Figure 1E solid arrow) or degeneracy (e.g.
+Figure 1E dashed arrow) can be measured directly from the second order derivative of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ called the ``Hessian."
 %Around the mode, eigenvalues of the Hessian are negative; probabilities decrease locally in all directions away from the mode.
@@
 \begin{center}
-\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
+\includegraphics[scale=.8]{fig2.pdf}
 \end{center}
@@
 
-As the number of neurons $N$ in the RNN, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, is scaled, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig.
-\ref{fig:RNN}A).
-It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
+As the number of neurons $N$ in the RNN, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, is scaled, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Figure \ref{fig:RNN}A).
+It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Figure \ref{fig:RNN}B).
 It is well known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
-Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig.
-\ref{fig:RNN1}), and tested more aggressive SNPE hyperparameter choices when SNPE failed to converge (Fig.
-\ref{fig:RNN3}).
+Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Figure 2-figure supplement 1), and tested more aggressive SNPE hyperparameter choices when SNPE failed to converge (Figure 2-figure supplement 2).
 In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time (elapsed real time) and simulation count.
 
-\q{rnn_constrained}{No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Fig.
-\ref{fig:RNN}C, blue).
-For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from the exemplar dataset $\mathbf{x}_0$ (Fig. \ref{fig:RNN}C, green).
-When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Fig. \ref{fig:RNN}C, orange).
-We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|\mathbf{x}(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions (Fig.
-\ref{fig:RNN}D).
+\q{rnn_constrained}{No matter the number of neurons, EPI always produces connectivity distributions with mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$ according to $\mathcal{X}$ (Figure \ref{fig:RNN}C, blue).
+For the dimensionalities in which SMC-ABC is tractable, the inferred parameters are concentrated and offset from the exemplar dataset $\mathbf{x}_0$ (Figure \ref{fig:RNN}C, green).
+When using SNPE, the predictions of the inferred parameters are highly concentrated at some RNN sizes and widely varied in others (Figure \ref{fig:RNN}C, orange).
+We see these properties reflected in simulations from the inferred distributions: EPI produces a consistent variety of stable, amplified activity norms $|\mathbf{x}(t)|$, SMC-ABC produces a limited variety of responses, and the changing variety of responses from SNPE emphasizes the control of EPI on parameter predictions (Figure \ref{fig:RNN}D).
 Even for moderate neuron counts, the predictions of the inferred distribution of SNPE are highly dependent on $N$ and $g$, while EPI maintains the emergent property across choices of RNN (see Section \ref{methods_RNN_params}).}
@@
 Furthermore, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
-composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig.
-\ref{fig:V1}A) \cite{pfeffer2013inhibition}.
+composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Figure \ref{fig:V1}A) \cite{pfeffer2013inhibition}.
 Here, we use EPI on a model of V1 with biologically realistic connectivity to show how the structure of input across neuron types affects the variability of the excitatory population -- the population largely responsible for projecting to other brain areas \cite{felleman1991distributed}.
 
-We considered response variability of a nonlinear dynamical V1 circuit model (Fig.
-\ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
+We considered response variability of a nonlinear dynamical V1 circuit model (Figure \ref{fig:V1}A) with a state comprised of each neuron-type population's rate $\mathbf{x} = \left[x_E, x_P , x_S, x_V \right]^\top$.
 Each population receives recurrent input $W \mathbf{x}$, where $W$ is the effective connectivity matrix (see Section \ref{methods_V1}) and
@@
 The external input has an additive noisy component $\bm{\epsilon}$ with variance $\bm{\sigma}^2 = \left[ \sigma_E^2, \sigma_P^2, \sigma_S^2, \sigma_V^2\right]$.
-This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Fig. \ref{fig:V1}B).
+This noise has a slower dynamical timescale $\tau_{\text{noise}} > \tau$ than the population rate, allowing fluctuations around a stimulus-dependent steady-state (Figure \ref{fig:V1}B).
 This model is the stochastic stabilized supralinear network (SSSN) \cite{hennequin2018dynamical} 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
+\includegraphics[scale=0.8]{fig3.pdf}
 \end{center}
@@
 \end{equation}
-\q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Fig. \ref{fig:V1}C).
+\q{v1_ep_text2}{where $s_E(\mathbf{x}; \mathbf{z})$ is the standard deviation of the stochastic $E$-population response about its steady state (Figure \ref{fig:V1}C).
 In the following analyses, we select 1Hz$^2$ variance such that the two emergent properties do not overlap in $s_E(\mathbf{z}; \mathbf{x})$.}
 
-First, we ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ producing E-population variability around 5Hz (Fig. \ref{fig:V1}D).
-From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Fig.
-\ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
-Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the unexpectedly high variability in those dimensions (Fig.
-\ref{fig:V1}D, bottom-right).
-Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5Hz, which is indicated by the modes along $\sigma_P$ (Fig. \ref{fig:V1}E).
+First, we ran EPI to obtain parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ producing E-population variability around 5Hz (Figure \ref{fig:V1}D).
+From the marginal distribution of $\sigma_E$ and $\sigma_P$ (Figure \ref{fig:V1}D, top-left), we can see that $s_E(\mathbf{x}; \mathbf{z})$ is sensitive to various combinations of $\sigma_E$ and $\sigma_P$.
+Alternatively, both $\sigma_S$ and $\sigma_V$ are degenerate with respect to $s_E(\mathbf{x}; \mathbf{z})$ evidenced by the unexpectedly high variability in those dimensions (Figure \ref{fig:V1}D, bottom-right).
+Together, these observations imply a curved path with respect to $s_E(\mathbf{x}; \mathbf{z})$ of 5Hz, which is indicated by the modes along $\sigma_P$ (Figure \ref{fig:V1}E).
 
 Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
-This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{Hz})$ (Fig. \ref{fig:V1}E and \ref{fig:V1_1}).
-Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig.
-\ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
-The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig.
-\ref{fig:V1_4}), meaning that the E-population is P-stabilized.
+This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{Hz})$ (Figure \ref{fig:V1}E and Figure 3-figure supplement 1).
+Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Figure 3-figure supplement 3, $p=.40$), in which parameters were bounded from 0 to 0.5.
+The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Figure 3-figure supplement 4), meaning that the E-population is P-stabilized.
 Future research may uncover a link between the population of network stabilization and compensatory interactions governing excitatory variability.
@@
 In this experimental setup, rats were given an explicit task cue on each trial, either Pro or Anti. 
-After a delay period, rats were shown a stimulus, and made a context  (task) dependent response (Fig. \ref{fig:SC}A). 
+After a delay period, rats were shown a stimulus, and made a context  (task) dependent response (Figure \ref{fig:SC}A). 
 In the Pro task, rats were required to orient towards the stimulus, while in the Anti task, rats were required to orient away from the stimulus.
-Pharmacological inactivation of the SC impaired rat performance, and time-specific optogenetic inactivation revealed a crucial role for the SC on the cognitively demanding Anti trials \cite{duan2019collicular}.
+Pharmacological inactivation of the SC impaired rat performance, and time-specific optogenetic inactivation revealed a crucial role for the SC on the cognitively demanding Anti trials \cite{duan2021collicular}.
 These results motivated a nonlinear dynamical model of the SC containing four functionally-defined neuron-type populations.
-In Duan et al. 2019, a computationally intensive procedure was used to obtain a set of 373 connectivity parameters that qualitatively reproduced these optogenetic inactivation results.
+In Duan et al. 2021, a computationally intensive procedure was used to obtain a set of 373 connectivity parameters that qualitatively reproduced these optogenetic inactivation results.
 To build upon the insights of this previous work, we use the probabilistic tools afforded by EPI to identify and characterize two linked, yet distinct regimes of rapid task switching connectivity.
 
-In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ \cite{duan2019collicular}.
-The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
+In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ \cite{duan2021collicular}.
+The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Figure \ref{fig:SC}B).
 The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_sc_model}).
@@
 \begin{center}
-\includegraphics[scale=0.75]{figures/fig4/fig4.pdf}
+\includegraphics[scale=0.75]{fig4.pdf}
 \end{center}
@@
 
-The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, bottom-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
-This parameter distribution has rich structure that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
-Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8}).
-This is most saliently observed in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
-By identifying the modes of the EPI distribution $\mathbf{z}^*(sW)$ at different values of $sW$ (Fig.
-\ref{fig:SC}C red/purple dots), we can quantify this change in distributional structure with the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$ (Fig.
-\ref{fig:SC}C red/purple arrows).
+The EPI inferred distribution  (Figure \ref{fig:SC}C) produces Pro and Anti task accuracies (Figure \ref{fig:SC}C, bottom-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
+This parameter distribution has rich structure that is not captured well by simple linear correlations (Figure 4-figure supplement 1).
+Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Figure 4-figure supplement 5).
+This is most saliently observed in the marginal distribution of  $sW$-$hW$ (Figure \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
+By identifying the modes of the EPI distribution $\mathbf{z}^*(sW)$ at different values of $sW$ (Figure \ref{fig:SC}C red/purple dots), we can quantify this change in distributional structure with the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$ (Figure \ref{fig:SC}C red/purple arrows).
 Note that the directionality of these sensitivity dimensions at $\mathbf{z}^*(sW)$ changes distinctly with $sW$, and are perpendicular to the robust dimensions of the EPI distribution that preserve rapid task switching.
-These two directionalities of sensitivity motivate the distinction of connectivity into two regimes, which produce different types of responses in the Pro and Anti tasks (Fig. \ref{fig:SC2}).
+These two directionalities of sensitivity motivate the distinction of connectivity into two regimes, which produce different types of responses in the Pro and Anti tasks (Figure 4-figure supplement 2).
 
@@
 \end{equation}
-Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
+Pro accuracy monotonically increases in both regimes (Figure \ref{fig:SC}D, top-left).
 However, there is a stark difference between regimes in Anti accuracy. 
-Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, bottom-left).
+Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Figure \ref{fig:SC}D, bottom-left).
 The sharp change in local structure of the EPI distribution is therefore explained by distinct sensitivities: Anti accuracy diminishes in only one or both directions of the sensitivity perturbation.
@@
 To understand the mechanisms differentiating the two regimes, we can make connectivity perturbations along dimensions that only modify a single eigenvalue of the connectivity matrix.
-These eigenvalues $\lambda_{\text{all}}$, $\lambda_{\text{side}}$, $\lambda_{\text{task}}$, and $\lambda_{\text{diag}}$ correspond to connectivity eigenmodes with intuitive roles in processing in this task (Fig.
-\ref{fig:SC3}A).
+These eigenvalues $\lambda_{\text{all}}$, $\lambda_{\text{side}}$, $\lambda_{\text{task}}$, and $\lambda_{\text{diag}}$ correspond to connectivity eigenmodes with intuitive roles in processing in this task (Figure 4-figure supplement 3A).
 For example, greater $\lambda_{\text{task}}$ will strengthen internal representations of task, while greater $\lambda_{\text{diag}}$ will amplify dominance of Pro and Anti pairs in opposite hemispheres (Section \ref{methods_sc_eig}).
@@
 \end{equation}
-Connectivity perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has a very similar effect on Anti accuracy as perturbations along the sensitivity dimension (Fig.
-\ref{fig:SC}D, middle).
+Connectivity perturbation analyses reveal that decreasing $\lambda_{\text{task}}$ has a very similar effect on Anti accuracy as perturbations along the sensitivity dimension (Figure \ref{fig:SC}D, middle).
 The similar effects of perturbations along the sensitivity dimension $\mathbf{v}_1(\mathbf{z}^*)$ and reduction of task eigenvalue (via perturbations along $-\mathbf{v}_{\text{task}}$) suggest that there is a carefully tuned strength of task representation in connectivity regime 1, which if disturbed results in random Anti trial responses.
-Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Fig. \ref{fig:SC}D, right).
+Finally, we recognize that increasing $\lambda_{\text{diag}}$ has opposite effects on Anti accuracy in each regime (Figure \ref{fig:SC}D, right).
 In the next section, we build on these mechanistic characterizations of each regime by examining their resilience to optogenetic inactivation.
@@
 Duan et al.
-found that bilateral optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Fig.
-\ref{fig:SC_opto}A) \cite{duan2019collicular}.
-The distribution of connectivities inferred by EPI exhibited this same effect in simulation at high optogenetic strengths $\gamma$, which reduce the network activities $\mathbf{x}(t)$ by a factor $1 - \gamma$ (Fig.
-\ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1.1]{figures/fig5/fig5.pdf}
+found that bilateral optogenetic inactivation of SC during the delay period consistently decreased performance in the Anti task, but had no effect on the Pro task (Figure \ref{fig:SC_opto}A) \cite{duan2021collicular}.
+The distribution of connectivities inferred by EPI exhibited this same effect in simulation at high optogenetic strengths $\gamma$, which reduce the network activities $\mathbf{x}(t)$ by a factor $1 - \gamma$ (Figure \ref{fig:SC_opto}B) (see Section \ref{methods_sc_opto}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.1]{fig5.pdf}
 \end{center}
@@
 
-The mean increase in Anti error of the EPI distribution is closest to the experimentally measured value of 7\% at $\gamma = 0.675$ (Fig. \ref{fig:SC_opto}B, black dot).
-At this level of optogenetic strength, regime 1 exhibits an increase in Anti error with delay period silencing (Fig. \ref{fig:SC_opto}C, left), while regime 2 does not.
-In regime 1, greater $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ decrease Anti error (Fig. \ref{fig:SC_opto}C, right).
+The mean increase in Anti error of the EPI distribution is closest to the experimentally measured value of 7\% at $\gamma = 0.675$ (Figure \ref{fig:SC_opto}B, black dot).
+At this level of optogenetic strength, regime 1 exhibits an increase in Anti error with delay period silencing (Figure \ref{fig:SC_opto}C, left), while regime 2 does not.
+In regime 1, greater $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ decrease Anti error (Figure \ref{fig:SC_opto}C, right).
 In other words, stronger task representations and diagonal amplification make the SC model more resilient to delay period silencing in the Anti task.
-This complements the finding from Duan et al. 2019 \cite{duan2019collicular} that $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ improve Anti accuracy.
-
-At roughly $\gamma = 0.85$ (Fig. \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero.
-Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Fig. \ref{fig:SC_opto}D, left).
+This complements the finding from Duan et al. 2021 \cite{duan2021collicular} that $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ improve Anti accuracy.
+
+At roughly $\gamma = 0.85$ (Figure \ref{fig:SC_opto}B, gray dot), the Anti error saturates, while Pro error remains at zero.
+Following delay period inactivation at this optogenetic strength, there are strong similarities in the responses of Pro and Anti trials during the choice period (Figure \ref{fig:SC_opto}D, left).
 We interpreted these similarities to suggest that delay period inactivation at this saturated level flips the internal representation of task (from Anti to Pro) in the circuit model.
 A flipped task representation would explain why the Anti error saturates at 50\%: the average Anti accuracy in EPI inferred connectivities is $75\%$, but is 25\% when the internal representation is flipped during delay period silencing.
-This hypothesis prescribes a model of Anti accuracy during delay period silencing of $p_{A,\text{opto}} = 100\% - p_P$, which is fit closely across both regimes of the EPI inferred connectivities (Fig.
-\ref{fig:SC_opto}D, right).
-Similarities between Pro and Anti trial responses were not present at the experiment-matching level of $\gamma=0.675$ (Fig.
-\ref{fig:SC6} left) and neither was anticorrelation in $p_P$ and $p_{A,\text{opto}}$ (Fig.
-\ref{fig:SC6} right).
+This hypothesis prescribes a model of Anti accuracy during delay period silencing of $p_{A,\text{opto}} = 100\% - p_P$, which is fit closely across both regimes of the EPI inferred connectivities (Figure \ref{fig:SC_opto}D, right).
+Similarities between Pro and Anti trial responses were not present at the experiment-matching level of $\gamma=0.675$ (Figure 5-figure supplement 2 left) and neither was anticorrelation in $p_P$ and $p_{A,\text{opto}}$ (Figure 5-figure supplement 2 right).
 
@@
 
-\bibliography{eLife2020}
+\bibliography{eLife2021}
 \bibliographystyle{unsrt}
@@
 \newpage 
-\startsupplement
+%\startsupplement
 \section{Methods}
@@
 Methodological advances to ABC methods have come through the use of Markov chain Monte Carlo (MCMC-ABC) \cite{marjoram2003markov} and sequential Monte Carlo (SMC-ABC) \cite{sisson2007sequential} sampling techniques.
-SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality \cite{sisson2018handbook} (cf. Fig. \ref{fig:RNN}).
+SMC-ABC is considered state-of-the-art ABC, yet this approach still struggles to scale in dimensionality \cite{sisson2018handbook} (cf. Figure \ref{fig:RNN}).
 Still, this method has enjoyed much success in systems biology \cite{liepe2014framework}.
@@
 EPI infers parameters in circuit models using the MEFN \cite{loaiza2017maximum} algorithm with a deep variational approximation.
-The deep neural network of EPI (Fig.
-\ref{fig:STG}E) defines the parametric form (with weights and biases as variational parameters $\bm{\theta}$) of the variational approximation of the inferred parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathbf{x})$.
+The deep neural network of EPI (Figure \ref{fig:STG}E) defines the parametric form (with weights and biases as variational parameters $\bm{\theta}$) of the variational approximation of the inferred parameter distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathbf{x})$.
 The EPI optimization is enabled using stochastic gradient techniques in the spirit of likelihood-free variational inference \cite{tran2017hierarchical}.
@@
 
-\begin{algorithm}\label{alg:epi}
+\begin{algorithm}
 \SetAlgoLined
@@
  \caption{Emergent property inference}
+ \label{alg:epi}
 \end{algorithm}
@@
 With each training epoch in which the update rule for $c$ is invoked, the constraint satisfaction terms are increasingly weighted, which generally results in decreased entropy (e.g.
-see Figure \ref{fig:LDS1}C).
+see Figure 1-figure supplement 1C).
 This encourages the discovery of suitable regions of parameter space, and the subsequent refinement of the distribution to produce the emergent property.
@@
 \subsubsection{Example: 2D LDS}\label{methods_2DLDS}
-To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. \ref{fig:LDS1}A):
+To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Figure 1-figure supplement 1A):
 \begin{equation} 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/figLDS1/figLDS1.pdf}
-\end{center}
-\caption{\textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
+\includegraphics[scale=0.8]{fig1supp1.pdf}
+\end{center}
+\caption*{Figure 1-figure supplement 1:~~\textbf{A}. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
 \textbf{B}.
@@
 At the beginning of each augmented lagrangian epoch, the emergent property moments adjust closer to their constraints.}
-\label{fig:LDS1}
 \end{figure}
@@
 This fact is due to the formally hard problem of the backward mapping: finding the natural parameters $\bm{\eta}$ from the mean parameters $\bm{\mu}$ of an exponential family distribution \cite{wainwright2008graphical}.
-Instead, we used EPI to approximate this distribution (Fig. \ref{fig:LDS1}B). 
+Instead, we used EPI to approximate this distribution (Figure 1-figure supplement 1B). 
 We used a real NVP normalizing flow architecture three coupling layers and two-layer neural networks of 50 units per layer, mapped onto a support of $z_i \in \left[-10, 10 \right]$.
@@
 the standard deviation $\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.25$, and the dashed gray line at twice the standard deviation
-$\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.5$ follow the contour of probability density of the samples (Fig. \ref{fig:LDS2}A). 
+$\text{real}(\lambda_1) = \frac{a_{1,1} + a_{2,2}}{2} \pm 0.5$ follow the contour of probability density of the samples (Figure 1-figure supplement 2A). 
 The distribution precisely reflects the desired statistical constraints and model degeneracy in the sum of $a_{1,1}$ and $a_{2,2}$.
@@
 \end{equation}
-In Figure \ref{fig:LDS2}B, we plot the contours of $\text{imag}(\lambda_1)$ where $a_{1,1} a_{2,2}$ is fixed to 0 at one standard deviation ($\frac{\pi}{5}$, black dashed) and two standard deviations ($\frac{2\pi}{5}$, gray dashed) from the mean of $2 \pi$.
+In Figure 1-figure supplement 2B, we plot the contours of $\text{imag}(\lambda_1)$ where $a_{1,1} a_{2,2}$ is fixed to 0 at one standard deviation ($\frac{\pi}{5}$, black dashed) and two standard deviations ($\frac{2\pi}{5}$, gray dashed) from the mean of $2 \pi$.
 This validates the curved multimodal structure of the inferred distribution learned through EPI.  
 Subtler combinations of model and emergent property will have more complexity, further motivating the use of EPI for understanding these systems.  
-As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Fig. \ref{fig:LDS3}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS2/figLDS2.pdf}
-\end{center}
-\caption{\textbf{A}. Probability contours in the $a_{1,1}$-$a_{2,2}$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
+As we expect, the distribution results in samples of two-dimensional linear systems oscillating near 1Hz (Figure 1-figure supplement 3).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{fig1supp2.pdf}
+\end{center}
+\caption*{Figure 1-figure supplement 2:~~\textbf{A}.
+Probability contours in the $a_{1,1}$-$a_{2,2}$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$.
 \textbf{B}. Probability contours in the $a_{1,2}$-$a_{2,1}$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
-\label{fig:LDS2}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figLDS3/figLDS3.pdf}
-\end{center}
-\caption{Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ and their simulated activity from $\mathbf{x}(t=0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
+%\label{fig:LDS2}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{fig1supp3.pdf}
+\end{center}
+\caption*{Figure 1-figure supplement 3:~~Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ and their simulated activity from $\mathbf{x}(t=0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability.
 \textbf{A}. Each dimension of the simulated trajectories throughout time.  
 \textbf{B}.  The simulated trajectories in phase space.}
-\label{fig:LDS3}
+%\label{fig:LDS3}
 \end{figure}
@@
 Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 5,000$ iterations (see Section \ref{methods_AL_opt}).  
-The optimization converged after five epochs (Fig. \ref{fig:STG1}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figSTG1/figSTG1.pdf}
-\end{center}
-\caption{EPI optimization of the STG model producing network syncing. 
+The optimization converged after five epochs (Figure 1-figure supplement 4).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{fig1supp4.pdf}
+\end{center}
+\caption*{Figure 1-figure supplement 4:~~EPI optimization of the STG model producing network syncing. 
 \textbf{A}. Entropy throughout optimization. 
 \textbf{B}. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented lagrangian epoch.}
-\label{fig:STG1}
+%\label{fig:STG1}
 \end{figure}
@@
 We did not include $g_{\text{synA}} < 0.01$, for numerical stability.
-EPI optimization was run using 5 different random seeds for architecture initialization $\bm{\theta}$ with an augmented lagrangian coefficient of $c_0 = 10^{5}$, $\beta = 2$, a batch size $n=400$, and we simulated one $\mathbf{x}^{(i)}$ per $\mathbf{z}^{(i)}$.
+EPI optimization was run with an augmented lagrangian coefficient of $c_0 = 10^{5}$, hyperparameter $\beta = 2$, a batch size $n=400$, and we simulated one $\mathbf{x}^{(i)}$ per $\mathbf{z}^{(i)}$.
 The architecture converged with criteria $N_{\text{test}} = 100$.
@@
 The length of the arrows is inversely proportional to the square root of the absolute value of their eigenvalues $\lambda_1 = -10.7$ and $\lambda_2 = -3.22$.
-For the same magnitude perturbation away from the mode, intermediate hub frequency only diminishes along the sensitivity dimension $\mathbf{v}_1$ (Fig. \ref{fig:STG}E-F).
-%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Fig. \ref{fig:STG}B, inset).
+For the same magnitude perturbation away from the mode, intermediate hub frequency only diminishes along the sensitivity dimension $\mathbf{v}_1$ (Figure \ref{fig:STG}E-F).
+%We quantitatively measured the sensitivity of the model with respect to network syncing along the eigenvectors of the Hessian (Figure \ref{fig:STG}B, inset).
 %Sensitivity was measured as the slope coefficient of linear regression fit to network syncing error (the sum of squared differences of each neuron's frequency from 0.53Hz) as a function of parametric perturbation magnitude (maximum 0.25) away from the mode along both orientations indicated by the eigenvector with 100 equally spaced samples.
@@
 %In contrast, eigenvector $v_2$ did not identify a dimension of parameter space significantly sensitive to network syncing ($\beta = 8.65 \times 10^{-4}$ with $p=.67$).
-%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Fig.
-\ref{fig:STG}B, inset).
+%These sensitivities were compared to all other dimensions of parameter space (100 equally spaced angles from 0 to $\pi$), revealing that the Hessian eigenvectors indeed identified the directions of greatest sensitivity and degeneracy (Figure \ref{fig:STG}B, inset).
 %The contours of Figure \ref{fig:STG} were calculated as error in $T(\mathbf{x})$ from $\mu$ in both the first and second moment emergent property statistics.
@@
 
-For EPI in Fig. \ref{fig:RNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
+For EPI in Figure \ref{fig:RNN}, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 100 units per layer.
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to the support of $\mathbf{z}_i \in [-1, 1]$. 
 We used an augmented lagrangian coefficient of $c_0 = 10^{3}$, a batch size $n=200$, $\beta=4$, and we simulated one $\mathbf{W}^{(i)}$ per $\mathbf{z}^{(i)}$.
-We chose to use $i_{\text{max}} = 500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Fig.
-\ref{fig:RNN}B blue line, and Fig.
-\ref{fig:RNN}C-D blue).
+We chose to use $i_{\text{max}} = 500$ iterations per augmented lagrangian epoch and emergent property constraint convergence was evaluated at $N_{\text{test}} = 200$ (Figure \ref{fig:RNN}B blue line, and Figure \ref{fig:RNN}C-D blue).
 It was fastest to initialize the EPI distribution on a Tesla V100 GPU, and then subsequently optimize it on a CPU with 32 cores.
@@
 For the deep probability distribution architecture, we chose a masked autoregressive flow with affine couplings (the default choice), three transforms, 50 hidden units, and a normalizing flow mapping to the support as in EPI.
-This architectural choice closely tracked the size of the architecture used by EPI (Fig. \ref{fig:RNN1}).
+This architectural choice closely tracked the size of the architecture used by EPI (Figure 2-figure supplement 1).
 As in SMC-ABC, we ran SNPE with $\mathbf{x}_0 = \mu$.
@@
 \begin{center}
-\includegraphics[scale=0.7]{figures/figRNN1/figRNN1.pdf}
-\end{center}
-\caption{Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
-}
-\label{fig:RNN1}
+\includegraphics[scale=0.7]{fig2supp1.pdf}
+\end{center}
+\caption*{Figure 2-figure supplement 1:~~Number of parameters in deep probability distribution architectures of EPI (blue) and SNPE (orange) by RNN size ($N$).
+}
+%\label{fig:RNN1}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=1.]{figures/figRNN3/figRNN3.pdf}
-\end{center}
-\caption{SNPE convergence was enabled by increasing $n_{\text{round}}$, not $n_{\text{atom}}$.
+\includegraphics[scale=1.]{fig2supp2.pdf}
+\end{center}
+\caption*{Figure 2-figure supplement 2:~~SNPE convergence was enabled by increasing $n_{\text{round}}$, not $n_{\text{atom}}$.
 \textbf{A}.
@@
 }
-\label{fig:RNN3}
+%\label{fig:RNN3}
 \end{figure}
@@
 To match EPI, we used a batch size of $n = 200$ for $N <= 25$, however we found $n=1,000$ to be helpful for SNPE in higher dimensions.
-While $n_{\text{round}} = 1,000$ yielded SNPE convergence for $N <= 25$, we found that a substantial increase to $n_{\text{round}} = 25,000$ yielded more consistent convergence at $N=50$ (Fig.
-\ref{fig:RNN3}A).
+While $n_{\text{round}} = 1,000$ yielded SNPE convergence for $N <= 25$, we found that a substantial increase to $n_{\text{round}} = 25,000$ yielded more consistent convergence at $N=50$ (Figure 2-figure supplement 2A).
 By increasing $n_{\text{round}}$, we also necessarily increase the duration of each round.
 At $N=100$, we tried two hyperparameter modifications.
-As suggested in \cite{greenberg2019automatic}, we increased $n_{\text{atom}}$ by an order of magnitude to improve gradient quality, but this had little effect on the optimization (much overlap between same random seeds) (Fig.
-\ref{fig:RNN3}B).
+As suggested in \cite{greenberg2019automatic}, we increased $n_{\text{atom}}$ by an order of magnitude to improve gradient quality, but this had little effect on the optimization (much overlap between same random seeds) (Figure Figure 2-figure supplement 2B).
 Finally, we increased $n_{\text{round}}$ by an order of magnitude, which yielded convergence in one case, but no others.
@@
 To clarify the difference in objectives of EPI and SNPE, we show their results on RNN models with different numbers of neurons $N$ and random strength $g$.  
-The parameters inferred by EPI consistently produces the same mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$, while those inferred by SNPE change according to the model definition (Fig.
-\ref{fig:RNN2}A).
-For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater randomness (Fig. \ref{fig:RNN2}B top, orange).
-At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Fig. \ref{fig:RNN2}B top).
-However at $N=10$, SNPE results in a predictive distribution of more widely dispersed eigenvalues (Fig.
-\ref{fig:RNN2}A bottom), and an inferred posterior with greater entropy than EPI (Fig.
-\ref{fig:RNN2}B bottom).
+The parameters inferred by EPI consistently produces the same mean and variance of $\text{real}(\lambda_1)$ and $\lambda_1^s$, while those inferred by SNPE change according to the model definition (Figure 2-figure supplement 3A).
+For $N=2$ and $g=0.01$, the SNPE posterior has greater concentration in eigenvalues around $\mathbf{x}_0$ than at $g=0.1$, where the model has greater randomness (Figure 2-figure supplement 3B top, orange).
+At both levels of $g$ when $N=2$, the posterior of SNPE has lower entropy than EPI at convergence (Figure 2-figure supplement 3B top).
+However at $N=10$, SNPE results in a predictive distribution of more widely dispersed eigenvalues (Figure 2-figure supplement 3A bottom), and an inferred posterior with greater entropy than EPI (Figure 2-figure supplement 3B bottom).
 We highlight these differences not to focus on an insightful trend, but to emphasize that these methods optimize different objectives with different implications.
 
-Note that SNPE converges when it's validation log probability has saturated after several rounds of optimization (Fig.
-\ref{fig:RNN2}C), and that EPI converges after several epochs of its own optimization to enforce the emergent property constraints (Fig.
-\ref{fig:RNN2}D blue).
-Importantly, as SNPE optimizes its posterior approximation, the predictive means change, and at convergence may be different than $\mathbf{x}_0$ (Fig. \ref{fig:RNN2}D orange, left).
+Note that SNPE converges when it's validation log probability has saturated after several rounds of optimization (Figure 2-figure supplement 3C), and that EPI converges after several epochs of its own optimization to enforce the emergent property constraints (Figure 2-figure supplement 3D blue).
+Importantly, as SNPE optimizes its posterior approximation, the predictive means change, and at convergence may be different than $\mathbf{x}_0$ (Figure 2-figure supplement 3D orange, left).
 It is sensible to assume that predictions of a well-approximated SNPE posterior should closely reflect the data on average (especially given a uniform prior and a low degree of stochasticity), however this is not a given.
-Furthermore, no aspect of the SNPE optimization controls the variance of the predictions (Fig. \ref{fig:RNN2}D orange, right).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{figures/figRNN2/figRNN2.pdf}
-\end{center}
-\caption{Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
+Furthermore, no aspect of the SNPE optimization controls the variance of the predictions (Figure 2-figure supplement 3D orange, right).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{fig2supp3.pdf}
+\end{center}
+\caption*{Figure 2-figure supplement 3:~~Model characteristics affect predictions of posteriors inferred by SNPE, while predictions of parameters inferred by EPI remain fixed.
 \textbf{A}.
@@
 }
-\label{fig:RNN2}
+%\label{fig:RNN2}
 \end{figure}
@@
 In particular three types of genetically identified inhibitory cell-types -- parvalbumin (P), somatostatin (S), VIP (V) --
-compose 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and follow specific connectivity patterns (Fig.
-\ref{fig:V1}A) \cite{pfeffer2013inhibition}, which lead to cell-type specific computations \cite{Mossing2021,palmigiano2020structure}.
+compose 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and follow specific connectivity patterns (Figure \ref{fig:V1}A) \cite{pfeffer2013inhibition}, which lead to cell-type specific computations \cite{Mossing2021,palmigiano2020structure}.
 Currently, how the subdivision of inhibitory cell-types, shapes correlated variability by reconfiguring  recurrent network dynamics is not understood.
@@
 
-For EPI in Figures \ref{fig:V1}D-E and \ref{fig:V1_1}, we used a real NVP architecture with three coupling layers and two-layer neural networks of 50 units per layer.
+For EPI in Figure \ref{fig:V1}D-E and Figure 3-figure supplement 1, we used a real NVP architecture with three coupling layers and two-layer neural networks of 50 units per layer.
 The normalizing flow architecture mapped $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ to a support of $\mathbf{z} = [\sigma_E, \sigma_P, \sigma_S, \sigma_V] \in [0.0, 0.5]^4$.
@@
 The distributions shown are those of the architectures converging with criteria $N_{\text{test}} = 100$ at greatest entropy across three random seeds.
-Optimization details are shown in Figure \ref{fig:V1_2}.
-The sums of squares of each pair of parameters are shown for each EPI distribution in Figure \ref{fig:V1_3}.
+Optimization details are shown in Figure 3-figure supplement 2.
+The sums of squares of each pair of parameters are shown for each EPI distribution in Figure 3-figure supplement 3.  
+The plots are histograms of 500 samples from each EPI distribution from which the significance $p$-values of Section \ref{results_V1} are determined.
 
 \begin{figure}[h]
-\caption{EPI inferred distribution for $\mathcal{X}(10\text{Hz})$.
- }
- \label{fig:V1_1}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_3/figV1_3.pdf}
+\caption*{Figure 3-figure supplement 1:~~EPI inferred distribution for $\mathcal{X}(10\text{Hz})$.
+ }
+% \label{fig:V1_1}
+\begin{center}
+\includegraphics[scale=.8]{fig3supp1.pdf}
 \end{center}
@@
 \begin{figure}[h]
-\caption{EPI optimization $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$
+\caption*{Figure 3-figure supplement 2:~~EPI optimization $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{Hz}))$
 \textbf{A}. Entropy throughout optimization. 
@@
  }
- \label{fig:V1_2}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_4/figV1_4.pdf}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\caption{EPI predictive distributions of the sum of squares of each pair of noise parameters.
- }
- \label{fig:V1_3}
-\begin{center}
-\includegraphics[scale=.8]{figures/figV1_2/figV1_2.pdf}
+ %\label{fig:V1_2}
+\begin{center}
+\includegraphics[scale=.8]{fig3supp2.pdf}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\caption*{Figure 3-figure supplement 3:~~EPI predictive distributions of the sum of squares of each pair of noise parameters.
+ }
+%\label{fig:V1_3}
+\begin{center}
+\includegraphics[scale=.8]{fig3supp3.pdf}
 \end{center}
@@
 \subsubsection{Sensitivity analyses}
-In Fig. \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
+In Figure \ref{fig:V1}E, we visualize the modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ throughout the $\sigma_E$-$\sigma_P$ marginal.
 At each local mode $\mathbf{z}^*(\sigma_P)$, where $\sigma_P$ is fixed, we calculated the Hessian and visualized the sensitivity dimension in the direction of positive $\sigma_E$.
@@
 The paradoxical effect occurs when a populations steady state rate is decreased (or increased) when an increase (decrease) in current is applied to that population \cite{tsodyks1997paradoxical}.
-To see which, if any, populations exhibited a paradoxical effect, we examined responses to changes in input (Fig. \ref{fig:V1_4}).
+To see which, if any, populations exhibited a paradoxical effect, we examined responses to changes in input (Figure 3-figure supplement 4).
 Input magnitudes were chosen so that the effect is salient ($0.002$ for E and P, but $0.02$ for S and V).
@@
 \begin{center}
-\includegraphics[scale=.8]{figures/figV1_1/figV1_1.pdf}
-\end{center}
-\caption{(Left) SSSN simulations for small increases in neuron-type population input.
+\includegraphics[scale=.8]{fig3supp4.pdf}
+\end{center}
+\caption*{Figure 3-figure supplement 4:~~(Left) SSSN simulations for small increases in neuron-type population input.
 (Right) Average (solid) and standard deviation (shaded) of stochastic fluctuations of responses.
  }
- \label{fig:V1_4}
+%\label{fig:V1_4}
 \end{figure}
@@
 The ability to switch between two separate tasks throughout randomly interleaved trials, or ``rapid task switching," has been studied in rats, and midbrain superior colliculus (SC) has been show to play an important in this computation \cite{duan2015requirement}.
-Neural recordings in SC exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side), which led to the distinction of two functional classes: the Pro/Contra and Anti/Ipsi neurons \cite{duan2019collicular}.
+Neural recordings in SC exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side), which led to the distinction of two functional classes: the Pro/Contra and Anti/Ipsi neurons \cite{duan2021collicular}.
 Given this evidence, Duan et al. proposed a model with four functionally-defined neuron-type populations: two in each hemisphere corresponding to the Pro/Contra and Anti/Ipsi populations.  
@@
 with time constant $\tau = 0.09s$, step size 24ms and Gaussian noise $d\mathbf{B}$ of variance $0.2^2$.
-These hyperparameter values are motivated by modeling choices and results from \cite{duan2019collicular}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.9]{figures/figSC1/figSC1.pdf}
-\end{center}
-\caption{\textbf{A}. Same pairplot as Fig. \ref{fig:SC}C colored by Pro task accuracy.  
+These hyperparameter values are motivated by modeling choices and results from \cite{duan2021collicular}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.9]{fig4supp1.pdf}
+\end{center}
+\caption*{Figure 4-figure supplement 1:~~\textbf{A}. Same pairplot as Figure \ref{fig:SC}C colored by Pro task accuracy.  
 \textbf{B}. Same as A colored by Anti task accuracy.  
@@
 }
-\label{fig:SC1}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1.]{figures/figSC2/figSC2.pdf}
-\end{center}
-\caption{
+%\label{fig:SC1}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.]{fig4supp2.pdf}
+\end{center}
+\caption*{Figure 4-figure supplement 2:~~
 \textbf{A}. Simulations in network regime 1: $\mathbf{z}^*(sW=-0.75)$.
@@
 }
-\label{fig:SC2}
+%\label{fig:SC2}
 \end{figure}
@@
 \begin{center}
-\includegraphics[scale=.9]{figures/figSC3/figSC3.pdf}
-\end{center}
-\caption{
+\includegraphics[scale=.9]{fig4supp3.pdf}
+\end{center}
+\caption*{Figure 4-figure supplement 3:~~
 \textbf{A}. Invariant eigenvectors of connectivity matrix $W$.
 \textbf{B}.
-Accuracies for connectivity perturbations when changing $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ ($\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ shown in Fig.
-\ref{fig:SC}D).
-}
-\label{fig:SC3}
+Accuracies for connectivity perturbations when changing $\lambda_{\text{all}}$ and $\lambda_{\text{side}}$ ($\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ shown in Figure \ref{fig:SC}D).
+}
+%\label{fig:SC3}
 \end{figure}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1.2]{figures/figSC5/figSC5.pdf}
-\end{center}
-\caption{
-(Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
-(Right) Correlations of connectivity eigenvalues with Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
-}
-\label{fig:SC5}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1.2]{figures/figSC6/figSC6.pdf}
-\end{center}
-\caption{
-(Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
-Accuracy in Pro (top) and Anti (bottom) task is shown as a percentage.
-(Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.
-}
-\label{fig:SC6}
-\end{figure}
-
 \subsubsection{EPI details for the SC model} \label{methods_sc_epi}
@@
 Throughout optimization, the augmented lagrangian parameters $\eta$ and $c$, were updated after each epoch of $i_{\text{max}} = 2,000$ iterations (see Section \ref{methods_AL_opt}).  
-The optimization converged after ten epochs (Fig. \ref{fig:SC6}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{figures/figSC7/figSC7.pdf}
-\end{center}
-\caption{EPI optimization of the SC model producing rapid task switching.
+The optimization converged after ten epochs (Figure 4-figure supplement 4).
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{fig4supp4.pdf}
+\end{center}
+\caption*{Figure 4-figure supplement 4:~~EPI optimization of the SC model producing rapid task switching.
 \textbf{A}. Entropy throughout optimization. 
@@
 }
-\label{fig:SC7}
-\end{figure}
-
-For EPI in Fig. \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
+%\label{fig:SC7}
+\end{figure}
+
+For EPI in Figure \ref{fig:SC}C, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 50 units per layer.
 The initial distribution was a standard isotropic gaussian $\mathbf{z}_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z}_i \in [-5, 5]$. 
@@
 \begin{center}
-\includegraphics[scale=0.8]{figures/figSC8/figSC8.pdf}
-\end{center}
-\caption{
+\includegraphics[scale=0.8]{fig4supp5.pdf}
+\end{center}
+\caption*{Figure 4-figure supplement 5:~~
 \textbf{A}.Rapid task switching SC connectivities obtained from random sampling.
@@
 }
-\label{fig:SC8}
+%\label{fig:SC8}
 \end{figure}
@@
 The bend in the EPI distribution is not a spurious result of the EPI optimization.
-The structure discovered by EPI matches the shape of the set of points returned from brute-force random sampling (Fig. \ref{fig:SC8}A) 
+The structure discovered by EPI matches the shape of the set of points returned from brute-force random sampling (Figure 4-figure supplement 5A) 
 These connectivities were sampled from a uniform distribution over the range of each connectivity parameter, and all parameters producing accuracy in each task within the range of 60\% to 90\% were kept.
 This set of connectivities will not match the distribution of EPI exactly, since it is not conditioned on the emergent property.
-For example the parameter set returned by the brute-force search is biased towards lower accuracies (Fig. \ref{fig:SC8}B).
+For example the parameter set returned by the brute-force search is biased towards lower accuracies (Figure 4-figure supplement 5B).
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.2]{fig5supp1.pdf}
+\end{center}
+\caption*{Figure 5-figure supplement 1:~~
+(Left) Mean and standard error of Pro and Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
+(Right) Correlations of connectivity eigenvalues with Anti error from regime 1 to regime 2 at $\gamma = 0.85$.
+}
+%\label{fig:SC5}
+\end{figure}
+
 \subsubsection{Sample grouping by mode} \label{methods_sc_group}
-For the analyses in Figure \ref{fig:SC_opto}C and Figure \ref{fig:SC5}, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampling from the EPI distribution.
+For the analyses in Figure \ref{fig:SC_opto}C and Figure 5-figure supplement 1, we obtained parameters for each step along the continuum between regimes 1 and 2 by sampling from the EPI distribution.
 Each sample was assigned to the closest mode $\mathbf{z}^*(sW)$.
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.2]{fig5supp2.pdf}
+\end{center}
+\caption*{Figure 5-figure supplement 2:~~
+(Left) Mean and standard deviation (shading) of responses of the SC model at the mode of the EPI distribution to delay period inactivation at $\gamma = 0.675$.
+Accuracy in Pro (top) and Anti (bottom) task is shown as a percentage.
+(Right) Anti accuracy following delay period inactivation at $\gamma = 0.675$ versus accuracy in the Pro task across connectivities in the EPI distribution.
+}
+%\label{fig:SC6}
+\end{figure}
+
 \subsubsection{Connectivity eigendecomposition and processing modes} \label{methods_sc_eig}
-To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same eigenmodes $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Fig.
-\ref{fig:SC3}A).
+To understand the connectivity mechanisms governing task accuracy, we took the eigendecomposition of the connectivity matrices $W = Q\Lambda Q^{-1}$, which results in the same eigenmodes $\mathbf{q}_i$ for all $W$ parameterized by $\mathbf{z}$ (Figure 4-figure supplement 3A).
 These eigenvectors are always the same, because the connectivity matrix is symmetric and the model also assumes symmetry across hemispheres, but the eigenvalues of connectivity (or degree of eigenmode amplification) change with $\mathbf{z}$.
@@
 %
-%These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Fig. \ref{fig:SC2}).
+%These eigenvalue changes are evident in the simulations of connectivity perturbations away from the modes (Figure \ref{fig:SC2}).
 %As the component of connectivity along $\mathbf{v}_1$ and $\mathbf{v}_2$ becomes stronger (left-to-right), there is less separation between Pro an Anti populations (lower $\lambda_{\text{task}}$) and greater separation between Left and Right populations following stimulus presentation (greater $\lambda_{\text{side}}$).
@@
 \subsubsection{Modeling optogenetic silencing.} \label{methods_sc_opto}
-We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2019collicular}.
+We tested whether the inferred SC model connectivities could reproduce experimental effects of optogenetic inactivation in rats \cite{duan2021collicular}.
 During periods of simulated optogenetic inactivation, activity was decreased proportional to the optogenetic strength $\gamma \in [0, 1]$

### 2021-06-08 tests/test_normalizing_flows.py
adjust spline filename check for default spline bins

@@ -217,7 +217,7 @@ def test_to_string():
         batch_norm=False,
         random_seed=20,
     )
-    assert nf.to_string() == "D100_C2_spline_L4_U200_bins=32_PA_rs20"
+    assert nf.to_string() == "D100_C2_spline_L4_U200_bins=4_PA_rs20"
 
     nf = NormalizingFlow("coupling", 4, 1, 2, 15, bn_momentum=0.999, post_affine=False)
     assert nf.to_string() == "D4_C1_affine_L2_U15_bnmom=9.99E-01_rs1"
@@ -407,4 +407,4 @@ def test_initialization():
 
 
 if __name__ == "__main__":
-    test_initialization()
+    test_to_string()

### 2021-06-17 written/eLifeRevision/Bittner_eLife2021.tex
latest edits

@@
 \textbf{Data availability statement}: \\
-The datasets generated during and/or analyzed during the current study are available from the corresponding author upon reasonable request.
+The datasets generated during this study have been made publicly available on Zenodo at this address: 
+https://doi.org/10.5281/zenodo.4910010 .
 
 \textbf{Code availability statement}: \\
-All software written for the current study is available at https://github.com/cunningham-lab/epi.
+All software written for this study is available at https://github.com/cunningham-lab/epi .
 
